---
title: NHANES
author: David Kane
tutorial:
  id: 06-nhanes
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #06 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(tidymodels)
library(primer.data)
library(broom)
library(marginaleffects)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

x <- nhanes |>
  filter(age >= 18 & age <= 27) |>
  select(height, sex) |>
  drop_na()

fit_height <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = x)
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

The world confronts us. Make decisions we must.

Imagine that you are in charge of ordering uniforms for next year's Marine Corps bootcamp recruits. There are many factors to consider: the cost of different designs, the number of male and female recruits, the distributions of heights and weights, and so on. There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

The best way to (try to) ensure that students remember these concepts more than a few months after the course ends is [spaced repetition](https://en.wikipedia.org/wiki/Spaced_repetition), although we focus more on the repetition than on the spacing. You will see the same questions more than once, across tutorials and within tutorials.

### Exercise 2

You should be working inside a GitHub repo named `nhanes`, opened in a Codespace from the [`PPBDS/codespace-starter`](https://github.com/PPBDS/codespace-starter) template. If you are not there yet, please create that repo and open it in a Codespace now --- see the [package README](https://github.com/PPBDS/primer/tree/main/primer.tutorials#working-environments-and-repo-setup) if you need setup instructions or are working locally instead.

Once you're inside the `nhanes` repo, create a new Quarto document titled `"NHANES"` with yourself as the author, render it, and save it as `analysis.qmd`.

Create a `.gitignore` file with `analysis_files` on the first line followed by a blank line. Save and push.

At the R prompt, run:

```
show_file(".gitignore")
```

If that fails, it's probably because you haven't loaded `library(tutorial.helpers)` at the R prompt.

"CP/CR" is short for *copy/paste the **c**ommand and the **r**esponse* --- both the command you sent to R and the response R gave back.

CP/CR.

```{r introduction-2}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Professionals keep their data science work in the cloud because laptops fail. Using GitHub for every project --- and rendering to GitHub Pages --- is the habit we want to build.

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.data)` in a new code chunk. Render the file.

Notice that the file does not look good because the code is visible and there are messages cluttering the output. To fix this, add `#| message: false` to this setup chunk to remove the messages. Also add the following to the YAML header, between the `---` delimiters, to hide the code itself:

```
execute:
  echo: false
```

Render again --- it should look cleaner. At the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    #| message: false\n    library(tidyverse)\n    library(primer.data)\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Rendering a QMD converts the code and prose into an HTML document that shows only what you want readers to see --- in this case, the eventual plot and summary paragraph, with the code hidden. Professional data-science deliverables almost never show raw code; the code is the path, not the destination.

### Exercise 4

Place your cursor in the QMD file on the `library(tidyverse)` line. Use `Cmd/Ctrl + Enter` to execute that line. Note that this sends the line to the R prompt and runs it.

CP/CR.

```{r introduction-4}
question_text(NULL,
  message = "    > library(tidyverse)\n    ── Attaching core tidyverse packages ─ tidyverse 2.0.0 ─\n    ✔ dplyr     ✔ forcats   ✔ ggplot2  ✔ lubridate\n    ✔ purrr     ✔ readr     ✔ stringr  ✔ tibble     ✔ tidyr\n    ── Conflicts ─ tidyverse_conflicts() ─\n    ✖ dplyr::filter() masks stats::filter()\n    ✖ dplyr::lag()    masks stats::lag()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Two worlds matter in data science: *QMD World* (the file where your code lives) and *R World* (the running R session at the R prompt, where objects and loaded packages live). `Cmd/Ctrl + Enter` moves a single line from QMD World into R World. You will do this hundreds of times.

### Exercise 5

Place your cursor in the QMD file on the `library(primer.data)` line. Use `Cmd/Ctrl + Enter` to execute that line.

CP/CR.

```{r introduction-5}
question_text(NULL,
  message = "    > library(primer.data)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

A version of the data from the National Health and Nutrition Examination Survey is available in the `nhanes` tibble from the **[primer.data](https://github.com/PPBDS/primer.data)** package. NHANES is an ongoing survey conducted by the Centers for Disease Control and Prevention.

### Exercise 6

At the R prompt, type `?nhanes` and paste the Description below.

```{r introduction-6}
question_text(NULL,
  message = "The National Health and Nutrition Examination Survey (NHANES) is a research survey distributed to adults and children across the United States to examine health and nutritional status nationwide. It is unique in that it conducts both interviews and physical examination to produce their data. The survey is run by the National Center for Health Statistics (NCHS), a part of the Centers for Disease Control and Prevention.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Note that NHANES intentionally oversamples some demographic groups --- older adults, racial and ethnic minorities --- so that each subgroup has enough observations to be studied on its own. The raw tibble is not representative of the general population; the survey ships sampling weights to correct for this. We will not use the weights in this tutorial, which we should flag as a concern later in Justice.

### Exercise 7

Height of young adults is the broad topic of this tutorial. Given that topic, which variable in `nhanes` should we use as our outcome variable?

```{r introduction-7}
question_text(NULL,
  message = "Our outcome variable is `height`. We care about the height of USMC recruits, who are young adults.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

We will use `height` as our outcome variable.

```{r}
#| echo: false
#| message: false
#| warning: false
nhanes |>
  filter(age >= 18 & age <= 27) |>
  drop_na(height) |>
  ggplot(aes(x = height)) +
  geom_density(fill = "grey70", color = "grey30") +
  labs(
    title = "Height of Young Adults",
    subtitle = "A broad, asymmetric peak with a heavy left shoulder --- a hint that sex matters for height",
    x = "Height (cm)",
    y = "Density",
    caption = "Source: NHANES via primer.data"
  ) +
  theme_minimal()
```

### Exercise 8

Which variable in `nhanes` do you think might have an important connection to `height`?

```{r introduction-8}
question_text(NULL,
  message = "One candidate is `sex`. Men and women tend to have systematically different average heights, so a model that includes `sex` should do better than one that does not.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

With a predictive model, each individual unit has only one observed outcome. Predictive models have no "treatments" --- only covariates.

### Exercise 9

Specify two different groups of young adults which have different values for `sex` and which might have different average values for `height`.

```{r introduction-9}
question_text(NULL,
  message = "Consider two groups: young adults with `sex` = Male, and young adults with `sex` = Female. These two groups might have different average heights.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

In predictive models, do not use "cause," "influence," "impact," or anything else which suggests causation. The best phrasing is in terms of "differences" between groups of units with different values for a covariate of interest.

### Exercise 10

Write a predictive question connecting the outcome `height` to `sex`, the covariate of interest.

```{r introduction-10}
question_text(NULL,
  message = "What is the average height of male and female USMC recruits?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The average height by sex is a convenient question --- a single number per sex, easy to interpret, easy to use for ordering uniforms. It isn't the only useful question. A logistics officer might also want the tallest recruit expected next year (so no one arrives without a uniform that fits), or the 10th and 90th percentiles of each sex's height distribution (to know how many small-versus-large sizes to order). We stick with the average question here because it exercises every step of the Cardinal Virtues cleanly; the matching chapter considers the broader family.


## Wisdom

Data science starts with some broad questions and a data set which might help us to answer them.

Imagine that you are in charge of ordering uniforms for next year's Marine Corps bootcamp recruits. There are many factors to consider: the cost of different designs, the number of male and female recruits, the distributions of heights and weights, and so on. There are many decisions to make.

The specific question: *What is the average height of male and female USMC recruits?*

We will work from the National Health and Nutrition Examination Survey (NHANES), conducted by the Centers for Disease Control and Prevention, available in the `nhanes` tibble of the **[primer.data](https://github.com/PPBDS/primer.data)** package.

###

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
  message = "Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

> *The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* --- John W. Tukey

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The Preceptor Table does not include all the covariates which you will eventually include in your model. It only includes, along with the outcome(s), covariates which are mentioned in your question.

### Exercise 3

Before we worry about defining the Preceptor Table for this problem, let's look at our data more carefully. Ask AI to write R code that plots the distribution of `height` in the `nhanes` tibble, filtered to young adults aged 18 through 27 (the USMC recruit range), using a density plot with a descriptive title, subtitle, axis labels, and caption. Paste the code below and run it.

```{r wisdom-3}
question_text(NULL,
  message = "    nhanes |>\n      filter(age >= 18 & age <= 27) |>\n      drop_na(height) |>\n      ggplot(aes(x = height)) +\n      geom_density(fill = \"grey70\", color = \"grey30\") +\n      labs(\n        title = \"Height of Young Adults\",\n        subtitle = \"A heavy left shoulder hints that sex matters\",\n        x = \"Height (cm)\", y = \"Density\",\n        caption = \"Source: NHANES via primer.data\"\n      ) +\n      theme_minimal()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
nhanes |>
  filter(age >= 18 & age <= 27) |>
  drop_na(height) |>
  ggplot(aes(x = height)) +
  geom_density(fill = "grey70", color = "grey30") +
  labs(
    title = "Height of Young Adults",
    subtitle = "A heavy left shoulder hints that sex matters",
    x = "Height (cm)",
    y = "Density",
    caption = "Source: NHANES via primer.data"
  ) +
  theme_minimal()
```

The distribution of `height` among young adults has an uneven peak with a noticeable shoulder on the low end --- a hint that `sex` matters for our predictions.

### Exercise 4

Now look at how `height` varies with `sex`. Ask AI to write R code that plots `height` by `sex` as overlapping density curves (same age filter), with a descriptive title, subtitle, axis labels, and caption. Paste the code below and run it.

```{r wisdom-4}
question_text(NULL,
  message = "    nhanes |>\n      filter(age >= 18 & age <= 27) |>\n      drop_na(height, sex) |>\n      ggplot(aes(x = height, fill = sex)) +\n      geom_density(alpha = 0.5) +\n      labs(\n        title = \"Height by Sex, Young Adults\",\n        subtitle = \"Male and female distributions are offset by about 15 cm\",\n        x = \"Height (cm)\", y = \"Density\", fill = \"Sex\",\n        caption = \"Source: NHANES via primer.data\"\n      ) +\n      theme_minimal()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
nhanes |>
  filter(age >= 18 & age <= 27) |>
  drop_na(height, sex) |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Height by Sex, Young Adults",
    subtitle = "Male and female distributions are offset by about 15 cm",
    x = "Height (cm)",
    y = "Density",
    fill = "Sex",
    caption = "Source: NHANES via primer.data"
  ) +
  theme_minimal()
```

The two distributions are clearly offset --- male young adults are on average taller than female young adults by roughly 15 centimeters. That systematic difference is what our model will try to capture.

### Exercise 5

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-5}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

This problem is predictive, so there is only one outcome column: `height`. We compare average heights between male and female young adults.

### Exercise 6

What are the units for this problem?

```{r wisdom-6}
question_text(NULL,
  message = "Each row represents one USMC recruit expected to enlist during the coming year. There are roughly 5,000 such recruits per year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Specifying the Preceptor Table forces us to think clearly about the units and outcomes implied by the question. No data science project follows a single direction. We always backtrack. There is always dialogue. We model units, but we only really care about aggregates.

### Exercise 7

What is the outcome variable for this problem?

```{r wisdom-7}
question_text(NULL,
  message = "Keep track of two 'outcome' variables: the one in our Preceptor Table and the one in our data. In this case, they share the name `height` --- both refer to a young adult's standing height in centimeters --- though the *measurement procedure* behind each column differs.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The outcome variable that we really care about is often not the outcome variable which our data includes. This compromise --- working with what we have rather than what we really want --- is a part of most data science work in the real world.

### Exercise 8

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r wisdom-8}
question_text(NULL,
  message = "Age at enlistment is a candidate. Height is nearly-but-not-quite stable by the age-18 window we care about, so a ten-year-older recruit is almost the same height as a fresh one. Another candidate is race or ethnicity, since height distributions vary across ancestry groups.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The term "covariates" is used in at least three ways in data science. First, it is all the variables which *might* be useful, regardless of whether or not we have the data. Second, it is all the variables for which we have data. Third, it is the set of variables in the data which we end up using in the model.

### Exercise 9

What are the treatments, if any, for this problem?

```{r wisdom-9}
question_text(NULL,
  message = "There are no treatments in this problem because we are building a predictive model. Nothing in `nhanes` can be manipulated --- we cannot assign a recruit a different sex or a different height. Predictive models have covariates, not treatments.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Remember that a treatment is just another covariate which, for the purposes of this specific problem, we are assuming can be manipulated --- thereby creating two or more different potential outcomes for each unit. Predictive models have no such variable.

### Exercise 10

What moment in time does the Preceptor Table refer to?

```{r wisdom-10}
question_text(NULL,
  message = "The coming year --- specifically, the moment at which each recruit arrives at bootcamp and has their height recorded at enlistment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

A Preceptor Table can never really refer to an exact instant in time since nothing is instantaneous in this fallen world. In practice, the Preceptor Table's "moment" is a narrow window: recruit intake for a single calendar year.

### Exercise 11

Describe in words the Preceptor Table for this problem.

```{r wisdom-11}
question_text(NULL,
  message = "The Preceptor Table has three columns: one identifying each USMC recruit, one for `height`, and one for `sex`. Each row represents one recruit expected to enlist during the coming year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

```{r}
#| echo: false
#| results: asis
pre_tibble <- tibble::tribble(
  ~`Recruit`,       ~`Height (cm)`, ~`Sex`,
  "Recruit 1",      "150",          "Female",
  "Recruit 2",      "172",          "Male",
  "...",            "...",          "...",
  "Recruit 5,000",  "160",          "Female"
)

pre_title_footnote     <- "A Preceptor Table is the smallest table such that, if every cell were filled with its true value, answering the question 'What is the average height of male and female USMC recruits?' would be easy."
pre_units_footnote     <- "Each row is one USMC recruit expected to enlist during the coming year. Missing rows represent the other recruits not shown."
pre_outcome_footnote   <- "Height in centimeters, measured at enlistment."
pre_covariate_footnote <- "Sex, taking values 'Male' or 'Female', as recorded at enlistment."

pre_gt_html <- gt::gt(pre_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`Recruit`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 12

Before we can fit a model, we need a cleaned tibble. Ask AI to write R code that creates a tibble called `x` by taking `nhanes`, filtering to ages 18 through 27 (the USMC recruit range), keeping only `height` and `sex`, and dropping rows with missing values. Paste the code below and run it.

```{r wisdom-12}
question_text(NULL,
  message = "    x <- nhanes |>\n      filter(age >= 18 & age <= 27) |>\n      select(height, sex) |>\n      drop_na()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

`filter(age >= 18 & age <= 27)` restricts the sample to the demographic range of new USMC recruits. `drop_na()` silently removes rows that are missing `height` or `sex` --- a simplification Justice will revisit.

### Exercise 13

In `analysis.qmd`, add a new code chunk and copy/paste the pipeline above. Render the QMD with `Cmd/Ctrl + Shift + K`. Then, at the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r wisdom-13}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    x <- nhanes |>\n      filter(age >= 18 & age <= 27) |>\n      select(height, sex) |>\n      drop_na()\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

You can never look at the data too much. --- Mark Engerman

### Exercise 14

We will be creating a summary paragraph over the course of this tutorial. Write the first two sentences. The first sentence is a general statement about the overall topic, mentioning the general class of outcome and at least one covariate. The second sentence introduces the data source and the specific question --- when/where gathered, how many observations, who collected it.

```{r wisdom-14}
question_text(NULL,
  message = "People vary in height, and that variation is strongly patterned by sex. Using data from the National Health and Nutrition Examination Survey (NHANES) conducted by the Centers for Disease Control and Prevention, we seek to estimate the average height of male and female USMC recruits for the coming year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Read our answer. It will not be the same as yours. You can change your answer to incorporate some of our ideas, but do not copy/paste our answer exactly. Add your two sentences to the summary paragraph portion of your QMD. `Cmd/Ctrl + Shift + K`, then commit/push.


## Justice

Wisdom gives us the Preceptor Table and the data.

```{r}
#| echo: false
#| results: asis
pre_tibble_j <- tibble::tribble(
  ~`Recruit`,       ~`Height (cm)`, ~`Sex`,
  "Recruit 1",      "150",          "Female",
  "Recruit 2",      "172",          "Male",
  "...",            "...",          "...",
  "Recruit 5,000",  "160",          "Female"
)

pre_title_footnote_j     <- "A Preceptor Table is the smallest table such that, if every cell were filled with its true value, answering the question 'What is the average height of male and female USMC recruits?' would be easy."
pre_units_footnote_j     <- "Each row is one USMC recruit expected to enlist during the coming year. Missing rows represent the other recruits not shown."
pre_outcome_footnote_j   <- "Height in centimeters, measured at enlistment."
pre_covariate_footnote_j <- "Sex, taking values 'Male' or 'Female', as recorded at enlistment."

pre_gt_html_j <- gt::gt(pre_tibble_j, id = "preceptor_tbl_justice") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`Recruit`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote_j,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl_justice .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html_j,
  "</div>\n",
  "```\n",
  sep = ""
)
```


```{r}
#| echo: false
#| results: asis
data_tibble <- tibble::tribble(
  ~`Respondent`, ~`Age`, ~`Height (cm)`, ~`Sex`,
  "NHANES 1",    "21",   "162",          "Female",
  "NHANES 2",    "25",   "178",          "Male",
  "NHANES 3",    "19",   "159",          "Female",
  "...",         "...",  "...",          "...",
  "NHANES 50",   "23",   "170",          "Male"
)

data_title_footnote     <- "National Health and Nutrition Examination Survey (NHANES), Centers for Disease Control and Prevention. Continuous survey data, 1999--present. We restrict to respondents aged 18 through 27 with non-missing height and sex --- about 1,300 rows. The original survey used a complex, stratified, multistage sampling design; we ignore the survey weights."
data_units_footnote     <- "Each row is one NHANES participant examined between 1999 and the most recent release cycle."
data_outcome_footnote   <- "Height in centimeters, measured by a trained examiner in NHANES's Mobile Examination Center using a fixed stadiometer. Each participant is measured once per survey visit with shoes removed."
data_covariate_footnote <- "Sex, recorded by the NHANES interviewer as 'Male' or 'Female'. NHANES does not separately record gender identity; the two-category coding dates from survey waves going back decades."

data_gt_html <- gt::gt(data_tibble, id = "data_tbl") |>
  gt::tab_header(title = "Data: NHANES Young Adults, Ages 18–27") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`Respondent`, `Age`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Respondent`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Age`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = data_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = data_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = data_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = data_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #data_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  data_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

###

### Exercise 1

In your own words, name the five key components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
  message = "Justice concerns the Population Table and the four key assumptions which underlie it: validity, stability, representativeness, and unconfoundedness.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Justice is about concerns that you (or your critics) might have, reasons why the model you create might not work as well as you hope.

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Validity is always about the columns in the Preceptor Table and the data. Just because columns from these two different tables have the same name does not mean that they are the same thing.

### Exercise 3

Provide one reason why the assumption of validity might not hold for the outcome variable `height` or for one of the covariates. Use the words "column" or "columns" in your answer.

```{r justice-3}
question_text(NULL,
  message = "The `height` column in the data is recorded by NHANES's trained examiners using a fixed stadiometer, with shoes removed. The `height` column in the Preceptor Table is recorded at USMC enlistment with whatever instrument the recruiting office has on hand --- possibly with shoes on, possibly under different instructions. Even though both columns are labeled 'Height (cm),' the two measurement procedures are not the same, so the columns may not represent the same underlying quantity.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

In order to consider the Preceptor Table and the data to be drawn from the same population, the columns from one must have a *valid correspondence* with the columns in the other. Validity, if true (or at least reasonable), allows us to construct the Population Table, which is the first step in Justice.

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The Population Table is almost always much bigger than the combination of the Preceptor Table and the data, because if we can really assume that both are part of the same population, then that population must cover a broad universe of time and units.

### Exercise 5

Specify the unit/time combinations which define each row in this Population Table.

```{r justice-5}
question_text(NULL,
  message = "Each row is one young adult (aged 18--27) in a given year. Data rows are NHANES participants examined between 1999 and the most recent release. Preceptor rows are the ~5,000 USMC recruits expected to enlist in the coming year. Other rows cover all the young adults alive during the overlapping span who appear in neither table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

```{r}
#| echo: false
#| results: asis
pop_tibble <- tibble::tribble(
  ~Source    , ~`Young Adult`    , ~`Year`, ~`Height (cm)`, ~`Sex`   ,
  "..."      , "..."             , "..."  , "..."         , "..."    ,
  "Data"     , "NHANES 1"        , "2002" , "162"         , "Female" ,
  "Data"     , "NHANES 2"        , "2005" , "178"         , "Male"   ,
  "Data"     , "..."             , "..."  , "..."         , "..."    ,
  "Data"     , "NHANES 50"       , "2015" , "170"         , "Male"   ,
  "..."      , "..."             , "..."  , "..."         , "..."    ,
  "Preceptor", "Recruit 1"       , "2026" , "150"         , "Female" ,
  "Preceptor", "Recruit 2"       , "2026" , "172"         , "Male"   ,
  "Preceptor", "..."             , "..."  , "..."         , "..."    ,
  "Preceptor", "Recruit 5,000"   , "2026" , "160"         , "Female" ,
  "..."      , "..."             , "..."  , "..."         , "..."
)

pop_title_footnote     <- "This table combines NHANES adult-height data from 1999 onward with the Preceptor Table's 2026 USMC recruits, drawn from the same broader population of US young adults aged 18--27."
pop_units_footnote     <- "Each row is one young adult at one point in time. Data rows are NHANES survey participants between the ages of 18 and 27. Preceptor rows are the ~5,000 USMC recruits expected to enlist in the coming year."
pop_outcome_footnote   <- "Height in centimeters. NHANES measures height directly in the Mobile Examination Center; USMC records height at enlistment. The two measurement procedures are close enough that we treat them as the same column, subject to the validity concern raised in Exercise 3."
pop_covariate_footnote <- "Sex, recorded as Male or Female in both data sources."

pop_gt_html <- gt::gt(pop_tibble, id = "population_tbl") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`Young Adult`, `Year`)) |>
  gt::tab_spanner(label = "Outcome"  , id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Source`, `Young Adult`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 6

In your own words, define the assumption of "stability" when employed in the context of data science.

```{r justice-6}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Stability is all about time. Is the relationship among the columns in the Population Table stable over time? The longer the time period between the data and the Preceptor Table, the more suspect stability becomes.

### Exercise 7

Provide one reason why the assumption of stability might not be true in this case.

```{r justice-7}
question_text(NULL,
  message = "Human height has drifted slowly upward over the last century --- each successive generation of young adults is on average slightly taller than the one before. The NHANES data spans roughly 25 years (1999 onward), and the Preceptor Table refers to next year's recruits. Over that quarter-century the male-female height gap has been roughly stable, but the baseline heights are still creeping up. If the drift is large enough, the Data and Preceptor rows of the Population Table reflect different parameters.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A change in time or the distribution of the data does not, in and of itself, demonstrate a violation of stability. Stability is about the parameters: $\beta_0$, $\beta_1$, and so on. Stability means these parameters are the same in the data as they are in the population as they are in the Preceptor Table.

### Exercise 8

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: Data → Population → Preceptor Table. In your own words, define the assumption of "representativeness."

```{r justice-8}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Ideally, we would like both the Preceptor Table and our data to be random samples from the population. Sadly, this is almost never the case.

### Exercise 9

We do not use the data directly to estimate missing values in the Preceptor Table. Instead, we use the data to learn about the overall population. Provide one reason, involving the relationship between the data and the population, why the assumption of representativeness might not be true in this case.

```{r justice-9}
question_text(NULL,
  message = "NHANES is a voluntary survey. People who agree to participate (and who show up for the Mobile Examination Center visit) are not a random sample of US young adults --- they skew toward people with time, transportation, and some trust in government research. If NHANES participants differ systematically in height from non-participants (say, because willingness to travel to a MEC correlates with health and income, which correlate with height), then the data's heights are a biased estimate of the population's heights.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

When representativeness is violated, the estimates for the model parameters will be biased.

### Exercise 10

We use information about the population to make inferences about the Preceptor Table. Provide one reason, involving the relationship between the Population and the Preceptor Table, why the assumption of representativeness might not be true in this case.

```{r justice-10}
question_text(NULL,
  message = "USMC recruits are not a random sample of the young-adult population. They have self-selected into military service, which means they are more physically fit, probably healthier, and, on average, likely taller than their peers in the broader young-adult population. The Preceptor Table row for a USMC recruit is drawn from a very specific subset of young adults, not the general population.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Stability looks across time periods. Representativeness looks within time periods. Validity is about the columns in our Population Table. Stability and representativeness are about the rows.

### Exercise 11

A statistical model consists of two parts: the probability family and the link function. The probability family is the probability distribution which generates the randomness in our data. The link function is the mathematical formula which links our data to the unknown parameters. Add `library(tidymodels)` to the QMD file. Place your cursor on that line. Use `Cmd/Ctrl + Enter` to execute. Note that this sends the line to the R prompt.

CP/CR.

```{r justice-11}
question_text(NULL,
  message = "    > library(tidymodels)\n    ── Attaching packages ──────── tidymodels 1.x.x ──\n    ✔ broom        ✔ recipes\n    ✔ dials        ✔ rsample\n    ✔ infer        ✔ tune\n    ✔ modeldata    ✔ workflows\n    ✔ parsnip      ✔ workflowsets\n    ✔ purrr        ✔ yardstick\n    ── Conflicts ──── tidymodels_conflicts() ──\n    ✖ purrr::discard() masks scales::discard()\n    ✖ dplyr::filter()  masks stats::filter()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Our outcome variable `height` is continuous. The probability family is Normal:

$$Y \sim N(\mu, \sigma^2)$$

### Exercise 12

Add `library(broom)` to the QMD file. Place your cursor on that line. Use `Cmd/Ctrl + Enter`.

CP/CR.

```{r justice-12}
question_text(NULL,
  message = "    > library(broom)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The link function for a continuous outcome is linear:

$$\mu = \beta_0 + \beta_1 X_1 + \ldots + \beta_n X_n$$

Our model will be a linear regression.

### Exercise 13

Write one sentence highlighting a potential weakness in your model. Derive it from possible problems with the assumptions above. We will add this to our summary paragraph. So far our version of the summary paragraph looks like this:

> People vary in height, and that variation is strongly patterned by sex. Using data from the National Health and Nutrition Examination Survey (NHANES) conducted by the Centers for Disease Control and Prevention, we seek to estimate the average height of male and female USMC recruits for the coming year.

Your version will be somewhat different.

```{r justice-13}
question_text(NULL,
  message = "One concern is that USMC recruits are more physically fit than the general young-adult population and so may be systematically taller, meaning that the NHANES-based estimates probably understate recruit heights.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a weakness sentence to the summary paragraph in your QMD. You can modify your paragraph, but don't copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`, then commit/push.


## Courage

Justice gives us the Population Table and the abstract data generating mechanism.

```{r}
#| echo: false
#| results: asis
pop_tibble_c <- tibble::tribble(
  ~Source    , ~`Young Adult`    , ~`Year`, ~`Height (cm)`, ~`Sex`   ,
  "..."      , "..."             , "..."  , "..."         , "..."    ,
  "Data"     , "NHANES 1"        , "2002" , "162"         , "Female" ,
  "Data"     , "NHANES 2"        , "2005" , "178"         , "Male"   ,
  "Data"     , "..."             , "..."  , "..."         , "..."    ,
  "Data"     , "NHANES 50"       , "2015" , "170"         , "Male"   ,
  "..."      , "..."             , "..."  , "..."         , "..."    ,
  "Preceptor", "Recruit 1"       , "2026" , "150"         , "Female" ,
  "Preceptor", "Recruit 2"       , "2026" , "172"         , "Male"   ,
  "Preceptor", "..."             , "..."  , "..."         , "..."    ,
  "Preceptor", "Recruit 5,000"   , "2026" , "160"         , "Female" ,
  "..."      , "..."             , "..."  , "..."         , "..."
)

pop_title_footnote_c     <- "This table combines NHANES adult-height data from 1999 onward with the Preceptor Table's 2026 USMC recruits, drawn from the same broader population of US young adults aged 18--27."
pop_units_footnote_c     <- "Each row is one young adult at one point in time. Data rows are NHANES survey participants between the ages of 18 and 27. Preceptor rows are the ~5,000 USMC recruits expected to enlist in the coming year."
pop_outcome_footnote_c   <- "Height in centimeters. NHANES measures height directly in the Mobile Examination Center; USMC records height at enlistment."
pop_covariate_footnote_c <- "Sex, recorded as Male or Female in both data sources."

pop_gt_html_c <- gt::gt(pop_tibble_c, id = "population_tbl_courage") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`Young Adult`, `Year`)) |>
  gt::tab_spanner(label = "Outcome"  , id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Source`, `Young Adult`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote_c,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote_c,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote_c,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote_c,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl_courage .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html_c,
  "</div>\n",
  "```\n",
  sep = ""
)
```

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n + \epsilon$$

with $\epsilon \sim N(0, \sigma^2)$.

###

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
  message = "Courage creates the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The abstract data generating mechanism shown in the preamble uses generic variables --- $Y$, $X_1$, $X_2$, and so on --- because our purpose is to describe the general mathematical structure. At the end of Justice we have decided the functional family (linear with normal errors), but not the number or identity of covariates we will ultimately include. Courage fills in the missing pieces.

### Exercise 2

Because our outcome variable is continuous, start to create the model by entering `linear_reg(engine = "lm")`.

```{r courage-2, exercise = TRUE}

```

```{r courage-2-hint-1, eval = FALSE}
linear_reg(engine = ...)
```

```{r courage-2-test, include = FALSE}
linear_reg(engine = "lm")
```

###

The **[tidymodels](https://www.tidymodels.org/)** framework is the most popular one in the R world for estimating models. *[Tidy Modeling with R](https://www.tmwr.org/)* by Max Kuhn and Julia Silge is a great introduction.

### Exercise 3

Continue the pipe to `fit(height ~ sex, data = x)`.

```{r courage-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-3-hint-1, eval = FALSE}
... |>
  fit(..., data = ...)
```

```{r courage-3-test, include = FALSE}
linear_reg(engine = "lm") |>
  fit(height ~ sex, data = x)
```

###

A categorical variable (whether character or factor) like `sex` is turned into a 0/1 "dummy" variable which is then renamed something like `sexMale`. We can't have words in a mathematical formula, hence the need for dummy variables.

### Exercise 4

Continue the pipe with `tidy(conf.int = TRUE)`.

```{r courage-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-4-hint-1, eval = FALSE}
...
  tidy(... = TRUE)
```

```{r courage-4-test, include = FALSE}
linear_reg(engine = "lm") |>
  fit(height ~ sex, data = x) |>
  tidy(conf.int = TRUE)
```

###

The intercept represents the expected value of `height` when `sexMale` is 0 --- that is, the average height for female recruits. The `sexMale` coefficient represents how much taller, on average, male recruits are than female recruits.

### Exercise 5

Behind the scenes of this tutorial, an object called `fit_height` has been created which is the result of the code above. Type `fit_height` and hit "Run Code." This prints the object.

```{r courage-5, exercise = TRUE}

```

```{r courage-5-hint-1, eval = FALSE}
fit_height
```

```{r courage-5-test, include = FALSE}
fit_height
```

###

In data science, **we deal with words, math, and code, but the most important of these is code.** We created the mathematical structure of the model and then wrote a model formula in order to estimate the unknown parameters.

### Exercise 6

We need `fit_height` to exist in R World so that later `marginaleffects` calls can use it. Copy/paste this code to the R prompt and execute it:

```
fit_height <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = x)
```

CP/CR.

```{r courage-6}
question_text(NULL,
  message = "    > fit_height <- linear_reg() |>\n    +   set_engine(\"lm\") |>\n    +   fit(height ~ sex, data = x)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Just because something exists in the tutorial (or in the QMD) does not mean that it is at the R prompt. You should be aware of what exists in R World, which is generally called your "workspace."

### Exercise 7

With the parameters estimated, we can write the fitted model as a concrete formula:

$$\hat{\text{height}} = 163 + 13.6 \cdot \text{sexMale}$$

Note the differences from the abstract form shown in the preamble. First, we have replaced the parameters $\beta_0$ and $\beta_1$ with our best estimates (rounded to three significant figures). Second, we have dropped the error term because this formula generates our estimated outcome, not a description of its randomness. Third, the left-hand side is $\widehat{\text{height}}$ instead of $\text{height}$ because this formula generates our estimated (or predicted) height --- a hat indicates an estimated value.

**This is our data generating mechanism.**

Read the formula above carefully, then continue.

```{r courage-7}
question_text(NULL,
  message = "(No command to run --- this exercise is a pause to make sure you have read the concrete formula above.)",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

A data generating mechanism is just a formula, something which we can write down and implement with computer code. Of course, there is randomness built into the DGM, but we won't worry about that detail for now.

### Exercise 8

Create a new code chunk in your QMD. Add the chunk option `#| cache: true`. Copy/paste the code for estimating the model into the chunk, assigning the result to `fit_height`:

```
fit_height <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = x)
```

`Cmd/Ctrl + Shift + K`. Rendering might take a moment the first time as Quarto caches the fit. At the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r courage-8}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    #| cache: true\n    fit_height <- linear_reg() |>\n      set_engine(\"lm\") |>\n      fit(height ~ sex, data = x)\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

Including `#| cache: true` causes Quarto to cache the chunk's results. The next time you render your QMD, as long as you have not changed the code, Quarto will just load up the saved fitted object. To confirm, `Cmd/Ctrl + Shift + K` again --- it should be quick.

### Exercise 9

Add `*_cache` to your `.gitignore` file. Cached objects are often large and don't belong on GitHub.

At the R prompt, run:

```
show_file(".gitignore")
```

CP/CR.

```{r courage-9}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files\n    *_cache",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Because of the change in your `.gitignore` (assuming you saved it), the cache directory should not appear in the Source Control panel --- Git is ignoring it. Commit and push.

### Exercise 10

At the R prompt, run `tidy()` on `fit_height` with the argument `conf.int` set to `TRUE`. This returns 95% intervals for all the parameters in our model.

```{r courage-10, exercise = TRUE}

```

```{r courage-10-hint-1, eval = FALSE}
tidy(..., conf.int = ...)
```

```{r courage-10-test, include = FALSE}
tidy(fit_height, conf.int = TRUE)
```

###

`tidy()` is part of the **[broom](https://broom.tidymodels.org/)** package, used to summarize information from a wide variety of models.

### Exercise 11

Here is a nicer-looking version of that parameter table, rendered with **[knitr](https://yihui.org/knitr/)**:

```{r}
#| echo: false
library(knitr)

tidy(fit_height, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) signif(x, 3))) |>
  kable(
    caption = "Estimates and 95% Confidence Intervals for Parameters.  Source: NHANES via primer.data"
  )
```

Rounding to three significant figures keeps the displayed values consistent with the fitted DGM formula from Exercise 7. At the very least, a table like this should include a title and a caption with the data source.

Read the table carefully, then continue.

```{r courage-11}
question_text(NULL,
  message = "(No command to run --- this exercise is a pause to make sure you have read the table above.)",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The more variables we add, the more difficult it is to interpret the meaning of any particular coefficient. But interpretation also becomes less important. We don't really care about coefficients. We care about using our model to estimate quantities of interest.

### Exercise 12

Add a sentence to your project summary explaining the structure of the model. Something like: "We model XX [concept of outcome, not variable name], [values of XX], as a [linear/logistic/multinomial/ordinal] function of XX [and maybe other covariates]."

Recall the beginning of our version of the summary:

> People vary in height, and that variation is strongly patterned by sex. Using data from the National Health and Nutrition Examination Survey (NHANES) conducted by the Centers for Disease Control and Prevention, we seek to estimate the average height of male and female USMC recruits for the coming year. One concern is that USMC recruits are more physically fit than the general young-adult population and so may be systematically taller, meaning that the NHANES-based estimates probably understate recruit heights.

```{r courage-12}
question_text(NULL,
  message = "We model height as a linear function of sex.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Read our answer. Do not copy/paste exactly. Add your two sentences to the summary paragraph. `Cmd/Ctrl + Shift + K`, then commit/push.


## Temperance

Courage provides the data generating mechanism. We can express the DGM in four ways:

**In words:** We describe height as a linear function of sex.

**In R code:**

```r
fit_height <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = x)
```

**In a parameter table:**

```{r}
#| echo: false
fit_height |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) signif(x, 3)))
```

**As a mathematical formula:**

$$\text{height} = 163 + 13.6 \cdot \text{sexMale} + \epsilon$$

with $\epsilon \sim N(0, 52.2)$.

###

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
  message = "Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Courage gave us the data generating mechanism. Temperance guides us in the use of the DGM --- or the "model" --- to answer the question(s) with which we began. Our tool for doing so is the **[marginaleffects](https://marginaleffects.com/)** package; its companion book *[Model to Meaning](https://marginaleffects.com/)* by Vincent Arel-Bundock is the reference we'll lean on throughout Temperance, starting with the [Predictions chapter](https://marginaleffects.com/chapters/predictions.html).

### Exercise 2

Before jumping to the full DGM, consider a simpler version of the model --- one that ignores `sex` entirely. Here is the fitted intercept-only model, `height ~ 1`:

```{r}
#| echo: false
linear_reg() |>
  set_engine("lm") |>
  fit(height ~ 1, data = x) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) signif(x, 3)))
```

Write a sentence interpreting the 170 estimate for `(Intercept)` in this simple model.

```{r temperance-2}
question_text(NULL,
  message = "The average height of a young adult in our sample is 170 centimeters. This is the expected value we would predict if we ignored every covariate --- including sex.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The simplest model has one parameter: the overall mean. That mean is the "best guess" when we know nothing about the individual beyond their being a young adult. The richer DGM breaks this mean down by sex, and that split is where the useful answer to our question lives.

### Exercise 3

Now consider the full DGM, `height ~ sex`:

```{r}
#| echo: false
fit_height |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) signif(x, 3)))
```

Write a sentence interpreting the 163 estimate for `(Intercept)`.

```{r temperance-3}
question_text(NULL,
  message = "The expected value of the height of a female USMC recruit is 163 centimeters because, for female recruits, `sexMale` is zero.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Most of the time parameters in a model have no direct relationship with any population value in which we might be interested. But in simple, small, linear models, it sometimes happens that a parameter does correspond to something real --- as the intercept does here.

### Exercise 4

Write a sentence interpreting the 13.6 estimate for `sexMale`.

```{r temperance-4}
question_text(NULL,
  message = "When compared with female recruits, male recruits have a 13.6 centimeter higher expected value for height.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Recall that a binary variable like `sex` is turned into a 0/1 "dummy" variable which is then renamed something like `sexMale`. We can't have words --- like "Male" or "Female" --- in a mathematical formula, hence the need for dummy variables.

### Exercise 5

Write a sentence interpreting the confidence interval for `sexMale`.

```{r temperance-5}
question_text(NULL,
  message = "We do not know the true value for the coefficient on `sexMale`, but we can be 95% confident that it lies somewhere between 12.8 and 14.4 centimeters.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

We care if the confidence interval for a given variable excludes zero. If not, we can't be sure whether the relationship between the variable and the outcome is positive or negative. Here, zero is well outside the interval --- male recruits are unambiguously taller than female recruits in this model.

### Exercise 6

In the end, we don't really care about parameters, much less how to interpret them. Parameters are *imaginary*, like unicorns. We care about answers to our questions. *In the modern world, all parameters are nuisance parameters.*

Add `library(marginaleffects)` to the QMD file.

Place your cursor in the QMD file on the `library(marginaleffects)` line. Use `Cmd/Ctrl + Enter` to execute that line.

CP/CR.

```{r temperance-6}
question_text(NULL,
  message = "    > library(marginaleffects)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

We should be modest in the claims we make. The posteriors we create are never the "truth." The assumptions we made to create the model are never perfect. Yet decisions made with flawed posteriors are almost always better than decisions made without them.

### Exercise 7

What is the specific question we are trying to answer?

```{r temperance-7}
question_text(NULL,
  message = "What is the expected height of male and female USMC recruits next year?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Data science projects begin with a decision which we face. To make that decision wisely, we would like to have good estimates of many unknown numbers. Yet, in order to make progress, we need to drill down to one specific question. This leads to the creation of a data generating mechanism, which can then be used to answer lots of questions.

### Exercise 8

At the R prompt, run `predictions()` on `fit_height`.

CP/CR.

```{r temperance-8}
question_text(NULL,
  message = "    > predictions(fit_height)\n    \n     Estimate Std. Error   z Pr(>|z|)  2.5 % 97.5 %\n        162.9      0.283 576   <0.001  162.4  163.5\n        176.6      0.285 620   <0.001  176.0  177.1\n        162.9      0.283 576   <0.001  162.4  163.5\n        ... (1,295 rows total, one per observation)\n    \n    Type: numeric",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

`predictions()` returns a data frame with one row for each observation in the data set used to fit the model. The 50 rows correspond to our 50-row sample. Each row's `estimate` is that row's model-predicted height.

### Exercise 9

At the R prompt, run `plot_predictions()` on `fit_height` with `condition = "sex"`.

CP/CR.

```{r temperance-9}
question_text(NULL,
  message = "    > plot_predictions(fit_height, condition = \"sex\")\n\n(A plot opens showing two point estimates --- one for `Female` near 163 cm and one for `Male` near 177 cm --- each with a narrow 95% confidence interval.)",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

```{r temperance-9-test, include = FALSE}
plot_predictions(fit_height, condition = "sex")
```

###

You should see two point estimates (one for Female, one for Male) with 95% confidence intervals. The gap between them is the model's estimate of the average height difference by sex.

### Exercise 10

At the R prompt, run `plot_predictions()` on `fit_height` with `condition = "sex"` and `draw = FALSE`.

CP/CR.

```{r temperance-10}
question_text(NULL,
  message = "    > plot_predictions(fit_height, condition = \"sex\", draw = FALSE)\n      rowid estimate std.error statistic p.value conf.low conf.high    sex\n    1     1    162.9      0.283       576       0    162.4     163.5 Female\n    2     2    176.6      0.285       620       0    176.0     177.1   Male",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

```{r temperance-10-test, include = FALSE}
plot_predictions(fit_height, condition = "sex", draw = FALSE)
```

###

Because `plot_predictions()` returns a ggplot object, you can continue with ggplot commands like `labs()`. But it can be useful to see the underlying values in the tibble and build your own plot directly.

### Exercise 11

Work with AI to create a beautiful plot starting from the output of `plot_predictions(fit_height, condition = "sex", draw = FALSE)`. Do this in your QMD (much easier than at the R prompt). Title: the key variables. Subtitle: the key takeaway. Caption: data source. Axis labels: human-readable with units.

Paste the plot code below.

```{r temperance-11}
question_text(NULL,
  message = "    plot_predictions(fit_height, condition = \"sex\", draw = FALSE) |>\n      ggplot(aes(x = sex, y = estimate)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high),\n                      color = \"grey20\", size = 1.1) +\n      labs(\n        title = \"Forecast Height for USMC Recruits\",\n        subtitle = \"Male recruits are expected to be about 14 cm taller\",\n        x = \"Sex\",\n        y = \"Expected Height (cm)\",\n        caption = \"Source: NHANES via primer.data\"\n      ) +\n      theme_minimal()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 12)
```

###

Here is our version:

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(fit_height, condition = "sex", draw = FALSE) |>
  ggplot(aes(x = sex, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                  color = "grey20", size = 1.1) +
  labs(
    title = "Forecast Height for USMC Recruits",
    subtitle = "Male recruits are expected to be about 14 cm taller than female recruits",
    x = "Sex",
    y = "Expected Height (cm)",
    caption = "Source: NHANES via primer.data"
  ) +
  theme_minimal()
```

Data science often involves this back-and-forth style of work. First, make a single chunk of code --- say, a new plot --- work well. This requires interactive work between the QMD and the R prompt. Second, ensure that the entire QMD runs correctly on its own.

### Exercise 12

Finalize the new graphics chunk in your QMD. `Cmd/Ctrl + Shift + K` to ensure it all works. At the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r temperance-12}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    plot_predictions(fit_height, condition = \"sex\", draw = FALSE) |>\n      ggplot(aes(x = sex, y = estimate)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +\n      labs(\n        title = \"Forecast Height for USMC Recruits\",\n        subtitle = \"Male recruits are expected to be about 14 cm taller\",\n        x = \"Sex\", y = \"Expected Height (cm)\",\n        caption = \"Source: NHANES via primer.data\"\n      ) +\n      theme_minimal()\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

Always remember: the map is not the territory. A beautiful graphic tells a story, but that story is always an imperfect representation of reality. Our models depend on assumptions that are never completely true.

### Exercise 13

Write the last sentence of your summary paragraph. It describes at least one Quantity of Interest and a measure of uncertainty. It is OK if this QoI differs from the one you began with.

```{r temperance-13}
question_text(NULL,
  message = "We estimate that the average male USMC recruit will be about 177 cm tall and the average female recruit about 163 cm, with 95% confidence intervals of roughly a few centimeters in each direction.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a final sentence to your summary paragraph, but don't copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`.

### Exercise 14

Write a few sentences explaining why the estimates for the quantities of interest, and the uncertainty, might be wrong. Suggest alternative estimates and a wider confidence interval if warranted.

```{r temperance-14}
question_text(NULL,
  message = "USMC recruits are more physically fit and probably taller than the general young-adult population measured by NHANES. A better point estimate would nudge male recruit heights upward by a centimeter or two and female recruit heights similarly. The confidence interval should be wider than what the model reports --- our sample of 50 is small, and the bigger sources of uncertainty (the NHANES-to-USMC validity gap and the voluntary-participation representativeness gap) aren't captured by the model's standard error.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Always go back to your Preceptor Table --- the information which, if you had it, would make answering your question easy. In almost all real-world cases, the Preceptor Table and the data are fairly different. So, even a perfectly estimated statistical model is rarely as useful as we might like.

### Exercise 15

Rearrange the material in your QMD so the order is graphic, then summary paragraph. The chunk that creates `fit_height` must occur before the chunk that creates the graphic. You can keep or discard the math at your discretion. `Cmd/Ctrl + Shift + K`. At the R prompt, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r temperance-15}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\")\n    ---\n    title: \"NHANES\"\n    author: \"<your name>\"\n    execute:\n      echo: false\n    ---\n    \n    ```{r}\n    #| message: false\n    library(tidyverse)\n    library(tidymodels)\n    library(primer.data)\n    library(broom)\n    library(marginaleffects)\n    ```\n    \n    ```{r}\n    x <- nhanes |>\n      filter(age >= 18 & age <= 27) |>\n      select(height, sex) |>\n      drop_na()\n    ```\n    \n    ```{r}\n    #| cache: true\n    fit_height <- linear_reg() |>\n      set_engine(\"lm\") |>\n      fit(height ~ sex, data = x)\n    ```\n    \n    ```{r}\n    plot_predictions(fit_height, condition = \"sex\", draw = FALSE) |>\n      ggplot(aes(x = sex, y = estimate)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +\n      labs(\n        title = \"Forecast Height for USMC Recruits\",\n        subtitle = \"Male recruits are expected to be about 14 cm taller\",\n        x = \"Sex\", y = \"Expected Height (cm)\",\n        caption = \"Source: NHANES via primer.data\"\n      ) +\n      theme_minimal()\n    ```\n    \n    People vary in height, and that variation is strongly patterned by sex. Using data from the National Health and Nutrition Examination Survey (NHANES) conducted by the Centers for Disease Control and Prevention, we seek to estimate the average height of male and female USMC recruits for the coming year. One concern is that USMC recruits are more physically fit than the general young-adult population and so may be systematically taller, meaning that the NHANES-based estimates probably understate recruit heights. We model height as a linear function of sex. We estimate that the average male USMC recruit will be about 177 cm tall and the average female recruit about 163 cm.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

This is the version of your QMD file at which your teacher is most likely to look closely.

### Exercise 16

Publish your rendered QMD to GitHub Pages. In the Terminal (not the R prompt!), run:

```
quarto publish gh-pages analysis.qmd
```

Copy/paste the resulting URL below.

```{r temperance-16}
question_text(NULL,
  message = "https://<your-github-username>.github.io/nhanes/analysis.html",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

Commit/push everything.

### Exercise 17

Copy/paste the URL to your GitHub repo.

```{r temperance-17}
question_text(NULL,
  message = "https://github.com/<your-github-username>/nhanes",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God. If all our assumptions are correct, then our DGM is true --- it accurately describes the way in which the world works. There is no better way to predict the future, or to model the past, than to use it. Sadly, this will only be the case with toy examples involving things like coins and dice.

The world confronts us. Make decisions we must.


## Summary

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(fit_height, condition = "sex", draw = FALSE) |>
  ggplot(aes(x = sex, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                  color = "grey20", size = 1.1) +
  labs(
    title = "Forecast Height for USMC Recruits",
    subtitle = "Male recruits are expected to be about 14 cm taller than female recruits",
    x = "Sex",
    y = "Expected Height (cm)",
    caption = "Source: NHANES via primer.data"
  ) +
  theme_minimal()
```

People vary in height, and that variation is strongly patterned by sex. Using data from the National Health and Nutrition Examination Survey (NHANES) conducted by the Centers for Disease Control and Prevention, we seek to estimate the average height of male and female USMC recruits for the coming year. One concern is that USMC recruits are more physically fit than the general young-adult population and so may be systematically taller, meaning that the NHANES-based estimates probably understate recruit heights. We model height as a linear function of sex. We estimate that the average male USMC recruit will be about 177 cm tall and the average female recruit about 163 cm, with 95% confidence intervals of roughly a few centimeters in each direction.

This forecast would help you estimate how many small-versus-large uniforms to order. It does not address every question the logistics office will face --- it says nothing about weight, shoulder width, or inseam. But it gives a defensible starting point.

The world is always more uncertain than our models would have us believe.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
