---
title: NHANES
author: David Kane and Aashna Patel
tutorial:
  id: 06-nhanes
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #06 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(tidymodels)
library(primer.data)
library(broom)
library(marginaleffects)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 600, tutorial.storage = "local")

# Because we want the uncertainty about the true mean to be
# large enough to show up in our plot in an interesting way, we pretend that we
# only have 50 observations.

set.seed(10)
x <- nhanes |>
  filter(age >= 18 & age <= 27) |>
  select(height, sex) |>
  drop_na() |>
  slice_sample(n = 50) 

fit_height <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = x)
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial supports [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). 

The world confronts us. Make decisions we must.

Imagine that you are in charge of ordering uniforms for next year's Marine Corps bootcamp recruits. There are many factors to consider: the cost of different designs, the number of male and female recruits, the distributions of heights and weights, and so on. There are many decisions to make. 

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

Why do we ask this, and a score more other questions, in each tutorial? Because the best way to (try to) ensure that students remember these concepts more than a few months after the course ends is [spaced repetition](https://en.wikipedia.org/wiki/Spaced_repetition), although we focus more on the repetition than on the spacing.

### Exercise 2

You should be working inside a GitHub repo named `nhanes`, opened in a Codespace from the [`PPBDS/codespace-starter`](https://github.com/PPBDS/codespace-starter) template. If you are not there yet, please create that repo and open it in a Codespace now — see the [package README](https://github.com/PPBDS/primer/tree/main/primer.tutorials#working-environments-and-repo-setup) if you need setup instructions or are working locally instead.

Once you're inside the `nhanes` repo, create a new Quarto document titled `"NHANES"` with yourself as the author, render it, and save it as `analysis.qmd`.

Create a `.gitignore` file with `analysis_files` on the first line and then a blank line. Save and push.

At the R prompt, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` at the R prompt.

CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

Professionals keep their data science work in the cloud because laptops fail.

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.data)` in a new code chunk. Render the file.

Notice that the file does not look good because the code is visible and there are annoying messages. To take care of this, add `#| message: false` to remove all the messages in this `setup` chunk. Also add the following to the YAML header to remove all code echos from the HTML:

```         
execute: 
  echo: false
```

At the R prompt, run:

```         
show_file("analysis.qmd", start = -5)
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

### 

Render again. Everything looks nice, albeit empty, because we have added code to make the file look better and more professional.

### Exercise 4

Place your cursor in the QMD file on the `library(tidyverse)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this sends `library(tidyverse)` to the R prompt and executes it.

CP/CR.

```{r introduction-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

You should practice moving from QMD World to R World and back.

### Exercise 5

Place your cursor in the QMD file on the next line. Use `Cmd/Ctrl + Enter` to execute that line.

This work flow --- writing things in the QMD so that you have a permanent copy and then sending them to the R prompt with `Cmd/Ctrl + Enter` --- is the most common approach to data science.

There is QMD World and R World. It is your responsibility to keep them in sync.

CP/CR.

```{r introduction-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

A version of the data from the National Health and Nutrition Examination Survey is available in the `nhanes` tibble from the **[primer.data](https://github.com/PPBDS/primer.data)** package.

### Exercise 6

At the R prompt, type `?nhanes` and paste the Description below.

```{r introduction-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

NHANES is a voluntary survey that oversamples groups like older adults and minorities. This can make the raw data unbalanced and inaccurate for the whole population, so sampling weights are used to correct it.

### Exercise 7

Height of young adults is the broad topic of this tutorial. Given that topic, which variable in `nhanes` should we use as our outcome variable?

```{r introduction-7}
question_text(NULL,
  message = "Our outcome variable is `height`. We care about the height of USMC recruits, who are young adults.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

We will use `height` as our outcome variable.

```{r}
#| echo: false
#| message: false
#| warning: false

nhanes |>
  filter(age >= 18 & age <= 27) |>
  drop_na(height) |>
  ggplot(aes(x = height)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "Heights of young adults in NHANES",
    subtitle = "Distribution of height among respondents aged 18–27",
    x = "Height (cm)",
    y = "Density",
    caption = "Source: National Health and Nutrition Examination Survey via primer.data"
  )
```

### Exercise 8

Let's consider a predictive model. Which variable in `nhanes` do you think might have an important connection to `height`?

```{r introduction-8}
question_text(NULL,
  message = "One candidate is `sex`. Men and women tend to have systematically different average heights, so a model that includes `sex` should do better than one that does not.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

With a predictive model, each individual unit has only one observed outcome. Predictive models have no "treatments" --- only covariates.

### Exercise 9

Specify two different groups of young adults which have different values for `sex` and which might have different average values for `height`.

```{r introduction-9}
question_text(NULL,
  message = "Consider two groups: young adults with `sex` = Male, and young adults with `sex` = Female. These two groups might have different average heights.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

In predictive models, do not use "cause," "influence," "impact," or anything else which suggests causation. The best phrasing is in terms of "differences" between groups of units with different values for a covariate of interest.

### Exercise 10

Write a predictive question connecting the outcome `height` to `sex`, the covariate of interest.

```{r introduction-10}
question_text(NULL,
  message = "What is the average height of male and female USMC recruits?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

This is the first version of the question. We will now create a Preceptor Table to answer the question. We may then revise the question given complexities discovered in the data. We then update the question and the Preceptor Table. And so on.

## Wisdom

Data science starts with some broad questions and a data set which might help us to answer them.

Imagine that you are in charge of ordering uniforms for next year's Marine Corps bootcamp recruits. There are many factors to consider: the cost of different designs, the number of male and female recruits, the distributions of heights and weights, and so on. There are many decisions to make.

The broad question: *What is the average height of male and female USMC recruits?*

We will work from the National Health and Nutrition Examination Survey (NHANES), conducted by the Centers for Disease Control and Prevention, available in the `nhanes` tibble of the **[primer.data](https://github.com/PPBDS/primer.data)** package.

###

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
  message = "Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

*The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* -- John W. Tukey

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The Preceptor Table does not include all the covariates which you will eventually include in your model. It only includes, along with the outcome(s), covariates which are mentioned in your question.

### Exercise 3

Let's explore the outcome variable. Ask AI to write R code that plots the distribution of `height` in `nhanes`, filtered to young adults aged 18 through 27 (the range for USMC recruits), using a density plot with a descriptive title, subtitle, axis labels, and caption. Paste the code below and run it.

```{r wisdom-3, exercise = TRUE}

```

```{r wisdom-3-hint-1, eval = FALSE}
nhanes |>
  filter(age >= 18 & age <= 27) |>
  drop_na(height) |>
  ggplot(aes(x = height)) +
  geom_density(...) +
  labs(...)
```

```{r wisdom-3-test, include = FALSE}
nhanes |>
  filter(age >= 18 & age <= 27) |>
  drop_na(height) |>
  ggplot(aes(x = height)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "Heights of young adults in NHANES",
    subtitle = "Distribution of height among respondents aged 18-27",
    x = "Height (cm)",
    y = "Density",
    caption = "Source: National Health and Nutrition Examination Survey via primer.data"
  )
```

###

The distribution of `height` among young adults shows two concentrations — roughly centered near the typical female and male heights. That bimodality is a hint that `sex` will matter a lot.

### Exercise 4

Now look at how height varies with our key covariate, `sex`. Ask AI to write R code that plots the distribution of `height` by `sex` (same age filter) as overlapping density curves, with a descriptive title, subtitle, axis labels, and caption. Paste the code below and run it.

```{r wisdom-4, exercise = TRUE}

```

```{r wisdom-4-hint-1, eval = FALSE}
nhanes |>
  filter(age >= 18 & age <= 27) |>
  drop_na(height, sex) |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(alpha = ...) +
  labs(...)
```

```{r wisdom-4-test, include = FALSE}
nhanes |>
  filter(age >= 18 & age <= 27) |>
  drop_na(height, sex) |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Heights of young adults by sex",
    subtitle = "Male heights sit systematically higher than female heights",
    x = "Height (cm)",
    y = "Density",
    caption = "Source: NHANES via primer.data"
  )
```

###

The two groups have clearly different average heights. This is the kind of systematic difference our model will try to capture.

### Exercise 5

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-5}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

This problem is predictive, so there is one outcome column. We compare average `height` between male and female young adults.

### Exercise 6

What are the units for this problem?

```{r wisdom-6}
question_text(NULL,
  message = "The units are USMC recruits enlisting next year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Specifying the Preceptor Table forces us to think clearly about the units and outcomes implied by the question. No data science project follows a single direction. We always backtrack. There is always dialogue. We model units, but we only really care about aggregates.

### Exercise 7

What is the outcome variable for this problem?

```{r wisdom-7}
question_text(NULL,
  message = "Keep track of two 'outcome' variables: the one in our Preceptor Table and the one in our data. In this case, `height`.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The outcome variable that we really care about is often not the outcome variable which our data includes. This compromise --- working with what we *have* rather than what we really *want* --- is a part of most data science work in the real world.

### Exercise 8

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r wisdom-8}
question_text(NULL,
  message = "The obvious covariate would be `sex`, since men and women differ in height, on average.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The term "covariates" is used in at least three ways in data science. First, it is all the variables which might be useful, regardless of whether or not we have the data. Second, it is all the variables for which we have data. Third, it is the set of variables in the data which we end up using in the model.

### Exercise 9

What are the treatments, if any, for this problem?

```{r wisdom-9}
question_text(NULL,
  message = "There are no treatments, since this is a predictive model.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Remember that a treatment is just another covariate which, for the purposes of this specific problem, we are assuming can be manipulated, thereby creating two or more different potential outcomes for each unit.

### Exercise 10

What moment in time does the Preceptor Table refer to?

```{r wisdom-10}
question_text(NULL,
  message = "The next year, or the period of time over which new recruits will arrive for training and need uniforms.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A Preceptor Table can never really refer to an exact instant in time since nothing is instantaneous in this fallen world. In almost all practical problems, the data was gathered at a time other than that to which the Preceptor Table refers.

```{r}
#| echo: false
#| message: false
#| warning: false

nhanes |>
  filter(age >= 18 & age <= 27) |>
  drop_na(height, sex) |>
  ggplot(aes(x = sex, y = height, fill = sex)) +
  geom_boxplot(alpha = 0.5) +
  labs(
    title = "Heights of young adults by sex",
    subtitle = "Male heights sit systematically higher than female heights",
    x = "Sex",
    y = "Height (cm)",
    caption = "Source: NHANES via primer.data"
  ) +
  theme(legend.position = "none")
```

> *You can never look at the data too much.* -- Mark Engerman

### Exercise 11

Describe in words the Preceptor Table for this problem.

```{r wisdom-11}
question_text(NULL,
  message = "The Preceptor Table has three columns: one identifying each USMC recruit, one for `height`, and one for `sex`. Each row represents one recruit expected to enlist during the coming year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

```{r}
#| echo: false
#| results: asis
pre_tibble <- tibble::tribble(
  ~`Recruit`,       ~`Height (cm)`, ~`Sex`,
  "Recruit 1",      "150",          "Female",
  "Recruit 2",      "172",          "Male",
  "...",            "...",          "...",
  "Recruit 5,000",  "160",          "Female"
)

pre_title_footnote     <- "A Preceptor Table is the smallest table such that, if every cell were filled with its true value, answering the question 'What is the average height of male and female USMC recruits?' would be easy."
pre_units_footnote     <- "Each row is one USMC recruit expected to enlist during the coming year. Missing rows represent the other recruits not shown."
pre_outcome_footnote   <- "Height in centimeters, measured at enlistment."
pre_covariate_footnote <- "Sex, taking values 'Male' or 'Female', as recorded at enlistment."

pre_gt_html <- gt::gt(pre_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`Recruit`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal",
    table.width               = "auto",
    container.width           = "auto"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  # Force footnote <td> cells to not drive the table's preferred (max-content)
  # width. Without this, the long footnote text makes gt's table have a
  # max-content width of ~1300px, so shrink-to-fit on any ancestor container
  # clamps to the available width and the table appears full-width. Capping
  # the footnote cell width and allowing words to break forces the table's
  # preferred width to come from the data columns only (~290px).
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

# Wrap in an inline-block div so the immediate containing block
# shrink-wraps to the table's intrinsic content width. Emit inside a
# pandoc raw-HTML fence (```{=html} ... ```) so pandoc passes it
# through verbatim instead of trying to parse the div content as
# markdown — that was what produced the earlier "Div unclosed ...
# closing implicitly" warnings and left the wrapper scoped wrong.
cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 12

Ask AI to write R code that creates a tibble called `x` by taking `nhanes`, filtering to ages 18 through 27 (USMC recruit range), keeping only `height` and `sex`, dropping rows with missing values, and taking a random sample of 50 rows. Set `set.seed(10)` on the line before the pipeline so the sample is reproducible. Paste the code below and run it.

```{r wisdom-12, exercise = TRUE}

```

```{r wisdom-12-hint-1, eval = FALSE}
set.seed(10)
x <- nhanes |>
  filter(age >= 18 & age <= 27) |>
  select(height, sex) |>
  drop_na() |>
  slice_sample(n = 50)
```

```{r wisdom-12-test, include = FALSE}
set.seed(10)
x <- nhanes |>
  filter(age >= 18 & age <= 27) |>
  select(height, sex) |>
  drop_na() |>
  slice_sample(n = 50)
```

###

We pretend we only have 50 observations so the uncertainty about the true average is large enough to show up in our plot in an interesting way.

### Exercise 13

In `analysis.qmd`, add a new code chunk and copy/paste the pipeline above (including the `set.seed(10)` line). Render the QMD with `Cmd/Ctrl + Shift + K`. Then, at the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r wisdom-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

Now you have an object `x`, a subset of `nhanes` cleaned and narrowed down to the units our question is about.

### Exercise 14

What is the narrow, specific question we will try to answer?

```{r wisdom-14}
question_text(NULL,
  message = "What is the average height of male and female USMC recruits?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The answer to this question is your Quantity of Interest. It is OK if your question differs from ours. Many similar questions lead to the creation of the same model.

### Exercise 15

We will be creating a summary paragraph over the course of this tutorial. Write the first two sentences.

The first sentence is a general statement about the overall topic, mentioning the general class of the outcome and at least one covariate. The second introduces the data source and the specific question --- when/where gathered, how many observations, who collected it.

```{r wisdom-15}
question_text(NULL,
  message = "People vary in height, and that variation is strongly patterned by sex. Using data from the National Health and Nutrition Examination Survey conducted from 2009 to 2011 by the Centers for Disease Control and Prevention, we seek to estimate the average height of male and female USMC recruits for the coming year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Read our answer. It will not be the same as yours. You can change your answer to incorporate some of our ideas, but do not copy/paste our answer exactly. Add your two sentences to your QMD, `Cmd/Ctrl + Shift + K`, and commit/push.

## Justice

Recall the Preceptor Table we built in Wisdom:

```{r}
#| echo: false
#| results: asis
pre_tibble_j <- tibble::tribble(
  ~`Recruit`,       ~`Height (cm)`, ~`Sex`,
  "Recruit 1",      "150",          "Female",
  "Recruit 2",      "172",          "Male",
  "...",            "...",          "...",
  "Recruit 5,000",  "160",          "Female"
)

pre_title_footnote_j     <- "A Preceptor Table is the smallest table such that, if every cell were filled with its true value, answering the question 'What is the average height of male and female USMC recruits?' would be easy."
pre_units_footnote_j     <- "Each row is one USMC recruit expected to enlist during the coming year. Missing rows represent the other recruits not shown."
pre_outcome_footnote_j   <- "Height in centimeters, measured at enlistment."
pre_covariate_footnote_j <- "Sex, taking values 'Male' or 'Female', as recorded at enlistment."

pre_gt_html_j <- gt::gt(pre_tibble_j, id = "preceptor_tbl_justice") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`Recruit`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote_j,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl_justice .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html_j,
  "</div>\n",
  "```\n",
  sep = ""
)
```

And here is the data we are working with:

```{r}
#| echo: false
#| results: asis
data_tibble <- tibble::tribble(
  ~`Respondent`, ~`Age`, ~`Height (cm)`, ~`Sex`,
  "NHANES 1",    "21",   "162",          "Female",
  "NHANES 2",    "25",   "178",          "Male",
  "NHANES 3",    "19",   "159",          "Female",
  "...",         "...",  "...",          "...",
  "NHANES 50",   "23",   "170",          "Male"
)

data_title_footnote     <- "National Health and Nutrition Examination Survey (NHANES), Centers for Disease Control and Prevention. Continuous survey data, 1999–present. We restrict to respondents aged 18 through 27 and take a random sample of 50 rows using `set.seed(10)` for reproducibility. Original survey used a complex, stratified multistage sampling design; our simple random subsample discards the survey weights."
data_units_footnote     <- "Each row is one NHANES participant examined between 1999 and the most recent release cycle."
data_outcome_footnote   <- "Height in centimeters, measured by a trained examiner in NHANES's Mobile Examination Center using a fixed stadiometer. Each participant is measured once per survey visit with shoes removed."
data_covariate_footnote <- "Sex, recorded by the NHANES interviewer as 'Male' or 'Female'. NHANES does not separately record gender identity; the two-category coding dates from survey waves going back decades."

data_gt_html <- gt::gt(data_tibble, id = "data_tbl") |>
  gt::tab_header(title = "Data: NHANES Young Adults, Ages 18–27") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`Respondent`, `Age`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Respondent`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Age`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = data_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = data_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = data_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = data_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #data_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  data_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

###

### Exercise 1

In your own words, name the key components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
  message = "Justice concerns the Population Table and the four key assumptions which underlie it: validity, stability, representativeness, and unconfoundedness.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Justice is about concerns that you (or your critics) might have, reasons why the model you create might not work as well as you hope. 

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Validity is always about the *columns* in the Preceptor Table and the data. Just because columns from these two different tables have the same *name* does not mean that they are the same *thing*.

### Exercise 3

Provide one reason why the assumption of validity might not hold for the outcome variable `height`. Use the words "column" or "columns" in your answer.

```{r justice-3}
question_text(NULL,
  message = "Is the way people measure height in the survey similar to the way it is being measured now -- e.g. Are measurements taken with shoes on or shoes off --. Such issues can cause the `height` column in NHANES in 2010 and the column `height` in the Preceptor Table not refer to the same thing.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

In order to consider the Preceptor Table and the data to be drawn from the same population, the columns from one must have a *valid correspondence* with the columns in the other. Validity, if true (or at least reasonable), allows us to construct the Population Table, which is the first step in Justice.

Because we control the Preceptor Table and, to a lesser extent, the original question, we can adjust those variables to be “closer” to the data that we actually have. This is another example of the iterative nature of data science. If the data is not close enough to the question, then we check with our boss/colleague/customer to see if we can modify the question in order to make the match between the data and the Preceptor Table close enough for validity to hold.

Despite these potential problems, we will assume that validity holds since it, mostly (?), does.

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The Population Table is almost always much bigger than the combination of the Preceptor Table and the data because, if we can really assume that both the Preceptor Table and the data are part of the same population, then that population must cover a broad universe of time and units since the Preceptor Table and the data are, themselves, often far apart from each other.

### Exercise 5

Specify the unit/time combinations which define each row in the Population Table for our problem.

```{r justice-5}
question_text(NULL,
  message = "Person/year is the unit/time combination which defines each row in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

In general, not all possible unit/time combinations are present in the Population Table. 

<!-- DK: Give an example if possible. -->

<!-- DK: Show what the population table looks like. -->

### Exercise 6

In your own words, define the assumption of "stability" when employed in the context of data science.

```{r justice-6}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The longer the time period between the Preceptor Table and the data, the more suspect the assumption of stability becomes. 

### Exercise 7

Provide one reason why the assumption of stability might not be true in this case.

```{r justice-7}
question_text(NULL,
  message = "One possible reason for the failure of the assumption of stability would be if immigrants since 2011 have a different average adult male height. This will change the overall average adult male height in America.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

A change in time or the distribution of the data does not, in and of itself, demonstrate a violation of stability. Stability is about the parameters: $\beta_0$, $\beta_1$ and so on. Stability means these parameters are the same in the data as they are in the population as they are in the Preceptor Table.

### Exercise 8

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: `Data -> Population -> Preceptor Table`. In your own words, define the assumption of "representativeness" when employed in the context of data science.

```{r justice-8}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Ideally, we would like both the Preceptor Table *and* our data to be random samples from the population. Sadly, this is almost never the case.

### Exercise 9

We do not use the data, directly, to estimate missing values in the Preceptor Table. Instead, we use the data to learn about the overall population. Provide one reason, involving the relationship between the data and the population, for why the assumption of representativeness might not be true in this case. 

```{r justice-9}
question_text(
  NULL,
  message = "Since participation is voluntary, some people may choose not to respond, which can cause small biases in the data. For example, it’s possible that taller men might be more likely to answer questions about their height.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6
)
```

### 

The reason that representativeness is important is because, when it is violated, the estimates for the model parameters might be biased. 

### Exercise 10

We use information about the population to make inferences about the Preceptor Table. Provide one reason, involving the relationship between the Population and the Preceptor Table, for why the assumption of representativeness might not be true in this case.

```{r justice-10}
question_text(
  NULL,
  message = "The population covers the last 20 years or more. The rows from the Preceptor Table are just for the coming year. If the height of recruits today is systematically different from the height of recruits over the 20 year period of the Population Table, then representativeness would not be true.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6
)
```

### 

Stability looks across time periods. Representativeness looks within time periods, for the most part. 

### Exercise 11

In your own words, define the assumption of "unconfoundedness" when employed in the context of data science.

```{r justice-11}
question_text(NULL,
  message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

This assumption is only relevant for causal models. We describe a model as "confounded" if this is not true. The easiest way to ensure unconfoundedness is to assign treatment randomly.

### Exercise 12

A statistical model consists of two parts: the *probability family* and the *link function*. The probability family is the probability distribution which generates the randomness in our data. The link function is the mathematical formula which links our data to the unknown parameters in the probability distribution. 

Add `library(tidymodels)` to the QMD file.

Place your cursor in the QMD file on the `library(tidymodels)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this sends `library(tidymodels)` to the R prompt and executes it.

CP/CR.

```{r justice-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The probability family is determined by the outcome variable $Y$. 

Since $Y$ is a continuous variable, the probability family is Normal, also known as Gaussian.

$$Y \sim N(\mu, \sigma^2)$$

### Exercise 13

Add `library(broom)` to the QMD file.

Place your cursor in the QMD file on the `library(broom)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this sends `library(broom)` to the R prompt and executes it.

CP/CR.

```{r justice-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The link function, the basic mathematical structure of the model, is (mostly) determined by the type of outcome variable. 

For a continuous outcome variable, we use a linear model for the link function:

$$\mu = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \ldots$$

### Exercise 14

<!-- DK: This question seems really hard! -->

Use AI to come up with a $\LaTeX$ representation of the mathematical structure of the model, with $Y$ as the dependent variable. This version will not have the values of any parameters since we have not, yet, estimated them. Confirm that the $\LaTeX$ code works by placing it in your QMD and then rendering. Paste that $\LaTeX$ code below.

```{r justice-14}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Our answer:

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n + \epsilon$$

with $\epsilon \sim \mathcal{N}(0, \sigma^2)$.

Which we created with $\LaTeX$ code that looks like this:

````
$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n + \epsilon$$

with $\epsilon \sim \mathcal{N}(0, \sigma^2)$.
````

This follows the linear regression form for continuous data.

We use generic variables --- $Y$, $X_1$ --- because our purpose is to describe the general mathematical structure of the model, independent of the specific variables we will eventually choose to use.

Having decided on the basic mathematical structure of the model, a choice mostly driven by the distribution of our outcome variable, we now turn toward estimating the model.

### Exercise 15

Write one sentence which highlights a potential weakness in your model. This will almost always be derived from possible problems with the assumptions discussed above. We will add this sentence to our summary paragraph. So far, our version of the summary paragraph looks like this:

> People vary in height. Using data from the National Health and Nutrition Examination Survey conducted from 2009 to 2011 by the Centers for Disease Control and Prevention, we seek to estimate the average height of USMC recruits for the coming year.

Of course, your version will be somewhat different.

```{r justice-15}
question_text(NULL,
  message = "Since the participation for the survey is voluntary, it could mean that taller people are more likely to answer this question, thus hurting representativeness of the data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Add a weakness sentence to the summary paragraph in your QMD. You can modify your paragraph as you see fit, but do not copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`, and then commit/push.

## Courage

Recall the Population Table we built in Justice:

```{r}
#| echo: false
#| results: asis
pop_tibble <- tibble::tribble(
  ~Source    , ~`Young Adult`    , ~`Year`, ~`Height (cm)`, ~`Sex`   ,
  "..."      , "..."             , "..."  , "..."         , "..."    ,
  "Data"     , "NHANES 1"        , "2002" , "162"         , "Female" ,
  "Data"     , "NHANES 2"        , "2005" , "178"         , "Male"   ,
  "Data"     , "..."             , "..."  , "..."         , "..."    ,
  "Data"     , "NHANES 50"       , "2015" , "170"         , "Male"   ,
  "..."      , "..."             , "..."  , "..."         , "..."    ,
  "Preceptor", "Recruit 1"       , "2026" , "150"         , "Female" ,
  "Preceptor", "Recruit 2"       , "2026" , "172"         , "Male"   ,
  "Preceptor", "..."             , "..."  , "..."         , "..."    ,
  "Preceptor", "Recruit 5,000"   , "2026" , "160"         , "Female" ,
  "..."      , "..."             , "..."  , "..."         , "..."
)

pop_title_footnote     <- "This table combines NHANES adult-height data from 1999 onward with the Preceptor Table's 2026 USMC recruits, drawn from the same broader population of US young adults aged 18-27."
pop_units_footnote     <- "Each row is one young adult at one point in time. Data rows are NHANES survey participants between the ages of 18 and 27. Preceptor rows are the ~5,000 USMC recruits expected to enlist in the coming year."
pop_outcome_footnote   <- "Height in centimeters. NHANES measures height directly in the mobile examination center; USMC records height at enlistment. The two measurement procedures are close enough that we treat them as the same column."
pop_covariate_footnote <- "Sex, recorded as Male or Female in both data sources."

pop_gt_html <- gt::gt(pop_tibble, id = "population_tbl") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`Young Adult`, `Year`)) |>
  gt::tab_spanner(label = "Outcome"  , id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Source`, `Young Adult`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

And the abstract mathematical structure of the model chosen at the end of Justice:

$$Y = \beta_0 + \beta_1 X_1 + \epsilon$$

with $\epsilon \sim N(0, \sigma^2)$.

We use generic variables --- $Y$, $X_1$, and so on --- because our purpose here is to describe the general mathematical structure of the model, independent of the specific variables we will eventually use.

###

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
  message = "Courage creates the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Having decided on the basic mathematical structure of the model at the end of *Justice*, a choice mostly driven by the distribution of our outcome variable, we now turn toward estimating the model.

### Exercise 2

Because our outcome variable is continuous, start to create the model by using `linear_reg()`.

```{r courage-2, exercise = TRUE}

```

```{r courage-2-hint-1, eval = FALSE}
linear_reg()
```

```{r courage-2-test, include = FALSE}
linear_reg()
```

### 

The **[tidymodels](https://www.tidymodels.org/)** framework is the most popular one in the R world for estimating models. *[Tidy Modeling with R](https://www.tmwr.org/)* by Max Kuhn and Julia Silge is a great introduction.

`linear_reg()` accepts a variety of "engines." In this case, the default engine, `lm`, is a good choice, so we can call `linear_reg()` without any arguments.

### Exercise 3

Continue the pipe to `fit(height ~ sex, data = x)`.

```{r courage-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-3-hint-1, eval = FALSE}
... |> 
  fit(..., data = ...)
```

```{r courage-3-test, include = FALSE}
linear_reg() |> 
  fit(height ~ sex, data = x)
```

### 

The only left-hand side variable in our model is the intercept, specified by `1` in the formula. In more complex models, we do not need to specify the intercept explicitly. It is included by default.

### Exercise 4

Continue the pipe with `tidy(conf.int = TRUE)`.

```{r courage-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-4-hint-1, eval = FALSE}
...
  tidy(... = TRUE)
```

```{r courage-4-test, include = FALSE}
linear_reg() |> 
  fit(height ~ sex, data = x) |> 
  tidy(conf.int = TRUE)
```

### 

In this simple model, the intercept is the expected value for the average height.

### Exercise 5

Behind the scenes of this tutorial, an object called `fit_height` has been created which is the result of the code above. Type `fit_height` and hit "Run Code." This generates the same results as using `print(fit_height)`.

```{r courage-5, exercise = TRUE}

```

```{r courage-5-hint-1, eval = FALSE}
fit_height
```

```{r courage-5-test, include = FALSE}
fit_height
```

### 

In data science, **we deal with words, math, and code, but the most important of these is code.** We created the mathematical structure of the model and then wrote a model formula in order to estimate the unknown parameters. 

###

With the parameters estimated, we can write the fitted model as a concrete formula:

$$\hat{\text{height}} = 162 + 15.9 \cdot \text{sexMale}$$

Note the differences from the abstract form shown in Courage's preamble. First, we have replaced the parameters $\beta_0$ and $\beta_1$ with our best estimates (rounded to three significant figures). Second, we have dropped the error term because this is a formula for predicting the value of our outcome variable, not for describing its randomness. Third, the left-hand side is $\widehat{\text{height}}$ instead of $\text{height}$ because this formula generates our estimated (or predicted) height — a hat indicates an estimated value.

**This is our data generating mechanism.**

A data generating mechanism is just a formula, something which we can write down and implement with computer code.

### Exercise 6

Create a new code chunk in your QMD. Add a code chunk option: `#| cache: true`. Copy/paste the code from above for estimating the model into the code chunk, assigning the result to `fit_height`. (This will include the call to `fit()` but not the call to `tidy()` because we want the entire fitted model, not just a table of the estimated parameter values.)

`Cmd/Ctrl + Shift + K`. It may take some time to render your QMD, depending on how complex your model is. But, by including `#| cache: true` you cause Quarto to cache the results of the chunk. The next time you render your QMD, as long as you have not changed the code, Quarto will just load up the saved fitted object.

At the R prompt, run:

```
tutorial.helpers::show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r courage-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

### 

To confirm, `Cmd/Ctrl + Shift + K` again. It should be quick.

### Exercise 7

Add `*_cache` to `.gitignore` file. Cached objects are often large. They don't belong on Github.

At the R prompt, run:

```
tutorial.helpers::show_file(".gitignore")
```

CP/CR.

```{r courage-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Because of the change in your `.gitignore` (assuming that you saved it), the cache directory should not appear in the Source Control panel because Git is ignoring it, as instructed. Commit and push. 

### Exercise 8

At the R prompt, run `tidy()` on `fit_height` with the argument `conf.int` set equal to `TRUE`. This returns 95% intervals for all the parameters in our model.

```{r courage-8, exercise = TRUE}

```

```{r courage-8-hint-1, eval = FALSE}
tidy(..., conf.int = ...)
```

```{r courage-8-test, include = FALSE}
tidy(fit_height, conf.int = TRUE)
```

### 

`tidy()` is part of the [**broom**](https://broom.tidymodels.org/) package, used to summarize information from a wide variety of models.

###

Here is a nicer-looking version of that parameter table, rendered with [**knitr**](https://yihui.org/knitr/):

```{r}
#| echo: false
library(knitr)

tidy(fit_height, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) signif(x, 3))) |>
  kable(
    caption = "Estimates and 95% Confidence Intervals for Parameters\nSource: NHANES 2009–2011"
  )
```

Rounding to three significant figures keeps the displayed values consistent with the fitted DGM formula above and with the interpretation exercises in Temperance. At the very least, a table like this should include a title and a caption with the data source.

### Exercise 9

Add a sentence to your project summary. 

Explain the structure of the model. Something like: "I/we model XX [the concept of the outcome, not the variable name], [insert description of values of XX], as a [linear/logistic/multinomial/ordinal] function of XX [and maybe other covariates]." 

Recall the beginning of our version of the summary:

> People vary in height. Using data from the National Health and Nutrition Examination Survey conducted from 2009 to 2011 by the Centers for Disease Control and Prevention, we seek to estimate the average height of USMC recruits for the coming year. Since the participation for the survey is voluntary, it could mean that taller people are more likely to answer this question, thus hurting representativeness of the data.

```{r courage-9}
question_text(NULL,
  message = "We model height as a linear function of sex.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Read our answer. It will not be the same as yours. You can, if you want, change your answer to incorporate some of our ideas. Do not copy/paste our answer exactly. Add your two sentences, edited or otherwise, to the summary paragraph portion of your QMD. `Cmd/Ctrl + Shift + K`, and then commit/push.

## Temperance

See the [Predictions chapter](https://marginaleffects.com/chapters/predictions.html) of the **[marginaleffects](https://marginaleffects.com/)** book *[Model to Meaning](https://marginaleffects.com/)* for background on the tools we use below.

The fitted model from Courage, `fit_height`, can be described in four complementary ways.

**In words:** We describe height as a linear function of sex.

**In R code:**

```r
fit_height <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = x)
```

**In a parameter table:**

```{r}
#| echo: false
fit_height |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) signif(x, 3)))
```

**As a mathematical formula:**

$$\text{height} = 162 + 15.9 \cdot \text{sexMale} + \epsilon$$

with $\epsilon \sim N(0, 40.3)$.

`sexMale` is a dummy variable: 1 for male recruits and 0 for female recruits. The intercept is the average height (in centimeters) for female recruits; the coefficient on `sexMale` is how much taller, on average, male recruits are. The residual variance is $\sigma^2 \approx 40.3$ (a residual SD of about 6.4 cm). All values are rounded to three significant figures to match the parameter table.

###

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
  message = "Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Courage gave us the data generating mechanism. Temperance guides us in the use of the DGM — or the “model” — we have created to answer the question(s) with which we began. We create posteriors for the quantities of interest.

### Exercise 2

Before jumping to the full DGM, consider a simpler version of the model --- one that ignores `sex` entirely. Here is the fitted intercept-only model, `height ~ 1`:

```{r}
#| echo: false
linear_reg() |>
  set_engine("lm") |>
  fit(height ~ 1, data = x) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) signif(x, 3)))
```

Write a sentence interpreting the 170 estimate for `(Intercept)` in this simple model.

```{r temperance-2}
question_text(NULL,
  message = "The average height of a young adult in our sample is 170 centimeters. This is the expected value we would predict if we ignored every covariate --- including sex.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The simplest model has one parameter: the overall mean. That mean is the "best guess" when we know nothing about the individual beyond their being a young adult. The richer DGM breaks this mean down by sex, and that split is where the useful answer to our question lives.

### Exercise 3

Now consider the full DGM, `height ~ sex`:

```{r}
fit_height |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) signif(x, 3)))
```

Write a sentence interpreting the 162 estimate for `(Intercept)`.

```{r temperance-3}
question_text(NULL,
  message = "The expected value of the height of a female USMC recruit is 162 centimeters because, for female recruits, `sexMale` is zero.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Most of the time parameters in a model have no direct relationship with any population value in which we might be interested. This is especially true in complex and/or non-linear models. That is, in those cases, a parameter like $\beta_0$ does not "mean" anything. But, in simple, small, linear models, it sometimes happens that a parameter does correspond to something real.

### Exercise 4

```{r}
fit_height |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) signif(x, 3)))
```

Write a sentence interpreting the 15.9 estimate for `sexMale`.

```{r temperance-4}
question_text(NULL,
  message = "When compared with female recruits, male recruits have a 15.9 higher expected value for height.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Recall that a binary variable (whether character or factor) like `sex` is turned into a $0/1$ "dummy" variable which is then re-named something like $sexMale$. After all, we can't have words --- like "Male" or "Female" --- in a mathematical formula, hence the need for dummy variables.


### Exercise 5

Write a sentence interpreting the confidence interval for `sexMale`.

```{r temperance-5}
question_text(NULL,
  message = "We do not know the true value for the coefficient for sexMale, but we can be 95% confident that it lies somewhere between 12.3 and 19.6.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

We care if the confidence interval for a given variable excludes zero. If not, then we can't be sure if the relationship between the variable and the outcome is positive or negative. In that case, then why would we include the variable in the model at all?


### Exercise 6

In the end, we don't really care about parameters, much less how to interpret them. Parameters are *imaginary*, like unicorns. We care about answers to our questions. Parameters are tools for answering questions. They aren't interesting in-and-of themselves. *In the modern world, all parameters are nuisance parameters.* 

Add `library(marginaleffects)` to the QMD file.

Place your cursor in the QMD file on the `library(marginaleffects)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this sends `library(marginaleffects)` to the R prompt and executes it.

CP/CR.

```{r temperance-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

We should be modest in the claims we make. The posteriors we create are never the “truth.” The assumptions we made to create the model are never perfect. Yet decisions made with flawed posteriors are almost always better than decisions made without them.

### Exercise 7

What is the specific question we are trying to answer? 

```{r temperance-7}
question_text(NULL,
  message = "What is the expected height of male and female USMC recruits next year?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Data science projects begin with a decision which we face. To make that decision wisely, we would like to have good estimates of many unknown numbers. Yet, in order to make progress, we need to drill down to one specific question. This leads to the creation of a data generating mechanism, which can now be used to answer lots of questions, thus allowing us to explore the original decision more broadly.

### Exercise 8

At the R prompt, run `predictions()` on `height_fit`.

CP/CR.

```{r temperance-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

```{r temperance-8-test, include = FALSE}
predictions(fit_height)
```

###

In the case of a linear model, `predictions()` returns a data frame with one row for each observation in the data set used to fit the model. In this case, `predictions()` returns 50 rows, because the input data has 50 observations. 

Because `sexMale` only has two possible values, there are only two predictions made by the model, one for men (178) and one for women (162).

### Exercise 9

At the R prompt, run `avg_predictions()` on `height_fit`.

CP/CR.

```{r temperance-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

```{r temperance-9-test, include = FALSE}
avg_predictions(fit_height)
```

### 

`avg_predictions()`, without any arguments, returns a single row, averaging all the individual predictions. In this case, the average prediction, 170 centimeters, is between the male and female predictions. 

### Exercise 10

At the R prompt, run `avg_predictions()` on `height_fit` with `by = "sex"`.

CP/CR.

```{r temperance-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

```{r temperance-10-test, include = FALSE}
avg_predictions(fit_height, by = "sex")
```

### 

`by = "sex"` causes `avg_predictions()` to return separate predictions for male and female recruits.

### Exercise 11

At the R prompt, run `plot_predictions()` on `height_fit` with `condition = "sex"`.

CP/CR.

```{r temperance-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

```{r temperance-11-test, include = FALSE}
plot_predictions(fit_height, condition = "sex")
```

### 


### Exercise 12

At the R prompt, run `plot_predictions()` on `fit_height` with `condition = "sex"` and `draw = FALSE`.

CP/CR.

<!-- DK: Really ought to explain condition versus by. -->

```{r temperance-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

```{r temperance-12-test, include = FALSE}
plot_predictions(fit_height, 
                 condition = "sex",
                 draw = FALSE)
```

### 

We use `draw = FALSE` to cause `plot_predictions()` to return a tibble with the numbers which underlie the plot instead of creating the plot itself.


### Exercise 13

Work with AI to create a beautiful plot with the output from this function. Do this in your QMD since that is much easier than working at the R prompt directly.

Your title should highlight the key variables. Your subtitle should describe the key takeaway, the sentence/conclusion which readers will, you hope, remember. Your caption should mention the data source. Your axis labels should look nice.

Copy the code for your plot here:

```{r temperance-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

### 

Our plot:

```{r}
plot_predictions(fit_height, 
                 condition = "sex",
                 draw = FALSE) |> 
ggplot(aes(x = sex, 
           y = estimate, 
           color = sex)) +
  geom_pointrange(aes(ymin = conf.low, 
                      ymax = conf.high),
                  size = 1.2, 
                  linewidth = 1.5,
                  fatten = 3) +
  scale_color_manual(values = c("Female" = "#E74C3C", "Male" = "#3498DB")) +
  labs(
    title = "Forecast Height for USMC Recruits",
    subtitle = "Men are about 16 (cm) taller than women",
    x = NULL,
    y = "Height (cm)",
    color = "Sex"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none") 
```


Created with this code:

````
plot_predictions(fit_height, 
                 condition = "sex",
                 draw = FALSE) |> 
ggplot(aes(x = sex, 
           y = estimate, 
           color = sex)) +
  geom_pointrange(aes(ymin = conf.low, 
                      ymax = conf.high),
                  size = 1.2, 
                  linewidth = 1.5,
                  fatten = 3) +
  scale_color_manual(values = c("Female" = "#E74C3C", "Male" = "#3498DB")) +
  labs(
    title = "Forecast Height for USMC Recruits",
    subtitle = "Men are about 16 (cm) taller than women",
    x = NULL,
    y = "Height (cm)",
    color = "Sex"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none") 
````


Data science often involves this back-and-forth style of work. First, we need to make a single chunk of code, in this case, a new plot, work well. This requires interactive work between the QMD and the R prompt. Second, we need to ensure that the entire QMD runs correctly on its own.

### Exercise 14

Finalize the new graphics code chunk in your QMD. `Cmd/Ctrl + Shift + K` to ensure that it all works as intended. 

At the R prompt, run:

```
tutorial.helpers::show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r temperance-14}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```


### 

Always [remember](https://en.wikipedia.org/wiki/Map%E2%80%93territory_relation): *The map is not the territory.* A beautiful graphic tells a story, but that story is always an imperfect representation of reality. Our models depend on assumptions, assumptions which are never completely true.


### Exercise 15

Write the last sentence of your summary paragraph. It describes at least one quantity of interest (QoI) and provides a measure of uncertainty about that QoI. (It is OK if this QoI is not the one that you began with. The focus of a data science project often changes over time.)

```{r temperance-15}
question_text(NULL,
  message = "Expected height of male (female) recruits will be 178 (162) centimeters with a 95% confidence interval of 176 to 181 (160 to 165).",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Add a final sentence to your summary paragraph in your QMD as you see fit, but do not copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`.

### Exercise 16

Write a few sentences which explain why the estimates for the quantities of interest, and the uncertainty thereof, might be wrong. Suggest an alternative estimate and confidence interval, if you think either might be warranted.

```{r temperance-16}
question_text(NULL,
  message = "USMC recruits are not necessarily representative of the larger population. They are more physically fit and, probably, taller. So, a better estimate for recruit heights would be slightly higher, with more uncertainty because of this issue.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Always go back to your Preceptor Table, the information which, if you had it, would make answering your question easy. In almost all real world cases, the Preceptor Table and the data are fairly different, not least because validity never holds perfectly. So, even a perfectly estimated statistical model is rarely as useful as we might like.

Whether or not to include these concerns in your QMD is up to you!

### Exercise 17

Rearrange the material in your QMD so that the order is graphic, followed by the paragraph. Doing so, of course, requires sensible judgment. For example, the code chunk which creates the fitted model must occur before the chunk which creates the graphic. You can keep or discard the math and any other material at your own discretion.

`Cmd/Ctrl + Shift + K` to ensure that everything works.

At the R prompt, run:

```
tutorial.helpers::show_file("analysis.qmd")
```

CP/CR.

```{r temperance-17}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 30)
```

### 

This is the version of your QMD file which your teacher is most likely to take a close look at.

### Exercise 18

Publish your rendered QMD to GitHub Pages. Copy/paste the resulting url below.

```{r temperance-18}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

Commit/push everything.

### Exercise 19

Copy/paste the url to your Github repo.

```{r temperance-19}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God. If all our assumptions are correct, then our DGM is true, it accurately describes the way in which the world works. There is no better way to predict the future, or to model the past, than to use it. Sadly, this will only be the case with toy examples involving things like coins and dice. We hope that our DGM is close to the true DGM but, since our assumptions are never perfectly correct, our DGM will always be different. The estimated magnitude and importance of that difference is a matter of judgment.

The world confronts us. Make decisions we must.

## Summary
### 

This tutorial supports [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). 

```{r}
plot_predictions(fit_height, 
                 condition = "sex",
                 draw = FALSE) |> 
ggplot(aes(x = sex, 
           y = estimate, 
           color = sex)) +
  geom_pointrange(aes(ymin = conf.low, 
                      ymax = conf.high),
                  size = 1.2, 
                  linewidth = 1.5,
                  fatten = 3) +
  scale_color_manual(values = c("Female" = "#E74C3C", "Male" = "#3498DB")) +
  labs(
    title = "Forecast Height for USMC Recruits",
    subtitle = "Men are about 16 (cm) taller than women",
    x = NULL,
    y = "Height (cm)",
    color = "Sex"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none") 
```

People vary in height. Using data from the National Health and Nutrition Examination Survey conducted from 2009 to 2011 by the Centers for Disease Control and Prevention, we seek to estimate the average height of USMC recruits for the coming year. Since the participation for the survey is voluntary, it could mean that taller people are more likely to answer this question, thus hurting representativeness of the data. Expected height of male (female) recruits will be 178 (162) centimeters with a 95% confidence interval of 176 to 181 (160 to 165).


USMC recruits are not necessarily representative of the larger population. They are more physically fit and, probably, taller. So, a better estimate for recruit heights would be slightly higher, with more uncertainty because of this issue.


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
