---
title: Cardinal Virtues
author: David Kane
tutorial:
  id: cardinal-virtues
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #04 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60,
        tutorial.storage = "local")

hatch <- function(x) {
  paste0(
    '<span style="background-image: repeating-linear-gradient(',
    '45deg, rgba(0,0,0,0.12) 0 1px, transparent 1px 6px);',
    ' padding: 6px 10px; margin: -6px -10px; display: block;">',
    x, '</span>'
  )
}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

The world confronts us. Make decisions we must.

Imagine that you are a data scientist working at a high school. The principal has a big-picture goal --- help her juniors gain admission to good colleges --- and many decisions to make in service of it: which juniors to recommend for SAT prep, which to nominate for selective scholarships, where to focus counseling time, whether to fund a school-wide SAT tutoring class. The principal trusts you to find evidence that helps her decide. This tutorial walks through the four Cardinal Virtues by applying them to two related questions using the same hypothetical SAT dataset: a predictive question (what SAT scores will next year's juniors get?) and a paired causal question (what would happen to SAT scores if she offered a school-wide tutoring class?). The exercises are conceptual --- no R coding here; the goal is to practice the framework before the full example tutorials that follow. There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

The four virtues are the order of operations for any data science problem. References and allusions to them throughout the *Primer* are a feature, not a bug; use them liberally.

### Exercise 2

Define a causal effect.

```{r introduction-2}
question_text(NULL,
  message = "A causal effect is the difference between two potential outcomes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

According to the Rubin Causal Model, there must be two (or more) potential outcomes for any discussion of causation to make sense. This is simplest when the treatment has two values, generating two potential outcomes per unit.

### Exercise 3

What is the fundamental problem of causal inference?

```{r introduction-3}
question_text(NULL,
  message = "The fundamental problem of causal inference is that we can only observe one potential outcome.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

For each junior, we will see the SAT score they actually got under whichever treatment arm they were actually in --- tutored or not tutored. The score they would have gotten in the other arm is the *counterfactual*, and we never see it.


## Wisdom
###

> *Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.*

Wisdom commits us to a question precise enough to be answered. The Preceptor Table makes the commitment concrete: it is the smallest table that, if every cell were filled with the truth, would make the question's answer easy to read off.

> *The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* --- John W. Tukey

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
  message = "Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The Preceptor Table includes only the variables the question forces us to include. Nothing more. Nothing less.

### Exercise 3

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-3}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

In this tutorial we work two questions in parallel: a predictive one and a paired causal one. Each gets its own Preceptor Table.

### The primary question (predictive)

The primary question is:

> What will be the average SAT score of next year's rising-junior class?

This is a predictive question. We want one specific number: an estimate of the average score, with its uncertainty.

### Exercise 4

What are the units for the primary question?

```{r wisdom-4}
question_text(NULL,
  message = "Each row is one rising junior expected to take the SAT next year. This high school has about 200 juniors per class.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

### Exercise 5

What is the outcome variable for the primary question?

```{r wisdom-5}
question_text(NULL,
  message = "The outcome is `sat_score` --- each junior's combined math-plus-verbal SAT score on the standard 400-to-1600 scale.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

### Exercise 6

What treatments, if any, are involved in the primary question?

```{r wisdom-6}
question_text(NULL,
  message = "There are no treatments. The primary question is predictive --- we are estimating an average, not asking what would change if we manipulated anything. Predictive models have covariates, not treatments.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

### Exercise 7

Describe in words what the Preceptor Table for the primary question would look like.

```{r wisdom-7}
question_text(NULL,
  message = "The Preceptor Table has two columns: one identifying each junior, and one for the junior's SAT score. Each row represents one rising junior expected to take the SAT next year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

```{r}
#| echo: false
#| results: asis
pt_primary_tibble <- tibble::tribble(
  ~`Junior`,            ~`SAT Score`,
  "Maria Alvarez",      "1320",
  "James Thompson",     "1450",
  "...",                "...",
  "Sarah Williams",     "1280"
)

pt_primary_html <- gt::gt(pt_primary_tibble, id = "pt_primary") |>
  gt::tab_header(title = "Preceptor Table --- Primary (predictive) question") |>
  gt::tab_spanner(label = "Unit",    id = "unit_span",    columns = c(`Junior`)) |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = c(`SAT Score`)) |>
  gt::cols_align(align = "left",  columns = c(`Junior`)) |>
  gt::cols_align(align = "right", columns = c(`SAT Score`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_column_spanners()) |>
  gt::tab_style(style = gt::cell_text(align = "left"),
                locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_style(style = gt::cell_text(align = "right"),
                locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(
    footnote = "If all the information in this table were available, we could answer the question: What will be the average SAT score of next year's rising-junior class?",
    locations = gt::cells_title()
  ) |>
  gt::tab_footnote(
    footnote = "Each row is one rising junior expected to take the SAT next year. About 200 juniors per class.",
    locations = gt::cells_column_spanners(spanners = "unit_span")
  ) |>
  gt::tab_footnote(
    footnote = "Combined math-plus-verbal SAT score on the 400-to-1600 scale.",
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::opt_css(css = "#pt_primary .gt_footnote { max-width: 1px; word-break: break-word; }") |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    pt_primary_html, "</div>\n```\n", sep = "")
```

### The paired question (causal)

The paired question is:

> What is the causal effect of a school-wide SAT tutoring class on a junior's SAT score?

This is a causal question. The principal could, in principle, offer or withhold the tutoring class. Each junior therefore has *two* potential SAT scores --- one if the school offered tutoring, one if it didn't --- and the causal effect is the difference between those two scores.

### Exercise 8

How many potential outcomes does each junior have in the paired causal question?

```{r wisdom-8}
question_text(NULL,
  message = "Two potential outcomes per junior: the SAT score the junior would get if the school offered the tutoring class, and the SAT score the junior would get if the school did not. The causal effect for that junior is the difference between the two. We can only ever observe one --- the one corresponding to the treatment arm the junior actually ended up in.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

### Exercise 9

What is the treatment in the paired causal question?

```{r wisdom-9}
question_text(NULL,
  message = "The treatment is the SAT tutoring class --- two possible values: `Tutored` (the junior was in a year when the school offered the class) and `Not Tutored` (the junior was not). The principal can, at least in principle, manipulate this by choosing whether or not to fund the program in any given year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

### Exercise 10

Describe in words what the Preceptor Table for the paired causal question would look like.

```{r wisdom-10}
question_text(NULL,
  message = "The Preceptor Table has four columns: one identifying each junior, two potential-outcome columns (`SAT if Tutored`, `SAT if Not Tutored`), and one treatment column (`Tutoring Class`). Each row represents one rising junior. For each junior, only one potential outcome is observable; the other --- the counterfactual --- is the SAT score the junior would have gotten under the treatment they did not receive, and we will mark it with diagonal hatching.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

```{r}
#| echo: false
#| results: asis
pt_paired_tibble <- tibble::tribble(
  ~`Junior`,            ~`SAT if Tutored`, ~`SAT if Not Tutored`, ~`Tutoring Class`,
  "Maria Alvarez",      "1380",            "1320",                "Tutored",
  "James Thompson",     "1500",            "1450",                "Not Tutored",
  "...",                "...",             "...",                 "...",
  "Sarah Williams",     "1340",            "1280",                "Tutored"
)

pt_paired_html <- gt::gt(pt_paired_tibble, id = "pt_paired") |>
  gt::tab_header(title = "Preceptor Table --- Paired (causal) question") |>
  gt::tab_spanner(label = "Unit",                id = "unit_span",
                  columns = c(`Junior`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",
                  columns = c(`SAT if Tutored`, `SAT if Not Tutored`)) |>
  gt::tab_spanner(label = "Treatment",           id = "treatment_span",
                  columns = c(`Tutoring Class`)) |>
  gt::cols_align(align = "left",  columns = c(`Junior`, `Tutoring Class`)) |>
  gt::cols_align(align = "right", columns = c(`SAT if Tutored`, `SAT if Not Tutored`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_column_spanners()) |>
  gt::tab_style(style = gt::cell_text(align = "left"),
                locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span"))) |>
  gt::tab_style(style = gt::cell_text(align = "right"),
                locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  # Hatch each row's unobservable counterfactual.
  # Row 1 (Tutored): `SAT if Not Tutored` is counterfactual.
  # Row 2 (Not Tutored): `SAT if Tutored` is counterfactual.
  # Row 4 (Tutored): `SAT if Not Tutored` is counterfactual.
  gt::text_transform(
    locations = gt::cells_body(columns = `SAT if Not Tutored`, rows = c(1, 4)),
    fn = function(x) hatch(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `SAT if Tutored`, rows = 2),
    fn = function(x) hatch(x)
  ) |>
  gt::tab_footnote(
    footnote = "If all the information in this table were available, we could answer the question: What is the causal effect of a school-wide SAT tutoring class on a junior's SAT score?",
    locations = gt::cells_title()
  ) |>
  gt::tab_footnote(
    footnote = "Each row is one rising junior. The cross-hatched cell in each row marks the unobservable counterfactual --- the SAT score the junior would have gotten under the treatment arm they did not actually end up in.",
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(
    footnote = "The treatment is whether the school offered the SAT tutoring class. Two possible values: Tutored and Not Tutored.",
    locations = gt::cells_column_spanners(spanners = "treatment_span")
  ) |>
  gt::opt_css(css = "#pt_paired .gt_footnote { max-width: 1px; word-break: break-word; }") |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    pt_paired_html, "</div>\n```\n", sep = "")
```

### Exercise 11

The two Preceptor Tables are nearly identical. What structural differences distinguish the causal version from the predictive one?

```{r wisdom-11}
question_text(NULL,
  message = "Two differences. First, the predictive table has one outcome column (`SAT Score`); the causal table has two potential-outcome columns (`SAT if Tutored`, `SAT if Not Tutored`), with the unobservable counterfactual hatched in each row. Second, the causal table has a `Treatment` column under its own spanner, naming which potential outcome was actually realized for each junior. The predictive table has no treatment column. Everything else --- the units, the row count, the role of covariates --- is the same. That single difference in outcome-and-treatment bookkeeping is the entire mechanical difference between predictive and causal models.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

The same fitted model will eventually serve both questions. The thing that makes one *predictive* and the other *causal* is the analyst's commitment about what the covariates *are*, not anything in the data itself.


## Justice
###

> *Justice reviews the Population Table and selects the formula for the data generating mechanism.*

Justice is the section of the analysis where you (or your critics) raise concerns about whether your model will work. The Population Table is where those concerns become concrete: it bridges between the data we actually have and the Preceptor Table we wish we had.

### Exercise 1

In your own words, describe the components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
  message = "Justice reviews the Population Table and selects the formula for the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The four classic Justice concerns are *validity*, *stability*, *representativeness*, and *unconfoundedness*. The first three apply to both predictive and causal models; the fourth applies only to causal models. We work each in turn.

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Validity is about the *columns*. Two columns can have the same name and still measure different things.

### Exercise 3

Suppose the data we have is from a different high school's class of 2015. Give one reason why the assumption of validity might not hold for the `sat_score` column.

```{r justice-3}
question_text(NULL,
  message = "The College Board redesigned the SAT in 2016 (and again later). A pre-2016 SAT score is not directly comparable to a 2026 SAT score; the test sections, the scoring, and the curve have all shifted. So the `sat_score` column in the 2015 data and the `sat_score` column in our 2026 Preceptor Table may not measure the same construct, even though both are called 'SAT score'.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The Population Table is the bridge from the data to the Preceptor Table. Each row is one unit at one moment in time. The table includes the rows we have (the Data block), the rows we want (the Preceptor block), and the much larger universe of unit/time combinations both blocks are samples from.

### Exercise 5

What unit/time combination defines each row in the Population Table for our SAT questions?

```{r justice-5}
question_text(NULL,
  message = "Each row is one junior at one year. Data rows are juniors from a previous year (when SAT scores were collected). Preceptor rows are the rising juniors who will take the SAT next year. The broader population covers juniors at this and comparable high schools across many years.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Below is the Population Table for the **primary (predictive)** question.

```{r}
#| echo: false
#| results: asis
pop_primary_tibble <- tibble::tribble(
  ~Source,     ~`Junior`,           ~`Year`, ~`SAT Score`,
  "...",       "...",               "...",   "...",
  "Data",      "Patricia Chen",     "2015",  "1380",
  "Data",      "Robert Davis",      "2015",  "1240",
  "Data",      "...",               "...",   "...",
  "Data",      "Linda Whitfield",   "2015",  "1410",
  "...",       "...",               "...",   "...",
  "Preceptor", "Maria Alvarez",     "2026",  "1320",
  "Preceptor", "James Thompson",    "2026",  "1450",
  "Preceptor", "...",               "...",   "...",
  "Preceptor", "Sarah Williams",    "2026",  "1280",
  "...",       "...",               "...",   "..."
)

pop_primary_html <- gt::gt(pop_primary_tibble, id = "pop_primary") |>
  gt::tab_header(title = "Population Table --- Primary question") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",    columns = c(`Junior`, `Year`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span", columns = c(`SAT Score`)) |>
  gt::cols_align(align = "left",  columns = c(`Source`, `Junior`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `SAT Score`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_column_spanners()) |>
  gt::tab_style(style = gt::cell_text(align = "left"),
                locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_style(style = gt::cell_text(align = "right"),
                locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(
    footnote = "Data rows are juniors at a comparable high school in 2015; Preceptor rows are the rising juniors at our high school in 2026. Both blocks share the broader population of U.S. high-school juniors but differ in era, in school, and in the version of the SAT they took.",
    locations = gt::cells_title()
  ) |>
  gt::opt_css(css = "#pop_primary .gt_footnote { max-width: 1px; word-break: break-word; }") |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    pop_primary_html, "</div>\n```\n", sep = "")
```

### Exercise 6

In your own words, define the assumption of "stability."

```{r justice-6}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Stability is best understood as a statement about *parameters*, not distributions. The relationship between covariates and outcome in our model is captured by a handful of coefficients. Stability says those coefficients are the same in our 2015 data as they are for the 2026 Preceptor Table.

### Exercise 7

Give one reason why the assumption of stability might not hold for our primary SAT question.

```{r justice-7}
question_text(NULL,
  message = "Test preparation has changed substantially between 2015 and 2026. Online prep services like Khan Academy's free Official SAT Practice have become widespread and shifted the relationship between student background and final SAT score. So the parameters governing the SAT-score distribution in 2026 may differ from those in 2015 --- not just because mean scores have changed (which is a distribution shift, not a stability violation) but because the slopes connecting student characteristics to scores have shifted (which is a stability violation).",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

A common confusion is to point at any change between two eras and call it a stability violation. It is not. A change in the *level* of SAT scores between 2015 and 2026 is a distribution shift; that on its own does not violate stability. A change in the *slope* connecting a student covariate to their score is a parameter shift; that is what violates stability.

### Exercise 8

In your own words, define the assumption of "representativeness."

```{r justice-8}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

### Exercise 9

Give one reason why the assumption of representativeness might not hold between the 2015 data and the broader population of U.S. juniors.

```{r justice-9}
question_text(NULL,
  message = "The 2015 data is from a single, specific high school. That school's student demographics --- income, race, parental education, prior academic preparation --- are not a random sample of U.S. high-school juniors. If our 2015 data's school skewed wealthier or more academically prepared than the U.S. average, the model fit on that data will overstate expected SAT scores for our 2026 juniors if they come from a different demographic mix.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A non-representative sample does not *guarantee* a biased estimate --- the bias might wash out by chance. But chance is the only mechanism left to defend us, and we have no principled reason to expect it to work in our favor.

### The paired causal Population Table and unconfoundedness

The paired causal question requires its own Population Table --- same structure as the primary, but with two potential-outcome columns and a treatment column.

```{r}
#| echo: false
#| results: asis
pop_paired_tibble <- tibble::tribble(
  ~Source,     ~`Junior`,           ~`Year`, ~`SAT if Tutored`, ~`SAT if Not Tutored`, ~`Tutoring Class`,
  "...",       "...",               "...",   "...",             "...",                 "...",
  "Data",      "Patricia Chen",     "2015",  "1380",            "...",                 "Tutored",
  "Data",      "Robert Davis",      "2015",  "...",             "1240",                "Not Tutored",
  "Data",      "...",               "...",   "...",             "...",                 "...",
  "Data",      "Linda Whitfield",   "2015",  "1410",            "...",                 "Tutored",
  "...",       "...",               "...",   "...",             "...",                 "...",
  "Preceptor", "Maria Alvarez",     "2026",  "1380",            "1320",                "Tutored",
  "Preceptor", "James Thompson",    "2026",  "1500",            "1450",                "Not Tutored",
  "Preceptor", "...",               "...",   "...",             "...",                 "...",
  "Preceptor", "Sarah Williams",    "2026",  "1340",            "1280",                "Tutored",
  "...",       "...",               "...",   "...",             "...",                 "..."
)

pop_paired_html <- gt::gt(pop_paired_tibble, id = "pop_paired") |>
  gt::tab_header(title = "Population Table --- Paired causal question") |>
  gt::tab_spanner(label = "Unit/Time",          id = "unit_span",      columns = c(`Junior`, `Year`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",   columns = c(`SAT if Tutored`, `SAT if Not Tutored`)) |>
  gt::tab_spanner(label = "Treatment",          id = "treatment_span", columns = c(`Tutoring Class`)) |>
  gt::cols_align(align = "left",  columns = c(`Source`, `Junior`, `Tutoring Class`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `SAT if Tutored`, `SAT if Not Tutored`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_column_spanners()) |>
  gt::tab_style(style = gt::cell_text(align = "left"),
                locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span"))) |>
  gt::tab_style(style = gt::cell_text(align = "right"),
                locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  # Hatch Preceptor rows' unobservable counterfactuals.
  # Row 8 (Tutored): hatch `SAT if Not Tutored`.
  # Row 9 (Not Tutored): hatch `SAT if Tutored`.
  # Row 11 (Tutored): hatch `SAT if Not Tutored`.
  gt::text_transform(
    locations = gt::cells_body(columns = `SAT if Not Tutored`, rows = c(8, 11)),
    fn = function(x) hatch(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `SAT if Tutored`, rows = 9),
    fn = function(x) hatch(x)
  ) |>
  gt::tab_footnote(
    footnote = "Same row structure as the primary Population Table, with two outcome columns instead of one. In Data rows, only one potential outcome is observed --- the one matching the junior's actual treatment arm; the counterfactual is '...' because no one measured it. In Preceptor rows, both potential outcomes are filled in (the unobservable one hatched), because Preceptor rows show the unobservable truth.",
    locations = gt::cells_title()
  ) |>
  gt::opt_css(css = "#pop_paired .gt_footnote { max-width: 1px; word-break: break-word; }") |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    pop_paired_html, "</div>\n```\n", sep = "")
```

### Exercise 10

In your own words, define the assumption of "unconfoundedness."

```{r justice-10}
question_text(NULL,
  message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Unconfoundedness applies only to the paired causal question. For the predictive question we don't even ask --- predictive models have no treatment, so the assumption is not relevant.

### Exercise 11

Give one reason why the assumption of unconfoundedness might be hard to defend for the SAT tutoring causal question.

```{r justice-11}
question_text(NULL,
  message = "If the tutoring class was offered school-wide in some years and not others (rather than randomly assigned to individual students), the years in which it was offered may differ from the years in which it was not in ways that also affect SAT scores --- changes in college-admissions selectivity, in counselor staffing, in the prior PSAT performance of the entering cohort. Unless we condition on those pre-treatment covariates, treatment assignment (whether the year offered tutoring) is correlated with potential outcomes through paths the model doesn't see, and unconfoundedness fails.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

The cleanest defense of unconfoundedness is random assignment of treatment. If the principal had randomly assigned individual juniors to a tutoring section vs. a study-hall control, unconfoundedness would hold by construction. With a school-wide year-on-year decision, the assumption is much harder.

### Probability family and link function

Justice also commits us to the *formula for the data generating mechanism* --- in later tutorials we will give this its technical name (*probability family* and *link function*). The outcome `sat_score` is continuous, so the probability family is Normal and the link function is the identity:

$$Y \sim \mathcal{N}(\mu, \sigma^2)$$

with $\mu = \beta_0 + \beta_1 X_1 + \cdots + \beta_n X_n$.


## Courage
###

> *Courage creates the data generating mechanism.*

Justice settled the structural choices --- Normal family, identity link, four assumptions to defend. Courage now picks specific covariates and fits the model. The lightweight version: *if we had data and code*, here is what Courage would do.

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
  message = "Courage creates the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

### Exercise 2

For the primary (predictive) question, what would the fitted model look like in plain language?

```{r courage-2}
question_text(NULL,
  message = "We would model SAT score as a normally distributed variable which is a linear function of zero or more covariates --- in the simplest case, an intercept-only model that produces a single expected SAT score and a measure of its spread. With covariates added (PSAT score, hours studied, prior GPA), the model would produce per-junior expected SAT scores.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

### Exercise 3

For the paired causal question, what would the *fitted model* look like? (How does it differ from the predictive case?)

```{r courage-3}
question_text(NULL,
  message = "The fitted model uses the same code and the same coefficients as the predictive version --- the only difference is the question we ask of it. The model includes the treatment variable as a covariate, and we read its coefficient as the average causal effect of the tutoring class on SAT score (under unconfoundedness and the other Justice assumptions). The math is identical to the predictive case; what changes is the analyst's interpretation.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The takeaway: *predictive and causal models can share the same fit*. The difference between them lives in the analyst's commitments, not in the data or the code. This is the deepest insight of the Cardinal Virtues framework, and it is why we work both questions in parallel in this tutorial.


## Temperance
###

> *Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.*

Temperance is where the fitted model becomes the answer. The lightweight version: *if we had a fit*, here is how Temperance would read it.

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
  message = "Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

### Exercise 2

For the primary question, what would Temperance produce? Use the word "uncertainty."

```{r temperance-2}
question_text(NULL,
  message = "An estimate of the expected average SAT score for next year's juniors, along with its associated uncertainty (a 95% confidence interval). The point estimate alone is not an answer; the interval around it is essential.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

### Exercise 3

For the paired causal question, what would Temperance produce? Be careful about the *language* you use to describe the answer.

```{r temperance-3}
question_text(NULL,
  message = "An estimate of the average causal effect of the SAT tutoring class on a junior's score, with its associated uncertainty. Causal language --- 'causal effect,' 'effect of,' 'caused by' --- is appropriate here only because we have committed (in Justice) to defending unconfoundedness. If unconfoundedness fails, the same number must be interpreted predictively --- 'the difference in SAT scores between juniors in tutored and not-tutored years' --- not as an effect of tutoring.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

> *We can never know the truth.* The posterior we report is, at best, the *Preceptor's Posterior* --- the best posterior achievable under our assumptions. The reported confidence interval captures only sampling variation under those assumptions; it does not capture our uncertainty about whether the assumptions themselves are correct. The reported interval is always too narrow.


## Summary
###

This tutorial walked through the four Cardinal Virtues --- *Wisdom*, *Justice*, *Courage*, and *Temperance* --- by applying them to a paired pair of questions about high-school SAT scores: a predictive question (next year's average) and a causal one (the effect of a school-wide tutoring class). The exercises were conceptual, not coded. The takeaway is the framework itself: every data science problem proceeds through the same four-step arc, and the predictive/causal distinction is something the analyst commits to in Wisdom and defends in Justice --- not something the data dictates.

The example tutorials that follow (`05-recruits`, `06-trains`, `07-colleges`, ...) walk through the same arc using real data and `tidymodels` code. The hardest concepts in those tutorials --- writing down the Preceptor Table, defending the four assumptions, choosing the probability family --- are exactly the conceptual moves rehearsed here.

The world is always more uncertain than our models would have us believe.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
