---
title: "Shaming"
author: David Kane
tutorial:
  id: 11-shaming
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #11 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(tidymodels)
library(primer.data)
library(broom)
library(marginaleffects)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

x <- shaming |>
  mutate(civ_engage = primary_00 + primary_02 + primary_04 +
                      general_00 + general_02 + general_04) |>
  select(primary_06, treatment, sex, age, civ_engage) |>
  drop_na() |>
  mutate(voter_class = factor(case_when(
    civ_engage %in% c(5, 6) ~ "Always Vote",
    civ_engage %in% c(3, 4) ~ "Sometimes Vote",
    civ_engage %in% c(1, 2) ~ "Rarely Vote"),
    levels = c("Rarely Vote", "Sometimes Vote", "Always Vote"))) |>
  mutate(voted = as.factor(primary_06)) |>
  drop_na()

fit_shaming <- logistic_reg(engine = "glm") |>
  fit(voted ~ age + sex + treatment*voter_class, data = x)
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

The world confronts us. Make decisions we must.

Imagine that you are the data scientist for a campaign for Governor of Texas. The candidate has a big-picture goal --- win the November election --- and trusts you to figure out where the campaign's limited budget should go. Many causal and predictive models would inform that allocation: which voters are persuadable, which postcards or robocalls or door-knocks actually move turnout, which media markets are the most cost-effective, how late in the cycle each tactic still works. This tutorial builds just one of them: the causal effect of social-pressure get-out-the-vote postcards on whether a registered voter shows up at the polls. The estimate alone won't determine where to spend, but the postcard-effectiveness piece is one good input to the larger campaign-allocation decision. There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

The best way to ensure that students remember these concepts more than a few months after the course ends is spaced repetition. By tutorial 11, repetition is doing the heavy lifting; the spacing has stretched out.

### Exercise 2

You should be working inside a GitHub repo named `shaming`, opened in a Codespace from the [`PPBDS/codespace-starter`](https://github.com/PPBDS/codespace-starter) template. If you are not there yet, create that repo and open it in a Codespace now --- see the [package README](https://github.com/PPBDS/primer/tree/main/primer.tutorials#working-environments-and-repo-setup) for setup instructions or for working locally.

Once you're inside the `shaming` repo, create a new Quarto document titled `"Shaming"` with yourself as the author, render it, and save it as `analysis.qmd`.

Create a `.gitignore` file with `analysis_files` on the first line followed by a blank line. Save and push.

At the R prompt, run:

```
show_file(".gitignore")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Professionals keep their data science work in the cloud because laptops fail.

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.data)` in a new code chunk. Render the file. Add `#| message: false` to suppress the messages, and add the following to the YAML header to hide the code:

```
execute:
  echo: false
```

Render again. At the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    #| message: false\n    library(tidyverse)\n    library(primer.data)\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The render runs in a fresh R session, so it doesn't see anything you have loaded at your interactive prompt. The isolation is a feature in modern data-science workflows: many R sessions run in parallel (your interactive prompt, the QMD render, this tutorial, sometimes an AI agent on the side), each with its own workspace. Parallelism is the norm; non-interaction between worlds is what makes it work.

### Exercise 4

Place your cursor on the `library(tidyverse)` line. Use `Cmd/Ctrl + Enter` to send the line to the R prompt.

CP/CR.

```{r introduction-4}
question_text(NULL,
  message = "    > library(tidyverse)\n    ── Attaching core tidyverse packages ─ tidyverse 2.0.0 ─\n    ✔ dplyr     ✔ forcats   ✔ ggplot2  ✔ lubridate\n    ✔ purrr     ✔ readr     ✔ stringr  ✔ tibble     ✔ tidyr\n    ── Conflicts ─ tidyverse_conflicts() ─\n    ✖ dplyr::filter() masks stats::filter()\n    ✖ dplyr::lag()    masks stats::lag()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

By tutorial 11, the masking message is old hat: `dplyr::filter()` shadows `stats::filter()`, and that's the version you want. When you write reusable code (a package, a script someone else will source), use the explicit `dplyr::filter` so the result no longer depends on load order.

### Exercise 5

Place your cursor on the `library(primer.data)` line. Use `Cmd/Ctrl + Enter`.

CP/CR.

```{r introduction-5}
question_text(NULL,
  message = "    > library(primer.data)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The data we will use is `shaming`, a 344,084-row tibble from a famous 2006 randomized field experiment in Michigan. Voters were randomly assigned to receive one of four social-pressure postcards (or no postcard at all) before the August 2006 primary, and the experimenters then matched their assignments to the public voter file to see who actually showed up.

### Exercise 6

At the R prompt, type `?shaming` and paste the Description below.

```{r introduction-6}
question_text(NULL,
  message = "Voter records and randomized treatment assignments from the Michigan 2006 social-pressure mailing experiment by Gerber, Green, and Larimer (American Political Science Review, 2008). The treatment column records which of five conditions a voter was randomly assigned to: no postcard, or one of four postcards of escalating social pressure --- Civic Duty, Hawthorne (informing recipients they are being studied), Self (showing the recipient's own voting history), or Neighbors (showing the recipient's voting history alongside their neighbors').",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

The Gerber-Green-Larimer paper is among the most cited studies in modern political science. The headline finding: the Neighbors postcard --- which lists the voter's own past turnout *and* their neighbors' --- raised primary turnout by about 8 percentage points relative to no postcard, far more than any prior GOTV intervention had managed. The other postcards moved turnout less.

### Exercise 7

Define a causal effect.

```{r introduction-7}
question_text(NULL,
  message = "A causal effect is the difference between two potential outcomes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Because the postcard treatment takes five values, each registered voter has *five* potential outcomes: whether they would have voted under No Postcard, under Civic Duty, under Hawthorne, under Self, and under Neighbors. We observe only the one that actually happened. A causal effect is the difference between any two of those potential outcomes for the same voter --- typically, in this setting, the difference between *Neighbors* and *No Postcard*.

### Exercise 8

Voter turnout in primary elections is the broad topic of this tutorial. Given that topic, which variable in `shaming` should we use as our outcome variable?

```{r introduction-8}
question_text(NULL,
  message = "Our outcome variable is `primary_06` --- a 0/1 indicator of whether the registered voter cast a ballot in the August 2006 Michigan primary, the election the postcard intervention targeted. We will recode it as a factor named `voted` so logistic regression in tidymodels accepts it.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

We will use `primary_06` as our outcome. The plot below shows the average turnout in the August 2006 primary, broken out by treatment arm.

```{r}
#| echo: false
#| message: false
#| warning: false
x |>
  group_by(treatment) |>
  summarize(turnout = mean(primary_06), .groups = "drop") |>
  ggplot(aes(x = treatment, y = turnout, fill = treatment)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = scales::label_percent(accuracy = 0.1)(turnout)),
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 0.45)) +
  scale_x_discrete(limits = c("No Postcard", "Civic Duty", "Hawthorne",
                              "Self", "Neighbors")) +
  labs(
    title    = "August 2006 Primary Turnout by Postcard Treatment",
    subtitle = "Neighbors raises turnout by about 8 percentage points; the weaker postcards move it less",
    x        = NULL,
    y        = NULL,
    caption  = "Source: Gerber, Green, Larimer 2008 via primer.data::shaming"
  ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.y     = element_blank(),
        axis.ticks.y    = element_blank(),
        panel.grid      = element_blank())
```

### Exercise 9

Pick a single voter from the data. Specify any two values the treatment variable could take for that voter, guess the potential outcome under each, and compute the causal effect for that voter.

```{r introduction-9}
question_text(NULL,
  message = "Take one specific voter, say Lisa Anderson. The treatment variable could take any of five values for her; pick two: `Neighbors` and `No Postcard`. If she received the Neighbors postcard, she would have voted (potential outcome `Yes`). If she received no postcard, she would not have voted (potential outcome `No`). The causal effect on her turnout of Neighbors versus No Postcard is `Yes` minus `No`, or `1 - 0 = 1`: the postcard moves her from non-voter to voter. Both potential outcomes are guesses; we observe only the one corresponding to her actual treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

A causal effect is the difference between two potential outcomes. *"Difference"* doesn't have to mean *"subtraction"* --- many potential outcomes are not numbers --- but here both are 0/1, so subtraction works and gives one of three values: `-1`, `0`, or `+1`. With only five postcard arms, every voter has $\binom{5}{2} = 10$ possible pairwise causal effects; the one most political scientists care about is *Neighbors vs. No Postcard*.

### Exercise 10

Write a causal question connecting the outcome `voted` to `treatment`. The question should specify a particular number to compute.

```{r introduction-10}
question_text(NULL,
  message = "What is the average causal effect of the Neighbors postcard, relative to no postcard, on the probability that a registered voter casts a ballot in the August 2006 Michigan primary?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The treatment variable here records the *experimentally assigned* postcard, not whether the voter actually read it or whether the postcard physically arrived. That distinction matters --- a few percent of mail does not get delivered --- so the *intent-to-treat* effect we will estimate is somewhat smaller than the effect on voters who actually read the postcard. For a campaign manager allocating budget, the intent-to-treat effect is usually the right number anyway: you can buy postcards delivered, not postcards read.


## Wisdom
###

Data science starts with some broad questions and a data set which might help us to answer them.

Imagine that you are the data scientist for a campaign for Governor of Texas. The candidate has a big-picture goal --- win the November election --- and trusts you to figure out where the campaign's limited budget should go. Many causal and predictive models would inform that allocation: which voters are persuadable, which postcards or robocalls or door-knocks actually move turnout, which media markets are the most cost-effective, how late in the cycle each tactic still works. This tutorial builds just one of them: the causal effect of social-pressure get-out-the-vote postcards on whether a registered voter shows up at the polls. The estimate alone won't determine where to spend, but the postcard-effectiveness piece is one good input to the larger campaign-allocation decision. There are many decisions to make.

The specific question: *What is the average causal effect of the Neighbors postcard, relative to no postcard, on the probability that a registered voter casts a ballot in the August 2006 Michigan primary?*

We will work from the Gerber-Green-Larimer 2006 Michigan experiment, available in the `shaming` tibble of the **[primer.data](https://github.com/PPBDS/primer.data)** package.

Remember that Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.

### Exercise 1

Define a Preceptor Table.

```{r wisdom-1}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The Preceptor Table only includes the variables mentioned in the question. For a causal model with five treatment arms, that means five potential-outcome columns plus the treatment column, plus any covariate the question asks us to break the analysis on.

### Exercise 2

The Introduction plot showed average turnout by treatment, pooled across all voters. Now look at how that picture changes by *voter class* --- a label we will derive shortly from each voter's recent turnout history. Ask AI to write R code that computes mean turnout by treatment and by voter class, and plots it as a grouped bar chart faceted by voter class. Use `x |>` as the starting tibble (it is already in the tutorial's environment). Paste the code below.

```{r wisdom-2}
question_text(NULL,
  message = "    x |>\n      group_by(treatment, voter_class) |>\n      summarize(turnout = mean(primary_06), .groups = \"drop\") |>\n      ggplot(aes(x = treatment, y = turnout, fill = treatment)) +\n      geom_col(width = 0.7) +\n      scale_y_continuous(labels = scales::label_percent(),\n                         limits = c(0, 0.7)) +\n      scale_x_discrete(limits = c(\"No Postcard\", \"Civic Duty\", \"Hawthorne\",\n                                  \"Self\", \"Neighbors\")) +\n      facet_wrap(~ voter_class) +\n      labs(\n        title    = \"Turnout by Treatment, Faceted by Voter Class\",\n        subtitle = \"The Neighbors postcard helps Sometimes/Always voters most; Rarely Vote barely moves\",\n        x = NULL, y = NULL,\n        caption  = \"Source: Gerber, Green, Larimer 2008 via primer.data::shaming\"\n      ) +\n      theme_minimal() +\n      theme(legend.position = \"none\",\n            axis.text.x = element_text(angle = 30, hjust = 1))",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 18)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
x |>
  group_by(treatment, voter_class) |>
  summarize(turnout = mean(primary_06), .groups = "drop") |>
  ggplot(aes(x = treatment, y = turnout, fill = treatment)) +
  geom_col(width = 0.7) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 0.7)) +
  scale_x_discrete(limits = c("No Postcard", "Civic Duty", "Hawthorne",
                              "Self", "Neighbors")) +
  facet_wrap(~ voter_class) +
  labs(
    title    = "Turnout by Treatment, Faceted by Voter Class",
    subtitle = "The Neighbors postcard helps Sometimes/Always voters most; Rarely Vote barely moves",
    x        = NULL,
    y        = NULL,
    caption  = "Source: Gerber, Green, Larimer 2008 via primer.data::shaming"
  ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x     = element_text(angle = 30, hjust = 1))
```

### Exercise 3

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-3}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

This problem is causal with a five-arm treatment, so the Preceptor Table has *five* potential-outcome columns, one per treatment arm. Each row holds the truth: every potential outcome is filled in. In reality we observe only one column per row (the one corresponding to the postcard the voter actually received), so the other four are unobservable --- which we will mark with diagonal hatching when we draw the table.

### Exercise 4

What are the units for this problem?

```{r wisdom-4}
question_text(NULL,
  message = "Each row is one registered Michigan voter at the time of the August 2006 primary. The Gerber-Green-Larimer experiment covered the universe of Michigan registered voters who had voted in the November 2004 general election --- about 344,000 individuals.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Specifying the unit forces a question precise enough to answer. We model individual voters here, but a campaign manager actually wants to know about *aggregates* --- how many extra voters a particular postcard would mobilize across a list of 50,000 mailing addresses, not which specific voter.

### Exercise 5

What is the outcome variable for this problem?

```{r wisdom-5}
question_text(NULL,
  message = "Keep track of two outcome variables: the one in our Preceptor Table and the one in our data. In this case, both correspond to whether the voter cast a ballot in the August 2006 Michigan primary --- in the data, recorded as the binary `primary_06` (0 = did not vote, 1 = voted), which we will recode as a factor `voted` for tidymodels.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The outcome we really care about is whether a postcard moves a *campaign-relevant* voter (a voter targetable by mail in a specific election cycle). The 2006 Michigan primary stand-in for that is fine for academic measurement, but a 2026 Texas campaign manager would want updated turnout records, a more recent voter file, and an outcome closer to *general-election* turnout where the postcard treatment would actually be deployed.

### Exercise 6

What is a covariate which you think might be useful for this problem, regardless of whether or not it is in the data?

```{r wisdom-6}
question_text(NULL,
  message = "Voter engagement history --- how often the voter has shown up in past elections --- is a strong candidate. Habitual voters and rare voters are likely to respond very differently to a social-pressure postcard, because the postcard is essentially appealing to a sense of civic duty that habitual voters already feel. The `shaming` data records past primary and general turnout for 2000, 2002, and 2004; we will summarize those into a single `voter_class` covariate.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The term *covariates* is used in three ways: variables that *might* matter, variables we have data for, and variables we end up putting in the model. Here we are lucky --- the variable that should matter (voter engagement history) is exactly the variable the experimenters recorded.

### Exercise 7

What is the treatment for this problem?

```{r wisdom-7}
question_text(NULL,
  message = "The treatment is `treatment`, a five-level factor recording which postcard the voter was randomly assigned: `No Postcard` (the control), `Civic Duty` (a generic civic-duty appeal), `Hawthorne` (informing the voter they are being studied), `Self` (showing the voter their own past turnout record), or `Neighbors` (showing both the voter's record and their immediate neighbors' records). Random assignment was performed at the household level by the experimenters.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Random assignment is what makes this a causal study. With five arms randomized, *any* pair of arms gives a valid causal contrast (Self vs. No Postcard, Neighbors vs. Civic Duty, etc.). Most published analyses focus on Neighbors vs. No Postcard because it is the largest contrast, but the others are equally well-identified.

### Exercise 8

What moment in time does the Preceptor Table refer to?

```{r wisdom-8}
question_text(NULL,
  message = "The August 2006 Michigan primary --- the election the postcard treatment was designed to move. Postcards were mailed during the eleven days before the primary; turnout is recorded as of the close of polls on August 8, 2006.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

A Preceptor Table can never really refer to an exact instant in time, but for an election where the polls close at a specific hour, it comes closer than most. The campaign-decision question we are imagining (Texas in 2026) is twenty years downstream of this exact moment, which is going to come up in Justice when we discuss stability.

### Exercise 9

Describe in words the Preceptor Table for this problem.

```{r wisdom-9}
question_text(NULL,
  message = "The Preceptor Table has eight columns: one identifying each voter, five potential-outcome columns (`Voted if No Postcard`, `Voted if Civic Duty`, `Voted if Hawthorne`, `Voted if Self`, `Voted if Neighbors`), one treatment column, and one covariate column for `Voter Class`. Each row represents one Michigan registered voter at the time of the August 2006 primary. For each voter, only one of the five potential-outcome columns is the value actually observed; the other four are unobservable counterfactuals --- the votes the voter would have cast under treatments they did not receive.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

```{r}
#| echo: false
#| results: asis
hatch <- function(x) {
  paste0(
    '<span style="background-image: repeating-linear-gradient(',
    '45deg, rgba(0,0,0,0.12) 0 1px, transparent 1px 6px);',
    ' padding: 6px 10px; margin: -6px -10px; display: block;">',
    x, '</span>'
  )
}

pre_tibble <- tibble::tribble(
  ~`Voter`,           ~`Voted if No Postcard`, ~`Voted if Civic Duty`, ~`Voted if Hawthorne`, ~`Voted if Self`, ~`Voted if Neighbors`, ~`Treatment`,  ~`Voter Class`,
  "Lisa Anderson",    "No",                    "No",                   "Yes",                 "Yes",            "Yes",                 "Neighbors",   "Always Vote",
  "Marcus Williams",  "No",                    "No",                   "No",                  "No",             "Yes",                 "No Postcard", "Sometimes Vote",
  "...",              "...",                   "...",                  "...",                 "...",            "...",                 "...",         "...",
  "Sandra Patel",     "No",                    "No",                   "No",                  "Yes",            "Yes",                 "Self",        "Rarely Vote"
)

pre_title_footnote     <- "If all the information in this table were available, we could answer the question: What is the average causal effect of the Neighbors postcard, relative to no postcard, on the probability that a registered voter casts a ballot in the August 2006 Michigan primary?"
pre_units_footnote     <- "Each row is one Michigan registered voter at the time of the August 2006 primary. The example rows above use plausible names spanning the relevant range of voter classes; missing rows represent the rest of the universe of about 344,000 registered voters."
pre_outcome_footnote   <- "Whether the voter cast a ballot in the August 2006 primary, under each of the five possible treatment values. The cross-hatched cells in each row mark the four unobservable counterfactual potential outcomes --- the votes that voter would have cast under the four treatments they did not receive. Only the un-hatched cell in each row corresponds to a value we could ever observe."
pre_treatment_footnote <- "The five-level postcard treatment that the voter was randomly assigned: No Postcard (control), or one of four social-pressure postcards. The cell shows the value the voter actually received --- which determines which of the five potential-outcome columns is the observable one for that row."
pre_covariate_footnote <- "Voter Class summarizes the voter's recent turnout history --- the count of past primary and general elections (2000, 2002, 2004) the voter showed up for. Rarely Vote = 1-2 elections, Sometimes Vote = 3-4, Always Vote = 5-6. Voter class is shown because the question's broader context (campaign targeting) cares about whether different voter types respond differently to the same postcard."

pre_gt_html <- gt::gt(pre_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",                id = "unit_span",
                  columns = c(`Voter`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",
                  columns = c(`Voted if No Postcard`, `Voted if Civic Duty`,
                              `Voted if Hawthorne`, `Voted if Self`,
                              `Voted if Neighbors`)) |>
  gt::tab_spanner(label = "Treatment",           id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::tab_spanner(label = "Covariate",           id = "covariates_span",
                  columns = c(`Voter Class`)) |>
  gt::cols_align(align = "left",
                 columns = c(`Voter`, `Voted if No Postcard`, `Voted if Civic Duty`,
                             `Voted if Hawthorne`, `Voted if Self`,
                             `Voted if Neighbors`, `Treatment`, `Voter Class`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "outcome_span",
                                                       "treatment_span", "covariates_span"))
  ) |>
  # Hatch each row's four unobservable potential outcomes (everything except the
  # PO column matching the row's actual treatment).
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted if No Postcard`, `Voted if Civic Duty`,
                  `Voted if Hawthorne`, `Voted if Self`),
      rows = 1),
    fn = function(x) hatch(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted if Civic Duty`, `Voted if Hawthorne`,
                  `Voted if Self`, `Voted if Neighbors`),
      rows = 2),
    fn = function(x) hatch(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted if No Postcard`, `Voted if Civic Duty`,
                  `Voted if Hawthorne`, `Voted if Neighbors`),
      rows = 4),
    fn = function(x) hatch(x)
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_treatment_footnote,
                   locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 10

The model in this tutorial uses a derived covariate, `voter_class`, built by summing past primary and general turnout in 2000, 2002, and 2004 and binning the total into three categories. Ask AI to write R code that takes `shaming`, builds a `civ_engage` column from those six turnout columns, recodes it into a 3-level factor `voter_class`, recodes `primary_06` as a factor `voted`, drops missing values, and assigns the result to `x`. Paste the code below.

```{r wisdom-10}
question_text(NULL,
  message = "    x <- shaming |>\n      mutate(civ_engage = primary_00 + primary_02 + primary_04 +\n                          general_00 + general_02 + general_04) |>\n      select(primary_06, treatment, sex, age, civ_engage) |>\n      drop_na() |>\n      mutate(voter_class = factor(case_when(\n        civ_engage %in% c(5, 6) ~ \"Always Vote\",\n        civ_engage %in% c(3, 4) ~ \"Sometimes Vote\",\n        civ_engage %in% c(1, 2) ~ \"Rarely Vote\"),\n        levels = c(\"Rarely Vote\", \"Sometimes Vote\", \"Always Vote\"))) |>\n      mutate(voted = as.factor(primary_06)) |>\n      drop_na()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 12)
```

###

Two things worth flagging: setting the `levels` of `voter_class` explicitly forces `Rarely Vote` to be the reference category in the regression (rather than alphabetical, which would put `Always Vote` first). And the second `drop_na()` after `case_when()` catches any voter with `civ_engage = 0`, who would otherwise be coded `NA` --- a deliberate choice, since voters with zero past elections are mostly newly registered and respond to GOTV very differently from the rest.

### Exercise 11

Add the data-prep code chunk to your QMD. Place your cursor on the line that assigns to `x` and use `Cmd/Ctrl + Enter` to execute. `Cmd/Ctrl + Shift + K`. At the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r wisdom-11}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    x <- shaming |>\n      mutate(civ_engage = primary_00 + primary_02 + primary_04 +\n                          general_00 + general_02 + general_04) |>\n      select(primary_06, treatment, sex, age, civ_engage) |>\n      drop_na() |>\n      mutate(voter_class = factor(case_when(\n        civ_engage %in% c(5, 6) ~ \"Always Vote\",\n        civ_engage %in% c(3, 4) ~ \"Sometimes Vote\",\n        civ_engage %in% c(1, 2) ~ \"Rarely Vote\"),\n        levels = c(\"Rarely Vote\", \"Sometimes Vote\", \"Always Vote\"))) |>\n      mutate(voted = as.factor(primary_06)) |>\n      drop_na()\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 14)
```

###

Commit and push. From this point on, every modeling step uses the cleaned tibble `x`, not raw `shaming`.

### Exercise 12

Write the first two sentences of the summary paragraph. The first introduces the topic; the second introduces the data and the specific question.

```{r wisdom-12}
question_text(NULL,
  message = "Voter mobilization is a perennial campaign question: which messaging --- if any --- gets registered voters who would otherwise stay home to actually show up. Using the 344,084-voter Gerber-Green-Larimer 2006 randomized field experiment in Michigan, we estimate the average causal effect of the Neighbors postcard on the probability that a registered voter casts a ballot in the August 2006 primary.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Read our answer. It will not be the same as yours. Add your two sentences to the summary paragraph in your QMD. `Cmd/Ctrl + Shift + K`, then commit/push.


## Justice
###

Wisdom gives us the Preceptor Table and the data.

```{r}
#| echo: false
#| results: asis
hatch_j <- function(x) {
  paste0(
    '<span style="background-image: repeating-linear-gradient(',
    '45deg, rgba(0,0,0,0.12) 0 1px, transparent 1px 6px);',
    ' padding: 6px 10px; margin: -6px -10px; display: block;">',
    x, '</span>'
  )
}

pre_tibble_j <- tibble::tribble(
  ~`Voter`,           ~`Voted if No Postcard`, ~`Voted if Civic Duty`, ~`Voted if Hawthorne`, ~`Voted if Self`, ~`Voted if Neighbors`, ~`Treatment`,  ~`Voter Class`,
  "Lisa Anderson",    "No",                    "No",                   "Yes",                 "Yes",            "Yes",                 "Neighbors",   "Always Vote",
  "Marcus Williams",  "No",                    "No",                   "No",                  "No",             "Yes",                 "No Postcard", "Sometimes Vote",
  "...",              "...",                   "...",                  "...",                 "...",            "...",                 "...",         "...",
  "Sandra Patel",     "No",                    "No",                   "No",                  "Yes",            "Yes",                 "Self",        "Rarely Vote"
)

pre_gt_html_j <- gt::gt(pre_tibble_j, id = "preceptor_tbl_j") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",                id = "unit_span",
                  columns = c(`Voter`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",
                  columns = c(`Voted if No Postcard`, `Voted if Civic Duty`,
                              `Voted if Hawthorne`, `Voted if Self`,
                              `Voted if Neighbors`)) |>
  gt::tab_spanner(label = "Treatment",           id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::tab_spanner(label = "Covariate",           id = "covariates_span",
                  columns = c(`Voter Class`)) |>
  gt::cols_align(align = "left",
                 columns = c(`Voter`, `Voted if No Postcard`, `Voted if Civic Duty`,
                             `Voted if Hawthorne`, `Voted if Self`,
                             `Voted if Neighbors`, `Treatment`, `Voter Class`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "outcome_span",
                                                       "treatment_span", "covariates_span"))
  ) |>
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted if No Postcard`, `Voted if Civic Duty`,
                  `Voted if Hawthorne`, `Voted if Self`),
      rows = 1),
    fn = function(x) hatch_j(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted if Civic Duty`, `Voted if Hawthorne`,
                  `Voted if Self`, `Voted if Neighbors`),
      rows = 2),
    fn = function(x) hatch_j(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted if No Postcard`, `Voted if Civic Duty`,
                  `Voted if Hawthorne`, `Voted if Neighbors`),
      rows = 4),
    fn = function(x) hatch_j(x)
  ) |>
  gt::opt_css(
    css = "
      #preceptor_tbl_j .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html_j,
  "</div>\n",
  "```\n",
  sep = ""
)
```

```{r}
#| echo: false
#| results: asis
data_tibble <- tibble::tribble(
  ~`Voter`,           ~`Year`, ~`Voted`,  ~`Treatment`,  ~`Voter Class`,
  "Lisa Anderson",    "2006",  "Yes",     "Neighbors",   "Always Vote",
  "Marcus Williams",  "2006",  "No",      "No Postcard", "Sometimes Vote",
  "...",              "...",   "...",     "...",         "...",
  "Sandra Patel",     "2006",  "Yes",     "Self",        "Rarely Vote"
)

data_title_footnote     <- "Public Michigan voter file matched to randomized treatment assignments from Gerber, Green, and Larimer's 2006 social-pressure mailing experiment. Treatment was assigned at the household level by the experimenters; turnout in the August 2006 primary was matched in afterwards from the official voter file."
data_units_footnote     <- "Each row is one Michigan registered voter who was eligible for the experiment (had voted in the November 2004 general election). The original file has 344,084 voters; we work with the full file because logistic regression with our 17-parameter model fits in about a second."
data_outcome_footnote   <- "Whether the voter showed up at the polls for the August 2006 primary, recorded by the state and matched to the experimental sample. Self-reported turnout would be inflated; the public voter-file match is the more reliable measure."
data_treatment_footnote <- "The five-level postcard treatment, randomly assigned at the household level. About 56% of voters were in the No Postcard arm; the remaining 44% were spread across the four social-pressure arms, roughly equally."
data_covariate_footnote <- "Voter Class is derived from the voter's turnout in the 2000, 2002, and 2004 primaries and general elections (six elections total). Rarely Vote = 1-2 of those, Sometimes Vote = 3-4, Always Vote = 5-6. Voters with zero past turnout are dropped because they are mostly newly registered and respond to GOTV very differently."

data_gt_html <- gt::gt(data_tibble, id = "data_tbl") |>
  gt::tab_header(title = "Data: Michigan 2006 Shaming Experiment") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`Voter`, `Year`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Voted`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Voter Class`)) |>
  gt::cols_align(align = "left",  columns = c(`Voter`, `Voted`, `Treatment`, `Voter Class`)) |>
  gt::cols_align(align = "right", columns = c(`Year`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "outcome_span",
                                                       "treatment_span", "covariates_span"))
  ) |>
  gt::tab_footnote(footnote = data_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = data_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = data_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = data_treatment_footnote,
                   locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = data_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #data_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  data_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

Remember that Justice concerns the Population Table, the four key assumptions which underlie it (validity, stability, representativeness, and unconfoundedness), and selects the formula for the data generating mechanism.

### Exercise 1

In your own words, define "validity" as we use the term.

```{r justice-1}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Validity is always about the columns. Two columns sharing a name across tables is no guarantee that they measure the same thing.

### Exercise 2

Provide one reason why the assumption of validity might not hold for the outcome variable `voted` or for the treatment.

```{r justice-2}
question_text(NULL,
  message = "The `voted` column in the data is the *match* between a voter's randomization record and the public voter file. Voter-file matches are imperfect: a voter who moved between registration and the primary may show up as `No` (didn't vote) in the file even if they cast a ballot in their new precinct, and rare clerical errors run in both directions. The Preceptor Table column is meant to capture the truth (did the voter cast a ballot, anywhere); the data column captures whatever the public file recorded. The two are close but not identical.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Validity, if reasonable, is what lets us treat the Preceptor Table and the data as drawn from the same population. Voter-file matching is one of the better outcome measurements available in social science --- it does not depend on self-report at all --- but it is still imperfect.

### Exercise 3

In your own words, define a Population Table.

```{r justice-3}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The Population Table is almost always larger than the Preceptor Table and the data combined. In our case the Data block and the Preceptor block share the August 2006 timeframe almost exactly, so the two are close, but the broader population still spans years of Michigan registration data.

### Exercise 4

Specify the unit/time combinations which define each row in this Population Table.

```{r justice-4}
question_text(NULL,
  message = "Each row is one Michigan registered voter at one point in time. Data rows are the 344,084 voters from the Gerber-Green-Larimer experimental sample, observed in August 2006. Preceptor rows are the same population --- Michigan registered voters at the time of the August 2006 primary --- which means our Data and Preceptor blocks essentially share era and target population. The realistic interpretation gap is between Michigan in 2006 and the campaign-decision context the analyst cares about (e.g., Texas in 2026), which lives outside the immediate Population Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

```{r}
#| echo: false
#| results: asis
hatch_p <- function(x) {
  paste0(
    '<span style="background-image: repeating-linear-gradient(',
    '45deg, rgba(0,0,0,0.12) 0 1px, transparent 1px 6px);',
    ' padding: 6px 10px; margin: -6px -10px; display: block;">',
    x, '</span>'
  )
}

pop_tibble <- tibble::tribble(
  ~Source    , ~`Voter`,            ~`Year`, ~`Voted (No Pc)`, ~`Voted (Civic)`, ~`Voted (Haw.)`, ~`Voted (Self)`, ~`Voted (Neigh.)`, ~`Treatment`,    ~`Voter Class`,
  "..."      , "...",               "..."  , "...",            "...",            "...",           "...",           "...",             "...",           "...",
  "Data"     , "Lisa Anderson",     "2006" , "...",            "...",            "Yes",           "...",           "...",             "Hawthorne",     "Always Vote",
  "Data"     , "Marcus Williams",   "2006" , "No",             "...",            "...",           "...",           "...",             "No Postcard",   "Sometimes Vote",
  "Data"     , "...",               "..."  , "...",            "...",            "...",           "...",           "...",             "...",           "...",
  "Data"     , "Sandra Patel",      "2006" , "...",            "...",            "...",           "Yes",           "...",             "Self",          "Rarely Vote",
  "..."      , "...",               "..."  , "...",            "...",            "...",           "...",           "...",             "...",           "...",
  "Preceptor", "Lisa Anderson",     "2006" , "No",             "No",             "Yes",           "Yes",           "Yes",             "Hawthorne",     "Always Vote",
  "Preceptor", "Marcus Williams",   "2006" , "No",             "No",             "No",            "No",            "Yes",             "No Postcard",   "Sometimes Vote",
  "Preceptor", "...",               "..."  , "...",            "...",            "...",           "...",           "...",             "...",           "...",
  "Preceptor", "Sandra Patel",      "2006" , "No",             "No",             "No",            "Yes",           "Yes",             "Self",          "Rarely Vote",
  "..."      , "...",               "..."  , "...",            "...",            "...",           "...",           "...",             "...",           "..."
)

pop_title_footnote     <- "Population: Michigan registered voters at the time of the August 2006 primary, drawn from the broader universe of Michigan-eligible voters. The Data block records what the experimenters observed (one potential outcome per voter, given the postcard arm they were assigned to); the Preceptor block records the unobservable truth (all five potential outcomes filled in, with the four counterfactual ones hatched in the Preceptor Table proper). The two blocks share era and target population --- the gap is essentially zero."
pop_units_footnote     <- "Each row is one voter at one point in time. Data rows: 344,084 Michigan voters from the experimental sample, August 2006. Preceptor rows: the same population, with all five potential outcomes filled in."
pop_outcome_footnote   <- "Whether the voter cast a ballot in the August 2006 primary, under each of the five possible treatments. In Data rows, only the value under the voter's actual treatment is observed --- the other four are unknown (`...`). In Preceptor rows, all five are filled in (counterfactuals were hatched in the Preceptor Table proper). Column labels here are abbreviated to fit; full names are in the Preceptor Table above."
pop_treatment_footnote <- "The five-level postcard arm the voter was randomly assigned: No Postcard (control), Civic Duty, Hawthorne, Self, or Neighbors."
pop_covariate_footnote <- "Voter Class is the same derived covariate (Rarely Vote / Sometimes Vote / Always Vote) used in the Preceptor Table; same column meaning across both blocks because both refer to the same voters."

pop_gt_html <- gt::gt(pop_tibble, id = "population_tbl") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time",          id = "unit_span",
                  columns = c(`Voter`, `Year`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",
                  columns = c(`Voted (No Pc)`, `Voted (Civic)`, `Voted (Haw.)`,
                              `Voted (Self)`, `Voted (Neigh.)`)) |>
  gt::tab_spanner(label = "Treatment",          id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::tab_spanner(label = "Covariate",          id = "covariates_span",
                  columns = c(`Voter Class`)) |>
  gt::cols_align(align = "left",
                 columns = c(`Source`, `Voter`, `Voted (No Pc)`, `Voted (Civic)`,
                             `Voted (Haw.)`, `Voted (Self)`, `Voted (Neigh.)`,
                             `Treatment`, `Voter Class`)) |>
  gt::cols_align(align = "right", columns = c(`Year`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "outcome_span",
                                                       "treatment_span", "covariates_span"))
  ) |>
  # Hatch each Preceptor row's four unobservable potential outcomes.
  # Row 8 (Lisa, Hawthorne) → hatch No Pc / Civic / Self / Neigh.
  # Row 9 (Marcus, No Postcard) → hatch Civic / Haw. / Self / Neigh.
  # Row 11 (Sandra, Self) → hatch No Pc / Civic / Haw. / Neigh.
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted (No Pc)`, `Voted (Civic)`, `Voted (Self)`, `Voted (Neigh.)`),
      rows = 8),
    fn = function(x) hatch_p(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted (Civic)`, `Voted (Haw.)`, `Voted (Self)`, `Voted (Neigh.)`),
      rows = 9),
    fn = function(x) hatch_p(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted (No Pc)`, `Voted (Civic)`, `Voted (Haw.)`, `Voted (Neigh.)`),
      rows = 11),
    fn = function(x) hatch_p(x)
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_treatment_footnote,
                   locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 5

In your own words, define the assumption of "stability."

```{r justice-5}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Stability is a statement about *parameters*, not about distributions. The relationship between treatment, voter class, and turnout in our model is captured by a handful of coefficients --- intercept, slopes for treatment, slopes for voter class, interactions. Stability says those coefficients are the same in our 2006 Michigan data as they would be in any other Population Table row drawn from the same broader population.

### Exercise 6

Provide one reason why stability might not hold in this case.

```{r justice-6}
question_text(NULL,
  message = "If we want to use this model to inform a 2026 Texas campaign, the relevant Population Table extends to a different state and twenty years later --- and the slope of turnout on the Neighbors postcard may have shifted. Voters in 2006 Michigan had not seen the social-pressure-postcard format before; by 2026, after twenty years of campaigns copying the design, the parameter linking *Neighbors postcard* to turnout may be smaller because voters have habituated to the technique. So the slope on `treatmentNeighbors` --- not just the marginal turnout level --- may have changed between when the data was collected and when we want to apply it.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

Note the careful phrasing: it is not the *level* of turnout that's the stability concern (turnout has fallen and risen many times since 2006, and that, by itself, is just a distribution shift). What violates stability is a change in the *slope* connecting the Neighbors treatment to the outcome. A 2026 voter shrugging at a postcard wording that worked in 2006 *is* a stability violation; a 2026 voter being more or less likely to vote overall is not.

### Exercise 7

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: Data → Population → Preceptor Table. In your own words, define the assumption of "representativeness."

```{r justice-7}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The Gerber-Green-Larimer experiment is one of the cleaner cases for representativeness in field experimental work. Treatment was randomly assigned at the household level within an enumerated list of registered Michigan voters --- so within the Data block, every voter is, by construction, comparable across treatment arms. The harder representativeness question is whether the *whole sample* (Michigan registered voters who had voted in 2004) stands in for whatever broader population the analyst cares about.

### Exercise 8

Provide one reason, involving the relationship between the data and the population, why representativeness might not hold.

```{r justice-8}
question_text(NULL,
  message = "The experiment was restricted to voters who had cast a ballot in the November 2004 general election. That selection criterion excludes everyone who registered after 2004 (mostly young voters and recent movers) and everyone who was registered but skipped 2004 (a smaller, demographically distinct group). Our 344,084-voter sample is therefore representative of *2004-active Michigan voters* in 2006, not of the full Michigan registered-voter list. If a campaign wants to use the model to estimate the effect of a Neighbors postcard on first-time or low-turnout voters, the Data block doesn't have those voters in any meaningful number.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A non-representative sample does not *guarantee* a biased estimate --- the bias might wash out by chance. But chance is the only mechanism that would save us, and we have no principled reason to expect it. *"Might be biased"* is the correct stance, not *"will be biased."*

### Exercise 9

We use information about the population to make inferences about the Preceptor Table. Provide one reason, involving the relationship between the population and the Preceptor Table, why representativeness might not hold.

```{r justice-9}
question_text(NULL,
  message = "If the campaign-decision Preceptor Table is *Texas registered voters in 2026*, then the relevant population --- 2026 Texas voters --- shares neither the geography nor the era with the Michigan-2006 sample the data is drawn from. Even if our model perfectly captures the 2006 Michigan population, the Preceptor Table's voters live in a different state, twenty years later, with a different ground game, different demographics, and a different baseline turnout pattern. Representativeness from the broader U.S.-voter population to the Texas-2026 Preceptor Table is a separate question from representativeness from the Michigan-2006 sample to the broader U.S.-voter population, and both have to hold for the inference to flow through.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

Stability looks across time periods; representativeness looks within them. Both are about whether the rows we have stand in for the rows we care about, but they pull in different directions. Here, both are stretched: stability worries that 2006 isn't 2026, representativeness worries that Michigan isn't Texas. The campaign-decision context strains both assumptions in different ways.

### Exercise 10

In your own words, define the assumption of "unconfoundedness."

```{r justice-10}
question_text(NULL,
  message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

This assumption is only relevant for causal models, and the easiest way to ensure it is to assign treatment randomly --- which the experimenters did here. By design, treatment assignment is independent of *anything* about the voter, including potential outcomes, with no covariate adjustment required.

### Exercise 11

Provide one reason why the assumption of unconfoundedness might be weaker than it looks in this case.

```{r justice-11}
question_text(NULL,
  message = "Random assignment was at the *household* level, not the voter level. If two registered voters lived at the same address, they were assigned to the same treatment arm. That keeps the experimenter's protocol clean (no awkward conversations about why one spouse got a postcard and the other didn't), but it means that within a household, the treatment assignment of voter A is perfectly correlated with the treatment assignment of voter B --- which technically violates the independence assumption at the individual level. The standard fix is to cluster standard errors at the household, but our model doesn't do that.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

The great advantage of randomization is that it guarantees unconfoundedness for the unit at which assignment happens. Here that unit is the household, not the voter. A more careful analysis would cluster standard errors at the household level; our 95% confidence intervals are very slightly too narrow as a result, though the magnitude is small with 344K voters.

### Exercise 12

A statistical model consists of two parts: the probability family and the link function. The probability family is the distribution which generates the randomness in our data; the link function is the mathematical formula connecting the data to the unknown parameters.

Add `library(tidymodels)` to the QMD file. Place your cursor on that line. Use `Cmd/Ctrl + Enter`.

CP/CR.

```{r justice-12}
question_text(NULL,
  message = "    > library(tidymodels)\n    ── Attaching packages ──────── tidymodels 1.x.x ──\n    ✔ broom        ✔ recipes\n    ✔ dials        ✔ rsample\n    ✔ infer        ✔ tune\n    ✔ modeldata    ✔ workflows\n    ✔ parsnip      ✔ workflowsets\n    ✔ purrr        ✔ yardstick\n    ── Conflicts ──── tidymodels_conflicts() ──\n    ✖ purrr::discard() masks scales::discard()\n    ✖ dplyr::filter()  masks stats::filter()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Our outcome variable `voted` is binary. The probability family is therefore Bernoulli: $Y \sim \text{Bernoulli}(\rho)$, where $\rho$ is the probability the voter cast a ballot.

### Exercise 13

Add `library(broom)` to the QMD file. Place your cursor on that line. Use `Cmd/Ctrl + Enter`.

CP/CR.

```{r justice-13}
question_text(NULL,
  message = "    > library(broom)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The link function for a binary outcome is the logit: $\log\!\big(\rho/(1-\rho)\big) = \beta_0 + \beta_1 X_1 + \cdots + \beta_n X_n$. Our model will be a logistic regression with an interaction term.

### Exercise 14

Write one sentence highlighting a potential weakness in your model. So far our summary paragraph looks like this:

> Voter mobilization is a perennial campaign question: which messaging --- if any --- gets registered voters who would otherwise stay home to actually show up. Using the 344,084-voter Gerber-Green-Larimer 2006 randomized field experiment in Michigan, we estimate the average causal effect of the Neighbors postcard on the probability that a registered voter casts a ballot in the August 2006 primary.

Your version will be somewhat different.

```{r justice-14}
question_text(NULL,
  message = "One concern is that random assignment happened at the household, not the voter, level: two voters at the same address always shared a treatment arm, which slightly understates our standard errors and means our 95% confidence intervals are mildly optimistic.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a weakness sentence to your summary paragraph in your QMD. Don't copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`, then commit/push.


## Courage
###

Justice gives us the Population Table and the abstract data generating mechanism.

```{r}
#| echo: false
#| results: asis
hatch_c <- function(x) {
  paste0(
    '<span style="background-image: repeating-linear-gradient(',
    '45deg, rgba(0,0,0,0.12) 0 1px, transparent 1px 6px);',
    ' padding: 6px 10px; margin: -6px -10px; display: block;">',
    x, '</span>'
  )
}

pop_tibble_c <- tibble::tribble(
  ~Source    , ~`Voter`,            ~`Year`, ~`Voted (No Pc)`, ~`Voted (Civic)`, ~`Voted (Haw.)`, ~`Voted (Self)`, ~`Voted (Neigh.)`, ~`Treatment`,    ~`Voter Class`,
  "..."      , "...",               "..."  , "...",            "...",            "...",           "...",           "...",             "...",           "...",
  "Data"     , "Lisa Anderson",     "2006" , "...",            "...",            "Yes",           "...",           "...",             "Hawthorne",     "Always Vote",
  "Data"     , "Marcus Williams",   "2006" , "No",             "...",            "...",           "...",           "...",             "No Postcard",   "Sometimes Vote",
  "Data"     , "...",               "..."  , "...",            "...",            "...",           "...",           "...",             "...",           "...",
  "Data"     , "Sandra Patel",      "2006" , "...",            "...",            "...",           "Yes",           "...",             "Self",          "Rarely Vote",
  "..."      , "...",               "..."  , "...",            "...",            "...",           "...",           "...",             "...",           "...",
  "Preceptor", "Lisa Anderson",     "2006" , "No",             "No",             "Yes",           "Yes",           "Yes",             "Hawthorne",     "Always Vote",
  "Preceptor", "Marcus Williams",   "2006" , "No",             "No",             "No",            "No",            "Yes",             "No Postcard",   "Sometimes Vote",
  "Preceptor", "...",               "..."  , "...",            "...",            "...",           "...",           "...",             "...",           "...",
  "Preceptor", "Sandra Patel",      "2006" , "No",             "No",             "No",            "Yes",           "Yes",             "Self",          "Rarely Vote",
  "..."      , "...",               "..."  , "...",            "...",            "...",           "...",           "...",             "...",           "..."
)

pop_gt_html_c <- gt::gt(pop_tibble_c, id = "population_tbl_c") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time",          id = "unit_span",
                  columns = c(`Voter`, `Year`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",
                  columns = c(`Voted (No Pc)`, `Voted (Civic)`, `Voted (Haw.)`,
                              `Voted (Self)`, `Voted (Neigh.)`)) |>
  gt::tab_spanner(label = "Treatment",          id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::tab_spanner(label = "Covariate",          id = "covariates_span",
                  columns = c(`Voter Class`)) |>
  gt::cols_align(align = "left",
                 columns = c(`Source`, `Voter`, `Voted (No Pc)`, `Voted (Civic)`,
                             `Voted (Haw.)`, `Voted (Self)`, `Voted (Neigh.)`,
                             `Treatment`, `Voter Class`)) |>
  gt::cols_align(align = "right", columns = c(`Year`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "outcome_span",
                                                       "treatment_span", "covariates_span"))
  ) |>
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted (No Pc)`, `Voted (Civic)`, `Voted (Self)`, `Voted (Neigh.)`),
      rows = 8),
    fn = function(x) hatch_c(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted (Civic)`, `Voted (Haw.)`, `Voted (Self)`, `Voted (Neigh.)`),
      rows = 9),
    fn = function(x) hatch_c(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(
      columns = c(`Voted (No Pc)`, `Voted (Civic)`, `Voted (Haw.)`, `Voted (Neigh.)`),
      rows = 11),
    fn = function(x) hatch_c(x)
  ) |>
  gt::opt_css(
    css = "
      #population_tbl_c .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html_c,
  "</div>\n",
  "```\n",
  sep = ""
)
```

$$P(Y = 1) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n)}}$$

with $Y \sim \text{Bernoulli}(\rho)$, where $\rho$ is the probability above.

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
  message = "Courage creates the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Justice settled the structural choices --- binary outcome, Bernoulli family, logit link, randomized treatment. Courage now picks the specific covariates, fits the model, and extracts the parameter values.

### Exercise 2

Because our outcome variable is binary, start to create the model by entering `logistic_reg(engine = "glm")`.

```{r courage-2, exercise = TRUE}

```

```{r courage-2-hint-1, eval = FALSE}
logistic_reg(engine = ...)
```

```{r courage-2-test, include = FALSE}
logistic_reg(engine = "glm")
```

###

**[tidymodels](https://www.tidymodels.org/)** is the most popular framework for estimating models among R users. The `logistic_reg()` function is the binary-outcome counterpart to `linear_reg()` --- different probability family, different link, same pipeline.

### Exercise 3

Continue the pipe to `fit(voted ~ treatment, data = x)`. This is the simplest causal model: outcome on the multi-arm treatment, no covariates.

```{r courage-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-3-hint-1, eval = FALSE}
... |>
  fit(voted ~ treatment, data = x)
```

```{r courage-3-test, include = FALSE}
logistic_reg(engine = "glm") |>
  fit(voted ~ treatment, data = x)
```

###

A multi-level categorical predictor like `treatment` (5 levels) becomes 4 dummy variables in the model, each one named for the level it represents. The reference level is whichever level is *first* in the factor's `levels` --- here `No Postcard` --- and the four estimated coefficients are interpreted relative to that reference.

### Exercise 4

Here is the parameter table for this first model:

```{r}
#| echo: false
logistic_reg(engine = "glm") |>
  fit(voted ~ treatment, data = x) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(v) round(v, 3)))
```

Write a sentence interpreting the `treatmentNeighbors` estimate (about 0.43).

```{r courage-4}
question_text(NULL,
  message = "Comparing two voters who differ only in postcard treatment, the log-odds of voting are about 0.43 higher for a voter who received the Neighbors postcard than for a voter who received no postcard. The 95% confidence interval excludes zero, so we are confident the average causal effect of Neighbors versus No Postcard is positive on the log-odds scale; we will translate that to a probability-scale answer in Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Causal language --- *causal effect* --- is appropriate here because random assignment of the postcard treatment earns it. Note that the four `treatment*` coefficients are all *vs. No Postcard*; if you want, say, *Neighbors vs. Self* you would either refit with a different reference category or use `marginaleffects::comparisons()` to compute the contrast directly. Log-odds remains a hard scale to read on its own: a coefficient of 0.43 is a substantial effect, but exactly how big in turnout terms requires the inverse-logit transformation we'll do in Temperance.

### Exercise 5

Now add an adjustment covariate: `fit(voted ~ treatment + age, data = x)`. In an observational study, adjusting for a real covariate often moves the treatment estimate substantially, because adjustment is how we deal with confounding. In a randomized experiment, treatment assignment is by construction independent of any pre-treatment covariate, so adjustment should leave the treatment estimate essentially unchanged. Let's check.

```{r courage-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-5-hint-1, eval = FALSE}
... |>
  fit(voted ~ treatment + age, data = x)
```

```{r courage-5-test, include = FALSE}
logistic_reg(engine = "glm") |>
  fit(voted ~ treatment + age, data = x)
```

###

If the experiment ran the way it was supposed to, age and treatment are uncorrelated; the `age` slope will be precisely estimated (older voters turn out more) and the treatment estimates will barely budge.

### Exercise 6

Here is the parameter table for the adjusted model:

```{r}
#| echo: false
logistic_reg(engine = "glm") |>
  fit(voted ~ treatment + age, data = x) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(v) round(v, 3)))
```

Write a sentence interpreting the `age` slope and noting how the `treatmentNeighbors` estimate moves under adjustment.

```{r courage-6}
question_text(NULL,
  message = "Comparing two voters in the same treatment arm differing by one year of age, the log-odds of voting are about 0.014 higher for the older voter --- a small per-year effect that compounds: a 70-year-old has roughly 0.014 * 50 = 0.70 higher log-odds than a 20-year-old. The treatmentNeighbors estimate barely moves between the unadjusted and adjusted models (about 0.43 in both), which is exactly what we'd expect from a randomized experiment: random assignment guarantees treatment is balanced on age, so adjustment doesn't change the treatment estimate.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

The fact that the treatment estimate is robust to adjustment is itself evidence that randomization worked. In an observational study where we worry about confounding, we would expect adjustment to *change* the estimate (and we'd be uncomfortable if it didn't). In a randomized experiment, we expect adjustment to *not* change the estimate (and we'd be uncomfortable if it did). The language for `age` is comparison language, not causal: `age` is not the variable whose assignment was randomized, so we say *"comparing two voters of different ages,"* not *"causing turnout to rise."*

### Exercise 7

Now fit the final model: `fit(voted ~ age + sex + treatment*voter_class, data = x)`. The `treatment*voter_class` term means *treatment effects are allowed to vary by voter class* --- the kind of heterogeneity the EDA plot in Wisdom hinted at.

```{r courage-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-7-hint-1, eval = FALSE}
... |>
  fit(voted ~ age + sex + treatment*voter_class, data = x)
```

```{r courage-7-test, include = FALSE}
logistic_reg(engine = "glm") |>
  fit(voted ~ age + sex + treatment*voter_class, data = x)
```

###

The `treatment*voter_class` shorthand expands to `treatment + voter_class + treatment:voter_class` --- the two main effects plus all 4 × 2 = 8 interaction terms. Add the intercept and the two unrelated covariates (`age`, `sex`) and the model has 17 parameters. With 344,084 voters, every coefficient is precisely estimated even at this many parameters.

### Exercise 8

Here is the parameter table for the final model:

```{r}
#| echo: false
fit_shaming |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(v) round(v, 3)))
```

Interpretation in an interaction model is *conditional*: the effect of one variable depends on the value of the other. Write a sentence interpreting the `treatmentNeighbors` main effect of about 0.36 in the final model, being careful to describe what voter class it applies to.

```{r courage-8}
question_text(NULL,
  message = "The `treatmentNeighbors` main effect of about 0.36 is the log-odds difference between Neighbors and No Postcard for a Rarely Vote voter (the reference level of voter_class), adjusting for age and sex. For Sometimes Vote and Always Vote voters, you have to add the relevant `treatmentNeighbors:voter_class*` interaction term to get the conditional effect: the Neighbors-vs-No-Postcard log-odds difference is about 0.36 + 0.04 = 0.40 for Sometimes Vote, and about 0.36 + 0.01 = 0.37 for Always Vote --- broadly similar, with Sometimes Vote responding marginally more strongly than the other two voter classes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

In an interaction model, the main effect of `treatment` is not *the* treatment effect --- it's the treatment effect *at the reference level of the moderator*. Reading off heterogeneous treatment effects from a parameter table requires manually adding the appropriate interaction coefficients, which is error-prone. The cleaner approach is `marginaleffects::avg_comparisons(..., by = "voter_class")` --- which we'll use in Temperance to read these effects on the probability scale rather than the log-odds scale.

### Exercise 9

Behind the scenes, an object called `fit_shaming` has been created which is the result of the final model. Type `fit_shaming` and hit "Run Code."

```{r courage-9, exercise = TRUE}

```

```{r courage-9-hint-1, eval = FALSE}
fit_shaming
```

```{r courage-9-test, include = FALSE}
fit_shaming
```

###

In data science, **we deal with words, math, and code, but the most important of these is code.**

### Exercise 10

We need `fit_shaming` to exist in R World. Copy/paste this code to the R prompt and execute it:

```
fit_shaming <- logistic_reg(engine = "glm") |>
  fit(voted ~ age + sex + treatment*voter_class, data = x)
```

CP/CR.

```{r courage-10}
question_text(NULL,
  message = "    > fit_shaming <- logistic_reg(engine = \"glm\") |>\n    +   fit(voted ~ age + sex + treatment*voter_class, data = x)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Just because the fit object exists in this tutorial does not mean it exists at your R prompt. QMD World, R World, and the tutorial's R session are three separate workspaces; keep them in sync.

### Exercise 11

At the R prompt, load the **[easystats](https://easystats.github.io/easystats/)** package.

CP/CR.

```{r courage-11}
question_text(NULL,
  message = "    > library(easystats)\n    # Attaching packages: easystats 0.x.x (red = needs update)\n    ✔ bayestestR     ✔ correlation     ✔ datawizard \n    ✔ effectsize    ✔ insight         ✔ modelbased \n    ✔ parameters    ✔ performance     ✔ report \n    ✔ see",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

We don't add `easystats` to the QMD because we are only using it for an interactive check of our fitted model. Its [check_predictions()](https://easystats.github.io/performance/reference/check_predictions.html) function expects a base-R `glm` object, not the `parsnip` wrapper that `tidymodels` produces, so we will pass it `extract_fit_engine(fit_shaming)`.

### Exercise 12

At the R prompt, run:

```
check_predictions(extract_fit_engine(fit_shaming))
```

CP/CR.

```{r courage-12}
question_text(NULL,
  message = "    > check_predictions(extract_fit_engine(fit_shaming))\n\n(A plot opens. Green = the actual outcome distribution; blue = simulated draws from the fitted DGM. For a binary outcome, the comparison is between the share of `1` in the data --- about 32% turnout overall in the experiment --- and the share simulated by the model. With 344K rows and a fitted model that includes the main predictors of turnout, the two should match almost exactly.)",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

`check_predictions()` compares your actual data with data simulated from your fitted model. For a binary outcome with this much data, the check is mostly a sanity check --- it confirms the model isn't catastrophically miscalibrated. It does *not* confirm the model is well-specified: a more thoughtful diagnostic would split the comparison by voter class or by treatment to see whether the model's calibration is uniform.

### Exercise 13

With the parameters estimated, we can write the fitted model as a concrete formula:

$$\widehat{P(\text{Voted} = 1)} = \frac{1}{1 + e^{-\eta}}$$

where the linear predictor $\eta$ is

$$\begin{aligned}
\eta = \, & -2.43 + 0.012 \cdot \text{Age} + 0.036 \cdot \text{sexMale} \\
&{} + 0.086 \cdot \text{Civic Duty} + 0.071 \cdot \text{Hawthorne} \\
&{} + 0.202 \cdot \text{Self} + 0.359 \cdot \text{Neighbors} \\
&{} + 0.820 \cdot \text{Sometimes Vote} + 1.610 \cdot \text{Always Vote} \\
&{} + 0.031 \cdot (\text{Civic Duty} \times \text{Sometimes}) + 0.059 \cdot (\text{Hawthorne} \times \text{Sometimes}) \\
&{} + 0.053 \cdot (\text{Self} \times \text{Sometimes}) + 0.041 \cdot (\text{Neighbors} \times \text{Sometimes}) \\
&{} - 0.051 \cdot (\text{Civic Duty} \times \text{Always}) + 0.063 \cdot (\text{Hawthorne} \times \text{Always}) \\
&{} - 0.006 \cdot (\text{Self} \times \text{Always}) + 0.006 \cdot (\text{Neighbors} \times \text{Always})
\end{aligned}$$

Each `treatment*` term is a 0/1 dummy: 1 if the voter received that postcard, 0 otherwise. Each `voter_class*` term is also a dummy. The interaction terms are products of two dummies --- 1 only if the voter has both that treatment and that voter class. The intercept (-2.43) is the log-odds for a 0-year-old female Rarely Vote voter assigned No Postcard --- mathematically defined but not a sensible prediction; the intercept is a reference point, not a forecast.

**This is our data generating mechanism.**

###

A data generating mechanism is just a formula, something we can write down and implement with computer code. The non-linearity ($1/(1 + e^{-\eta})$) and the seventeen parameters together are why coefficient-by-coefficient interpretation is awkward; `marginaleffects` in Temperance gives us the cleaner reading on the outcome scale.

### Exercise 14

Create a new code chunk in your QMD with `#| cache: true`. Copy/paste the code for estimating the model:

```
fit_shaming <- logistic_reg(engine = "glm") |>
  fit(voted ~ age + sex + treatment*voter_class, data = x)
```

`Cmd/Ctrl + Shift + K`. At the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r courage-14}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    #| cache: true\n    fit_shaming <- logistic_reg(engine = \"glm\") |>\n      fit(voted ~ age + sex + treatment*voter_class, data = x)\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

`#| cache: true` tells Quarto to reuse the saved fit on subsequent renders. Re-render with `Cmd/Ctrl + Shift + K` --- it should be quick the second time. Even at 344K rows our fit takes only about a second, but caching is the right habit to build for the heavier models you'll meet later in the curriculum.

### Exercise 15

Add `*_cache` to your `.gitignore`. Cached objects don't belong on GitHub.

At the R prompt, run:

```
show_file(".gitignore")
```

CP/CR.

```{r courage-15}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files\n    *_cache",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Commit and push.

### Exercise 16

At the R prompt, run `tidy()` on `fit_shaming` with `conf.int = TRUE`.

```{r courage-16, exercise = TRUE}

```

```{r courage-16-hint-1, eval = FALSE}
tidy(fit_shaming, conf.int = TRUE)
```

```{r courage-16-test, include = FALSE}
tidy(fit_shaming, conf.int = TRUE)
```

###

`tidy()` is part of **[broom](https://broom.tidymodels.org/)**, used to summarize information from a wide variety of models. The `estimate`, `conf.low`, and `conf.high` columns are all on the log-odds scale --- the same scale as the model's link function.

### Exercise 17

Create a new code chunk in your QMD. Ask AI to make a nice-looking table from the tibble returned by `tidy()`. Insert your table code, `Cmd/Ctrl + Shift + K`, and at the R prompt run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r courage-17}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    tidy(fit_shaming, conf.int = TRUE) |>\n      select(term, estimate, conf.low, conf.high) |>\n      knitr::kable(\n        caption = \"Logistic-regression estimates (log-odds scale) and 95% CIs.  Source: Gerber, Green, Larimer 2008 via primer.data::shaming\",\n        digits = 3\n      )\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

A nice table includes a title, a caption with the data source, and reasonable rounding. For a 17-parameter logistic regression, three decimal places keeps the small interaction coefficients visible without over-precising the larger main-effect ones.

### Exercise 18

Add a sentence to your project summary explaining the model structure.

```{r courage-18}
question_text(NULL,
  message = "We model voter turnout as a Bernoulli variable which is a logistic function of treatment, voter class, the treatment-by-voter-class interaction, age, and sex.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add your sentence to the summary paragraph. `Cmd/Ctrl + Shift + K`, then commit/push.


## Temperance
###

Courage provides the data generating mechanism. We can express the DGM in four ways:

**In words:** We model voter turnout as a Bernoulli variable which is a logistic function of treatment, voter class, the treatment-by-voter-class interaction, age, and sex.

**In R code:**

```r
fit_shaming <- logistic_reg(engine = "glm") |>
  fit(voted ~ age + sex + treatment*voter_class, data = x)
```

**In a parameter table:**

```{r}
#| echo: false
fit_shaming |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(v) round(v, 3)))
```

**As a mathematical formula:**

$$\widehat{P(\text{Voted} = 1)} = \frac{1}{1 + e^{-\eta}}$$

where the linear predictor $\eta$ is

$$\begin{aligned}
\eta = \, & -2.43 + 0.012 \cdot \text{Age} + 0.036 \cdot \text{sexMale} \\
&{} + 0.086 \cdot \text{Civic Duty} + 0.071 \cdot \text{Hawthorne} \\
&{} + 0.202 \cdot \text{Self} + 0.359 \cdot \text{Neighbors} \\
&{} + 0.820 \cdot \text{Sometimes Vote} + 1.610 \cdot \text{Always Vote} \\
&{} + 0.031 \cdot (\text{Civic Duty} \times \text{Sometimes}) + 0.059 \cdot (\text{Hawthorne} \times \text{Sometimes}) \\
&{} + 0.053 \cdot (\text{Self} \times \text{Sometimes}) + 0.041 \cdot (\text{Neighbors} \times \text{Sometimes}) \\
&{} - 0.051 \cdot (\text{Civic Duty} \times \text{Always}) + 0.063 \cdot (\text{Hawthorne} \times \text{Always}) \\
&{} - 0.006 \cdot (\text{Self} \times \text{Always}) + 0.006 \cdot (\text{Neighbors} \times \text{Always})
\end{aligned}$$

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
  message = "Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Courage gave us the DGM, including its 17 estimated parameters. Temperance puts those parameters to work --- not by interpreting them coefficient-by-coefficient (a hopeless task on the log-odds scale with eight interaction terms), but by using **[marginaleffects](https://marginaleffects.com/)** to translate the model into outcome-scale answers. The companion book is *[Model to Meaning](https://marginaleffects.com/)* by Vincent Arel-Bundock.

### Exercise 2

Add `library(marginaleffects)` to the QMD file. Place your cursor on that line. Use `Cmd/Ctrl + Enter`.

CP/CR.

```{r temperance-2}
question_text(NULL,
  message = "    > library(marginaleffects)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

In the end, we don't really care about parameters, much less how to interpret seventeen of them. Parameters are *imaginary*, like unicorns. We care about answers to our questions, on the outcome scale.

### Exercise 3

What is the specific question we are trying to answer?

```{r temperance-3}
question_text(NULL,
  message = "What is the average causal effect of the Neighbors postcard, relative to no postcard, on the probability that a registered voter casts a ballot in the August 2006 Michigan primary?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

A specific question must specify a number to compute. *"How does the postcard treatment affect turnout?"* is too vague. *"The average causal effect of Neighbors versus No Postcard on the probability of voting"* is one specific number we can produce.

### Exercise 4

`avg_comparisons()` computes average causal effects on the outcome scale. At the R prompt, run:

```
avg_comparisons(extract_fit_engine(fit_shaming), variables = "treatment")
```

The `extract_fit_engine()` step is required because `marginaleffects` needs the underlying `glm` object, not the `parsnip` wrapper.

CP/CR.

```{r temperance-4}
question_text(NULL,
  message = "    > avg_comparisons(extract_fit_engine(fit_shaming), variables = \"treatment\")\n\n                     Contrast Estimate Std. Error     z Pr(>|z|)     S  2.5 % 97.5 %\n     Civic Duty - No Postcard   0.0183    0.00253  7.24   <0.001  41.0 0.0133 0.0232\n     Hawthorne - No Postcard    0.0254    0.00253 10.04   <0.001  76.4 0.0204 0.0303\n     Neighbors - No Postcard    0.0811    0.00261 31.08   <0.001 702.2 0.0760 0.0862\n     Self - No Postcard         0.0482    0.00257 18.76   <0.001 258.5 0.0432 0.0533\n\n    Term: treatment\n    Type: response",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 12)
```

```{r temperance-4-test, include = FALSE}
avg_comparisons(extract_fit_engine(fit_shaming), variables = "treatment")
```

###

Each row is the average difference in *predicted probability of voting* between that treatment arm and No Postcard, on the response scale (not the log-odds scale of the parameter table). Reading the four contrasts: Neighbors raises turnout by about 8.1 percentage points; Self by about 4.8; Hawthorne by about 2.5; Civic Duty by about 1.8. *Percentage points*, not *percent*: a difference of 8.1 percentage points (from 30% to 38.1% baseline turnout) is about a 27% relative increase, but those are different numbers.

### Exercise 5

Now break the comparisons down by voter class. At the R prompt, run:

```
avg_comparisons(extract_fit_engine(fit_shaming),
                variables = "treatment",
                by = "voter_class")
```

CP/CR.

```{r temperance-5}
question_text(NULL,
  message = "    > avg_comparisons(extract_fit_engine(fit_shaming),\n    +                 variables = \"treatment\",\n    +                 by = \"voter_class\")\n\n                     Contrast    voter_class Estimate Std. Error\n     Civic Duty - No Postcard Rarely Vote     0.00948    0.00500\n     Hawthorne - No Postcard  Rarely Vote     0.00774    0.00504\n     Neighbors - No Postcard  Rarely Vote     0.04369    0.00549\n     Self - No Postcard       Rarely Vote     0.02327    0.00519\n     Civic Duty - No Postcard Sometimes Vote  0.02382    0.00323\n     Hawthorne - No Postcard  Sometimes Vote  0.02649    0.00322\n     Neighbors - No Postcard  Sometimes Vote  0.08590    0.00334\n     Self - No Postcard       Sometimes Vote  0.05355    0.00330\n     Civic Duty - No Postcard Always Vote     0.00859    0.00577\n     Hawthorne - No Postcard  Always Vote     0.03301    0.00578\n     Neighbors - No Postcard  Always Vote     0.09041    0.00579\n     Self - No Postcard       Always Vote     0.04872    0.00576\n\n    Term: treatment\n    Type: response",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 18)
```

```{r temperance-5-test, include = FALSE}
avg_comparisons(extract_fit_engine(fit_shaming),
                variables = "treatment",
                by = "voter_class")
```

###

This is the heterogeneous-treatment-effect picture. The Neighbors postcard moves *Rarely Vote* voters by only about 4.4 percentage points, but moves *Sometimes Vote* and *Always Vote* voters by 8.6 and 9.0 percentage points respectively. Counterintuitively, the postcard has a *smaller* effect on the people who needed a nudge most (Rarely Vote) and a *larger* effect on people who were already inclined to vote anyway. The campaign-allocation implication is meaningful: targeting the Sometimes/Always voters with Neighbors is the most cost-effective use of a fixed mailing budget.

### Exercise 6

`plot_predictions(condition = c(...))` draws the model's predicted probabilities across two covariates simultaneously. At the R prompt, run:

```
plot_predictions(extract_fit_engine(fit_shaming),
                 condition = c("treatment", "voter_class"))
```

CP/CR.

```{r temperance-6}
question_text(NULL,
  message = "    > plot_predictions(extract_fit_engine(fit_shaming),\n    +                  condition = c(\"treatment\", \"voter_class\"))\n\n(A plot opens with three panels --- one per voter class --- each showing predicted turnout for the five treatment arms with confidence intervals. Always Vote sits highest (around 60-70%); Sometimes Vote in the middle (about 30-40%); Rarely Vote lowest (about 10-15%). Within each panel, the Neighbors postcard sits clearly above the others.)",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

```{r temperance-6-test, include = FALSE}
plot_predictions(extract_fit_engine(fit_shaming),
                 condition = c("treatment", "voter_class"))
```

###

Two-variable conditioning shows the effect surface of the model: predicted probability across both axes at once. The size of the gap between Neighbors and No Postcard varies visibly across the three voter classes, which is exactly what the interaction term in the model captures.

### Exercise 7

Run the same call with `draw = FALSE` to see the underlying tibble.

CP/CR.

```{r temperance-7}
question_text(NULL,
  message = "    > plot_predictions(extract_fit_engine(fit_shaming),\n    +                  condition = c(\"treatment\", \"voter_class\"),\n    +                  draw = FALSE)\n\n  rowid  estimate ...   treatment    voter_class  conf.low  conf.high\n1     1     0.123 ...  No Postcard    Rarely Vote     0.118      0.128\n2     2     0.132 ...   Civic Duty    Rarely Vote     0.122      0.143\n...                                                              \n15   15     0.687 ...    Neighbors    Always Vote     0.682      0.692",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

```{r temperance-7-test, include = FALSE}
plot_predictions(extract_fit_engine(fit_shaming),
                 condition = c("treatment", "voter_class"),
                 draw = FALSE)
```

###

`plot_predictions()` returns a ggplot object, but it's often easier to work from the underlying tibble (`draw = FALSE`) and build the final plot from scratch.

### Exercise 8

Work with AI to create a beautiful plot starting from the output of `plot_predictions(..., draw = FALSE)`. Do this in your QMD. Title: key variables. Subtitle: takeaway. Caption: data source. Y-axis: percentages. Paste the plot code below.

```{r temperance-8}
question_text(NULL,
  message = "    plot_predictions(extract_fit_engine(fit_shaming),\n                     condition = c(\"treatment\", \"voter_class\"),\n                     draw = FALSE) |>\n      ggplot(aes(x = treatment, y = estimate, color = voter_class)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high),\n                      size = 0.7) +\n      scale_x_discrete(limits = c(\"No Postcard\", \"Civic Duty\", \"Hawthorne\",\n                                  \"Self\", \"Neighbors\")) +\n      scale_y_continuous(labels = scales::label_percent(),\n                         limits = c(0, 0.75)) +\n      facet_wrap(~ voter_class) +\n      labs(\n        title    = \"Predicted Turnout by Postcard and Voter Class\",\n        subtitle = \"Neighbors moves Sometimes/Always voters most; Rarely Vote barely budges\",\n        x        = NULL,\n        y        = NULL,\n        color    = \"Voter Class\",\n        caption  = \"Source: Gerber, Green, Larimer 2008 via primer.data::shaming\"\n      ) +\n      theme_minimal() +\n      theme(legend.position = \"none\",\n            axis.text.x     = element_text(angle = 30, hjust = 1))",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 22)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(extract_fit_engine(fit_shaming),
                 condition = c("treatment", "voter_class"),
                 draw = FALSE) |>
  ggplot(aes(x = treatment, y = estimate, color = voter_class)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                  size = 0.7) +
  scale_x_discrete(limits = c("No Postcard", "Civic Duty", "Hawthorne",
                              "Self", "Neighbors")) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 0.75)) +
  facet_wrap(~ voter_class) +
  labs(
    title    = "Predicted Turnout by Postcard and Voter Class",
    subtitle = "Neighbors moves Sometimes/Always voters most; Rarely Vote barely budges",
    x        = NULL,
    y        = NULL,
    color    = "Voter Class",
    caption  = "Source: Gerber, Green, Larimer 2008 via primer.data::shaming"
  ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x     = element_text(angle = 30, hjust = 1))
```

### Exercise 9

Finalize the new graphics chunk in your QMD. `Cmd/Ctrl + Shift + K`. At the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r temperance-9}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    plot_predictions(extract_fit_engine(fit_shaming),\n                     condition = c(\"treatment\", \"voter_class\"),\n                     draw = FALSE) |>\n      ggplot(aes(x = treatment, y = estimate, color = voter_class)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high), size = 0.7) +\n      scale_x_discrete(limits = c(\"No Postcard\", \"Civic Duty\", \"Hawthorne\",\n                                  \"Self\", \"Neighbors\")) +\n      scale_y_continuous(labels = scales::label_percent(), limits = c(0, 0.75)) +\n      facet_wrap(~ voter_class) +\n      labs(title = \"Predicted Turnout by Postcard and Voter Class\",\n           subtitle = \"Neighbors moves Sometimes/Always voters most; Rarely Vote barely budges\",\n           x = NULL, y = NULL, color = \"Voter Class\",\n           caption = \"Source: Gerber, Green, Larimer 2008 via primer.data::shaming\") +\n      theme_minimal() +\n      theme(legend.position = \"none\", axis.text.x = element_text(angle = 30, hjust = 1))\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 18)
```

###

The map is not the territory. A clean graphic tells a story, but the story always rests on assumptions that are at best approximately true.

### Exercise 10

Write the last sentence of your summary paragraph. It describes at least one Quantity of Interest and a measure of uncertainty.

```{r temperance-10}
question_text(NULL,
  message = "We estimate that the Neighbors postcard, relative to no postcard, raises the probability of voting by about 8.1 percentage points on average (95% CI: 7.6 to 8.6 pp), with the effect concentrated among Sometimes Vote (8.6 pp) and Always Vote (9.0 pp) voters and notably smaller among Rarely Vote voters (4.4 pp).",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a final sentence to your summary paragraph, but don't copy/paste exactly. `Cmd/Ctrl + Shift + K`.

### Exercise 11

Write a few sentences explaining why the estimate might be wrong. Suggest alternative estimates if warranted.

```{r temperance-11}
question_text(NULL,
  message = "Three reasons our estimate may be optimistic. First, our 95% confidence intervals are computed at the voter level, but treatment was assigned at the household level --- so the true intervals are mildly wider than what we report. Second, the experiment was conducted in 2006 in Michigan, on voters who had cast a ballot in 2004; the relevance to a 2026 Texas campaign depends on stability holding across both era and geography, and the slope of turnout on Neighbors-style postcards has likely flattened since 2006 as voters have habituated to the technique. Third, we report intent-to-treat (assignment, not delivery) effects --- for the campaign manager weighing the cost-effectiveness of the postcards, this is the right number; for understanding why Neighbors works, it understates the per-recipient effect. A defensible 2026-Texas estimate is a Neighbors effect somewhere between 3 and 8 percentage points, not the 8.1 we measure.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

Always go back to your Preceptor Table. Even a perfectly estimated statistical model is rarely as useful as we'd like.

### Exercise 12

Rearrange the material in your QMD so the order is graphic, then summary paragraph. The chunk that creates `fit_shaming` must occur before the chunk that creates the graphic. `Cmd/Ctrl + Shift + K`. At the R prompt, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r temperance-12}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\")\n    ---\n    title: \"Shaming\"\n    author: \"<your name>\"\n    execute:\n      echo: false\n    ---\n    \n    ```{r}\n    #| message: false\n    library(tidyverse)\n    library(tidymodels)\n    library(primer.data)\n    library(broom)\n    library(marginaleffects)\n    ```\n    \n    ```{r}\n    x <- shaming |>\n      mutate(civ_engage = primary_00 + primary_02 + primary_04 +\n                          general_00 + general_02 + general_04) |>\n      select(primary_06, treatment, sex, age, civ_engage) |>\n      drop_na() |>\n      mutate(voter_class = factor(case_when(\n        civ_engage %in% c(5, 6) ~ \"Always Vote\",\n        civ_engage %in% c(3, 4) ~ \"Sometimes Vote\",\n        civ_engage %in% c(1, 2) ~ \"Rarely Vote\"),\n        levels = c(\"Rarely Vote\", \"Sometimes Vote\", \"Always Vote\"))) |>\n      mutate(voted = as.factor(primary_06)) |>\n      drop_na()\n    ```\n    \n    ```{r}\n    #| cache: true\n    fit_shaming <- logistic_reg(engine = \"glm\") |>\n      fit(voted ~ age + sex + treatment*voter_class, data = x)\n    ```\n    \n    ```{r}\n    plot_predictions(extract_fit_engine(fit_shaming),\n                     condition = c(\"treatment\", \"voter_class\"),\n                     draw = FALSE) |>\n      ggplot(aes(x = treatment, y = estimate, color = voter_class)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high), size = 0.7) +\n      scale_x_discrete(limits = c(\"No Postcard\", \"Civic Duty\", \"Hawthorne\",\n                                  \"Self\", \"Neighbors\")) +\n      scale_y_continuous(labels = scales::label_percent(), limits = c(0, 0.75)) +\n      facet_wrap(~ voter_class) +\n      labs(title = \"Predicted Turnout by Postcard and Voter Class\",\n           subtitle = \"Neighbors moves Sometimes/Always voters most\",\n           x = NULL, y = NULL, color = \"Voter Class\",\n           caption = \"Source: Gerber, Green, Larimer 2008 via primer.data::shaming\") +\n      theme_minimal() +\n      theme(legend.position = \"none\", axis.text.x = element_text(angle = 30, hjust = 1))\n    ```\n    \n    Voter mobilization is a perennial campaign question. Using the 344,084-voter Gerber-Green-Larimer 2006 randomized field experiment in Michigan, we estimate the average causal effect of the Neighbors postcard on the probability that a registered voter casts a ballot in the August 2006 primary. One concern is that random assignment happened at the household level, so our 95% intervals are mildly optimistic. We model voter turnout as a Bernoulli variable which is a logistic function of treatment, voter class, the treatment-by-voter-class interaction, age, and sex. We estimate that the Neighbors postcard raises the probability of voting by about 8.1 percentage points on average, with the effect concentrated among Sometimes/Always voters (8-9 pp) and notably smaller among Rarely Vote voters (4 pp).",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 14)
```

###

This is the version of your QMD your teacher is most likely to look at closely.

### Exercise 13

Publish your rendered QMD to GitHub Pages. In the Terminal (not the R prompt!), run:

```
quarto publish gh-pages analysis.qmd
```

Copy/paste the resulting URL below.

```{r temperance-13}
question_text(NULL,
  message = "https://<your-github-username>.github.io/shaming/analysis.html",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

Commit/push everything.

### Exercise 14

Copy/paste the URL to your GitHub repo.

```{r temperance-14}
question_text(NULL,
  message = "https://github.com/<your-github-username>/shaming",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God.

The world confronts us. Make decisions we must.


## Summary
###

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(extract_fit_engine(fit_shaming),
                 condition = c("treatment", "voter_class"),
                 draw = FALSE) |>
  ggplot(aes(x = treatment, y = estimate, color = voter_class)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                  size = 0.7) +
  scale_x_discrete(limits = c("No Postcard", "Civic Duty", "Hawthorne",
                              "Self", "Neighbors")) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 0.75)) +
  facet_wrap(~ voter_class) +
  labs(
    title    = "Predicted Turnout by Postcard and Voter Class",
    subtitle = "Neighbors moves Sometimes/Always voters most; Rarely Vote barely budges",
    x        = NULL,
    y        = NULL,
    color    = "Voter Class",
    caption  = "Source: Gerber, Green, Larimer 2008 via primer.data::shaming"
  ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x     = element_text(angle = 30, hjust = 1))
```

Voter mobilization is a perennial campaign question: which messaging --- if any --- gets registered voters who would otherwise stay home to actually show up. Using the 344,084-voter Gerber-Green-Larimer 2006 randomized field experiment in Michigan, we estimate the average causal effect of the Neighbors postcard on the probability that a registered voter casts a ballot in the August 2006 primary. One concern is that random assignment happened at the household level, not the voter level, so our 95% confidence intervals are mildly optimistic. We model voter turnout as a Bernoulli variable which is a logistic function of treatment, voter class, the treatment-by-voter-class interaction, age, and sex. We estimate that the Neighbors postcard raises the probability of voting by about 8.1 percentage points on average (95% CI: 7.6 to 8.6 pp), with the effect concentrated among Sometimes Vote (8.6 pp) and Always Vote (9.0 pp) voters and notably smaller among Rarely Vote voters (4.4 pp).

This finding would inform a campaign-allocation decision: which voters to target with the Neighbors-style postcard, given a fixed mailing budget. The headline implication is counterintuitive --- the people who needed a turnout nudge most (Rarely Vote) are also the ones a postcard moves least, while the people most likely to vote anyway (Always Vote) are the ones whose marginal turnout is most movable by social pressure. A campaign manager who applied this finding directly would mail Neighbors-style postcards to Sometimes/Always voters and skip the Rarely Vote list. The finding does not address every campaign question --- it says nothing about door-knocking, phone banking, digital ads, or how the effect of social-pressure mail interacts with other GOTV tools running simultaneously. But it gives a defensible starting point.

The world is always more uncertain than our models would have us believe.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
