---
title: Stops
author: David Kane and Aashna Patel
tutorial:
  id: 141-stops
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #13 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(tidymodels)
library(broom)
library(marginaleffects)
library(primer.data)
library(tidytext)   # Needed (?) for scale_x_reordered()

library(easystats)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 600, 
        tutorial.storage = "local") 

fit_arrested <- linear_reg() |>
    set_engine("lm") |>
    fit(arrested ~ zone, data = stops)

x <- stops |>
  filter(race %in% c("black", "white")) |>
  mutate(race = str_to_title(race), 
         sex = str_to_title(sex))

fit_stops <- linear_reg() |>
    set_engine("lm") |>
    fit(arrested ~ sex + race*zone, data = x)

fit_stops_logistic <- logistic_reg() |>
  set_engine("glm") |>
  fit(as.factor(arrested) ~ sex + race, data = x)
# This file is too large for GitHub to be written so it must be calculated on the fly
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial supports [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). 

The world confronts us. Make decisions we must.

Imagine you’re a member of your city’s police department, and you believe strongly in fair and equal treatment for everyone. You want to ensure that a person’s race or age doesn’t unfairly affect their chances of being arrested during a traffic stop. There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
	message = "Wisdom, Justice, Courage, and Temperance.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 2)
```

###

Why do we ask this, and a score more other questions, in each tutorial? Because the best way to (try to) ensure that students remember these concepts more than a few months after the course ends is [spaced repetition](https://en.wikipedia.org/wiki/Spaced_repetition), although we focus more on the repetition than on the spacing.

### Exercise 2

Create a Github repo called `stops`. Make sure to click the "Add a README file" check box.

Connect the repo to a project on your computer using `File -> New Folder from Git ...`. Make sure to select the "Open in a new window" box. 

You need two Positon windows: this one for running the tutorial and the one you just created for writing your code and interacting with the Console.

Select `File -> New File -> Quarto Document ...`. Provide a title -- `"Stops"` -- and an author (you). Render the document and save it as `stops.qmd`.

Create a `.gitignore` file with `stops_files` on the first line and then a blank line. Save and push.

In the Console, run:

```         
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the Console.

CP/CR.

```{r introduction-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Professionals keep their data science work in the cloud because laptops fail.

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.data)` in a new code chunk. Render the file.

Notice that the file does not look good because the code is visible and there are annoying messages. To take care of this, add `#| message: false` to remove all the messages in this `setup` chunk. Also add the following to the YAML header to remove all code echos from the HTML:

```         
execute: 
  echo: false
```

In the Console, run:

```         
show_file("stops.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 6)
```

### 

Render again. Everything looks nice, albeit empty, because we have added code to make the file look better and more professional.

### Exercise 4

Place your cursor in the QMD file on the `library(tidyverse)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(tidyverse)` to be copied down to the Console and then executed. 

CP/CR.

```{r introduction-4}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

The data that we will use was sourced from the [Open Policing project](https://openpolicing.stanford.edu/findings/). Based at Stanford University, the project aims to improve police accountability and transparency by providing data on traffic stops across the United States. They have many downloadable datasets, and our data is specifically derived from their [New Orleans dataset.](https://openpolicing.stanford.edu/data/)

###

A version of the data from the Open Policing project is available in the `stops` tibble.

### Exercise 5

In the Console, type `?stops`, and paste the Description below.

```{r introduction-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 8)
```

### 

`stops` contains data from over 400,000 traffic stops in New Orleans from July 1, 2011 to July 18, 2018. The dataset includes information about the date, time, and location of each stop, as well as demographic details about the driver and the outcomes of the stop.

### Exercise 6

Define a causal effect.

```{r introduction-6}
question_text(NULL,
	message = "A causal effect is the difference between two potential outcomes.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

According to the Rubin Causal Model, there must be two (or more) potential outcomes for any discussion of causation to make sense. This is simplest to discuss when the treatment only has two different values, thereby generating only two potential outcomes. 

### Exercise 7

What is the fundamental problem of causal inference?

```{r introduction-7}
question_text(NULL,
	message = "The fundamental problem of causal inference is that we can only observe one potential outcome.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

If the treatment variable is continuous (like a lottery payment), then there are lots and lots of potential outcomes, one for each possible value of the treatment variable. 

### Exercise 8

Arrests in traffic stops are the broad topic of this tutorial. Given that topic, which variable in `stops` should we use as our outcome variable? 

```{r introduction-8}
question_text(NULL,
	message = "We should be using the `arrested` variable.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 2)
```

### 

We will use `arrested` as our outcome variable.

###

```{r}
stops |>
  mutate(arrested = if_else(arrested == 1, "Arrested", "Not Arrested")) |>
  ggplot(aes(x = arrested)) +
  geom_bar(fill = "#2c7fb8") +
  labs(
    title = "Distribution of Arrest Outcomes",
    subtitle = "Most traffic stops do not result in arrest",
    x = NULL,
    y = "Number of Stops"
  ) +
  theme_minimal(base_size = 13)
```

Most traffic stops in the dataset do not result in an arrest. This imbalance suggests that arrest is a relatively rare outcome, which supports using a logistic regression model to predict the probability of arrest.

### Exercise 9

Let's imagine a brand new variable which **does not exist** in the data. This variable should be binary, meaning that it only takes on one of two values. It should also, at least in theory, be manipulable. In other words, if the value of the variable is "3," or whatever, then it generates one potential outcome and if it is "9," or whatever, it generates another potential outcome.

Describe this imaginary variable and how might we manipulate its value.

```{r introduction-9}
question_text(NULL,
	message = "Imagine a variable called `mask`, indicating whether or not the person is wearing a mask. We can manipulate this, at least in theory, by giving out masks to half the motorists that we plan on studying, and seeing if the mask affects their chances of getting arrested.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Any data set can be used to construct a causal model as long as there is at least one covariate that we can, at least in theory, manipulate. It does not matter whether or not anyone did, in fact, manipulate it. 

### Exercise 10

Given our (imaginary) treatment variable `mask`, how many potential outcomes are there for each arrest? Explain why.

```{r introduction-10}
question_text(NULL,
	message = "There are 2 potential outcomes because the treatment variable `mask` takes on 2 posible values: either wearing a mask and getting arrested or not wearing one and getting arrested.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The same data set can be used to create, separately, lots and lots of different models, both causal and predictive. We can just use different outcome variables and/or specify different treatment variables. This is a *conceptual* framework we apply to the data. It is never inherent in the data itself.

### Exercise 11

In a few sentences, specify the two different values for the imaginary treatment variable `mask`, for a single unit, guess at the potential outcomes which would result, and then determine the causal effect for that unit given those guesses.

```{r introduction-11}
question_text(NULL,
	message = "For each person, there are only two possible values for arrested: 0 (meaning not arrested) and 1 (meaning arrested). So, there might be a person who, if she had been wearing a mask, would have been arrested, but, since she was not wearing a mask, she was not arrested. For a given arrest, assume that the value of the treatment variable might be `is wearing a mask` or `isn't wearing a mask`. If the person `is wearing a mask`, and gets arrested, but the person `isn't wearing a mask`, and does not get arrested. The causal effect on the outcome of a treatment of `mask = TRUE` versus `mask = FALSE` is getting arrested the difference between two potential outcomes --- which is the causal effect.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

A causal effect is defined as the *difference* between two potential outcomes. Keep two things in mind.

First, *difference* does not necessarily mean *subtraction*. Many potential outcome are not numbers. For example, it makes no sense to subtract a potential outcome, like who you would vote for if you saw a Facebook ad, from another potential outcome, like who you vote for if you did not see the ad.

Second, even in the case of numeric outcomes, you can’t simply say the effect is 10 without specifying the order of subtraction, although there is, perhaps, a default sense in which the causal effect is defined as potential outcome under treatment minus potential outcome under control.

### Exercise 12

Let's consider a *predictive* model. Which variable in `stops` do you think might have an important connection to `arrested`? 

```{r introduction-12}
question_text(NULL,
	message = "Let's consider the variable `race`. Does race have an effect on the rate people get arrested at?",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 2)
```

### 

With a predictive model, each individual unit has only one observed outcome. There are not two potential outcomes because none of the covariates are treated as treatment variables. Instead, all covariates are assumed to be "fixed."

Predictive models have no "treatments" -—- only covariates.

### Exercise 13

Specify two different groups of people which have different values for race and which might have different average values for the arrest.  

```{r introduction-13}
question_text(NULL,
	message = "Consider two groups, the first with a value for race of white and the second with value black. Those two groups might have different average values for the outcome variable.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

In predictive models, do not use words like "cause," "influence," "impact," or anything else which suggests causation. The best phrasing is in terms of "differences" between groups of units with different values for a covariate of interest.

Any causal connection means exploring the *within row* difference between two potential outcomes. There's no need to consider other rows.

### Exercise 14

Write a predictive question which connects the outcome variable `arrested` to `race`, the covariate of interest. 

```{r introduction-14}
question_text(NULL,
	message = "What is the difference in arrest rate between Black and White drivers adjusting for other variables?",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

This is the first version of the question. We will now create a Preceptor Table to answer the question. We may then revise the question given complexities discovered in the data. We then update the question and the Preceptor Table. And so on.

## Wisdom
### 

*The power to question is the basis of all human progress.* - Indira Gandhi

Our goal is to ensure fair and equal treatment during traffic stops. We know arrests can depend on many factors, but we’re especially concerned about whether race plays a role. Are Black drivers more likely to be arrested than White drivers, even after accounting for age, gender, and reason for the stop? Does this pattern change depending on the location or time of day? Understanding these differences won’t fix everything, but it can help us make more informed, equitable decisions.

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
    message = "Wisdom requires a question, the creation of a Preceptor Table, and an examination of our data.",
    answer(NULL, correct = TRUE),
    allow_retry = FALSE,
    incorrect = NULL,
    rows = 3)
```

###

*The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* -- John W. Tukey

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
	message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The Preceptor Table does not include all the covariates which you will eventually include in your model. It only includes covariates which you need to answer your question.

> Are Black drivers arrested at higher rates than White drivers, even after adjusting for age, gender, and stop reason?

> Does the difference in arrest rates between Black and White drivers vary by location or time of day?

### Exercise 3

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-3}
question_text(NULL,
	message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

 This problem is predictive so we are comparing the likelihood of arrest for Black and White drivers, adjusting for other factors like age, gender, and stop reason.

### Exercise 4

What are the units for this problem?

```{r wisdom-4}
question_text(NULL,
	message = "The units are the people who have been pulled over.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Specifying the Preceptor Table forces us to think clearly about the units and outcomes implied by the question. The resulting discussion sometimes leads us to modify the question with which we started. No data science project follows a single direction. We always backtrack. There is always dialogue.

We model units, but we only really care about aggregates.

### Exercise 5

What is the outcome variable for this problem?

```{r wisdom-5}
question_text(NULL,
	message = "Our outcome variable is `arrested`.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The outcome variable that we really care about is often not the outcome variable which our data includes. This compromise --- working with what we *have* rather than what we really *want* --- is a part of most data science work in the real world.

### Exercise 6

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r wisdom-6}
question_text(NULL,
	message = "A useful covariate is `race`. Other characteristics like the person's age, sex, and the type of car they're driving might also affect whether someone is arrested. It would also be beneficial to include `zone`, since the officer's zone may affect arrest rates.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The term "covariates" is used in at least three ways in data science. First, it is all the variables which might be useful, regardless of whether or not we have the data. Second, it is all the variables for which we have data. Third, it is the set of variables in the data which we end up using in the model.

### Exercise 7

What are the treatments, if any, for this problem?

```{r wisdom-7}
question_text(NULL,
	message = "There are no treatments.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Remember that a treatment is just another covariate which, for the purposes of this specific problem, we are assuming can be manipulated, thereby, creating two or more different potential outcomes for each unit.

### Exercise 8

What moment in time does the Preceptor Table refer to?

```{r wisdom-8}
question_text(NULL,
	message = "Our Preceptor Table should represent traffic stops sometime between the late 2000s and the mid 2010s for maximum compatibility with our data.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

A Preceptor Table can never really refer to an exact instant in time since nothing is instantaneous in this fallen world.

In almost all practical problems, the data was gathered at a time other than that to which the Preceptor Table refers.

```{r}
plot_predictions(fit_arrested,
                      newdata = expand_grid(zone = unique(x$zone)),
                      condition = c("zone")) +
  coord_flip() +
  scale_x_reordered() +
  labs(x = "Zone", 
       y = "Estimated Arrest Probability", 
       title = "Predicted Arrest Rate of New Orleans Motorists by Zones",
       subtitle = "The bars represent the 95% Confidence Intervals.",
       caption = "Data from the Open Policing Project")
```

> *You can never look at the data too much. -- Mark Engerman*

### Exercise 9

Describe in words the Preceptor Table for this problem.

```{r wisdom-9}
question_text(NULL,
	message = "The Preceptor Table has one row for each traffic stop. The columns are an ID, `arrested` (the outcome variable), and covariates including `race`, `sex`, and `zone`.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The Preceptor Table for this problem looks something like this:

```{r}
tibble(ID = c("1", "2", "...", "10", "11", "...", "105,852,176"),
       arrested = c("TRUE", "FALSE", "FALSE", "FALSE", "TRUE", "FALSE", "FALSE"),
       sex = c("M", "F", "...", "F", "F", "...", "M"),
       race = c("Black", "White", "...", "Black", "White", "Black", "Black"),
       zone = c("...", "B", "M", "...", "W", "...", "A"))|>

  gt() |>
  tab_header(title = "Preceptor Table") |>
  cols_label(ID = md("ID"),
             arrested = md("Arrested"),
             sex = md("Sex"),
             race = md("Race"),
             zone = md("Zone")) |>
  tab_style(cell_borders(sides = "right"),
            location = cells_body(columns = c(ID))) |>
  tab_style(style = cell_text(align = "left", v_align = "middle", size = "large"),
            locations = cells_column_labels(columns = c(ID))) |>
  cols_align(align = "center", columns = everything()) |>
  cols_align(align = "left", columns = c(ID)) |>
  fmt_markdown(columns = everything()) |>
  tab_spanner(label = "Outcome", columns = c(arrested)) |>
  tab_spanner(label = "Covariate", columns = c(sex, race, zone))
```

Like all aspects of a data science problem, the Preceptor Table evolves as we continue our work. 

### Exercise 10

What is the narrow, specific question we will try to answer?

```{r wisdom-10}
question_text(NULL,
	message = "What is the difference in arrest rate between Black and White drivers adjusting for other variables?",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

###

The answer to this question is your "Quantity of Interest." It is OK if your question differs from ours. Many similar questions lead to the creation of the same model. For the purpose of this tutorial, let's use our question.

Our Quantity of Interest might appear too specific, too narrow to capture the full complexity of the topic. There are many, many numbers in which we are interested, many numbers that we want to know. But we don't need to list them all here! We just need to choose one of them since our goal is just to have a specific question which helped to guide us in the creation of the Preceptor Table and, soon, the model. 

### Exercise 11

Over the course of this tutorial, we will be creating a summary paragraph. The purpose of this exercise is to write the first two sentences of that paragraph.

The first sentence is a general statement about the overall topic, mentioning both the general class of the outcome variable and of at least one of the covariates. It is **not** connected to the initial "Imagine that you are XX" which set the stage for this project. That sentence can be rhetorical. It can be trite, or even a platitude. The purpose of the sentence is to let the reader know, gently, about our topic.

The second sentence does two things. First, it introduces the data source. Second, it introduces the specific question. The sentence can't be that long. Important aspects of the data include when/where it was gathered, how many observations it includes and the organization (if famous) which collected it.

Type your two sentences below.

```{r wisdom-11}
question_text(NULL,
	message = "Racial disparities in policing outcomes remain a pressing concern, particularly when examining how factors like race and location influence the likelihood of arrest during traffic stops. Using data from a study of New Orleans drivers, we seek to understand the relationship between driver race and the probability of getting arrested during a traffic stop.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Read our answer. It will not be the same as yours. You can, if you want, change your answer to incorporate some of our ideas. Do not copy/paste our answer exactly. Add your two sentences, edited or otherwise, to your QMD, `Cmd/Ctrl + Shift + K`, and then commit/push.

## Justice
### 

*Justice delayed is justice denied.* - William E. Gladstone

### Exercise 1

In your own words, name the five key components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
	message = "Justice concerns the Population Table and the four key assumptions which underlie it: validity, stability, representativeness, and unconfoundedness.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Justice is about concerns that you (or your critics) might have, reasons why the model you create might not work as well as you hope. 

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
	message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Validity is always about the *columns* in the Preceptor Table and the data. Just because columns from these two different tables have the same *name* does not mean that they are the same *thing*.

### Exercise 3

Provide one reason why the assumption of validity might not hold for the outcome variable `arrested` or for one of the covariates. Use the words "column" or "columns" in your answer.

```{r justice-3}
question_text(NULL,
	message = "The `arrested` column in our data records whether a person was formally arrested. But the `arrested` column in the Preceptor Table might capture a broader category of police actions — including detainments or summonses that fall short of a formal arrest. If the two columns are using different definitions of what counts as an arrest, they are not measuring the same thing.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

In order to consider the Preceptor Table and the data to be drawn from the same population, the columns from one must have a *valid correspondence* with the columns in the other. Validity, if true (or at least reasonable), allows us to construct the Population Table, which is the first step in Justice.

Because we control the Preceptor Table and, to a lesser extent, the original question, we can adjust those variables to be “closer” to the data that we actually have. This is another example of the iterative nature of data science. If the data is not close enough to the question, then we check with our boss/colleague/customer to see if we can modify the question in order to make the match between the data and the Preceptor Table close enough for validity to hold.

Despite these potential problems, we will assume that validity holds since it, mostly (?), does.

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
	message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The Population Table is almost always much bigger than the combination of the Preceptor Table and the data because, if we can really assume that both the Preceptor Table and the data are part of the same population, than that population must cover a broad universe of time and units since the Preceptor Table and the data are, themselves, often far apart from each other.

### Exercise 5

Specify the unit/time combinations which define each row in this Population Table.

```{r justice-5}
question_text(NULL,
	message = "Each row in the Population Table represents a single traffic stop involving a driver, observed at the time the stop occurred. So, the unit is a driver during a specific traffic stop, and the time is the moment or date of that stop.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The exact time period used --- whether hour, day, month, year, or whatever --- is relatively arbitrary. The important thing to note is that the Population Table, unlike the Preceptor Table, covers a period of time over which things may change.

```{r}
stops |>
  select(race, age, sex, zone, arrested) |>
  mutate(
    race = str_to_title(race),
    sex = str_to_title(sex),
    arrested = if_else(arrested == 1, "Yes", "No")
  ) |>
  slice(1:5) |>
  gt()
```

### Exercise 6

In your own words, define the assumption of "stability" when employed in the context of data science.

```{r justice-6}
question_text(NULL,
	message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Stability is all about *time*. Is the relationship among the columns in the Population Table stable over time? In particular, is the relationship --- which is another way of saying "mathematical formula" --- at the time the data was gathered the same as the relationship at the (generally later) time referenced by the Preceptor Table.

### Exercise 7

Provide one reason why the assumption of stability might not be true in this case.

```{r justice-7}
question_text(NULL,
	message = "We don't know the exact time period of the Preceptor Table, so we don't know if any laws were passed between the time periods of both data sources that affected the basis of arrests during traffic stops. Therefore, we don't know if the arrests reported in the `arrested` column in both data sources were made on the same basis and whether or not both data sources had the same arrest rate.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

A change in time or the distribution of the data does not, in and of itself, demonstrate a violation of stability. Stability is about the parameters: $\beta_0$, $\beta_1$ and so on. Stability means these parameters are the same in the data as they are in the population as they are in the Preceptor Table.

### Exercise 8

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: `Data -> Population -> Preceptor Table`. In your own words, define the assumption of "representativeness" when employed in the context of data science.

```{r justice-8}
question_text(NULL,
	message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Ideally, we would like both the Preceptor Table *and* our data to be random samples from the population. Sadly, this is almost never the case.

### Exercise 9

We do not use the data, directly, to estimate missing values in the Preceptor Table. Instead, we use the data to learn about the overall population. Provide one reason, involving the relationship between the data and the population, for why the assumption of representativeness might not be true in this case. 

```{r justice-9}
question_text(NULL,
	message = "The dataset `stops` is a heavily modified version of the data from the actual study, and therefore has left out nearly 3.1 million entries from the real data, shortening it to roughly 400,000 entries. The deletion of the entries may have led to a misrepresentation of the population, in that a lot of the current data may only be from select areas with select conditions present, and could be from biased officers who are more likely to arrest drivers compared to other officers in the zone. This would mess up our final predictions because it would be providing values that are unrealistic to the area, causing our entire model to be unreastic.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The reason that representativeness is important is because, when it is violated, the estimates for the model parameters might be biased. 

### Exercise 10

We use information about the population to make inferences about the Preceptor Table. Provide one reason, involving the relationship between the Population and the Preceptor Table, for why the assumption of representativeness might not be true in this case.

<!-- AR: This answer refers to the data and not so much the Preceptor Table in my opinion. Edit? -->

```{r justice-10}
question_text(NULL,
	message = "Our data may not be collected fairly, in that some of the data could have just been collected from corrupt officers who would be more likely to arrest drivers than other officers in the zone. The opposite could be inferred as well. This, again, would mess up our final predictions because it would be providing values that are unrealistic to the area, causing our entire model to be unreastic.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Stability looks across time periods. Representativeness looks within time periods, for the most part. 

### Exercise 11

In your own words, define the assumption of "unconfoundedness" when employed in the context of data science.

```{r justice-11}
question_text(NULL,
	message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

This assumption is only relevant for causal models. We describe a model as "confounded" if this is not true. The easiest way to ensure unconfoundedness is to assign treatment randomly.

### Exercise 12

A statistical model consists of two parts: the *probability family* and the *link function*. The probability family is the probability distribution which generates the randomness in our data. The link function is the mathematical formula which links our data to the unknown parameters in the probability distribution. 

Add `library(tidymodels)` to the QMD file.

Place your cursor in the QMD file on the `library(tidymodels)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(tidymodels)` to be copied down to the Console and then executed. 

CP/CR.

```{r justice-12}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

The probability family is determined by the outcome variable $Y$. 

Since $Y$ is a binary variable (with exactly two possible values), the probability family is Bernoulli.

$$Y \sim \text{Bernoulli}(\rho)$$

where $\rho$ is the probability that one of the two possible values --- conventionally referred to as `1` or `TRUE` --- occurs. By definition, $1 - \rho$ is the probability of the other value.

### Exercise 13

Add `library(broom)` to the QMD file.

Place your cursor in the QMD file on the `library(broom)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(broom)` to be copied down to the Console and then executed. 

CP/CR.

```{r justice-13}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

The link function, the basic mathematical structure of the model, is (mostly) determined by the type of outcome variable. 

<!-- AR: In the courage section we say we use a linear model for a binary outcome variable. Need to make consistent. -->

For a binary outcome variable, we use a log-odds model:

$$
\log\left[ \frac { \rho }{ 1 - \rho } \right] = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \ldots 
$$

### Exercise 14

Use AI to come up with a $\LaTeX$ representation of the mathematical structure of the model, with $Y$ as the dependent variable and $X_1$, $X_2$ and so on as the independent variables. This version will not have the values of any parameters since we have not, yet, estimated them. Confirm that the $\LaTeX$ code works by placing it in your QMD and then rendering. Paste that $\LaTeX$ code below.

```{r justice-14}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

Our answer:

$$
P(Y = 1) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n)}}
$$

with $( Y \sim \text{Bernoulli}(\rho) )$ where $( \rho = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n)}} )$



Which we created with $\LaTeX$ code that looks like this:

````
$$
P(Y = 1) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n)}}
$$

with \( Y \sim \text{Bernoulli}(\rho) \) where \( \rho = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n)}} \)
````

This follows the logistic regression form for binary data, where the $\beta$ coefficients represent the effect of predictors on the log-odds of the outcome.

We use generic variables --- $Y$, $X_1$ and so on --- because our purpose is to describe the general mathematical structure of the model, independent of the specific variables we will eventually choose to use.

Having decided on the basic mathematical structure of the model, a choice mostly driven by the distribution of our outcome variable, we now turn toward estimating the model.

### Exercise 15

Write one sentence which highlights a potential weakness in your model. This will almost always be derived from possible problems with the assumptions discussed above. We will add this sentence to our summary paragraph. So far, our version of the summary paragraph looks like this:

> Racial disparities in policing outcomes remain a pressing concern, particularly when examining how factors like race and location influence the likelihood of arrest during traffic stops. Using data from a study of New Orleans drivers, we seek to understand the relationship between driver race and the probability of getting arrested during a traffic stop. 

Of course, your version will be somewhat different.

```{r justice-15}
question_text(NULL,
	message = "Racial disparities in policing outcomes remain a pressing concern, particularly when examining how factors like race and location influence the likelihood of arrest during traffic stops. Using data from a study of New Orleans drivers, we seek to understand the relationship between driver race and the probability of getting arrested during a traffic stop. However, our data from both our Preceptor Table and our dataset may not fully represent the population as both may not be from the same time frame and some of our data may come from biased officers, who may target certain groups of individuals.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Add a weakness sentence to the summary paragraph in your QMD. You can modify your paragraph as you see fit, but do not copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`, and then commit/push.

## Courage
### 

*Courage is going from failure to failure without losing enthusiasm.* - Winston Churchill

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
	message = "Courage creates the data generating mechanism.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Having decided on the basic mathematical structure of the model at the end of *Justice*, a choice mostly driven by the distribution of our outcome variable, we now turn toward estimating the model.

### Exercise 2

<!--AR: in justice-13 knowledge drop we say that we use a log-odds model for a binary outcome variable! -->

Because our outcome variable is binary, start to create the model by using `linear_reg(engine = "lm")`.

```{r courage-2, exercise = TRUE}

```

```{r courage-2-hint-1, eval = FALSE}
linear_reg(engine = "...")
```

```{r courage-2-test, include = FALSE}
linear_reg(engine = "lm")
```

**Note:** This will give you an error that we will be fixing later.

### 

The *[tidymodels](https://www.tidymodels.org/)* framework is the most popular one in the R world for estimating models. *[Tidy Modeling with R](https://www.tmwr.org/)* by Max Kuhn and Julia Silge is a great introduction.

### Exercise 3

Continue the pipe to `fit(arrested ~ sex, data = x)`.

```{r courage-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-3-hint-1, eval = FALSE}
... |> 
  fit(arrested ~ ..., data = ...)
```

```{r courage-3-test, include = FALSE}
linear_reg() |>
  set_engine("lm") |>
  fit(arrested ~ sex, data = x)
```

### 

Recall that a categorical variable (whether character or factor) like `sex` is turned into a $0/1$ "dummy" variable which is then re-named something like $sexMale$. After all, we can't have words --- like "Male" or "Female" --- in a mathematical formula, hence the need for dummy variables.

### Exercise 4

Continue the pipe with `tidy(conf.int = TRUE)`.

```{r courage-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-4-hint-1, eval = FALSE}
linear_reg() |>
  set_engine("lm") |>
  fit(arrested ~ sex, data = x) |>
  tidy(... = TRUE)
```

```{r courage-4-test, include = FALSE}
linear_reg() |>
  set_engine("lm") |>
  fit(arrested ~ sex, data = x) |>
  tidy(conf.int = TRUE)
```

### 

The meaning of the magnitude of the coefficient values is much harder to interpret in logistic models than in linear models. But the fact that the coefficient of sexMale is positive means that men are more likely to be arrested than women, and the fact that the confidence interval excludes zero means that we can be fairly certain of that claim.

### Exercise 5

Change the call to `fit()` to `fit(arrested ~ race, data = x)` and remove the call to `tidy()`. Note that behind the scenes, we've filtered stops to only include the race values "Black" and "White" behind the scenes.

```{r courage-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-5-hint-1, eval = FALSE}
linear_reg() |>
  set_engine("lm") |>
  fit(arrested ~ ..., data = x) 
```

```{r courage-5-test, include = FALSE}
linear_reg() |>
  set_engine("lm") |>
  fit(arrested ~ race, data = x)
```

### 

A categorical variable with $N$ levels is represented using $N−1$ dummy $(0/1)$ variables. Because models typically include an intercept, one category is omitted and becomes the reference group, absorbed into the intercept. For example, if race has values “Black” and “White,” the model includes one dummy variable (e.g., $race_white$), and the intercept represents the reference category (“Black”).

### Exercise 6

Add back the call to `tidy(conf.int = TRUE)`.

```{r courage-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-6-hint-1, eval = FALSE}
linear_reg() |>
  set_engine("lm") |>
  fit(arrested ~ race, data = x) |>
   tidy(conf.int = TRUE)

```

```{r courage-6-test, include = FALSE}
linear_reg() |>
  set_engine("lm") |>
  fit(arrested ~ race, data = x) |>
  tidy(conf.int = TRUE)
```

### 

White respondents, the group for which a dummy variable is created here (raceWhite), are less likely to be arrested than respondents of other races. The coefficient is −0.0438, and the 95% confidence interval ranges from −0.0469 to −0.0408, which does not include 0. That means this difference is statistically significant — we can be fairly confident it's not due to random chance.

The exact size of the coefficient is harder to interpret since we’re working with a logistic model, but the direction is clear: identifying as white is associated with a lower probability of getting arrested. 

### Exercise 7

Change the call to `fit()` to `fit(arrested ~ sex + race, data = x)`.

```{r courage-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-7-hint-1, eval = FALSE}
... |> 
  fit(... ~ sex + ..., ... = x) |>
  ...
```

```{r courage-7-test, include = FALSE}
linear_reg(engine = "lm") |> 
  fit(arrested ~ sex + race, data = x) |> 
  tidy(conf.int = TRUE)
```

### 

```{r}
linear_reg(engine = "lm") |> 
  fit(arrested ~ sex + race, data = x) |> 
  tidy(conf.int = TRUE) |> 
  select(term, estimate, conf.low, conf.high)
```

The more variables we add, the more difficult it is to interpret the meaning of any particular coefficient. But interpretation also becomes less important. We don't really care about coefficients. We care about using our model to estimate quantities of interest.

In this case, the coefficient for `raceWhite` is slightly more negative than before, presumably because we’ve now included `sex`, and men are somewhat more likely to be arrested. By adjusting for `sex`, we’re isolating the effect of race more precisely — and the effect of being white remains statistically significant and negative.

### Exercise 8

<!--AR: Haven't really seen interactions or had them explained in tutorials yet, so from my perspective I don't really understand what this means or that race*zone is an interaction term -->

Change the formula in the call to `fit()` to `arrested ~ sex + race*zone`.

```{r courage-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-8-hint-1, eval = FALSE}
... |> 
  fit(arrested ~ sex + race*zone, data = x) |>
  ...
```

```{r courage-8-test, include = FALSE}
linear_reg(engine = "lm") |> 
  fit(arrested ~ sex + race*zone, data = x) |> 
  tidy(conf.int = TRUE)
```

### 

We included the interaction between race and zone because the effect of race on the likelihood of being arrested may differ depending on the zone. For example, the coefficient for raceWhite:zoneD is negative and its confidence interval excludes zero, suggesting that white individuals may be treated differently in that zone compared to others. And because we’ve included interaction terms, we keep the main effects for both race and zone as well, even if not all of them are individually significant.

### Exercise 9

Behind the scenes of this tutorial, an object called `fit_stops` has been created which is the result of the code above. Type `fit_stops` and hit "Run Code." This generates the same results as using `print(fit_stops)`.

```{r courage-9, exercise = TRUE}

```

```{r courage-9-hint-1, eval = FALSE}
fit_stops
```

```{r courage-9-test, include = FALSE}
fit_stops
```

### 

In data science, **we deal with words, math, and code, but the most important of these is code.** We created the mathematical structure of the model and then wrote a model formula in order to estimate the unknown parameters. 

### Exercise 10

We need `fit_stops` to exit in Console World. Copy/paste this code into the Console and execute it.

````
x <- stops |>
  filter(race %in% c("black", "white")) |>
  mutate(race = str_to_title(race), 
         sex = str_to_title(sex))

fit_stops <- linear_reg() |>
    set_engine("lm") |>
    fit(arrested ~ sex + race*zone, data = x)
````

CP/CR.

```{r courage-10}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

Just because something exists in the tutorial (or in the QMD) does not mean that it is in the Console. You should be aware of what exists in Console World, which is generally called your "workspace."

### Exercise 11

In the Console, load the **[easystats](https://easystats.github.io/easystats/)** package. CP/CR.

```{r courage-11}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

We don't add **easystats** to the QMD because we are only using it for an interactive check of our fitted model. However, the [easystats ecosystem](https://easystats.github.io/easystats/) has a variety of interesting functions and packages which you might want to explore.

### Exercise 12

You may skip this question.

<!-- DK: Why does check_predictions() produce an empty plot? Too many observations? -->

<!-- In the Console, run `check_predictions(extract_fit_engine(fit_stops))`. CP/CR. -->

```{r courage-12}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

The purpose of `check_predictions()` is to compare your actual data (in green) with data that has been simulated from your fitted model, i.e., from your data generating mechanism. If your DGM is reasonable, then data simulated from it should not look too dissimilar from your actual data. Of course, it won't look exactly the same because of randomness, both in the world and in your simulation. But the actual data should be within the range of outcomes that your DGM simulates with `check_predictions()`.

### Exercise 13

Ask AI to create $\LaTeX$ code for this model, including our variable names and estimates for all the coefficients. Because this is a fitted model, the dependent variable will have a "hat" and the formula will not include an error term. 

Add the code to your QMD. `Cmd/Ctrl + Shift + K`.

Make sure the resulting display looks good. For example, you don't want an absurd number of figures to the right of the decimal. If the model is too long, you will need to spread it across several lines. You may need to go back-and-forth with the AI a few times.

Once the $\LaTeX$ code looks good, paste it below.

```{r courage-13}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

Our formula looks like:

$$
\widehat{\text{arrested}} = 0.177 
+ 0.0614 \cdot \text{sex}_{\text{Male}} 
- 0.0445 \cdot \text{race}_{\text{White}} 
+ 0.0146 \cdot \text{zone}_{\text{B}} 
+ 0.00610 \cdot \text{zone}_{\text{C}} 
+ 0.0781 \cdot \text{zone}_{\text{D}} 
+ 0.00190 \cdot \text{zone}_{\text{E}} 
- 0.00271 \cdot \text{zone}_{\text{F}} 
+ 0.0309 \cdot \text{zone}_{\text{G}} 
+ 0.0757 \cdot \text{zone}_{\text{H}} 
+ \text{(interaction terms for race and zone)}
$$

It was created with:

````
$$
\widehat{\text{arrested}} = 0.177 
+ 0.0614 \cdot \text{sex}_{\text{Male}} 
- 0.0445 \cdot \text{race}_{\text{White}} 
+ 0.0146 \cdot \text{zone}_{\text{B}} 
+ 0.00610 \cdot \text{zone}_{\text{C}} 
+ 0.0781 \cdot \text{zone}_{\text{D}} 
+ 0.00190 \cdot \text{zone}_{\text{E}} 
- 0.00271 \cdot \text{zone}_{\text{F}} 
+ 0.0309 \cdot \text{zone}_{\text{G}} 
+ 0.0757 \cdot \text{zone}_{\text{H}} 
+ \text{(interaction terms for race and zone)}
$$
```` 

First, we have replaced the parameters with our best estimates. Second, the left-hand side variable is $\widehat{\text{arrested}}$ instead of $\text{arrested}$ because this formula generates our estimated `arrested`. A hat indicates an estimated value. 

**This is our data generating mechanism.**

Of course, there is randomness built into the DGM, but we won't worry about that detail for now.

A data generating mechanism is just a formula, something which we can write down and implement with computer code.

### Exercise 14

Create a new code chunk in your QMD. Add a code chunk option: `#| cache: true`. Copy/paste the R code for the final model into the code chunk, assigning the result to `fit_stops`. (This will include the call to `fit()` but not the call to `tidy()` because we want the entire fitted model, not just a table of the estimated parameter values.)

Place your cursor in the QMD file on the `fit_stops` line. Use `Cmd/Ctrl + Enter` to execute that line. Strictly speaking, this step is unnecessary because we already added `fit_stops` to our workspace above. But ensuring that everything in the QMD is also in the Console is a good habit.

`Cmd/Ctrl + Shift + K`. It may take some time to render your QMD, depending on how complex your model is. But, by including `#| cache: true` you cause Quarto to cache the results of the chunk. The next time you render your QMD, as long as you have not changed the code, Quarto will just load up the saved fitted object.

At the Console, run:

```
tutorial.helpers::show_file("stops.qmd", chunk = "Last")
```

CP/CR.

```{r courage-14}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 8)
```

### 

To confirm, `Cmd/Ctrl + Shift + K` again. It should be quick.

### Exercise 15

Add `*_cache` to `.gitignore` file. Cached objects are often large. They don't belong on Github.

At the Console, run:

```
tutorial.helpers::show_file(".gitignore")
```

CP/CR.

```{r courage-15}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 6)
```

###

Because of the change in your `.gitignore` (assuming that you saved it), the cache directory should not appear in the Source Control panel because Git is ignoring it, as instructed. Commit and push. 

### Exercise 16

In the Console, run `tidy()` on `fit_stops` with the argument `conf.int` set equal to `TRUE`. This returns 95% intervals for all the parameters in our model.

`tidy()` is part of the [**broom**](https://broom.tidymodels.org/) package, used to summarize information from a wide variety of models.


```{r courage-16, exercise = TRUE}

```

```{r courage-16-hint-1, eval = FALSE}
tidy(..., conf.int = ...)
```

```{r courage-16-test, include = FALSE}
tidy(fit_stops, conf.int = TRUE)
```

### Exercise 17

Create a new code chunk in your QMD. Ask AI to help you make a nice looking table from the tibble which is returned by `tidy()`. You don't have to include all the variables which `tidy()` produces. We often just show the estimate and the confidence intervals.

You might need to add some new libraries, e.g., **[tinytable](https://vincentarelbundock.github.io/tinytable/)**, **[knitr](https://yihui.org/knitr/)**, **[gt](https://gt.rstudio.com/)**, **[kableExtra](https://haozhu233.github.io/kableExtra/)**, **[flextable](https://davidgohel.github.io/flextable/)**, **[modelsummary](https://modelsummary.com/)**, et cetera, to the `setup` code chunk, if you use any functions from these packages, all of which have strengths and weaknesses for making tables.

Insert your table code into the QMD. (If you need a new library, and you probably will, place it in the `setup` code chunk along with the other libraries you have already loaded.)

`Cmd/Ctrl + Shift + K`. 

At the Console, run:

```
tutorial.helpers::show_file("stops.qmd", chunk = "Last")
```

CP/CR.

```{r courage-17}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 12)
```

###

This is our table:

```{r}
tidy(fit_stops_logistic, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), ~round(., 3))) |>
  knitr::kable(
    caption = "Logistic Regression Estimates for Arrest Probability (Source: Traffic stops dataset filtered for Black and White drivers)"
  )
```

It was created with:

```
fit_stops_logistic <- logistic_reg() |>
  set_engine("glm") |>
  fit(as.factor(arrested) ~ sex + race, data = x)

tidy(fit_stops_logistic, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), ~round(., 3))) |>
  knitr::kable(
    caption = "Logistic Regression Estimates for Arrest Probability (Source: Traffic stops dataset filtered for Black and White drivers)"
  )
```

At the very least, your table should include a title and a caption with the data source. The more you use AI, the better you will get at doing so.

### Exercise 18

Add a sentence to your project summary. 

Explain the structure of the model. Something like: "I/we model XX [the concept of the outcome, not the variable name], [insert description of values of XX], as a [linear/logistic/multinomial/ordinal] function of XX [and maybe other covariates]." 


Recall the beginning of our version of the summary:

> Using data from a study of New Orleans drivers, we seek to understand the relationship between driver race and the probabilty of getting arrested during a traffic stop. However, our data from both our Preceptor Table and our dataset may not fully represent the population as both may not be from the same time frame and some of our data may come from biased officers, who may target certain groups of individuals.

<!-- DK: Below summary is very awkward. -->
<!-- AR: Summary should not yet discuss findings? Isn't that supposed to happen after temperance? -->
```{r courage-18}
question_text(NULL,
	message = "Using data from a study of New Orleans drivers, we seek to understand the relationship between driver race and the probabilty of getting arrested during a traffic stop. However, our data from both our Preceptor Table and our dataset may not fully represent the population as both may not be from the same time frame and some of our data may come from biased officers, who may target certain groups of individuals. However, these concerns don't appear to be valid in a substantial manner in either dataset, allowing us to continue in our process. We modeled `arrested` as a linear function of both `sex` and the product of `race` and `zone`. From this, we examined that Males are more likely to get arrested than Females.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Read our answer. It will not be the same as yours. You can, if you want, change your answer to incorporate some of our ideas. Do not copy/paste our answer exactly. Add your two sentences, edited or otherwise, to the summary paragraph portion of your QMD. `Cmd/Ctrl + Shift + K`, and then commit/push.

## Temperance
### 

*Temperance is a bridle of gold; he, who uses it rightly, is more like a god than a man.* - Robert Burton

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
	message = "Temperance uses the data generating mechanism to answer the question with which we began. Humility reminds us that this answer is always a lie. We can also use the DGM to calculate many similar quantities of interest, displaying the results graphically.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Courage gave us the data generating mechanism. Temperance guides us in the use of the DGM — or the “model” — we have created to answer the question(s) with which we began. We create posteriors for the quantities of interest. 

### Exercise 2

Before using the DGM, we should make sure that we can interpret it.

Recall the values for the parameters in our data generating mechanism:

```{r}
fit_stops |> 
  tidy(conf.int = TRUE) |> 
  select(term, estimate, conf.low, conf.high) |> 
  slice(1:5)
```
(Table truncated: only the first 5 rows shown)

Write a sentence interpreting the 0.06 estimate for `sexMale`.

```{r temperance-2}
question_text(NULL,
	message = "When comparing men with women, men have a 0.06 higher value for arrested, meaning that they are more likely to get arrested, relative to women, conditional on the other variables in the model.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

###

### Exercise 3

```{r}
fit_stops |> 
  tidy(conf.int = TRUE) |> 
  select(term, estimate, conf.low, conf.high) |> 
  slice(1:5)
```
(Table truncated: only the first 5 rows shown)

Write a sentence interpreting the -0.04 estimate for `raceWhite`.

```{r temperance-3}
question_text(NULL,
	message = "When comparing White people with Black people, White people have a lower value for arrested, meaning that they are less likely to get arrested, relative to Black people conditional on the other variables in the model.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

###

### Exercise 4

```{r}
fit_stops |> 
  tidy(conf.int = TRUE) |> 
  select(term, estimate, conf.low, conf.high) |> 
  slice(1:5)
```
(Table truncated: only the first 5 rows shown)

Write a sentence interpreting the 0.18 estimate for `(Intercept)`.

```{r temperance-4}
question_text(NULL,
	message = "The intercept represents the baseline predicted value of the outcome when all the predictor variables are at their reference levels. In our case this means that there is a 18% chance that a Black female living in Zone A will get arrested.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### Exercise 5

In the end, we don't really care about parameters, much less how to interpret them. Parameters are *imaginary*, like unicorns. We care about answers to our questions. Parameters are tools for answering questions. They aren't interesting in-and-of themselves. *In the modern world, all parameters are nuisance parameters.* 

Add `library(marginaleffects)` to the QMD file.

Place your cursor in the QMD file on the `library(marginaleffects)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(marginaleffects)` to be copied down to the Console and then executed. 

CP/CR.

```{r temperance-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

We should be modest in the claims we make. The posteriors we create are never the “truth.” The assumptions we made to create the model are never perfect. Yet decisions made with flawed posteriors are almost always better than decisions made without them.

### Exercise 6

What is the specific question we are trying to answer? 

```{r temperance-6}
question_text(NULL,
	message = "We are generally investigating the likelihood of getting arrested during a Traffic Stop in New Orleans. We are specifically interested in the probabilty of arrest for a Black driver.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Data science projects begin with a decision which we face. To make that decision wisely, we would like to have good estimates of many unknown numbers. Yet, in order to make progress, we need to drill down to one specific question. This leads to the creation of a data generating mechanism, which can now be used to answer lots of questions, thus allowing us to explore the original decision more broadly.

### Exercise 7

In the Console, run `predictions()` on `fit_stops`.

CP/CR.

```{r temperance-7}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

```{r temperance-7-test, include = FALSE}
predictions(fit_stops)
```


###

The `predictions()` function returns a data frame with one row for each observation in the dataset used to fit the model. In this case, `predictions(fit_stops)` returns **378,467 rows**, because the input data has **378,467 observations**. Each row contains the model’s estimated probability of arrest for that individual stop, along with standard errors and confidence intervals. These values reflect the uncertainty around each prediction and are useful for understanding the model’s expected outcomes across different groups.

### Exercise 8
In the Console, run `plot_predictions()` on `fit_stops` with `by` equal to `"sex"`.

CP/CR.

```{r temperance-8}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

```{r temperance-8-test, include = FALSE}
plot_predictions(fit_stops, by = "sex")
```


###

Using plot_predictions(fit_stops, by = "sex"), this plot shows the average predicted arrest rate for each sex group based on their actual data. It reflects real-world differences, since other variables like race and zone are not balanced. This helps us see how outcomes differ across sex in practice.

### Exercise 9

In the Console, run `plot_predictions()` on `fit_stops` with `condition` equal to `"sex"`.

CP/CR.

```{r temperance-9}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

```{r temperance-9-test, include = FALSE}
plot_predictions(fit_stops, condition = "sex")
```


###

<!-- AR: I wish that we could look at the plots here and better examine the differences that result from the "condition", versus "by" calls -->

Using plot_predictions(fit_stops, condition = "sex"), this plot predicts arrest rates for males and females while holding other variables constant or balanced. It isolates the effect of sex by ensuring all else is equal. This shows what the model predicts if everyone had the same background variables.

###

When you use `by = "sex"`, the graph shows the real arrest rates for men and women as they actually are in the data, including all other differences like age or location. Using `condition = "sex"` shows what arrest rates would look like if men and women were the same in every other way except sex, isolating the effect of sex alone. The graphs look almost the same because sex is the main factor affecting arrest rates, and other variables don’t change the results much.

### Exercise 10

In the Console, run `plot_predictions()` on `fit_stops` with `condition` equal to `c("sex", "race")`.

CP/CR.

```{r temperance-10}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

```{r temperance-10-test, include = FALSE}
plot_predictions(fit_stops, condition = c("sex", "race"))
```

### 

This plot shows predicted probabilities of being arrested by sex and race, with vertical lines representing uncertainty (95% confidence intervals). Black males have the highest predicted probability (around 0.32), and White females the lowest (around 0.24). While the estimates are fairly precise—indicated by the short confidence intervals—there is still some overlap, especially between Black and White males, suggesting some uncertainty in how distinct those groups are.

### Exercise 11

Work with AI to create a beautiful plot starting with the output to the above call to `plot_predictions()`. Do this in your QMD since that is much easier than working in the Console directly.

Your title should highlight the key variables. Your subtitle should describe an important takeaway, the sentence/conclusion which readers will, you hope, remember. Your caption should mention the data source. Your axis labels should look nice.

This plot is not directly connected to your question. It answers lots of questions! It might be used by lots of different people. 

Copy the code for your plot here:

```{r temperance-11}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 15)
```

### 

Our plot:

```{r}
plot_predictions(fit_stops$fit,
                 newdata = "balanced",
                 condition = c("zone", "race", "sex"),
                 draw = FALSE) |> as_tibble() |> 
  group_by(zone, sex) |>
  mutate(sort_order = estimate[race == "Black"]) |>
  ungroup() |>
  mutate(zone = reorder_within(zone, sort_order, sex)) |>
  ggplot(aes(x = zone, 
             color = race)) +
  geom_errorbar(aes(ymin = conf.low, 
                    ymax = conf.high), 
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  geom_point(aes(y = estimate), 
             size = 1, 
             position = position_dodge(width = 0.5)) +
  facet_wrap(~ sex, scales = "free_x") +
  scale_x_reordered() +
  theme(axis.text.x = element_text(size = 8)) +
  scale_y_continuous(labels = percent_format())
```


Our code:

````
plot_predictions(fit_stops$fit,
                 newdata = "balanced",
                 condition = c("zone", "race", "sex"),
                 draw = FALSE) |> as_tibble() |> 
  group_by(zone, sex) |>
  mutate(sort_order = estimate[race == "Black"]) |>
  ungroup() |>
  mutate(zone = reorder_within(zone, sort_order, sex)) |>
  ggplot(aes(x = zone, 
             color = race)) +
  geom_errorbar(aes(ymin = conf.low, 
                    ymax = conf.high), 
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  geom_point(aes(y = estimate), 
             size = 1, 
             position = position_dodge(width = 0.5)) +
  facet_wrap(~ sex, scales = "free_x") +
  scale_x_reordered() +
  theme(axis.text.x = element_text(size = 8)) +
  scale_y_continuous(labels = percent_format())
````

Data science often involves this-back-and-forth style of work. First, we need to make a single chunk of code, in this case, a new plot, work well. This requires interactive work between the QMD and the Console. Second, we need to ensure that the entire QMD runs correctly on its own.

### Exercise 12

Finalize the new graphics code chunk in your QMD. `Cmd/Ctrl + Shift + K` to ensure that it all works as intended. 

At the Console, run:

```
tutorial.helpers::show_file("stops.qmd", chunk = "Last")
```

CP/CR.

```{r temperance-12}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 10)
```


### 

Always [remember](https://en.wikipedia.org/wiki/Map%E2%80%93territory_relation): *The map is not the territory.* A beautiful graphic tells a story, but that story is always an imperfect representation of reality. Our models depend on assumptions, assumptions which are never completely true.

### Exercise 13

Write the last sentence of your summary paragraph. It describes at least one quantity of interest (QoI) and provides a measure of uncertainty about that QoI. It is OK if this QoI is not the one that you began with. The focus of a data science project often changes over time. It is also OK to discuss more than one QoI. Do whatever seems reasonable, given the context.

```{r temperance-13}
question_text(NULL,
	message = "Using data from a study of New Orleans drivers, we seek to understand the relationship between driver race and the probabilty of getting arrested during a traffic stop. However, our data from both our Preceptor Table and our dataset may not fully represent the population as both may not be from the same time frame and some of our data may come from biased officers, who may target certain groups of individuals. However, these concerns don't appear to be valid in a substantial manner in either dataset, allowing us to continue in our process. We modeled `arrested` as a linear function of both `sex` and the product of `race` and `zone`. From this, we examined that Males are less likely of getting arrested than Females. Focusing on our question, there is no uncertainty associated with our estimate because it, itself, is an expression of uncertainty. We estimate that the probaibility of a Black driver in New Orleans getting arrested during a traffic stop is roughly 25%, compared to roughly 20% for White drivers.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Add a final sentence to your summary paragraph in your QMD as you see fit, but do not copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`.

### Exercise 14

Write a few sentences which explain why the estimates for the quantities of interest, and the uncertainty thereof, might be wrong. Suggest an alternative estimate and confidence interval, if you think either might be warranted.

<!-- AR: I find this question to be broad and not that helpful. Also, this answer is suspect. I especially don't think we should suggest decreasing the floor of confidence intervals based on hunches. -->

```{r temperance-14}
question_text(NULL,
	message = "Based on all of the errors in our data that we collected during Justice, our original data source to base our model on was already flawed. Additionally, our model can only predict estimates on paper, based on an ideal world. It doesn't account for many factors of the real world that we, ourselves, can't even describe. However, our estimate was formed to the best of our abilities given the data on hand, and all I would do to modify it would be to modify the confidence intervals: I would keep the upper confidence interval the same, but would singifcantly decrease the lower confidence interval to roughly 15%-17% to account for the future, where the values would most likely decrease based on the pattern of racism in the US decreasing over time.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Always go back to your Preceptor Table, the information which, if you had it, would make answering your question easy. In almost all real world cases, the Preceptor Table and the data are fairly different, not least because validity never holds perfectly. So, even a perfectly estimated statistical model is rarely as useful as we might like.

### Exercise 15

Rearrange the material in your QMD so that the order is graphic, followed by the paragraph. Doing so, of course, requires sensible judgment. For example, the code chunk which creates the fitted model must occur before the chunk which creates the graphic. You can keep or discard the math and any other material at your own discretion.

`Cmd/Ctrl + Shift + K` to ensure that everything works.

At the Console, run:

```
tutorial.helpers::show_file("stops.qmd")
```

CP/CR.

```{r temperance-15}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 30)
```

### 

This is the version of your QMD file at which your teacher is most likely to look closely.

### Exercise 16

Publish your rendered QMD to GitHub Pages. In the Terminal --- not the Console! --- run:

````
quarto publish gh-pages stops.qmd
````

Copy/paste the resulting URL below.

```{r temperance-16}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Commit/push everything.

### Exercise 17

Copy/paste the URL to your Github repo.

```{r temperance-17}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God. If all our assumptions are correct, then our DGM is true, it accurately describes the way in which the world works. There is no better way to predict the future, or to model the past, than to use it. Sadly, this will only be the case with toy examples involving things like coins and dice. We hope that our DGM is close to the true DGM but, since our assumptions are never perfectly correct, our DGM will always be different. The estimated magnitude and importance of that difference is a matter of judgment.

The world confronts us. Make decisions we must.

## Summary
### 

This tutorial supports [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). 

```{r}
plot_predictions(fit_stops$fit,
                 newdata = "balanced",
                 condition = c("zone", "race", "sex"),
                 draw = FALSE) |> as_tibble() |> 
  group_by(zone, sex) |>
  mutate(sort_order = estimate[race == "Black"]) |>
  ungroup() |>
  mutate(zone = reorder_within(zone, sort_order, sex)) |>
  ggplot(aes(x = zone, 
             color = race)) +
  geom_errorbar(aes(ymin = conf.low, 
                    ymax = conf.high), 
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  geom_point(aes(y = estimate), 
             size = 1, 
             position = position_dodge(width = 0.5)) +
  facet_wrap(~ sex, scales = "free_x") +
  scale_x_reordered() +
  theme(axis.text.x = element_text(size = 8)) +
  scale_y_continuous(labels = percent_format())
```

The plot shows the model-predicted probability of arrest for Black and White drivers across different zones, separated by sex. Within each sex group, zones are ordered by the estimated arrest probability for Black drivers. Across most zones, Black drivers have a higher predicted probability of arrest than White drivers, and this pattern appears consistently for both men and women. The size of the difference varies by zone.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
