---
title: "Smokes"
author: David Kane
tutorial:
  id: 09-smokes
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #09 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(tidymodels)
library(primer.tutorials)
library(broom)
library(marginaleffects)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = "90%")
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

fit_smokes <- logistic_reg(engine = "glm") |>
  fit(smoke ~ age + sex, data = smokes)
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports the [Smokes](https://ppbds.github.io/primer/09-smokes.html) chapter of [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

The world confronts us. Make decisions we must.

Imagine that you are a public-health analyst at a state health department. The Public Health Commissioner has a big-picture goal --- reduce the smoking-attributable disease burden in the state --- and trusts you to figure out where the next anti-smoking campaign's outreach budget should go. Many causal and predictive models would help: who is most likely to be smoking now, which cessation messages move which subgroups, which geographic and demographic segments are the most cost-effective targets, how outreach effects fade once a campaign ends. This tutorial builds just one of them: the probability of being an ever-smoker as a function of age and sex --- the basic demographics on which outreach can actually be segmented when buying ads or staffing clinics. The estimate alone won't determine the campaign, but the demographic-targeting piece is one good input. There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

The four virtues structure every Primer chapter; Smokes is our first *logistic* example. The model's coefficients sit on the log-odds scale --- not a scale anyone reads directly --- so most of Temperance will be about turning the fit into probability-scale answers with `marginaleffects`.

### Exercise 2

You should be working inside a GitHub repo named `smokes`, opened in a Codespace from the [`PPBDS/codespace-starter`](https://github.com/PPBDS/codespace-starter) template. If you are not there yet, create that repo and open it in a Codespace now --- see the [package README](https://github.com/PPBDS/primer/tree/main/primer.tutorials#working-environments-and-repo-setup) for setup instructions or for working locally.

Once you're inside the `smokes` repo, create a new Quarto document titled `"Smokes"` with yourself as the author, render it, and save it as `analysis.qmd`.

Create a `.gitignore` file with `analysis_files` on the first line followed by a blank line. Save and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The `smokes` tibble is a 1,000-row teaching cut of NHANES 2009--2012 restricted to adults 20--80, with `smoke` as a Yes/No factor following the canonical NHANES "ever-smoker" definition (≥ 100 cigarettes in a lifetime).

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.tutorials)` in a new code chunk. Render the file. Add `#| message: false` to suppress the messages, and add the following to the YAML header to hide the code:

```
execute:
  echo: false
```

Render again. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    #| message: false\n    library(tidyverse)\n    library(primer.tutorials)\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The `smokes` tibble in **[primer.tutorials](https://github.com/PPBDS/primer.tutorials)** is a 1,000-row teaching cut from NHANES 2009--2012, restricted to adults aged 20--80 with non-missing values on smoking status, age, and sex. The outcome `smoke` is coded `Yes` if the respondent reports having smoked at least 100 cigarettes in their lifetime --- the canonical NHANES "ever-smoker" definition, and the bar essentially every U.S. tobacco-policy researcher uses.

### Exercise 4

In the R Terminal, type `?smokes` and paste the Description below.

```{r introduction-4}
question_text(NULL,
  message = "A 1,000-row sample drawn from the educational NHANES dataset bundled in the NHANES CRAN package (a 10,000-row subset of the 2009-2012 NHANES waves). Restricted to adults aged 20-80 with non-missing values on smoking status, age, and sex. Used by the 09-smokes tutorial as a Medium-tier predictive logistic regression problem: who has ever smoked at least 100 cigarettes, as a function of age and sex.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The threshold *"ever smoked at least 100 cigarettes"* is the standard public-health definition of an *ever-smoker*. It captures lifetime exposure rather than current status: someone who smoked daily as a college student and quit twenty years ago is still an ever-smoker.

### Exercise 5

Smoking status is the broad topic of this tutorial. Given that topic, which variable in `smokes` should we use as our outcome variable?

```{r introduction-5}
question_text(NULL,
  message = "Our outcome variable is `smoke` --- a factor with levels `\"No\"` and `\"Yes\"` indicating whether the respondent has ever smoked at least 100 cigarettes in their lifetime. The 100-cigarette threshold is the standard public-health definition of an ever-smoker.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

We will use `smoke` as our outcome variable. Below is the share of respondents in the sample who answer `Yes`, broken out by sex.

```{r}
#| echo: false
#| message: false
#| warning: false
smokes |>
  count(sex, smoke) |>
  group_by(sex) |>
  mutate(share = n / sum(n)) |>
  filter(smoke == "Yes") |>
  ggplot(aes(x = sex, y = share, fill = sex)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = scales::label_percent(accuracy = 1)(share)),
            vjust = -0.5, size = 5) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 0.6)) +
  labs(
    title    = "Share of Ever-Smokers in Our 1,000-Adult Sample",
    subtitle = "Men are about 13 percentage points more likely to be ever-smokers than women",
    x        = NULL,
    y        = NULL,
    caption  = "Source: 1,000-row NHANES 2009-2012 cut via primer.tutorials::smokes"
  ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.y     = element_blank(),
        axis.ticks.y    = element_blank(),
        panel.grid      = element_blank())
```

### Exercise 6

Specify two different groups of adults which have different values for `sex` and which might have different probabilities of being ever-smokers.

```{r introduction-6}
question_text(NULL,
  message = "Consider two groups: women aged 30 and men aged 70. The two groups differ both on `sex` (Female vs. Male) and on `age` (30 vs. 70). The plot above already suggests men are more likely than women to be ever-smokers; if smoking propensity also rises with age, the older men in this comparison may be even more likely than the younger women to be ever-smokers.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

In predictive models, do not use words like *"cause,"* *"influence,"* or *"impact"* --- the best phrasing is in terms of *differences* between groups of adults with different values for the covariates. The fact that older men are more often ever-smokers does not mean that being older or being male *causes* smoking; it means that, in this sample, that group has a higher share. Predictive models have no treatments --- only covariates.

### Exercise 7

Write a predictive question connecting the outcome `smoke` to `age` and `sex`. The question should specify a particular number to compute.

```{r introduction-7}
question_text(NULL,
  message = "What is the difference in the probability of being an ever-smoker between a 30-year-old woman and a 70-year-old man?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The 100-cigarette threshold is what makes the outcome a clean lifetime measure: someone who smoked heavily for two years in their twenties and quit will still register as `Yes`, even though they are no longer a current smoker. The question we are asking is about *ever-having-smoked*, not about *currently-smoking* --- and a state health department deciding where to put cessation outreach probably cares about both.


## Wisdom
###

Data science starts with some broad questions and a data set which might help us to answer them.

The specific question: *What is the difference in the probability of being an ever-smoker between a 30-year-old woman and a 70-year-old man?*

We will work from a 1,000-adult sample of NHANES 2009--2012, available in the `smokes` tibble of the **primer.tutorials** package.

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
  message = "Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

> *The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* --- John W. Tukey

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The Preceptor Table only includes the variables mentioned in the question. For a predictive model, that is one outcome column plus the covariates we are conditioning on; no treatment, no potential outcomes.

### Exercise 3

The Introduction plot showed the share of ever-smokers split by sex. Now look at how `smoke` varies with `age`, our other covariate. Write R code that plots the share of ever-smokers vs. age in five-year bins, faceted by sex, with a smoothed line and the y-axis labeled as a percentage. Paste the code below.

```{r wisdom-3}
question_text(NULL,
  message = "    smokes |>\n      mutate(age_bin = 5 * (age %/% 5)) |>\n      group_by(sex, age_bin) |>\n      summarize(share_yes = mean(smoke == \"Yes\"), .groups = \"drop\") |>\n      ggplot(aes(x = age_bin, y = share_yes, color = sex)) +\n      geom_point(size = 3) +\n      geom_smooth(se = FALSE, method = \"loess\", linewidth = 1.1) +\n      scale_y_continuous(labels = scales::label_percent(),\n                         limits = c(0, 1)) +\n      facet_wrap(~ sex) +\n      labs(\n        title    = \"Share of Ever-Smokers by Age and Sex\",\n        subtitle = \"Older respondents are more likely to be ever-smokers, especially among men\",\n        x        = \"Age (binned to 5 years)\",\n        y        = NULL,\n        caption  = \"Source: 1,000-row NHANES cut via primer.tutorials::smokes\"\n      ) +\n      theme_minimal() +\n      theme(legend.position = \"none\")",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 18)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
smokes |>
  mutate(age_bin = 5 * (age %/% 5)) |>
  group_by(sex, age_bin) |>
  summarize(share_yes = mean(smoke == "Yes"), .groups = "drop") |>
  ggplot(aes(x = age_bin, y = share_yes, color = sex)) +
  geom_point(size = 3) +
  geom_smooth(se = FALSE, method = "loess", linewidth = 1.1) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 1)) +
  facet_wrap(~ sex) +
  labs(
    title    = "Share of Ever-Smokers by Age and Sex",
    subtitle = "Older respondents are more likely to be ever-smokers, especially among men",
    x        = "Age (binned to 5 years)",
    y        = NULL,
    caption  = "Source: 1,000-row NHANES cut via primer.tutorials::smokes"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Exercise 4

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-4}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

This problem is predictive, so there is exactly one outcome column (`smoke`) and no treatment. We compare two groups of adults with different covariate values --- in our case, a 30-year-old woman versus a 70-year-old man.

### Exercise 5

What are the units for this problem?

```{r wisdom-5}
question_text(NULL,
  message = "Each row is one U.S. adult aged 20-80, observed at the time of the NHANES interview. The full educational NHANES dataset covers about 6,000 such adults across the 2009-2012 waves; our analysis sample is 1,000 of them, drawn at random with `set.seed(2026)` so the sample is reproducible.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Specifying the unit forces a question precise enough to answer. We model individual adults here, but a state health department actually wants to know about *aggregates* --- how many ever-smokers are in a target ZIP code, not which specific person.

### Exercise 6

What is the outcome variable for this problem?

```{r wisdom-6}
question_text(NULL,
  message = "Keep track of two outcome variables: the one in our Preceptor Table and the one in our data. In this case, both are `smoke` --- a factor taking the values `\"Yes\"` and `\"No\"` based on the standard public-health threshold of having smoked at least 100 cigarettes in one's lifetime.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The outcome we really care about for an anti-smoking campaign --- *"who is currently smoking and could plausibly be moved to quit?"* --- is fuzzier than what `smoke` measures. `smoke` records *ever-smoker* status. Someone who smoked heavily in their twenties and quit thirty years ago still answers `Yes`, but is not the target of a cessation program.

### Exercise 7

What is a covariate which you think might be useful for this problem, regardless of whether or not it is in the data?

```{r wisdom-7}
question_text(NULL,
  message = "Educational attainment is a strong candidate. In U.S. data, smoking rates have fallen much faster among college graduates than among adults without a four-year degree, so education would help separate two adults of the same age and sex. We do not have education in this trimmed `smokes` tibble, but the original NHANES dataset does.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The term *covariates* is used in three ways: variables that *might* matter, variables we have data for, and variables we end up putting in the model. Education would matter; it is in the source data; we did not include it in this teaching cut. Most candidates a researcher lists never make it into the final model.

### Exercise 8

What are the treatments, if any, for this problem?

```{r wisdom-8}
question_text(NULL,
  message = "There are no treatments. This is a predictive model: we compare ever-smoker status between groups of adults defined by their age and sex, and do not assume that any of those covariates can be manipulated. Predictive models have no treatments --- only covariates.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Without a treatment, the language we use changes. We talk about *differences in expected probabilities between groups*, not about *causal effects of one variable on another*. Same data, same model, different framing.

### Exercise 9

What moment in time does the Preceptor Table refer to?

```{r wisdom-9}
question_text(NULL,
  message = "The present, conventionally taken to be the year 2026 in which we are designing the campaign. Our data, by contrast, was collected during the 2009-2012 NHANES waves, so there is more than a decade between when the data was gathered and when our question is being asked.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

A Preceptor Table can never really refer to an exact instant in time --- *"the present"* is a smear, not a moment. The 13-year gap between data and Preceptor Table is going to come up again in Justice when we discuss stability.

### Exercise 10

Describe in words the Preceptor Table for this problem.

```{r wisdom-10}
question_text(NULL,
  message = "The Preceptor Table has three columns: one identifying each adult, one outcome column (`Ever Smoked`), and two covariate columns (`Age` and `Sex`). Each row represents one U.S. adult in 2026. Because this is a predictive model, the outcome column is single-valued and observable in principle for every adult --- there are no potential outcomes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

```{r}
#| echo: false
#| results: asis
pre_tibble <- tibble::tribble(
  ~`Adult`,            ~`Ever Smoked`, ~`Age`, ~`Sex`,
  "Maria Alvarez",     "No",           "30",   "Female",
  "Anthony Carter",    "Yes",          "70",   "Male",
  "...",               "...",          "...",  "...",
  "Linda Whitfield",   "No",           "55",   "Female"
)

pre_title_footnote     <- "If all the information in this table were available, we could answer the question: What is the difference in the probability of being an ever-smoker between a 30-year-old woman and a 70-year-old man?"
pre_units_footnote     <- "Each row is one U.S. adult aged 20-80 in 2026, the year the campaign would be deployed. The example rows above use plausible names chosen to span the relevant age and sex range; missing rows represent the rest of the adult population we would target."
pre_outcome_footnote   <- "Whether the adult has ever smoked at least 100 cigarettes in their lifetime --- the standard public-health definition of an ever-smoker. Factor with values `Yes` and `No`."
pre_covariate_footnote <- "Age in years (integer, 20-80) and self-reported sex (Female or Male). Sex is shown as a binary because that is what the data collected; non-binary respondents would not be representable in the table as drawn."

pre_gt_html <- gt::gt(pre_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",       id = "unit_span",
                  columns = c(`Adult`)) |>
  gt::tab_spanner(label = "Outcome",    id = "outcome_span",
                  columns = c(`Ever Smoked`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Age`, `Sex`)) |>
  gt::cols_align(align = "left",  columns = c(`Adult`, `Ever Smoked`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Age`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "outcome_span", "covariates_span"))
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 11

Write the first two sentences of the summary paragraph. The first introduces the topic; the second introduces the data and the specific question.

```{r wisdom-11}
question_text(NULL,
  message = "Lifetime smoking is one of the strongest predictors of long-term health outcomes, and a state health department designing an anti-smoking campaign needs to know who its likely targets are. Using a 1,000-adult sample from the 2009-2012 NHANES waves, we estimate the difference in the probability of being an ever-smoker between a 30-year-old woman and a 70-year-old man.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Read our answer. It will not be the same as yours. You can change your answer to incorporate some of our ideas, but do not copy/paste exactly. Add your two sentences to the summary paragraph in your QMD. Render the QMD, then commit/push.


## Justice
###

Wisdom gives us the Preceptor Table and the data.

```{r}
#| echo: false
#| results: asis
pre_tibble_j <- tibble::tribble(
  ~`Adult`,            ~`Ever Smoked`, ~`Age`, ~`Sex`,
  "Maria Alvarez",     "No",           "30",   "Female",
  "Anthony Carter",    "Yes",          "70",   "Male",
  "...",               "...",          "...",  "...",
  "Linda Whitfield",   "No",           "55",   "Female"
)

pre_gt_html_j <- gt::gt(pre_tibble_j, id = "preceptor_tbl_j") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",       id = "unit_span",
                  columns = c(`Adult`)) |>
  gt::tab_spanner(label = "Outcome",    id = "outcome_span",
                  columns = c(`Ever Smoked`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Age`, `Sex`)) |>
  gt::cols_align(align = "left",  columns = c(`Adult`, `Ever Smoked`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Age`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "outcome_span", "covariates_span"))
  ) |>
  gt::opt_css(
    css = "
      #preceptor_tbl_j .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html_j,
  "</div>\n",
  "```\n",
  sep = ""
)
```

```{r}
#| echo: false
#| results: asis
data_tibble <- tibble::tribble(
  ~`Respondent`,        ~`Year`, ~`Ever Smoked`, ~`Age`, ~`Sex`,
  "Sarah Chen",         "2010",  "No",           "32",   "Female",
  "Robert Davis",       "2011",  "Yes",          "68",   "Male",
  "...",                "...",   "...",          "...",  "...",
  "Patricia Williams",  "2012",  "No",           "44",   "Female"
)

data_title_footnote     <- "National Health and Nutrition Examination Survey (NHANES) 2009--2012, conducted by the U.S. Centers for Disease Control and Prevention. Adults aged 20--80 are interviewed in person; smoking status is self-reported as part of the questionnaire."
data_units_footnote     <- "Each row is one NHANES respondent at the time of their interview. NHANES is a multi-year, multi-stage probability sample drawn from the U.S. civilian non-institutionalized population; our 1,000-row teaching cut takes a uniform random subsample of the educational NHANES file (`set.seed(2026)`)."
data_outcome_footnote   <- "Self-reported response to the question *Have you smoked at least 100 cigarettes in your entire life?* Stored as a factor with levels `No` and `Yes`. The 100-cigarette threshold is what NHANES uses to define an ever-smoker, so all respondents are coded the same way."
data_covariate_footnote <- "Age is reported in completed years; respondents at the upper end (80+) are top-coded at 80 in the educational file. Sex is reported by the respondent and recorded as Female or Male."

data_gt_html <- gt::gt(data_tibble, id = "data_tbl") |>
  gt::tab_header(title = "Data: NHANES 2009--2012, Adults Aged 20--80") |>
  gt::tab_spanner(label = "Unit/Time",  id = "unit_span",
                  columns = c(`Respondent`, `Year`)) |>
  gt::tab_spanner(label = "Outcome",    id = "outcome_span",
                  columns = c(`Ever Smoked`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Age`, `Sex`)) |>
  gt::cols_align(align = "left",  columns = c(`Respondent`, `Ever Smoked`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Age`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "outcome_span", "covariates_span"))
  ) |>
  gt::tab_footnote(footnote = data_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = data_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = data_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = data_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #data_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  data_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 1

In your own words, describe the components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
  message = "Justice concerns the Population Table, the four key assumptions which underlie it (validity, stability, representativeness, and unconfoundedness), and selects the formula for the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Justice surfaces the worries you (or your critics) might have about whether your model will work. The "formula for the data generating mechanism" has two pieces with formal names: a probability family (the distribution of the outcome --- here, Bernoulli, since the outcome is binary) and a link function (how the outcome's expected value depends on the covariates --- here, the logit). We will use those terms freely starting in the Hard tutorials.

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Validity is always about the columns, never the rows. The names of two columns matching across tables is no guarantee that they measure the same thing.

### Exercise 3

Provide one reason why the assumption of validity might not hold for the outcome variable `smoke` or for one of the covariates.

```{r justice-3}
question_text(NULL,
  message = "The `smoke` column in NHANES is *self-reported*: respondents answer the question themselves rather than being verified against pharmacy records or biomarkers. Adults who smoked occasionally as teenagers may misremember whether they ever crossed the 100-cigarette threshold; current smokers may slightly underreport in clinical settings. The Preceptor Table column `Ever Smoked` is meant to capture the truth, not the answer respondents give to a survey, so the two columns measure the same concept only if recall and reporting bias roughly cancel out across respondents.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Validity, if reasonable, is what lets us treat the Preceptor Table and the data as drawn from the same population. NHANES is one of the better-validated U.S. health surveys --- it includes physical examinations and lab work alongside self-report --- but the smoking question is still self-reported, so this validity worry is real even if not catastrophic.

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The Population Table is almost always larger than the Preceptor Table and the data combined: it covers the full universe of unit/time combinations both samples are drawn from.

### Exercise 5

Specify the unit/time combinations which define each row in this Population Table.

```{r justice-5}
question_text(NULL,
  message = "Each row is one U.S. adult aged 20-80 at one point in time. Data rows are 1,000 NHANES respondents interviewed during 2009-2012. Preceptor rows are U.S. adults aged 20-80 in 2026, the year we are designing the campaign. The same adult could appear in both blocks if they were 20 in 2009 and are still alive in 2026, but most rows are different people.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

```{r}
#| echo: false
#| results: asis
pop_tibble <- tibble::tribble(
  ~Source    , ~`Adult`,             ~`Year`, ~`Ever Smoked`, ~`Age`, ~`Sex`,
  "..."      , "...",                "..."  , "...",          "...",  "...",
  "Data"     , "Sarah Chen",         "2010" , "No",           "32",   "Female",
  "Data"     , "Robert Davis",       "2011" , "Yes",          "68",   "Male",
  "Data"     , "...",                "..."  , "...",          "...",  "...",
  "Data"     , "Patricia Williams",  "2012" , "No",           "44",   "Female",
  "..."      , "...",                "..."  , "...",          "...",  "...",
  "Preceptor", "Maria Alvarez",      "2026" , "No",           "30",   "Female",
  "Preceptor", "Anthony Carter",     "2026" , "Yes",          "70",   "Male",
  "Preceptor", "...",                "..."  , "...",          "...",  "...",
  "Preceptor", "Linda Whitfield",    "2026" , "No",           "55",   "Female",
  "..."      , "...",                "..."  , "...",          "...",  "..."
)

pop_title_footnote     <- "Population: U.S. adults aged 20--80, drawn from a much larger universe that the NHANES sample (Data rows) and the 2026 campaign-target population (Preceptor rows) are both subsets of. The 13--17 year gap between the two blocks is the heart of the stability discussion below."
pop_units_footnote     <- "Each row is one adult at one point in time. Data rows are NHANES respondents (2009--2012); Preceptor rows are U.S. adults in 2026, the year of the campaign."
pop_outcome_footnote   <- "Whether the adult has smoked at least 100 cigarettes in their lifetime. In Data rows, the answer is what the respondent told the NHANES interviewer; in Preceptor rows, the answer is the actual ever-smoker status of each adult in 2026."
pop_covariate_footnote <- "Age in years and self-reported sex. The covariate definitions are the same in both blocks; what differs is the era and the population the rows are drawn from."

pop_gt_html <- gt::gt(pop_tibble, id = "population_tbl") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time",  id = "unit_span",
                  columns = c(`Adult`, `Year`)) |>
  gt::tab_spanner(label = "Outcome",    id = "outcome_span",
                  columns = c(`Ever Smoked`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Age`, `Sex`)) |>
  gt::cols_align(align = "left",  columns = c(`Source`, `Adult`, `Ever Smoked`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Age`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "outcome_span", "covariates_span"))
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 6

In your own words, define the assumption of "stability."

```{r justice-6}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Stability is best understood as a statement about *parameters*, not distributions. The relationship between age, sex, and ever-smoker status in our model is captured by a handful of coefficients --- an intercept, a slope on age, an offset for sex. Stability says those coefficients are the same in our 2009--2012 NHANES sample as they are for U.S. adults in 2026.

### Exercise 7

Provide one reason why stability might not hold in this case.

```{r justice-7}
question_text(NULL,
  message = "Smoking has been declining in the United States for decades. The slope of ever-smoker status on age in 2026 may be steeper than it was in 2009-2012 because the older cohorts in 2026 grew up when smoking was much more common, while younger 2026 adults grew up after most of the cultural and regulatory crackdowns. So the parameter governing the relationship between age and ever-smoker status may have shifted between the era our data was collected and the era we want to predict --- not just because the share of smokers has changed (a distribution shift) but because the way age maps onto smoking status has itself changed (a slope shift).",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

The most common student mistake on stability is to conflate it with a distribution shift. *"Smoking rates have fallen since 2010"* is true, but it does not, by itself, violate stability --- a lower share of ever-smokers across the board would not change the slope of `smoke` on `age`. What violates stability is a change in the *relationship* between covariates and outcome: the slope on age, the offset for sex, the intercept. A change in the *level* of smoking (the marginal share) is not the same as a change in the *parameter* connecting age to smoking. Distributional change is everywhere; parameter change is what hurts us.

### Exercise 8

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: Data → Population → Preceptor Table. In your own words, define the assumption of "representativeness."

```{r justice-8}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Ideally we would like both the Preceptor Table and the data to be random samples from the population. NHANES is closer to that ideal than most surveys --- it is a stratified probability sample drawn from the U.S. civilian non-institutionalized population --- but it is not perfect, and the gaps matter for our question.

### Exercise 9

Provide one reason, involving the relationship between the data and the population, why representativeness might not hold.

```{r justice-9}
question_text(NULL,
  message = "NHANES deliberately oversamples certain demographic subgroups (older adults, Hispanic and non-Hispanic Black respondents, low-income households) so that those subgroups have enough observations for stable estimates. The educational version we draw from also drops the survey weights that would let us reweight back to a representative sample. Our 1,000-row sample therefore over-represents the oversampled groups relative to the U.S. adult population, which might bias our estimate of the relationship between age, sex, and ever-smoker status in directions we cannot easily check.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A non-representative sample does not *guarantee* a biased estimate --- the bias might wash out by chance. But chance is the only mechanism that would save us, and we have no principled reason to expect it to work in our favor. *"Might be biased"* is the correct stance, not *"will be biased."*

### Exercise 10

We use information about the population to make inferences about the Preceptor Table. Provide one reason, involving the relationship between the population and the Preceptor Table, why representativeness might not hold.

```{r justice-10}
question_text(NULL,
  message = "Our Preceptor Table represents adults the state campaign would actually target --- residents of the state, in 2026, reachable by radio, mail, or social-media advertising. The U.S. adult population that NHANES samples from is different in two ways: it is national, not state-specific, and it includes adults who are unreachable by any reasonable outreach effort (incarcerated adults, undocumented adults who avoid census contact, adults who do not speak English or Spanish). Even if our model fits the U.S. adult population perfectly, the Preceptor Table's narrower target population may have a different age-and-sex relationship to ever-smoker status than the broader U.S. population does.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

Stability looks across time periods; representativeness looks within time periods. Both are about whether the rows we have stand in for the rows we care about, but they pull in different directions: stability worries that 2009 isn't 2026, while representativeness worries that NHANES respondents aren't the same kind of people as our state campaign's target list.

### Exercise 11

A statistical model consists of two parts: the probability family and the link function. The probability family is the probability distribution that generates the randomness in our data. Our outcome variable is `smoke` --- binary. What probability family should we use?

```{r justice-11}
question_text(NULL,
  message = "Bernoulli: $Y \\sim \\text{Bernoulli}(\\rho)$, where $\\rho$ is the probability that `smoke = \"Yes\"` for a given adult.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

While you're here, also add `library(tidymodels)` to the setup chunk in `analysis.qmd` so the rest of Courage can call `logistic_reg()` and `fit()`.

### Exercise 12

The link function is the mathematical formula that links our data to the unknown parameters. What is the link function for a binary outcome?

```{r justice-12}
question_text(NULL,
  message = "Logit: $\\log\\!\\big(\\rho/(1-\\rho)\\big) = \\beta_0 + \\beta_1 X_1 + \\cdots + \\beta_n X_n$. Our model will be a logistic regression --- the first one in the curriculum.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

While you're here, also add `library(broom)` to the setup chunk in `analysis.qmd` so we can call `tidy()` on the fitted model in Courage.

### Exercise 13

Write one sentence highlighting a potential weakness in your model. So far our summary paragraph looks like this:

> Lifetime smoking is one of the strongest predictors of long-term health outcomes, and a state health department designing an anti-smoking campaign needs to know who its likely targets are. Using a 1,000-adult sample from the 2009--2012 NHANES waves, we estimate the difference in the probability of being an ever-smoker between a 30-year-old woman and a 70-year-old man.

Your version will be somewhat different.

```{r justice-13}
question_text(NULL,
  message = "One concern is that the relationship between age and ever-smoker status has shifted between when the data was gathered (2009-2012) and the present (2026): smoking is dropping unevenly across age cohorts, so applying a 13-year-old slope to today's adults may misstate the magnitudes by enough to matter for campaign targeting.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a weakness sentence to your summary paragraph in your QMD. Don't copy/paste our answer exactly. Render the QMD, then commit/push.


## Courage
###

Justice gives us the Population Table and the abstract data generating mechanism.

```{r}
#| echo: false
#| results: asis
pop_tibble_c <- tibble::tribble(
  ~Source    , ~`Adult`,             ~`Year`, ~`Ever Smoked`, ~`Age`, ~`Sex`,
  "..."      , "...",                "..."  , "...",          "...",  "...",
  "Data"     , "Sarah Chen",         "2010" , "No",           "32",   "Female",
  "Data"     , "Robert Davis",       "2011" , "Yes",          "68",   "Male",
  "Data"     , "...",                "..."  , "...",          "...",  "...",
  "Data"     , "Patricia Williams",  "2012" , "No",           "44",   "Female",
  "..."      , "...",                "..."  , "...",          "...",  "...",
  "Preceptor", "Maria Alvarez",      "2026" , "No",           "30",   "Female",
  "Preceptor", "Anthony Carter",     "2026" , "Yes",          "70",   "Male",
  "Preceptor", "...",                "..."  , "...",          "...",  "...",
  "Preceptor", "Linda Whitfield",    "2026" , "No",           "55",   "Female",
  "..."      , "...",                "..."  , "...",          "...",  "..."
)

pop_gt_html_c <- gt::gt(pop_tibble_c, id = "population_tbl_c") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time",  id = "unit_span",
                  columns = c(`Adult`, `Year`)) |>
  gt::tab_spanner(label = "Outcome",    id = "outcome_span",
                  columns = c(`Ever Smoked`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Age`, `Sex`)) |>
  gt::cols_align(align = "left",  columns = c(`Source`, `Adult`, `Ever Smoked`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Age`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "outcome_span", "covariates_span"))
  ) |>
  gt::opt_css(
    css = "
      #population_tbl_c .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html_c,
  "</div>\n",
  "```\n",
  sep = ""
)
```

$$P(Y = 1) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n)}}$$

with $Y \sim \text{Bernoulli}(\rho)$, where $\rho$ is the probability above.

Remember that Courage creates the data generating mechanism.

### Exercise 1

Because our outcome variable is binary, add a code chunk to `analysis.qmd` that begins building the model with `logistic_reg(engine = "glm")`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-1-test}
#| echo: true
logistic_reg(engine = "glm")
```

###

**[tidymodels](https://www.tidymodels.org/)** is the most popular framework for estimating models among R users. The `logistic_reg()` function is the binary-outcome counterpart to the `linear_reg()` function we used in earlier tutorials --- different probability family, different link, same pipeline.

### Exercise 2

Extend the chunk in `analysis.qmd` to `linear_reg(engine = "lm") |> fit(smoke ~ sex, data = smokes)`. This is the simplest candidate model: `smoke` regressed on the binary covariate `sex`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-2-test}
#| echo: true
logistic_reg(engine = "glm") |>
  fit(smoke ~ sex, data = smokes)
```

###

The `tidymodels` framework requires the outcome of a `logistic_reg()` to be a *factor*, not a 0/1 integer. The `smoke` column in our `smokes` tibble is already a factor with levels `c("No", "Yes")` --- the order matters: `"Yes"` is the second level, so it is the level the model treats as the *event* whose probability the coefficients describe.

### Exercise 3

Here is the parameter table for this first model:

```{r}
#| echo: false
logistic_reg(engine = "glm") |>
  fit(smoke ~ sex, data = smokes) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 2)))
```

Write a sentence interpreting the `sexMale` estimate of 0.54.

```{r courage-3}
question_text(NULL,
  message = "Comparing two adults of different sex (one Male, one Female), the log-odds of being an ever-smoker are about 0.54 higher for Male than for Female. The 95% confidence interval (0.29 to 0.80) excludes zero, so we are reasonably confident the difference goes in the direction shown. Note the language: this is a comparison between two groups, not a claim that being male causes smoking.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The character variable `sex` becomes a 0/1 dummy called `sexMale`: 1 for Male, 0 for Female. The intercept (-0.54) is the log-odds of being an ever-smoker for the reference level (Female), and `sexMale` is the offset for switching to Male. *Log-odds* is not the scale we actually want to interpret; we will leave that translation to `marginaleffects` in the Temperance section. For now, the take-aways are: (1) the coefficient is positive, so Male adults have a higher probability of being ever-smokers, and (2) the confidence interval excludes zero, so this difference is not just sampling noise.

### Exercise 4

Fit a candidate that uses age instead, with the formula `fit(smoke ~ age, data = smokes)`. This swaps in the continuous covariate. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-4-test}
#| echo: true
logistic_reg(engine = "glm") |>
  fit(smoke ~ age, data = smokes)
```

###

When a logistic-regression covariate is continuous (like `age`), the dummy-variable machinery does not apply --- the coefficient is the change in log-odds *per one-year increase* in age. Whether that one-unit comparison is the natural one to report is a different question (a one-year age comparison is small; a forty-year comparison is big).

### Exercise 5

Here is the parameter table for the age-only model:

```{r}
#| echo: false
logistic_reg(engine = "glm") |>
  fit(smoke ~ age, data = smokes) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

Write a sentence interpreting the `age` estimate.

```{r courage-5}
question_text(NULL,
  message = "Comparing two adults differing by one year of age, the log-odds of being an ever-smoker are about 0.007 higher for the older adult. The 95% confidence interval (-0.001 to 0.014) just barely crosses zero, so we cannot quite tell with this much data whether the relationship between age and ever-smoker status is positive, negative, or essentially flat.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A coefficient whose CI just barely crosses zero is the harder pedagogical case. *"Barely"* is not the same as *"clearly insignificant."* The reasonable read is that age has a small positive association with ever-smoker status that we cannot quite pin down at this sample size. We will keep `age` in the final model anyway --- not because its CI clears zero, but because the question explicitly asks about a 30-year-old vs. a 70-year-old, so we need a slope on age to answer it.

### Exercise 6

Fit the final model in `analysis.qmd` with the formula `fit(smoke ~ age + sex, data = smokes)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-6-test}
#| echo: true
logistic_reg(engine = "glm") |>
  fit(smoke ~ age + sex, data = smokes)
```

###

This is the model the rest of the tutorial uses. Now that the model has more than one covariate, every interpretation needs an *adjustment clause*: *"comparing two adults of different sex but the same age,"* not just *"comparing two adults of different sex."*

### Exercise 7

Here is the parameter table for the final model:

```{r}
#| echo: false
logistic_reg(engine = "glm") |>
  fit(smoke ~ age + sex, data = smokes) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

Write a sentence interpreting the `sexMale` estimate of 0.545 in the final model.

```{r courage-7}
question_text(NULL,
  message = "Comparing two adults of the same age but different sex, the log-odds of being an ever-smoker are about 0.55 higher for Male than for Female (95% CI: 0.29 to 0.80, adjusting for age). The estimate is essentially unchanged from the sex-only model, which makes sense: in this dataset, age and sex are roughly uncorrelated, so adjusting for age does not move the sex coefficient.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The more variables we add, the harder it gets to interpret any one coefficient --- but interpretation also matters less, because what we ultimately want is not coefficients but predictions on the outcome scale. The *adjustment clause* (*"adjusting for age"*) is mandatory whenever the model has more than one covariate. Skipping it is the single most common student mistake on multi-covariate interpretation.

### Exercise 8

We've settled on the final fit. In `analysis.qmd`, add a new code chunk that assigns the result of `logistic_reg(engine = "glm") |> fit(smoke ~ age + sex, data = smokes)` to `fit_smokes`, then on the next line prints `fit_smokes`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-8-test}
#| echo: true
fit_smokes <- logistic_reg(engine = "glm") |>
  fit(smoke ~ age + sex, data = smokes)
fit_smokes
```

###

Logistic-regression coefficients live on the log-odds scale, which is not a scale anyone reads directly. A coefficient of +0.05 on `age` means that the *log* of the *odds* of being an ever-smoker goes up by 0.05 for each additional year of age --- a statement most people cannot intuit. The real interpretation happens in Temperance, where we use `marginaleffects::predictions()` to turn the log-odds back into probabilities.

### Exercise 9

Re-fitting the model on every render is wasteful. `fit_smokes` earns its own cached create chunk. The next three exercises set that up: clean the chunk, turn caching on, add the cache directory to `.gitignore`.

First, clean up. In your chunk from Exercise 8, delete the `fit_smokes` line at the bottom, leaving just the `fit_smokes <- logistic_reg(...) |> fit(...)` assignment. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

<pre><code>fit_smokes <- logistic_reg(engine = "glm") |>
  fit(smoke ~ age + sex, data = smokes)
</code></pre>

###

A chunk ready to cache should contain only the artifact it builds.

### Exercise 10

Add `#| cache: true` as the first line inside the chunk. Render the QMD. Rendering creates an `analysis_cache/` directory next to your `analysis.qmd`, which holds the saved fit so Quarto can reload it on subsequent renders rather than re-fitting every time. To confirm the directory is there, run `ls` in the **bash Terminal**. CP/CR.

```{r courage-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<pre><code>$ ls
analysis.qmd  analysis_cache  analysis_files  analysis.html
</code></pre>

###

The savings here are modest; the same pattern matters more when the cached fit takes seconds rather than milliseconds.

### Exercise 11

The Source Control change count just jumped, because `analysis_cache/` and everything inside it now look like untracked work to git. Cached files should never go to GitHub --- they're machine-specific and regenerated on every render.

Add `analysis_cache` to your `.gitignore` on its own line. In the R Terminal, run `show_file(".gitignore")`. CP/CR.

```{r courage-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<pre><code>analysis_files
analysis_cache
</code></pre>

###

The Source Control change count should drop back to whatever it was before the render.

### Exercise 12

To check whether our model is reasonable, we compare our actual outcome distribution to draws simulated from the fitted model. In `analysis.qmd`, add a chunk that loads **[easystats](https://easystats.github.io/easystats/)** (`library(easystats)`) and runs `check_predictions(extract_fit_engine(fit_smokes))`. The `extract_fit_engine()` step is necessary because `easystats::check_predictions()` expects a `glm` object, not the `parsnip` wrapper that `tidymodels` produces. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-12-test, message = FALSE}
#| echo: true
library(easystats)
check_predictions(extract_fit_engine(fit_smokes))
```

###

The purpose of `check_predictions()` is to compare your actual data with data simulated from your fitted model. If your DGM is reasonable, the simulated and actual distributions should look similar. For a binary outcome the comparison is necessarily simple: is the share of `Yes` close to what the model would generate? Here, yes --- and that is the floor, not the ceiling, of what model-checking does.

### Exercise 13

With the parameters estimated, we can write the fitted model as a concrete formula:

$$\widehat{P(\text{Ever Smoked} = \text{Yes})} = \frac{1}{1 + e^{-(-0.875 + 0.007 \cdot \text{Age} + 0.545 \cdot \text{sexMale})}}$$

`sexMale` is a dummy variable: 1 for Male, 0 for Female. The intercept (-0.875) is the log-odds of being an ever-smoker for a 0-year-old Female (an extrapolation: the data only includes adults aged 20--80, so the intercept is mathematically defined but not a sensible prediction on its own). The coefficient on `Age` (0.007) is the per-year change in log-odds; the coefficient on `sexMale` (0.545) is the offset between Female and Male, adjusting for age.

**This is our data generating mechanism.**

###

A data generating mechanism is just a formula, something we can write down and implement with computer code. The non-linearity in logistic regression --- the $1/(1 + e^{-\eta})$ wrapper --- is what makes log-odds coefficients hard to read on the outcome scale. Translating coefficients back to probabilities is what `marginaleffects` does in Temperance.

### Exercise 14

Add `*_cache` to your `.gitignore`. Cached objects don't belong on GitHub.

In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r courage-14}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files\n    *_cache",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Commit and push.

### Exercise 15

In `analysis.qmd`, add a chunk that runs `tidy(fit_smokes, conf.int = TRUE)`. This returns 95% intervals for all the parameters in our model. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-15}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-15-test}
#| echo: true
tidy(fit_smokes, conf.int = TRUE)
```

###

`tidy()` is part of **[broom](https://broom.tidymodels.org/)**, used to summarize information from a wide variety of models. The `estimate`, `conf.low`, and `conf.high` columns are all on the log-odds scale --- the same scale as the model's link function.

### Exercise 16

Create a new code chunk in your QMD. Make a nice-looking table from the tibble returned by `tidy()`. Insert your table code, Render the QMD, and in the R Terminal run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r courage-16}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    tidy(fit_smokes, conf.int = TRUE) |>\n      select(term, estimate, conf.low, conf.high) |>\n      knitr::kable(\n        caption = \"Logistic-regression estimates (log-odds scale) and 95% CIs.  Source: NHANES 2009-2012 via primer.tutorials::smokes\",\n        digits = 3\n      )\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

A nice table includes a title, a caption with the data source, and reasonable rounding. For logistic-regression coefficients on the log-odds scale, three decimal places usually keeps the small slopes (like the 0.007 on `age`) visible while not over-precising the bigger ones.

### Exercise 17

Add a sentence to your project summary explaining the model structure.

```{r courage-17}
question_text(NULL,
  message = "We model whether an adult is an ever-smoker as a Bernoulli variable which is a logistic function of age and sex.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add your sentence to the summary paragraph. Render the QMD, then commit/push.


## Temperance
###

Courage provides the data generating mechanism. We can express the DGM in four ways:

**In words:** We model whether an adult is an ever-smoker as a Bernoulli variable which is a logistic function of age and sex.

**In R code:**

```r
fit_smokes <- logistic_reg(engine = "glm") |>
  fit(smoke ~ age + sex, data = smokes)
```

**In a parameter table:**

```{r}
#| echo: false
fit_smokes |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

**As a mathematical formula:**

$$\widehat{P(\text{Ever Smoked} = \text{Yes})} = \frac{1}{1 + e^{-(-0.875 + 0.007 \cdot \text{Age} + 0.545 \cdot \text{sexMale})}}$$

Remember that Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.

### Exercise 1

In the end, we don't really care about parameters, much less how to interpret them. Parameters are *imaginary*, like unicorns. We care about answers to our questions, on the outcome scale.

Add `library(marginaleffects)` to the setup chunk in `analysis.qmd` and render.

CP/CR.

```{r temperance-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

<pre><code>
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(marginaleffects)
</code></pre>

###

The tool for translating log-odds coefficients into outcome-scale probabilities is the **[marginaleffects](https://marginaleffects.com/)** package, with companion book *[Model to Meaning](https://marginaleffects.com/)* by Vincent Arel-Bundock.

### Exercise 2

What is the specific question we are trying to answer?

```{r temperance-2}
question_text(NULL,
  message = "What is the difference in the probability of being an ever-smoker between a 30-year-old woman and a 70-year-old man?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

A specific question must specify a number to compute. *"How does ever-smoker status vary with age and sex?"* is too vague --- it doesn't tell us what to subtract from what. *"The difference in probability between a 30-year-old woman and a 70-year-old man"* is one specific number we can produce.

### Exercise 3

In `analysis.qmd`, add a chunk that runs

```
predictions(extract_fit_engine(fit_smokes),
            newdata = data.frame(age = c(30, 70, 30, 70),
                                 sex = c("Female", "Female", "Male", "Male")))
```

The `extract_fit_engine()` step is required because `marginaleffects` needs the underlying `glm` object, not the `parsnip` wrapper. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r temperance-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r temperance-3-test}
#| echo: true
predictions(extract_fit_engine(fit_smokes),
            newdata = data.frame(age = c(30, 70, 30, 70),
                                 sex = c("Female", "Female", "Male", "Male")))
```

###

`predictions()` returns the model's predicted probability for each combination of covariate values. Reading the four rows above: a 30-year-old woman has a predicted probability of 0.34 of being an ever-smoker; a 70-year-old man has a predicted probability of 0.54. The difference is 0.20, or 20 percentage points --- the answer to our specific question.

### Exercise 4

In `analysis.qmd`, add a chunk that runs `plot_predictions(extract_fit_engine(fit_smokes), condition = c("age", "sex"))`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r temperance-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r temperance-4-test}
#| echo: true
plot_predictions(extract_fit_engine(fit_smokes), condition = c("age", "sex"))
```

###

Two-variable conditioning (`condition = c("age", "sex")`) shows the effect surface of the model: how predicted probability moves jointly across both covariates. The shallow slope on age and the larger gap between sexes are the two stories the model is telling.

### Exercise 5

Predictions tell us the level of the outcome at specific covariate values. *Comparisons* tell us how the outcome shifts when we change one covariate while holding others fixed. Most of the questions we ultimately ask --- *"how big is the sex difference?"*, *"how much does a 40-year age gap move things?"* --- are comparisons, not predictions.

In `analysis.qmd`, add a chunk that runs `avg_comparisons(extract_fit_engine(fit_smokes), variables = "sex")`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r temperance-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r temperance-5-test}
#| echo: true
avg_comparisons(extract_fit_engine(fit_smokes), variables = "sex")
```

###

The estimate of 0.133 is the average difference in predicted probability of being an ever-smoker between Male and Female adults, on the *response* scale (i.e., probabilities, not log-odds), averaged across the actual age distribution in our sample. About 13 percentage points. *Percentage points*, not *percent*: a difference of 13 percentage points (from, say, 37% to 50%) is a 35% relative increase, but those are different numbers and they answer different questions.

### Exercise 6

Modify the previous `plot_predictions` chunk in `analysis.qmd` to add `draw = FALSE` to the call. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r temperance-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r temperance-6-test}
#| echo: true
plot_predictions(extract_fit_engine(fit_smokes), condition = c("age", "sex"), draw = FALSE)
```

###

`plot_predictions()` returns a ggplot object, so you can keep adding ggplot layers. But it is often easier to start from the underlying tibble (`draw = FALSE`) and build the plot from scratch --- which gives you complete control over titles, labels, and color choices.

### Exercise 7

Create a beautiful plot starting from the output of `plot_predictions(..., draw = FALSE)`. Do this in your QMD. Title: key variables. Subtitle: takeaway. Caption: data source. Y-axis: percentages. Paste the plot code below.

```{r temperance-7}
question_text(NULL,
  message = "    plot_predictions(extract_fit_engine(fit_smokes),\n                     condition = c(\"age\", \"sex\"),\n                     draw = FALSE) |>\n      ggplot(aes(x = age, y = estimate, color = sex, fill = sex)) +\n      geom_ribbon(aes(ymin = conf.low, ymax = conf.high),\n                  alpha = 0.2, color = NA) +\n      geom_line(linewidth = 1.1) +\n      scale_y_continuous(labels = scales::label_percent(),\n                         limits = c(0, 0.7)) +\n      labs(\n        title    = \"Probability of Being an Ever-Smoker, by Age and Sex\",\n        subtitle = \"Men of any age are about 13 percentage points more likely than women to be ever-smokers\",\n        x        = \"Age\",\n        y        = NULL,\n        color    = \"Sex\",\n        fill     = \"Sex\",\n        caption  = \"Source: NHANES 2009-2012 via primer.tutorials::smokes\"\n      ) +\n      theme_minimal()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 18)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(extract_fit_engine(fit_smokes),
                 condition = c("age", "sex"),
                 draw = FALSE) |>
  ggplot(aes(x = age, y = estimate, color = sex, fill = sex)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.2, color = NA) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 0.7)) +
  labs(
    title    = "Probability of Being an Ever-Smoker, by Age and Sex",
    subtitle = "Men of any age are about 13 percentage points more likely than women to be ever-smokers",
    x        = "Age",
    y        = NULL,
    color    = "Sex",
    fill     = "Sex",
    caption  = "Source: NHANES 2009-2012 via primer.tutorials::smokes"
  ) +
  theme_minimal()
```

### Exercise 8

Finalize the new graphics chunk in your QMD. Render the QMD. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r temperance-8}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    plot_predictions(extract_fit_engine(fit_smokes),\n                     condition = c(\"age\", \"sex\"),\n                     draw = FALSE) |>\n      ggplot(aes(x = age, y = estimate, color = sex, fill = sex)) +\n      geom_ribbon(aes(ymin = conf.low, ymax = conf.high),\n                  alpha = 0.2, color = NA) +\n      geom_line(linewidth = 1.1) +\n      scale_y_continuous(labels = scales::label_percent(),\n                         limits = c(0, 0.7)) +\n      labs(\n        title    = \"Probability of Being an Ever-Smoker, by Age and Sex\",\n        subtitle = \"Men of any age are about 13 percentage points more likely than women to be ever-smokers\",\n        x = \"Age\", y = NULL, color = \"Sex\", fill = \"Sex\",\n        caption = \"Source: NHANES 2009-2012 via primer.tutorials::smokes\"\n      ) +\n      theme_minimal()\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 18)
```

###

The map is not the territory. A clean graphic tells a story, but the story always rests on assumptions that are at best approximately true.

### Exercise 9

Write the last sentence of your summary paragraph. It describes at least one Quantity of Interest and a measure of uncertainty.

```{r temperance-9}
question_text(NULL,
  message = "We estimate that a 70-year-old man has a probability of about 54% of being an ever-smoker, compared to about 34% for a 30-year-old woman --- a gap of roughly 20 percentage points, with the underlying sex-by-age model showing a much larger sex gap (~13 pp, CI roughly 7 to 19 pp) than age gap (~6 pp over a 40-year span, CI crossing zero).",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a final sentence to your summary paragraph, but don't copy/paste exactly. Render the QMD.

### Exercise 10

Write a few sentences explaining why the estimate might be wrong. Suggest alternative estimates if warranted.

```{r temperance-10}
question_text(NULL,
  message = "Our 1,000-row sample is drawn from NHANES 2009-2012, which oversamples certain demographic groups and which is a decade behind the year we want to predict for. Smoking has been falling unevenly across age cohorts since the 2010s, so the slope on age may understate how strongly *current* younger adults differ from older adults. The sex gap is more stable in the U.S. literature than the age slope, but it is shrinking, so 13 percentage points may be an overestimate for 2026. We should widen the confidence interval substantially: a defensible report would say the male-female gap is somewhere between 5 and 20 percentage points, and the 30-vs-70 difference is somewhere between -5 and +30 percentage points.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

Always go back to your Preceptor Table. Even a perfectly estimated statistical model is rarely as useful as we'd like.

### Exercise 11

Rearrange the material in your QMD so the order is graphic, then summary paragraph. The chunk that creates `fit_smokes` must occur before the chunk that creates the graphic. Render the QMD. In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r temperance-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

<pre><code>
---
title: "Smokes"
author: "&lt;your name&gt;"
execute:
  echo: false
---

#| message: false
library(tidyverse)
library(tidymodels)
library(primer.tutorials)
library(broom)
library(marginaleffects)

#| cache: true
fit_smokes <- logistic_reg(engine = "glm") |>
  fit(smoke ~ age + sex, data = smokes)

plot_predictions(extract_fit_engine(fit_smokes),
                 condition = c("age", "sex"),
                 draw = FALSE) |>
  ggplot(aes(x = age, y = estimate, color = sex, fill = sex)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.2, color = NA) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 0.7)) +
  labs(
    title    = "Probability of Being an Ever-Smoker, by Age and Sex",
    subtitle = "Men of any age are about 13 percentage points more likely than women to be ever-smokers",
    x = "Age", y = NULL, color = "Sex", fill = "Sex",
    caption = "Source: NHANES 2009-2012 via primer.tutorials::smokes"
  ) +
  theme_minimal()

Lifetime smoking is one of the strongest predictors of long-term health outcomes. Using a 1,000-adult sample from the 2009-2012 NHANES waves, we estimate the probability of being an ever-smoker as a function of age and sex. One concern is that smoking rates have shifted unevenly across cohorts since the data was collected, so the slope on age may misstate the relationship for 2026 adults. We model ever-smoker status as a Bernoulli variable which is a logistic function of age and sex. We estimate that a 70-year-old man has about a 54% chance of being an ever-smoker compared to about 34% for a 30-year-old woman, with the male-female gap (~13 pp) much more precisely estimated than the age gap.
</code></pre>

###

This is the version of your QMD your teacher is most likely to look at closely.

### Exercise 12

Publish your rendered QMD to GitHub Pages. In the bash Terminal (not the R Terminal!), run:

```
quarto publish gh-pages analysis.qmd
```

Copy/paste the resulting URL below.

```{r temperance-12}
question_text(NULL,
  message = "https://<your-github-username>.github.io/smokes/analysis.html",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

Commit/push everything.

### Exercise 13

Copy/paste the URL to your GitHub repo.

```{r temperance-13}
question_text(NULL,
  message = "https://github.com/<your-github-username>/smokes",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God.

The world confronts us. Make decisions we must.


## Summary
###

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(extract_fit_engine(fit_smokes),
                 condition = c("age", "sex"),
                 draw = FALSE) |>
  ggplot(aes(x = age, y = estimate, color = sex, fill = sex)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.2, color = NA) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 0.7)) +
  labs(
    title    = "Probability of Being an Ever-Smoker, by Age and Sex",
    subtitle = "Men of any age are about 13 percentage points more likely than women to be ever-smokers",
    x        = "Age",
    y        = NULL,
    color    = "Sex",
    fill     = "Sex",
    caption  = "Source: NHANES 2009-2012 via primer.tutorials::smokes"
  ) +
  theme_minimal()
```

Lifetime smoking is one of the strongest predictors of long-term health outcomes, and a state health department designing an anti-smoking campaign needs to know which adults are most likely to be smokers. Using a 1,000-adult sample from the 2009--2012 NHANES waves, we estimate the probability of being an ever-smoker as a function of age and sex. One concern is that smoking rates in the United States have shifted unevenly across age cohorts since the data was collected, so the slope on age may misstate the relationship between age and smoking for 2026 adults. We model ever-smoker status as a Bernoulli variable which is a logistic function of age and sex. We estimate that a 70-year-old man has about a 54% probability of being an ever-smoker, compared to about 34% for a 30-year-old woman --- a gap of roughly 20 percentage points, with the male-female component of the gap (~13 percentage points, 95% CI of about 7 to 19 percentage points) much more precisely estimated than the age component.

This finding would inform a campaign-targeting decision: where to put cessation outreach, how to scale advertising buys by demographic segment, and how to weigh the relative importance of age and sex when designing creative material. It does not address every question a public-health analyst would face --- it says nothing about *current* smokers (who are a strict subset of ever-smokers and the actual cessation target), about race or income, or about how outreach should differ within sex by socioeconomic status. But it gives a defensible starting point.

The world is always more uncertain than our models would have us believe.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
