---
title: Rubin Causal Model
author: David Kane and Ivy Spratt
tutorial:
  id: rubin-causal-model
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #03 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60,
        tutorial.storage = "local")

hatch <- function(x) {
  paste0(
    '<span style="background-image: repeating-linear-gradient(',
    '45deg, rgba(0,0,0,0.12) 0 1px, transparent 1px 6px);',
    ' padding: 6px 10px; margin: -6px -10px; display: block;">',
    x, '</span>'
  )
}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). It focuses on the [Rubin Causal Model](https://ppbds.github.io/primer/rubin-causal-model.html) --- the framework most modern statistics uses to think about cause and effect through the lens of *potential outcomes*.

The world confronts us. Make decisions we must.

Imagine that you are a data scientist at a mid-size technology company. The Head of HR has a big-picture goal --- improve employees' long-term financial wellbeing --- and many decisions to make in service of it: which benefits to offer, which to cut, how aggressively to nudge employees toward higher retirement-account contributions, and whether to fund a new one-on-one financial-counseling program where employees can meet with a certified financial planner. The HR Director trusts you to find the evidence that helps her decide. This tutorial walks through the four Cardinal Virtues by applying them to two related questions about the same employees: a causal one (what is the effect of the financial-counseling program on 401(k) contribution rates?) and a paired predictive one (what will be the average 401(k) contribution rate among next year's new hires?). The exercises are conceptual --- no R coding here; the goal is to practice the Rubin framework before the example tutorials that follow. There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

### Exercise 2

In your own words, describe the Rubin Causal Model.

```{r introduction-2}
question_text(NULL,
  message = "The Rubin Causal Model is an approach to the statistical analysis of cause and effect based on the framework of potential outcomes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Rubin's framework is the dominant one in modern statistics. Its vocabulary --- *potential outcomes*, *treatment*, *unconfoundedness*, *fundamental problem of causal inference* --- is what the rest of this tutorial (and the rest of the *Primer*) uses to talk about causation.

### Exercise 3

Define a potential outcome.

```{r introduction-3}
question_text(NULL,
  message = "A potential outcome is the outcome for an individual under a specified treatment. In a causal model there are at least two potential outcomes for each unit: the outcome under treatment and the outcome under control.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

For our problem, each employee has two potential 401(k) contribution rates: the rate they *would* contribute if the company offered them the financial-counseling program, and the rate they *would* contribute if it did not. Both potential outcomes are well-defined attributes of the employee; what makes the situation hard is that we only ever see one.

### Exercise 4

Define a causal effect.

```{r introduction-4}
question_text(NULL,
  message = "A causal effect is the difference between two potential outcomes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

According to the Rubin Causal Model, there must be two (or more) potential outcomes for any discussion of causation to make sense. This is simplest when the treatment has two values, generating two potential outcomes per unit. With a continuous treatment (e.g., *how many hours of counseling?*), there is a potential outcome for every possible value of the treatment variable.

### Exercise 5

What is the fundamental problem of causal inference?

```{r introduction-5}
question_text(NULL,
  message = "The fundamental problem of causal inference is that we can only observe one potential outcome.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

For each employee, we see the contribution rate they actually had under whichever treatment they actually got --- counseling or no counseling. The contribution rate they would have had under the other arm is the *counterfactual*, and we never see it. Causal inference is therefore a missing-data problem: we are estimating an average of unit-level effects we cannot observe for any single unit.

### Exercise 6

In your own words, describe the difference between predictive and causal models.

```{r introduction-6}
question_text(NULL,
  message = "Predictive models have only one outcome column. Causal models have more than one (potential) outcome column because we need more than one potential outcome in order to estimate a causal effect.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The distinction is *not* a property of the data; it is a commitment made by the analyst. The same dataset --- and the same R code --- can fit a predictive model or a causal model. What differs is whether the analyst is willing to treat one of the covariates as a *treatment* (a variable that can in principle be manipulated). Yes: causal. No: predictive.


## Wisdom
###

> *Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.*

Wisdom commits us to a question precise enough to be answered. The Preceptor Table makes the commitment concrete: it is the smallest table that, if every cell were filled with the truth, would make the question's answer easy to read off.

> *The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* --- John W. Tukey

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
  message = "Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

### Exercise 3

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-3}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

In this tutorial we work two questions in parallel. The primary is the causal one (it sits at the heart of the Rubin Causal Model); the paired is the predictive one.

### The primary question (causal)

The primary question is:

> What is the average causal effect of the financial-counseling program on an employee's 401(k) contribution rate?

This is a causal question. Each employee has two potential outcomes --- the contribution rate they would have under the counseling program, and the rate they would have without it --- and the causal effect for that employee is the difference. The average causal effect across all employees is the number the HR Director wants to weigh against the program's cost.

### Exercise 4

What are the units for the primary question?

```{r wisdom-4}
question_text(NULL,
  message = "Each row is one current employee, observed during the calendar year in which the financial-counseling program would run. The company has about 2,400 benefits-eligible employees.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

### Exercise 5

What is the outcome variable for the primary question?

```{r wisdom-5}
question_text(NULL,
  message = "The outcome is the employee's annual 401(k) contribution as a percent of salary --- a continuous variable on a 0-to-100 scale, in practice clustered between 0 and 15.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

### Exercise 6

What is the treatment in the primary causal question?

```{r wisdom-6}
question_text(NULL,
  message = "The treatment is whether the employee was enrolled in the company's financial-counseling program. Two values: `Counseled` (the employee attended at least one one-on-one session with a certified financial planner during the calendar year) or `Not Counseled` (the employee did not). The HR Director can, at least in principle, manipulate this by funding the program for some employees and not others, or for some years and not others.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

### Exercise 7

For a single employee, write down the *causal effect* in terms of potential outcomes.

```{r wisdom-7}
question_text(NULL,
  message = "Take Alice, an employee at the company. Alice has two potential contribution rates: the rate she would have under counseling (say 8%) and the rate she would have without (say 5%). Her causal effect is 8 - 5 = 3 percentage points. We can never see both numbers for Alice; the experiment can only put her in one arm, so the other potential outcome is unobservable for her.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

### Exercise 8

Describe in words what the Preceptor Table for the primary causal question would look like.

```{r wisdom-8}
question_text(NULL,
  message = "The Preceptor Table has four columns: one identifying each employee, two potential-outcome columns (`Contribution if Counseled`, `Contribution if Not Counseled`), and one treatment column (`Counseling Status`). Each row represents one employee. For each employee, only one potential outcome is observable; the other --- the counterfactual --- is the contribution rate the employee would have had under the treatment they did not receive, marked with diagonal hatching.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

```{r}
#| echo: false
#| results: asis
pt_primary_tibble <- tibble::tribble(
  ~`Employee`,         ~`Contribution if Counseled`, ~`Contribution if Not Counseled`, ~`Counseling Status`,
  "Alice Becker",      "8.0",                        "5.0",                            "Counseled",
  "Brendan Cohen",     "10.0",                       "10.0",                           "Not Counseled",
  "...",               "...",                        "...",                            "...",
  "Diego Ramirez",     "12.5",                       "9.5",                            "Counseled"
)

pt_primary_html <- gt::gt(pt_primary_tibble, id = "pt_primary") |>
  gt::tab_header(title = "Preceptor Table --- Primary (causal) question") |>
  gt::tab_spanner(label = "Unit",                id = "unit_span",      columns = c(`Employee`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",   columns = c(`Contribution if Counseled`, `Contribution if Not Counseled`)) |>
  gt::tab_spanner(label = "Treatment",          id = "treatment_span", columns = c(`Counseling Status`)) |>
  gt::cols_align(align = "left",  columns = c(`Employee`, `Counseling Status`)) |>
  gt::cols_align(align = "right", columns = c(`Contribution if Counseled`, `Contribution if Not Counseled`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_column_spanners()) |>
  gt::tab_style(style = gt::cell_text(align = "left"),
                locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span"))) |>
  gt::tab_style(style = gt::cell_text(align = "right"),
                locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  # Hatch each row's unobservable counterfactual.
  # Row 1 (Counseled): hatch `Contribution if Not Counseled`.
  # Row 2 (Not Counseled): hatch `Contribution if Counseled`.
  # Row 4 (Counseled): hatch `Contribution if Not Counseled`.
  gt::text_transform(
    locations = gt::cells_body(columns = `Contribution if Not Counseled`, rows = c(1, 4)),
    fn = function(x) hatch(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `Contribution if Counseled`, rows = 2),
    fn = function(x) hatch(x)
  ) |>
  gt::tab_footnote(
    footnote = "If all the information in this table were available, we could answer the question: What is the average causal effect of the financial-counseling program on an employee's 401(k) contribution rate?",
    locations = gt::cells_title()
  ) |>
  gt::tab_footnote(
    footnote = "Each row is one of the company's roughly 2,400 benefits-eligible employees. The cross-hatched cell in each row marks the unobservable counterfactual --- the contribution rate the employee would have had under the treatment they did not receive.",
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(
    footnote = "The treatment is whether the employee attended at least one one-on-one session with a certified financial planner during the calendar year.",
    locations = gt::cells_column_spanners(spanners = "treatment_span")
  ) |>
  gt::opt_css(css = "#pt_primary .gt_footnote { max-width: 1px; word-break: break-word; }") |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    pt_primary_html, "</div>\n```\n", sep = "")
```

Notice the Preceptor Table shows the *unobservable truth* --- the actual potential outcomes for each employee under both treatment arms. In reality we will only ever see one cell per row (the un-hatched one). The Rubin framework's whole job is to estimate the missing cells.

### The paired question (predictive)

The paired question is:

> What will be the average 401(k) contribution rate of new hires next year?

This is a predictive question. We want one specific number: an estimate of the average contribution rate for the next cohort of new hires, with its associated uncertainty.

### Exercise 9

How many potential outcomes does each new hire have in the paired predictive question?

```{r wisdom-9}
question_text(NULL,
  message = "One. The paired question is predictive: there is no treatment to manipulate, so each new hire has a single observable contribution rate (whatever rate they end up choosing). Predictive models have no potential outcomes --- just one outcome column.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

```{r}
#| echo: false
#| results: asis
pt_paired_tibble <- tibble::tribble(
  ~`New Hire`,           ~`Contribution Rate (%)`,
  "Erica Foster",        "6.0",
  "Frank Gutierrez",     "10.0",
  "...",                 "...",
  "Hannah Park",         "4.5"
)

pt_paired_html <- gt::gt(pt_paired_tibble, id = "pt_paired") |>
  gt::tab_header(title = "Preceptor Table --- Paired (predictive) question") |>
  gt::tab_spanner(label = "Unit",    id = "unit_span",    columns = c(`New Hire`)) |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = c(`Contribution Rate (%)`)) |>
  gt::cols_align(align = "left",  columns = c(`New Hire`)) |>
  gt::cols_align(align = "right", columns = c(`Contribution Rate (%)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_column_spanners()) |>
  gt::tab_style(style = gt::cell_text(align = "left"),
                locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_style(style = gt::cell_text(align = "right"),
                locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(
    footnote = "If all the information in this table were available, we could answer the question: What will be the average 401(k) contribution rate of new hires next year?",
    locations = gt::cells_title()
  ) |>
  gt::tab_footnote(
    footnote = "Each row is one new hire expected next year. The company adds about 300 net new hires annually.",
    locations = gt::cells_column_spanners(spanners = "unit_span")
  ) |>
  gt::tab_footnote(
    footnote = "Annual 401(k) contribution as a percent of salary.",
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::opt_css(css = "#pt_paired .gt_footnote { max-width: 1px; word-break: break-word; }") |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    pt_paired_html, "</div>\n```\n", sep = "")
```

### Exercise 10

The two Preceptor Tables are nearly identical in structure. What differs?

```{r wisdom-10}
question_text(NULL,
  message = "Two structural differences. First, the causal table has two potential-outcome columns (`Contribution if Counseled`, `Contribution if Not Counseled`) with the unobservable counterfactual hatched; the predictive table has one outcome column (`Contribution Rate (%)`) with nothing hatched. Second, the causal table has a `Treatment` column under its own spanner; the predictive table has no treatment column. Everything else --- the role of the unit identifier, the place of covariates if we had any --- is the same. That single difference in outcome-and-treatment bookkeeping is the entire mechanical difference between predictive and causal models.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

Same data could fit either model. What makes one *causal* and the other *predictive* is the analyst's commitment about what the covariates are, defended in Justice.


## Justice
###

> *Justice reviews the Population Table and selects the formula for the data generating mechanism.*

Justice is the section of the analysis where you (or your critics) raise concerns about whether your model will work. The Population Table is where those concerns become concrete: it bridges between the data we actually have and the Preceptor Table we wish we had.

### Exercise 1

In your own words, describe the components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
  message = "Justice reviews the Population Table and selects the formula for the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The four classic Justice concerns are *validity*, *stability*, *representativeness*, and *unconfoundedness*. The first three apply to both predictive and causal models; the fourth applies only to causal models --- which means it is central to the primary question in this tutorial.

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Validity is about the *columns*. Two columns can have the same name and still measure different things.

### Exercise 3

Suppose our data is from a similar tech company that ran a comparable counseling program three years ago. Give one reason why the assumption of validity might not hold for the `contribution_rate` column.

```{r justice-3}
question_text(NULL,
  message = "Two companies' 401(k) plans may differ in their default contribution rate (the rate an employee is automatically enrolled at if they take no action). If the comparison company auto-enrolls at 5% and our company auto-enrolls at 3%, the same employee would show different `contribution_rate` values across the two companies even with no behavioral difference. The column called `contribution_rate` therefore does not measure the same underlying construct in the two firms.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The Population Table is the bridge from the data to the Preceptor Table. Each row is one unit at one moment in time. The table includes the rows we have (the Data block), the rows we want (the Preceptor block), and the much larger universe of unit/time combinations both blocks are samples from.

### Exercise 5

What unit/time combination defines each row in the Population Table for our primary causal question?

```{r justice-5}
question_text(NULL,
  message = "Each row is one employee at one calendar year. Data rows are employees at a similar tech company from a previous year. Preceptor rows are our company's roughly 2,400 employees in the year the counseling program would run.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The Population Table for the **primary (causal)** question shows two potential outcomes, with the unobserved counterfactual marked `...` in Data rows and hatched in Preceptor rows.

```{r}
#| echo: false
#| results: asis
pop_primary_tibble <- tibble::tribble(
  ~Source,     ~`Employee`,         ~`Year`, ~`Contribution if Counseled`, ~`Contribution if Not Counseled`, ~`Counseling Status`,
  "...",       "...",               "...",   "...",                        "...",                            "...",
  "Data",      "Sarah Kim",         "2022",  "9.0",                        "...",                            "Counseled",
  "Data",      "Michael Lee",       "2022",  "...",                        "4.5",                            "Not Counseled",
  "Data",      "...",               "...",   "...",                        "...",                            "...",
  "Data",      "Priya Patel",       "2022",  "11.0",                       "...",                            "Counseled",
  "...",       "...",               "...",   "...",                        "...",                            "...",
  "Preceptor", "Alice Becker",      "2026",  "8.0",                        "5.0",                            "Counseled",
  "Preceptor", "Brendan Cohen",     "2026",  "10.0",                       "10.0",                           "Not Counseled",
  "Preceptor", "...",               "...",   "...",                        "...",                            "...",
  "Preceptor", "Diego Ramirez",     "2026",  "12.5",                       "9.5",                            "Counseled",
  "...",       "...",               "...",   "...",                        "...",                            "..."
)

pop_primary_html <- gt::gt(pop_primary_tibble, id = "pop_primary") |>
  gt::tab_header(title = "Population Table --- Primary causal question") |>
  gt::tab_spanner(label = "Unit/Time",          id = "unit_span",      columns = c(`Employee`, `Year`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",   columns = c(`Contribution if Counseled`, `Contribution if Not Counseled`)) |>
  gt::tab_spanner(label = "Treatment",          id = "treatment_span", columns = c(`Counseling Status`)) |>
  gt::cols_align(align = "left",  columns = c(`Source`, `Employee`, `Counseling Status`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Contribution if Counseled`, `Contribution if Not Counseled`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_column_spanners()) |>
  gt::tab_style(style = gt::cell_text(align = "left"),
                locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span"))) |>
  gt::tab_style(style = gt::cell_text(align = "right"),
                locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  # Hatch the Preceptor rows' unobservable counterfactuals (rows 8, 9, 11).
  gt::text_transform(
    locations = gt::cells_body(columns = `Contribution if Not Counseled`, rows = c(8, 11)),
    fn = function(x) hatch(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `Contribution if Counseled`, rows = 9),
    fn = function(x) hatch(x)
  ) |>
  gt::tab_footnote(
    footnote = "Data rows are 2022 employees at a comparable tech company; Preceptor rows are our company's 2026 employees. Two notations mark unobservable potential outcomes: in Data rows, '...' means we don't have the value; in Preceptor rows, the cross-hatched cell means the truth exists but no empirical process could ever reveal it.",
    locations = gt::cells_title()
  ) |>
  gt::opt_css(css = "#pop_primary .gt_footnote { max-width: 1px; word-break: break-word; }") |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    pop_primary_html, "</div>\n```\n", sep = "")
```

### Exercise 6

In your own words, define the assumption of "stability."

```{r justice-6}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Stability is best understood as a statement about *parameters*, not distributions. The relationship between treatment, covariates, and contribution rate in our model is captured by a handful of coefficients. Stability says those coefficients are the same in our 2022 data as they are for our 2026 Preceptor Table.

### Exercise 7

Give one reason why the assumption of stability might not hold for our primary causal question.

```{r justice-7}
question_text(NULL,
  message = "The macro-economic environment differs between 2022 and 2026. Inflation, interest rates, and tax policy on retirement accounts have changed; the slope connecting financial counseling to contribution behavior likely depends on the broader environment. An employee in 2022 facing low inflation and high market returns may respond to a counseling session very differently from an employee in 2026 facing different conditions. The parameter governing the counseling effect may therefore have shifted, even if the marginal contribution rate has not changed much.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A change in the *level* of contribution rates between 2022 and 2026 (because of, say, broader wage gains) is a distribution shift; that on its own does not violate stability. A change in the *slope* connecting counseling to contribution rate is a parameter shift; that is what violates stability.

### Exercise 8

In your own words, define the assumption of "representativeness."

```{r justice-8}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

### Exercise 9

Give one reason why representativeness might not hold between the 2022 data and our 2026 Preceptor Table.

```{r justice-9}
question_text(NULL,
  message = "Our data comes from a comparable but different tech company. Their employees may differ from ours in age, tenure, compensation level, gender mix, and remote-vs-in-office balance --- all of which plausibly affect 401(k) contribution behavior. Our 2026 employees are not a random sample from the same broader population that the 2022 data was drawn from, so estimates fit on the 2022 data may not transport to our 2026 employees without bias.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

### Exercise 10

In your own words, define the assumption of "unconfoundedness."

```{r justice-10}
question_text(NULL,
  message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Unconfoundedness applies only to the causal question. The cleanest defense is random assignment: if the comparison company randomly assigned employees to attend financial counseling (treatment) versus no counseling (control), unconfoundedness would hold by construction.

### Exercise 11

Suppose the comparison company offered the counseling program voluntarily --- any employee could sign up. Give one reason why unconfoundedness would be hard to defend.

```{r justice-11}
question_text(NULL,
  message = "Employees who self-select into voluntary financial counseling differ from those who don't. They may be more financially literate, more risk-averse, more conscientious, more concerned about retirement, or going through a recent life event (marriage, birth, parent illness) that prompts financial planning. All of these characteristics also affect 401(k) contribution rates through paths that have nothing to do with the counseling itself. So treatment assignment (signing up for counseling) is correlated with potential outcomes (the contribution rates the employee would set in either arm), and unconfoundedness fails.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

This is the classic self-selection problem in observational causal inference. The Rubin framework makes the problem precise: voluntary takers and non-takers are not exchangeable, so the difference in their average contribution rates is not the average causal effect. Defending against this requires either randomization (best) or careful conditioning on the right pre-treatment covariates (second-best, with the analyst defending why those covariates suffice).

### Paired Population Table

The paired predictive question's Population Table has the same row structure but a single outcome column and no treatment column.

```{r}
#| echo: false
#| results: asis
pop_paired_tibble <- tibble::tribble(
  ~Source,     ~`Employee`,         ~`Year`, ~`Contribution Rate (%)`,
  "...",       "...",               "...",   "...",
  "Data",      "Sarah Kim",         "2022",  "9.0",
  "Data",      "Michael Lee",       "2022",  "4.5",
  "Data",      "...",               "...",   "...",
  "Data",      "Priya Patel",       "2022",  "11.0",
  "...",       "...",               "...",   "...",
  "Preceptor", "Erica Foster",      "2027",  "6.0",
  "Preceptor", "Frank Gutierrez",   "2027",  "10.0",
  "Preceptor", "...",               "...",   "...",
  "Preceptor", "Hannah Park",       "2027",  "4.5",
  "...",       "...",               "...",   "..."
)

pop_paired_html <- gt::gt(pop_paired_tibble, id = "pop_paired") |>
  gt::tab_header(title = "Population Table --- Paired predictive question") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",    columns = c(`Employee`, `Year`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span", columns = c(`Contribution Rate (%)`)) |>
  gt::cols_align(align = "left",  columns = c(`Source`, `Employee`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Contribution Rate (%)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_column_spanners()) |>
  gt::tab_style(style = gt::cell_text(align = "left"),
                locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_style(style = gt::cell_text(align = "right"),
                locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(
    footnote = "Data rows are employees at the comparable tech company in 2022; Preceptor rows are our company's expected new hires in 2027 (one year out, after the counseling program would have launched). No hatching --- predictive models have no potential outcomes.",
    locations = gt::cells_title()
  ) |>
  gt::opt_css(css = "#pop_paired .gt_footnote { max-width: 1px; word-break: break-word; }") |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    pop_paired_html, "</div>\n```\n", sep = "")
```

### Probability family and link function

Justice also commits us to the *formula for the data generating mechanism* --- in later tutorials we will give this its technical name (*probability family* and *link function*). The outcome `contribution_rate` is continuous (though bounded below by 0 and above by 100); we will model it as Normal with the identity link:

$$Y \sim \mathcal{N}(\mu, \sigma^2)$$

with $\mu = \beta_0 + \beta_1 X_1 + \cdots + \beta_n X_n$.


## Courage
###

> *Courage creates the data generating mechanism.*

Justice settled the structural choices --- Normal family, identity link, four assumptions to defend. Courage now picks specific covariates and fits the model. The lightweight version: *if we had data and code*, here is what Courage would do.

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
  message = "Courage creates the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

### Exercise 2

For the primary (causal) question, what would the fitted model look like in plain language?

```{r courage-2}
question_text(NULL,
  message = "We would model contribution rate as a normally distributed variable which is a linear function of the treatment (whether the employee was counseled) and zero or more pre-treatment covariates (age, tenure, salary, prior contribution rate). The fitted coefficient on the treatment variable, under unconfoundedness, would be our estimate of the average causal effect of the counseling program on contribution rate. The model would also produce uncertainty around that estimate --- a 95% confidence interval.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

### Exercise 3

For the paired predictive question, how would the fitted model differ from the causal one?

```{r courage-3}
question_text(NULL,
  message = "It wouldn't, mechanically. The same R code --- `linear_reg() |> fit(contribution_rate ~ treatment + covariates, data = x)` --- fits both. The difference is what the analyst is willing to claim. For the predictive question, we would not call the treatment-arm comparison a 'causal effect' and would not need to defend unconfoundedness; we would just call it a difference in expected contribution rate between the two groups, and report it with comparison language rather than causal language.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The same fit, the same coefficient, two different readings. This is the deepest insight of the Rubin framework: the data does not tell you what kind of model you are running --- the analyst does.


## Temperance
###

> *Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.*

Temperance is where the fitted model becomes the answer. The lightweight version: *if we had a fit*, here is how Temperance would read it.

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
  message = "Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

### Exercise 2

For the primary (causal) question, what would Temperance produce? Be careful about the language you use.

```{r temperance-2}
question_text(NULL,
  message = "An estimate of the average causal effect of the financial-counseling program on an employee's 401(k) contribution rate, with its associated uncertainty (a 95% confidence interval). Causal language --- 'causal effect,' 'effect of,' 'change caused by' --- is appropriate here only because we have committed (in Justice) to defending unconfoundedness. If unconfoundedness fails, the same number must be interpreted predictively as 'the difference in contribution rates between counseled and not-counseled employees,' not as an effect of counseling.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

### Exercise 3

For the paired predictive question, what would Temperance produce?

```{r temperance-3}
question_text(NULL,
  message = "An estimate of the expected average 401(k) contribution rate among next year's new hires, with its associated uncertainty. The point estimate alone is not an answer; the interval around it is essential. Comparison language --- 'expected contribution rate,' 'on average' --- is appropriate; causal language is not, because the predictive question makes no claim about manipulating anything.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

> *We can never know the truth.* The posterior we report is, at best, the *Preceptor's Posterior* --- the best posterior achievable under our assumptions. The reported confidence interval captures only sampling variation under those assumptions; it does not capture our uncertainty about whether the assumptions themselves are correct. The reported interval is always too narrow.


## Summary
###

This tutorial walked through the Rubin Causal Model framework by applying the four Cardinal Virtues to a paired pair of questions about an employee financial-counseling program: a causal one (the effect of the program on 401(k) contribution rates) and a predictive one (next year's average contribution rate among new hires). The exercises were conceptual, not coded. The key Rubin Causal Model takeaways:

* **Each unit has at least two potential outcomes** under a causal framing, one per treatment value.
* **A causal effect is the difference between two potential outcomes** --- a property of the unit, not of the data.
* **The fundamental problem of causal inference is that we only ever observe one potential outcome.** The other is the counterfactual, and we can never see it.
* **The predictive/causal distinction is an analyst's commitment**, not a property of the data or the model. The same fit serves both questions; what changes is the language and the assumptions you are willing to defend.

The example tutorials that follow (`05-recruits`, `06-trains`, `07-colleges`, ...) walk through the same arc using real data and `tidymodels` code. The conceptual moves rehearsed here --- writing down the Preceptor Table, distinguishing potential outcomes from observable outcomes, defending unconfoundedness or acknowledging where it fails --- are exactly the moves you will use there.

The world is always more uncertain than our models would have us believe.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
