---
title: Sampling
author: David Kane
tutorial:
  id: sampling
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #02 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(primer.data)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = "90%")
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

# Pre-computed 1,000-sample tibbles for the unbiased and biased
# demonstrations live in data/precomputed.rds. Rebuilt by
# data-raw/precompute.R; last regenerated 2026-05-31.
precomputed <- readRDS(system.file(
  "tutorials/02-sampling/data/precomputed.rds",
  package = "primer.tutorials"
))
truth          <- precomputed$truth
my_sample      <- precomputed$my_sample
sample_mean    <- precomputed$sample_mean
sample_se      <- precomputed$sample_se
samples        <- precomputed$samples
biased_sample  <- precomputed$biased_sample
biased_mean    <- precomputed$biased_mean
biased_se      <- precomputed$biased_se
biased_samples <- precomputed$biased_samples
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports the [Sampling](https://ppbds.github.io/primer/02-sampling.html) chapter of [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

The world confronts us. Make decisions we must.

The Sampling chapter starts from a simple fact: we almost never measure a whole population, so we take a *sample* and use it to estimate what we cannot see. A different sample would give a different answer, so every estimate carries uncertainty --- and the *standard error* measures it, shrinking only with the square root of the sample size (four times the data buys just half the error). A rough 95% confidence interval is the estimate give or take two standard errors. The chapter's sharpest point: all of this assumes the sample was drawn *fairly* --- when the sampling mechanism is biased, more data does not help, it only sharpens a wrong answer. This tutorial rebuilds these ideas by hand, first with fair sampling and then with biased sampling.

### Exercise 1

You should be connected to a repo named `sampling`. If you are not, create one and connect to it. Create a new Quarto document titled `"Sampling"` with yourself as the author, save it as `analysis.qmd`, render it, and open `analysis.html` with Live Server so the rendered HTML auto-refreshes on every later render. Going forward, we will just tell you to "Render" when we want you to take these steps. Create a `.gitignore` file with `analysis_files` on the first line followed by a blank line. Commit and push. Use AI however you like. In the R Terminal, run `show_file(".gitignore")`. If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

A *census* measures every unit in a population --- and it is usually too expensive, and often impossible: any population that reaches into the future can never be fully counted. So almost all of data science works instead from a *sample* and reasons back to the population. How much we can trust that reasoning is the subject of this tutorial.

### Exercise 2

In your QMD, put `library(tidyverse)` and `library(primer.data)` in a new code chunk. Render the file. Notice that the document does not look good because the code is visible and there are package-attach messages. Add `#| message: false` to suppress messages in this setup chunk. Also add the following to the YAML header to suppress all code echos:

````
execute:
  echo: false
````

Render again. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

The one number we wish we knew about the whole population --- here, the average age of every registered voter --- is the *population parameter*. We never get to see it directly; we compute an estimate from a sample and treat it as our best guess. Keeping the parameter (the unknown truth) separate from the estimate (a guess from one sample) is the central discipline of this chapter.

### Exercise 3

Add a code chunk to `analysis.qmd` that prints the `shaming` tibble (just `shaming` on its own line). Render, and copy the first few rows from the rendered HTML below.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r introduction-3-test}
#| echo: true
shaming
```

###

The `shaming` tibble holds registered Michigan voters from the 2006 get-out-the-vote experiment of Gerber, Green, and Larimer --- one row per voter, with `age`, `sex`, and past-turnout columns. For this tutorial we will treat the whole tibble as our *universe* of voters: not a sample of it, but the entire population. That lets us pretend, for teaching purposes, to know things --- like the true average age --- that we could never know in real life. Our question reaches for just one column: `age`.


## Sampling and the Standard Error
###

In real life we never see the whole universe. We see only a sample. From that sample we compute a number --- here, the average age --- and we want to know how close that number is to the truth. The tool that answers that question is the **standard error**. Over this section we will build, step by step, a working chunk in `analysis.qmd` that draws samples, computes their averages, and visualizes how often those averages land near the truth.

### Exercise 1

Start a working chunk in `analysis.qmd` that computes the mean of `age` in the `shaming` tibble (skipping missing values), stores it as `truth`, and prints it. Render. Copy the printed value from the rendered HTML below.

```{r sampling-and-the-standard-error-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 2)
```

###

```{r sampling-and-the-standard-error-1-test}
#| echo: true
truth <- mean(shaming$age, na.rm = TRUE)
truth
```

###

For the rest of the tutorial, treat this number --- about 49.79 years --- as the **truth**: the average age of every voter in our universe. In an ordinary project this number would be unknowable; we only see it here because we declared the entire `shaming` tibble to be the universe.

### Exercise 2

Modify the working chunk to draw a reproducible sample of 50 voters from `shaming`, store it as `my_sample`, compute its sample mean as `sample_mean`, and print `sample_mean`. Use `set.seed(10)`. Render. Copy the printed sample mean from the rendered HTML below.

```{r sampling-and-the-standard-error-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 2)
```

###

```{r sampling-and-the-standard-error-2-test}
#| echo: true
set.seed(10)
my_sample <- shaming |> slice_sample(n = 50)
sample_mean <- mean(my_sample$age, na.rm = TRUE)
sample_mean
```

###

Our sample mean is about 50.02 years; the truth is 49.79. Close but not identical. That gap --- between the number a sample gives us and the unknown truth --- is what every conversation about uncertainty is really about. A different sample of 50 voters would have given you a different mean. The next step is to put a number on *how much* the sample mean would jump around.

### Exercise 3

The **standard error** of a sample mean is the sample's standard deviation divided by the square root of the sample size. Modify the working chunk to compute the standard error of `my_sample`'s mean, store it as `sample_se`, and print it. Render. Copy the printed value from the rendered HTML below.

```{r sampling-and-the-standard-error-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 2)
```

###

```{r sampling-and-the-standard-error-3-test}
#| echo: true
sample_se <- sd(my_sample$age, na.rm = TRUE) / sqrt(50)
sample_se
```

###

You can do this by hand. The standard deviation of the 50 ages in our sample is about 14.32. The square root of 50 is about 7.07. Divide them: $14.32 / 7.07 \approx 2.03$. The standard error is about two years. A rough rule of thumb: roughly 95% of the time, the truth sits within two standard errors of the sample mean. In our case, two standard errors is four years, so a 95% confidence interval for the truth based on this one sample is about (46, 54). The actual truth is 49.79, which is inside that interval --- but we got lucky on this particular draw, and a different sample could miss.

The square root in the denominator is doing important work. Suppose we had drawn a sample of 1,000 voters instead of 50. The sample standard deviation would be roughly the same --- around 14 --- because that is a property of the population, not of the sample size. But the divisor changes: $\sqrt{1000} \approx 31.6$, so the standard error becomes $14 / 31.6 \approx 0.45$ years, almost a factor of five smaller than the 2.03 we got from 50 voters. The general rule: **to cut the standard error in half you need four times the sample size, because $\sqrt{4} = 2$**. Doubling the sample size only shrinks the standard error by about 30%. Confidence is expensive.

### Exercise 4

Time to draw the picture. Modify the working chunk to plot `sample_mean` as a point with a horizontal error bar covering mean ± 1.96·`sample_se`, and a red dashed vertical line at `truth`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r sampling-and-the-standard-error-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r sampling-and-the-standard-error-4-test}
#| echo: true
ggplot() +
  geom_point(aes(x = sample_mean, y = 1), size = 3) +
  geom_errorbarh(aes(xmin = sample_mean - 1.96 * sample_se,
                     xmax = sample_mean + 1.96 * sample_se,
                     y = 1),
                 height = 0.2) +
  geom_vline(xintercept = truth, color = "red",
             linetype = "dashed") +
  labs(title = "One sample of 50 voters: mean ± 1.96·SE",
       subtitle = "Red dashed line is the truth",
       x = "Age (years)",
       y = NULL) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

###

That horizontal bar is the **95% confidence interval**. It is a statement about the *method*, not about this particular sample: if we repeated the experiment many times, about 95% of the intervals we draw would contain the truth.

### Exercise 5

To see the "about 95%" claim land, we need not one sample but many. `samples` will be reused across the next three exercises, so build it once and cache it. Add a cached create chunk to `analysis.qmd` that constructs a tibble `samples` with 1,000 rows --- each row representing one 50-voter draw from `shaming` and carrying that draw's sample mean, standard error, and lower and upper bounds of the 95% confidence interval. Use `set.seed(10)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r sampling-and-the-standard-error-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

###

```{r sampling-and-the-standard-error-5-test}
#| echo: true
set.seed(10)
samples <- tibble(trial = 1:1000) |>
  mutate(
    sample      = map(trial, ~ shaming |> slice_sample(n = 50)),
    sample_mean = map_dbl(sample, ~ mean(.x$age, na.rm = TRUE)),
    sample_se   = map_dbl(sample, ~ sd(.x$age, na.rm = TRUE) / sqrt(50)),
    lower       = sample_mean - 1.96 * sample_se,
    upper       = sample_mean + 1.96 * sample_se
  ) |>
  select(-sample)
head(samples)
```

###

Each row is one of the thousand hypothetical "what if we had drawn a different 50 voters?" experiments, carrying its own mean, standard error, and 95% interval. Caching the chunk (`#| cache: true`) means Quarto reuses the saved 1,000-row tibble on subsequent renders rather than redoing the simulation each time. For 50,000 draws that's a couple of seconds saved; for the heavier model fits in later tutorials, it can be the difference between a one-second and a one-minute render.

### Exercise 6

Now plot the first 100 of those intervals as horizontal error bars stacked vertically, with `truth` marked as a vertical red line. (We use 100 rather than 1,000 so the picture stays readable.) Modify the working chunk to draw this plot. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r sampling-and-the-standard-error-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r sampling-and-the-standard-error-6-test}
#| echo: true
ggplot(samples |> slice(1:100), aes(y = trial)) +
  geom_errorbarh(aes(xmin = lower, xmax = upper),
                 height = 0.3, alpha = 0.5) +
  geom_point(aes(x = sample_mean), size = 0.6) +
  geom_vline(xintercept = truth, color = "red",
             linetype = "dashed") +
  labs(title = "100 samples of 50 voters each",
       subtitle = "Most 95% intervals contain the truth; a few miss",
       x = "Age (years)",
       y = "Sample number") +
  theme_minimal()
```

###

Most intervals cross the red line --- those contain the truth. A handful do not. The misses are not a sign the method is broken; they are exactly what we should expect when each interval is meant to catch the truth 95% of the time.

### Exercise 7

Let's count. Modify the working chunk to compute and print the fraction of the 1,000 intervals in `samples` that contain `truth` (`lower < truth & upper > truth`). Render. Copy the printed value from the rendered HTML below.

```{r sampling-and-the-standard-error-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 2)
```

###

```{r sampling-and-the-standard-error-7-test}
#| echo: true
samples |>
  summarise(coverage = mean(lower < truth & upper > truth))
```

###

About 94.5% of the intervals contained the truth --- close to the 95% the method advertises, not exactly 95%. That gap (94.5% vs. 95%) is normal sampling noise. Run the whole procedure a second time with a different seed and you would get a different number close to but not equal to 95%.

### Exercise 8

Polish the 100-sample plot for the final document: clear title, a one-line subtitle naming the takeaway, a caption pointing to `primer.data`, axis labels with units, and `theme_minimal()`. Then stage `analysis.qmd`, commit with the message `"Add sampling and standard error analysis"`, and push to GitHub. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r sampling-and-the-standard-error-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

The standard error is the bridge between "the number this one sample gave me" and "what could the truth reasonably be?" Every time you meet uncertainty in an estimate later on --- the confidence interval on a regression coefficient, the width of a predicted probability, the spread of a posterior --- the underlying machinery is the same one you just saw: draw samples, watch how the answer jumps around, summarize the jumping with a standard error. The standard error gets shorter when the sample is bigger, and longer when the data is noisier. Everything else is detail.


## Biased Sampling
###

In the previous section the sampling was *random* --- every voter in the universe had the same chance of ending up in the sample. That assumption is what made the 95% confidence intervals do what they claimed. Real survey work is almost never that clean. Door-knockers find more people at home in older neighborhoods. App-based outreach reaches younger users. Phone-only sampling skips cell-only households. In this section we will rebuild the previous machinery --- but with a sampling mechanism that systematically over-represents older voters --- and watch the coverage collapse.

### Exercise 1

Suppose a get-out-the-vote campaign's door-knock app, by quirk of its routing, makes voters over the age of 40 *three times as likely* to appear in a sample as voters 40 or younger. The voters themselves haven't changed; the universe is the same `shaming` tibble. Only the *mechanism* that pulls voters into the sample has changed.

Start a new working chunk in `analysis.qmd` that draws a biased sample of 50 voters from `shaming`, where voters over 40 are three times as likely to be picked as voters aged 40 or younger. Store the result as `biased_sample`. Use `set.seed(10)`. Print the first few rows of `biased_sample`'s `age`, `primary_06`, and `sex` columns. Render. Copy the printed rows from the rendered HTML below.

```{r biased-sampling-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r biased-sampling-1-test}
#| echo: true
set.seed(10)
biased_sample <- shaming |>
  mutate(wt = if_else(age > 40, 3, 1)) |>
  slice_sample(n = 50, weight_by = wt)
head(biased_sample |> select(age, primary_06, sex))
```

###

A weighting argument to `slice_sample()` makes each row's probability of being picked proportional to that row's weight. Rows with weight `3` are three times as likely to land in the sample as rows with weight `1`.

### Exercise 2

Before trusting any downstream results, we want to *see* that the bias is doing what we wrote it to do. Modify the working chunk to also draw an unbiased reference sample of 50 voters from `shaming` (with `set.seed(11)`) and plot the age distributions of the biased and unbiased samples side by side. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r biased-sampling-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r biased-sampling-2-test}
#| echo: true
set.seed(11)
unbiased_sample <- shaming |> slice_sample(n = 50)

bind_rows(
  biased_sample   |> mutate(origin = "Biased (age > 40 weighted 3x)"),
  unbiased_sample |> mutate(origin = "Unbiased")
) |>
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "grey50", color = "white") +
  facet_wrap(~ origin) +
  labs(title = "Age distribution: biased vs. unbiased samples of 50",
       subtitle = "The biased sample is visibly skewed toward older voters",
       x = "Age (years)", y = NULL) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

###

In the universe (`shaming`), about 76% of voters are over 40. In the unbiased sample of 50, the share over 40 will sit near 76% give-or-take sampling noise. In the *biased* sample of 50, about 98% are over 40 --- you can read it straight off the histogram. The mechanism is doing exactly what we wrote.

### Exercise 3

Now compute and print the mean and standard error of `age` in `biased_sample` the same way we did in the previous section. Modify the working chunk so the last two lines print these two numbers. Render. Copy both printed values from the rendered HTML below.

```{r biased-sampling-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```{r biased-sampling-3-test}
#| echo: true
mean(biased_sample$age, na.rm = TRUE)
sd(biased_sample$age, na.rm = TRUE) / sqrt(50)
```

###

Two things to notice. First, the biased sample's mean is 55.26 --- about 5.5 years above the truth of 49.79. The mechanism that over-sampled older voters pushed the sample mean up. Second, the standard error is **smaller** than in the unbiased section (1.51 vs. 2.03), because the biased sample is more homogeneous --- almost everyone is over 40 --- so the within-sample standard deviation is smaller. **The standard error formula has no way to know our sample is biased**. It measures the spread inside the sample we got; it cannot see what the sampling mechanism left out.

### Exercise 4

We now want a thousand biased samples, parallel to the thousand unbiased ones in `samples`. `biased_samples` will be reused across the next three exercises, so build it once and cache it. Add a cached create chunk to `analysis.qmd` that constructs a tibble `biased_samples` with 1,000 rows --- each row representing one biased 50-voter draw from `shaming` (with the same age > 40 weighting) and carrying that draw's sample mean, standard error, and lower and upper bounds of the 95% confidence interval. Use `set.seed(10)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r biased-sampling-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

###

```{r biased-sampling-4-test}
#| echo: true
set.seed(10)
shaming_wt <- shaming |> mutate(wt = if_else(age > 40, 3, 1))
biased_samples <- tibble(trial = 1:1000) |>
  mutate(
    sample      = map(trial, ~ shaming_wt |> slice_sample(n = 50, weight_by = wt)),
    sample_mean = map_dbl(sample, ~ mean(.x$age, na.rm = TRUE)),
    sample_se   = map_dbl(sample, ~ sd(.x$age, na.rm = TRUE) / sqrt(50)),
    lower       = sample_mean - 1.96 * sample_se,
    upper       = sample_mean + 1.96 * sample_se
  ) |>
  select(-sample)
head(biased_samples)
```

###

The downstream code is identical to what built `samples` in the previous section --- only the draw itself is biased. `biased_samples` will be heavy to recompute on every render, so caching it earns its keep.

### Exercise 5

Plot the first 100 rows of `biased_samples` as horizontal error bars with a red dashed vertical line at `truth` --- same picture as the unbiased version, but on `biased_samples`. Modify the working chunk to draw this plot. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r biased-sampling-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r biased-sampling-5-test}
#| echo: true
ggplot(biased_samples |> slice(1:100), aes(y = trial)) +
  geom_errorbarh(aes(xmin = lower, xmax = upper),
                 height = 0.3, alpha = 0.5) +
  geom_point(aes(x = sample_mean), size = 0.6) +
  geom_vline(xintercept = truth, color = "red",
             linetype = "dashed") +
  labs(title = "100 biased samples of 50 voters each",
       subtitle = "Most 95% intervals sit to the RIGHT of the truth",
       x = "Age (years)",
       y = "Sample number") +
  theme_minimal()
```

###

Compare this to the unbiased picture from the previous section. There the error bars were *centered* on the truth and the misses split roughly half above and half below. Here the error bars are *shifted right* --- they cluster above the truth, and the misses are almost always too high. Bias has a *direction*; sampling noise does not.

### Exercise 6

Modify the working chunk to compute and print the fraction of the 1,000 intervals in `biased_samples` that contain `truth`. Render. Copy the printed value from the rendered HTML below.

```{r biased-sampling-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 2)
```

###

```{r biased-sampling-6-test}
#| echo: true
biased_samples |>
  summarise(coverage = mean(lower < truth & upper > truth))
```

###

Coverage dropped from 94.5% in the unbiased section to **46%** here. Confidence intervals that *claim* to capture the truth 95% of the time are missing more than half the time --- and when they miss, they miss systematically high. Every single miss in this run was a CI whose entire interval sat above the truth.

This is the single most important lesson of the whole tutorial: **a bigger sample does not fix a biased sampling mechanism**. If we had used 1,000 biased voters instead of 50, the standard error would have shrunk by roughly a factor of $\sqrt{1000/50} \approx 4.5$ --- intervals about a fifth as wide --- and coverage would have *gotten worse*, because the narrower intervals would now miss the truth almost every time. More data sharpens the wrong answer.

### Exercise 7

To see *where* the biased sample means are clustering, modify the working chunk to plot a density curve of `biased_samples$sample_mean` with a red dashed vertical line at `truth` and a black solid vertical line at the average of `sample_mean`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r biased-sampling-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r biased-sampling-7-test}
#| echo: true
ggplot(biased_samples, aes(x = sample_mean)) +
  geom_density(fill = "grey70", color = "grey30") +
  geom_vline(xintercept = truth, color = "red",
             linetype = "dashed") +
  geom_vline(xintercept = mean(biased_samples$sample_mean)) +
  labs(title = "Distribution of 1,000 biased sample means",
       subtitle = "Black: average of biased means. Red dashed: truth.",
       x = "Sample mean of age (years)",
       y = NULL) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

###

The bell sits to the right of the truth. The average of the biased sample means is about 53.4 --- roughly 3.6 years above the truth of 49.79. That gap, between where the sample means cluster and where the truth lives, is **bias**. Sampling noise spreads each individual estimate around the cluster center; bias moves the cluster center itself.

### Exercise 8

Stage `analysis.qmd`, commit with the message `"Add biased sampling demonstration"`, and push to GitHub. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r biased-sampling-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

The standard error and the confidence interval tell you about the variability of an estimator under *the sampling mechanism you actually used*. If that mechanism is unbiased, they are the right intervals to report. If it is biased, the intervals are still mathematically correct under their own assumptions --- and still systematically wrong about the truth. This is what a failure of *representativeness* looks like: the difference between the sample mean and the population mean is the bias the sampling mechanism introduced, and no amount of arithmetic on the sample alone can recover it.


## Summary
###

(One more section to come. This is a placeholder so the tutorial renders end-to-end.)

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
