---
title: Sampling
author: David Kane
tutorial:
  id: sampling
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #02 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)

library(tidyverse)
library(primer.data)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

The world confronts us. Make decisions we must.

Imagine you are a voter-file analyst at a Michigan civic-engagement nonprofit preparing the next door-knock campaign. The state's registered-voter file lists more than seven million people, and you need to know the average age of those voters so that the campaign can plan its outreach --- older voters get printed materials, younger voters get app-based messaging. Walking the whole file every campaign cycle is impossible. Your boss, the campaign director, will use the number you report to decide how to split the printing budget against the app development budget. The estimate alone won't decide anything, but it is one good input. This tutorial focuses on one specific estimate: the average age of registered voters, with its associated uncertainty. There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

The best way to ensure that students remember these concepts more than a few months after the course ends is spaced repetition, although we focus more on the repetition than on the spacing.

### Exercise 2

You should be working inside a GitHub repo named `sampling`, opened in a Codespace from the [`PPBDS/codespace-starter`](https://github.com/PPBDS/codespace-starter) template. If you are not there yet, please create that repo and open it in a Codespace now --- see the [package README](https://github.com/PPBDS/primer/tree/main/primer.tutorials#working-environments-and-repo-setup) if you need setup instructions or are working locally instead. Once you're inside the `sampling` repo, create a new Quarto document titled `"sampling"` with yourself as the author, render it, and save it as `analysis.qmd`. Create a `.gitignore` file with `analysis_files` on the first line followed by a blank line. Save and push. At the R prompt, run `show_file(".gitignore")`. If that fails, it's probably because you haven't loaded `library(tutorial.helpers)` at the R prompt. CP/CR means *copy/paste the command you sent to R along with the response R gave back*; you will see this shorthand throughout the tutorials.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Professionals keep their data science work in the cloud because laptops fail.

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.data)` in a new code chunk. Render the file. Notice that the file does not look good because the code is visible and there are messages. Add `#| message: false` to remove the messages in this setup chunk. Also add the following to the YAML header to remove all code echos from the HTML:

````
execute:
  echo: false
````

At the R prompt, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Render again. Everything looks nice, albeit empty, because we have added code to make the file look better and more professional.

### Exercise 4

Place your cursor in the QMD file on the `library(tidyverse)` line. Use `Cmd/Ctrl + Enter` to execute that line. Note that this sends the line to the R prompt and runs it. CP/CR.

```{r introduction-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Place your cursor on the `library(primer.data)` line and do the same. The workflow --- write code in the QMD so you have a permanent copy, then send each line to the R prompt to run it interactively --- is the most common approach to data science. There is QMD World and R World. It is your responsibility to keep them in sync.

### Exercise 5

At the R prompt, type `?shaming`, and paste the Description below.

```{r introduction-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

The `shaming` tibble contains records on more than 300,000 Michigan registered voters from the 2006 study by Gerber, Green and Larimer. For this tutorial, we will pretend that the `shaming` tibble *is* the universe of voters our analyst is responsible for --- not a sample of it, but the entire population. We use it that way so that we can pretend, for pedagogical purposes, to know things we would never know in real life.


## Sampling and the Standard Error
###

In real life, we never get to see the whole universe. We see only a sample. From that sample, we compute a number --- here, the average age --- and we want to know how close that number is to the truth. The tool that answers that question is the **standard error**. Over this section we will build, step by step, a single growing code chunk in your QMD that draws samples, computes their averages, and visualizes how often those averages land near the truth.

### Exercise 1

In your QMD, create a new R code chunk. Tell your AI assistant: *"Compute the mean of the `age` column in the `shaming` tibble, removing missing values."* Paste the result you see when you run the line. CP/CR.

```{r sampling-and-the-standard-error-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
mean(shaming$age, na.rm = TRUE)
#> [1] 49.79
```

For the rest of the tutorial, we will treat this number --- 49.79 years --- as the **truth**: the average age of every voter in our universe. In an ordinary project this number would be unknowable; we only see it here because we declared the entire `shaming` tibble to be the universe.

### Exercise 2

Tell your AI assistant: *"In the same chunk, draw a random sample of 50 rows from `shaming` (call the result `my_sample`) and compute the mean of `age` in that sample. Add `set.seed(10)` at the top of the chunk so the sample is reproducible."* Run it and paste the mean. CP/CR.

```{r sampling-and-the-standard-error-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
set.seed(10)
my_sample <- shaming |> slice_sample(n = 50)
mean(my_sample$age, na.rm = TRUE)
#> [1] 50.02
```

Our sample mean is 50.02; the truth is 49.79. The two are close but not identical. That gap --- between the number a sample gives us and the unknown truth --- is what every conversation about uncertainty is really about.

### Exercise 3

A different sample of 50 voters would have given you a different mean. To say how much the sample mean would jump around if we kept drawing new samples, we use the **standard error**. For a sample mean, the standard error is the sample's standard deviation divided by the square root of the sample size.

Tell your AI assistant: *"In the same chunk, compute the standard error of `my_sample$age` --- the standard deviation divided by the square root of 50."* Run it and paste the result. CP/CR.

```{r sampling-and-the-standard-error-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
sd(my_sample$age, na.rm = TRUE) / sqrt(50)
#> [1] 2.03
```

You can do this calculation by hand. The standard deviation of the 50 ages in our sample is about 14.32. The square root of 50 is about 7.07. Divide them: $14.32 / 7.07 \approx 2.03$. That is what the line above computes. The standard error is about 2 years. A rough rule of thumb: roughly 95% of the time, the truth sits within two standard errors of the sample mean. In our case, two standard errors is four years, so a 95% confidence interval for the truth based on this one sample is about (46, 54). The actual truth is 49.79, which is inside that interval --- but we got lucky on this particular draw, and a different sample could miss.

The square root in the denominator is doing important work. Suppose we had drawn a sample of 1,000 voters instead of 50. The sample standard deviation would be roughly the same --- around 14 --- because that is a property of the population, not of the sample size. But the divisor changes: $\sqrt{1000} \approx 31.6$, so the standard error becomes $14 / 31.6 \approx 0.45$ years, almost a factor of five smaller than the 2.03 we got from 50 voters. The general rule: **to cut the standard error in half, you need four times the sample size, because $\sqrt{4} = 2$**. Doubling the sample size only shrinks the standard error by about 30%. Confidence is expensive.

### Exercise 4

Let's draw the picture. We will plot the sample mean as a point, draw a horizontal error bar covering the mean ± 1.96 standard errors, and mark the truth with a vertical line.

Tell your AI assistant: *"Add a ggplot to the same chunk that shows the single sample mean as a point at `y = 1`, with a horizontal error bar covering mean ± 1.96·SE, and a red dashed vertical line at the truth (the mean of `shaming$age`). Label the axes nicely."* Run it. You will see a plot. Paste a one-sentence description of what you see.

```{r sampling-and-the-standard-error-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
sample_mean <- mean(my_sample$age, na.rm = TRUE)
sample_se   <- sd(my_sample$age, na.rm = TRUE) / sqrt(50)
truth       <- mean(shaming$age, na.rm = TRUE)

ggplot() +
  geom_point(aes(x = sample_mean, y = 1), size = 3) +
  geom_errorbarh(aes(xmin = sample_mean - 1.96 * sample_se,
                     xmax = sample_mean + 1.96 * sample_se,
                     y = 1),
                 height = 0.2) +
  geom_vline(xintercept = truth, color = "red", linetype = "dashed") +
  labs(title = "One sample of 50 voters: mean ± 1.96·SE",
       subtitle = "Red dashed line is the truth",
       x = "Age (years)",
       y = NULL) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

That horizontal bar is the **95% confidence interval**. It is a statement about the *method*, not about this particular sample: if we repeated the experiment many times, about 95% of the intervals we draw would contain the truth.

### Exercise 5

To see the "about 95%" claim land, we need to draw not one sample but many. Tell your AI assistant: *"In the same chunk, draw 1,000 samples of 50 voters each from `shaming`, and for each sample compute its mean, its standard error, and the lower and upper 95% confidence bounds (mean ± 1.96·SE). Use `set.seed(10)` and store the results in a tibble called `samples` with one row per sample."* Run it. Paste the first few rows of `samples` (run `head(samples)` and copy the output). CP/CR.

```{r sampling-and-the-standard-error-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Our version:

```r
set.seed(10)
samples <- tibble(trial = 1:1000) |>
  mutate(data        = map(trial, ~ slice_sample(shaming, n = 50))) |>
  mutate(sample_mean = map_dbl(data, ~ mean(.$age, na.rm = TRUE))) |>
  mutate(sample_se   = map_dbl(data, ~ sd(.$age, na.rm = TRUE) / sqrt(50))) |>
  mutate(lower       = sample_mean - 1.96 * sample_se,
         upper       = sample_mean + 1.96 * sample_se)
```

Each row is one of the thousand hypothetical "what if we had drawn a different 50 voters?" experiments. Each carries its own sample mean, its own standard error, and its own 95% confidence interval. Most are near the truth; a few are far away.

### Exercise 6

Now plot 100 of those intervals as horizontal error bars stacked vertically, with the truth marked as a vertical red line. (We use 100 rather than 1,000 so the picture stays readable; we still count coverage on all 1,000 in the next step.)

Tell your AI assistant: *"Add a ggplot to the same chunk that shows the first 100 rows of `samples`: a horizontal error bar from `lower` to `upper` at `y = trial`, a point at `sample_mean`, and a red dashed vertical line at the truth. Label the axes."* Run it. Paste a one-sentence description of what you see.

```{r sampling-and-the-standard-error-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
ggplot(samples |> slice(1:100), aes(y = trial)) +
  geom_errorbarh(aes(xmin = lower, xmax = upper),
                 height = 0.3, alpha = 0.5) +
  geom_point(aes(x = sample_mean), size = 0.6) +
  geom_vline(xintercept = truth, color = "red", linetype = "dashed") +
  labs(title = "100 samples of 50 voters each",
       subtitle = "Most 95% intervals contain the truth; a few miss",
       x = "Age (years)",
       y = "Sample number") +
  theme_minimal()
```

Most of the intervals cross the red line --- those contain the truth. A handful do not. The misses are not a sign that the method is broken; they are exactly what we should expect when each interval is meant to catch the truth 95% of the time.

### Exercise 7

Let's count. Tell your AI assistant: *"In the same chunk, compute the fraction of the 1,000 confidence intervals in `samples` that contain the truth --- that is, intervals whose `lower` is less than the truth and whose `upper` is greater."* Run it and paste the result. CP/CR.

```{r sampling-and-the-standard-error-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
samples |>
  summarise(coverage = mean(lower < truth & upper > truth))
#> # A tibble: 1 × 1
#>   coverage
#>      <dbl>
#> 1    0.945
```

About 94.5% of the intervals contained the truth --- close to the 95% the method advertises, not exactly 95%. That gap (94.5% vs. 95%) is normal sampling noise. Run the whole procedure a second time with a different seed and you would get a different number close to but not equal to 95%.

### Exercise 8

Now polish the 100-sample plot for your final QMD. Tell your AI assistant: *"Update the 100-sample plot to have a clear title, a one-line subtitle naming the takeaway, a caption pointing to the data source (`primer.data`), and axis labels with units. Use `theme_minimal()`."*

Render your QMD (`Cmd/Ctrl + Shift + K`) and then at the R prompt run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r sampling-and-the-standard-error-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Every figure you make in this course should follow the same four-slot pattern: title names the plot, subtitle names the takeaway, caption names the data source, axis labels carry units. A plot whose subtitle doesn't tell the reader what to take away is a plot that hasn't earned its space.

### Exercise 9

Commit and push your changes. At the R prompt, run `tutorial.helpers::show_file("analysis.qmd")` to confirm the final state of your QMD. CP/CR.

```{r sampling-and-the-standard-error-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

The standard error is the bridge between "the number this one sample gave me" and "what could the truth reasonably be?" When the Cardinal Virtues chapters talk about uncertainty in an estimate --- the confidence interval on a regression coefficient, the width of a predicted probability, the spread of a posterior --- the underlying machinery is the same one you just saw: draw samples, watch how the answer jumps around, summarize the jumping with a standard error. The standard error gets shorter when the sample is bigger, and longer when the data is noisier. Everything else is detail.


## Biased Sampling
###

In the previous section the sampling was *random* --- every voter in the universe had the same chance of ending up in your sample. That assumption is what made the 95% confidence intervals do what they claimed. Real survey work is almost never that clean. Door-knockers find more people at home in older neighborhoods. App-based outreach reaches younger users. Phone-only sampling skips the cell-phone-only households. In this section we will rebuild the previous machinery --- but with a sampling mechanism that systematically over-represents older voters --- and watch the coverage collapse.

### Exercise 1

Suppose the campaign's door-knock app, by quirk of its routing, makes voters over the age of 40 *three times as likely* to appear in a sample as voters 40 or younger. The voters themselves haven't changed; the universe is the same `shaming` tibble. Only the *mechanism* that pulls voters into the sample has changed.

Tell your AI assistant: *"In the same growing code chunk, create a new biased sample of 50 voters by first adding a column `wt` to `shaming` that is `3` when `age > 40` and `1` otherwise, and then drawing 50 rows using `slice_sample(n = 50, weight_by = wt)`. Use `set.seed(10)`. Call the result `biased_sample` and show its first few rows."* Run it and paste the result. CP/CR.

```{r biased-sampling-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Our version:

```r
set.seed(10)
shaming_w <- shaming |> mutate(wt = ifelse(age > 40, 3, 1))
biased_sample <- shaming_w |> slice_sample(n = 50, weight_by = wt)
```

`weight_by` tells `slice_sample()` to make each row's probability of being picked proportional to that row's weight. Rows with `wt = 3` are three times as likely to land in the sample as rows with `wt = 1`.

### Exercise 2

Before trusting any of the downstream results, we want to *see* that the bias is doing what we wrote it to do. Tell your AI assistant: *"In the same chunk, make a side-by-side histogram comparing the age distribution of `biased_sample` against an unbiased sample of the same size. Draw the unbiased sample as `unbiased_sample <- slice_sample(shaming, n = 50)` with `set.seed(11)`. Use `bind_rows()` with an `origin` column and `facet_wrap(~ origin)` so the two histograms appear side by side."* Run it. Describe in one sentence what you see.

```{r biased-sampling-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
set.seed(11)
unbiased_sample <- shaming |> slice_sample(n = 50)

bind_rows(
  biased_sample   |> mutate(origin = "Biased (age > 40 weighted 3x)"),
  unbiased_sample |> mutate(origin = "Unbiased")
) |>
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "grey50", color = "white") +
  facet_wrap(~ origin) +
  labs(title = "Age distribution: biased vs. unbiased samples of 50",
       subtitle = "The biased sample is visibly skewed toward older voters",
       x = "Age (years)", y = NULL) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

In the universe (`shaming`), about 76% of voters are over 40. In the unbiased sample of 50, the share over 40 will sit near 76% give-or-take sampling noise. In the *biased* sample of 50, about 98% are over 40 --- you can read it straight off the histogram. The mechanism is doing exactly what we wrote.

### Exercise 3

Now compute the mean and standard error of the biased sample, the same way we did before. Tell your AI assistant: *"In the same chunk, compute and print the mean and standard error of `age` in `biased_sample` (standard deviation divided by the square root of 50)."* Run it. CP/CR.

```{r biased-sampling-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
mean(biased_sample$age, na.rm = TRUE)
#> [1] 55.26
sd(biased_sample$age, na.rm = TRUE) / sqrt(50)
#> [1] 1.51
```

Two things to notice. First, the biased sample's mean is 55.26 --- about 5.5 years above the truth of 49.79. The mechanism that over-sampled older voters pushed the sample mean up. Second, the standard error is **smaller** than it was in the unbiased section (1.51 vs. 2.03). That is because the biased sample is more homogeneous --- almost everyone is over 40 --- so the within-sample standard deviation is smaller. **The standard error formula has no way to know that our sample is biased**. It measures the spread inside the sample we got; it cannot see what the sampling mechanism left out.

### Exercise 4

Repeat the exercise from the previous section: draw 1,000 biased samples, compute the mean, standard error, and 95% confidence bounds for each, and store them in a tibble. Tell your AI assistant: *"In the same chunk, draw 1,000 biased samples of 50 voters each (using the same `weight_by = wt` mechanism and `set.seed(10)`) and store the means, standard errors, and 95% confidence bounds in a tibble called `biased_samples`."* Run it and paste `head(biased_samples)`. CP/CR.

```{r biased-sampling-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Our version:

```r
set.seed(10)
biased_samples <- tibble(trial = 1:1000) |>
  mutate(data        = map(trial, ~ slice_sample(shaming_w, n = 50, weight_by = wt))) |>
  mutate(sample_mean = map_dbl(data, ~ mean(.$age, na.rm = TRUE))) |>
  mutate(sample_se   = map_dbl(data, ~ sd(.$age, na.rm = TRUE) / sqrt(50))) |>
  mutate(lower       = sample_mean - 1.96 * sample_se,
         upper       = sample_mean + 1.96 * sample_se)
```

This is exactly the same code as before, with one change: `slice_sample(..., weight_by = wt)` instead of `slice_sample(..., n = 50)`. The downstream calculations don't know they're operating on biased samples; they cannot know.

### Exercise 5

Plot 100 of those confidence intervals as horizontal error bars, with the truth marked as a vertical red line --- same picture as the unbiased version, but on `biased_samples`.

Tell your AI assistant: *"Plot the first 100 rows of `biased_samples` as horizontal error bars (`lower` to `upper`) at `y = trial`, points at `sample_mean`, and a red dashed vertical line at the truth. Label the axes."* Run it. Describe in one sentence what you see.

```{r biased-sampling-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
ggplot(biased_samples |> slice(1:100), aes(y = trial)) +
  geom_errorbarh(aes(xmin = lower, xmax = upper),
                 height = 0.3, alpha = 0.5) +
  geom_point(aes(x = sample_mean), size = 0.6) +
  geom_vline(xintercept = truth, color = "red", linetype = "dashed") +
  labs(title = "100 biased samples of 50 voters each",
       subtitle = "Most 95% intervals sit to the RIGHT of the truth",
       x = "Age (years)",
       y = "Sample number") +
  theme_minimal()
```

Compare this to the unbiased picture from the previous section. There the error bars were *centered* on the truth and the misses split roughly half above and half below. Here the error bars are *shifted right* --- they cluster above the truth, and the misses are almost always too high. Bias has a *direction*; sampling noise does not.

### Exercise 6

Count the coverage. Tell your AI assistant: *"Compute the fraction of the 1,000 intervals in `biased_samples` that contain the truth."* Run it and paste the result. CP/CR.

```{r biased-sampling-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
biased_samples |>
  summarise(coverage = mean(lower < truth & upper > truth))
#> # A tibble: 1 × 1
#>   coverage
#>      <dbl>
#> 1    0.464
```

Coverage dropped from 94.5% in the unbiased section to **46%** here. Confidence intervals that *claim* to capture the truth 95% of the time are missing more than half the time --- and when they miss, they miss systematically high. Every single miss in this run was a CI whose entire interval sat above the truth.

This is the single most important lesson of the whole tutorial: **a bigger sample does not fix a biased sampling mechanism**. If we had used 1,000 biased voters instead of 50, the standard error would have shrunk by roughly a factor of $\sqrt{1000/50} \approx 4.5$ --- intervals about a fifth as wide --- and coverage would have *gotten worse*, because the narrow intervals would now miss the truth almost every time. More data sharpens the wrong answer.

### Exercise 7

To see *where* the biased sample means are clustering, tell your AI assistant: *"Plot a density curve of `biased_samples$sample_mean` with a vertical red dashed line at the truth and a vertical black solid line at the average of `sample_mean`. Use `theme_minimal()` and drop the y-axis since density-plot y-values are not meaningful."* Run it. Describe in one sentence what you see.

```{r biased-sampling-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our version:

```r
ggplot(biased_samples, aes(x = sample_mean)) +
  geom_density(fill = "grey70", color = "grey30") +
  geom_vline(xintercept = truth, color = "red", linetype = "dashed") +
  geom_vline(xintercept = mean(biased_samples$sample_mean)) +
  labs(title = "Distribution of 1,000 biased sample means",
       subtitle = "Black line is the average of the biased means; red dashed line is the truth",
       x = "Sample mean of age (years)",
       y = NULL) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

The bell sits to the right of the truth. The average of the biased sample means is about 53.4 --- roughly 3.6 years above the truth of 49.79. That gap, between where the sample means cluster and where the truth lives, is **bias**. Sampling noise spreads each individual estimate around the cluster center; bias moves the cluster center itself.

### Exercise 8

Render your QMD with `Cmd/Ctrl + Shift + K`, then at the R prompt run `show_file("analysis.qmd", chunk = "Last")`. Commit and push. CP/CR.

```{r biased-sampling-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

The standard error and the confidence interval tell you about the variability of an estimator under *the sampling mechanism you actually used*. If that mechanism is unbiased, they are the right intervals to report. If it is biased, the intervals are still mathematically correct under their own assumptions --- and still systematically wrong about the truth. When the Cardinal Virtues chapters talk about *representativeness*, this is the picture they have in mind: the difference between the sample mean and the population mean is the bias the sampling mechanism introduced, and no amount of arithmetic on the sample alone can recover it.


## Summary
###

(One more section to come. This is a placeholder so the tutorial renders end-to-end.)

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
