---
title: Ordered Factors
author: David Kane and Ivy Spratt
tutorial:
  id: ordered-factors
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial: #121 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(marginaleffects)


knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 600, 
        tutorial.storage = "local") 

x <- colleges |>
  select(tuition, grad_rate, selectivity) |>
   filter(tuition > 2)

fit_colleges <- linear_reg() |>
    set_engine("lm") |>
    fit(grad_rate ~ tuition + selectivity, 
        data = x)

```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial supports [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). 

The world confronts us. Make decisions we must.

Imagine that you are a data scientist working for a non-profit organization that helps students find the best college for them. You have been tasked with creating a model that predicts the graduation rate of colleges based on their tuition fee. 

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
	message = "Wisdom, Justice, Courage, and Temperance.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 2)
```

###

Why do we ask this, and a score more other questions, in each tutorial? Because the best way to (try to) ensure that students remember these concepts more than a few months after the course ends is [spaced repetition](https://en.wikipedia.org/wiki/Spaced_repetition), although we focus more on the repetition than on the spacing.

### Exercise 2

Create a Github repo called `ordered`. Make sure to click the "Add a README file" check box.

Connect the repo to a project on your computer using `File -> New Folder from Git ...`. Make sure to select the "Open in a new window" box. 

You need two Positon windows: this one for running the tutorial and the one you just created for writing your code and interacting with the Console.

Select `File -> New File -> Quarto Document ...`. Provide a title -- `"Ordered"` -- and an author (you). Render the document and save it as `ordered.qmd`.

Create a `.gitignore` file with `ordered_files` on the first line and then a blank line. Save and push.

In the Console, run:

```         
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the Console.

CP/CR.

```{r introduction-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Professionals keep their data science work in the cloud because laptops fail.

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.data)` in a new code chunk. Render the file.

Notice that the file does not look good because the code is visible and there are annoying messages. To take care of this, add `#| message: false` to remove all the messages in this `setup` chunk. Also add the following to the YAML header to remove all code echos from the HTML:

```         
execute: 
  echo: false
```

In the Console, run:

```         
show_file("ordered.qmd", start = -5)
```

CP/CR.

```{r introduction-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 6)
```

### 

Render again. Everything looks nice, albeit empty, because we have added code to make the file look better and more professional.

### Exercise 4

Place your cursor in the QMD file on the `library(tidyverse)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(tidyverse)` to be copied down to the Console and then executed. 

CP/CR.

```{r introduction-4}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

The data that we will use was downloaded from Opportunity Insights. Based at Harvard University, Opportunity Insights focuses on using big data to understand and improve economic mobility. They have many downloadable data sets, check it out for yourself [here!](https://opportunityinsights.org/data/)

### Exercise 5

In the Console, type `?colleges`, and paste the Description below.

```{r introduction-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 8)
```

### 

`colleges` contains data from over 900 colleges and universities in the United States. These draw primarily on data from the Department of Education’s (DOE) [IPEDS](https://nces.ed.gov/ipeds/) database in 2013 and the College Scorecard. 

### Exercise 6

Define a causal effect.

```{r introduction-6}
question_text(NULL,
	message = "A causal effect is the difference between two potential outcomes.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

According to the Rubin Causal Model, there must be two (or more) potential outcomes for any discussion of causation to make sense. This is simplest to discuss when the treatment only has two different values, thereby generating only two potential outcomes. 

### Exercise 7

Colleges are the broad topic of this tutorial. Given that topic, which variable in `colleges` should we use as our outcome variable? 

```{r introduction-7}
question_text(NULL,
	message = "We will use `grad_rate`.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 2)
```

### 

`grad_rate` is the rate of graduation at that college/university from the year 2013.

There are many acceptable answers that we could measure, but, for this tutorial we will stick with `grad_rate`.

### Exercise 8

Let's imagine a brand new variable which **does not exist** in the data. This variable should be binary, meaning that it only takes on one of two values. It should also, at least in theory, be manipulable. In other words, if the value of the variable is "X," or whatever, then it generates one potential outcome and if it is "Y," or whatever, it generates another potential outcome.

How might we manipulate this variable?


```{r introduction-8}
question_text(NULL,
	message = "We could create a new variable called `impv_food` where we could send funds to certain schools for the purpose of improving their food, for which `1` means funds were sent and `0` means funds were not send. We could compare the graduation rates after 5 years and see if improving food had any effect on it.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Any data set can be used to construct a causal model as long as there is at least one covariate that we can, at least in theory, manipulate. It does not matter whether or not anyone did, in fact, manipulate it. 

### Exercise 9

Given our (imaginary) treatment variable `impv_food`, how many potential outcomes are there for each college? Explain why.


```{r introduction-9}
question_text(NULL,
	message = "There are 2 potential outcomes because the treatment variable `impv_food` takes on 2 posible values: a college improved their food or a college didn't improve their food.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The same data set can be used to create, separately, lots and lots of different models, both causal and predictive. We can just use different outcome variables and/or specify different treatment variables. All of this stuff is a *conceptual* framework we apply to the data. It is never inherent in the data itself.

### Exercise 10

In a few sentences, specify the two different values for the imaginary treatment variable `impv_food`, for a single unit, guess at the potential outcomes which would result, and then determine the causal effect for that unit given those guesses.

```{r introduction-10}
question_text(NULL,
	message = "For a given college, assume that the value of the treatment variable might be 1 or 0. If the college gets 1 (meaning they improved their food), then the graduation rate (`grad_rate`) would be 0.85. If the college gets o (meaning they did not improve their food), then the graduation rate would be 0.60. The causal effect on the outcome of a treatment of 1 versus 0 is 0.85 - 0.6 --- i.e., the difference between two potential outcomes --- which equals 0.25, which is the causal effect.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The the definition of a causal effect is the difference between two potential outcomes. Of course, you can't just say that the causal effect is 10. The exact value depends on which potential outcome comes first in the subtraction and which second. There is, perhaps, a default sense in which the causal effect is defined as treatment minus control. 

Any causal connection means exploring the *within row* difference between two potential outcomes. We don't need to look at any other rows to have that conversation. 

### Exercise 11

Let's consider a *predictive* model. Which variable in `colleges` do you think might have an important connection to `grad_rate`? 

```{r introduction-11}
question_text(NULL,
	message = "One variable which might have an important connection could be `tuition`. People who pay more for a college may be more willing to spend time there.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 2)
```

### 

With a predictive model, there is only one outcome for each individual unit. There are not two potential outcomes because we are not considering any of the covariates to be treatment variables. We assuming that all covariates are "fixed." 

### Exercise 12

Specify two different groups of colleges which have specific value for `tuition` and which might have different average values for the `grad_rate`. 

```{r introduction-12}
question_text(NULL,
	message = "Some colleges might charge a tuition of $90,000 per year. Others might have a tuition of $30,000. Those two groups will, on average, have graduation rate.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

In predictive models, do not use words like "cause," "influence," "impact," or anything else which suggests causation. The best phrasing is in terms of "differences" between groups of units with different values for the covariate of interest.

### Exercise 13

Write a predicative question which connects the outcome variable `grad_rate` to a covariate of interest. 

```{r introduction-13}
question_text(NULL,
	message = "How does graduation rate differ between colleges at the 75th and 25th percentile of tuition?",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

We call the answer to this question is your "Quantity of Interest." 

Our Quantity of Interest might appear too specific, too narrow to capture the full complexity of the topic. There are many, many numbers which we are interested in, many numbers that we want to know. But we don't need to list them all here! We just need to choose one of them since our goal is just to have a specific number which helps to guide us in the creation of the Preceptor Table and, then, the model.

## Wisdom
### 

*All we can know is that we know nothing. And that’s the height of human wisdom.* - Leo Tolstoy

Imagine that you are a data scientist working for a non-profit organization that helps students find the best college for them. You have been tasked with creating a model that predicts the graduation rate of colleges based on their tuition fee. 

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
	message = "Wisdom requires a question, the creation of a Preceptor Table, and an examination of our data.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

*The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* -- John W. Tukey

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
	message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The Preceptor Table does not include all the covariates which you will eventually including in your model. It only includes covariates which you need to answer your question.

### Exercise 3

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-3}
question_text(NULL,
	message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be a treatment.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Our question is predicative. Not all predicative questions need covariates. If our question asked "What is the probability that a student will graduate Harvard University?" There would be no covariates in this Preceptor Table. Just one column stating whether or not the student graduated.

### Exercise 4

What are the units for this problem?

```{r wisdom-4}
question_text(NULL,
	message = "Each individual college in the United States.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Specifying the Preceptor Table forces us to think clearly about the units and outcomes implied by the question. The resulting discussion sometimes leads us to modify the question with which we started. No data science project follows a single direction. We always backtrack. There is always dialogue.

We model units, but we only really care about aggregates.

### Exercise 5

What is the outcome variable for this problem?

```{r wisdom-5}
question_text(NULL,
	message = "The outcome variable is `grad_rate`, which measures the graduation rate for each college.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The outcome variable that we really care about is often not the outcome variable which our data includes. This compromise --- working with what we *have* rather than what we really *want* --- is a part of most data science work in the real world.

### Exercise 6

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r wisdom-6}
question_text(NULL,
	message = "A useful covariate is `tuition`. Others that might be helpful include `selectivity`, size of campus, location of campus (could be `region`, `state`, or `zip`), and the prestige (or `tier`) of the school.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The term "covariates" is used in at least three ways in data science. First, it is all the variables which might be useful, regardless of whether or not we have the data. Second, it is all the variables for which we have data. Third, it is the set of variables in the data which we end up using in the model.

### Exercise 7

What are the treatments, if any, for this problem?

```{r wisdom-7}
question_text(NULL,
	message = "Since this is a predictive model, there are no treatments.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Remember that a treatment is just another covariate which, for the purposes of this specific problem, we are assuming can be manipulated, thereby, creating two or more different potential outcomes for each unit.

### Exercise 8

What moment in time does the Preceptor Table refer to?

```{r wisdom-8}
question_text(NULL,
	message = "The Preceptor Table refers to the current academic year, when students are choosing colleges.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

A Preceptor Table can never really refer to an exact instant in time since nothing is instantaneous in this fallen world.

In almost all practical problems, the data was gathered at a time other than that to which the Preceptor Table refers.

```{r}
colleges |>
  select(tuition, grad_rate, selectivity) |>
  filter(tuition > 2) |>
  ggplot(aes(x = tuition, y  = grad_rate)) + 
  geom_point() + 
  geom_smooth() + 
  labs(title = "Graduation Rate vs Tuition", 
       subtitle = "For colleges with tuition greater than $20,000",
       x = "Tuition (in tens of thousands)", 
       y = "Graduation Rate")
```

### Exercise 9

Describe in words the Preceptor Table for this problem.

```{r wisdom-9}
question_text(NULL,
	message = "The Preceptor Table has one row for each college. The columns are ID, `grad_rate` (the outcome variable), and `tuition` (the main covariate).",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The Preceptor Table for this problem looks something like this:

```{r}
library(gt)
library(tidyverse)


data <- tibble(
  ID = c("1", "2", "...", "10", "11", "...", "950"),
  graduation_rate = c(0.92, 0.55, "...", 0.42, 0.78, "...", 0.87),
  tuition = c(4.5, 1.3, "...", 2.9, 2.2, "...", 3.7)
)
data |>
  gt() |>
  tab_header(title = "Preceptor Table") |>
  cols_label(
    ID = md("ID"),
    graduation_rate = md("Graduation Rate"),
    tuition = md("Tuition (tens of thousands)")
  ) |>
  tab_style(
    cell_borders(sides = "right"),
    location = cells_body(columns = c(ID))
  ) |>
  tab_style(
    style = cell_text(align = "left", v_align = "middle", size = "large"), 
    locations = cells_column_labels(columns = c(ID))
  ) |>
  cols_align(align = "center", columns = everything()) |>
  cols_align(align = "left", columns = c(ID)) |>
  fmt_markdown(columns = everything()) |>
  tab_spanner(label = "Outcome", columns = c(graduation_rate)) |>
  tab_spanner(label = "Covariate", columns = c(tuition))

```

### 

Like all aspects of a data science problem, the Preceptor Table evolves as we continue our work. For example, at the start, we aren't sure what right-hand side variables will be included in the model, so we are not yet sure which columns must be in the Preceptor Table.

### Exercise 10

What is the narrow, specific question we will try to answer?

```{r wisdom-10}
question_text(NULL,
	message = "What effect does the tuition of a college have on its graduation rate?",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

###

The answer to this question is your "Quantity of Interest." It is OK if your question differs from ours. Many similar questions lead to the creation of the same model. For the purpose of this tutorial, let's use our question.

Our Quantity of Interest might appear too specific, too narrow to capture the full complexity of the topic. There are many, many numbers in which we are interested, many numbers that we want to know. But we don't need to list them all here! We just need to choose one of them since our goal is just to have a specific question which helped to guide us in the creation of the Preceptor Table and, soon, the model. 

### Exercise 11

Over the course of this tutorial, we will be creating a summary paragraph. The purpose of this exercise is to write the first two sentences of that paragraph.

The first sentence is a general statement about the overall topic, mentioning both the general class of the outcome variable and of at least one of the covariates. It is **not** connected to the initial "Imagine that you are XX" which set the stage for this project. That sentence can be rhetorical. It can be trite, or even a platitude. The purpose of the sentence is to let the reader know, gently, about our topic.

The second sentence does two things. First, it introduces the data source. Second, it introduces the specific question. The sentence can't be that long. Important aspects of the data include when/where it was gathered, how many observations it includes and the organization (if famous) which collected it.

Type your two sentences below.

```{r wisdom-11}
question_text(NULL,
	message = "Tuition and educational quality are among the main concerns for many students and their families when choosing a college. Using data from IPEDS, a system of interrelated surveys on postsecondary education in 2013, we seek to understand the relationship between graduation rates and tuition for colleges and universities in the United States.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Read our answer. It will not be the same as yours. You can, if you want, change your answer to incorporate some of our ideas. Do not copy/paste our answer exactly. Add your two sentences, edited or otherwise, to your QMD, `Cmd/Ctrl + Shift + K`, and then commit/push.

## Justice
### 

*The arc of the moral universe is long, but it bends toward justice.* - Theodore Parker

### Exercise 1

In your own words, name the five key components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
	message = "Justice concerns the Population Table and the four key assumptions which underlie it: validity, stability, representativeness, and unconfoundedness.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Justice is about concerns that you (or your critics) might have, reasons why the model you create might not work as well as you hope. 

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
	message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Validity is always about the *columns* in the Preceptor Table and the data. Just because columns from these two different tables have the same *name* does not mean that they are the same *thing*.

### Exercise 3

Provide one reason why the assumption of validity might not hold for the outcome variable: `grad_rate`. Use the words "column" or "columns" in your answer.

```{r justice-3}
question_text(NULL,
	message = "The `grad_rate` column in our data is reported by colleges themselves to the federal government, and colleges have an incentive to make their graduation rates look high. The `grad_rate` column in the Preceptor Table represents the true graduation rate. These two columns may not be measuring exactly the same thing.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

In order to consider the Preceptor Table and the data to be drawn from the same population, the columns from one must have a *valid correspondence* with the columns in the other. Validity, if true (or at least reasonable), allows us to construct the Population Table, which is the first step in Justice.

Because we control the Preceptor Table and, to a lesser extent, the original question, we can adjust those variables to be “closer” to the data that we actually have. This is another example of the iterative nature of data science. If the data is not close enough to the question, then we check with our boss/colleague/customer to see if we can modify the question in order to make the match between the data and the Preceptor Table close enough for validity to hold.

Despite these potential problems, we will assume that validity holds since it, mostly (?), does.

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
	message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The Population Table is almost always much bigger than the combination of the Preceptor Table and the data because, if we can really assume that both the Preceptor Table and the data are part of the same population, than that population must cover a broad universe of time and units since the Preceptor Table and the data are, themselves, often far apart from each other.

### Exercise 5

In your own words, define the assumption of "stability" when employed in the context of data science.

```{r justice-5}
question_text(NULL,
	message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Stability is all about *time*. Is the relationship among the columns in the Population Table stable over time? In particular, is the relationship --- which is another way of saying "mathematical formula" --- at the time the data was gathered the same as the relationship at the (generally later) time references by the Preceptor Table.


### Exercise 6

Provide one reason why the assumption of stability might not be true in this case.

```{r justice-6}
question_text(NULL,
	message = "College graduation requirements may have changed over time. This may result in different rates resulting from factors unaccounted for in our data.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

A change in time or the distribution of the data does not, in and of itself, demonstrate a violation of stability. Stability is about the parameters: $\beta_1$, $\beta_1$ and so on. Stability means these parameters are the same in the data as they are in the population as they are in the Preceptor Table.

### Exercise 7

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: `Data -> Population -> Preceptor Table`. In your own words, define the assumption of "representativeness" when employed in the context of data science.

```{r justice-7}
question_text(NULL,
	message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Ideally, we would like both the Preceptor Table *and* our data to be random samples from the population. Sadly, this is almost never the case.

### Exercise 8

We do not use the data, directly, to estimate missing values in the Preceptor Table. Instead, we use the data to learn about the overall population. Provide one reason, involving the relationship between the data and the population, for why the assumption of representativeness might not be true in this case.

```{r justice-8}
question_text(NULL,
	message = "Smaller schools may have been excluded from our dataset.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The reason that representativeness is important is because, when it is violated, the estimates for the model parameters might be biased. 


### Exercise 9

We use information about the population to make inferences about the Preceptor Table. Provide one reason, involving the relationship between the population and the Preceptor Table, for why the assumption of representativeness might not be true in this case.

```{r justice-9}
question_text(NULL,
	message = "Because the Preceptor Table includes the entire population, there is no problem with representativeness in using the population to draw inferences about the Precetor Table. They are one and the same.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Stability looks across time periods. Representativeness looks within time periods, for the most part. 

Of course, it is sometimes the case that the Preceptor Table includes every row from the Population Table for that moment in time. In that case, the assumption representativeness is met, by definition, if we only consider that moment. So, in that case, the only possible violation of representativeness must involve a claim that this moment in time is not representative of the rest of the Population Table.

### Exercise 10

In your own words, define the assumption of "unconfoundedness" when employed in the context of data science.

```{r justice-10}
question_text(NULL,
	message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

This assumption is only relevant for causal models. We describe a model as "confounded" if this is not true. The easiest way to ensure unconfoundedness is to assign treatment randomly.

### Exercise 11

Write one sentence which highlights a potential weakness in your model. This will almost always be derived from possible problems with the assumptions discussed above. We will add this sentence to our summary paragraph. So far, our version of the summary paragraph looks like this:

> Tuition fee and educational quality are one of the main concerns for many students and their families when going to college. Using data from a IPEDS, a system of interrelated surveys on postsecondary eduation in 2013, we seek to understand the relationship between graduation rates tuition for colleges/universities in the United States.

Of course, your version will be somewhat different.


```{r justice-11}
question_text(NULL,
	message = "Tuition fee and educational quality are one of the main concerns for many students and their families when going to college. Using data from a IPEDS, a system of interrelated surveys on postsecondary eduation in 2013, we seek to understand the relationship between graduation rates tuition for colleges/universities in the United States. Graduation rates could've changed tremendously since 2013 thus potentially weaken our model.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Add a weakness sentence to the summary paragraph in your QMD. You can modify your paragraph as you see fit, but do not copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`, and then commit/push.

## Courage
### 

*Courage is being scared to death, but saddling up anyway.* - John Wayne

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
	message = "Courage starts with math, explores models, and then creates the data generating mechanism.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Having decided on the basic mathematical structure of the model at the end of *Justice*, a choice mostly driven by the distribution of our outcome variable, we now turn toward estimating the model.

### Exercise 2

In the setup code chunk in your QMD, put `library(tidymodels)`. Run this line in the console.

CP/CR.

```{r courage-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 6)
```

### 

The *[tidymodels](https://www.tidymodels.org/)* framework is the most popular one in the R world for estimating models. *[Tidy Modeling with R](https://www.tmwr.org/)* by Max Kuhn and Julia Silge is a great introduction.


### Exercise 3

Because our outcome variable is continuous, start to create the model by using `linear_reg(engine = "lm")`.

```{r courage-3, exercise = TRUE}

```

```{r courage-3-hint-1, eval = FALSE}
linear_reg(engine = "lm")
```

```{r courage-3-test, include = FALSE}
linear_reg(engine = "lm")
```


### Exercise 4

Continue the pipe to `fit(grad_rate ~ tuition + region, data = colleges)`.

```{r courage-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-4-hint-1, eval = FALSE}
... |> 
  fit(..., 
      data = ...)
```

```{r courage-4-test, include = FALSE}
linear_reg() |>
    set_engine("lm") |>
    fit(grad_rate ~ tuition + region, 
        data = colleges)
```

### 

Recall that a categorical variable (whether character or factor) like region is turned into a set of $0/1$ "dummy" variables, each representing a region such as $region_{South}$ or $region_{Midwest}$. This is necessary because we can’t include words—like "South" or "Midwest"—directly in a mathematical formula, so dummy variables convert these categories into numerical indicators.

### Exercise 5

Continue the pipe with `tidy(conf.int = TRUE)`.

```{r courage-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-5-hint-1, eval = FALSE}
...
  tidy(... = TRUE)
```

```{r courage-5-test, include = FALSE}
linear_reg() |>
    set_engine("lm") |>
    fit(grad_rate ~ tuition + region, 
        data = colleges) |> 
	tidy(conf.int = TRUE)
```

### 

The intercept is the predicted graduation rate for the baseline category of region ("Midwest") when tuition is zero. The tuition coefficient ($\beta_1$) shows how graduation rate changes with each additional $10,000 of tuition. If $\beta_1 = 0.07$, then a $10,000 increase in tuition is associated with a 0.7-point increase in predicted graduation rate.

### Exercise 6

Change the call to `fit()` to `fit(grad_rate ~ tuition + public, data = colleges)`.

```{r courage-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-6-hint-1, eval = FALSE}
... |> 
  fit(..., data = ...)
  ...
```

```{r courage-6-test, include = FALSE}
linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition + public, data = colleges) |>
  tidy(conf.int = TRUE)
```

### 

The same dummy variable approach applies to a categorical covariate with $N$ values. Since public is a binary variable (TRUE or FALSE), the model creates one dummy variable (e.g., $public_{TRUE}$). Because most models include an intercept, the "missing" category (public = FALSE, i.e., private colleges) is incorporated into the intercept. The coefficient for public represents the difference in predicted graduation rate between public and private colleges, relative to this baseline.

### Exercise 7

<!-- XX: Fit the model one last time. -->

Change the call to `fit()` to `fit(grad_rate ~ tuition + selectivity, data = x)`.

```{r courage-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-7-hint-1, eval = FALSE}
... |> 
  fit(..., data = ...)
  ...
```

```{r courage-7-test, include = FALSE}
linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition + selectivity, data = colleges) |>
  tidy(conf.int = TRUE)
```

### 

The more variables we add, the more difficult it is to interpret the meaning of any particular coefficient. But interpretation also becomes less important. We don't really care about coefficients. We care about using our model to estimate quantities of interest.

### Exercise 8

Behind the scenes of this tutorial, an object called `fit_colleges` has been created which is the result of the code above, after filtering for colleges with tuition greater than $20k. Type `fit_colleges` and hit "Run Code." This generates the same results as using `print(fit_colleges)`.

```{r courage-8, exercise = TRUE}

```

```{r courage-8-hint-1, eval = FALSE}
fit_colleges
```

```{r courage-8-test, include = FALSE}
fit_colleges
```

### 

In data science, we deal with words, math, and code, but the most important of these is code. We created the mathematical structure of the model and then wrote a model formula in order to estimate the unknown parameters. 

### Exercise 9

We need `fit_colleges` to exit in Console World. Copy/paste this code into the Console and execute it.

````
x <- colleges |>
select(grad_rate, tuition, selectivity) |>
filter(tuition > 2)

linear_reg() |>
    set_engine("lm") |>
    fit(grad_rate ~ tuition + selectivity, 
        data = x)
````

CP/CR.

```{r courage-9}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

Just because something exists in the tutorial (or in the QMD) does not mean that it is in the Console. You should be aware of what exists in Console World, which is generally called your "workspace."

### Exercise 10

In your final code chunk in your QMD. Add the option: `#| cache: true`. Assign the final model result to `fit_colleges`.

Place your cursor in the QMD file on the `fit_colleges` line. Use `Cmd/Ctrl + Enter` to execute that line. Strictly speaking, this step is unnecessary because we already added `fit_colleges` to our workspace above. But ensuring that everything in the QMD is also in the Console is a good habit.

`Cmd/Ctrl + Shift + K`. It may take some time to render your QMD, depending on how complex your model is. But, by including `#| cache: true` you cause Quarto to cache the results of the chunk. The next time you render your QMD, as long as you have not changed the code, Quarto will just load up the saved fitted object.

At the Console, run:

```
tutorial.helpers::show_file("ordered.qmd", chunk = "Last")
```

CP/CR.

```{r courage-10}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 8)
```

### 

To confirm, `Cmd/Ctrl + Shift + K` again. It should be quick.

### Exercise 11

Add `*_cache` to `.gitignore` file. Cached objects are often large. They don't belong on Github.

At the Console, run:

```
tutorial.helpers::show_file(".gitignore")
```

CP/CR.

```{r courage-11}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 6)
```

###

Because of the change in your `.gitignore` (assuming that you saved it), the cache directory should not appear in the Source Control panel because Git is ignoring it, as instructed. Commit and push. 


### Exercise 12

In the Console, load the **[easystats](https://easystats.github.io/easystats/)** package. CP/CR.

```{r courage-12}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

We don't add **easystats** to the QMD because we are only using it for an interactive check of our fitted model. However, the [easystats ecosystem](https://easystats.github.io/easystats/) has a variety of interesting functions and packages which you might want to explore.

### Exercise 13

In the Console, run `check_predictions(extract_fit_engine(fit_colleges))`. CP/CR.

```{r courage-13}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

The purpose of `check_predictions()` is to compare your actual data (in green) with data that has been simulated from your fitted model, i.e., from your data generating mechanism. If your DGM is reasonable, then data simulated from it should not look too dissimilar from your actual data. Of course, it won't look exactly the same because of randomness, both in the world and in your simulation. But the actual data should be within the range of outcomes that your DGM simulates with `check_predictions()`.

### Exercise 14

Ask AI to create $\LaTeX$ code for this model, including our variable names and estimates for all the coefficients. Because this is a fitted model, the dependent variable will have a "hat" and the formula will not include an error term. 

Add the code to your QMD. `Cmd/Ctrl + Shift + K`.

Make sure the resulting display looks good. For example, you don't want an absurd number of figures to the right of the decimal. If the model is too long, you will need to spread it across several lines. You may need to go back-and-forth with the AI a few times.

Once the $\LaTeX$ code looks good, paste it below.

```{r courage-14}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

Our formula looks like:

$$
\widehat{\text{grad\_rate}} = 50.2 - 0.0031 \cdot \text{tuition} + 7.5 \cdot \text{selectivity}_{\text{Selective}} + 15.8 \cdot \text{selectivity}_{\text{HighlySelective}}
$$

It was created with:

```
$$
\widehat{\text{grad\_rate}} = 50.2 - 0.0031 \cdot \text{tuition} + 7.5 \cdot \text{selectivity}_{\text{Selective}} + 15.8 \cdot \text{selectivity}_{\text{HighlySelective}}
$$
```

First, we have replaced the parameters with our best estimates. Second, the left-hand side variable is $\widehat{\text{grad\_rate}}$ instead of $\text{grad_rate}$ because this formula generates our estimated `grad_rate`. A hat indicates an estimated value. 

**This is our data generating mechanism.**

Of course, there is randomness built into the DGM, but we won't worry about that detail for now.

A data generating mechanism is just a formula, something which we can write down and implement with computer code.

### Exercise 15

In the Console, run `tidy()` on `fit_colleges` with the argument `conf.int` set equal to `TRUE`. This returns 95% intervals for all the parameters in our model.

```{r courage-15, exercise = TRUE}

```

```{r courage-15-hint-1, eval = FALSE}
tidy(..., conf.int = ...)
```

```{r courage-15-test, include = FALSE}
tidy(fit_colleges, conf.int = TRUE)
```

### 

`tidy()` is part of the [**broom**](https://broom.tidymodels.org/) package, used to summarize information from a wide variety of models.

### Exercise 16

Create a new code chunk in your QMD. Ask AI to help you make a nice looking table from the tibble which is returned by `tidy()`. You don't have to include all the variables which `tidy()` produces. We often just show the estimate and the confidence intervals.

You might need to add some new libraries, e.g., **[tinytable](https://vincentarelbundock.github.io/tinytable/)**, **[knitr](https://yihui.org/knitr/)**, **[gt](https://gt.rstudio.com/)**, **[kableExtra](https://haozhu233.github.io/kableExtra/)**, **[flextable](https://davidgohel.github.io/flextable/)**, **[modelsummary](https://modelsummary.com/)**, et cetera, to the `setup` code chunk, if you use any functions from these packages, all of which have strengths and weaknesses for making tables.

Insert your table code into the QMD. (If you need a new library, and you probably will, place it in the `setup` code chunk along with the other libraries you have already loaded.)

`Cmd/Ctrl + Shift + K`. 

At the Console, run:

```
tutorial.helpers::show_file("ordered.qmd", chunk = "Last")
```

CP/CR.

```{r courage-16}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 12)
```

###

Our model:

```{r}
library(broom)
library(gt)

tidy(fit_colleges, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), round, digits = 2)) |>
  gt() |>
  cols_label(
    term = "Variable",
    estimate = "Estimate",
    conf.low = "95% CI (Lower)",
    conf.high = "95% CI (Upper)"
  ) |>
  tab_header(
    title = "Regression Coefficients"
  )
```

Our code:

```
library(broom)
library(gt)

tidy(fit_colleges, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), round, digits = 2)) |>
  gt() |>
  cols_label(
    term = "Variable",
    estimate = "Estimate",
    conf.low = "95% CI (Lower)",
    conf.high = "95% CI (Upper)"
  ) |>
  tab_header(
    title = "Regression Coefficients"
  )
```

At the very least, your table should include a title and a caption with the data source. The more you use AI, the better you will get at doing so.

### Exercise 17

Add a sentence to your project summary.

Explain the structure of the model. Something like: "I/we model Y [the concept of the outcome, not the variable name] as a [linear/logistic/multinomial/ordinal] function of X [and maybe other covariates]."

Recall the beginning of our version of the summary:

> Tuition fee and educational quality are one of the main concerns for many students and their families when going to college. Using data from a IPEDS, a system of interrelated surveys on postsecondary eduation in 2013, we seek to understand the relationship between graduation rates tuition for colleges/universities in the United States. Graduation rates could've changed tremendously since 2013 thus potentially weaken our model.

```{r courage-17}
question_text(NULL,
	message = "Tuition fee and educational quality are one of the main concerns for many students and their families when going to college. Using data from a IPEDS, a system of interrelated surveys on postsecondary eduation in 2013, we seek to understand the relationship between graduation rates tuition for colleges/universities in the United States. Graduation rates could've changed tremendously since 2013 thus potentially weaken our model. We modeled graduation rate, a continuous variable, as a linear function of both tuition and selectivity. Graduation rates seem to rise as tuition increases.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Read our answer. It will not be the same as yours. You can, if you want, change your answer to incorporate some of our ideas. Do not copy/paste our answer exactly. Add your two sentences, edited or otherwise, to summary paragraph portion of your QMD. `Cmd/Ctrl + Shift + K`, and then commit/push.

## Temperance
### 

*Temperance is a tree which as for its root very little contentment, and for its fruit calm and peace.* - Buddha

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
	message = "Temperance uses the data generating mechanism to answer the questions with which we began. Humility reminds us that this answer is always a lie. We can also use the DGM to calculate many similar quantities of interest, displaying the results graphically.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Courage gave us the data generating mechanism. Temperance guides us in the use of the DGM — or the “model” — we have created to answer the questions with which we began. We create posteriors for the quantities of interest. We should be modest in the claims we make. The posteriors we create are never the “truth.” The assumptions we made to create the model are never perfect. Yet decisions made with flawed posteriors are almost always better than decisions made without them.

### Exercise 2

Before using the DGM, we should make sure that we can interpret it.

What does the estimate for selectivityHighlySelective tell us?

Recall the values for the parameters in our data generating mechanism:

```{r}
 fit_colleges |> 
 	tidy(conf.int = TRUE) |> 
	select(term, estimate, conf.low, conf.high)
```

<!-- I find the selectivity variable names to be difficult to interpret. Consider renaming in the code. -->

```{r temperance-2}
question_text(NULL,
	message = "The estimate for selectivityHighlySelective is 0.16, with a 95% confidence interval from 0.13 to 0.19. This means that, when comparing colleges that are highly selective to the baseline category, we predict a 16 percentage point higher graduation rate, adjusting for tuition. Since the confidence interval does not include zero, we can be reasonably confident that this positive association is not due to random chance.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

###

Dummy variables must always be interpreted in the context of the base value for that variable, which is generally included in the intercept. The confidence interval excluding zero tells us this relationship is statistically significant.

### Exercise 3

What does the estimate for tuition suggest about the relationship between tuition and graduation rate?

```{r}
 fit_colleges |> 
 	tidy(conf.int = TRUE) |> 
	select(term, estimate, conf.low, conf.high)
```

```{r temperance-3}
question_text(NULL,
	message = "The estimate for tuition is 0.09, meaning that for each additional $10k of tuition, graduation rate is predicted to increase by 0.9 percentage points, adjusting for selectivity. The confidence interval (0.074, 0.106) does not include zero, so the effect is statistically significant.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

###

Numeric variables are harder to interpret directly. A helpful strategy is to compare two plausible values of the numeric variable (e.g., $10,000 apart) and interpret the difference in outcomes.

### Exercise 4

What does it mean that the confidence intervals for selectivitySelective and selectivityHighlySelective do not overlap?

```{r}
 fit_colleges |> 
 	tidy(conf.int = TRUE) |> 
	select(term, estimate, conf.low, conf.high)
```

```{r temperance-4}
question_text(NULL,
	message = "The confidence interval for selectivitySelective is [0.04, 0.10], and for selectivityHighlySelective it is [0.13, 0.19]. Because these intervals do not overlap, we can reasonably conclude that graduation rates are higher for highly selective colleges than for moderately selective ones, adjusting for tuition. The difference is not only positive but also statistically distinguishable.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

###

If the confidence intervals for dummy variables from the same categorical variable do not overlap, we can be more confident about the ordering of their effects.

### Exercise 5

In the end, we don't really care about parameters, much less how to interpret them. Parameters are *imaginary*, like unicorns. We care about answers to our questions. Parameters are tools for answering questions. They aren't interesting in-and-of themselves. *In the modern world, all parameters are nuisance parameters.* 

Add `library(marginaleffects)` to the QMD file.

Place your cursor in the QMD file on the `library(marginaleffects)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(marginaleffects)` to be copied down to the Console and then executed. 

CP/CR.

```{r temperance-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

We should be modest in the claims we make. The posteriors we create are never the “truth.” The assumptions we made to create the model are never perfect. Yet decisions made with flawed posteriors are almost always better than decisions made without them.

### Exercise 6

What is the specific question we are trying to answer? 

```{r temperance-6}
question_text(NULL,
	message = "What is the relationship between tuition and graduation rates for colleges and universities in the United States?",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Data science projects begin with a decision which we face. To make that decision wisely, we would like to have good estimates of many unknown numbers. Yet, in order to make progress, we need to drill down to one specific question. This leads to the creation of a data generating mechanism, which can now be used to answer lots of questions, thus allowing us to explore the original decision more broadly.

### Exercise 7

In the Console, `predictions()` on `fit_colleges`.

CP/CR.

```{r temperance-7}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

```{r temperance-7-test, include = FALSE}
predictions(fit_colleges)
```

###

`predictions()` returns a data frame with one row for each observation in the data set used to fit the model. In this case, predictions() returns 542 rows, because the input data has 542 observations. Each row contains the model's predicted value for that specific observation, along with a standard error and confidence interval for the estimate.

<!-- XX: You can insert other marginaleffects questions here. Below, we just show one plot_predictions() question, but your tutorial should really have several. -->

### Exercise 8

At the Console, run `plot_predictions()` on `fit_colleges` with `condition` set to `"selectivity"`. 

CP/CR.

```{r temperance-8}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

```{r temperance-8-test, include = FALSE}
plot_predictions(fit_colleges, 
                condition = "selectivity")
```

### 

The `plot_predictions()` function visualizes the estimated effect of a categorical variable—in this case, `selectivity`—on the predicted outcome. Each point on the plot represents the model’s predicted graduation rate for a level of selectivity, and the vertical lines represent the 95% confidence intervals, showing the uncertainty around each estimate. For example, if the point for “Highly Selective” is around 0.72 with error bars ranging from about 0.70 to 0.74, this tells us that we’re 95% confident the true average graduation rate for highly selective colleges lies in that range.

### Exercise 9

In the Console, run the same `plot_predictions()` call as above again, but with `draw = FALSE`.

CP/CR.

```{r temperance-9}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

### 

Because `plot_predictions()` returns a ggplot object, you can just continue with ggplot commands like `labs()`. However, it can be useful to see the actual underlying values in the tibble and then build your own plot directly from them.

### Exercise 10

Work with AI to create a beautiful plot starting with the output to the above call to `plot_predictions()`. Do this in your QMD since that is much easier than working in the Console directly.

Your title should highlight the key variables. Your subtitle should describe an important takeaway, the sentence/conclusion which readers will, you hope, remember. Your caption should mention the data source. Your axis labels should look nice.

This plot is not directly connected to your question. It answers lots of questions! It might be used by lots of different people. 

Copy the code for your plot here:

```{r temperance-10}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 15)
```

### 

Our plot:

```{r}
library(ggplot2)

preds <- plot_predictions(fit_colleges, condition = "selectivity", draw = FALSE)

ggplot(preds, aes(x = selectivity, y = estimate)) +
  geom_point(size = 4, color = "#1f77b4") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = "#1f77b4") +
  labs(
    title = "Predicted Graduation Rate by College Selectivity",
    subtitle = "Highly selective colleges are predicted to have the highest graduation rates",
    x = "College Selectivity Level",
    y = "Predicted Graduation Rate",
    caption = "Source: College dataset used in fit_colleges model"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 13, margin = margin(b = 10)),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )
```


Our code:

````
library(ggplot2)

preds <- plot_predictions(fit_colleges, condition = "selectivity", draw = FALSE)

ggplot(preds, aes(x = selectivity, y = estimate)) +
  geom_point(size = 4, color = "#1f77b4") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = "#1f77b4") +
  labs(
    title = "Predicted Graduation Rate by College Selectivity",
    subtitle = "Highly selective colleges are predicted to have the highest graduation rates",
    x = "College Selectivity Level",
    y = "Predicted Graduation Rate",
    caption = "Source: College dataset used in fit_colleges model"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 13, margin = margin(b = 10)),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )
````

Data science often involves this-back-and-forth style of work. First, we need to make a single chunk of code, in this case, a new plot, work well. This requires interactive work between the QMD and the Console. Second, we need to ensure that the entire QMD runs correctly on its own.

### Exercise 11

Finalize the new graphics code chunk in your QMD. `Cmd/Ctrl + Shift + K` to ensure that it all works as intended. 

At the Console, run:

```
tutorial.helpers::show_file("ordered.qmd", chunk = "Last")
```

CP/CR.

```{r temperance-11}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 10)
```


### 

Always [remember](https://en.wikipedia.org/wiki/Map%E2%80%93territory_relation): *The map is not the territory.* A beautiful graphic tells a story, but that story is always an imperfect representation of reality. Our models depend on assumptions, assumptions which are never completely true.

### Exercise 12

Write the last sentence of your summary paragraph. It describes at least one quantity of interest (QoI) and provides a measure of uncertainty about that QoI. It is OK if this QoI is not the one that you began with. The focus of a data science project often changes over time. It is also OK to discuss more than one QoI. Do whatever seems reasonable, given the context.

```{r temperance-12}
question_text(NULL,
	message = "Tuition fee and educational quality are one of the main concerns for many students and their families when going to college. Using data from a IPEDS, a system of interrelated surveys on postsecondary eduation in 2013, we seek to understand the relationship between graduation rates tuition for colleges/universities in the United States. Graduation rates could've changed tremendously since 2013 thus potentially weaken our model. We modeled graduation rate, a continuous variable, as a linear function of both tuition and selectivity. Graduation rates seem to rise as tuition increases. Colleges in the 75th percentile are have on average an 11% higher graduation rate as opposed to colleges in the 25th percentile, although that number may be anywhere from 9-13%.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Add a final sentence to your summary paragraph in your QMD as you see fit, but do not copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`.


### Exercise 13

Rearrange the material in your QMD so that the order is graphic, followed by the paragraph. Doing so, of course, requires sensible judgment. For example, the code chunk which creates the fitted model must occur before the chunk which creates the graphic. You can keep or discard the math and any other material at your own discretion.

`Cmd/Ctrl + Shift + K` to ensure that everything works.

At the Console, run:

```
tutorial.helpers::show_file("ordered.qmd")
```

CP/CR.

```{r temperance-13}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 30)
```

### 

This is the version of your QMD file at which your teacher is most likely to look closely.

### Exercise 14

Publish your rendered QMD to GitHub Pages. In the Terminal --- not the Console! --- run:

````
quarto publish gh-pages ordered.qmd
````

Copy/paste the resulting URL below.

```{r temperance-14}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Commit/push everything.

### Exercise 15

Copy/paste the URL to your Github repo.

```{r temperance-15}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God. If all our assumptions are correct, then our DGM is true, it accurately describes the way in which the world works. There is no better way to predict the future, or to model the past, than to use it. Sadly, this will only be the case with toy examples involving things like coins and dice. We hope that our DGM is close to the true DGM but, since our assumptions are never perfectly correct, our DGM will always be different. The estimated magnitude and importance of that difference is a matter of judgment.

The world confronts us. Make decisions we must.

## Summary
### 

This tutorial supports [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). 

We started the Tutorial with a question:

> *What is the relationship between tuition and graduation rates for colleges and universities in the United States?*

Using data from a IPEDS, a system of interrelated surveys on postsecondary eduation in 2013, we seek to understand the relationship between graduation rates tuition for colleges/universities in the United States. Graduation rates could've changed tremendously since 2013 thus potentially weaken our model. We modeled graduation rate, a continuous variable, as a linear function of both tuition and selectivity. Graduation rates seem to rise as tuition increases. Colleges in the 75th percentile are have on average an 11% higher graduation rate as opposed to colleges in the 25th percentile, although that number may be anywhere from 9-13%.

```{r}
plot_predictions(fit_colleges, 
                condition = "selectivity") +
  labs(
    title = "Graduation Rates Based on Selectivity",
    subtitle = "Graduation rates seem to rise as selectivity increases",
    caption = "Data from IPEDS, 2013",
    x = "Tuition",
    y = "Graduation Rate",
    color = "Tuition Percentile") +
    scale_color_manual(values = c("3.73" = "cyan3", "2.63" = "red"),
                       labels = c("3.73" = "75th Percentile", "2.63" = "25th Percentile")) + 
      guides(color = guide_legend(reverse = TRUE)) +
      theme_minimal()
```

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
