---
title: Trains
author: David Kane and Aashna Patel
tutorial:
  id: 07-trains
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #07 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(marginaleffects)

library(easystats)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 600, 
        tutorial.storage = "local") 

fit_att <- linear_reg(engine = "lm") |> 
  fit(att_end ~ treatment, data = trains)
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial supports [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). 

The world confronts us. Make decisions we must.

Imagine that you are a campaign manager for a Republican congressional candidate in Georgia. Your goal is to elect your candidate. Your candidate is conservative about immigration. You know that the more conservative a voter is about immigration, the more likely she is to vote for your boss. How should you spend your campaign funds to increase your odds of winning the election? There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

Why do we ask this, and a score more other questions, in each tutorial? Because the best way to (try to) ensure that students remember these concepts more than a few months after the course ends is [spaced repetition](https://en.wikipedia.org/wiki/Spaced_repetition), although we focus more on the repetition than on the spacing.

### Exercise 2

Create a Github repo called `trains`. Make sure to click the "Add a README file" check box.

Connect the repo to a project on your computer using `File -> New Folder from Git ...`. Make sure to select the "Open in a new window" box. 

You need two Positron windows: this one for running the tutorial and the one you just created for writing your code and interacting with the Console.

Select `File -> New File -> Quarto Document ...`. Provide a title -- `"Trains"` -- and an author (you). Render the document and save it as `trains.qmd`.

Create a `.gitignore` file with `trains_files` on the first line and then a blank line. Save and push.

In the Console, run:

```         
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the Console.

CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

Professionals keep their data science work in the cloud because laptops fail.

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.data)` in a new code chunk. Render the file.

Notice that the file does not look good because the code is visible and there are annoying messages. To take care of this, add `#| message: false` to remove all the messages in this `setup` chunk. Also add the following to the YAML header to remove all code echos from the HTML:

```         
execute: 
  echo: false
```

In the Console, run:

```         
show_file("trains.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

### 

Render again. Everything looks nice, albeit empty, because we have added code to make the file look better and more professional.

### Exercise 4

Place your cursor in the QMD file on the `library(tidyverse)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(tidyverse)` to be copied down to the Console and then executed. 

CP/CR.

```{r introduction-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The `trains` tibble measures attitudes toward immigration among Boston commuters. Individuals were exposed to one of two possible conditions, and then their attitudes towards immigrants were recorded. One condition was waiting on a train platform near individuals speaking Spanish. The other was being on a train platform without Spanish-speakers.

### Exercise 5

Place your cursor in the QMD file on the next line. Use `Cmd/Ctrl + Enter` to execute that line.

This work flow --- writing things in the QMD so that you have a permanent copy and then executing them in the Console with `Cmd/Ctrl + Enter` --- is the most common approach to data science. 

There is QMD World and R World. It is your responsibility to keep them in sync.

CP/CR.

```{r introduction-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

A version of the data from **primer.data** is available in the `trains` tibble.

### Exercise 6

In the Console, type `?trains`, and paste the Description below.

```{r introduction-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

### 

The data we have was collected by Enos (2014) from commuters on nine train platforms around Boston, Massachusetts in 2012. The data was for attitudes toward immigration-related policies, both before and after an experiment which randomly exposed a treated group to Spanish-speakers on a Boston commuter train platform.

### Exercise 7

Define a causal effect.

```{r introduction-7}
question_text(NULL,
  message = "A causal effect is the difference between two potential outcomes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

According to the Rubin Causal Model, there must be two (or more) potential outcomes for any discussion of causation to make sense. This is simplest to discuss when the treatment only has two different values, thereby generating only two potential outcomes. 

### Exercise 8

What is the fundamental problem of causal inference?

```{r introduction-8}
question_text(NULL,
  message = "The fundamental problem of causal inference is that we can only observe one potential outcome.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

If the treatment variable is continuous (like income), then there are lots and lots of potential outcomes, one for each possible value of the treatment variable. 

### Exercise 9

Voter attitudes toward immigration is the broad topic of this tutorial. Given that topic, which variable in `trains` should we use as our outcome variable? 

```{r introduction-9}
question_text(NULL,
  message = "`att_end` is the commuter's attitude toward immigration after the experiment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

### 

We will use `att_end` as our outcome variable.

```{r}
ggplot(trains, aes(x = att_end)) +
  geom_bar(color = "white", 
           alpha = 0.8) +
  scale_x_continuous(breaks = 3:15, 
                     limits = c(2.4, 15.6)) +
  labs(
    title = "Distribution of Attitudes Toward Immigration",
    subtitle = "Most individuals cluster toward the middle",
    caption = "Source: Enos (2014)",
    x = "Final Attitude Score (att_end)",
    y = "Count"
  ) +
  theme_minimal()
```

### Exercise 10

Let's imagine a brand new variable which **does not exist** in the data. This variable should be binary, meaning that it only takes on one of two values. It should also, at least in theory, be manipulable. In other words, if the value of the variable is "3," or whatever, then it generates one potential outcome and if it is "9," or whatever, it generates another potential outcome.

Describe this imaginary variable and how might we manipulate its value.

For now, ignore the actual treatment variable `treatment` which we will be using later in the analysis. The point of this exercise is to reinforce our understanding of the [Rubin Causal Model](https://ppbds.github.io/primer/rubin-causal-model.html).


```{r introduction-10}
question_text(NULL,
  message = "Imagine a variable called `phone_call` which has a value of `1` if the person received a phone call listing problems with immigration and `0` if they did not receive such a phone call. We, meaning the organization in charge of making such phone calls, can manipulate this variable by deciding, either randomly or otherwise, whether or not we will call a specific individual.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Any data set can be used to construct a causal model as long as there is at least one covariate that we can, at least in theory, manipulate. It does not matter whether or not anyone did, in fact, manipulate it. 

### Exercise 11

Given our (imaginary) treatment variable `phone_call`, how many potential outcomes are there for each person? Explain why.

```{r introduction-11}
question_text(NULL,
  message = "There are two potential outcomes because the treatment variable `phone_call` takes on two possible values: `1` or `0`, i.e., each person either receives a phone call or does not.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The same data set can be used to create, separately, lots and lots of different models, both causal and predictive. We can just use different outcome variables and/or specify different treatment variables. This is a *conceptual* framework we apply to the data. It is never inherent in the data itself.

### Exercise 12

In a few sentences, specify the two different values for the imaginary treatment variable `phone_call`, for a single unit, guess at the potential outcomes which would result, and then determine the causal effect for that unit given those guesses.

```{r introduction-12}
question_text(NULL,
  message = "For a given person, assume that the value of the treatment variable might be `1` or `0`. If the person gets `1`, then attitude toward immigration would be 9 on a 3 to 15 unit scale. If the person gets `0`, then the attitude would be 7. The causal effect on the outcome of a treatment of `1` versus `0` is `9 - 7` --- i.e., the difference between two potential outcomes --- which equals 2, which is the causal effect.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A causal effect is defined as the difference between two potential outcomes. "Difference" does not necessarily mean "subtraction" — many potential outcomes are not numbers. Even for numeric outcomes, you can’t simply say the effect is 10 without specifying the order of subtraction.

### Exercise 13

Let's consider a *predictive* model. Which variable in `trains` do you think might have an important connection to `att_end`? 

```{r introduction-13}
question_text(NULL,
  message = "`sex` might have a connection to `att_end`.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

### 

With a predictive model, each individual unit has only one observed outcome. There are not two potential outcomes because none of the covariates are treated as treatment variables. Instead, all covariates are assumed to be "fixed."

Predictive models have no "treatments" --- only covariates.

### Exercise 14

Specify two different groups of people who have different values for `sex` and which might have different average values for the `att_end`.  

```{r introduction-14}
question_text(NULL,
  message = "Consider two groups, the first with a value for `sex` of 'Female' and the second with value 'Male'. Those two groups might have different average values for `att_end`.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

In predictive models, do not use words like "cause," "influence," "impact," or anything else which suggests causation. The best phrasing is in terms of "differences" between groups of units with different values for a covariate of interest.

Any causal connection means exploring the *within row* difference between two potential outcomes. There's no need to consider other rows.

### Exercise 15

Write a causal question which connects the outcome variable `att_end` to `treatment`, the covariate of interest. 

```{r introduction-15}
question_text(NULL,
  message = "What is the average causal effect of exposure to Spanish-speakers on attitudes toward immigration?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

This is the first version of the question. We will now create a Preceptor Table to answer the question. We may then revise the question given complexities discovered in the data. We then update the question and the Preceptor Table. And so on.

## Wisdom
###

Data science starts with some broad questions and a data set which might help us to answer them.

Imagine that you are a campaign manager for a Republican congressional candidate in Georgia. Your goal is to elect your candidate. Your candidate is conservative about immigration. You know that the more conservative a voter is about immigration, the more likely she is to vote for your boss. How should you spend your campaign funds to increase your odds of winning the election? There are many decisions to make.

What is the average causal effect of exposure to Spanish-speakers on attitudes toward immigration?

We will work from data collected by [Enos (2014)](https://www.pnas.org/doi/10.1073/pnas.1317670111) from commuters on nine train platforms around Boston, Massachusetts in 2012, available in the `trains` tibble of the `primer.data` package.

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
    message = "Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.",
    answer(NULL, correct = TRUE),
    allow_retry = FALSE,
    incorrect = NULL,
    rows = 3)
```

###

*The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* -- John W. Tukey

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The Preceptor Table does not include all the covariates which you will eventually include in your model. It only includes, along with the outcome(s), covariates which are mentioned in your question.

### Exercise 3

Use AI to produce a plot of the distribution of `att_end` in the `trains` data.

```{r wisdom-3, exercise = TRUE}

```

```{r wisdom-3-hint-1, eval = FALSE}
trains |>
  ggplot(aes(x = att_end)) +
  geom_bar()
```

```{r wisdom-3-test, include = FALSE}
trains |>
  ggplot(aes(x = att_end)) +
  geom_bar(color = "white") +
  labs(title = "Distribution of Post-Experiment Immigration Attitudes",
       subtitle = "Most commuters score between 8 and 12",
       caption = "Source: Enos (2014)",
       x = "Attitude Score (att_end)",
       y = "Count") +
  theme_minimal()
```

###

`att_end` ranges from 3 to 15, where higher values indicate more conservative attitudes toward immigration. Most commuters cluster in the middle of the scale.

### Exercise 4

Use AI to produce a plot showing the number of commuters in each level of the `treatment` variable.

```{r wisdom-4, exercise = TRUE}

```

```{r wisdom-4-hint-1, eval = FALSE}
trains |>
  ggplot(aes(x = treatment)) +
  geom_bar()
```

```{r wisdom-4-test, include = FALSE}
trains |>
  ggplot(aes(x = treatment)) +
  geom_bar(color = "white", fill = "steelblue") +
  labs(title = "Treatment Assignment in the Trains Experiment",
       subtitle = "Roughly equal numbers of treated and control commuters",
       caption = "Source: Enos (2014)",
       x = "Treatment Condition",
       y = "Count") +
  theme_minimal()
```

###

The experiment randomly assigned commuters to `Treated` (exposed to Spanish-speakers) or `Control` (no exposure). Random assignment is what makes it possible to interpret the effect causally.

### Exercise 5

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-5}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

This problem is causal so one of the covariates is a treatment. In our problem, the treatment is `treatment`. There is a potential outcome for each of the 2 possible values of the treatment.

### Exercise 6

What are the units for this problem?

```{r wisdom-6}
question_text(NULL,
  message = "The units are the individual registered voters in our congressional district in Georgia.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Specifying the Preceptor Table forces us to think clearly about the units and outcomes implied by the question. The resulting discussion sometimes leads us to modify the question with which we started. No data science project follows a single direction. We always backtrack. There is always dialogue.

We model units, but we only really care about aggregates.

### Exercise 7

What is the outcome variable for this problem?

```{r wisdom-7}
question_text(NULL,
  message = "Keep track of two outcome variables: the one in our Preceptor Table and the one in our data. In the Preceptor Table, the outcome is a Georgia voter's true attitude toward immigration. In our data, `att_end` measures each Boston commuter's self-reported attitude after the experiment — a reasonable stand-in, but not identical.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The outcome variable that we really care about is often not the outcome variable which our data includes. This compromise --- working with what we *have* rather than what we really *want* --- is a part of most data science work in the real world.


### Exercise 8

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r wisdom-8}
question_text(NULL,
  message = "Past survey results about immigration, from a year or more ago, might be useful. The data does include `att_start` but that was measured just a few days before the experiment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The term "covariates" is used in at least three ways in data science. First, it is all the variables which might be useful, regardless of whether or not we have the data. Second, it is all the variables for which we have data. Third, it is the set of variables in the data which we end up using in the model.

### Exercise 9

What are the treatments, if any, for this problem?

```{r wisdom-9}
question_text(NULL,
  message = "The treatment variable is, conveniently, called `treatment`. Its value is `Treated` if the individual was exposed to Spanish-speakers on the train platform and `Control` if the individual was not.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Remember that a treatment is just another covariate which, for the purposes of this specific problem, we are assuming can be manipulated, thereby, creating two or more different potential outcomes for each unit.

### Exercise 10

What moment in time does the Preceptor Table refer to?

```{r wisdom-10}
question_text(NULL,
  message = "The Preceptor Table is about *now*, the period leading up to the election.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

A Preceptor Table can never really refer to an exact instant in time since nothing is instantaneous in this fallen world.

```{r}
trains |> 
  ggplot(aes(x = att_end, fill = treatment)) +
    geom_bar(aes(y = after_stat(count/sum(count))),
                   position = "dodge") +
    labs(title = "Ending Attitude Toward Immigration",
         subtitle = "Treated Individuals Are More Conservative",
         caption = "Source: Enos (2014)",
         x = "Attitude",
         y = "Probability",
         fill = NULL) +
    scale_y_continuous(labels = scales::percent_format()) + 
    theme_classic()

```

> *You can never look at the data too much. -- Mark Engerman*

### Exercise 11

Describe in words the Preceptor Table for this problem.

```{r wisdom-11}
question_text(NULL,
  message = "There is a row for each registered voter in our Georgia congressional district. There are two potential outcome columns — attitude toward immigration if exposed to Spanish-speakers, and attitude if not exposed. There is one treatment column indicating whether the voter was actually exposed.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

```{r, results = 'asis'}
hatch <- function(x) {
  paste0(
    '<span style="background-image: repeating-linear-gradient(',
    '45deg, rgba(0,0,0,0.12) 0 1px, transparent 1px 6px);',
    ' padding: 6px 10px; margin: -6px -10px; display: block;">',
    x,
    '</span>'
  )
}

p_tibble <- tibble::tribble(
  ~`Voter`         , ~`Attitude if Exposed`, ~`Attitude if Not Exposed`, ~`Spanish Exposure`,
  "James Carter"   , "10"                  , "8"                       , "Exposed"          ,
  "Maria Gonzalez" , "9"                   , "7"                       , "Not exposed"      ,
  "..."            , "..."                 , "..."                     , "..."              ,
  "Robert Kim"     , "11"                  , "9"                       , "Exposed"
)

pre_title_footnote     <- "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest. If all the information in this table were available, we could answer the question: What is the average causal effect of exposure to Spanish-speakers on attitudes toward immigration?"
pre_units_footnote     <- "Each row represents one registered voter in the Georgia congressional district during the current election cycle — a voter who could plausibly be exposed to Spanish-speaking campaign volunteers."
pre_outcome_footnote   <- "Attitude toward immigration on an integer scale from 3 to 15, where higher values indicate more conservative attitudes. The cross-hatched cells mark each voter's unobservable potential outcome — the one we could never see, because the voter actually experienced the other condition."
pre_treatment_footnote <- "Whether the voter was exposed to Spanish-speakers — here, campaign volunteers speaking Spanish in their vicinity — taking the value 'Exposed' or 'Not exposed'."
pre_effect_footnote    <- "The causal effect for Maria Gonzalez is the difference between her two potential outcomes: 9 − 7 = 2 points on the 3–15 scale. Because the Preceptor Table shows the unobservable truth, this is the true causal effect for her, not an estimate."

pre_gt_html <- gt::gt(p_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit"              , id = "unit_span",
                  columns = c(`Voter`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",
                  columns = c(`Attitude if Exposed`, `Attitude if Not Exposed`)) |>
  gt::tab_spanner(label = "Treatment"         , id = "treatment_span",
                  columns = c(`Spanish Exposure`)) |>
  gt::cols_align(align = "left" , columns = c(`Voter`, `Spanish Exposure`)) |>
  gt::cols_align(align = "right", columns = c(`Attitude if Exposed`, `Attitude if Not Exposed`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "center"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `Attitude if Not Exposed`, rows = c(1, 4)),
    fn        = function(x) hatch(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `Attitude if Exposed`, rows = 2),
    fn        = function(x) hatch(x)
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_treatment_footnote,
                   locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = pre_effect_footnote,
                   locations = gt::cells_body(columns = `Voter`, rows = 2)) |>
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 12

What is the narrow, specific question we will try to answer?

```{r wisdom-12}
question_text(NULL,
  message = "What is the average causal effect of exposure to Spanish-speakers on attitudes toward immigration?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The answer to this question is your Quantity of Interest. It is OK if your question differs from ours. Many similar questions lead to the creation of the same model.

### Exercise 13

Over the course of this tutorial, we will be creating a summary paragraph. Write its first two sentences below.

###

The first sentence is a general statement about the overall topic — mentioning the outcome and at least one covariate. It can be trite or even a platitude; its job is to gently orient the reader.

###

The second sentence introduces the data: when and where it was gathered, how many observations, who collected it, and what specific question we are asking.

```{r wisdom-13}
question_text(NULL,
  message = "Attitudes toward immigration are an increasingly important part of US congressional elections. Using data from a 2012 experiment among Boston commuters, we seek to explore the causal effect of exposure to Spanish speakers on attitudes toward immigration among registered voters in Georgia today.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Read our answer. It will not be the same as yours. You can, if you want, change your answer to incorporate some of our ideas. Do not copy/paste our answer exactly. Add your two sentences, edited or otherwise, to your QMD, `Cmd/Ctrl + Shift + K`, and then commit/push.

## Justice
###

Recall the Preceptor Table we built in Wisdom:

```{r, results = 'asis'}
hatch <- function(x) {
  paste0(
    '<span style="background-image: repeating-linear-gradient(',
    '45deg, rgba(0,0,0,0.12) 0 1px, transparent 1px 6px);',
    ' padding: 6px 10px; margin: -6px -10px; display: block;">',
    x,
    '</span>'
  )
}

p_tibble <- tibble::tribble(
  ~`Voter`         , ~`Attitude if Exposed`, ~`Attitude if Not Exposed`, ~`Spanish Exposure`,
  "James Carter"   , "10"                  , "8"                       , "Exposed"          ,
  "Maria Gonzalez" , "9"                   , "7"                       , "Not exposed"      ,
  "..."            , "..."                 , "..."                     , "..."              ,
  "Robert Kim"     , "11"                  , "9"                       , "Exposed"
)

pre_title_footnote     <- "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest. If all the information in this table were available, we could answer the question: What is the average causal effect of exposure to Spanish-speakers on attitudes toward immigration?"
pre_units_footnote     <- "Each row represents one registered voter in the Georgia congressional district during the current election cycle — a voter who could plausibly be exposed to Spanish-speaking campaign volunteers."
pre_outcome_footnote   <- "Attitude toward immigration on an integer scale from 3 to 15, where higher values indicate more conservative attitudes. The cross-hatched cells mark each voter's unobservable potential outcome — the one we could never see, because the voter actually experienced the other condition."
pre_treatment_footnote <- "Whether the voter was exposed to Spanish-speakers — here, campaign volunteers speaking Spanish in their vicinity — taking the value 'Exposed' or 'Not exposed'."
pre_effect_footnote    <- "The causal effect for Maria Gonzalez is the difference between her two potential outcomes: 9 − 7 = 2 points on the 3–15 scale. Because the Preceptor Table shows the unobservable truth, this is the true causal effect for her, not an estimate."

pre_gt_html <- gt::gt(p_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit"              , id = "unit_span",
                  columns = c(`Voter`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",
                  columns = c(`Attitude if Exposed`, `Attitude if Not Exposed`)) |>
  gt::tab_spanner(label = "Treatment"         , id = "treatment_span",
                  columns = c(`Spanish Exposure`)) |>
  gt::cols_align(align = "left" , columns = c(`Voter`, `Spanish Exposure`)) |>
  gt::cols_align(align = "right", columns = c(`Attitude if Exposed`, `Attitude if Not Exposed`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "center"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `Attitude if Not Exposed`, rows = c(1, 4)),
    fn        = function(x) hatch(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `Attitude if Exposed`, rows = 2),
    fn        = function(x) hatch(x)
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_treatment_footnote,
                   locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = pre_effect_footnote,
                   locations = gt::cells_body(columns = `Voter`, rows = 2)) |>
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

###

```{r, results = 'asis'}
data_tibble <- tibble::tribble(
  ~`Commuter`, ~`Attitude`, ~`Treatment`,
  "001"      , "10"       , "Treated"  ,
  "002"      , "8"        , "Control"  ,
  "..."      , "..."      , "..."      ,
  "115"      , "11"       , "Treated"
)

dat_title_footnote     <- "Enos, Ryan D. (2014). Causal effect of intergroup contact on exclusionary attitudes. Proceedings of the National Academy of Sciences, 111(10), 3699–3704. Data collected on Boston MBTA commuter rail platforms in spring 2012; 115 commuters randomized to treatment or control."
dat_outcome_footnote   <- "Immigration attitude on a scale from 3 (most liberal) to 15 (most conservative), measured after the two-week treatment window."
dat_treatment_footnote <- "Whether the subject rode on a platform alongside Spanish-speaking confederates ('Treated') or not ('Control'). Assignment was randomized."

dat_gt_html <- gt::gt(data_tibble, id = "data_tbl") |>
  gt::tab_header(title = "Data: Enos Boston MBTA Experiment, 2012") |>
  gt::tab_spanner(label = "Unit"     , id = "unit_span",
                  columns = c(`Commuter`)) |>
  gt::tab_spanner(label = "Outcome"  , id = "outcome_span",
                  columns = c(`Attitude`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::cols_align(align = "left" , columns = c(`Commuter`, `Treatment`)) |>
  gt::cols_align(align = "right", columns = c(`Attitude`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "center"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = dat_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = dat_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = dat_treatment_footnote,
                   locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::opt_css(
    css = "
      #data_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  dat_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

###

### Exercise 1

In your own words, name the five key components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
  message = "Justice concerns the Population Table and the four key assumptions which underlie it: validity, stability, representativeness, and unconfoundedness.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Justice is about concerns that you (or your critics) might have, reasons why the model you create might not work as well as you hope. 

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Validity is always about the *columns* in the Preceptor Table and the data. Just because columns from these two different tables have the same *name* does not mean that they are the same *thing*.

### Exercise 3

Provide one reason why the assumption of validity might not hold for the outcome variable `att_end` or for one of the covariates. Use the words "column" or "columns" in your answer.

```{r justice-3}
question_text(NULL,
  message = "The `att_end` column in the data records how commuters responded to a survey question about immigration attitudes. But the `att_end` column in the Preceptor Table represents their true, underlying attitudes — not necessarily what they would report on a survey. People do not always answer surveys honestly, so these two columns may not be measuring the same thing.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

In order to consider the Preceptor Table and the data to be drawn from the same population, the columns from one must have a *valid correspondence* with the columns in the other. Validity, if true (or at least reasonable), allows us to construct the Population Table, which is the first step in Justice.

Because we control the Preceptor Table and, to a lesser extent, the original question, we can adjust those variables to be “closer” to the data that we actually have. This is another example of the iterative nature of data science. If the data is not close enough to the question, then we check with our boss/colleague/customer to see if we can modify the question in order to make the match between the data and the Preceptor Table close enough for validity to hold.

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The Population Table is almost always much bigger than the combination of the Preceptor Table and the data because, if we can really assume that both the Preceptor Table and the data are part of the same population, then that population must cover a broad universe of time and units since the Preceptor Table and the data are, themselves, often far apart from each other.

### Exercise 5

Specify the unit/time combinations which define each row in this Population Table.

```{r justice-5}
question_text(NULL,
  message = "The unit/time combination is person/year. Of course, other time measures are reasonable. And, to be clear, in the Preceptor Table we are most concerned with the period right before the election, not with the entire calendar year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The exact time period used --- whether hour, day, month, year, or whatever --- is relatively arbitrary. The important thing to note is that the Population Table, unlike the Preceptor Table, covers a period of time over which things may change.

```{r}
trains |> 
  group_by(treatment) |> 
  summarise(
    N = n(),
    mean_att_end = mean(att_end, na.rm = TRUE),
    sd_att_end = sd(att_end, na.rm = TRUE)
  ) |> 
  gt() |> 
  tab_header(
    title = "Population Table: Immigration Attitudes by Treatment",
    subtitle = "Outcome: att_end (Attitudes toward Immigration)"
  ) |> 
  fmt_number(
    columns = c(mean_att_end, sd_att_end),
    decimals = 2
  )
```

### Exercise 6

In your own words, define the assumption of "stability" when employed in the context of data science.

```{r justice-6}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Stability is all about *time*. Is the relationship among the columns in the Population Table stable over time? The longer the time period between the data and the Preceptor Table, the more suspect stability becomes.

### Exercise 7

Provide one reason why the assumption of stability might not be true in this case.

```{r justice-7}
question_text(NULL,
  message = "The US is a much more Hispanic country today than it was in 2012, so being exposed to Spanish speakers might have a much smaller effect on immigration attitudes today.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

A change in time or the distribution of the data does not, in and of itself, demonstrate a violation of stability. Stability is about the parameters: $\beta_0$, $\beta_1$ and so on. Stability means these parameters are the same in the data as they are in the population as they are in the Preceptor Table.

### Exercise 8

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: `Data -> Population -> Preceptor Table`. In your own words, define the assumption of "representativeness" when employed in the context of data science.

```{r justice-8}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Ideally, we would like both the Preceptor Table *and* our data to be random samples from the population. Sadly, this is almost never the case.

### Exercise 9

We do not use the data, directly, to estimate missing values in the Preceptor Table. Instead, we use the data to learn about the overall population. Provide one reason, involving the relationship between the data and the population, for why the assumption of representativeness might not be true in this case. 

```{r justice-9}
question_text(NULL,
  message = "We only have data for Boston commuters. Our population includes people who do not commute. We might have a problem with representativeness if commuters are very different from non-commuters.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The reason that representativeness is important is because, when it is violated, the estimates for the model parameters might be biased. 

### Exercise 10

We use information about the population to make inferences about the Preceptor Table. Provide one reason, involving the relationship between the Population and the Preceptor Table, for why the assumption of representativeness might not be true in this case.

```{r justice-10}
question_text(NULL,
  message = "The Preceptor Table only includes registered voters in one Georgia congressional district. The population includes voters throughout the United States. Georgia is quite different from many other states.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Stability looks across time periods. Representativeness looks within time periods, for the most part. 

### Exercise 11

In your own words, define the assumption of "unconfoundedness" when employed in the context of data science.

```{r justice-11}
question_text(NULL,
  message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

This assumption is only relevant for causal models. We describe a model as "confounded" if this is not true. The easiest way to ensure unconfoundedness is to assign treatment randomly.

### Exercise 12

Provide one reason why the assumption of unconfoundedness might not be true (or relevant) in this case.

```{r justice-12}
question_text(NULL,
  message = "Although the exposure to Spanish-speakers is randomized, the difficulty of ensuring that this treatment was actually received might have caused problems. For example, what if some people are more likely to move away from strangers than others? In that case, those people, moving away from the experimenters, might never have heard any Spanish, even though they were officially 'assigned' to the treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The great advantage of randomized assignment of treatment is that it guarantees unconfoundedness, *if the randomization is done correctly*. There is no way for treatment assignment to be correlated with anything, including potential outcomes, if treatment assignment is random, and *if the experimental set up worked as designed.* Sadly, in the real world, there are sometimes problems.

### Exercise 13

A statistical model consists of two parts: the *probability family* and the *link function*. The probability family is the probability distribution which generates the randomness in our data. The link function is the mathematical formula which links our data to the unknown parameters in the probability distribution. 

Add `library(tidymodels)` to the QMD file.

Place your cursor in the QMD file on the `library(tidymodels)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(tidymodels)` to be copied down to the Console and then executed. 

CP/CR.

```{r justice-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The probability family is determined by the outcome variable $Y$. 

Since $Y$ is a continuous variable, the probability family is Normal, also known as Gaussian.

$$Y \sim N(\mu, \sigma^2)$$

### Exercise 14

Add `library(broom)` to the QMD file.

Place your cursor in the QMD file on the `library(broom)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(broom)` to be copied down to the Console and then executed. 

CP/CR.

```{r justice-14}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The link function, the basic mathematical structure of the model, is (mostly) determined by the type of outcome variable. 

For a continuous outcome variable, we use a linear model for the link function:

$$\mu = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \ldots$$

### Exercise 15

Write one sentence which highlights a potential weakness in your model. This will almost always be derived from possible problems with the assumptions discussed above. We will add this sentence to our summary paragraph. So far, our version of the summary paragraph looks like this:

> Attitudes toward immigration are an increasingly important part of US congressional elections. Using data from a 2012 experiment among Boston commuters, we seek to explore the causal effect of exposure to Spanish speakers on attitudes toward immigration among registered voters in Georgia today.

Of course, your version will be somewhat different.

```{r justice-15}
question_text(NULL,
  message = "The relationship between exposure to Spanish speakers and attitudes toward immigration may have changed since 2012.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Add a weakness sentence to the summary paragraph in your QMD. You can modify your paragraph as you see fit, but do not copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`, and then commit/push.


## Courage
### 

*Courage is found in unlikely places.* - J.R.R. Tolkien

Since our outcome variable $Y$ (`att_end`) is continuous, the probability family is Normal and the link function is linear. Together, the full abstract form of the model is:

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n + \epsilon$$

with $\epsilon \sim N(0, \sigma^2)$.

We use generic variables --- $Y$, $X_1$, and so on --- because our purpose is to describe the general mathematical structure of the model, independent of the specific variables we will eventually use.

###

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
  message = "Courage creates the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Having decided on the basic mathematical structure of the model at the end of *Justice*, a choice mostly driven by the distribution of our outcome variable, we now turn toward estimating the model.

### Exercise 2

Because our outcome variable is continuous, start to create the model by using `linear_reg(engine = "lm")`.

```{r courage-2, exercise = TRUE}

```

```{r courage-2-hint-1, eval = FALSE}
linear_reg(... = "lm")
```

```{r courage-2-test, include = FALSE}
linear_reg(engine = "lm")
```

### 

The *[tidymodels](https://www.tidymodels.org/)* framework is the most popular one in the R world for estimating models. *[Tidy Modeling with R](https://www.tmwr.org/)* by Max Kuhn and Julia Silge is a great introduction.

### Exercise 3

Continue the pipe to `fit(att_end ~ sex, data = trains)`.

```{r courage-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-3-hint-1, eval = FALSE}
... |> 
  fit(..., data = ...)
```

```{r courage-3-test, include = FALSE}
linear_reg(engine = "lm") |> 
  fit(att_end ~ sex, data = trains)
```

### 

Recall that a categorical variable (whether character or factor) like `sex` is turned into a $0/1$ "dummy" variable which is then re-named something like $sexMale$. After all, we can't have words --- like "Male" or "Female" --- in a mathematical formula, hence the need for dummy variables.

### Exercise 4

Continue the pipe with `tidy(conf.int = TRUE)`.

```{r courage-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-4-hint-1, eval = FALSE}
...
  tidy(... = TRUE)
```

```{r courage-4-test, include = FALSE}
linear_reg(engine = "lm") |> 
  fit(att_end ~ sex, data = trains) |> 
  tidy(conf.int = TRUE)
```

### 

The average `att_end` for women is around 9.7. For men, it is about 1 unit lower, meaning, perhaps surprisingly, less conservative about immigration.

### Exercise 5

Change the call to `fit()` to `fit(att_end ~ station, data = trains)`.

```{r courage-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-5-hint-1, eval = FALSE}
... |> 
  fit(..., data = ...)
  ...
```

```{r courage-5-test, include = FALSE}
linear_reg(engine = "lm") |> 
  fit(att_end ~ station, data = trains) |> 
  tidy(conf.int = TRUE)
```

### 

The same dummy variable approach applies to a categorical covariate with $N$ values. Such cases produce $N-1$ dummy $0/1$ variables. The presence of an intercept in most models means that we can't have $N$ categories. The "missing" category is incorporated into the intercept. Since `station` has nice values, the model creates 8 0/1 dummy variables, giving them names like $stationGrafton$ and $stationNatick$. The results for the *first* category are included in the intercept, which becomes the reference case, relative to which the other coefficients are applied.

### Exercise 6

Change the call to `fit()` to `fit(att_end ~ treatment, data = trains)`.

```{r courage-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-6-hint-1, eval = FALSE}
... |> 
  fit(..., data = ...)
  ...
```

```{r courage-6-test, include = FALSE}
linear_reg(engine = "lm") |> 
  fit(att_end ~ treatment, data = trains) |> 
  tidy(conf.int = TRUE)
```

### 

In a more advanced tutorial, we would include multiple independent variables in this regression. For now, however, we keep things simple.

### Exercise 7

Behind the scenes of this tutorial, an object called `fit_att` has been created which is the result of the code above. Type `fit_att` and hit "Run Code." This generates the same results as using `print(fit_att)`.

```{r courage-7, exercise = TRUE}

```

```{r courage-7-hint-1, eval = FALSE}
fit_att
```

```{r courage-7-test, include = FALSE}
fit_att
```

### 

In data science, **we deal with words, math, and code, but the most important of these is code.** We created the mathematical structure of the model and then wrote a model formula in order to estimate the unknown parameters. 

### Exercise 8

We need `fit_att` to exist in R World. Copy/paste this code into the Console and execute it.

````
fit_att <- linear_reg(engine = "lm") |> 
  fit(att_end ~ treatment, data = trains)
````

CP/CR.

```{r courage-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Just because something exists in the tutorial (or in the QMD) does not mean that it is in the Console. You should be aware of what exists in R World, which is generally called your "workspace."


### Exercise 9

In the Console, load the **[easystats](https://easystats.github.io/easystats/)** package. CP/CR.

```{r courage-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

We don't add **easystats** to the QMD because we are only using it for an interactive check of our fitted model. However, the [easystats ecosystem](https://easystats.github.io/easystats/) has a variety of interesting functions and packages which you might want to explore.

### Exercise 10

In the Console, run `check_predictions(extract_fit_engine(fit_att))`. CP/CR.

```{r courage-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

The purpose of `check_predictions()` is to compare your actual data (in green) with data that has been simulated from your fitted model, i.e., from your data generating mechanism. If your DGM is reasonable, then data simulated from it should not look too dissimilar from your actual data. Of course, it won't look exactly the same because of randomness, both in the world and in your simulation. But the actual data should be within the range of outcomes that your DGM simulates with `check_predictions()`.

### Exercise 11

Ask AI to create $\LaTeX$ code for this model, including our variable names and estimates for all the coefficients. Because this is a fitted model, the dependent variable will have a "hat" and the formula will not include an error term. 

Add the code to your QMD. `Cmd/Ctrl + Shift + K`.

Make sure the resulting display looks good. For example, you don't want an absurd number of figures to the right of the decimal. If the model is too long, you will need to spread it across several lines. You may need to go back-and-forth with the AI a few times.

Once the $\LaTeX$ code looks good, paste it below.

```{r courage-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our formula looks like:

$$\widehat{\text{att_end}} = 8.45 + 1.55 \cdot \text{treatment}_{\text{Treated}}$$

It was created with:

````
$$\widehat{\text{att_end}} = 8.45 + 1.55 \cdot \text{treatment}_{\text{Treated}}$$
```` 

Note the differences. First, we’ve replaced the parameters with our best estimates. Second, we’ve dropped the error term because this formula is used to predict the outcome. Third, the left-hand side is $\widehat{\text{att_end}}$ instead of $\text{att_end}$ because it represents an estimated value—indicated by the “hat.”

**This is our data generating mechanism.**

Of course, there is randomness built into the DGM, but we won't worry about that detail for now.

A data generating mechanism is just a formula, something which we can write down and implement with computer code.

### Exercise 12

Create a new code chunk in your QMD. Add a code chunk option: `#| cache: true`. Copy/paste the code from above for estimating the model into the code chunk, assigning the result to `fit_att`. (This will include the call to `fit()` but not the call to `tidy()` because we want the entire fitted model, not just a table of the estimated parameter values.)

Place your cursor in the QMD file on the `fit_att` line. Use `Cmd/Ctrl + Enter` to execute that line. Strictly speaking, this step is unnecessary because we already added `fit_att` to our workspace above. But ensuring that everything in the QMD is also in the Console is a good habit.

`Cmd/Ctrl + Shift + K`. It may take some time to render your QMD, depending on how complex your model is. But, by including `#| cache: true` you cause Quarto to cache the results of the chunk. The next time you render your QMD, as long as you have not changed the code, Quarto will just load up the saved fitted object.

At the Console, run:

```
tutorial.helpers::show_file("trains.qmd", chunk = "Last")
```

CP/CR.

```{r courage-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

### 

To confirm, `Cmd/Ctrl + Shift + K` again. It should be quick.

### Exercise 13

Add `*_cache` to `.gitignore` file. Cached objects are often large. They don't belong on Github.

At the Console, run:

```
tutorial.helpers::show_file(".gitignore")
```

CP/CR.

```{r courage-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Because of the change in your `.gitignore` (assuming that you saved it), the cache directory should not appear in the Source Control panel because Git is ignoring it, as instructed. Commit and push. 

### Exercise 14

In the Console, run `tidy()` on `fit_att` with the argument `conf.int` set equal to `TRUE`. This returns 95% intervals for all the parameters in our model.

```{r courage-14, exercise = TRUE}

```

```{r courage-14-hint-1, eval = FALSE}
tidy(..., conf.int = ...)
```

```{r courage-14-test, include = FALSE}
tidy(fit_att, conf.int = TRUE)
```

### 

`tidy()` is part of the [**broom**](https://broom.tidymodels.org/) package, used to summarize information from a wide variety of models.

### Exercise 15

Create a new code chunk in your QMD. Ask AI to help you make a nice looking table from the tibble which is returned by `tidy()`. You don't have to include all the variables which `tidy()` produces. We often just show the estimate and the confidence intervals.

You might need to add some new libraries, e.g., **[tinytable](https://vincentarelbundock.github.io/tinytable/)**, **[knitr](https://yihui.org/knitr/)**, **[gt](https://gt.rstudio.com/)**, **[kableExtra](https://haozhu233.github.io/kableExtra/)**, **[flextable](https://davidgohel.github.io/flextable/)**, **[modelsummary](https://modelsummary.com/)**, et cetera, to the `setup` code chunk, if you use any functions from these packages, all of which have strengths and weaknesses for making tables.

Insert your table code into the QMD. (If you need a new library, and you probably will, place it in the `setup` code chunk along with the other libraries you have already loaded.)

`Cmd/Ctrl + Shift + K`. 

At the Console, run:

```
tutorial.helpers::show_file("trains.qmd", chunk = "Last")
```

CP/CR.

```{r courage-15}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

Here is ours:

```{r}
library(knitr)

tidy(fit_att, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  kable(
    caption = "Estimates and 95% Confidence Intervals for Parameters\nSource: Enos (2014)"
  )
```

Generated by this code:

````
library(knitr)

tidy(fit_att, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  kable(
    caption = "Estimates and 95% Confidence Intervals for Parameters\nSource: Enos (2014)"
  )
````

At the very least, your table should include a title and a caption with the data source. The more you use AI, the better you will get at doing so.

### Exercise 16

Add a sentence to your project summary. 

Explain the structure of the model. Something like: "I/we model XX [the concept of the outcome, not the variable name], [insert description of values of XX], as a [linear/logistic/multinomial/ordinal] function of XX [and maybe other covariates]." 


Recall the beginning of our version of the summary:

> Attitudes toward immigration are an increasingly important part of US congressional elections. Using data from a 2012 experiment among Boston commuters, we seek to explore the causal effect of exposure to Spanish speakers on attitudes toward immigration among registered voters in Georgia today. The relationship between exposure to Spanish speakers and attitudes toward immigration may have changed since 2012.

```{r courage-16}
question_text(NULL,
  message = "We model attitude toward immigration, an integer value ranging from 3 to 15, as a linear function of exposure to Spanish-speakers.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Read our answer. It will not be the same as yours. You can, if you want, change your answer to incorporate some of our ideas. Do not copy/paste our answer exactly. Add your two sentences, edited or otherwise, to the summary paragraph portion of your QMD. `Cmd/Ctrl + Shift + K`, and then commit/push.

## Temperance
### 

*Temperance is a tree which has for its root very little contentment, and for its fruit calm and peace.* - Buddha

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
  message = "Temperance uses the data generating mechanism to answer the question with which we began. Humility reminds us that this answer is always a lie. We can also use the DGM to calculate many similar quantities of interest, displaying the results graphically.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Courage gave us the data generating mechanism. Temperance guides us in the use of the DGM — or the “model” — we have created to answer the question(s) with which we began. We create posteriors for the quantities of interest. 

### Exercise 2

Before using the DGM, we should make sure that we can interpret it.

Recall the values for the parameters in our data generating mechanism:

```{r}
fit_att |> 
    tidy(conf.int = TRUE) |> 
  select(term, estimate, conf.low, conf.high)
```

Interpret the 1.55 estimate for `treatmentTreated`.

```{r temperance-2}
question_text(NULL,
  message = "Since this is a causal model, then the average causal effect of receiving the treatment of hearing Spanish-speakers on the train platform, relative to the control, is to have a 1.55 higher value for att_end, meaning that the treatment, relative to the control, makes one more conservative about immigration.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The interpretation of a treatment variable is very different from the interpretation of a standard covariate. The key point is that there is no such thing as a causal (versus predictive) data set nor a causal (versus predictive) R code formula. You can use the same data set (and the same R code!) for both causal and predictive models. The difference lies in the assumptions you make.

### Exercise 3

```{r}
fit_att |> 
  tidy(conf.int = TRUE) |> 
  select(term, estimate, conf.low, conf.high)
```

Interpret the 0.51 to 2.58 confidence interval for the `treatmentTreated` coefficient.

```{r temperance-3}
question_text(NULL,
  message = "Our best estimate for the average causal effect is 1.55, meaning that being treated makes someone 1.55 units more conservative about immigration. Yet, the true value could be lower or higher. We are 95% certain that the true effect is somewhere between 0.51 to 2.58.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

We care if the confidence interval for a given variable excludes zero. If not, then we can't be sure if the relationship between the variable and the outcome is positive or negative. In this case, we can be fairly certain that the causal effect of treatment is positive.

### Exercise 4

```{r}
fit_att |> 
  tidy(conf.int = TRUE) |> 
  select(term, estimate, conf.low, conf.high)
```

Interpret the estimate of 8.45 for the `(Intercept)` term.

```{r temperance-4}
question_text(NULL,
  message = "The intercept of 8.45 represents the average immigration attitude (`att_end`) for people in the control group—those who were not exposed to Spanish speakers. Since treatment is a binary variable, this is the expected outcome when `treatment = 0`.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The intercept shows the baseline outcome for the control group. It helps put the treatment effect in context by telling us where we started.

### Exercise 5

In the end, we don't really care about parameters, much less how to interpret them. Parameters are *imaginary*, like unicorns. We care about answers to our questions. Parameters are tools for answering questions. They aren't interesting in-and-of themselves. *In the modern world, all parameters are nuisance parameters.* 

Add `library(marginaleffects)` to the QMD file.

Place your cursor in the QMD file on the `library(marginaleffects)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(marginaleffects)` to be copied down to the Console and then executed. 

CP/CR.

```{r temperance-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

We should be modest in the claims we make. The posteriors we create are never the “truth.” The assumptions we made to create the model are never perfect. Yet decisions made with flawed posteriors are almost always better than decisions made without them.

### Exercise 6

What is the specific question we are trying to answer? 

```{r temperance-6}
question_text(NULL,
  message = "What is the average causal effect of exposure to Spanish-speakers on attitudes toward immigration?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Data science projects begin with a decision which we face. To make that decision wisely, we would like to have good estimates of many unknown numbers. Yet, in order to make progress, we need to drill down to one specific question. This leads to the creation of a data generating mechanism, which can now be used to answer lots of questions, thus allowing us to explore the original decision more broadly.

### Exercise 7

In the Console, run `predictions()` on `fit_att`.

CP/CR.

```{r temperance-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

```{r temperance-7-test, include = FALSE}
predictions(fit_att)
```

###

Because this is a linear model, `predictions()` produces one prediction for each data point in `trains`. But the fitted model only includes one variable, `treatment`, which takes on only two possible values. So, there are only two predictions: one for "Treated" (10) and one for "Control" (8.45).

### Exercise 8

In the Console, run `avg_predictions()` on `fit_att` with `by = "treatment"`.

CP/CR.

```{r temperance-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

```{r temperance-8-test, include = FALSE}
avg_predictions(fit_att, by = "treatment")
```

### 

This shows the average prediction for "Treated" and "Control" separately.

### Exercise 9

In the Console, run `plot_predictions()` on `fit_att` with `by = "treatment"`.

CP/CR.

```{r temperance-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

```{r temperance-9-test, include = FALSE}
plot_predictions(fit_att, by = "treatment")
```

### 

This is interesting, but our primary focus is not the outcomes for "Treated" and "Control" but the expected difference in those outcomes. That is the causal effect.

### Exercise 10

In the Console, run `avg_comparisons()` on `fit_att`.

CP/CR.

```{r temperance-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

```{r temperance-10-test, include = FALSE}
avg_comparisons(fit_att)
```

### 

This produces the number that we really care about --- the average causal effect of treatment --- along with a confidence interval for that number.

### Exercise 11

Work with AI to create a beautiful plot starting with the output to the above call to `plot_predictions()`. Do this in your QMD since that is much easier than working in the Console directly.

Your title should highlight the key variables. Your subtitle should describe an important takeaway, the sentence/conclusion which readers will, you hope, remember. Your caption should mention the data source. Your axis labels should look nice.

This plot is not directly connected to your question. It answers lots of questions! It might be used by lots of different people. 

Copy the code for your plot here:

```{r temperance-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

### 

Our plot:

```{r}
avg_comparisons(fit_att) |>
 ggplot(aes(x = 1, y = estimate)) +
 geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                 size = 1.2,
                 linewidth = 1.5,
                 color = "#2E86AB") +
 scale_y_continuous(breaks = seq(0, 3, 0.5), limits = c(0, 3)) +
 labs(
   title = "Causal Effect on Attitudes Toward Immigration",
   subtitle = "Hearing Spanish-speakers makes people more conservative about immigration",
   caption = "Source: Enos (2012)",
   y = "Effect on 3 to 15 Scale",
   x = NULL
 ) +
 theme_minimal() +
 theme(
   axis.text.x = element_blank(),
   axis.ticks.x = element_blank()
 )
```

Our code:

````{verbatim}
avg_comparisons(fit_att) |>
 ggplot(aes(x = 1, y = estimate)) +
 geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                 size = 1.2,
                 linewidth = 1.5,
                 color = "#2E86AB") +
 scale_y_continuous(breaks = seq(0, 3, 0.5), limits = c(0, 3)) +
 labs(
   title = "Causal Effect on Attitudes Toward Immigration",
   subtitle = "Hearing Spanish-speakers makes people more conservative about immigration",
   caption = "Source: Enos (2012)",
   y = "Effect on 3 to 15 Scale",
   x = NULL
 ) +
 theme_minimal() +
 theme(
   axis.text.x = element_blank(),
   axis.ticks.x = element_blank()
 )
````

Data science often involves this-back-and-forth style of work. First, we need to make a single chunk of code, in this case, a new plot, work well. This requires interactive work between the QMD and the Console. Second, we need to ensure that the entire QMD runs correctly on its own.

### Exercise 12

Finalize the new graphics code chunk in your QMD. `Cmd/Ctrl + Shift + K` to ensure that it all works as intended. 

At the Console, run:

```
tutorial.helpers::show_file("trains.qmd", chunk = "Last")
```

CP/CR.

```{r temperance-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```


### 

Always [remember](https://en.wikipedia.org/wiki/Map%E2%80%93territory_relation): *The map is not the territory.* A beautiful graphic tells a story, but that story is always an imperfect representation of reality. Our models depend on assumptions, assumptions which are never completely true.

### Exercise 13

Write the last sentence of your summary paragraph. It describes at least one quantity of interest (QoI) and provides a measure of uncertainty about that QoI. It is OK if this QoI is not the one that you began with. The focus of a data science project often changes over time. It is also OK to discuss more than one QoI. Do whatever seems reasonable, given the context.

```{r temperance-13}
question_text(NULL,
  message = "Briefly exposing voters to Spanish speakers makes them more conservative on immigration, by about 1.5 (plus/minus 1), which is about 2/3s of the difference between Democrats and Republicans.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Add a final sentence to your summary paragraph in your QMD as you see fit, but do not copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`.

### Exercise 14

Write a few sentences which explain why the estimates for the quantities of interest, and the uncertainty thereof, might be wrong. Suggest an alternative estimate and confidence interval, if you think either might be warranted.

```{r temperance-14}
question_text(NULL,
  message = "The estimates may be misleading because we’re modeling immigration attitudes, not actual voting behavior. Without data linking the two, our inferences could be biased or overconfident. A better approach would model voting directly or use wider confidence intervals to reflect this uncertainty.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Always go back to your Preceptor Table, the information which, if you had it, would make answering your question easy. In almost all real world cases, the Preceptor Table and the data are fairly different, not least because validity never holds perfectly. So, even a perfectly estimated statistical model is rarely as useful as we might like.

### Exercise 15

Rearrange the material in your QMD so that the order is graphic, followed by the paragraph. Doing so, of course, requires sensible judgment. For example, the code chunk which creates the fitted model must occur before the chunk which creates the graphic. You can keep or discard the math and any other material at your own discretion.

`Cmd/Ctrl + Shift + K` to ensure that everything works.

At the Console, run:

```
tutorial.helpers::show_file("trains.qmd")
```

CP/CR.

```{r temperance-15}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 30)
```

### 

This is the version of your QMD file at which your teacher is most likely to look closely.

### Exercise 16

Publish your rendered QMD to GitHub Pages. In the Terminal --- not the Console! --- run:

````
quarto publish gh-pages trains.qmd
````

Copy/paste the resulting URL below.

```{r temperance-16}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

Commit/push everything.

### Exercise 17

Copy/paste the URL to your Github repo.

```{r temperance-17}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God. If all our assumptions are correct, then our DGM is true, it accurately describes the way in which the world works. There is no better way to predict the future, or to model the past, than to use it. Sadly, this will only be the case with toy examples involving things like coins and dice. We hope that our DGM is close to the true DGM but, since our assumptions are never perfectly correct, our DGM will always be different. The estimated magnitude and importance of that difference is a matter of judgment.

The world confronts us. Make decisions we must.

## Summary
### 

This tutorial supports [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). 

```{r}
avg_comparisons(fit_att) |>
 ggplot(aes(x = 1, y = estimate)) +
 geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                 size = 1.2,
                 linewidth = 1.5,
                 color = "#2E86AB") +
 scale_y_continuous(breaks = seq(0, 3, 0.5), limits = c(0, 3)) +
 labs(
   title = "Causal Effect on Attitudes Toward Immigration",
   subtitle = "Hearing Spanish-speakers makes people more conservative about immigration",
   caption = "Source: Enos (2012)",
   y = "Effect on 3 to 15 Scale",
   x = NULL
 ) +
 theme_minimal() +
 theme(
   axis.text.x = element_blank(),
   axis.ticks.x = element_blank()
 )
```

Attitudes toward immigration are an increasingly important part of US congressional elections. Using data from a 2012 experiment among Boston commuters, we seek to explore the causal effect of exposure to Spanish speakers on attitudes toward immigration among registered voters in Georgia today. The relationship between exposure to Spanish speakers and attitudes toward immigration may have changed since 2012. We model attitude toward immigration, an integer value ranging from 3 to 15, as a linear function of exposure to Spanish-speakers. Briefly exposing voters to Spanish speakers makes them more conservative on immigration, by about 1.5 (plus/minus 1), which is about 2/3s of the difference between Democrats and Republicans.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
