---
title: "Wisdom and Recruits (Answer Key)"
format:
  html: default
  pdf: default
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(tidymodels)
library(marginaleffects)
library(gt)
library(primer.tutorials)
```


## Wisdom

You are in charge of ordering uniforms for next year's Marine Corps bootcamp recruits.


What will be the average height of male and female USMC recruits next year?


**What are the units, outcome, and covariates for this question?**

- Units: 

One recruit

- Outcomes:

 `height` in centimeters

- Covariates: 

`sex`

- Treatment (if any): 

None

- Quantity of interest: 

The average height for male and female recruits, and the difference between them.


Describe the key components of a Preceptor table for this problem; Use words like "units," "outcomes," and "covariates."

  The Preceptor Table should have one row per recruit. In this exercise, the columns are unit, outcome, and covariates.

```{r}
#| echo: false
#| results: asis

preceptor_tibble <- tibble::tribble(
  ~`Recruit`            , ~`Height (cm)`, ~`Sex`   ,
  "Emma Rodriguez"      , "166"         , "Female",
  "Marcus Johnson"      , "180"         , "Male"  ,
  "..."                 , "..."         , "..."   ,
  "David Chen"          , "177"         , "Male"
)

pre_title_footnote <- "If all the information in this table were available, we could answer the question: What is the average height of male and female USMC recruits next year?"

pre_units_footnote <- "Each row represents one recruit entering USMC bootcamp in 2026. The total population of annual recruits is roughly 35,000–40,000 (about 10% female, 90% male by historical rates), though screening policies and self-selection may shift these proportions."

pre_outcome_footnote <- "Height measured in centimeters from the soles of the feet to the top of the head, taken at standard military induction physical examination. This is a continuous measurement with no artificial rounding or truncation in our analysis."

pre_covariate_footnote <- "Sex as recorded on military identification: Male or Female. This is the trait the question uses to partition recruits into two comparison groups."

pre_gt_html <- gt::gt(preceptor_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit"     , id = "unit_span",
    columns = c(`Recruit`)) |>
  gt::tab_spanner(label = "Outcome"  , id = "outcome_span",
    columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
    columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
    locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
    locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
    locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote,
    locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```


Create a plot that shows the relationship between height and sex.

```{r}
#| echo: true
sex_counts <- recruits |> count(sex)

recruits |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(aes(y = after_stat(count)), alpha = 0.6) +
  labs(
    title = "Height Distribution by Sex",
    subtitle = paste0(
      "Area is scaled by sample size: n = ", sex_counts$n[sex_counts$sex == "Male"],
      " male vs. n = ", sex_counts$n[sex_counts$sex == "Female"], " female recruits"
    ),
    x = "Height (cm)",
    y = "Count-scaled density",
    fill = "Sex"
  ) +
  theme_minimal()
```

Male recruits outnumber female recruits four to one in this sample, so the male curve covers roughly four times as much area as the female curve. That's a feature of the sample size, not evidence that male and female heights are equally common or equally variable — it's worth noticing before fitting any model.

Create a plot that shows the relationship between height by age.

```{r}
#| echo: true
age_fits <- recruits |>
  group_by(sex) |>
  summarize(slope = coef(lm(height ~ age))[["age"]], .groups = "drop")

recruits |>
  ggplot(aes(x = age, y = height, color = sex)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Height vs. Age by Sex",
    subtitle = paste0(
      "Slope is ", round(age_fits$slope[age_fits$sex == "Female"], 2),
      " cm/yr for females and ", round(age_fits$slope[age_fits$sex == "Male"], 2),
      " cm/yr for males — both near zero, as expected for adults"
    ),
    x = "Age (years)",
    y = "Height (cm)",
    color = "Sex"
  ) +
  theme_minimal()
```

Among adults aged 18-27, height has mostly stopped changing, so neither line should show a strong trend. Whatever slope appears is sampling noise, not a real age effect — and there is no reason the two sexes' noise should point in the same direction, so the two slopes are computed and reported separately rather than assumed.


---

## Justice

Describe the key components of a Population table for this problem. 


```{r}
#| echo: false
#| results: asis

population_tibble <- tibble::tribble(
  ~Source    , ~`Recruit`        , ~`Year`, ~`Height (cm)`, ~`Sex`   ,
  "..."      , "..."             , "..."  , "..."          , "..."   ,
  "Data"     , "Emma Rodriguez"  , "2005" , "165"          , "Female",
  "Data"     , "Marcus Johnson"  , "2005" , "179"          , "Male"  ,
  "Data"     , "..."             , "..."  , "..."          , "..."   ,
  "Data"     , "David Chen"      , "2005" , "176"          , "Male"  ,
  "..."      , "..."             , "..."  , "..."          , "..."   ,
  "Preceptor", "Emma Rodriguez"  , "2026" , "166"          , "Female",
  "Preceptor", "Marcus Johnson"  , "2026" , "180"          , "Male"  ,
  "Preceptor", "..."             , "..."  , "..."          , "..."   ,
  "Preceptor", "David Chen"      , "2026" , "177"          , "Male"  ,
  "..."      , "..."             , "..."  , "..."          , "..."
)

pop_title_footnote <- "This table bridges from NHANES data (2003–2006) to next year's USMC recruits. Both come from the population of young adults aged 18–27 living in the United States."

pop_units_footnote <- "Each row is a young adult at a specific time. Data rows are 50 participants in the National Health and Nutrition Examination Survey (NHANES), a cross-sectional sample of the US non-institutionalized civilian population, surveyed between 2003 and 2006. Preceptor rows are recruits entering USMC bootcamp in 2026, drawn from volunteers aged 18–27. The 20-year gap between the data era and the prediction year, and the shift from a population sample to military volunteers, both raise representativeness concerns."

pop_outcome_footnote <- "Height in centimeters, measured from sole to crown. NHANES used standardized protocols. Preceptor rows show the (unobservable) height each recruit will have in 2026, which we model by assuming NHANES heights remain stable across the population."

pop_covariate_footnote <- "Sex as recorded in the two data sources. NHANES sex is self-reported or interviewer-recorded at survey time. USMC recruits' sex comes from military identification at induction. Both are treated as a binary Male/Female classification."

pop_gt_html <- gt::gt(population_tibble, id = "population_tbl") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time"  , id = "unit_span",
    columns = c(`Recruit`, `Year`)) |>
  gt::tab_spanner(label = "Outcome"    , id = "outcome_span",
    columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate"  , id = "covariates_span",
    columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Source`, `Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote,
    locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote,
    locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote,
    locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote,
    locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```


Counter-example to the assumption of validity:

  The height column in NHANES is measured differently than the height column in the Preceptor Table, so the two columns may not represent exactly the same construct.

Counter-example to the assumption of stability:

  The relationship between height and sex may have changed over time, so the model parameters (such as the intercept or the slope on sexMale) may not be the same for the data and the Preceptor Table.

Counter-example to the assumption of representativeness:

  The NHANES sample is not representative of future USMC recruits because recruits are self-selected and screened for military service.

Counter-example to the assumption of unconfoundedness:

  Sex is not a manipulable treatment, and treatment assignment is not independent of the potential outcomes, so the unconfoundedness assumption fails.

What probability family should we use?

  Normal Distribution

Why should we use that probability family?

  The Normal distribution is the standard choice for continuous outcomes.

What is the link function for our outcome?

  Our model will be a linear regression.

## Courage

Create a candidate model, modeling the outcome height as a linear function of the binary covariate sex.

```{r}
fit_recruits <- linear_reg(engine = "lm") |>
  fit(height ~ sex, data = recruits)
```

Create another candidate model, modeling the outcome height as a linear function of the continuous covariate age. 

```{r courage-5-test}
fit_age <- linear_reg(engine = "lm") |>
  fit(height ~ age, data = recruits)
```


## Temperance

Use `predictions(fit_recruits)` to estimate the average height of male and female USMC recruits next year.

```{r}
predictions(fit_recruits)
```

What is the expected height of a female recruit?

  The expected height of a female recruit is about **165.9 cm**, with a 95% confidence interval of approximately **161.8 cm to 170.0 cm**.

What is the expected height of a male recruit?

  The expected height of a male recruit is about **174.8 cm**, with a 95% confidence interval of approximately **172.5 cm to 177.0 cm**.

Create a visualization of these predictions using `plot_predictions()`.

```{r}
#| echo: true
avg_preds <- avg_predictions(fit_recruits, variables = "sex")
taller_sex <- avg_preds$sex[which.max(avg_preds$estimate)]
height_diff <- round(diff(range(avg_preds$estimate)), 1)

plot_predictions(
  fit_recruits,
  condition = "sex",
  draw = FALSE
) |>
  ggplot(aes(x = sex, y = estimate)) +
  geom_pointrange(
    aes(
      ymin = conf.low,
      ymax = conf.high
    ),
    linewidth = 0.7
  ) +
  labs(
    title = "Predicted Average Height of USMC Recruits by Sex",
    subtitle = paste0(taller_sex, " recruits are predicted to be about ", height_diff, " cm taller, on average"),
    x = "Sex",
    y = "Predicted Height (cm)",
    caption = "Source: 50-row teaching sample from NHANES via primer.tutorials::recruits"
  ) +
  theme_minimal(base_size = 13)
```

This plot turns the regression into a direct answer for the Quartermaster: predicted heights with their precision, not just two numbers in a table.

What measure of uncertainty is shown?

  The vertical error bars show **95% confidence intervals** for the estimated average height of each group.

Why is the confidence interval wider for female recruits?

  There are only **10 female observations** compared with **40 male observations**, so the female average is estimated with less precision.

Besides the average height, identify two additional quantities of interest the Quartermaster might care about.

- The **tallest recruit**, to ensure the largest uniform size is available.
- The **10th and 90th percentiles** of recruit height, to estimate how many extra-small and extra-large uniforms should be ordered.

Why might these estimated average heights be wrong?

  The estimates rely on NHANES data rather than actual USMC recruits. Because recruits are self-selected, screened for military service, and likely taller on average than the general young-adult population, the model probably understates recruit height. In addition, the reported confidence intervals reflect only sampling uncertainty and do not account for possible violations of validity, stability, or representativeness.
