---
title: "Wisdom and Recruits (Answer Key)"
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(primer.tutorials)
```



## Wisdom

What will be the average height of male and female USMC recruits next year?

**What are the units, outcome, and covariates in this question?**

- Unit: 

One recruit

- Outcome:

 `height` in centimeters

- Covariates: 

`sex`

- Treatment (if any): 

None

- Quantity of interest: 

The average height for male and female recruits, and the difference between them.



Describe the key components of a Preceptor table for this problem; Use words like "units," "outcomes," and "covariates."

The Preceptor Table should have one row per recruit. In this exercise, the columns are unit, outcome, and covariates.




```{r}
#| echo: false
#| results: asis
preceptor_tibble <- tibble::tribble(
  ~Unit, ~Outcome, ~Covariate,
  "Recruit 1", "150", "Female",
  "Recruit 2", "172", "Male",
  "...", "...", "...",
  "Recruit 5,000", "160", "Female"
)

preceptor_html <- gt::gt(preceptor_tibble, id = "preceptor_table") |>
  gt::tab_header(title = "Completed Preceptor Table") |>
  gt::tab_spanner(label = "Unit", id = "unit_span", columns = "Unit") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Unit, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "Each row is one recruit whose height we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariates are sex and age, known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    preceptor_html, "</div>\n```", sep = "")
```

## Justice

Describe the key components of a Population table for this problem. 


```{r}
#| echo: false
#| results: asis
population_tibble <- tibble::tribble(
  ~Source, ~`Unit / Time`, ~Outcome, ~Covariate,
  "Data", "NHANES 1 (2002)", "162", "Female",
  "Data", "NHANES 2 (2005)", "178", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 1 (2026)", "150", "Female",
  "Preceptor", "Recruit 2 (2026)", "172", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 5,000 (2026)", "160", "Female"
)

population_html <- gt::gt(population_tibble, id = "population_table") |>
  gt::tab_header(title = "Completed Population Table") |>
  gt::tab_spanner(label = "Unit / Time", id = "unit_span", columns = "Unit / Time") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Source, `Unit / Time`, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "The data rows are the observed sample; the preceptor rows are the recruits we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariate is sex, which is known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    population_html, "</div>\n```", sep = "")
```


Each row is one young adult observed at one moment in time, and the table connects the observed data to the recruits we want to forecast. The rows labeled "Data" come from the sample we observed; the rows labeled "Preceptor" represent the future recruits we want to predict.



Counter-example to the assumption of validity:



Counter-example to the assumption of stability:



Counter-example to the assumption of representativeness:



Counter-example to the assumption of unconfoundedness:



## Courage

What outcome are we modeling? What predictors will we use?

The outcome is `height`, a continuous numeric variable. We will model it with a Gaussian linear model and an identity link. A linear model is appropriate because height is continuous and the mean height can be described as a linear function of the predictors.

Why is this model reasonable?

A linear model is reasonable because the mean height can be expressed as a linear function of the predictors, and the outcome is continuous rather than categorical.


```{r}
fit_recruits <- lm(height ~ sex, data = recruits)
summary(fit_recruits)
```

## Exploratory Data Analysis Exercises


### Exercise 1: Visualize height by sex

Use AI to create a plot that shows the distribution of height separately for male and female recruits. Include appropriate labels and a title. What do you notice about the two distributions?

**Sample output:**

```{r}
recruits |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(alpha = 0.6) +
  labs(
    title = "Height Distribution by Sex",
    x = "Height (cm)",
    y = "Density",
    fill = "Sex"
  ) +
  theme_minimal()
```

**What students should notice:** The male distribution is clearly shifted to the right (taller) compared to the female distribution. There is overlap but clear separation between the two groups.

**Follow-up question for students:** If you were a logistics analyst ordering uniforms, why would it be important to break down the height distribution by sex rather than treating all recruits as one group?

---

### Exercise 2: Summary statistics by sex

Use AI to calculate the mean, standard deviation, and count of recruits for each sex. Organize the output in a clean table. Why might the difference in sample sizes (40 vs. 10) matter for our model?

**Sample output:**

```{r}
recruits |>
  group_by(sex) |>
  summarise(
    N = n(),
    Mean_Height = mean(height),
    SD_Height = sd(height),
    .groups = "drop"
  )
```

```
# A tibble: 2 × 4
  sex        N Mean_Height SD_Height
  <fct>    <int>       <dbl>     <dbl>
1 Female      10        162.      4.67
2 Male        40        175.      6.45
```

**What students should notice:** Males are on average 13 cm taller. The female group is much smaller (10 vs. 40), so estimates for female recruits will have more uncertainty.

**Follow-up question for students:** Which group's estimate (male or female average height) will have the larger standard error? Why?

**Teaching fellow prompt:** This is a key insight for Temperance --- the unequal sample sizes mean the confidence intervals will have different widths. This deliberate imbalance in the data is there to make students notice and reason about precision.

---

### Exercise 3: Age and height relationship

Use AI to create a scatter plot showing the relationship between age and height, colored by sex. Does age appear to be related to height in this sample?

**Sample output:**

```{r}
recruits |>
  ggplot(aes(x = age, y = height, color = sex)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Height vs. Age by Sex",
    x = "Age (years)",
    y = "Height (cm)",
    color = "Sex"
  ) +
  theme_minimal()
```

**What students should notice:** Age shows almost no relationship to height in this sample (the lines are nearly flat). This suggests that sex is a much stronger predictor of height than age in this age range (18--27). The sex difference is consistent across ages.

**Follow-up question for students:** Given this plot, do you think the model `height ~ sex` will be more predictive than `height ~ age`? Why?

---

## Temperance


What is the main conclusion from the fitted model?

Male recruits are taller on average than female recruits. The difference is about 13 cm, with male recruits averaging around 175 cm and female recruits averaging around 162 cm.

Ask AI to create a plot that summarizes the fitted model and the data. Use AI for the coding, and then write a short Quarto document that includes the plot and a brief interpretation. Publish the result to GitHub Pages.

```{r}
recruits |>
  ggplot(aes(x = sex, y = height)) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  geom_point(
    data = recruits |>
      group_by(sex) |>
      summarise(mean_height = mean(height), .groups = "drop"),
    aes(y = mean_height),
    color = "steelblue",
    size = 3
  ) +
  labs(
    title = "Observed heights and average height by sex",
    x = "Sex",
    y = "Height (cm)"
  ) +
  theme_minimal()
```
