---
title: "Wisdom and Recruits (Answer Key)"
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(gt)
library(primer.tutorials)
```

## Imagine

Imagine that you are a logistics analyst at U.S. Marine Corps Supply, planning next year's bootcamp uniform order. You want to use the available data to make a sensible forecast for the recruits who will arrive next year.

The data describe a sample of recruits with information such as height, sex, and age. Each row is one recruit, and the goal is to use what is known before the forecast to reason about what will happen next year.

## Wisdom

We are working with the `recruits` dataset. This exercise is a predictive modeling exercise. We will use it to reason about how to forecast the height of next year's recruits from information we already know about them.

What question might you try to answer if you are a logistics analyst planning next year's uniform order?

We are trying to forecast the average height of next year's Marine Corps recruits from information such as sex and age.

What are the units, outcome, and covariates in this problem?

- Unit: one recruit
- Outcome: `height` in centimeters
- Covariates: `sex` and `age`
- Treatment (if any): none
- Quantity of interest: the expected height for future recruits, summarized by the fitted model

What belongs in the conceptual Preceptor Table for this problem? You do not need to construct it yet.

The conceptual Preceptor Table should have one row per recruit, a height outcome column, and the covariates we know before prediction. In this exercise, the columns are unit, outcome, and covariates.

Teaching fellow prompt: Ask the class to define Wisdom and a Preceptor Table in their own words before they answer.

The instructor will display the completed Preceptor Table after the discussion.

```{r}
#| echo: false
#| results: asis
preceptor_tibble <- tibble::tribble(
  ~Unit, ~Outcome, ~Covariate,
  "Recruit 1", "150", "Female",
  "Recruit 2", "172", "Male",
  "...", "...", "...",
  "Recruit 5,000", "160", "Female"
)

preceptor_html <- gt::gt(preceptor_tibble, id = "preceptor_table") |>
  gt::tab_header(title = "Completed Preceptor Table") |>
  gt::tab_spanner(label = "Unit", id = "unit_span", columns = "Unit") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Unit, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "Each row is one recruit whose height we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariates are sex and age, known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    preceptor_html, "</div>\n```", sep = "")
```

## Justice

Now that we know the question, we need to think about the population and the assumptions that justify using the data to answer it.

Describe the Population Table in your own words.

Each row is one young adult observed at one moment in time, and the table connects the observed data to the recruits we want to forecast. The rows labeled "Data" come from the sample we observed; the rows labeled "Preceptor" represent the future recruits we want to predict.

What is one realistic violation of validity or representativeness in this problem?

A realistic violation of validity would be if measured heights were recorded with different rules than the heights we care about for recruits. A realistic violation of representativeness would be if the sample were heavily unbalanced by sex or if the future recruits were a very different group from the people in the data.

Why is unconfoundedness not part of this predictive modeling problem?

Unconfoundedness is not part of this problem because we are not asking about a treatment effect. There is no intervention being assigned, so there is no need to reason about treatment assignment being independent of potential outcomes.

Teaching fellow prompt: Ask the class to define Justice and a Population Table before they answer.

```{r}
#| echo: false
#| results: asis
population_tibble <- tibble::tribble(
  ~Source, ~`Unit / Time`, ~Outcome, ~Covariate,
  "Data", "NHANES 1 (2002)", "162", "Female",
  "Data", "NHANES 2 (2005)", "178", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 1 (2026)", "150", "Female",
  "Preceptor", "Recruit 2 (2026)", "172", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 5,000 (2026)", "160", "Female"
)

population_html <- gt::gt(population_tibble, id = "population_table") |>
  gt::tab_header(title = "Completed Population Table") |>
  gt::tab_spanner(label = "Unit / Time", id = "unit_span", columns = "Unit / Time") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Source, `Unit / Time`, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "The data rows are the observed sample; the preceptor rows are the recruits we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariate is sex, which is known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    population_html, "</div>\n```", sep = "")
```

## Courage

Before we fit a model, we should understand what model we are using and why.

What outcome are we modeling? What predictors will we use?

The outcome is `height`, a continuous numeric variable. We will model it with a Gaussian linear model and an identity link. A linear model is appropriate because height is continuous and the mean height can be described as a linear function of the predictors.

Why is this model reasonable?

A linear model is reasonable because the mean height can be expressed as a linear function of the predictors, and the outcome is continuous rather than categorical.

Teaching fellow prompt: Ask the class to explain why a linear model is the right choice for a continuous height outcome.

The instructor will show the fitted model and its output.

```{r}
fit_recruits <- lm(height ~ sex, data = recruits)
summary(fit_recruits)
```

## Temperance

Now that the model is fitted, we need to interpret it carefully and turn the result into a short Quarto report.

What is the main conclusion from the fitted model?

The fitted model says that female recruits are shorter on average than male recruits, and the model can be used to estimate the expected height for a recruit with given covariates.

Ask AI to create a plot that summarizes the fitted model and the data. Use AI for the coding, and then write a short Quarto document that includes the plot and a brief interpretation. Publish the result to GitHub Pages.

If you finish early, ask AI to help you explore one of these ideas for the recruits data:

- suggest a better-looking plot with clearer labels, colors, or annotations for the height-by-sex relationship;
- identify any unusual recruits or outliers and explain whether they change the main conclusion;
- use functions such as `summarise()`, `group_by()`, `filter()`, or `arrange()` to examine the data more closely and then describe what you learn;
- ask AI to propose a few follow-up questions about the data that could shape a stronger final Quarto report.

Teaching fellow prompt: Ask the class to explain the difference between a fitted mean and a future observation.

```{r}
recruits |>
  ggplot(aes(x = sex, y = height)) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  geom_point(
    data = recruits |>
      group_by(sex) |>
      summarise(mean_height = mean(height), .groups = "drop"),
    aes(y = mean_height),
    color = "steelblue",
    size = 3
  ) +
  labs(
    title = "Observed heights and average height by sex",
    x = "Sex",
    y = "Height (cm)"
  ) +
  theme_minimal()
```
