---
title: "Wisdom and Recruits (Answer Key)"
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(primer.tutorials)
```



## Wisdom

What will be the average height of male and female USMC recruits next year?

**What are the units, outcome, and covariates for this question?**

- Units: 

One recruit

- Outcomes:

 `height` in centimeters

- Covariates: 

`sex`

- Treatment (if any): 

None

- Quantity of interest: 

The average height for male and female recruits, and the difference between them.



Describe the key components of a Preceptor table for this problem; Use words like "units," "outcomes," and "covariates."

The Preceptor Table should have one row per recruit. In this exercise, the columns are unit, outcome, and covariates.




```{r}
#| echo: false
#| results: asis
preceptor_tibble <- tibble::tribble(
  ~Unit, ~Outcome, ~Covariate,
  "Recruit 1", "150", "Female",
  "Recruit 2", "172", "Male",
  "...", "...", "...",
  "Recruit 5,000", "160", "Female"
)

preceptor_html <- gt::gt(preceptor_tibble, id = "preceptor_table") |>
  gt::tab_header(title = "Completed Preceptor Table") |>
  gt::tab_spanner(label = "Unit", id = "unit_span", columns = "Unit") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Unit, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "Each row is one recruit whose height we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariates are sex and age, known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    preceptor_html, "</div>\n```", sep = "")
```



Create a plot that shows the relationship between height and sex.

```{r}
recruits |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(alpha = 0.6) +
  labs(
    title = "Height Distribution by Sex",
    x = "Height (cm)",
    y = "Density",
    fill = "Sex"
  ) +
  theme_minimal()
```


Create a plot that shows the relationship between height by age.

```{r}
recruits |>
  ggplot(aes(x = age, y = height, color = sex)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Height vs. Age by Sex",
    x = "Age (years)",
    y = "Height (cm)",
    color = "Sex"
  ) +
  theme_minimal()
```


---

## Justice

Describe the key components of a Population table for this problem. 


```{r}
#| echo: false
#| results: asis
population_tibble <- tibble::tribble(
  ~Source, ~`Unit / Time`, ~Outcome, ~Covariate,
  "Data", "NHANES 1 (2002)", "162", "Female",
  "Data", "NHANES 2 (2005)", "178", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 1 (2026)", "150", "Female",
  "Preceptor", "Recruit 2 (2026)", "172", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 5,000 (2026)", "160", "Female"
)

population_html <- gt::gt(population_tibble, id = "population_table") |>
  gt::tab_header(title = "Completed Population Table") |>
  gt::tab_spanner(label = "Unit / Time", id = "unit_span", columns = "Unit / Time") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Source, `Unit / Time`, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "The data rows are the observed sample; the preceptor rows are the recruits we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariate is sex, which is known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    population_html, "</div>\n```", sep = "")
```


Counter-example to the assumption of validity:



Counter-example to the assumption of stability:



Counter-example to the assumption of representativeness:



Counter-example to the assumption of unconfoundedness:




What probability family should we use?



Why should we use that probability family?



What is the link function for our outcome?



## Courage

Create a candidate model, modeling the outcome height as a linear function of the binary covariate sex.

```{r}
#| echo: true
linear_reg(engine = "lm") |>
  fit(height ~ sex, data = recruits)
```


Create another candidate model, modeling the outcome height as a linear function of the continuous covariate age. 

```{r courage-5-test}
#| echo: true
linear_reg(engine = "lm") |>
  fit(height ~ age, data = recruits)
```


## Temperance


What is the main conclusion from the fitted model?

Male recruits are taller on average than female recruits. The difference is about 13 cm, with male recruits averaging around 175 cm and female recruits averaging around 162 cm.

Ask AI to create a plot that summarizes the fitted model and the data. Use AI for the coding, and then write a short Quarto document that includes the plot and a brief interpretation. Publish the result to GitHub Pages.

```{r}
recruits |>
  ggplot(aes(x = sex, y = height)) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  geom_point(
    data = recruits |>
      group_by(sex) |>
      summarise(mean_height = mean(height), .groups = "drop"),
    aes(y = mean_height),
    color = "steelblue",
    size = 3
  ) +
  labs(
    title = "Observed heights and average height by sex",
    x = "Sex",
    y = "Height (cm)"
  ) +
  theme_minimal()
```
