---
title: "Wisdom and Recruits (Answer Key)"
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(tidymodels)
library(marginaleffects)
library(gt)
library(primer.tutorials)
```



## Wisdom

What will be the average height of male and female USMC recruits next year?



**What are the units, outcome, and covariates for this question?**

- Units: 

One recruit

- Outcomes:

 `height` in centimeters

- Covariates: 

`sex`

- Treatment (if any): 

None

- Quantity of interest: 

The average height for male and female recruits, and the difference between them.


Describe the key components of a Preceptor table for this problem; Use words like "units," "outcomes," and "covariates."

  The Preceptor Table should have one row per recruit. In this exercise, the columns are unit, outcome, and covariates.

```{r}
#| echo: false
#| results: asis
preceptor_tibble <- tibble::tribble(
  ~Unit, ~Outcome, ~Covariate,
  "Recruit 1", "150", "Female",
  "Recruit 2", "172", "Male",
  "...", "...", "...",
  "Recruit 5,000", "160", "Female"
)

preceptor_html <- gt::gt(preceptor_tibble, id = "preceptor_table") |>
  gt::tab_header(title = "Completed Preceptor Table") |>
  gt::tab_spanner(label = "Unit", id = "unit_span", columns = "Unit") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Unit, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "Each row is one recruit whose height we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariates are sex and age, known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    preceptor_html, "</div>\n```", sep = "")
```


Create a plot that shows the relationship between height and sex.

```{r}
recruits |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(alpha = 0.6) +
  labs(
    title = "Height Distribution by Sex",
    x = "Height (cm)",
    y = "Density",
    fill = "Sex"
  ) +
  theme_minimal()
```


Create a plot that shows the relationship between height by age.

```{r}
recruits |>
  ggplot(aes(x = age, y = height, color = sex)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Height vs. Age by Sex",
    x = "Age (years)",
    y = "Height (cm)",
    color = "Sex"
  ) +
  theme_minimal()
```


---

## Justice

Describe the key components of a Population table for this problem. 


```{r}
#| echo: false
#| results: asis
population_tibble <- tibble::tribble(
  ~Source, ~`Unit / Time`, ~Outcome, ~Covariate,
  "Data", "NHANES 1 (2002)", "162", "Female",
  "Data", "NHANES 2 (2005)", "178", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 1 (2026)", "150", "Female",
  "Preceptor", "Recruit 2 (2026)", "172", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 5,000 (2026)", "160", "Female"
)

population_html <- gt::gt(population_tibble, id = "population_table") |>
  gt::tab_header(title = "Completed Population Table") |>
  gt::tab_spanner(label = "Unit / Time", id = "unit_span", columns = "Unit / Time") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Source, `Unit / Time`, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "The data rows are the observed sample; the preceptor rows are the recruits we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariate is sex, which is known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    population_html, "</div>\n```", sep = "")
```


Counter-example to the assumption of validity:

  The height column in NHANES is measured differently than the height column in the Preceptor Table, so the two columns may not represent exactly the same construct.

Counter-example to the assumption of stability:

  The relationship between height and sex may have changed over time, so the model parameters (such as the intercept or the slope on sexMale) may not be the same for the data and the Preceptor Table.

Counter-example to the assumption of representativeness:

  The NHANES sample is not representative of future USMC recruits because recruits are self-selected and screened for military service.

Counter-example to the assumption of unconfoundedness:

  Sex is not a manipulable treatment, and treatment assignment is not independent of the potential outcomes, so the unconfoundedness assumption fails.

What probability family should we use?

  Normal Distribution

Why should we use that probability family?

  The Normal distribution is the standard choice for continuous outcomes.

What is the link function for our outcome?

  Our model will be a linear regression.

## Courage

Create a candidate model, modeling the outcome height as a linear function of the binary covariate sex.

```{r}
#| echo: true
fit_recruits <- linear_reg(engine = "lm") |>
  fit(height ~ sex, data = recruits)
```

Create another candidate model, modeling the outcome height as a linear function of the continuous covariate age. 

```{r courage-5-test}
#| echo: true
fit_age <- linear_reg(engine = "lm") |>
  fit(height ~ age, data = recruits)
```


## Temperance

Use `predictions(fit_recruits)` to estimate the average height of male and female USMC recruits next year.

```{r}
predictions(fit_recruits)
```

What is the expected height of a female recruit?

  The expected height of a female recruit is about **165.9 cm**, with a 95% confidence interval of approximately **161.8 cm to 170.0 cm**.

What is the expected height of a male recruit?

  The expected height of a male recruit is about **174.8 cm**, with a 95% confidence interval of approximately **172.5 cm to 177.0 cm**.

Create a visualization of these predictions using `plot_predictions()`.

```{r}
plot_predictions(
  fit_recruits,
  condition = "sex",
  draw = FALSE
) |>
  ggplot(aes(x = sex, y = estimate)) +
  geom_pointrange(
    aes(
      ymin = conf.low,
      ymax = conf.high
    ),
    linewidth = 0.7
  ) +
  labs(
    title = "Predicted Average Height of USMC Recruits by Sex",
    subtitle = "Male recruits are predicted to be taller on average than female recruits.",
    x = "Sex",
    y = "Predicted Height (cm)",
    caption = "Source: 50-row teaching sample from NHANES via primer.tutorials::recruits"
  ) +
  theme_minimal(base_size = 13)
```


What measure of uncertainty is shown?

  The vertical error bars show **95% confidence intervals** for the estimated average height of each group.

Why is the confidence interval wider for female recruits?

  There are only **10 female observations** compared with **40 male observations**, so the female average is estimated with less precision.

Besides the average height, identify two additional quantities of interest the Quartermaster might care about.

- The **tallest recruit**, to ensure the largest uniform size is available.
- The **10th and 90th percentiles** of recruit height, to estimate how many extra-small and extra-large uniforms should be ordered.

Why might these estimated average heights be wrong?

  The estimates rely on NHANES data rather than actual USMC recruits. Because recruits are self-selected, screened for military service, and likely taller on average than the general young-adult population, the model probably understates recruit height. In addition, the reported confidence intervals reflect only sampling uncertainty and do not account for possible violations of validity, stability, or representativeness.
