---
title: "Wisdom and Recruits (Answer Key)"
format:
  html: default
  pdf: default
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(tidymodels)
library(marginaleffects)
library(gt)
library(primer.tutorials)
```

## Wisdom

You are in charge of ordering uniforms for next year's Marine Corps bootcamp recruits.

What will be the average height of male and female USMC recruits next year?

**What are the units, outcome, and covariates for this question?**

- Units:

Each row represents one USMC recruit expected to enlist during the coming year. There are roughly 5,000 such recruits per year.

- Outcomes:

Keep track of two 'outcome' variables: the one in our Preceptor Table and the one in our data. In this case, they share the name `height` --- both refer to a young adult's standing height in centimeters --- though the *measurement procedure* behind each column differs.

- Covariates:

`sex` and `age` are candidates. Age at enlistment is a candidate. Height is nearly-but-not-quite stable by the age-18 window we care about, so a ten-year-older recruit is almost the same height as a fresh one. Another candidate is race or ethnicity, since height distributions vary across ancestry groups.
  

- Treatment (if any):

There are no treatments in this problem because we are building a predictive model. Nothing in `recruits` can be manipulated --- we cannot assign a recruit a different sex or a different height. Predictive models have covariates, not treatments.
  

- Quantity of interest:

The average height for male and female recruits, and the difference between them.

Describe the key components of a Preceptor table for this problem:

The Preceptor Table should have one row per recruit. In this exercise, the columns are Recruit, Height, and Sex. 

```{r}
#| echo: false
#| results: asis

preceptor_tibble <- tibble::tribble(
  ~`Recruit`            , ~`Height (cm)`, ~`Sex`   ,
  "Emma Rodriguez"      , "166"         , "Female",
  "Marcus Johnson"      , "180"         , "Male"  ,
  "..."                 , "..."         , "..."   ,
  "David Chen"          , "177"         , "Male"
)

pre_title_footnote <- "If all the information in this table were available, we could answer the question: What is the average height of male and female USMC recruits next year?"

pre_units_footnote <- "Each row represents one recruit entering USMC bootcamp in 2026. The total population of annual recruits is roughly 35,000–40,000 (about 10% female, 90% male by historical rates), though screening policies and self-selection may shift these proportions."

pre_outcome_footnote <- "Height measured in centimeters from the soles of the feet to the top of the head, taken at standard military induction physical examination. This is a continuous measurement with no artificial rounding or truncation in our analysis."

pre_covariate_footnote <- "Sex as recorded on military identification: Male or Female. This is the trait the question uses to partition recruits into two comparison groups."

pre_gt_html <- gt::gt(preceptor_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit"     , id = "unit_span",
    columns = c(`Recruit`)) |>
  gt::tab_spanner(label = "Outcome"  , id = "outcome_span",
    columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
    columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
    locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
    locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
    locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote,
    locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

Create a plot that shows the relationship between height and sex.

```{r}
#| echo: true
sex_counts <- recruits |> count(sex)

recruits |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(aes(y = after_stat(count)), alpha = 0.6) +
  labs(
    title = "Height Distribution by Sex",
    subtitle = paste0(
      "Area is scaled by sample size: n = ", sex_counts$n[sex_counts$sex == "Male"],
      " male vs. n = ", sex_counts$n[sex_counts$sex == "Female"], " female recruits"
    ),
    x = "Height (cm)",
    y = "Count-scaled density",
    fill = "Sex"
  ) +
  theme_minimal()
```

Male recruits outnumber female recruits four to one in this sample, so the male curve covers roughly four times as much area as the female curve. This is due to the sample size, and should not be taken as evidence that male and female heights are equally common or equally variable.

Create a plot that shows the relationship between height by age.

```{r}
#| echo: true
age_fits <- recruits |>
  group_by(sex) |>
  summarize(slope = coef(lm(height ~ age))[["age"]], .groups = "drop")

recruits |>
  ggplot(aes(x = age, y = height, color = sex)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Height vs. Age by Sex",
    subtitle = paste0(
      "Slope is ", round(age_fits$slope[age_fits$sex == "Female"], 2),
      " cm/yr for females and ", round(age_fits$slope[age_fits$sex == "Male"], 2),
      " cm/yr for males — both near zero, as expected for adults"
    ),
    x = "Age (years)",
    y = "Height (cm)",
    color = "Sex"
  ) +
  theme_minimal()
```

Among adults aged 18-27, height has mostly stopped changing, so neither line should show a strong trend. Whatever slope appears is sampling noise, not a real age effect.

------------------------------------------------------------------------

## Justice

```{r}
#| echo: false
#| results: asis

population_tibble <- tibble::tribble(
  ~Source    , ~`Recruit`        , ~`Year`, ~`Height (cm)`, ~`Sex`   ,
  "..."      , "..."             , "..."  , "..."          , "..."   ,
  "Data"     , "Emma Rodriguez"  , "2005" , "165"          , "Female",
  "Data"     , "Marcus Johnson"  , "2005" , "179"          , "Male"  ,
  "Data"     , "..."             , "..."  , "..."          , "..."   ,
  "Data"     , "David Chen"      , "2005" , "176"          , "Male"  ,
  "..."      , "..."             , "..."  , "..."          , "..."   ,
  "Preceptor", "Emma Rodriguez"  , "2026" , "166"          , "Female",
  "Preceptor", "Marcus Johnson"  , "2026" , "180"          , "Male"  ,
  "Preceptor", "..."             , "..."  , "..."          , "..."   ,
  "Preceptor", "David Chen"      , "2026" , "177"          , "Male"  ,
  "..."      , "..."             , "..."  , "..."          , "..."
)

pop_title_footnote <- "This table bridges from NHANES data (2003–2006) to next year's USMC recruits. Both come from the population of young adults aged 18–27 living in the United States."

pop_units_footnote <- "Each row is a young adult at a specific time. Data rows are 50 participants in the National Health and Nutrition Examination Survey (NHANES), a cross-sectional sample of the US non-institutionalized civilian population, surveyed between 2003 and 2006. Preceptor rows are recruits entering USMC bootcamp in 2026, drawn from volunteers aged 18–27. The 20-year gap between the data era and the prediction year, and the shift from a population sample to military volunteers, both raise representativeness concerns."

pop_outcome_footnote <- "Height in centimeters, measured from sole to crown. NHANES used standardized protocols. Preceptor rows show the (unobservable) height each recruit will have in 2026, which we model by assuming NHANES heights remain stable across the population."

pop_covariate_footnote <- "Sex as recorded in the two data sources. NHANES sex is self-reported or interviewer-recorded at survey time. USMC recruits' sex comes from military identification at induction. Both are treated as a binary Male/Female classification."

pop_gt_html <- gt::gt(population_tibble, id = "population_tbl") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time"  , id = "unit_span",
    columns = c(`Recruit`, `Year`)) |>
  gt::tab_spanner(label = "Outcome"    , id = "outcome_span",
    columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate"  , id = "covariates_span",
    columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Source`, `Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote,
    locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote,
    locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote,
    locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote,
    locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

Counter-example to the assumption of validity:

The height column in NHANES might be measured differently than the height column in the Preceptor Table, so the two columns may not represent exactly the same construct. NHANES heights are measured in the Mobile Examination Center by trained examiners using a fixed stadiometer, with shoes removed, instructions standardized across all participants. USMC heights are recorded at the recruiting office or at bootcamp intake, with whatever stadiometer is available, possibly with shoes on, possibly with whatever instructional drift has accumulated since the last standardization. A systematic offset of a centimeter or two between the two procedures would not be surprising.


Counter-example to the assumption of stability:

The relationship between height and sex may have changed over time, so the model parameters (such as the intercept or the slope on sexMale) may not be the same for the data and the Preceptor Table. For Recruits, the relevant stability worry is small. Adult human height has been stable for the broad U.S. young-adult population for several decades — the intercept may have drifted upward by a centimeter or two as nutrition has improved, but the male-female height gap (the parameter we care about most) has barely moved.

Counter-example to the assumption of representativeness:

Data → Population. NHANES is much closer to a probability sample than most U.S. health surveys because it uses a stratified, multistage sampling design to represent the U.S. civilian, non-institutionalized population. However, the educational version of NHANES used here omits the survey weights, and our 50-row teaching sample further over-represents males (40 males, 10 females) to highlight differences in estimation uncertainty. As a result, our data are not representative of U.S. young adults. This does not necessarily bias our estimates, but it weakens our justification for treating the sample as representative of the target population.

Population → Preceptor Table. Even if NHANES perfectly represented U.S. young adults, that population is still different from the population of future USMC recruits. Military recruits are volunteers who must meet eligibility and fitness requirements, so they are not a random sample of young adults. If recruits differ systematically from the general population in characteristics related to height, then a model trained on NHANES may not accurately predict the average height of next year's recruits. Any resulting prediction error would arise because the target population differs from the population represented by the data.

Counter-example to the assumption of unconfoundedness:

None. Unconfoundedness applies only to causal questions, ours is predictive. We are not trying to estimate the causal effect of sex on height. 

What probability family should we use?

Normal Distribution

Why should we use that probability family?

The Normal distribution is the standard choice for continuous outcomes.

What is the link function for our outcome?

Our model will be a linear regression.

## Courage

Create a candidate model, modeling the outcome height as a linear function of the binary covariate sex.

```{r}
fit_recruits <- linear_reg(engine = "lm") |>
  fit(height ~ sex, data = recruits)
fit_recruits
```

Create another candidate model, modeling the outcome height as a linear function of the continuous covariate age.

```{r courage-5-test}
fit_age <- linear_reg(engine = "lm") |>
  fit(height ~ age, data = recruits)
fit_age
```

## Temperance

Use `predictions(fit_recruits)` to estimate the average height of male and female USMC recruits next year.

```{r}
predictions(fit_recruits)
```

What is the expected height of a female recruit?

The expected height of a female recruit is about **165.9 cm**, with a 95% confidence interval of approximately **161.8 cm to 170.0 cm**.

What is the expected height of a male recruit?

The expected height of a male recruit is about **174.8 cm**, with a 95% confidence interval of approximately **172.5 cm to 177.0 cm**.

Create a visualization of these predictions using `plot_predictions()`.

```{r}
#| echo: true
avg_preds <- avg_predictions(fit_recruits, variables = "sex")
taller_sex <- avg_preds$sex[which.max(avg_preds$estimate)]
height_diff <- round(diff(range(avg_preds$estimate)), 1)

plot_predictions(
  fit_recruits,
  condition = "sex",
  draw = FALSE
) |>
  ggplot(aes(x = sex, y = estimate)) +
  geom_pointrange(
    aes(
      ymin = conf.low,
      ymax = conf.high
    ),
    linewidth = 0.7
  ) +
  labs(
    title = "Predicted Average Height of USMC Recruits by Sex",
    subtitle = paste0(taller_sex, " recruits are predicted to be about ", height_diff, " cm taller, on average"),
    x = "Sex",
    y = "Predicted Height (cm)",
    caption = "Source: 50-row teaching sample from NHANES via primer.tutorials::recruits"
  ) +
  theme_minimal(base_size = 13)
```

What measure of uncertainty is shown?

The vertical error bars show **95% confidence intervals** for the estimated average height of each group.

Why is the confidence interval wider for female recruits?

There are only **10 female observations** compared with **40 male observations**, so the female average is estimated with less precision.


Why might these estimated average heights be wrong?

The estimates may be wrong because the model is based on NHANES, which surveys the general U.S. population, rather than people who actually enlist in the Marine Corps. For example, joining the Marines is voluntary, so people who choose to enlist may differ from the average young adult in important ways, such as physical fitness or body size. In addition, recruits must meet medical and physical eligibility requirements before they can serve. If the people who enlist and qualify tend to be taller (or shorter) than the general population, then our predicted average heights will be biased. Finally, the confidence intervals only measure sampling uncertainty—they do not account for problems such as using the wrong population, omitted variables, or changes in recruit characteristics over time.
