---
title: "Wisdom and Recruits (Answer Key)"
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(gt)
library(primer.tutorials)
```

## Wisdom

We are working with the `recruits` dataset. This exercise is a predictive modeling exercise. We will use it to reason about how to forecast the height of next year's recruits from information we already know about them.

What question are we trying to answer?

We are trying to forecast the average height of next year's Marine Corps recruits from information such as sex and age.

What are the units in this problem? What is the outcome? What covariates are relevant? Is there a treatment here? What quantity of interest do we want to summarize?

The units are individual recruits. The outcome is `height`, measured in centimeters. The relevant covariates are `sex` and `age`. This exercise does not include a treatment, because we are making a predictive forecast rather than asking about an intervention. The quantity of interest is the expected height of recruits, or the mean height for groups defined by the available covariates.

What belongs in the conceptual Preceptor Table for this problem? You do not need to construct it yet.

The conceptual Preceptor Table should have one row per recruit, a height outcome column, and the covariates we know before prediction. In this exercise, the columns are unit, outcome, and covariates.

Teaching fellow prompt: Ask the class to define Wisdom and a Preceptor Table in their own words before they answer.

The instructor will display the completed Preceptor Table after the discussion.

```{r}
#| echo: false
#| results: asis
preceptor_tibble <- tibble::tribble(
  ~Unit, ~Outcome, ~Covariate,
  "Recruit 1", "150", "Female",
  "Recruit 2", "172", "Male",
  "...", "...", "...",
  "Recruit 5,000", "160", "Female"
)

preceptor_html <- gt::gt(preceptor_tibble, id = "preceptor_table") |>
  gt::tab_header(title = "Completed Preceptor Table") |>
  gt::tab_spanner(label = "Unit", id = "unit_span", columns = "Unit") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Unit, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "Each row is one recruit whose height we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariates are sex and age, known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    preceptor_html, "</div>\n```", sep = "")
```

## Justice

Now that we know the question, we need to think about the population and the assumptions that justify using the data to answer it.

Describe the Population Table in your own words. What does each row represent in this context?

Each row is one young adult observed at one moment in time, and the table connects the observed data to the recruits we want to forecast. The rows labeled "Data" come from the sample we observed; the rows labeled "Preceptor" represent the future recruits we want to predict.

For each of the key assumptions, define it in your own words and give one realistic counterexample for the current problem.

- Validity: The measurement and the meaning of the outcome should be the same in the data and in the target population. A counterexample would be if the observed heights were recorded with different measurement rules than the heights we care about for recruits.
- Stability: The relationship between the covariates and the outcome should be similar across the data and the target recruits. A counterexample would be if the sex gap in height changed substantially between the sample period and the coming year.
- Representativeness: The data should be a reasonable stand-in for the recruits we want to predict. A counterexample would be if the sample were heavily unbalanced by sex or if the recruits were a very different group from the people in the data.

Why is unconfoundedness not part of this predictive modeling problem?

Unconfoundedness is not part of this problem because we are not asking about a treatment effect. There is no intervention being assigned, so there is no need to reason about treatment assignment being independent of potential outcomes.

Teaching fellow prompt: Ask the class to define Justice and a Population Table before they answer.

```{r}
#| echo: false
#| results: asis
population_tibble <- tibble::tribble(
  ~Source, ~`Unit / Time`, ~Outcome, ~Covariate,
  "Data", "NHANES 1 (2002)", "162", "Female",
  "Data", "NHANES 2 (2005)", "178", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 1 (2026)", "150", "Female",
  "Preceptor", "Recruit 2 (2026)", "172", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 5,000 (2026)", "160", "Female"
)

population_html <- gt::gt(population_tibble, id = "population_table") |>
  gt::tab_header(title = "Completed Population Table") |>
  gt::tab_spanner(label = "Unit / Time", id = "unit_span", columns = "Unit / Time") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Source, `Unit / Time`, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "The data rows are the observed sample; the preceptor rows are the recruits we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariate is sex, which is known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    population_html, "</div>\n```", sep = "")
```

## Courage

Before we fit a model, we should understand what model we are using and why.

What outcome are we modeling? What predictors will we use? Which probability family and link function are appropriate here? Explain why a linear model is reasonable for this outcome.

The outcome is `height`, a continuous numeric variable. We will model it with a Gaussian linear model and an identity link. A linear model is appropriate because height is continuous and the mean height can be described as a linear function of the predictors.

What do the model parameters mean in this setting?

The intercept is the expected height for the reference group, and the coefficient for `sexFemale` measures the expected change in height for female recruits relative to male recruits.

Teaching fellow prompt: Ask the class to explain why a linear model is the right choice for a continuous height outcome.

The instructor will show the fitted model and its output.

```{r}
fit_recruits <- lm(height ~ sex, data = recruits)
summary(fit_recruits)
```

## Temperance

Now that the model is fitted, we need to interpret it carefully.

How should we interpret the fitted coefficients? What does the fitted model say about the average height of male and female recruits?

The fitted model says that the average height for male recruits is the intercept, while the average height for female recruits is the intercept plus the coefficient for `sexFemale`. In this sample, the fitted model implies that female recruits are shorter on average than male recruits.

How would we use the fitted model to make a prediction for a new recruit? What uncertainty remains? What are the limitations of this model? How does the answer to the original question depend on those limitations?

We would plug the covariates for a new recruit into the fitted model to obtain a predicted mean height. The model does not tell us exactly what the recruit's height will be; it gives an expected value and leaves residual variation unexplained. The model is limited by the sample size, by the fact that we only used `sex`, and by the assumption that the same pattern will hold for next year's recruits. The answer to the original question depends on whether those assumptions are reasonable.

Teaching fellow prompt: Ask the class to explain the difference between a fitted mean and a future observation.

```{r}
recruits |>
  ggplot(aes(x = sex, y = height)) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  geom_point(
    data = recruits |>
      group_by(sex) |>
      summarise(mean_height = mean(height), .groups = "drop"),
    aes(y = mean_height),
    color = "steelblue",
    size = 3
  ) +
  labs(
    title = "Observed heights and average height by sex",
    x = "Sex",
    y = "Height (cm)"
  ) +
  theme_minimal()
```
