---
title: "Wisdom and Recruits (Answer Key)"
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(tidymodels)
library(marginaleffects)
library(primer.tutorials)
```



## Wisdom

What will be the average height of male and female USMC recruits next year?


**What are the units, outcome, and covariates for this question?**

- Units: 

One recruit

- Outcomes:

 `height` in centimeters

- Covariates: 

`sex`

- Treatment (if any): 

None

- Quantity of interest: 

The average height for male and female recruits, and the difference between them.



Describe the key components of a Preceptor table for this problem; Use words like "units," "outcomes," and "covariates."

The Preceptor Table should have one row per recruit. In this exercise, the columns are unit, outcome, and covariates.

```{r}
#| echo: false
#| results: asis
preceptor_tibble <- tibble::tribble(
  ~Unit, ~Outcome, ~Covariate,
  "Recruit 1", "150", "Female",
  "Recruit 2", "172", "Male",
  "...", "...", "...",
  "Recruit 5,000", "160", "Female"
)

preceptor_html <- gt::gt(preceptor_tibble, id = "preceptor_table") |>
  gt::tab_header(title = "Completed Preceptor Table") |>
  gt::tab_spanner(label = "Unit", id = "unit_span", columns = "Unit") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Unit, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "Each row is one recruit whose height we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariates are sex and age, known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    preceptor_html, "</div>\n```", sep = "")
```



Create a plot that shows the relationship between height and sex.

```{r}
recruits |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(alpha = 0.6) +
  labs(
    title = "Height Distribution by Sex",
    x = "Height (cm)",
    y = "Density",
    fill = "Sex"
  ) +
  theme_minimal()
```


Create a plot that shows the relationship between height by age.

```{r}
recruits |>
  ggplot(aes(x = age, y = height, color = sex)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Height vs. Age by Sex",
    x = "Age (years)",
    y = "Height (cm)",
    color = "Sex"
  ) +
  theme_minimal()
```


---

## Justice

Describe the key components of a Population table for this problem. 


```{r}
#| echo: false
#| results: asis
population_tibble <- tibble::tribble(
  ~Source, ~`Unit / Time`, ~Outcome, ~Covariate,
  "Data", "NHANES 1 (2002)", "162", "Female",
  "Data", "NHANES 2 (2005)", "178", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 1 (2026)", "150", "Female",
  "Preceptor", "Recruit 2 (2026)", "172", "Male",
  "...", "...", "...", "...",
  "Preceptor", "Recruit 5,000 (2026)", "160", "Female"
)

population_html <- gt::gt(population_tibble, id = "population_table") |>
  gt::tab_header(title = "Completed Population Table") |>
  gt::tab_spanner(label = "Unit / Time", id = "unit_span", columns = "Unit / Time") |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = "Outcome") |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = "Covariate") |>
  gt::cols_align(align = "left", columns = c(Source, `Unit / Time`, Covariate)) |>
  gt::cols_align(align = "right", columns = "Outcome") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_column_spanners()) |>
  gt::tab_footnote(footnote = "The data rows are the observed sample; the preceptor rows are the recruits we want to predict.", locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = "The outcome is height in centimeters.", locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = "The covariate is sex, which is known before prediction.", locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    population_html, "</div>\n```", sep = "")
```


Counter-example to the assumption of validity:

The height column in NHANES is measured differently than the height column in the Preceptor Table, so the two columns may not represent exactly the same construct.

Counter-example to the assumption of stability:

The relationship between height and sex may have changed over time, so the model parameters (such as the intercept or the slope on sexMale) may not be the same for the data and the Preceptor Table.

Counter-example to the assumption of representativeness:

The NHANES sample is not representative of future USMC recruits because recruits are self-selected and screened for military service.

Counter-example to the assumption of unconfoundedness:

Sex is not a manipulable treatment, and treatment assignment is not independent of the potential outcomes, so the unconfoundedness assumption fails.

What probability family should we use?

Normal Distribution

Why should we use that probability family?

The Normal distribution is the standard choice for continuous outcomes.

What is the link function for our outcome?

Our model will be a linear regression.

## Courage

Create a candidate model, modeling the outcome height as a linear function of the binary covariate sex.

```{r}
#| echo: true
fit_recruits <- linear_reg(engine = "lm") |>
  fit(height ~ sex, data = recruits)
```

Create another candidate model, modeling the outcome height as a linear function of the continuous covariate age. 

```{r courage-5-test}
#| echo: true
fit_age <- linear_reg(engine = "lm") |>
  fit(height ~ age, data = recruits)
```


## Temperance

Use `predictions(fit_recruits)` and `plot_predictions(fit_recruits, condition = "sex", draw = FALSE)` to forecast the average height of USMC recruits next year in a visualization.

```{r}
#| message: false
library(tidyverse)
library(tidymodels)
library(primer.tutorials)
library(broom)
library(marginaleffects)

#| cache: true
fit_recruits <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = recruits)

plot_predictions(fit_recruits, condition = "sex", draw = FALSE) |>
  ggplot(aes(x = sex, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  labs(
    title = "Forecast Height for USMC Recruits",
    subtitle = "Male recruits are expected to be about 9 cm taller than female recruits",
    x = "Sex",
    y = "Expected Height (cm)",
    caption = "Source: 50-row sample from NHANES via primer.tutorials::recruits"
  ) +
  theme_minimal()
```

The expected height of a male recruit: 

The expected height of a male recruit is about 174.8 cm, with a 95% confidence interval of roughly [172.5, 177.0].

The expected height of a female recruit:

The expected height of a female recruit is about 165.9 cm, with a 95% confidence interval of roughly [161.8, 170.0].

Using your fitted model, estimate the 10th and 90th percentiles of recruit height by sex.

```{r}
predictions(
  fit_recruits,
  newdata = tibble(
    sex = c("Female", "Male")
  )
)
```

Based on your estimates, explain how the USMC could use these percentiles when ordering uniforms.

 The USMC could use these estimates to decide how many extra-small and extra-large uniforms should be stocked.



Why might your estimate of the average recruit height be wrong? 


USMC recruits are likely taller than the NHANES population because recruits are self-selected and screened for military service.
