---
title: "Recruits Answer Guideline"
format:
  html: default
  pdf: default
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(tidymodels)
library(marginaleffects)
library(gt)
library(broom)
library(primer.tutorials)
```

## The Question

You are in charge of ordering uniforms for next year's Marine Corps bootcamp recruits.

What will be the average height of male and female USMC recruits next year?

## Wisdom

**What are the units, outcome, and covariates for this question?**

- Units:

Each row represents one USMC recruit expected to enlist during the coming year. There are roughly 25,000 to 30,000 such recruits per year.

- Outcomes:

`height` in centimeters. Both our data and our Preceptor Table use this name, though the *measurement procedure* behind each column differs --- a concern we return to under validity.

- Covariates:

`sex` is the covariate the question forces us to include. `age` is a candidate we consider in Courage, but height is nearly stable across the 18--27 adult window, so we expect it to add little.

- Treatment (if any):

None. This is a predictive model. Nothing in `recruits` can be manipulated --- we cannot assign a recruit a different sex. Predictive models have covariates, not treatments.

- Quantity of interest:

The average height for male and female recruits, and the difference between them.

Describe the key components of a Preceptor table for this problem:

  The Preceptor Table should have one row per recruit. In this exercise, the columns are Recruit, Height, and Sex. 

```{r}
#| label: preceptor-table
#| echo: false
#| results: asis

preceptor_tibble <- tibble::tribble(
  ~`Recruit`,        ~`Height (cm)`, ~`Sex`,
  "Emma Rodriguez",  "166",          "Female",
  "Marcus Johnson",  "180",          "Male",
  "...",             "...",          "...",
  "David Chen",      "177",          "Male"
)

pre_title_footnote <- "If all the information in this table were available, we could answer the question: What will be the average height of male and female USMC recruits next year?"
pre_units_footnote <- "Each row represents one recruit entering USMC bootcamp in 2026. The total population of annual recruits is roughly 25,000-30,000."
pre_outcome_footnote <- "Height measured in centimeters from the soles of the feet to the top of the head, taken at standard military induction physical examination."
pre_covariate_footnote <- "Sex as recorded on military identification: Male or Female. This is the trait the question uses to partition recruits into two comparison groups."

preceptor_html <- gt::gt(preceptor_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit", id = "unit_span", columns = c(`Recruit`)) |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = c(`Sex`)) |>
  gt::cols_align(align = "left", columns = c(`Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_column_spanners()) |>
  gt::tab_style(style = gt::cell_text(align = "left"),
                locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))) |>
  gt::tab_style(style = gt::cell_text(align = "right"),
                locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_title_footnote, locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote, locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote, locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote, locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(css = "#preceptor_tbl .gt_footnote { max-width: 1px; word-break: break-word; }") |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    preceptor_html, "</div>\n```\n", sep = "")
```

The Preceptor Table is the smallest table that, if every cell were filled in with the truth, would make the question's answer easy to read off. 


EDA:

Before fitting a model, look at the data. The hour spent on a careful EDA almost always saves a day downstream. We look at height-by-sex three complementary ways, then height-by-age three complementary ways.

**Plots that show the relationship between height and sex.**

```{r}
#| label: eda-height-sex-boxplot
#| echo: true

recruits |>
  ggplot(aes(x = sex, y = height, fill = sex)) +
  geom_boxplot(width = 0.5, alpha = 0.7, show.legend = FALSE) +
  labs(
    title = "Height by Sex: Boxplot",
    subtitle = "The median male recruit is taller than essentially every female recruit",
    x = "Sex",
    y = "Height (cm)"
  ) +
  theme_minimal()
```

The boxplot shows medians, spread, and the near-complete separation between the two boxes' interquartile ranges -- a first signal that sex carries real information about height.

```{r}
#| label: eda-height-sex-violin
#| echo: true

recruits |>
  ggplot(aes(x = sex, y = height, fill = sex)) +
  geom_violin(alpha = 0.6, show.legend = FALSE) +
  geom_jitter(width = 0.08, alpha = 0.5, size = 1.5) +
  labs(
    title = "Height by Sex: Violin Plot with Jittered Points",
    subtitle = "Each dot is one recruit; the violin width shows where heights cluster",
    x = "Sex",
    y = "Height (cm)"
  ) +
  theme_minimal()
```

The violin plot adds shape to the boxplot's summary, and the jittered points let us see exactly how few female recruits (10) there are compared to male recruits (40) -- a fact the boxplot alone hides.

```{r}
#| label: eda-height-sex-density
#| echo: true

sex_counts <- recruits |> count(sex)

recruits |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(aes(y = after_stat(count)), alpha = 0.6) +
  labs(
    title = "Height Distribution by Sex",
    subtitle = paste0(
      "Area is scaled by sample size: n = ", sex_counts$n[sex_counts$sex == "Male"],
      " male vs. n = ", sex_counts$n[sex_counts$sex == "Female"], " female recruits"
    ),
    x = "Height (cm)",
    y = "Count-scaled density",
    fill = "Sex"
  ) +
  theme_minimal()
```

Male recruits outnumber female recruits four to one in this sample, so the male curve covers roughly four times as much area as the female curve. This is due to the sample size, and should not be taken as evidence that male and female heights are equally common or equally variable.

**Plots that show the relationship between height and age.**

```{r}
#| label: eda-height-age-scatter
#| echo: true

age_fits <- recruits |>
  group_by(sex) |>
  summarize(slope = coef(lm(height ~ age))[["age"]], .groups = "drop")

recruits |>
  ggplot(aes(x = age, y = height, color = sex)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Height vs. Age by Sex",
    subtitle = paste0(
      "Slope is ", round(age_fits$slope[age_fits$sex == "Female"], 2),
      " cm/yr for females and ", round(age_fits$slope[age_fits$sex == "Male"], 2),
      " cm/yr for males -- both near zero, as expected for adults"
    ),
    x = "Age (years)",
    y = "Height (cm)",
    color = "Sex"
  ) +
  theme_minimal()
```

Among adults aged 18-27, height has mostly stopped changing, so none of these views show a strong age trend. Whatever slope appears is sampling noise, not a real age effect.

------------------------------------------------------------------------

## Justice

```{r}
#| label: population-table
#| echo: false
#| results: asis

population_tibble <- tibble::tribble(
  ~Source,     ~`Recruit`,       ~`Year`, ~`Height (cm)`, ~`Sex`,
  "...",       "...",            "...",   "...",          "...",
  "Data",      "Emma Rodriguez", "2005",  "165",          "Female",
  "Data",      "Marcus Johnson", "2005",  "179",          "Male",
  "Data",      "...",            "...",   "...",          "...",
  "Data",      "David Chen",     "2005",  "176",          "Male",
  "...",       "...",            "...",   "...",          "...",
  "Preceptor", "Emma Rodriguez", "2026",  "166",          "Female",
  "Preceptor", "Marcus Johnson", "2026",  "180",          "Male",
  "Preceptor", "...",            "...",   "...",          "...",
  "Preceptor", "David Chen",     "2026",  "177",          "Male",
  "...",       "...",            "...",   "...",          "..."
)

pop_title_footnote <- "This table bridges from NHANES data (circa 2005) to next year's USMC recruits. Both come from the population of young adults aged 18-27 living in the United States."
pop_units_footnote <- "Data rows are 50 participants in NHANES, surveyed circa 2003-2006. Preceptor rows are recruits entering USMC bootcamp in 2026. The same three individuals appear in both blocks at different ages -- the 20-year gap is the stability concern."
pop_outcome_footnote <- "Height in centimeters, measured from sole to crown using standardized NHANES protocols."
pop_covariate_footnote <- "Sex as recorded in each source: self-reported/interviewer-recorded at NHANES survey time; military-identification-recorded at USMC enlistment."

population_html <- gt::gt(population_tibble, id = "population_tbl") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span", columns = c(`Recruit`, `Year`)) |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span", columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span", columns = c(`Sex`)) |>
  gt::cols_align(align = "left", columns = c(`Source`, `Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_column_spanners()) |>
  gt::tab_style(style = gt::cell_text(align = "left"),
                locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))) |>
  gt::tab_style(style = gt::cell_text(align = "right"),
                locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_title_footnote, locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote, locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote, locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote, locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(css = "#population_tbl .gt_footnote { max-width: 1px; word-break: break-word; }") |>
  gt::as_raw_html()

cat("```{=html}\n",
    '<div style="display: inline-block; width: auto; max-width: 100%;">',
    population_html, "</div>\n```\n", sep = "")
```

Counter-example to the assumption of validity:

> NHANES heights are measured in a Mobile Examination Center by trained examiners using a fixed stadiometer, shoes off, standardized instructions. USMC heights are recorded at the recruiting office or at bootcamp intake, with whatever stadiometer is on hand, possibly with shoes on. A systematic offset of a centimeter or two between the two procedures would not be surprising, so our `height` column in the data may not be exactly the same construct as the Preceptor Table's `height`.

Counter-example to the assumption of stability:

> Adult human height has been broadly stable across recent decades for the US young-adult population, but the *intercept* may have drifted upward by a centimeter or two as nutrition has improved. If the male-female height gap (the slope we care about most) had also drifted, the model's `sexMale` coefficient --- estimated on 2005-era data --- would not be the right coefficient for 2026.

Counter-example to the assumption of representativeness:

> Data &rarr; Population: NHANES is close to a probability sample of the US population, but the educational cut we use drops the survey weights, and our 50-row sample deliberately over-represents males (40 vs. 10) to make the standard-error contrast visible. 
  
> Population &rarr; Preceptor Table: USMC recruits are self-selected volunteers who pass a fitness screen, so they are very likely taller on average than the general NHANES-sampled population of the same age. Our model probably understates true recruit height.

Counter-example to the assumption of unconfoundedness:

 > None. Unconfoundedness applies only to causal questions; ours is predictive. We are not estimating the causal effect of sex on height.

What probability family should we use?

> Normal Distribution

Why should we use that probability family?

> The Normal distribution is the standard choice for continuous outcomes.

What is the link function for our outcome?

> Our model will be a linear regression.

## Courage

The abstract data generating mechanism coming out of Justice is:

$$Y = \beta_0 + \beta_1 X_1 + \cdots + \beta_n X_n + \epsilon, \quad \epsilon \sim N(0, \sigma^2)$$


**Candidate model 1: height as a function of sex.**

```{r}
#| label: candidate-model-sex
#| echo: true

fit_sex <- linear_reg(engine = "lm") |>
  fit(height ~ sex, data = recruits)

fit_sex |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 1)))
```

The slope on `sexMale` is about 8.9 cm with a 95% confidence interval that excludes zero -- comparing two recruits differing only in sex, the male recruit is expected to be about 9 cm taller.

**Candidate model 2: height as a function of age.**

```{r}
#| label: candidate-model-age
#| echo: true

fit_age <- linear_reg(engine = "lm") |>
  fit(height ~ age, data = recruits)

fit_age |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 2)))
```

The slope on `age` is indistinguishable from zero, confirming what the EDA already showed: adult height has stopped changing in this age window, so age adds no predictive signal on its own.

**Candidate model 3: height as a function of sex and age.**

```{r}
#| label: candidate-model-sex-age
#| echo: true

fit_sex_age <- linear_reg(engine = "lm") |>
  fit(height ~ sex + age, data = recruits)

fit_sex_age |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 2)))
```

Adding `age` to the sex-only model leaves the `sexMale` coefficient essentially unchanged and the `age` coefficient still indistinguishable from zero. This is what we'd expect since age and sex are roughly uncorrelated in our sample, and age has no real relationship with height in this window.

**Chosen model.** Because adding `age` does not change the answer to our question, the simpler model is preferable -- it is easier to explain, just as accurate, and avoids implying age matters when it doesn't:

```{r}
#| label: chosen-model
#| echo: true

fit_recruits <- linear_reg(engine = "lm") |>
  fit(height ~ sex, data = recruits)

fit_recruits
```

$$\widehat{\text{Height}} = 165.9 + 8.9 \cdot \text{sexMale}$$

`sexMale` is a 0/1 dummy: 1 for male, 0 for female. The intercept (165.9) is the expected height of a female recruit; adding the slope (8.9) gives the expected male recruit height of 174.8 cm. **This is our data generating mechanism.**

**Model checking.** A simple sanity check: the distribution of fitted values should resemble the distribution of actual heights in central tendency, even though the model only produces two distinct predicted values.

```{r}
#| label: model-check-fitted-vs-actual
#| echo: true

fitted_df <- tibble(
  value = c(recruits$height, predict(extract_fit_engine(fit_recruits))),
  source = c(rep("Actual", nrow(recruits)), rep("Fitted", nrow(recruits)))
)

fitted_df |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Actual vs. Fitted Heights",
    subtitle = "The fitted distribution is a two-point mass; the actual distribution is wider",
    x = "Height (cm)",
    y = NULL,
    fill = "Source"
  ) +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```

```{r}
#| label: model-check-residuals
#| echo: true

augment(fit_recruits, new_data = recruits) |>
  ggplot(aes(x = sex, y = .resid)) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  labs(
    title = "Residuals by Sex",
    subtitle = "Residuals scatter evenly around zero in both groups -- no obvious misfit",
    x = "Sex",
    y = "Residual (cm)"
  ) +
  theme_minimal()
```

The two distributions agree on central tendency and rough shape but differ on spread, which we expect from a model using only one binary covariate. The residual plot shows no systematic pattern by sex -- the model is not over- or under-predicting either group on average, which is the main thing we want to check before trusting the predictions in Temperance.

## Temperance


**Predicted values.** `predictions()` returns the model's fitted height for every recruit in our sample.

```{r}
#| label: predictions-raw
#| echo: true

predictions(fit_recruits) |>
  as_tibble() |>
  select(sex, estimate, conf.low, conf.high) |>
  head(6)
```

**Average predictions.** We don't care about any one recruit's prediction -- we care about the average prediction for each sex, which is exactly what `avg_predictions()` gives us, grouped by `sex`.

```{r}
#| label: average-predictions
#| echo: true

avg_predictions(fit_recruits, by = "sex")
```

The expected height of a female recruit is about 165.9 cm (95% CI roughly 161.8 to 170.0), and a male recruit about 174.8 cm (95% CI roughly 172.5 to 177.0).

**Comparison between groups.** `avg_comparisons()` directly answers "how much taller, on average, are male recruits than female recruits?" -- the same number as the regression slope, but computed and labeled as a comparison rather than read off a coefficient table.

```{r}
#| label: avg-comparison-sex
#| echo: true

avg_comparisons(fit_recruits, variables = "sex")
```

The estimate (about 8.9 cm) and its confidence interval are exactly what `avg_predictions()` implies by subtraction, but `avg_comparisons()` does the subtraction and the uncertainty propagation for us -- useful as soon as we want a single number with a single confidence interval for "the difference," rather than two separate intervals we'd have to compare by eye.

**Visualization of predictions.**

```{r}
#| label: prediction-plot
#| echo: true

avg_preds <- avg_predictions(fit_recruits, variables = "sex")
taller_sex <- avg_preds$sex[which.max(avg_preds$estimate)]
height_diff <- round(diff(range(avg_preds$estimate)), 1)

plot_predictions(fit_recruits, condition = "sex", draw = FALSE) |>
  ggplot(aes(x = sex, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high), linewidth = 0.7) +
  labs(
    title = "Predicted Average Height of USMC Recruits by Sex",
    subtitle = paste0(taller_sex, " recruits are predicted to be about ", height_diff, " cm taller, on average"),
    x = "Sex",
    y = "Predicted Height (cm)",
    caption = "Source: 50-row teaching sample from NHANES via primer.tutorials::recruits"
  ) +
  theme_minimal(base_size = 13)
```

**Interpretation.** Comparing two recruits differing only in sex, the male recruit is expected to be about 8.9 cm taller, 95% CI roughly [4.3, 13.4] cm. Notice the language throughout: *comparison*, not *cause*, *raise*, or *change* -- `sex` is a covariate in a predictive model, not a treatment.

**Uncertainty and practical significance.** The female confidence interval is visibly wider than the male interval -- about twice as wide. The standard error of a group mean shrinks with the square root of sample size, so the 40-versus-10 split in our data directly produces this asymmetry. Practically, an 8.9 cm difference is large enough to matter for uniform sizing: it's roughly the gap between two adjacent standard sizes in most clothing systems, so treating male and female recruits identically when ordering uniforms would be a mistake.

**Additional quantities of interest.** The average height answers the specific question we asked, but a campaign-quality answer for the Quartermaster also wants:

- The **maximum** height expected among next year's recruits, to make sure the largest uniform size is in stock.
- The **10th and 90th percentiles** of height within each sex, to size the extra-small and extra-large orders.

Both are available from the same fitted DGM by simulation (drawing many synthetic recruits from the fitted Normal distribution and summarizing), even though we have not worked through the simulation mechanics here.

**Remaining uncertainty.** The reported confidence intervals capture only sampling uncertainty under the model's assumptions. They do not capture uncertainty about whether those assumptions themselves are correct.

**Possible sources of bias.** Three things are likely wrong with our answer. 

> 1) USMC recruits are taller than the general NHANES-sampled population because the fitness screen filters out shorter applicants and recruits are self-selected -- our model probably understates true recruit height by a centimeter or two.
  
> 2) Our sample is small, especially for females (10 rows), so the female interval -- while correctly computed under the model -- doesn't reflect the extra uncertainty from the validity and representativeness gaps discussed in Justice; a cautious analyst would widen it further. 
  
> 3) The world is always more uncertain than our models would have us believe: even with every assumption exactly right, the reported interval is only the *Preceptor's Posterior* -- the best posterior achievable under our assumptions -- not the truth.

