---
title: Models
author: David Kane and Anish Talla
tutorial:
  id: models
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #06 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt) 

library(tidyverse)
library(tidymodels)
library(marginaleffects)
library(broom)

library(easystats)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 600, 
        tutorial.storage = "local") 

set.seed(10)
poll_data <- tibble(biden = as.factor(c(rep("Yes", 655), 
                                        rep("No", 904)))) |> 
        slice_sample(prop = 1)

fit_biden <- logistic_reg() |> 
  fit(biden ~ 1, data = poll_data)
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- DK: Put the data somewhere, either in primer.data as a tibble or just a csv in raw data which can be read in. Probably want a rds so that biden is a factor. -->

## Introduction
### 

This tutorial supports [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). 

The world confronts us. Make decisions we must.

Imagine you are the chair of the Democratic National Committee in early 2024. To which political contests should you devote your time and money? You could spend most of your money on the presidential campaign or, instead, devote those resources toward Congressional races. It would be good to know the voter support for different candidates. There are many decisions to make.

```{r}
# Edit the following tibbles and footnotes, look at the vignette for more details
  
p_tibble <- tibble::tribble(
  ~`Patient ID`, ~`Visit Number`, ~`Recovery Score`, ~`Drug Dosage Group`, ~`Baseline Health Score`,
  "..."        , "..."          , "..."            , "..."               , "..."                   ,
  "..."        , "..."          , "..."            , "..."               , "..."                   ,
  "..."        , "..."          , "..."            , "..."               , "..."                   
)

d_tibble <- tibble::tribble(
  ~`Patient ID`, ~`Visit Number`, ~`Recovery Score`, ~`Drug Dosage Group`, ~`Baseline Health Score`,
  "..."        , "..."          , "..."            , "..."               , "..."                   ,
  "..."        , "..."          , "..."            , "..."               , "..."                   ,
  "..."        , "..."          , "..."            , "..."               , "..."                   
)
  
pre_title_footnote <- "..."
pre_units_footnote <- "..."
pre_outcome_footnote <- "..."
pre_treatment_footnote <- "..."
pre_covariates_footnote <- "..."

pop_title_footnote <- "..."
pop_units_footnote <- "..."
pop_outcome_footnote <- "..."
pop_treatment_footnote <- "..."
pop_covariates_footnote <- "..."
```

```{r}
# This code chunk will generate the Preceptor Table
  
p_tibble_full <- expand_input_tibble(list(p_tibble), "preceptor")

gt::gt(p_tibble_full) |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span", columns = c(`Patient ID`, `Visit Number`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span", columns = c(`Recovery Score`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span", columns = c(`Drug Dosage Group`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span", columns = c(`Baseline Health Score`, "More")) |>
  gt::cols_align(align = "center", columns = gt::everything()) |>
  gt::cols_align(align = "left", columns = c(`Patient ID`)) |>
  gt::cols_width("Patient ID" ~ gt::px(80), "Visit Number" ~ gt::px(100), "Recovery Score" ~ gt::px(120), "Drug Dosage Group" ~ gt::px(120), "Baseline Health Score" ~ gt::px(136), "More" ~ gt::px(168), "Patient ID" ~ gt::px(60)) |>
  gt::cols_label(More = "...") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_footnote(footnote = pre_title_footnote, locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote, locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote, locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_treatment_footnote, locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = pre_covariates_footnote, locations = gt::cells_column_spanners(spanners = "covariates_span"))
```

```{r}
# This code chunk will generate the Population Table
    
data_tibble <- dplyr::bind_rows(
  d_tibble[1:2, , drop = FALSE],  
  d_tibble[1, , drop = FALSE] |> dplyr::mutate(dplyr::across(dplyr::everything(), ~ "...")),  
  d_tibble[3, , drop = FALSE]     
) |>
  dplyr::mutate(Source = "Data", .before = 1)

preceptor_tibble <- p_tibble_full |>
  dplyr::select(-More) |>
  dplyr::mutate(Source = "Preceptor", .before = 1)

empty_row <- data_tibble[1, , drop = FALSE]
empty_row[,] <- "..."

population_tibble <- dplyr::bind_rows(
  empty_row,              # Row 1: blank
  data_tibble,            # Rows 2-5: 4 data rows (3rd is blank)
  empty_row,              # Row 6: blank  
  preceptor_tibble,       # Rows 7-10: 4 preceptor rows (3rd is blank)
  empty_row               # Row 11: blank
)

population_tibble$More <- "..."

gt::gt(population_tibble) |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span", columns = c(`Patient ID`, `Visit Number`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span", columns = c(`Recovery Score`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span", columns = c(`Drug Dosage Group`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span", columns = c(`Baseline Health Score`, "More")) |>
  gt::cols_align(align = "center", columns = gt::everything()) |>
  gt::cols_align(align = "left", columns = c(`Patient ID`)) |>
  gt::cols_width("Source" ~ gt::px(80), "Patient ID" ~ gt::px(100), "Visit Number" ~ gt::px(120), "Recovery Score" ~ gt::px(120), "Drug Dosage Group" ~ gt::px(136), "Baseline Health Score" ~ gt::px(168), "More" ~ gt::px(60)) |>
  gt::cols_label(More = "...") |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_footnote(footnote = pop_title_footnote, locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote, locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote, locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_treatment_footnote, locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = pop_covariates_footnote, locations = gt::cells_column_spanners(spanners = "covariates_span"))
```

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

Why do we ask this, and a score more other questions, in each tutorial? Because the best way to (try to) ensure that students remember these concepts more than a few months after the course ends is [spaced repetition](https://en.wikipedia.org/wiki/Spaced_repetition), although we focus more on the repetition than on the spacing.

### Exercise 2

Create a Github repo called `models`. Make sure to click the "Add a README file" check box.

Connect the repo to a project on your computer using `File -> New Folder from Git ...`. Make sure to select the "Open in a new window" box. 

You need two Positon windows: this one for running the tutorial and the one you just created for writing your code and interacting with the Console.

Select `File -> New File -> Quarto Document ...`. Provide a title -- `"Biden Polling"` -- and an author (you). Render the document and save it as `analysis.qmd`.

Create a `.gitignore` file with `analysis_files` on the first line and then a blank line. Save and push.

In the Console, run:

```         
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the Console.

CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

Professionals keep their data science work in the cloud because laptops fail.

### Exercise 3

In your QMD, load the **tidyverse** package in a new code chunk. Render the file.

Notice that the file does not look good because the code is visible and there are annoying messages. To take care of this, add `#| message: false` to remove all the messages in this `setup` chunk. Also add the following to the YAML header to remove all code echos from the HTML:

```         
execute: 
  echo: false
```

In the Console, run:

```         
show_file("analysis.qmd", start = -5)
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

### 

Render again. Everything looks nice, albeit empty, because we have added code to make the file look better and more professional.

### Exercise 4

Place your cursor in the QMD file on the `library(tidyverse)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(tidyverse)` to be copied down to the Console and then executed. 

CP/CR.

```{r introduction-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The data we are interested in using does not reside in any R package. So we need to create it ourselves, both in the QMD and in the Console.

### Exercise 5

Copy/paste this code into your `setup` code chunk in your QMD.

````
poll_data <- tibble(biden = as.factor(c(rep("Yes", 655), 
                                        rep("No", 904)))) |> 
        slice_sample(prop = 1)
````        


Place your cursor in the QMD file next to `poll_data`. Use `Cmd/Ctrl + Enter` to execute. It will automatically execute all three lines in the Console.

This work flow --- writing things in the QMD so that you have a permanent copy and then executing them in the Console with `Cmd/Ctrl + Enter` --- is the most common approach to data science. 

There is QMD World and R World. It is your responsibility to keep them in sync.

CP/CR.

```{r introduction-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The polling data comes from a YouGov survey of 1,559 U.S. adult citizens conducted March 10-12, 2024. The tibble `poll_data` --- which you have just created, has also been loaded up in the workspace of this tutorial --- contains that data.

## Wisdom
### 

*Wonder is the beginning of wisdom.* - Socrates

With this project, we’re not trying to predict every detail about the 2024 election. That would be impossible. Instead, we’re focusing on a question that we can actually begin to answer: What proportion of all votes will be cast for Joe Biden? 

The general question we are interested in is the future results of the 2024 election, as seen from a survey conducted March of 2024.

Our question:

> *What proportion of all votes will be cast for Joe Biden in the 2024 election?*

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
    message = "Wisdom requires a question, the creation of a Preceptor Table, and an examination of our data.",
    answer(NULL, correct = TRUE),
    allow_retry = FALSE,
    incorrect = NULL,
    rows = 3)
```

*The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* -- John W. Tukey

### Exercise 2

Run `glimpse(poll_data)`. 

```{r wisdom-2, exercise = TRUE}

```

```{r wisdom-2-hint-1, eval = FALSE}
glimpse(poll_data)
```

```{r wisdom-2-test, include = FALSE}
glimpse(poll_data)
```

###

For this poll, the specific question asked was: 

"If an election for president were going to be held now and the Democratic nominee was Joe Biden and the Republican nominee was Donald Trump, would you vote for…"  

For our analysis, we focus only on whether respondents indicated they would vote for Biden (coded as "Yes") or not (coded as "No").

### Exercise 3

Which variable in `poll_data` should we use as our outcome variable? (A trick question since there is only one variable!)


```{r wisdom-3}
question_text(NULL,
  message = "biden: A binary outcome variable indicating whether a respondent supports Joe Biden ('Yes') or not ('No') in the 2024 election.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

### 

We will use `biden` as our outcome variable.

```{r}
poll_data |> 
  ggplot(aes(x = factor(biden))) +
  geom_bar(fill = "steelblue") +
  labs(
    title = "Distribution of Support for Joe Biden",
    subtitle = "About 42% of respondents said they would vote for Biden",
    x = "Would Vote for Biden (1 = Yes, 0 = No)",
    y = "Number of Respondents"
  ) +
  theme_minimal(base_size = 14)
```

### Exercise 4

Let's imagine a brand new variable which **does not exists** in the data. This variable should be binary, meaning that it only takes on one of two values. It should also, at least in theory, be manipulable. In other words, if the value of the variable is "3," or whatever, then it generates one potential outcome and if it is "9," or whatever, it generates another potential outcome.

Describe this imaginary variable and how might we manipulate its value.

```{r wisdom-4}
question_text(NULL,
  message = "Imagine a variable called `trump_ad` which has a value of `1` if the person was shown a pro-Trump advertisement and `0` if they were not shown the ad. We can manipulate this variable by deciding, either randomly or otherwise, whether or not a specific individual sees the advertisement.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Any dataset can be used to construct a causal model as long as there is at least one covariate that we can, at least in theory, manipulate. It does not matter whether or not anyone did, in fact, manipulate it. 

### Exercise 5

Given our (imaginary) treatment variable `trump_ad`, how many potential outcomes are there for each unit? Explain why.

```{r wisdom-5}
question_text(NULL,
  message = "There are two potential outcomes because the treatment variable `trump_ad` takes on two possible values: `1` (exposure to a pro-Trump advertisement) versus `0` (no such exposure).",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The same dataset can be used to create, separately, lots and lots of different models, both causal and predictive. We can just use different outcome variables and/or specify different treatment variables. This is a *conceptual* framework we apply to the data. It is never inherent in the data itself.

### Exercise 6

In a few sentences, specify the two different values for the imaginary treatment variable `trump_ad`, for a single unit, guess at the potential outcomes which would result, and then determine the causal effect for that unit given those guesses.

```{r wisdom-6}
question_text(NULL,
  message = "For a given voter, the treatment variable would be **exposure to a pro-Trump advertisement** or **no exposure**. If exposed, her support for `biden` is `Yes`; if not, it's `No`. However, the causal effect is simply the difference between these two potential outcomes, which would be the difference between 'Yes' and 'No' for this individual.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

A causal effect is defined as the *difference* between two potential outcomes. Keep two things in mind.

First, *difference* does not necessarily mean *subtraction*. Many potential outcome are not numbers. For example, it makes no sense to subtract a potential outcome, like who you would vote for if you saw a Facebook ad, from another potential outcome, like who you vote for if you did not see the ad.

Second, even in the case of numeric outcomes, you can’t simply say the effect is 10 without specifying the order of subtraction. There is, perhaps, a default sense in which the causal effect is defined as potential outcome under treatment minus potential outcome under control.

### Exercise 7

Let's consider a *predictive model*. Which variable, even if it is not in `poll_data`, do you think might have an important connection to `biden`?

```{r wisdom-7}
question_text(NULL,
  message = "One key covariate to explore might be `age`, since voter preferences often vary by age.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

### 

With a predictive model, there is only one outcome for each individual unit. There are not two potential outcomes because we are not considering any of the covariates to be a treatment variable. We assuming that the values of all covariates are "fixed." 

There are no "treatments" in predictive models. There are only covariates.

### Exercise 8

Specify two different groups of **voters** which have different values for **age** and which might have different average values for the `biden` outcome.

```{r wisdom-8}
question_text(NULL,
  message = "Consider two groups, the first with a value for age of **25** and the second with a value of **65**. Those two groups might have different average values for the outcome variable `biden`.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

In predictive models, do not use words like "cause," "influence," "impact," or anything else which suggests causation. The best phrasing is in terms of "differences" between groups of units with different values for a covariate of interest.

### Exercise 9

Define a Preceptor Table.

```{r wisdom-9}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantities of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The Preceptor Table does not include all the covariates which you will eventually include in your model. It only includes covariates which you need to answer your question.

### Exercise 10

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-10}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

This problem is predictive so there is only one outcome. 

### Exercise 11

What are the units for this problem?

```{r wisdom-11}
question_text(NULL,
  message = "The units are all voting-eligible citizens in October/November 2024.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Specifying the Preceptor Table forces us to think clearly about the units and outcomes implied by the question. The resulting discussion sometimes leads us to modify the question with which we started. No data science project follows a single direction. We always backtrack. There is always dialogue.

We model units, but we only really care about aggregates.

### Exercise 12

What is the outcome variable for this problem?

```{r wisdom-12}
question_text(NULL,
  message = "The outcome is whether a person votes for Joe Biden: 'Yes' or 'No'.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Looking at the question, having the outcome as "the candidate for whom the vote was cast" (or something similar) is too broad. We don't actually need to know the name of the candidate. We just need to know if the vote was cast for Biden or not. The name of the candidate, if it was not Biden, is irrelevant to our question. So, we can represent the outcome as 'No' (did not vote for Biden) and 'Yes' (did vote for Biden).

### Exercise 13

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r wisdom-13}
question_text(NULL,
  message = "A plausible covariate is `age`, as older people may have different voting preferences than younger people.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The term "covariates" is used in at least three ways in data science. First, it is all the variables which might be useful, regardless of whether or not we have the data. Second, it is all the variables for which we have data. Third, it is the set of variables in the data which we end up using in the model.

### Exercise 14

What are the treatments, if any, for this problem?

```{r wisdom-14}
question_text(NULL,
  message = "There are no treatments.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

### 

Remember that a treatment is just another covariate which, for the purposes of this specific problem, we are assuming can be manipulated, thereby, creating two or more different potential outcomes for each unit.

### Exercise 15

What moment in time does the Preceptor Table refer to?

```{r wisdom-15}
question_text(NULL,
  message = "When votes are cast in the 2024 election.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

A Preceptor Table can never really refer to an exact instant in time since nothing is instantaneous in this fallen world.

In almost all practical problems, the data was gathered at a time other than that to which the Preceptor Table refers.

> *You can never look at the data too much. -- Mark Engerman*

### Exercise 16

Define a causal effect. (Note that the model in this tutorial is predictive, not causal. We just want to make sure you understand what a causal model is.)

```{r wisdom-16}
question_text(NULL,
  message = "A causal effect is the difference between two potential outcomes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

According to the Rubin Causal Model, there must be two (or more) potential outcomes for any discussion of causation to make sense. This is simplest to discuss when the treatment only has two different values, thereby generating only two potential outcomes. 

### Exercise 17

What is the fundamental problem of causal inference?

```{r wisdom-17}
question_text(NULL,
  message = "The fundamental problem of causal inference is that we can only observe one potential outcome.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

If the treatment variable is continuous (like income), then there are lots and lots of potential outcomes, one for each possible value of the treatment variable. 

### Exercise 18

How does the motto "No causation without manipulation" apply in this problem?

```{r wisdom-18}
question_text(NULL,
  message = "This tutorial uses a predictive model, so the motto 'No causation without manipulation' does not apply directly. However, you can still reflect on whether it's theoretically possible to manipulate the covariates or treatment to create a causal scenario.?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Always consider if issues like manipulation and causation apply to both your Preceptor Table and your actual data.

### Exercise 19

Describe in words the Preceptor Table for this problem.

```{r wisdom-19}
question_text(NULL,
  message = "The Preceptor Table consists of rows representing individual voters in the 2024 U.S. presidential election. The outcome column shows whether or not the individual voted for Biden. There is also at least one covariate column, such as party affiliation.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The Preceptor Table for this problem looks something like this:

```{r}
tibble(
  voter_ID = c("1", "2", "...", "10", "11", "...", "1,559"),
  biden = c("0", "0", "...", "1", "0", "...", "1")
) |>
  gt() |>
  tab_header(title = "Preceptor Table") |>
  cols_label(
    voter_ID = md("Voter ID"),
    biden = md("Voted for Biden")
  ) |>
  tab_style(cell_borders(sides = "right"),
            location = cells_body(columns = c(voter_ID))) |>
  tab_style(
    style = cell_text(align = "left", v_align = "middle", size = "large"),
    locations = cells_column_labels(columns = c(voter_ID))
  ) |>
  cols_align(align = "center", columns = everything()) |>
  cols_align(align = "left", columns = c(voter_ID)) |>
  fmt_markdown(columns = everything()) |>
  tab_spanner(label = "Outcome", columns = c(biden))
```

Like all aspects of a data science problem, the Preceptor Table evolves as we continue our work.

### Exercise 20

Over the course of this tutorial, we will be creating a summary paragraph. The purpose of this exercise is to write the first two sentences of that paragraph.

The first sentence is a general statement about the overall topic, mentioning both the general class of the outcome variable and of at least one of the covariates. It is **not** connected to the initial "Imagine that you are XX" which set the stage for this project. That sentence can be rhetorical. It can be trite, or even a platitude. The purpose of the sentence is to let the reader know, gently, about our topic.

The second sentence does two things. First, it introduces the data source. Second, it introduces the specific question. The sentence can't be that long. Important aspects of the data include when/where it was gathered, how many observations it includes and the organization (if famous) which collected it.

Type your two sentences below.

```{r wisdom-20}
question_text(NULL,
  message = "Public opinion polls measure voter sentiment during political campaigns. Using March 2024 YouGov survey data from 1,559 U.S. adults, we estimate the proportion of votes Joe Biden will receive in the November election.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Read our answer. It will not be the same as yours. You can, if you want, change your answer to incorporate some of our ideas. Do not copy/paste our answer exactly. Add your two sentences, edited or otherwise, to your QMD, `Cmd/Ctrl + Shift + K`, and then commit/push.

## Justice
### 

*Justice is truth in action.* - Benjamin Disraeli

### Exercise 1

In your own words, name the five key components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
  message = "Justice concerns the Population Table and the four assumptions which underlie it: validity, stability, representativeness, and unconfoundedness.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Justice is about concerns that you (or your critics) might have, reasons why the model you create might not work as well as you hope. 

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the dataset and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Validity is always about the *columns* in the Preceptor Table and the data. Just because columns from these two different tables have the same *name* does not mean that they are the same *thing*.

### Exercise 3

Provide one reason why the assumption of validity might not hold for the outcome (Biden support). Use the words "column" or "columns" in your answer.

```{r justice-3}
question_text(NULL,
  message = "Validity may fail because the poll's 'would vote for Biden' column measures intentions months before the election, while the Preceptor Table's 'voted for Biden' column captures actual behavior - these can differ substantially.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

In order to consider the Preceptor Table and the data to be drawn from the same population, the columns from one must have a *valid correspondence* with the columns in the other. Validity, if true (or at least reasonable), allows us to construct the Population Table, which is the first step in Justice.

Because we control the Preceptor Table and, to a lesser extent, the original question, we can adjust those variables to be “closer” to the data that we actually have. This is another example of the iterative nature of data science. If the data is not close enough to the question, then we check with our boss/colleague/customer to see if we can modify the question in order to make the match between the data and the Preceptor Table close enough for validity to hold.

Despite these potential problems, we will assume that validity holds since it, mostly (?), does.

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The Population Table is almost always much bigger than the combination of the Preceptor Table and the data because, if we can really assume that both the Preceptor Table and the data are part of the same population, than that population must cover a broad universe of time and units since the Preceptor Table and the data are, themselves, often far apart from each other.

```{r}
tibble(source = c("...", "...", "...","...",  
                  "Data", "Data", "Data", "Data", "...",
                  "...", "...", "...","...", 
                  "Preceptor Table", "Preceptor Table", "Preceptor Table", "...",
                  "...", "...", "..."),
       time = c("February 2024", "February 2024", "February 2024", "...",
                "March 2024", "March 2024", "March 2024", "March 2024", "...",
                "October 2024", "October 2024", "October 2024", "...",
                "November 2024", "November 2024", "November 2024", "...",
                "December 2024", "December 2024", "December 2024"),
       id = c("1", "200", "976", "...",
              "1", "200", "...", "1559", "...",
              "1", "200", "2025", "...",
              "1", "200", "2078", "...",
              "1", "200", "2300"),
       biden = c("?", "?", "?", "...",
              "0", "1", "...", "1", "...",
              "?", "?", "?", "...",
              "1", "0", "1", "...",
              "?", "?", "?")) |>
  gt() |>
  cols_label(source = md("Source"),
             time = md("Time"),
             id = md("ID"),
             biden = md("Biden")) |>
  tab_style(cell_borders(sides = "right"),
            location = cells_body(columns = c(source))) |>
  tab_style(style = cell_text(align = "left", v_align = "middle", size = "large"),
            locations = cells_column_labels(columns = c(source))) |>
  cols_align(align = "center", columns = everything()) |>
  cols_align(align = "left", columns = c(source)) |>
  fmt_markdown(columns = everything())
```

### Exercise 5

In your own words, define the assumption of "stability" when employed in the context of data science.

```{r justice-5}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Stability is all about *time*. Is the relationship among the columns in the Population Table stable over time? In particular, is the relationship --- which is another way of saying "mathematical formula" --- at the time the data was gathered the same as the relationship at the (generally later) time referenced by the Preceptor Table. 

### Exercise 6

Provide one reason why the assumption of stability might not be true in this case. 

```{r justice-6}
question_text(NULL,
  message = "If Biden's support changes over time, than the stability assumption does not hold.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Stability is about the parameters: $\beta_0$, $\beta_1$ and so on. Stability means these parameters are the same in the data as they are in the population as they are in the Preceptor Table.

### Exercise 7

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: `Data -> Population -> Preceptor Table`. In your own words, define the assumption of "representativeness" when employed in the context of data science.

```{r justice-7}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Ideally, we would like both the Preceptor Table *and* our data to be random samples from the population. Sadly, this is almost never the case.

### Exercise 8

We do not use the data, directly, to estimate missing values in the Preceptor Table. Instead, we use the data to learn about the overall population. Provide one reason, involving the relationship between the data and the population, for why the assumption of representativeness might not be true in this case. 

```{r justice-8}
question_text(NULL,
  message = "The poll respondents might systematically differ from the overall population in their political engagement - frequent voters are overrepresented in surveys while occasional voters who decide late are harder to reach, potentially skewing our estimates of Biden's support.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

The reason that representativeness is important is because, when it is violated, the estimates for the model parameters might be biased. 

### Exercise 9

We use information about the population to make inferences about the Preceptor Table. Provide one reason, involving the relationship between the population and the Preceptor Table, for why the assumption of representativeness might not be true in this case.

```{r justice-9}
question_text(NULL,
  message = "The Preceptor Table only includes actual voters, while the population contains all eligible voters - groups like young people and minorities often vote at different rates than their share of the population, creating a representativeness gap.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Stability looks across time periods. Representativeness looks within time periods, for the most part. 

### Exercise 10

In your own words, define the assumption of "unconfoundedness" in the context of data science.

```{r justice-10}
question_text(NULL,
  message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates. This assumption is only relevant for casual models.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

This assumption is only relevant for causal models. We describe a model as "confounded" if this is not true. The easiest way to ensure unconfoundedness is to assign treatment randomly.

### Exercise 11

A statistical model consists of two parts: the *probability family* and the *link function*. The probability family is the probability distribution which generates the randomness in our data. The link function is the mathematical formula which links our data to the unknown parameters in the probability distribution. 

Add `library(tidymodels)` to the QMD file.

Place your cursor in the QMD file on the `library(tidymodels)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(tidymodels)` to be copied down to the Console and then executed. 

CP/CR.

```{r justice-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The probability family is determined by the outcome variable $Y$. 

Since $Y$ is a binary variable (with exactly two possible values), the probability family is Bernoulli.

$$Y \sim \text{Bernoulli}(\rho)$$

where $\rho$ is the probability that one of the two possible values --- conventionally referred to as `1` or `TRUE` --- occurs. By definition, $1 - \rho$ is the probability of the other value.

### Exercise 12

Add `library(broom)` to the QMD file.

Place your cursor in the QMD file on the `library(broom)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(broom)` to be copied down to the Console and then executed. 

CP/CR.

```{r justice-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The link function, the basic mathematical structure of the model, is (mostly) determined by the type of outcome variable.

For a binary outcome variable, we use a log-odds model:

$$
\log\left[ \frac { \rho }{ 1 - \rho } \right] = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \ldots 
$$

### Exercise 13

Use AI to come up with a $\LaTeX$ representation of the mathematical structure of the model, with $Y$ as the dependent variable and $X_1$, $X_2$ and so on as the independent variables. This version will not have the values of any parameters since we have not, yet, estimated them. Confirm that the $\LaTeX$ code works by placing it in your QMD and then rendering. Paste that $\LaTeX$ code below.

No worries if you have trouble with this question!

```{r justice-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Our answer:

$$P(Y = 1) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n)}}$$

with $Y \sim \text{Bernoulli}(\rho)$ where $\rho$ is the probability above.

Which we created with $\LaTeX$ code that looks like this:

````
$$P(Y = 1) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n)}}$$

with $Y \sim \text{Bernoulli}(\rho)$ where $\rho$ is the probability above.
````

This follows the logistic regression form for binary data, where the $\beta$ coefficients represent the effect of predictors on the log-odds of the outcome.

We use generic variables --- $Y$, $X_1$ and so on --- because our purpose is to describe the general mathematical structure of the model, independent of the specific variables we will eventually choose to use.

Having decided on the basic mathematical structure of the model, a choice mostly driven by the distribution of our outcome variable, we now turn toward estimating the model.

### Exercise 14

Write one sentence which highlights a potential weakness in your model. This will almost always be derived from possible problems with the assumptions discussed above. We will add this sentence to our summary paragraph. So far, our version of the summary paragraph looks like this:

> Public opinion polls measure voter sentiment during political campaigns. Using March 2024 YouGov survey data from 1,559 U.S. adults, we estimate the proportion of votes Joe Biden will receive in the November election. 

Of course, your version will be somewhat different.

```{r justice-14}
question_text(NULL,
  message = "Our estimate may be biased if actual voters are not representative of the overall population of registered voters.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

### 

Add a weakness sentence to the summary paragraph in your QMD. You can modify your paragraph as you see fit, but do not copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`, and then commit/push.

## Courage
### 

*Courage is found in unlikely places.* - J.R.R. Tolkien

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
  message = "Courage creates the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Having decided on the basic mathematical structure of the model at the end of *Justice*, a choice mostly driven by the distribution of our outcome variable, we now turn toward estimating the model.

### Exercise 2

Because our outcome variable is **binary**, start to create the model by entering `logistic_reg()`.

```{r courage-2, exercise = TRUE}

```

```{r courage-2-hint-1, eval = FALSE}
logistic_reg()
```

```{r courage-2-test, include = FALSE}
logistic_reg()
```

### 

The default `engine` for `logistic_reg()` is `"glm"`. Since this is what we want, we don't need to specify it. 

The *[tidymodels](https://www.tidymodels.org/)* framework is the most popular one in the R world for estimating models. *[Tidy Modeling with R](https://www.tmwr.org/)* by Max Kuhn and Julia Silge is a great introduction.

### Exercise 3

We want to fit a model with a single categorical predictor. We'll use `biden`, which contains "Yes"/"No" responses, and fit a logistic regression model with it as the outcome variable.

Continue the pipe to `fit(biden ~ 1, data = poll_data).`

```{r courage-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-3-hint-1, eval = FALSE}
logistic_reg() |> 
  fit(..., data = ...)
```

```{r courage-3-test, include = FALSE}
logistic_reg() |> 
  fit(biden ~ 1, data = poll_data)
```

###

Note that, if `biden` were not a factor, this would generate an error. The "glm" engine requires that the outcome variable be a factor.

<!-- DK: Or is it logistic_reg() which requires a factor outcome? -->

### Exercise 4

Continue the pipe with `tidy(conf.int = TRUE)`.

```{r courage-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-4-hint-1, eval = FALSE}
...
  tidy(conf.int = TRUE)
```

```{r courage-4-test, include = FALSE}
logistic_reg() |> 
  fit(biden ~ 1, data = poll_data) |> 
    tidy(conf.int = TRUE)
```

###

In a logistic regression model, the intercept ($\beta_0$) represents the log-odds of voting for Biden when all predictor variables are 0. 

### Exercise 5

Behind the scenes, we’ve created an object called `fit_biden` that stores the model you just fit. Type `fit_biden` and hit "Run Code" to view the results.

```{r courage-5, exercise = TRUE}

```

```{r courage-5-hint-1, eval = FALSE}
fit...
```

```{r courage-5-test, include = FALSE}
fit_biden
```

###

In data science, **we deal with words, math, and code, but the most important of these is code.** We created the mathematical structure of the model and then wrote a model formula in order to estimate the unknown parameters.

### Exercise 6

We need `fit_biden` to exit in R World. Copy/paste this code into the Console and execute it.

````
fit_biden <- logistic_reg() |> 
  fit(biden ~ 1, data = poll_data)
````

CP/CR.

```{r courage-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Just because something exists in the tutorial (or in the QMD) does not mean that it is in the Console. You should be aware of what exists in R World, which is generally called your "workspace."

### Exercise 7

In the Console, load the **[easystats](https://easystats.github.io/easystats/)** package. CP/CR.

```{r courage-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

We don't add **easystats** to the QMD because we are only using it for an interactive check of our fitted model. However, the [easystats ecosystem](https://easystats.github.io/easystats/) has a variety of interesting functions and packages which you might want to explore.

### Exercise 8

In the Console, run `check_predictions(extract_fit_engine(fit_biden))`. CP/CR.

```{r courage-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

The purpose of `check_predictions()` is to compare your actual data (in green) with data that has been simulated from your fitted model, i.e., from your data generating mechanism. If your DGM is reasonable, then data simulated from it should not look too dissimilar from your actual data. Of course, it won't look exactly the same because of randomness, both in the world and in your simulation. But the actual data should be within the range of outcomes that your DGM simulates with `check_predictions()`.

The model uses only an intercept, so it predicts the same probability of Biden support for everyone—just the overall average. 

### Exercise 9

Ask AI to create $\LaTeX$ code for this model, including our variable names and estimates for all the coefficients. Because this is a fitted model, the dependent variable will have a "hat" and the formula will not include an error term.

Add the code to your QMD. `Cmd/Ctrl + Shift + K`.

Make sure the resulting display looks good. For example, you don't want an absurd number of figures to the right of the decimal. If the model is too long, you will need to spread it across several lines. You may need to go back-and-forth with the AI a few times.

Once the $\LaTeX$ code looks good, paste it below.

```{r courage-9}

question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Our formula looks like:

$$
\log\left( \frac{\widehat{p}}{1 - \widehat{p}} \right) = -0.32
$$


It was created with:

````
\widehat{p} = \frac{1}{1 + e^{-(-0.32)}} \approx 0.42
````

Note the differences. First, we have replaced the parameters with our best estimate from the model. Second, we have dropped the error term because this is a formula for predicting the value of our outcome variable. Third, the left-hand side variable is $\widehat{p}$ instead of $p$ or `biden` because this formula generates the estimated percentage of people who support Biden. The “hat” indicates an estimated value. The symbol $\approx 0.42$ means the calculated value is approximately equal to 0.42, or 42%.

**This is our data generating mechanism.**

A data generating mechanism is just a formula, something which we can write down and implement with computer code.

### Exercise 10

Create a new code chunk in your QMD. Add a code chunk option: `#| cache: true.` Copy/paste the code from above for estimating the model into the code chunk, assigning the result to `fit_biden.` 

````
fit_biden <- logistic_reg() |> 
  fit(biden ~ 1, data = poll_data)
````

This includes the call to `fit()` but not the call to `tidy()` because we want the entire fitted model, not just a table of the estimated parameter values.

`Cmd/Ctrl + Shift + K`. By including `#| cache: true` you cause Quarto to cache the results of the chunk. The next time you render your QMD, as long as you have not changed the code, Quarto will just load up the saved fitted object.

At the Console, run:

```
tutorial.helpers::show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r courage-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

To confirm, `Cmd/Ctrl + Shift + K` again. It should be quick.

### Exercise 11

Add `*_cache` to `.gitignore` file. Cached objects are often large. They don't belong on Github.

At the Console, run:

```
tutorial.helpers::show_file(".gitignore")
```

CP/CR.


```{r courage-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Because of the change in your `.gitignore` (assuming that you saved it), the cache directory should not appear in the Source Control panel because Git is ignoring it, as instructed. Commit and push.

### Exercise 12

In the Console, run `tidy()` on `fit_biden` with the argument `conf.int` set equal to `TRUE`. The returns 95% intervals for all the parameters in our model.


```{r courage-12, exercise = TRUE}

```

```{r courage-12-hint-1, eval = FALSE}
tidy(..., conf.int = ...)
```

```{r courage-12-test, include = FALSE}
tidy(fit_biden, conf.int = TRUE)


```

###

`tidy()` is part of the [**broom**](https://broom.tidymodels.org/) package, used to summarize information from a wide variety of models.


### Exercise 13

Create a new code chunk in your QMD. Ask AI to help you make a nice looking table from the tibble which is returned by `tidy()`. You don't have to include all the variables which `tidy()` produces. We often just show the estimate and the confidence intervals.

Insert that code into the QMD.

`Cmd/Ctrl + Shift + K`.

Make sure it works. You might need to add some new libraries, e.g., **[tinytable](https://vincentarelbundock.github.io/tinytable/)**, **[knitr](https://yihui.org/knitr/)**, **[gt](https://gt.rstudio.com/)**, **[kableExtra](https://haozhu233.github.io/kableExtra/)**, **[flextable](https://davidgohel.github.io/flextable/)**, **[modelsummary](https://modelsummary.com/)**, et cetera, to the `setup` code chunk, if you use any functions from these packages, all of which have strengths and weaknesses for making tables.

At the Console, run:

```
tutorial.helpers::show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r courage-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

Here is ours, along with the code:

```{r}
library(knitr)

tidy(fit_biden, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  kable(
    caption = "Estimates and 95% Confidence Intervals for Intercept\nSource: March 2024 YouGov Poll "
  )
```

```
library(knitr)

tidy(fit_biden, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  kable(
    caption = "Estimates and 95% Confidence Intervals for Intercept\nSource: March 2024 YouGov Poll "
  )
```

At the very least, your table should include a title and a caption with the data source. The more you use AI, the better you will get at making tables.

<!-- DK: Not a great table. Really ought to give the probability not some nonsense intercept. -->


### Exercise 14

Add a sentence to your project summary.

Explain the structure of the model. Something like: "I/we model XX [the concept of the outcome, not the variable name], [insert description of values of XX], as a [linear/logistic/multinomial/ordinal] function of XX [and maybe other covariates]."

Recall the beginning of our version of the summary:

> Public opinion polls measure voter sentiment during political campaigns. Using March 2024 YouGov survey data from 1,559 U.S. adults, we estimate the proportion of votes Joe Biden will receive in the November election. Our estimate may be biased if actual voters are not representative of the overall population of registered voters.

```{r courage-14}
question_text(NULL,
  message = "We model voter preference —--- the probability that a respondent supports Biden --— as a logistic function with no predictors, capturing the overall proportion of support in the population. The outcome variable is coded as 'Yes' for Biden supporters and 'No' for non-supporters.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Read our answer. It will not be the same as yours. You can, if you want, change your answer to incorporate some of our ideas. Do not copy/paste our answer exactly. Add your two sentences, edited or otherwise, to the summary paragraph portion of your QMD. `Cmd/Ctrl + Shift + K`, and then commit/push.

## Temperance
###

*Temperance is a tree which as for its root very little contentment, and for its fruit calm and peace.* - Buddha

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
  message = "Temperance uses the data generating mechanism to answer the question with which we began. Humility reminds us that this answer is always a lie. We can also use the DGM to calculate many similar quantities of interest, displaying the results graphically.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Courage gave us the data generating mechanism. Temperance guides us in the use of the DGM — or the “model” — we have created to answer the question(s) with which we began. We create posteriors for the quantities of interest. 

### Exercise 2

Before using the DGM, we should make sure that we can interpret it.

Recall the values for the parameters in our data generating mechanism:

```{r}
fit_biden |>   
  tidy(conf.int = TRUE) |>   
  select(term, estimate, conf.low, conf.high)
```

Interpret the -0.32 value for the estimate of the `(Intercept)`. Feel free to use AI to help with this or any other question.

```{r temperance-2}
question_text(NULL,
  message = "The intercept of -0.32 is the log-odds of supporting Biden. Log-odds are the natural log of odds, and you can convert log-odds to probability with p = 1/(1 + exp(-log-odds)). Here, -0.32 means a probability of about 42%.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Most of the time parameters in a model have no direct relationship with any population value in which we might be interested. This is especially true in complex and/or non-linear models. That is, in those cases, a coefficient like $\beta_0$ does not "mean" anything. But, in simple, small, linear models, it sometimes happens that a parameter does correspond to something real.

### Exercise 3

```{r}
fit_biden |> 
  tidy(conf.int = TRUE) |> 
  select(term, estimate, conf.low, conf.high)
```

Interpret the -0.42 value for the conf.low of the `(Intercept)`.

```{r temperance-3}
question_text(NULL,
  message = "The -0.42 lower bound means we are 95% confident the true log-odds of supporting Biden is at least -0.42, which corresponds to a minimum probability of about 40%.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

We look for two things in the confidence interval. First, does it exclude zero? If not, then we can't be sure if the relationship is positive or negative. Second, does it overlap with the confidence intervals for other dummy columns derived from this variable? If so, then we can't be sure that the ordering as to which comparisons are bigger.

### Exercise 4

```{r}
fit_biden |> 
  tidy(conf.int = TRUE) |> 
  select(term, estimate, conf.low, conf.high)
```

Interpret the -0.22 value for the conf.high of the `(Intercept)`.

```{r temperance-4}
question_text(NULL,
  message = "The -0.22 upper bound suggests the true log-odds of supporting Biden is likely no more than -0.22, which translates to about a 45% probability.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The interpretation of a treatment variable is very different from the interpretation of a standard covariate. The key point is that there is no such thing as a causal (versus predictive) data set nor a causal (versus predictive) R code formula. You can use the same data set (and the same R code!) for both causal and predictive models. The difference lies in the assumptions you make.


### Exercise 5

In the end, we don't really care about parameters, much less how to interpret them. Parameters are *imaginary*, like unicorns. We care about answers to our questions. Parameters are tools for answering questions. They aren't interesting in-and-of themselves. *In the modern world, all parameters are nuisance parameters.* 

Add `library(marginaleffects)` to the QMD file.

Place your cursor in the QMD file on the `library(marginaleffects)` line. Use `Cmd/Ctrl + Enter` to execute that line.

Note that this causes `library(marginaleffects)` to be copied down to the Console and then executed. 

CP/CR.

```{r temperance-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

### 

We should be modest in the claims we make. The posteriors we create are never the “truth.” The assumptions we made to create the model are never perfect. Yet decisions made with flawed posteriors are almost always better than decisions made without them.

### Exercise 6

<!-- DK: Keep this question? -->

What is the general topic we are investigating? What is the specific question we are trying to answer?

```{r temperance-6}
question_text(NULL,
  message = "The general question we are interested in is the future results of the 2024 election, as seen from a survey conducted March 2024. Our specific question: What proportion of all votes will be cast for Joe Biden in the 2024 election?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Data science projects begin with a decision which we face. To make that decision wisely, we would like to have good estimates of many unknown numbers. Yet, in order to make progress, we need to drill down to one specific question. This leads to the creation of a data generating mechanism, which can now be used to answer lots of questions, thus allowing us to explore the original decision more broadly.

### Exercise 7

In the Console, run `predictions()` on `fit_biden` with `type = "prob"`.

`type = "prob"` is required because `fit_biden` is a logistic model so the outputs are **prob**abilities.

CP/CR.

```{r temperance-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

```{r temperance-7-test, include = FALSE}
predictions(fit_biden, type = "prob")
```

### 

`predictions()` returns a data frame with one row for each observation in the dataset used to fit the model. In this case, because there are about 1560 responses in the sample and two possible outcomes (`yes` and `no`), the result has about 3120 rows -- two for each voter. For an intercept-only logistic regression, these probabilities are the same for every observation, but split across the two outcome groups.

### Exercise 8

In the Console, run `plot_predictions()` on `fit_biden` with `type = "prob"` and `condition = "group"`.

`condition = "group"` shows us the Yes's and the No's separately.

CP/CR.

```{r temperance-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

```{r temperance-8-test, include = FALSE}
predictions(fit_biden, type = "prob", newdata = poll_data) |>
  ggplot(aes(x = group, y = estimate)) +
  geom_col()
```

### 

This plot shows predicted support for Biden by group, but it's not very informative. Since the model only includes group, the bars just reflect the estimated support within each group. Showing both is redundant because one is just 1 minus the other. 

### Exercise 9

In the Console, run `avg_predictions()` on `fit_biden` with `type = "prob"`.

CP/CR.

```{r temperance-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 20)
```

```{r temperance-9-test, include = FALSE}
avg_predictions(fit_biden, type = "prob")
```

###

<!-- DK: How does avg_prediction really work? How does it know to average across different values of the outcome variable? -->

`avg_predictions()` does what you might expect, taking the average of each prediction within the two different groups. In this case, that calculation is trivial since the predictions within each group are identical. In more complex models, they won't be.

### Exercise 10

Work interactively within your QMD to create a plot which looks something like this:

```{r}
avg_predictions(fit_biden, type = "prob") |>
  filter(group == "Yes") |>
  ggplot(aes(x = group, 
         y = estimate, 
       ymin = conf.low, 
       ymax = conf.high)) +
  geom_pointrange(color = "#2C3E50", size = 1.0,) +
  labs(
    title = "Support for President Biden in March 2024",
    subtitle = "42% is not a good sign for an incumbent seeking reelection",
    caption = "Source: YouGov poll, March 2024.",
    y = NULL,
    x = NULL
  ) +
  scale_x_discrete(labels = NULL, breaks = NULL) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0.38, 0.46)
  ) 
```

Use AI to help. Your plot does not have to look exactly like ours. 

Note that this plot takes the output of `avg_predictions(fit_biden, type = "prob")`, filters out just the row for `group == "Yes"` and then pipes to `ggplot()`, using `geom_pointrange()`.

Copy the code for your plot here:

```{r temperance-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 20)
```

###

Our code:

```
avg_predictions(fit_biden, type = "prob") |>
  filter(group == "Yes") |>
  ggplot(aes(x = group, 
         y = estimate, 
       ymin = conf.low, 
       ymax = conf.high)) +
  geom_pointrange(color = "#2C3E50", size = 1.0,) +
  labs(
    title = "Support for President Biden in March 2024",
    subtitle = "42% is not a good sign for an incumbent seeking reelection",
    caption = "Source: YouGov poll, March 2024.",
    y = NULL,
    x = NULL
  ) +
  scale_x_discrete(labels = NULL, breaks = NULL) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0.38, 0.46)
  ) 
```

Data science often involves this-back-and-forth style of work. First, we need to make a single chunk of code, in this case, a new plot, work well. This requires interactive work between the QMD and the Console. Second, we need to ensure that the entire QMD runs correctly on its own.

### Exercise 11

Finalize the new graphics code chunk in your QMD.

`Cmd/Ctrl + Shift + K` to ensure that it all works as intended. 

At the Console, run:

```
tutorial.helpers::show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r temperance-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

The model assumes everyone has the same probability of supporting Biden. It’s overly simple, so we must be humble about its limits.

### Exercise 12

Write the last sentence of your summary paragraph. It describes at least one quantity of interest (QoI) and provides a measure of uncertainty about that QoI. (It is OK if this QoI is not the one that you began with. The focus of a data science project often changes over time.)

Our summary paragraph so far:

> Public opinion polls measure voter sentiment during political campaigns. Using March 2024 YouGov survey data from 1,559 U.S. adults, we estimate the proportion of votes Joe Biden will receive in the November election. Our estimate may be biased if actual voters are not representative of the overall population of registered voters. We model voter preference —--- the probability that a respondent supports Biden --— as a logistic function with no predictors, capturing the overall proportion of support in the population. The outcome variable is coded as 'Yes' for Biden supporters and 'No' for non-supporters.

```{r temperance-12}
question_text(NULL,
  message = "We estimate that approximately 42% of Americans will support Joe Biden, with a 95% confidence interval of 39% to 45%.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Add a final sentence to your summary paragraph in your QMD as you see fit, but do not copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`.

### Exercise 13

Write a few sentences which explain why the estimates for the quantities of interest, and the uncertainty thereof, might be wrong. Suggest an alternative estimate and confidence interval, if you think either might be warranted.

```{r temperance-13}
question_text(NULL,
  message = "The estimates might be wrong because the stability assumption is flawed. If things change alot between March and November, then the support for Biden might end up much higher or lower than 42%.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Always go back to your Preceptor Table, the information which, if you had it, would make answering your question easy. In almost all real world cases, the Preceptor Table and the data are fairly different, not least because validity never holds perfectly. So, even a perfectly estimated statistical model is rarely as useful as we might like.

### Exercise 14

Rearrange the material in your QMD so that the order is graphic, followed by the paragraph. Doing so, of course, requires sensible judgment. For example, the code chunk which creates the fitted model must occur before the chunk which creates the graphic. You can keep or discard the math and any other material at your own discretion.

`Cmd/Ctrl + Shift + K` to ensure that everything works.

At the Console, run:

```
tutorial.helpers::show_file("analysis.qmd")
```

CP/CR.

```{r temperance-14}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 30)
```

###

This is the version of your QMD file which your teacher is most likely to take a close look at.

### Exercise 15

Publish your rendered QMD to GitHub Pages. In the Terminal --- not the Console! --- run:

````
quarto publish gh-pages analysis.qmd
````

Copy/paste the resulting URL below.

```{r temperance-15}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Commit/push everything.

### Exercise 16

Copy/paste the URL to your Github repo.

```{r temperance-16}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God. If all our assumptions are correct, then our DGM is true, it accurately describes the way in which the world works. There is no better way to predict the future, or to model the past, than to use it. Sadly, this will only be the case with toy examples involving things like coins and dice. We hope that our DGM is close to the true DGM but, since our assumptions are never perfectly correct, our DGM will always be different. The estimated magnitude and importance of that difference is a matter of judgment.

The world confronts us. Make decisions we must.

## Summary
###

This tutorial covered [Chapter 4: Models](https://ppbds.github.io/primer/models.html) of [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

```{r}
avg_predictions(fit_biden, type = "prob") |>
  filter(group == "Yes") |>
  ggplot(aes(x = group, 
         y = estimate, 
       ymin = conf.low, 
       ymax = conf.high)) +
  geom_pointrange(color = "#2C3E50", size = 1.0,) +
  labs(
    title = "Support for President Biden in March 2024",
    subtitle = "42% is not a good sign for an incumbent seeking reelection",
    caption = "Source: YouGov poll, March 2024.",
    y = NULL,
    x = NULL
  ) +
  scale_x_discrete(labels = NULL, breaks = NULL) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0.38, 0.46)
  ) 
```


Public opinion polls measure voter sentiment during political campaigns. Using March 2024 YouGov survey data from 1,559 U.S. adults, we estimate the proportion of votes Joe Biden will receive in the November election. Our estimate may be biased if actual voters are not representative of the overall population of registered voters. We model voter preference —--- the probability that a respondent supports Biden --— as a logistic function with no predictors, capturing the overall proportion of support in the population. The outcome variable is coded as 'Yes' for Biden supporters and 'No' for non-supporters. We estimate that approximately 42% of Americans will support Joe Biden, with a 95% confidence interval of 39% to 45%.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
