---
title: Recruits
author: David Kane
tutorial:
  id: 05-recruits
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #05 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(tidymodels)
library(primer.tutorials)
library(broom)
library(marginaleffects)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = "90%")
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

fit_recruits <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = recruits)
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports the [Recruits](https://ppbds.github.io/primer/05-recruits.html) chapter of [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

The world confronts us. Make decisions we must.

Imagine that you are a logistics analyst at U.S. Marine Corps Supply, planning next year's bootcamp uniform order. The Quartermaster General has a big-picture goal --- every recruit kitted out at the right size on day one, within a fixed procurement budget --- and trusts you to figure out what to order and in what quantities. Many models would inform the order: distributions of height, weight, chest, foot length; the male-female mix in incoming recruits; the size-tail dynamics that drive how many extra-large uniforms to keep in inventory. This tutorial builds just one of them: the average height of male and female recruits. The number alone won't determine the order, but the height-by-sex split is one good input to the larger forecast. There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

The four virtues are the four sections of every Primer chapter and tutorial: a question and a Preceptor Table (Wisdom), an audit of assumptions (Justice), the model (Courage), and the answer with its uncertainty (Temperance). Recruits is the first time we run that arc end to end, on a 50-row sample of height by sex.

### Exercise 2

You should be working inside a GitHub repo named `recruits`, opened in a Codespace from the [`PPBDS/codespace-starter`](https://github.com/PPBDS/codespace-starter) template. If you are not there yet, please create that repo and open it in a Codespace now --- see the [package README](https://github.com/PPBDS/primer/tree/main/primer.tutorials#working-environments-and-repo-setup) if you need setup instructions or are working locally instead.

Once you're inside the `recruits` repo, create a new Quarto document titled `"Recruits"` with yourself as the author, render it, and save it as `analysis.qmd`.

Create a `.gitignore` file with `analysis_files` on the first line followed by a blank line. Save and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it's probably because you haven't loaded `library(tutorial.helpers)` in the R Terminal.

"CP/CR" is short for *copy/paste the **c**ommand and the **r**esponse* --- both the command you sent to R and the response R gave back.

CP/CR.

```{r introduction-2}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The `recruits` tibble is a 50-row teaching cut of NHANES with 40 male and 10 female young adults aged 18--27; the deliberately uneven split will matter when we get to the precision of group means in Temperance.

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.tutorials)` in a new code chunk. Render the file.

Notice that the file does not look good because the code is visible and there are messages cluttering the output. To fix this, add `#| message: false` to this setup chunk to remove the messages. Also add the following to the YAML header, between the `---` delimiters, to hide the code itself:

```
execute:
  echo: false
```

Render again --- it should look cleaner. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    #| message: false\n    library(tidyverse)\n    library(primer.tutorials)\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Rendering a QMD converts the code and prose into an HTML document that shows only what you want readers to see --- in this case, the eventual plot and summary paragraph, with the code hidden. Professional data-science deliverables almost never show raw code; the code is the path, not the destination.

### Exercise 4

In the R Terminal, type `?recruits` and paste the Description below.

```{r introduction-4}
question_text(NULL,
  message = "A 50-row sample drawn from the National Health and Nutrition Examination Survey (NHANES, via primer.data::nhanes) for the 05-recruits tutorial. Restricted to ages 18--27 and to the columns relevant for the tutorial's question. The split between sexes (40 male, 10 female) is deliberately uneven so the two group means have visibly different standard errors --- a feature the tutorial's Temperance section asks students to notice and explain.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Note that NHANES intentionally oversamples some demographic groups --- older adults, racial and ethnic minorities --- so that each subgroup has enough observations to be studied on its own. The raw NHANES survey is not representative of the general population; the survey ships sampling weights to correct for this. Our `recruits` sample inherits these issues and amplifies them: we drew 40 men and 10 women rather than matching their representation in NHANES. We should flag both as concerns later in Justice.

### Exercise 5

Height of young adults is the broad topic of this tutorial. Given that topic, which variable in `recruits` should we use as our outcome variable?

```{r introduction-5}
question_text(NULL,
  message = "Our outcome variable is `height`. We care about the height of USMC recruits, who are young adults.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

We will use `height` as our outcome variable.

```{r}
#| echo: false
#| message: false
#| warning: false
recruits |>
  ggplot(aes(x = height)) +
  geom_density(fill = "grey70", color = "grey30") +
  labs(
    title = "Height in Our 50-Recruit Sample",
    subtitle = "A broad, asymmetric peak with a heavy left shoulder --- a hint that sex matters for height",
    x = "Height (cm)",
    y = NULL,
    caption = "Source: 50-row sample from NHANES via primer.tutorials::recruits"
  ) +
  theme_minimal() +
  theme(axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())
```

### Exercise 6

Which variable in `recruits` do you think might have an important connection to `height`?

```{r introduction-6}
question_text(NULL,
  message = "One candidate is `sex`. Men and women tend to have systematically different average heights, so a model that includes `sex` should do better than one that does not.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

With a predictive model, each individual unit has only one observed outcome. Predictive models have no "treatments" --- only covariates.

### Exercise 7

Specify two different groups of young adults which have different values for `sex` and which might have different average values for `height`.

```{r introduction-7}
question_text(NULL,
  message = "Consider two groups: young adults with `sex` = Male, and young adults with `sex` = Female. These two groups might have different average heights.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

In predictive models, do not use "cause," "influence," "impact," or anything else which suggests causation. The best phrasing is in terms of "differences" between groups of units with different values for a covariate of interest.

### Exercise 8

Write a predictive question connecting the outcome `height` to `sex`, the covariate of interest.

```{r introduction-8}
question_text(NULL,
  message = "What is the average height of male and female USMC recruits?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The average height by sex is a convenient question --- a single number per sex, easy to interpret, easy to use for ordering uniforms. It isn't the only useful question. A logistics officer might also want the tallest recruit expected next year (so no one arrives without a uniform that fits), or the 10th and 90th percentiles of each sex's height distribution (to know how many small-versus-large sizes to order). We stick with the average question here because it exercises every step of the Cardinal Virtues cleanly; the matching chapter considers the broader family.


## Wisdom
###

Data science starts with some broad questions and a data set which might help us to answer them.

The specific question: *What is the average height of male and female USMC recruits?*

We will work from the National Health and Nutrition Examination Survey (NHANES), conducted by the Centers for Disease Control and Prevention, available in the `nhanes` tibble of the **[primer.data](https://github.com/PPBDS/primer.data)** package.

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
  message = "Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

> *The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* --- John W. Tukey

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The Preceptor Table does not include all the covariates which you will eventually include in your model. It only includes, along with the outcome(s), covariates which are mentioned in your question.

### Exercise 3

We saw the overall distribution of `height` in the Introduction. Now look at how `height` varies with `sex`. Write R code that plots `height` by `sex` as overlapping density curves, with a descriptive title, subtitle, axis labels, and caption. Paste the code below and run it.

```{r wisdom-3}
question_text(NULL,
  message = "    recruits |>\n      ggplot(aes(x = height, fill = sex)) +\n      geom_density(alpha = 0.5) +\n      labs(\n        title = \"Height by Sex in Our 50-Recruit Sample\",\n        subtitle = \"Male and female distributions are offset by about 9 cm\",\n        x = \"Height (cm)\", y = NULL, fill = \"Sex\",\n        caption = \"Source: 50-row sample from NHANES via primer.tutorials::recruits\"\n      ) +\n      theme_minimal() +\n      theme(axis.text.y  = element_blank(),\n            axis.ticks.y = element_blank())",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
recruits |>
  ggplot(aes(x = height, fill = sex)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Height by Sex in Our 50-Recruit Sample",
    subtitle = "Male and female distributions are offset by about 9 cm",
    x = "Height (cm)",
    y = NULL,
    fill = "Sex",
    caption = "Source: 50-row sample from NHANES via primer.tutorials::recruits"
  ) +
  theme_minimal() +
  theme(axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())
```

### Exercise 4

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-4}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

This problem is predictive, so there is only one outcome column: `height`. We compare average heights between male and female young adults.

### Exercise 5

What are the units for this problem?

```{r wisdom-5}
question_text(NULL,
  message = "Each row represents one USMC recruit expected to enlist during the coming year. There are roughly 5,000 such recruits per year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Specifying the Preceptor Table forces us to think clearly about the units and outcomes implied by the question. No data science project follows a single direction. We always backtrack. There is always dialogue. We model units, but we only really care about aggregates.

### Exercise 6

What is the outcome variable for this problem?

```{r wisdom-6}
question_text(NULL,
  message = "Keep track of two 'outcome' variables: the one in our Preceptor Table and the one in our data. In this case, they share the name `height` --- both refer to a young adult's standing height in centimeters --- though the *measurement procedure* behind each column differs.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The outcome variable that we really care about is often not the outcome variable which our data includes. This compromise --- working with what we have rather than what we really want --- is a part of most data science work in the real world.

### Exercise 7

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r wisdom-7}
question_text(NULL,
  message = "Age at enlistment is a candidate. Height is nearly-but-not-quite stable by the age-18 window we care about, so a ten-year-older recruit is almost the same height as a fresh one. Another candidate is race or ethnicity, since height distributions vary across ancestry groups.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The term "covariates" is used in at least three ways in data science. First, it is all the variables which *might* be useful, regardless of whether or not we have the data. Second, it is all the variables for which we have data. Third, it is the set of variables in the data which we end up using in the model.

### Exercise 8

What are the treatments, if any, for this problem?

```{r wisdom-8}
question_text(NULL,
  message = "There are no treatments in this problem because we are building a predictive model. Nothing in `recruits` can be manipulated --- we cannot assign a recruit a different sex or a different height. Predictive models have covariates, not treatments.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Remember that a treatment is just another covariate which, for the purposes of this specific problem, we are assuming can be manipulated --- thereby creating two or more different potential outcomes for each unit. Predictive models have no such variable.

### Exercise 9

What moment in time does the Preceptor Table refer to?

```{r wisdom-9}
question_text(NULL,
  message = "The coming year --- specifically, the moment at which each recruit arrives at bootcamp and has their height recorded at enlistment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

A Preceptor Table can never really refer to an exact instant in time since nothing is instantaneous in this fallen world. In practice, the Preceptor Table's "moment" is a narrow window: recruit intake for a single calendar year.

### Exercise 10

Describe in words the Preceptor Table for this problem.

```{r wisdom-10}
question_text(NULL,
  message = "The Preceptor Table has three columns: one identifying each USMC recruit, one for `height`, and one for `sex`. Each row represents one recruit expected to enlist during the coming year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

```{r}
#| echo: false
#| results: asis
pre_tibble <- tibble::tribble(
  ~`Recruit`,       ~`Height (cm)`, ~`Sex`,
  "Recruit 1",      "150",          "Female",
  "Recruit 2",      "172",          "Male",
  "...",            "...",          "...",
  "Recruit 5,000",  "160",          "Female"
)

pre_title_footnote     <- "A Preceptor Table is the smallest table such that, if every cell were filled with its true value, answering the question 'What is the average height of male and female USMC recruits?' would be easy."
pre_units_footnote     <- "Each row is one USMC recruit expected to enlist during the coming year. Missing rows represent the other recruits not shown."
pre_outcome_footnote   <- "Height in centimeters, measured at enlistment."
pre_covariate_footnote <- "Sex, taking values 'Male' or 'Female', as recorded at enlistment."

pre_gt_html <- gt::gt(pre_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`Recruit`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 11

Because `recruits` is already prepared (filtered to ages 18--27, restricted to `height`, `sex`, and `age`, and missing rows dropped), there is no data-prep code to write. Take a moment to verify the breakdown that matters most for what comes next. In `analysis.qmd`, add a chunk that runs `recruits |> count(sex)`. Render. Copy the printed tibble from the rendered HTML below.

```{r wisdom-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r wisdom-11-test}
#| echo: true
recruits |> count(sex)
```

###

The split between sexes is deliberate: 40 male rows and 10 female. We did not choose this to reflect the male/female proportions among real USMC recruits, but to make a particular feature of the model visible later --- when we estimate the average height for each group, the larger group will give a tighter estimate (smaller confidence interval) than the smaller group. You can never look at the data too much. --- Mark Engerman

### Exercise 12

We will be creating a summary paragraph over the course of this tutorial. Write the first two sentences. The first sentence is a general statement about the overall topic, mentioning the general class of outcome and at least one covariate. The second sentence introduces the data source and the specific question --- when/where gathered, how many observations, who collected it.

```{r wisdom-12}
question_text(NULL,
  message = "People vary in height, and that variation is strongly patterned by sex. Using data from the National Health and Nutrition Examination Survey (NHANES) conducted by the Centers for Disease Control and Prevention, we seek to estimate the average height of male and female USMC recruits for the coming year.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Read our answer. It will not be the same as yours. You can change your answer to incorporate some of our ideas, but do not copy/paste our answer exactly. Add your two sentences to the summary paragraph portion of your QMD. Render the QMD, then commit/push.


## Justice
###

Wisdom gives us the Preceptor Table and the data.

```{r}
#| echo: false
#| results: asis
pre_tibble_j <- tibble::tribble(
  ~`Recruit`,       ~`Height (cm)`, ~`Sex`,
  "Recruit 1",      "150",          "Female",
  "Recruit 2",      "172",          "Male",
  "...",            "...",          "...",
  "Recruit 5,000",  "160",          "Female"
)

pre_title_footnote_j     <- "A Preceptor Table is the smallest table such that, if every cell were filled with its true value, answering the question 'What is the average height of male and female USMC recruits?' would be easy."
pre_units_footnote_j     <- "Each row is one USMC recruit expected to enlist during the coming year. Missing rows represent the other recruits not shown."
pre_outcome_footnote_j   <- "Height in centimeters, measured at enlistment."
pre_covariate_footnote_j <- "Sex, taking values 'Male' or 'Female', as recorded at enlistment."

pre_gt_html_j <- gt::gt(pre_tibble_j, id = "preceptor_tbl_justice") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`Recruit`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Recruit`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote_j,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl_justice .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html_j,
  "</div>\n",
  "```\n",
  sep = ""
)
```


```{r}
#| echo: false
#| results: asis
data_tibble <- tibble::tribble(
  ~`Respondent`, ~`Age`, ~`Height (cm)`, ~`Sex`,
  "NHANES 1",    "21",   "162",          "Female",
  "NHANES 2",    "25",   "178",          "Male",
  "NHANES 3",    "19",   "159",          "Female",
  "...",         "...",  "...",          "...",
  "NHANES 50",   "23",   "170",          "Male"
)

data_title_footnote     <- "National Health and Nutrition Examination Survey (NHANES), Centers for Disease Control and Prevention. Continuous survey data, 1999--present. We restrict to respondents aged 18 through 27 with non-missing height and sex --- about 1,300 rows. The original survey used a complex, stratified, multistage sampling design; we ignore the survey weights."
data_units_footnote     <- "Each row is one NHANES participant examined between 1999 and the most recent release cycle."
data_outcome_footnote   <- "Height in centimeters, measured by a trained examiner in NHANES's Mobile Examination Center using a fixed stadiometer. Each participant is measured once per survey visit with shoes removed."
data_covariate_footnote <- "Sex, recorded by the NHANES interviewer as 'Male' or 'Female'. NHANES does not separately record gender identity; the two-category coding dates from survey waves going back decades."

data_gt_html <- gt::gt(data_tibble, id = "data_tbl") |>
  gt::tab_header(title = "Data: NHANES Young Adults, Ages 18–27") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`Respondent`, `Age`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Respondent`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Age`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = data_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = data_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = data_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = data_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #data_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  data_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 1

In your own words, describe the components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
  message = "Justice reviews the Population Table and selects the formula for the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Justice is about concerns that you (or your critics) might have, reasons why the model you create might not work as well as you hope. The exercises that follow walk through several such concerns by name. Justice also picks the formula for the data generating mechanism --- different outcome variables call for different formulas, a piece that gets more technical names in later tutorials.

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Validity is always about the columns in the Preceptor Table and the data. Just because columns from these two different tables have the same name does not mean that they are the same thing.

### Exercise 3

Provide one reason why the assumption of validity might not hold for the outcome variable `height` or for one of the covariates. Use the words "column" or "columns" in your answer.

```{r justice-3}
question_text(NULL,
  message = "The `height` column in the data is recorded by NHANES's trained examiners using a fixed stadiometer, with shoes removed. The `height` column in the Preceptor Table is recorded at USMC enlistment with whatever instrument the recruiting office has on hand --- possibly with shoes on, possibly under different instructions. Even though both columns are labeled 'Height (cm),' the two measurement procedures are not the same, so the columns may not represent the same underlying quantity.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

In order to consider the Preceptor Table and the data to be drawn from the same population, the columns from one must have a *valid correspondence* with the columns in the other. Validity, if true (or at least reasonable), allows us to construct the Population Table, which is the first step in Justice.

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The Population Table is almost always much bigger than the combination of the Preceptor Table and the data, because if we can really assume that both are part of the same population, then that population must cover a broad universe of time and units.

### Exercise 5

Specify the unit/time combinations which define each row in this Population Table.

```{r justice-5}
question_text(NULL,
  message = "Each row is one young adult (aged 18--27) in a given year. Data rows are NHANES participants examined between 1999 and the most recent release. Preceptor rows are the ~5,000 USMC recruits expected to enlist in the coming year. Other rows cover all the young adults alive during the overlapping span who appear in neither table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

```{r}
#| echo: false
#| results: asis
pop_tibble <- tibble::tribble(
  ~Source    , ~`Young Adult`    , ~`Year`, ~`Height (cm)`, ~`Sex`   ,
  "..."      , "..."             , "..."  , "..."         , "..."    ,
  "Data"     , "NHANES 1"        , "2002" , "162"         , "Female" ,
  "Data"     , "NHANES 2"        , "2005" , "178"         , "Male"   ,
  "Data"     , "..."             , "..."  , "..."         , "..."    ,
  "Data"     , "NHANES 50"       , "2015" , "170"         , "Male"   ,
  "..."      , "..."             , "..."  , "..."         , "..."    ,
  "Preceptor", "Recruit 1"       , "2026" , "150"         , "Female" ,
  "Preceptor", "Recruit 2"       , "2026" , "172"         , "Male"   ,
  "Preceptor", "..."             , "..."  , "..."         , "..."    ,
  "Preceptor", "Recruit 5,000"   , "2026" , "160"         , "Female" ,
  "..."      , "..."             , "..."  , "..."         , "..."
)

pop_title_footnote     <- "This table combines NHANES adult-height data from 1999 onward with the Preceptor Table's 2026 USMC recruits, drawn from the same broader population of US young adults aged 18--27."
pop_units_footnote     <- "Each row is one young adult at one point in time. Data rows are NHANES survey participants between the ages of 18 and 27. Preceptor rows are the ~5,000 USMC recruits expected to enlist in the coming year."
pop_outcome_footnote   <- "Height in centimeters. NHANES measures height directly in the Mobile Examination Center; USMC records height at enlistment. The two measurement procedures are close enough that we treat them as the same column, subject to the validity concern raised in Exercise 3."
pop_covariate_footnote <- "Sex, recorded as Male or Female in both data sources."

pop_gt_html <- gt::gt(pop_tibble, id = "population_tbl") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`Young Adult`, `Year`)) |>
  gt::tab_spanner(label = "Outcome"  , id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Source`, `Young Adult`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 6

In your own words, define the assumption of "stability" when employed in the context of data science.

```{r justice-6}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The most natural way for stability to fail is *time*: our data was gathered in earlier NHANES cycles, our recruits enlist next year. If the world that produced the data behaves the same as the world the recruits enter, stability holds. The 25-year span of NHANES against next year's recruit cohort is a long enough window for the relationship between sex and adult height to drift. We will sharpen this idea --- the *parameters* of the model, not the levels of any single variable --- in the next exercise.

### Exercise 7

Provide one reason why the assumption of stability might not be true in this case.

```{r justice-7}
question_text(NULL,
  message = "The *intercept* of the height-on-sex regression --- the expected height of a female young adult --- has likely drifted upward over the 25-year span of NHANES, while the *slope* (the male-female height gap, the `sexMale` coefficient) has stayed roughly constant. By 2026 the underlying $\\beta_0$ for our recruits is probably a centimeter or two higher than the average across NHANES's reporting years. That is a *parameter* shift; if it is large enough, stability is violated. The fact that average heights in the population have changed is not by itself the issue --- a distribution change in `height` is something a stable model can absorb. What stability requires is that $\\beta_0$ and $\\beta_1$ stay put.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A common confusion is to point at any change between two eras and call it a stability violation. It isn't. The distribution of `height` in the population may have shifted upward; the spread of heights may have narrowed; the mix of ethnicities in the recruiting pool may differ from NHANES participants. None of those, on their own, is a stability violation. What stability requires is that the *parameters* of our model --- $\beta_0$ and $\beta_1$ --- be the same in the data, in the Preceptor Table, and in the broader population. Distribution shifts are everywhere; parameter shifts are what hurt us.

### Exercise 8

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: Data → Population → Preceptor Table. In your own words, define the assumption of "representativeness."

```{r justice-8}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Ideally, we would like both the Preceptor Table and our data to be random samples from the population. Sadly, this is almost never the case.

### Exercise 9

We do not use the data directly to estimate missing values in the Preceptor Table. Instead, we use the data to learn about the overall population. Provide one reason, involving the relationship between the data and the population, why the assumption of representativeness might not be true in this case.

```{r justice-9}
question_text(NULL,
  message = "NHANES is a voluntary survey. People who agree to participate (and who show up for the Mobile Examination Center visit) are not a random sample of US young adults --- they skew toward people with time, transportation, and some trust in government research. On top of that, the `recruits` tibble we are fitting on is itself a non-random subsample of NHANES: 40 male rows and 10 female rows, deliberately unbalanced. Both layers --- the survey participation and our 40/10 choice --- mean the data we model is not a random sample of US young adults.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

When representativeness is violated, our parameter estimates *might* be biased --- not "will be," because by pure luck a non-representative sample can still produce the right number, but we have no principled reason to expect that to happen. There is a *second* representativeness relationship --- between the population and the Preceptor Table --- that later tutorials will exercise; for now, the data ↔ population link is what we focus on.

### Exercise 10

A statistical model consists of two parts: the probability family and the link function. The probability family is the probability distribution that generates the randomness in our data. Our outcome variable is `height` --- continuous. What probability family should we use?

```{r justice-10}
question_text(NULL,
  message = "Normal: $Y \\sim N(\\mu, \\sigma^2)$. The Normal distribution is the standard choice for continuous outcomes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

While you're here, also add `library(tidymodels)` to the setup chunk in `analysis.qmd` so the rest of Courage can call `linear_reg()` and `fit()`.

### Exercise 11

The link function is the mathematical formula that links our data to the unknown parameters. What is the link function for a continuous outcome?

```{r justice-11}
question_text(NULL,
  message = "Linear: $\\mu = \\beta_0 + \\beta_1 X_1 + \\ldots + \\beta_n X_n$. Our model will be a linear regression.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

While you're here, also add `library(broom)` to the setup chunk in `analysis.qmd` so we can call `tidy()` on the fitted model in Courage.

### Exercise 12

Write one sentence highlighting a potential weakness in your model. Derive it from possible problems with the assumptions above. We will add this to our summary paragraph. So far our version of the summary paragraph looks like this:

> People vary in height, and that variation is strongly patterned by sex. Using data from the National Health and Nutrition Examination Survey (NHANES) conducted by the Centers for Disease Control and Prevention, we seek to estimate the average height of male and female USMC recruits for the coming year.

Your version will be somewhat different.

```{r justice-12}
question_text(NULL,
  message = "One concern is that USMC recruits are more physically fit than the general young-adult population and so may be systematically taller, meaning that the NHANES-based estimates probably understate recruit heights.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a weakness sentence to the summary paragraph in your QMD. You can modify your paragraph, but don't copy/paste our answer exactly. Render the QMD, then commit/push.


## Courage
###

Justice gives us the Population Table and the abstract data generating mechanism.

```{r}
#| echo: false
#| results: asis
pop_tibble_c <- tibble::tribble(
  ~Source    , ~`Young Adult`    , ~`Year`, ~`Height (cm)`, ~`Sex`   ,
  "..."      , "..."             , "..."  , "..."         , "..."    ,
  "Data"     , "NHANES 1"        , "2002" , "162"         , "Female" ,
  "Data"     , "NHANES 2"        , "2005" , "178"         , "Male"   ,
  "Data"     , "..."             , "..."  , "..."         , "..."    ,
  "Data"     , "NHANES 50"       , "2015" , "170"         , "Male"   ,
  "..."      , "..."             , "..."  , "..."         , "..."    ,
  "Preceptor", "Recruit 1"       , "2026" , "150"         , "Female" ,
  "Preceptor", "Recruit 2"       , "2026" , "172"         , "Male"   ,
  "Preceptor", "..."             , "..."  , "..."         , "..."    ,
  "Preceptor", "Recruit 5,000"   , "2026" , "160"         , "Female" ,
  "..."      , "..."             , "..."  , "..."         , "..."
)

pop_title_footnote_c     <- "This table combines NHANES adult-height data from 1999 onward with the Preceptor Table's 2026 USMC recruits, drawn from the same broader population of US young adults aged 18--27."
pop_units_footnote_c     <- "Each row is one young adult at one point in time. Data rows are NHANES survey participants between the ages of 18 and 27. Preceptor rows are the ~5,000 USMC recruits expected to enlist in the coming year."
pop_outcome_footnote_c   <- "Height in centimeters. NHANES measures height directly in the Mobile Examination Center; USMC records height at enlistment."
pop_covariate_footnote_c <- "Sex, recorded as Male or Female in both data sources."

pop_gt_html_c <- gt::gt(pop_tibble_c, id = "population_tbl_courage") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`Young Adult`, `Year`)) |>
  gt::tab_spanner(label = "Outcome"  , id = "outcome_span",
                  columns = c(`Height (cm)`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Sex`)) |>
  gt::cols_align(align = "left" , columns = c(`Source`, `Young Adult`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Height (cm)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote_c,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote_c,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote_c,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote_c,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl_courage .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html_c,
  "</div>\n",
  "```\n",
  sep = ""
)
```

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n + \epsilon$$

with $\epsilon \sim N(0, \sigma^2)$.

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
  message = "Courage creates the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The abstract data generating mechanism shown in the preamble uses generic variables --- $Y$, $X_1$, $X_2$, and so on --- because our purpose is to describe the general mathematical structure. At the end of Justice we have decided the functional family (linear with normal errors), but not the number or identity of covariates we will ultimately include. Courage fills in the missing pieces.

### Exercise 2

Because our outcome variable is continuous, add a code chunk to `analysis.qmd` that begins building the model with `linear_reg(engine = "lm")`. Render. Copy the printed model object from the rendered HTML.

```{r courage-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-2-test}
#| echo: true
linear_reg(engine = "lm")
```

###

**[tidymodels](https://www.tidymodels.org/)** is the most popular framework for estimating models among R users. *[Tidy Modeling with R](https://www.tmwr.org/)* by Max Kuhn and Julia Silge is a great introduction.

### Exercise 3

Extend the chunk in `analysis.qmd` to `linear_reg(engine = "lm") |> fit(height ~ sex, data = recruits)`. This is our first candidate model: outcome on the binary `sex`. Render. Copy the printed fit object from the rendered HTML.

```{r courage-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-3-test}
#| echo: true
linear_reg(engine = "lm") |>
  fit(height ~ sex, data = recruits)
```

###

A two-level character variable like `sex` is turned into a 0/1 "dummy" variable, renamed `sexMale` (the level for which the dummy is 1). Math doesn't speak text; dummies do the translation.

### Exercise 4

Here is the parameter table for this candidate model:

```{r}
#| echo: false
linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = recruits) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 1)))
```

Write a sentence interpreting the `sexMale` estimate of 8.9.

```{r courage-4}
question_text(NULL,
  message = "When we compare two groups of recruits differing only in sex, the male group has an expected height about 8.9 centimeters taller than the female group. The intercept of 165.9 is the expected height of a female recruit; adding `sexMale` (= 1 for male) gives the expected male height of about 174.8 cm.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Notice the language: *"when we compare two groups,"* not *"raising sex by one unit increases height."* For a binary covariate the comparison framing is the only honest reading. The same comparison framing applies to continuous covariates too --- which we can see by trying one. The 95% confidence interval for `sexMale` is [4.3, 13.4] centimeters; zero is well outside that interval, so the model says the *expected* height of male recruits is taller than the expected height of female recruits. That does not mean every male is taller than every female --- the two distributions overlap, and there are plenty of female recruits taller than some male recruits.

### Exercise 5

Change the formula to `fit(height ~ age, data = recruits)`. This refits with `age`, a continuous covariate, instead of `sex`. Render. Copy the printed fit object from the rendered HTML.

```{r courage-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-5-test}
#| echo: true
linear_reg(engine = "lm") |>
  fit(height ~ age, data = recruits)
```

###

Continuous covariates produce a single slope coefficient. The reading is per-unit comparison: how much do two groups, differing by one unit of the covariate, differ in expected outcome?

### Exercise 6

Here is the parameter table for the second candidate model:

```{r}
#| echo: false
linear_reg() |>
  set_engine("lm") |>
  fit(height ~ age, data = recruits) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 1)))
```

Write a sentence interpreting the slope on `age`.

```{r courage-6}
question_text(NULL,
  message = "The slope on `age` is essentially zero (0.0 cm per year, with a 95% confidence interval that crosses zero in both directions). When we compare two groups of recruits differing by one year of age (within our 18--27 window), their expected heights are indistinguishable. Adult height has stopped changing by the late teens; this fit confirms that.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The wide confidence interval that includes zero is exactly what *"this covariate doesn't help"* looks like in a parameter table. We tried `age`, and it brought no signal beyond what `sex` already gives us. Our final model is therefore `height ~ sex` --- the candidate from Exercise 4 --- which is the model the rest of the tutorial uses.

### Exercise 7

We've settled on the final fit. In `analysis.qmd`, add a new code chunk that assigns the result of `linear_reg() |> set_engine("lm") |> fit(height ~ sex, data = recruits)` to `fit_recruits`, then on the next line prints `fit_recruits`. Render. Copy the printed model summary from the rendered HTML below.

```{r courage-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-7-test}
#| echo: true
fit_recruits <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = recruits)
fit_recruits
```

###

In data science, **we deal with words, math, and code, but the most important of these is code.** We created the mathematical structure of the model and then wrote a model formula in order to estimate the unknown parameters.

### Exercise 8

Re-fitting the model on every render is wasteful --- the fit's the same each time. `fit_recruits` earns its own cached create chunk. The next three exercises set that up: clean the chunk, turn caching on, add the cache directory to `.gitignore`.

First, clean up. In your chunk from Exercise 7, delete the `fit_recruits` line at the bottom, leaving just the `fit_recruits <- linear_reg() ...` assignment. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

<pre><code>fit_recruits <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = recruits)
</code></pre>

###

A chunk ready to cache should contain only the artifact it builds --- nothing else.

### Exercise 9

Add `#| cache: true` as the first line inside the chunk. Render the QMD. Rendering creates an `analysis_cache/` directory next to your `analysis.qmd`, which holds the saved fit so Quarto can reload it on subsequent renders rather than re-fitting every time. To confirm the directory is there, run `ls` in the **bash Terminal**. CP/CR.

```{r courage-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<pre><code>$ ls
analysis.qmd  analysis_cache  analysis_files  analysis.html
</code></pre>

###

For a 50-row linear regression the savings are small; for the larger model fits in later tutorials, caching can be the difference between a one-second and a one-minute render.

### Exercise 10

The Source Control change count just jumped, because `analysis_cache/` and everything inside it now look like untracked work to git. Cached files should never go to GitHub --- they're machine-specific and regenerated on every render.

Add `analysis_cache` to your `.gitignore` on its own line. In the R Terminal, run `show_file(".gitignore")`. CP/CR.

```{r courage-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<pre><code>analysis_files
analysis_cache
</code></pre>

###

The Source Control change count should drop back to whatever it was before the render. Cached objects live next to the QMD, not in version control.

### Exercise 11

With the parameters estimated, we can write the fitted model as a concrete formula:

$$\hat{\text{height}} = 165.9 + 8.9 \cdot \text{sexMale}$$

Note the differences from the abstract form shown in the preamble. First, we have replaced the parameters $\beta_0$ and $\beta_1$ with our best estimates (rounded to one decimal place, fine enough that the estimates and their confidence-interval bounds remain visibly distinct). Second, we have dropped the error term because this formula generates our estimated outcome, not a description of its randomness. Third, the left-hand side is $\widehat{\text{height}}$ instead of $\text{height}$ because this formula generates our estimated (or predicted) height --- a hat indicates an estimated value.

**This is our data generating mechanism.**

###

A data generating mechanism is just a formula, something which we can write down and implement with computer code. Of course, there is randomness built into the DGM, but we won't worry about that detail for now.

### Exercise 12

Start a new working chunk below the cached `fit_recruits` chunk that runs `tidy(fit_recruits, conf.int = TRUE)`. This returns 95% intervals for all the parameters in our model. Render. Copy the printed tibble from the rendered HTML.

```{r courage-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-12-test}
#| echo: true
tidy(fit_recruits, conf.int = TRUE)
```

###

`tidy()` is part of the **[broom](https://broom.tidymodels.org/)** package, used to summarize information from a wide variety of models.

### Exercise 13

Here is a nicer-looking version of that parameter table, rendered with **[knitr](https://yihui.org/knitr/)**:

```{r}
#| echo: false
library(knitr)

tidy(fit_recruits, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 1))) |>
  kable(
    caption = "Estimates and 95% Confidence Intervals for Parameters.  Source: 50-row sample from NHANES via primer.tutorials::recruits"
  )
```

Rounding to one decimal place keeps the displayed values consistent with the fitted DGM formula from Exercise 7 and is fine enough that the estimate and the confidence-interval bounds are visibly distinguishable in every row. A blanket "three significant figures" would risk collapsing values close to one another --- like an intercept of 165.9 and its upper bound 170.0 --- into ambiguous displays. You need rounding finer than the standard error to keep adjacent values apart. At the very least, a table like this should include a title and a caption with the data source.

###

The more variables we add, the more difficult it is to interpret the meaning of any particular coefficient. But interpretation also becomes less important. We don't really care about coefficients. We care about using our model to estimate quantities of interest.

### Exercise 14

Add a sentence to your project summary explaining the structure of the model. Something like: "We model XX [concept of outcome, not variable name], [values of XX], as a [linear/logistic/multinomial/ordinal] function of XX [and maybe other covariates]."

Recall the beginning of our version of the summary:

> People vary in height, and that variation is strongly patterned by sex. Using data from the National Health and Nutrition Examination Survey (NHANES) conducted by the Centers for Disease Control and Prevention, we seek to estimate the average height of male and female USMC recruits for the coming year. One concern is that USMC recruits are more physically fit than the general young-adult population and so may be systematically taller, meaning that the NHANES-based estimates probably understate recruit heights.

```{r courage-14}
question_text(NULL,
  message = "We model height as a normally distributed variable which is a linear function of sex.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Read our answer. Do not copy/paste exactly. Add your two sentences to the summary paragraph. Render the QMD, then commit/push.


## Temperance
###

Courage provides the data generating mechanism. We can express the DGM in four ways:

**In words:** We model height as a normally distributed variable which is a linear function of sex.

**In R code:**

```r
fit_recruits <- linear_reg() |>
  set_engine("lm") |>
  fit(height ~ sex, data = recruits)
```

**In a parameter table:**

```{r}
#| echo: false
fit_recruits |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 1)))
```

**As a mathematical formula:**

$$\text{height} = 165.9 + 8.9 \cdot \text{sexMale} + \epsilon$$

with $\epsilon \sim N(0, 41.0)$.

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
  message = "Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Courage gave us the data generating mechanism, including the parameter values that Temperance now puts to work. The tool for that work is the **[marginaleffects](https://marginaleffects.com/)** package, with companion book *[Model to Meaning](https://marginaleffects.com/)* by Vincent Arel-Bundock.

### Exercise 2

In the end, we don't really care about parameters, much less how to interpret them. Parameters are *imaginary*, like unicorns. We care about answers to our questions. *In the modern world, all parameters are nuisance parameters.*

Add `library(marginaleffects)` to the setup chunk in `analysis.qmd` and render. CP/CR.

```{r temperance-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

<pre><code>
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(marginaleffects)
</code></pre>

###

We should be modest in the claims we make. The posteriors we create are never the "truth." The assumptions we made to create the model are never perfect. Yet decisions made with flawed posteriors are almost always better than decisions made without them.

### Exercise 3

What is the specific question we are trying to answer?

```{r temperance-3}
question_text(NULL,
  message = "What is the expected height of male and female USMC recruits next year?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Data science projects begin with a decision which we face. To make that decision wisely, we would like to have good estimates of many unknown numbers. Yet, in order to make progress, we need to drill down to one specific question. This leads to the creation of a data generating mechanism, which can then be used to answer lots of questions.

### Exercise 4

In `analysis.qmd`, add a chunk that runs `predictions(fit_recruits)`. Render. Copy the first few rows of the printed output from the rendered HTML below.

```{r temperance-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r temperance-4-test}
#| echo: true
predictions(fit_recruits)
```

###

`predictions()` returns a data frame with one row for each observation in the data set used to fit the model. The 50 rows correspond to our 50-row `recruits` sample. Each row's `estimate` is that row's model-predicted height. Note already that the standard error column has two distinct values --- about 1.01 for male rows and about 2.02 for female rows. We will come back to this.

### Exercise 5

In `analysis.qmd`, add a chunk that runs `plot_predictions(fit_recruits, condition = "sex")`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r temperance-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r temperance-5-test}
#| echo: true
plot_predictions(fit_recruits, condition = "sex")
```

###

You should see two point estimates (one for Female, one for Male) with 95% confidence intervals. The gap between them is the model's estimate of the average height difference by sex. The Female interval is visibly wider than the Male interval --- the predicted height for a female recruit is less certain than the predicted height for a male recruit. That is the consequence of having drawn 10 female rows but 40 male rows: the standard error of a group mean shrinks roughly with the square root of the group's sample size, so quadrupling the sample size halves the standard error, and the female interval ends up about twice as wide as the male interval.

### Exercise 6

Modify the previous chunk to add `draw = FALSE` to the `plot_predictions()` call. Render. Copy the printed tibble from the rendered HTML below.

```{r temperance-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r temperance-6-test}
#| echo: true
plot_predictions(fit_recruits, condition = "sex", draw = FALSE)
```

###

Because `plot_predictions()` returns a ggplot object, you can continue with ggplot commands like `labs()`. But it can be useful to see the underlying values in the tibble and build your own plot directly.

### Exercise 7

Create a beautiful plot starting from the output of `plot_predictions(fit_recruits, condition = "sex", draw = FALSE)`. Do this in your QMD (much easier than typing in the R Terminal directly). Title: the key variables. Subtitle: the key takeaway. Caption: data source. Axis labels: human-readable with units.

Paste the plot code below.

```{r temperance-7}
question_text(NULL,
  message = "    plot_predictions(fit_recruits, condition = \"sex\", draw = FALSE) |>\n      ggplot(aes(x = sex, y = estimate)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high),\n                      color = \"grey20\", size = 1.1) +\n      labs(\n        title = \"Forecast Height for USMC Recruits\",\n        subtitle = \"Male recruits are expected to be about 9 cm taller\",\n        x = \"Sex\",\n        y = \"Expected Height (cm)\",\n        caption = \"Source: 50-row sample from NHANES via primer.tutorials::recruits\"\n      ) +\n      theme_minimal()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 12)
```

###

Here is our version:

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(fit_recruits, condition = "sex", draw = FALSE) |>
  ggplot(aes(x = sex, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                  color = "grey20", size = 1.1) +
  labs(
    title = "Forecast Height for USMC Recruits",
    subtitle = "Male recruits are expected to be about 9 cm taller than female recruits",
    x = "Sex",
    y = "Expected Height (cm)",
    caption = "Source: 50-row sample from NHANES via primer.tutorials::recruits"
  ) +
  theme_minimal()
```

Data science often involves this back-and-forth style of work. First, make a single chunk of code --- say, a new plot --- work well. This requires interactive work between the QMD and the R Terminal. Second, ensure that the entire QMD runs correctly on its own.

### Exercise 8

Finalize the new graphics chunk in your QMD. Render to ensure it all works. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r temperance-8}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    plot_predictions(fit_recruits, condition = \"sex\", draw = FALSE) |>\n      ggplot(aes(x = sex, y = estimate)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +\n      labs(\n        title = \"Forecast Height for USMC Recruits\",\n        subtitle = \"Male recruits are expected to be about 9 cm taller\",\n        x = \"Sex\", y = \"Expected Height (cm)\",\n        caption = \"Source: 50-row sample from NHANES via primer.tutorials::recruits\"\n      ) +\n      theme_minimal()\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

Always remember: the map is not the territory. A beautiful graphic tells a story, but that story is always an imperfect representation of reality. Our models depend on assumptions that are never completely true.

### Exercise 9

Write the last sentence of your summary paragraph. It describes at least one Quantity of Interest and a measure of uncertainty. It is OK if this QoI differs from the one you began with.

```{r temperance-9}
question_text(NULL,
  message = "We estimate that the average male USMC recruit will be about 175 cm tall and the average female recruit about 166 cm, with the female interval visibly wider than the male interval because we have far fewer female observations in our sample.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a final sentence to your summary paragraph, but don't copy/paste our answer exactly. Render the QMD.

### Exercise 10

Write a few sentences explaining why the estimates for the quantities of interest, and the uncertainty, might be wrong. Suggest alternative estimates and a wider confidence interval if warranted.

```{r temperance-10}
question_text(NULL,
  message = "USMC recruits are more physically fit and probably taller than the general young-adult population measured by NHANES. A better point estimate would nudge male recruit heights upward by a centimeter or two and female recruit heights similarly. The confidence intervals should be wider than what the model reports --- our 50-row sample is small (only 10 of those rows are female), and the bigger sources of uncertainty (the NHANES-to-USMC validity gap and the voluntary-participation representativeness gap) aren't captured by the model's standard error.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Always go back to your Preceptor Table --- the information which, if you had it, would make answering your question easy. In almost all real-world cases, the Preceptor Table and the data are fairly different. So, even a perfectly estimated statistical model is rarely as useful as we might like.

### Exercise 11

Rearrange the material in your QMD so the order is graphic, then summary paragraph. The chunk that creates `fit_recruits` must occur before the chunk that creates the graphic. You can keep or discard the math at your discretion. Render the QMD. In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r temperance-11}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\")\n    ---\n    title: \"Recruits\"\n    author: \"<your name>\"\n    execute:\n      echo: false\n    ---\n    \n    ```{r}\n    #| message: false\n    library(tidyverse)\n    library(tidymodels)\n    library(primer.tutorials)\n    library(broom)\n    library(marginaleffects)\n    ```\n    \n    ```{r}\n    #| cache: true\n    fit_recruits <- linear_reg() |>\n      set_engine(\"lm\") |>\n      fit(height ~ sex, data = recruits)\n    ```\n    \n    ```{r}\n    plot_predictions(fit_recruits, condition = \"sex\", draw = FALSE) |>\n      ggplot(aes(x = sex, y = estimate)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +\n      labs(\n        title = \"Forecast Height for USMC Recruits\",\n        subtitle = \"Male recruits are expected to be about 9 cm taller\",\n        x = \"Sex\", y = \"Expected Height (cm)\",\n        caption = \"Source: 50-row sample from NHANES via primer.tutorials::recruits\"\n      ) +\n      theme_minimal()\n    ```\n    \n    People vary in height, and that variation is strongly patterned by sex. Using a 50-row teaching sample drawn from the National Health and Nutrition Examination Survey (NHANES, conducted by the Centers for Disease Control and Prevention), we seek to estimate the average height of male and female USMC recruits for the coming year. One concern is that USMC recruits are more physically fit than the general young-adult population and so may be systematically taller, meaning that the NHANES-based estimates probably understate recruit heights. We model height as a normally distributed variable which is a linear function of sex. We estimate that the average male recruit will be about 175 cm tall and the average female recruit about 166 cm, with confidence intervals that are visibly wider for the female estimate because there are only ten female rows in our sample.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

This is the version of your QMD file at which your teacher is most likely to look closely.

### Exercise 12

Publish your rendered QMD to GitHub Pages. In the bash Terminal (not the R Terminal!), run:

```
quarto publish gh-pages analysis.qmd
```

Copy/paste the resulting URL below.

```{r temperance-12}
question_text(NULL,
  message = "https://<your-github-username>.github.io/recruits/analysis.html",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

Commit/push everything.

### Exercise 13

Copy/paste the URL to your GitHub repo.

```{r temperance-13}
question_text(NULL,
  message = "https://github.com/<your-github-username>/recruits",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God. If all our assumptions are correct, then our DGM is true --- it accurately describes the way in which the world works. There is no better way to predict the future, or to model the past, than to use it. Sadly, this will only be the case with toy examples involving things like coins and dice.

The world confronts us. Make decisions we must.


## Summary
###

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(fit_recruits, condition = "sex", draw = FALSE) |>
  ggplot(aes(x = sex, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                  color = "grey20", size = 1.1) +
  labs(
    title = "Forecast Height for USMC Recruits",
    subtitle = "Male recruits are expected to be about 9 cm taller than female recruits",
    x = "Sex",
    y = "Expected Height (cm)",
    caption = "Source: 50-row sample from NHANES via primer.tutorials::recruits"
  ) +
  theme_minimal()
```

People vary in height, and that variation is strongly patterned by sex. Using a 50-row teaching sample drawn from the National Health and Nutrition Examination Survey (NHANES, conducted by the Centers for Disease Control and Prevention), we seek to estimate the average height of male and female USMC recruits for the coming year. One concern is that USMC recruits are more physically fit than the general young-adult population and so may be systematically taller, meaning that the NHANES-based estimates probably understate recruit heights. We model height as a normally distributed variable which is a linear function of sex. We estimate that the average male recruit will be about 175 cm tall and the average female recruit about 166 cm, with the female interval visibly wider than the male interval because we have only ten female rows in our sample.

This forecast would help you estimate how many small-versus-large uniforms to order. It does not address every question the logistics office will face --- it says nothing about weight, shoulder width, or inseam. But it gives a defensible starting point.

The world is always more uncertain than our models would have us believe.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
