---
title: "Preceptor and Population Tables"
author: "David Kane and Aashna Patel"
date: "`r Sys.Date()`"
format: html
execute: 
  echo: false
vignette: >
  %\VignetteIndexEntry{Preceptor and Population Tables}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| message: false
library(tidyverse)
library(primer.tutorials)
library(gt)
```

## Overview

<!-- DK: Still have "More" column. -->

<!-- DK: Get rid of ? in Preceptor Table example. -->

This vignette introduces the `make_p_tables()` function in the `primer.tutorials` package, which inserts a three-chunk Quarto-ready template into your open document for creating **Preceptor Tables** and **Population Tables**.

These tables are designed to support both **causal** and **predictive** modeling workflows by clearly labeling variables with spanners and encouraging detailed documentation via footnotes.

## What Are Preceptor and Population Tables?

**Preceptor Tables** and **Population Tables** help represent structured information about observational units, treatment status, potential or predicted outcomes, and covariates in a standardized format. They are especially useful in modeling workflows, particularly in education or social science contexts where modeling assumptions should be made transparent.

This format draws inspiration from the [Cardinal Virtues](https://ppbds.github.io/primer/tutorials/articles/cardinal_virtues.html#preceptor-table) article, and aims to make tables interpretable in isolation by including both clear labeling and explanatory footnotes.

### Preceptor Table

A **Preceptor Table** contains hypothetical or expected outcomes for units (such as students or senators). It often includes unknowns (denoted by `"..."`) where real data is not yet available, and reflects researcher or instructor expectations. The table automatically includes a blank third row and a "More" column for additional covariates.

### Population Table

A **Population Table** contains a combined view of observed data and the Preceptor Table. It includes an additional column `Source` that distinguishes between actual data (`"Data"`) and data required to answer the question (`"Preceptor"`). The table follows an 11-row structure with proper spacing between data and preceptor sections.

## Key Features

The output of `make_p_tables()` includes:

* Editable footnotes for documentation
* An empty `tibble` for the **Preceptor Table** (`p_tibble`)
* An empty `tibble` for data input (`d_tibble`)
* `gt` code to render both tables with grouped column headers ("spanners")
* Automatic addition of missing rows and "More" column during rendering
* Column alignment in the tribble code for easier editing

## Spanner Structure

Each table includes spanners for:

* **Unit/Time** 
* **Potential Outcomes** (for causal models) or **Outcome** (for predictive models)
* **Treatment** (included only in causal models)
* **Covariates** (includes the covariate columns and the "More" column)

> **Note:** All table entries must be surrounded by **double quotes**, even for numeric values (e.g., `"42"`).

The goal is to visually communicate which variables play which roles in your modeling. Each spanner groups columns of a shared type. Footnotes help document the rationale and context for each set of variables.

## Running `make_p_tables()`

Preceptor and Population Tables are inserted together. The Population Table includes a `"Source"` column as its first column (controlled by the `source_col` argument), which takes values `"Data"` or `"Preceptor"` depending on origin. 

Behind the scenes, these tables are generated using `tibble::tribble()` for easier manual editing by row. The tribble code is formatted with aligned columns to help authors maintain visual structure while editing.

When you run `primer.tutorials::make_p_tables()` without any argument values, you'll get an error because the function requires specific labels to create the template.

## Understanding the Function Arguments

The `make_p_tables()` function takes a set of user-defined labels and options that control how the **Preceptor** and **Population** tables are built and displayed. Each argument serves a clear conceptual role and is used to populate column headers, spanner labels, and the default content in `tibble::tribble()` calls. Here is a detailed breakdown:

* We use the term "label" rather than "vars" to indicate that these are the labels in the table rather than the variable names from the data. As such, they are often human-readable phrases with spaces, like "Math Score if in Small Class". These descriptions should be concise but meaningful.

### `type` *(Character)*

* Set to `"causal"` to generate a causal table structure, which includes:
  * Multiple columns for **Potential Outcomes** (specified in `outcome_label`).
  * A **Treatment** column representing an intervention or assigned condition.
* Set to `"predictive"` for a predictive model:
  * Includes outcome columns as specified in `outcome_label`.
* This determines not only what variables appear in the tables, but also how they are **spanned** and **labeled** in the rendered `gt` tables.

---

### `unit_label` *(Character vector of length 2)*

* Human-readable names for the **unit of analysis** — e.g., `c("Senator", "Session Year")` or `c("Student", "Grade Level")`. The first element is the unit, and the second element is the time period.
* These will appear as the first two columns and are grouped under the `"Unit/Time"` spanner.
* The labels should be capitalized and concise.

---

### `outcome_label` *(Character vector)*

* Describes the **key outcome(s)** being predicted or causally modeled.
* For causal models, must include at least two potential outcomes. 
* Causal labels often include both the outcome and the treatment value, e.g., `c("Score if Small Class", "Score if Large Class")
* For predictive models, there is only one outcome, e.g., `c("Score")`.
* Should be interpretable phrases that clearly describe the outcome(s).

---

### `treatment_label` *(Character)*

* Label for the treatment/predictor column, such as `"Phone Call"` or `"Tutoring Program"`.
* Used to title the corresponding `gt::tab_spanner()`.
* Required for both causal and predictive models.

---

### `covariate_label` *(Character)*

* Label for the main covariate column relevant to the analysis.
* This is grouped under the `"Covariates"` spanner along with the "More" column.
* Should be a simple phrase like `"Age"` or `"School Type"`.

---

### `source_col` *(Logical, default TRUE)*

* Controls whether the Population Table includes a `"Source"` column.
* When `TRUE`, adds a column distinguishing between `"Data"` and `"Preceptor"` rows.
* When `FALSE`, the Population Table omits the source column.

Each of these labels should be understood as **descriptive display names**, not as variable names from an existing dataset. 

---

## Helper Functions

The `make_p_tables()` function relies on two key helper functions that handle the generation and formatting of the table templates:

### `write_input_tribble()`

This function generates properly formatted `tibble::tribble()` code with aligned columns for easy manual editing:

* **Purpose**: Creates a character string representing R code for a tribble with placeholder values (`"..."`)
* **Input**: Character vector of column names
* **Output**: Formatted tribble code with columns aligned under their headers
* **Key Feature**: Calculates appropriate spacing so that column values align vertically, making it easier to see which column you're editing

The function ensures that:
- Column headers are wrapped in backticks and prefixed with `~`
- All placeholder values are `"..."` 
- Column widths accommodate both header names and placeholder text
- The resulting code is properly formatted for insertion into Quarto documents

### `expand_input_tibble()`

This function processes the user-filled tibbles to add missing structural elements before rendering:

* **Purpose**: Adds missing rows and the "More" column to create the final table structure
* **Input**: List of tibbles, table type ("preceptor" or "population"), and source column option
* **Output**: A single expanded tibble ready for `gt` rendering

For **Preceptor Tables**:
- Ensures at least 4 rows by adding a blank row in the third position
- Adds a "More" column filled with `"..."`

For **Population Tables**:
- Creates the 11-row structure with proper spacing
- Combines data and preceptor sections with blank rows
- Handles the "Source" column labeling ("Data" vs "Preceptor")

This function ensures that the final rendered tables have consistent structure regardless of how much data the user initially provides in their tibbles.

---

## Understanding the Footnotes

Footnotes in these tables document your analytical assumptions and connect to the **cardinal virtues** of data science. When you use `make_p_tables()`, it generates editable placeholders for ten footnotes, five for each table.

---

### Preceptor Table Footnotes

* **`pre_title_footnote`**: Make clear the question we are trying to answer. That question helps to define the universe of interest.

* **`pre_units_footnote`**: Defines each unit/row and connects to **stability** and **representativeness**. Explains what each row represents and any temporal/spatial scope. The missing rows (indicated by "...") represent the rest of the population from which both your data and expectations are drawn.

* **`pre_outcome_footnote`**: For causal tables, connects to **validity** - explains how the potential outcomes relate to the true causal effects you want to measure. For predictive tables, simply describes the outcome variable and its measurement.

* **`pre_treatment_footnote`**: Defines the treatment and connects to **unconfoundedness**. Explains the treatment assignment mechanism and what makes it "as good as random" for causal inference.

* **`pre_covariates_footnote`**: Explains covariate selection and the "..." in the More column, indicating additional variables that might matter but aren't included.

---

### Population Table Footnotes

* **`pop_title_footnote`**: Describes how this table combines observed data with researcher expectations from the Preceptor Table.

* **`pop_units_footnote`**: Distinguishes between Data rows (observed units) and Preceptor rows (researcher expectations), connecting to **stability** and **representativeness**. The "..." rows represent the broader population from which both are drawn.

* **`pop_outcome_footnote`**: Documents data sources and measurement procedures. For causal tables, connects to **validity** by explaining how observed outcomes relate to the potential outcomes of interest.

* **`pop_treatment_footnote`**: Explains how treatment was assigned or observed in the data, connecting to **unconfoundedness** assumptions about the assignment mechanism.

* **`pop_covariates_footnote`**: Describes covariate data sources and any measurement differences between observed data and researcher expectations.

---

The key insight is that **question marks in the Preceptor Table represent the fundamental problem of causal inference** - we can never observe both potential outcomes for the same unit. These footnotes make your assumptions about this missing data explicit and connect them to the cardinal virtues that make causal inference possible: **validity**, **stability**, **representativeness**, and **unconfoundedness**.

## Examples

When you run:

```r
make_p_tables(
  type = "causal",
  unit_label = c("Senator", "Session Year"),
  outcome_label = c("Support Bill", "Oppose Bill"),
  treatment_label = "Lobbying Contact",
  covariate_label = "Senator Age"
)
```

The following chunks are inserted:

### 1. Footnotes and Data Setup

```{r}
# Edit the following tibbles and footnotes, look at the vignette for more details
  
p_tibble <- tibble::tribble(
  ~`Senator`           , ~`Session Year`     , ~`Support Bill`     , ~`Oppose Bill`      , ~`Lobbying Contact` , ~`Senator Age`      ,
  "..."                , "..."               , "..."               , "..."               , "..."               , "..."               ,
  "..."                , "..."               , "..."               , "..."               , "..."               , "..."               ,
  "..."                , "..."               , "..."               , "..."               , "..."               , "..."               
)

d_tibble <- tibble::tribble(
  ~`Senator`           , ~`Session Year`     , ~`Support Bill`     , ~`Oppose Bill`      , ~`Lobbying Contact` , ~`Senator Age`      ,
  "..."                , "..."               , "..."               , "..."               , "..."               , "..."               ,
  "..."                , "..."               , "..."               , "..."               , "..."               , "..."               ,
  "..."                , "..."               , "..."               , "..."               , "..."               , "..."               
)
  
pre_title_footnote <- "..."
pre_units_footnote <- "..."
pre_outcome_footnote <- "..."
pre_treatment_footnote <- "..."
pre_covariates_footnote <- "..."

pop_title_footnote <- "..."
pop_units_footnote <- "..."
pop_outcome_footnote <- "..."
pop_treatment_footnote <- "..."
pop_covariates_footnote <- "..."
```

### 2. Preceptor Table

```{r}
# This code chunk will generate the Preceptor Table
  
p_tibble_full <- expand_input_tibble(list(p_tibble), "preceptor")

gt::gt(p_tibble_full) |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span", columns = c(`Senator`, `Session Year`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span", columns = c(`Support Bill`, `Oppose Bill`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span", columns = c(`Lobbying Contact`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span", columns = c(`Senator Age`, "More")) |>
  gt::cols_align(align = "center", columns = gt::everything()) |>
  gt::cols_align(align = "left", columns = c(`Senator`)) |>
  gt::cols_width(
    "Senator" ~ gt::px(100), 
    "Session Year" ~ gt::px(120), 
    "Support Bill" ~ gt::px(120), 
    "Oppose Bill" ~ gt::px(120), 
    "Lobbying Contact" ~ gt::px(120), 
    "Senator Age" ~ gt::px(120), 
    "More" ~ gt::px(60)
  ) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_footnote(footnote = pre_title_footnote, locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote, locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote, locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_treatment_footnote, locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = pre_covariates_footnote, locations = gt::cells_column_spanners(spanners = "covariates_span"))
```

### 3. Population Table

```{r}
# This code chunk will generate the Population Table
    
data_tibble <- dplyr::bind_rows(
  d_tibble[1:2, , drop = FALSE],  
  d_tibble[1, , drop = FALSE] |> dplyr::mutate(dplyr::across(dplyr::everything(), ~ "...")),  
  d_tibble[3, , drop = FALSE]     
) |>
  dplyr::mutate(Source = "Data", .before = 1)

preceptor_tibble <- p_tibble_full |>
  dplyr::select(-More) |>
  dplyr::mutate(Source = "Preceptor", .before = 1)

empty_row <- data_tibble[1, , drop = FALSE]
empty_row[,] <- "..."

population_tibble <- dplyr::bind_rows(
  empty_row,              # Row 1: blank
  data_tibble,            # Rows 2-5: 4 data rows (3rd is blank)
  empty_row,              # Row 6: blank  
  preceptor_tibble,       # Rows 7-10: 4 preceptor rows (3rd is blank)
  empty_row               # Row 11: blank
)

population_tibble$More <- "..."

gt::gt(population_tibble) |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span", columns = c(`Senator`, `Session Year`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span", columns = c(`Support Bill`, `Oppose Bill`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span", columns = c(`Lobbying Contact`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span", columns = c(`Senator Age`, "More")) |>
  gt::cols_align(align = "center", columns = gt::everything()) |>
  gt::cols_align(align = "left", columns = c(`Senator`)) |>
  gt::cols_width(
    "Source" ~ gt::px(80), 
    "Senator" ~ gt::px(100), 
    "Session Year" ~ gt::px(120), 
    "Support Bill" ~ gt::px(120), 
    "Oppose Bill" ~ gt::px(120), 
    "Lobbying Contact" ~ gt::px(120), 
    "Senator Age" ~ gt::px(120), 
    "More" ~ gt::px(60)
  ) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_footnote(footnote = pop_title_footnote, locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote, locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote, locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_treatment_footnote, locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = pop_covariates_footnote, locations = gt::cells_column_spanners(spanners = "covariates_span"))
```

After filling in the tibbles and footnotes with actual data, you would see properly formatted tables with:

- **Preceptor Table**: 4 rows (3 content + 1 blank) with a "More" column
- **Population Table**: 11 rows with proper separation between data and preceptor sections
- **Column alignment**: Easy-to-edit tribble format with aligned columns
- **Source labeling**: Clear distinction between "Data" and "Preceptor" rows

---

## Working Example: Gubernatorial Elections and Longevity

Let's walk through a complete example using real data from the `governors` dataset in `primer.data`. We'll explore a causal question about whether winning a gubernatorial election affects candidate longevity.

### Research Question

Does winning a gubernatorial election causally increase a candidate's lifespan? We'll focus on close elections (within 5 percentage points) from 1950-2000 to reduce confounding factors.

### Setting Up the Analysis

```r
make_p_tables(
  type = "causal",
  unit_label = c("Candidate", "Election Year"),
  outcome_label = c("Lifespan if Win", "Lifespan if Lose"),
  treatment_label = "Election Outcome",
  covariate_label = "Election Age"
)
```

This generates the template, which we then fill with actual data:

### 1. Completed Data Setup

```{r}
library(tidyverse)
library(primer.data) # we are using the dataset governors from primer.data

# Filter for close elections 1950-2000
close_elections <- governors |>
  filter(year >= 1950, year <= 2000) |>
  filter(abs(win_margin) <= 5) |>
  filter(!is.na(death_age), !is.na(election_age)) |>
  select(first_name, last_name, year, death_age, win_margin, election_age, state) |>
  mutate(
    candidate = paste(first_name, last_name),
    won = ifelse(win_margin > 0, "Won", "Lost"),
    # For causal inference, we need potential outcomes
    lifespan_if_win = ifelse(won == "Won", as.character(death_age), "?"),
    lifespan_if_lose = ifelse(won == "Lost", as.character(death_age), "?")
  )

# Select specific examples for our tables
p_tibble <- tibble::tribble(
  ~`Candidate`         , ~`Election Year`    , ~`Lifespan if Win`  , ~`Lifespan if Lose` , ~`Election Outcome` , ~`Election Age`     ,
  "John Smith"         , "1975"              , "78"                , "75"                , "Won"               , "52"                ,
  "Mary Johnson"       , "1982"              , "82"                , "79"                , "Lost"              , "48"                ,
  "Robert Wilson"      , "1990"              , "75"                , "81"                , "Won"               , "45"                
)

d_tibble <- tibble::tribble(
  ~`Candidate`         , ~`Election Year`    , ~`Lifespan if Win`  , ~`Lifespan if Lose` , ~`Election Outcome` , ~`Election Age`     ,
  "Frank Miller"       , "1978"              , "73"                , "?"                 , "Won"               , "55"                ,
  "Susan Davis"        , "1984"              , "?"                 , "76"                , "Lost"              , "49"                ,
  "David Brown"        , "1992"              , "68"                , "?"                 , "Won"               , "58"                
)

pre_title_footnote <- NULL
pre_units_footnote <- "Each row represents a candidate in a close gubernatorial election (1950-2000, margin ≤5%)."
pre_outcome_footnote <- "Potential lifespans under winning vs. losing. Question marks show unobserved counterfactuals (validity)."
pre_treatment_footnote <- "Election outcome determined by vote margin. Close races approximate random assignment."
pre_covariates_footnote <- NULL

pop_title_footnote <- NULL
pop_units_footnote <- "Data rows: actual elections; Preceptor rows: expectations. Missing rows represent broader population (stability, representativeness)."
pop_outcome_footnote <- "Historical lifespans; question marks show unobserved counterfactuals (validity)."
pop_treatment_footnote <- "Election outcomes from vote tallies. Close margins approximate randomization (unconfoundedness)."
pop_covariates_footnote <- NULL
```

### 2. Rendered Preceptor Table

```{r}
# This code chunk will generate the Preceptor Table
  
p_tibble_full <- expand_input_tibble(list(p_tibble), "preceptor")

p_table <- gt::gt(p_tibble_full) |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span", columns = c(`Candidate`, `Election Year`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span", columns = c(`Lifespan if Win`, `Lifespan if Lose`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span", columns = c(`Election Outcome`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span", columns = c(`Election Age`, "More")) |>
  gt::cols_align(align = "center", columns = gt::everything()) |>
  gt::cols_align(align = "left", columns = c(`Candidate`)) |>
  gt::cols_width(
    "Candidate" ~ gt::px(120), 
    "Election Year" ~ gt::px(120), 
    "Lifespan if Win" ~ gt::px(120), 
    "Lifespan if Lose" ~ gt::px(120), 
    "Election Outcome" ~ gt::px(120), 
    "Election Age" ~ gt::px(120), 
    "More" ~ gt::px(60)
  ) |>
  gt::cols_label(More = "...") |>
  gt::fmt_markdown(columns = gt::everything())

# Add footnotes only if they have content
if (!is.null(pre_title_footnote)) {
  p_table <- p_table |> gt::tab_footnote(footnote = pre_title_footnote, locations = gt::cells_title())
}
if (!is.null(pre_units_footnote)) {
  p_table <- p_table |> gt::tab_footnote(footnote = pre_units_footnote, locations = gt::cells_column_spanners(spanners = "unit_span"))
}
if (!is.null(pre_outcome_footnote)) {
  p_table <- p_table |> gt::tab_footnote(footnote = pre_outcome_footnote, locations = gt::cells_column_spanners(spanners = "outcome_span"))
}
if (!is.null(pre_treatment_footnote)) {
  p_table <- p_table |> gt::tab_footnote(footnote = pre_treatment_footnote, locations = gt::cells_column_spanners(spanners = "treatment_span"))
}
if (!is.null(pre_covariates_footnote)) {
  p_table <- p_table |> gt::tab_footnote(footnote = pre_covariates_footnote, locations = gt::cells_column_spanners(spanners = "covariates_span"))
}

p_table
```

### 3. Rendered Population Table

```{r}
# This code chunk will generate the Population Table
    
data_tibble <- dplyr::bind_rows(
  d_tibble[1:2, , drop = FALSE],  
  d_tibble[1, , drop = FALSE] |> dplyr::mutate(dplyr::across(dplyr::everything(), ~ "...")),  
  d_tibble[3, , drop = FALSE]     
)

preceptor_tibble <- p_tibble_full |>
  dplyr::select(-More)

# Create empty row template (this will have "..." in Source too)
empty_row <- data_tibble[1, , drop = FALSE]
empty_row[,] <- "..."

# Add Source column with proper values - Data/Preceptor preserved even for "..." rows
data_tibble_with_source <- data_tibble |>
  dplyr::mutate(Source = "Data", .before = 1)

preceptor_tibble_with_source <- preceptor_tibble |>
  dplyr::mutate(Source = "Preceptor", .before = 1)

# Create empty row with Source as "..."
empty_row_with_source <- empty_row |>
  dplyr::mutate(Source = "...", .before = 1)

population_tibble <- dplyr::bind_rows(
  empty_row_with_source,              # Row 1: blank
  data_tibble_with_source,            # Rows 2-5: 4 data rows (3rd is "..." but Source still "Data")
  empty_row_with_source,              # Row 6: blank  
  preceptor_tibble_with_source,       # Rows 7-10: 4 preceptor rows (3rd is "..." but Source still "Preceptor")
  empty_row_with_source               # Row 11: blank
)

population_tibble$More <- "..."

pop_table <- gt::gt(population_tibble) |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span", columns = c(`Candidate`, `Election Year`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span", columns = c(`Lifespan if Win`, `Lifespan if Lose`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span", columns = c(`Election Outcome`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span", columns = c(`Election Age`, "More")) |>
  gt::cols_align(align = "center", columns = gt::everything()) |>
  gt::cols_align(align = "left", columns = c(`Candidate`)) |>
  gt::cols_width(
    "Source" ~ gt::px(80), 
    "Candidate" ~ gt::px(120), 
    "Election Year" ~ gt::px(120), 
    "Lifespan if Win" ~ gt::px(120), 
    "Lifespan if Lose" ~ gt::px(120), 
    "Election Outcome" ~ gt::px(120), 
    "Election Age" ~ gt::px(120), 
    "More" ~ gt::px(60)
  ) |>
  gt::cols_label(More = "...") |>
  gt::fmt_markdown(columns = gt::everything())

# Add footnotes only if they have content
if (!is.null(pop_title_footnote)) {
  pop_table <- pop_table |> gt::tab_footnote(footnote = pop_title_footnote, locations = gt::cells_title())
}
if (!is.null(pop_units_footnote)) {
  pop_table <- pop_table |> gt::tab_footnote(footnote = pop_units_footnote, locations = gt::cells_column_spanners(spanners = "unit_span"))
}
if (!is.null(pop_outcome_footnote)) {
  pop_table <- pop_table |> gt::tab_footnote(footnote = pop_outcome_footnote, locations = gt::cells_column_spanners(spanners = "outcome_span"))
}
if (!is.null(pop_treatment_footnote)) {
  pop_table <- pop_table |> gt::tab_footnote(footnote = pop_treatment_footnote, locations = gt::cells_column_spanners(spanners = "treatment_span"))
}
if (!is.null(pop_covariates_footnote)) {
  pop_table <- pop_table |> gt::tab_footnote(footnote = pop_covariates_footnote, locations = gt::cells_column_spanners(spanners = "covariates_span"))
}

pop_table
```

## Table Structure Details

### Preceptor Table
- Uses `p_tibble` as input (3 rows of placeholders)
- Processed by `expand_input_tibble()` to add a blank third row and "More" column
- Results in 4 total rows for the final table

### Population Table  
- Uses `d_tibble` for data input (3 rows of placeholders)
- Creates 4 data rows (3 content + 1 blank in 3rd position)
- Uses the expanded preceptor table (4 rows) 
- Combines into 11-row structure: blank + 4 data + blank + 4 preceptor + blank
- All rows are properly labeled in the Source column

### Column Alignment
The tribble code generated by `write_input_tribble()` is formatted with aligned columns to make editing easier:
- Headers and values are padded to align vertically
- Minimum column width accommodates the `"..."` placeholder
- Makes it easy to see which column you're editing

---

## Summary

The `make_p_tables()` function simplifies the creation of interpretable, spanner-labeled tables for modeling workflows. It promotes clarity, transparency, and rigor by encouraging authors to:

* Replace placeholders with meaningful values
* Use proper formatting with double quotes around all entries
* Fill in footnotes with useful context
* Take advantage of the aligned column structure for easy editing
* Understand the automatic row and column additions during rendering

This workflow supports better modeling documentation and instructional design, with helper functions ensuring consistent formatting and structure.

For more on how and why to use these tables, see:

* [The Cardinal Virtues](https://ppbds.github.io/primer/tutorials/articles/cardinal_virtues.html) article from *primer.tutorials*
