---
title: Probability
author: David Kane
tutorial:
  id: probability
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #01 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = "90%")
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

# throw_dice() simulates rolling two dice n times and returning the sum of
# each pair. The weights argument lets us model a biased die; the default
# is the fair die with each face equally likely (1/6).
throw_dice <- function(n = 1, weights = rep(1/6, 6)) {
  d1 <- sample(1:6, n, replace = TRUE, prob = weights)
  d2 <- sample(1:6, n, replace = TRUE, prob = weights)
  d1 + d2
}

# Unfair weights used in the second half of the body section: face 6 has a
# 25% chance, the other five faces share the remaining 75% equally (15% each).
unfair_weights <- c(rep(0.15, 5), 0.25)

# roll_die() simulates rolling a single die n times. Used in the
# Measuring Surprise section. The default is the fair die; the loaded
# casino die uses loaded_weights below.
roll_die <- function(n = 1, weights = rep(1/6, 6)) {
  sample(1:6, n, replace = TRUE, prob = weights)
}

# Loaded casino die: P(6) = 0.30, other faces 0.14 each. Real-world
# casino cheating is much subtler than this (typically 1--3% bias) but
# at 0.30 the discrimination is fast enough to be pedagogically tidy.
loaded_weights <- c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30)

# Helper used in the Measuring Surprise section to compute per-roll
# log-likelihood ratios (loaded vs. fair) and convert cumulative log-
# odds into posterior probability of "loaded".
log_lr <- function(rolls) {
  ifelse(rolls == 6,
         log(loaded_weights[6] / (1/6)),
         log(loaded_weights[1] / (1/6)))
}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports the [Probability](https://ppbds.github.io/primer/01-probability.html) chapter of [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

The world confronts us. Make decisions we must.

In any data analysis, three different probability distributions are in play. The first is the *PDF of the data* --- the distribution of the values we have actually observed, often shown as a histogram. The second is the *PDF of the data generating mechanism* --- the unobservable process the data came from, the "true" distribution we wish we knew. The third is the *PDF over possible data generating mechanisms* --- the set of mechanisms that are still consistent with what we have seen. Data analysis is the act of using the first to learn about the second, with the third tracking how confident we should be. This tutorial walks through some of these ideas by hand with a simple example: rolling two dice.

### Exercise 1

You should be connected to a repo named `probability`. If you are not, create one and connect to it. Create a new Quarto document titled `"Probability"` with yourself as the author, save it as `analysis.qmd`, render it, and open `analysis.html` with Live Server so the rendered HTML auto-refreshes on every later render. Going forward, we will just tell you to "Render" when we want you to take these steps. Create a `.gitignore` file with `analysis_files` on the first line followed by a blank line. Commit and push. Use AI however you like. In the R Terminal, run `show_file(".gitignore")`. If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
analysis_files

```

###

A *probability* is a number between 0 and 1 expressing how confident we are about something. The first kind of probability we will meet is the chance of a single die landing on a particular face --- a clean 1/6 if the die is fair.

### Exercise 2

In your QMD, put `library(tidyverse)` in a new code chunk. Render the file. Notice that the document does not look good because the code is visible and there are package-attach messages. Add `#| message: false` to suppress messages in this setup chunk. Also add the following to the YAML header to suppress all code echos:

````
execute:
  echo: false
````

Render again. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

````
```{r}
#| message: false
library(tidyverse)
```
````

###

"Random" in a computer is not really random --- it is a long deterministic sequence designed to *look* random. The `set.seed()` calls you will see throughout this tutorial pin the simulator to a known starting point, so the histograms in your rendered HTML will line up exactly with the ones we show as our version. Without `set.seed()`, every render would give different results --- still valid, just not reproducible.


## The Data Generating Mechanism
###

When we look at a histogram of an experiment's results, what we are really asking is *what process produced these numbers?* In this section we will build the simplest possible example of that question by hand. We have a pair of dice. We can roll them as many times as we like. Each roll gives us the sum of the two faces. From a finite stack of those sums, what can we say about the dice?

### Exercise 1

A finished `analysis.qmd` has a setup chunk (libraries) and --- as the analysis grows --- a small number of working chunks plus, when an artifact will be reused across exercises, one or two "create" chunks dedicated to those artifacts. We will need a reusable function for throwing two dice. We will build it the natural way: write the function and use it in one chunk first, and refactor for caching in the next exercise.

In `analysis.qmd`, add a new code chunk that defines a function `throw_dice(n, weights = rep(1/6, 6))` simulating `n` rolls of two dice (with the per-face probability vector `weights` defaulting to a fair die) and returning the sum of each pair as a numeric vector of length `n`. In the same chunk, set the seed with `set.seed(10)` and call `throw_dice(5)`. Render. Copy the printed five-number vector from the rendered HTML below.

```{r the-data-generating-mechanism-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 2)
```

###

```{r the-data-generating-mechanism-1-test}
#| echo: true
throw_dice <- function(n = 1, weights = rep(1/6, 6)) {
  d1 <- sample(1:6, n, replace = TRUE, prob = weights)
  d2 <- sample(1:6, n, replace = TRUE, prob = weights)
  d1 + d2
}
set.seed(10)
throw_dice(5)
```

###

The five sums you see are not the "true" probabilities of anything --- they are just five draws from the *data generating mechanism* (DGM) that produced them. The DGM here is *roll two fair dice and add the faces*; the data is the five-number vector. With only five rolls, almost any DGM could have produced the same vector. Knowing the truth (fair dice, sum from 2 to 12 with a peak at 7) does not by itself tell us anything about how clearly that truth will show up in a small sample --- and in a real problem we would not know the truth at all. The `weights` argument lets us later swap a fair die for an unfair one without redefining the function.

### Exercise 2

`throw_dice()` will be called from many later exercises, so it earns its own cached "create" chunk: a chunk that defines the function once and reuses the saved definition on every subsequent render.

Refactor your chunk from Exercise 1 into two chunks. Remove the `set.seed(10)` and `throw_dice(5)` lines from the chunk that defines `throw_dice()`, leaving only the function definition, and add `#| cache: true` to that chunk's options. Then add a new working chunk below it containing `set.seed(10)` and `throw_dice(5)`. Render. In the R Terminal, run `show_file("analysis.qmd")`. CP/CR.

```{r the-data-generating-mechanism-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

###

````
```{r}
#| cache: true
throw_dice <- function(n = 1, weights = rep(1/6, 6)) {
  d1 <- sample(1:6, n, replace = TRUE, prob = weights)
  d2 <- sample(1:6, n, replace = TRUE, prob = weights)
  d1 + d2
}
```

```{r}
set.seed(10)
throw_dice(5)
```
````

###

`#| cache: true` tells Quarto to store the function and reload it from disk on subsequent renders rather than re-defining it from scratch. The benefit is tiny for a four-line function but huge for the model fits you will meet in later tutorials --- the same discipline applies, so we get the habit in early. Cached objects live in an `analysis_cache/` directory that Quarto creates next to your QMD; we add that directory to `.gitignore` in a moment.

### Exercise 3

Modify the working chunk so that, instead of printing the five sums, it draws a histogram of 20 throws of two fair dice with the x-axis ticked at each integer sum from 2 to 12. Keep `set.seed(10)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r the-data-generating-mechanism-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r the-data-generating-mechanism-3-test}
#| echo: true
set.seed(10)
tibble(sum = throw_dice(20)) |>
  ggplot(aes(x = sum)) +
  geom_bar(fill = "grey50") +
  scale_x_continuous(breaks = 2:12) +
  labs(title = "Twenty throws of two fair dice",
       subtitle = "The shape is choppy --- 20 rolls is not enough to see the true distribution",
       x = "Sum of two dice", y = "Count") +
  theme_minimal()
```

###

A reader who saw only this histogram, without the title telling them the dice were fair, would have a hard time guessing the underlying DGM. The bumps and gaps look like they might mean *something* --- maybe the dice favor a particular sum, maybe they avoid another --- but with twenty rolls each of those features could just be noise. The histogram is the *PDF of the data*; we still know almost nothing about the *PDF of the DGM*.

### Exercise 4

Now run the same 20-throw experiment ten different times --- ten separate "shifts" of an inspector rolling the same pair of fair dice --- and plot all ten histograms in one panel so we can compare them. Modify the working chunk to plot ten such histograms together, faceted by trial. Keep `set.seed(10)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r the-data-generating-mechanism-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r the-data-generating-mechanism-4-test}
#| echo: true
set.seed(10)
tibble(trial = 1:10) |>
  mutate(throws = map(trial, ~ throw_dice(20))) |>
  unnest(throws) |>
  ggplot(aes(x = throws)) +
  geom_bar(fill = "grey50") +
  facet_wrap(~ trial, nrow = 2) +
  scale_x_continuous(breaks = 2:12) +
  labs(title = "Ten samples of 20 throws each --- fair dice",
       subtitle = "Each panel is one possible \"experiment\" we might have run",
       x = "Sum", y = "Count") +
  theme_minimal()
```

###

The ten panels are *strikingly different from each other*, even though they were all generated by the exact same DGM. Some panels are peaked, others are flat. Some have a missing sum (no 7s!) that the true DGM says should be the most common outcome. **At twenty rolls, many wildly different DGMs are consistent with the data.** This is the *PDF over data generating mechanisms* the chapter mentions: with little data, that PDF is wide --- lots of dice configurations could have produced any one of those panels.

### Exercise 5

Now do the opposite experiment: keep the dice fair and watch what happens as the number of rolls grows. Modify the working chunk to plot histograms for `n = 10, 20, 50, 100, 1000` throws of two fair dice, each panel side by side, with the y-axis allowed to vary by panel so the smaller panels are not crushed. Keep `set.seed(10)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r the-data-generating-mechanism-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r the-data-generating-mechanism-5-test}
#| echo: true
set.seed(10)
tibble(n = c(10, 20, 50, 100, 1000)) |>
  mutate(throws = map(n, throw_dice)) |>
  unnest(throws) |>
  mutate(n = factor(n, levels = c(10, 20, 50, 100, 1000))) |>
  ggplot(aes(x = throws)) +
  geom_bar(fill = "grey50") +
  facet_wrap(~ n, nrow = 1, scales = "free_y",
             labeller = labeller(n = \(x) paste0("n = ", x))) +
  scale_x_continuous(breaks = 2:12) +
  labs(title = "Sum of two fair dice as the sample grows",
       subtitle = "The empirical PDF converges to the true triangular shape",
       x = "Sum", y = "Count") +
  theme_minimal()
```

###

At `n = 10` the histogram looks almost random. At `n = 1000` it is a clean triangle, peaking at 7 with the familiar 2-and-12 corners. The DGM has not changed --- the dice were fair throughout. What changed is how clearly the *PDF of the data* (the histograms) reflects the *PDF of the DGM* (the triangle). **As `n` grows, the PDF over data generating mechanisms narrows: only DGMs that produce something close to a triangle remain consistent with what we saw.** The data has rejected a lot of candidates.

### Exercise 6

Now flip the setup. Suppose the dice are *not* fair --- each die's 6 face has a 25% chance instead of the usual 1/6, and the other five faces share the remaining 75% equally. Modify the working chunk to re-run Exercise 4's ten-panel experiment using this unfair die. Keep `set.seed(10)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r the-data-generating-mechanism-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r the-data-generating-mechanism-6-test}
#| echo: true
set.seed(10)
tibble(trial = 1:10) |>
  mutate(throws = map(trial,
                      ~ throw_dice(20, weights = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.25)))) |>
  unnest(throws) |>
  ggplot(aes(x = throws)) +
  geom_bar(fill = "grey50") +
  facet_wrap(~ trial, nrow = 2) +
  scale_x_continuous(breaks = 2:12) +
  labs(title = "Ten samples of 20 throws each --- unfair dice (P(6) = 0.25)",
       subtitle = "Can you tell at a glance that these came from unfair dice?",
       x = "Sum", y = "Count") +
  theme_minimal()
```

###

These panels look *just as messy as the fair-dice panels in Exercise 4 did*. A reader handed only these ten histograms, with no caption telling them the dice were rigged, would not be able to tell the dice apart from a fair pair. The slight bias (P(6) = 0.25 instead of 1/6) shifts the average sum by only half a point --- but the noise from drawing twenty rolls per panel is larger than that. **Many DGMs are consistent with the same sparse data**: the true unfair-dice DGM is one of them, but so is the fair-dice DGM. Twenty rolls per inspector is not enough to tell them apart.

### Exercise 7

Finally, repeat Exercise 5's `n` progression with the same unfair dice. Modify the working chunk to plot the `n = 10, 20, 50, 100, 1000` faceted histogram with the unfair-die weights. Keep `set.seed(10)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r the-data-generating-mechanism-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r the-data-generating-mechanism-7-test}
#| echo: true
set.seed(10)
tibble(n = c(10, 20, 50, 100, 1000)) |>
  mutate(throws = map(n, ~ throw_dice(.x, weights = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.25)))) |>
  unnest(throws) |>
  mutate(n = factor(n, levels = c(10, 20, 50, 100, 1000))) |>
  ggplot(aes(x = throws)) +
  geom_bar(fill = "grey50") +
  facet_wrap(~ n, nrow = 1, scales = "free_y",
             labeller = labeller(n = \(x) paste0("n = ", x))) +
  scale_x_continuous(breaks = 2:12) +
  labs(title = "Sum of two unfair dice (P(6) = 0.25) as the sample grows",
       subtitle = "The bias shows up clearly only at large n",
       x = "Sum", y = "Count") +
  theme_minimal()
```

###

At `n = 10` and `n = 20` the histograms look chaotic --- the unfair-dice pattern is indistinguishable from fair-dice noise. By `n = 100` the picture starts to lean right --- there are visibly more high sums than the fair triangle predicts. At `n = 1000` the bias is obvious: the peak is no longer at 7 but is shifted toward 8, the 12 is much taller than 1/36 of the rolls, and the whole shape is asymmetric. **The same data that could not reject "fair dice" at twenty rolls can reject it cleanly at a thousand.** Data does not directly tell us the DGM --- it tells us which DGMs are no longer plausible. That is what "learning from data" really means.


## Measuring Surprise
###

The last section ended on a slogan: *data tells us which DGMs are no longer plausible.* This section makes that idea precise. We will start from histograms --- the picture-level intuition --- and end with a number: the probability that one DGM is true rather than another, updated roll by roll.

Imagine you are at a casino's craps table. The pit boss has just brought out a single die. You have two suspicions:

- The die is **fair**: each face shows up with probability 1/6.
- The die is **loaded**: the 6 face shows up with probability 0.30, and each of the other five faces with probability 0.14. (This is more extreme than any real casino would risk --- they would get caught --- but the math is identical.)

At this moment, before you have seen any rolls, you think these two possibilities are equally likely: a *prior* of 50/50. Your job is to update that belief as the rolls come in.

### Exercise 1

Before looking at any actual rolls, we need to know what each DGM *looks like* in the limit. Start a new working chunk in `analysis.qmd` that plots two bar charts side by side: the theoretical PDF of the fair die (six bars at 1/6 each) and the theoretical PDF of the loaded die (five bars at 0.14 and a sixth bar at 0.30). Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r measuring-surprise-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r measuring-surprise-1-test}
#| echo: true
tibble(
  face = rep(1:6, 2),
  prob = c(rep(1/6, 6), c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30)),
  die  = rep(c("Fair", "Loaded"), each = 6)
) |>
  ggplot(aes(x = factor(face), y = prob)) +
  geom_col(fill = "grey50") +
  facet_wrap(~ die) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  labs(title = "Theoretical PDFs: fair die vs. loaded die",
       subtitle = "The loaded die's only difference is a taller 6 bar",
       x = "Face", y = "Probability") +
  theme_minimal()
```

###

The fair die's PDF is flat: every face at exactly 1/6. The loaded die's PDF is *almost* flat --- the five non-6 bars sit at 0.14, just a touch under 1/6 --- but the 6 bar is taller, at 0.30, about 80% taller than a fair 6. The cheat is concentrated entirely in one face. **A bar chart of the *theoretical* PDF is the reference shape. A histogram of actual rolls is a noisy approximation to that shape, and the smaller the sample, the noisier the approximation.**

### Exercise 2

Now look at twenty rolls of the loaded die. Modify the working chunk to draw 20 rolls of the loaded die, store the result as `rolls_20`, and plot a bar chart of how often each face appeared. Use `set.seed(10)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r measuring-surprise-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r measuring-surprise-2-test}
#| echo: true
set.seed(10)
rolls_20 <- sample(1:6, 20, replace = TRUE,
                   prob = c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30))

tibble(face = rolls_20) |>
  ggplot(aes(x = factor(face, levels = 1:6))) +
  geom_bar(fill = "grey50") +
  labs(title = "Twenty rolls of the loaded die",
       subtitle = "The 6 bar looks tall --- but the sample is small enough that this could be luck",
       x = "Face", y = "Count") +
  theme_minimal()
```

###

The 6 bar is the tallest, which is what the loaded DGM would predict --- but with only 20 rolls, "tallest" is not strong evidence. A fair die over 20 rolls produces tall bars in random places all the time. You might be at 70% or 80% on "loaded" looking at this picture, but you would not bet at 95%. **Twenty rolls is the "suggestive but not determinative" region the chapter warns about: many DGMs are still consistent with the data.**

### Exercise 3

Now look at two hundred rolls of the same loaded die. Modify the working chunk to repeat Exercise 2's exercise with 200 rolls instead of 20, storing the result as `rolls_200`. Use `set.seed(10)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r measuring-surprise-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r measuring-surprise-3-test}
#| echo: true
set.seed(10)
rolls_200 <- sample(1:6, 200, replace = TRUE,
                    prob = c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30))

tibble(face = rolls_200) |>
  ggplot(aes(x = factor(face, levels = 1:6))) +
  geom_bar(fill = "grey50") +
  labs(title = "Two hundred rolls of the loaded die",
       subtitle = "The 6 bar is now clearly taller than the other five",
       x = "Face", y = "Count") +
  theme_minimal()
```

###

The 6 bar is plainly taller than the others, and the five non-6 bars are roughly the same height as each other. This is exactly the loaded-die PDF showing through. **But "plainly" is not a probability.** If a friend asked you *"how sure are you?"* the histogram cannot give a number --- it can only say "very." To produce a real percentage we need to leave the picture and start scoring individual rolls.

### Exercise 4

Each individual roll either supports the loaded-die hypothesis or undermines it. Modify the working chunk to print a small tibble with one row per face listing the **likelihood ratio** of that face under loaded versus fair: $P(\text{face} \mid \text{loaded}) / P(\text{face} \mid \text{fair})$. Render. Copy the printed tibble from the rendered HTML below.

```{r measuring-surprise-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r measuring-surprise-4-test}
#| echo: true
tibble(face = 1:6,
       p_loaded = c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30),
       p_fair   = 1/6,
       likelihood_ratio = round(p_loaded / p_fair, 3))
```

###

Rolling a 6 is **1.8 times more likely** under loaded than under fair; rolling any other face is **0.84 times as likely** under loaded. Every individual roll is a vote: a 6 votes for "loaded" (multiply our belief by 1.8); any non-6 votes for "fair" (multiply by 0.84). **This is the per-roll surprise the chapter introduces.** Histograms hide it inside a picture; the likelihood ratio puts the surprise on a number line.

### Exercise 5

The next few exercises combine many per-roll likelihood ratios into a single posterior probability. The cleanest way to do that is to *sum the logs* (not multiply the ratios): logs add instead of multiply, which is numerically stabler over hundreds of rolls and easier to plot. With a 50/50 prior and independent rolls, the math is simple: sum the per-roll log likelihood ratios, then convert the resulting log-odds to a probability via $1 / (1 + e^{-\text{log-odds}})$.

Modify the working chunk to define a helper function `log_lr(rolls)` that, given a vector of rolls (each value 1-6), returns a vector of per-roll log likelihood ratios of loaded over fair (using the loaded weights P(6) = 0.30, others 0.14). Then in the same chunk, use `log_lr()` on `rolls_20` from Exercise 2 to print the posterior probability that the die is loaded, rounded to four decimals. Render. Copy the printed number from the rendered HTML below.

```{r measuring-surprise-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```{r measuring-surprise-5-test}
#| echo: true
log_lr <- function(rolls) {
  loaded <- c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30)
  log(loaded[rolls] / (1/6))
}
set.seed(10)
rolls_20 <- sample(1:6, 20, replace = TRUE,
                   prob = c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30))
posterior_loaded_20 <- 1 / (1 + exp(-sum(log_lr(rolls_20))))
round(posterior_loaded_20, 4)
```

###

Twenty rolls put our posterior at about 0.86 --- meaningfully past 50/50, but well short of the 95% threshold we set for ourselves. Our intuition from the bar chart in Exercise 2 ("looks loaded, but I would not bet") corresponds, on the number line, to "0.86." The picture and the number now agree, but the number is the thing you can act on.

### Exercise 6

`log_lr()` will be reused across the next three exercises, so it earns its own cached create chunk.

Refactor your working chunk into two chunks. Move the `log_lr` definition into its own create chunk (with `#| cache: true`) and leave the `rolls_20` regeneration plus the posterior calculation in the working chunk. Render. In the R Terminal, run `show_file("analysis.qmd")`. CP/CR.

```{r measuring-surprise-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

###

````
```{r}
#| cache: true
log_lr <- function(rolls) {
  loaded <- c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30)
  log(loaded[rolls] / (1/6))
}
```

```{r}
set.seed(10)
rolls_20 <- sample(1:6, 20, replace = TRUE,
                   prob = c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30))
posterior_loaded_20 <- 1 / (1 + exp(-sum(log_lr(rolls_20))))
round(posterior_loaded_20, 4)
```
````

###

`log_lr()` is now in its own create chunk: defined once, cached, reused by every later exercise without re-defining the function on every render. Same refactoring pattern as `throw_dice()` in the previous section --- and the model fits in later tutorials.

### Exercise 7

Now do the same calculation for the 200-roll sample. Repeat Exercise 5's posterior calculation with `rolls_200` from Exercise 3, rounded to six decimals. Render. Copy the printed number from the rendered HTML below.

```{r measuring-surprise-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```{r measuring-surprise-7-test}
#| echo: true
set.seed(10)
rolls_200 <- sample(1:6, 200, replace = TRUE,
                    prob = c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30))
posterior_loaded_200 <- 1 / (1 + exp(-sum(log_lr(rolls_200))))
round(posterior_loaded_200, 6)
```

###

Two hundred rolls leave essentially zero doubt --- the posterior is indistinguishable from 1. The data has rejected the fair-die DGM as effectively as the data ever could. **Notice what the calculation produces: not "the die is loaded" but "the probability that the die is loaded, given what we have seen, is 0.9999."** Belief is on a continuum; certainty is a useful limit, not a starting point.

### Exercise 8

The most interesting question is not what the posterior is at the *end* of 200 rolls but how it *evolves*, roll by roll, from 0.50. Modify the working chunk to plot the running posterior over the 200 rolls in `rolls_200` --- one point per roll, with a horizontal reference line at 0.95. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r measuring-surprise-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r measuring-surprise-8-test}
#| echo: true
set.seed(10)
rolls_200 <- sample(1:6, 200, replace = TRUE,
                    prob = c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30))

tibble(
  roll      = 1:200,
  log_odds  = cumsum(log_lr(rolls_200)),
  posterior = 1 / (1 + exp(-log_odds))
) |>
  ggplot(aes(x = roll, y = posterior)) +
  geom_line() +
  geom_hline(yintercept = 0.95, color = "red", linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1),
                     labels = scales::label_percent(accuracy = 1)) +
  labs(title = "Posterior probability that the die is loaded, roll by roll",
       subtitle = "Starts at 50%; each roll nudges the line up (a 6) or down (anything else)",
       x = "Roll number",
       y = "Posterior probability of \"loaded\"") +
  theme_minimal()
```

###

The line starts at 50%, wobbles a bit while every individual roll matters, then climbs --- with steps up at every 6 and small steps down at every non-6 --- past the 95% line somewhere in the first few dozen rolls and stays there. **This plot *is* the third PDF the chapter talked about**: our belief about which DGM is true, updated in real time. The fair-die line would be the mirror image, climbing down to zero on the same data.

### Exercise 9

One run is one run. To see how *typical* that run was, modify the working chunk to repeat the experiment 1,000 times: for each repetition, draw a long sequence of rolls from the loaded die, find the first roll at which the running posterior first crosses 0.95, and record how many rolls that took. Then plot a histogram of those 1,000 stopping times. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r measuring-surprise-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r measuring-surprise-9-test}
#| echo: true
set.seed(42)
stop_times <- replicate(1000, {
  rolls    <- sample(1:6, 500, replace = TRUE,
                     prob = c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30))
  log_odds <- cumsum(log_lr(rolls))
  post     <- 1 / (1 + exp(-log_odds))
  which(post >= 0.95)[1]
})

tibble(stop_time = stop_times[!is.na(stop_times)]) |>
  ggplot(aes(x = stop_time)) +
  geom_histogram(binwidth = 10, fill = "grey50", color = "white") +
  geom_vline(xintercept = median(stop_times, na.rm = TRUE),
             color = "red", linetype = "dashed") +
  labs(title = "Rolls needed to reach 95% posterior on \"loaded\"",
       subtitle = "1,000 simulated runs; red dashed line is the median",
       x = "Rolls until posterior >= 0.95",
       y = "Count") +
  theme_minimal()
```

###

The median is about 42 rolls, but the distribution stretches from the teens out past 100. Some runs hit 95% in 20 rolls because the early sixes piled up; others took 80 or more because the early rolls happened to be non-sixes. **The DGM is the same in every run --- but how long it takes the data to *prove* it varies enormously.** This is the third PDF in motion one more time: the same truth, refracted through different finite samples, gives different histories of belief.

The whole arc you just walked --- two competing DGMs, a histogram-level intuition, a per-roll likelihood ratio, an updating posterior, a time series of belief, and a distribution over how long belief takes to settle --- is the engine that the rest of the Primer rides on. Every later chapter is a more elaborate version of this same story: write down candidate DGMs, see what the data implies for each, and let the posterior do the choosing.


## Summary
###

This tutorial walked through the engine that every later chapter of the Primer rides on. You wrote a function that turns randomness into data and watched many small samples scramble the underlying *data generating mechanism* while a single large sample reveals it cleanly (Topic 1). Then you put two competing DGMs side by side, scored each individual roll for how surprising it would be under each, and watched the posterior probability of "loaded" walk from 50% to near-certainty as the rolls accumulated (Topic 2). The three PDFs the chapter named --- *of the data*, *of the DGM*, *over the DGMs* --- are no longer just labels; you saw all three appear in actual plots.

### Exercise 1

Render `analysis.qmd` one last time to confirm everything works end to end. In the R Terminal, run `show_file("analysis.qmd")`. CP/CR.

```{r summary-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

````
---
title: "Probability"
author: "<your name>"
execute:
  echo: false
---

```{r}
#| message: false
library(tidyverse)
```

```{r}
#| cache: true
throw_dice <- function(n = 1, weights = rep(1/6, 6)) {
  d1 <- sample(1:6, n, replace = TRUE, prob = weights)
  d2 <- sample(1:6, n, replace = TRUE, prob = weights)
  d1 + d2
}
```

```{r}
#| cache: true
log_lr <- function(rolls) {
  loaded <- c(0.14, 0.14, 0.14, 0.14, 0.14, 0.30)
  log(loaded[rolls] / (1/6))
}
```

(The final working chunk from Measuring Surprise: 1,000 simulated runs
of rolls-to-95%-posterior, plotted as a histogram.)
````

###

The published version of your QMD is the one your instructor --- and your future self --- will look at most closely. Make sure it tells a clean story end to end: the histograms in Topic 1, the loaded-die plots in Topic 2, the time-series of belief, the stopping-time distribution.

### Exercise 2

Publish your rendered QMD to GitHub Pages. In the bash Terminal, run:

```
quarto publish gh-pages analysis.qmd
```

Paste the resulting URL below.

```{r summary-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
https://<your-github-username>.github.io/probability/analysis.html
```

###

This URL is now a permanent, shareable record of your probability work --- live data science in the open.

### Exercise 3

Commit and push any remaining changes. Paste the URL of your GitHub repo below.

```{r summary-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
https://github.com/<your-github-username>/probability
```

###

The world confronts us. Make decisions we must.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
