---
title: Probability
author: David Kane and Ivy Spratt
tutorial:
  id: probability
output:
  learnr::tutorial:
    progressive: yes
    allow_skip:: yes
runtime: shiny_prerendered
description: "Tutorial #01 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(primer.data)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 



```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```



<!-- At a later date, we might go and add material about probability distributions, standard errors and so on. But probably not this year. -->


## Introduction
### 

This tutorial covers [Chapter 2: Probability](https://ppbds.github.io/primer/probability.html) of [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/). 

The world confronts us. Make decisions we must.

This tutorial will introduce the four Cardinal Virtues of Wisdom, Justice, Courage, and Temperance. It will cover key concepts such as validity, stability, representativeness, and unconfoundedness.

## Example 1
###

Imagine you are a high school principal today who wants to forecast how well the juniors in his school will do on the SAT. (Some of the students have taken the PSAT and some have not.) 

The data is from a different high school from the class of 2015. It includes all the members of the graduating class who went on to college. The outcome of interest is each student's SAT score (math + verbal, scores can range from 400 to 1600). The covariates include their sex, high school grades, PSAT scores, the average SAT scores of their parents, their family income, how many hours they studied for the SAT, whether or not they took an SAT tutoring class, and where they were going to college.

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r example-1-1}
question_text(NULL,
	message = "Wisdom requires the creation of a Preceptor Table, an examination of our data, and a determination, using the concept of validity, as to whether or not we can (reasonably!) assume that the two come from the same population.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The central problem for Wisdom is: Can we use data from a different high school's class of 2015 to predict the variables/relationships in juniors' SAT scores?

### Exercise 2

Define a Preceptor Table.

```{r example-1-2}
question_text(NULL,
	message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The Preceptor Table does not include all the covariates which you will eventually include in your model. It only includes covariates which you need to answer your question.

### Exercise 3

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r example-1-3}
question_text(NULL,
	message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will considered a treatment.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

This problem is predictive, so we are trying to determine the SAT scores (outcomes), knowing that different sorts of students, as defined by various covariates, have diffferent final GPAs.

### Exercise 4

What are the units for this problem?

```{r example-1-4}
question_text(NULL,
	message = "The individual students.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Specifying the Preceptor Table forces us to think clearly about the units and outcomes implied by the question. The resulting discussion sometimes leads us to modify the question with which we started. No data science project follows a single direction. We always backtrack. There is always dialogue.

### Exercise 5

What is the outcome variable for this problem?

```{r example-1-5}
question_text(NULL,
	message = "SAT scores.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### Exercise 6

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r example-1-6}
question_text(NULL,
	message = "PSAT scores give a good indication of how well a student is going to do on the SAT, given that it is a pre-test and usually covers the same format and similar content.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The term "covariates" is used in at least three ways in data science. First, it is all the variables which might be useful, regardless of whether or not we have the data. Second, it is all the variables for which we have data. Third, it is the set of variables in the data which we end up using in the model.

### Exercise 7

What are the treatments, if any, for this problem?

```{r example-1-7}
question_text(NULL,
	message = "Since this is a predictive model, there is no treatment.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Remember that a treatment is just another covariate which, for the purposes of this specific problem, we are assuming can be manipulated, thereby, creating two or more different potential outcomes for each unit.

### Exercise 8

What moment in time does the Preceptor Table refer to?

```{r example-1-8}
question_text(NULL,
	message = "The Preceptor Table refers to the current moment, since we are trying to predict the future SAT scores of current juniors.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```


### Exercise 9

Define a causal effect. Note that the model in this tutorial is predictive, not causal. We just want to make sure you understand what a causal model is.

```{r example-1-9}
question_text(NULL,
	message = "A causal effect is the difference between two potential outcomes.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

According to the Rubin Causal Model, there must be two (or more) potential outcomes for any discussion of causation to make sense. This is simplest to discuss when the treatment only has two different values, thereby generating only two potential outcomes. 

### Exercise 10

What is the fundamental problem of causal inference?

```{r example-1-10}
question_text(NULL,
	message = "The fundamental problem of causal inference is that we can only observe one potential outcome.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

If the treatment variable is continuous (like income), then there are lots and lots of potential outcomes, one for each possible value of the treatment variable. 

### Exercise 11

How does the motto "No causation without manipulation" apply in this problem?

```{r example-1-11}
question_text(NULL,
	message = "The motto does not apply because this is a predictive, not causal, model.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```


### Exercise 12

Describe in words the Preceptor Table for this problem.

```{r example-1-12}
question_text(NULL,
	message = "The rows (units) are individual students. The outcome is the SAT score. Key covariates include high school grades, PSAT scores, how many hours they studied for the SAT, and whether or not they took an SAT tutoring class.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```


### Exercise 13

In your own words, define "validity" as we use the term.

```{r example-1-13}
question_text(NULL,
	message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Validity is always about the *columns* in the Preceptor Table and the data. Just because columns from these two different tables have the same *name* does not mean that they are the same *thing*.

### Exercise 14

Provide one reason why the assumption of validity might not hold for the outcome variable (SAT score) or for one of the covariates. Use the words "column" or "columns" in your answer.

```{r example-1-14}
question_text(NULL,
	message = "The high school grades column in the data covers all four years of high school — the class of 2015 has already graduated. But the high school grades column in the Preceptor Table can only include grades through junior year, since the current students have not yet finished high school. These two columns are measuring different things — cumulative four-year GPA versus cumulative three-year GPA — and so may not correspond to the same concept.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

In order to consider the Preceptor Table and the data to be drawn from the same population, the columns from one must have a *valid correspondence* with the columns in the other. Validity, if true (or at least reasonable), allows us to construct the Population Table, which is the first step in Justice.

Because we control the Preceptor Table and, to a lesser extent, the original question, we can adjust those variables to be “closer” to the data that we actually have. This is another example of the iterative nature of data science. If the data is not close enough to the question, then we check with our boss/colleague/customer to see if we can modify the question in order to make the match between the data and the Preceptor Table close enough for validity to hold.

Despite these potential problems, we will assume that validity holds since it, mostly (?), does.

### Exercise 15

In your own words, name the four key components of Justice when working on a data science problem.

```{r example-1-15}
question_text(NULL,
	message = "Justice concerns four topics: the Population Table, stability, representativeness, and unconfoundedness.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Justice is about concerns that you (or your critics) might have, reasons why the model you create might not work as well as you hope. 

### Exercise 16

In your own words, define a Population Table.

```{r example-1-16}
question_text(NULL,
	message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The Population Table is almost always much bigger than the combination of the Preceptor Table and the data because, if we can really assume that both the Preceptor Table and the data are part of the same population, then that population must cover a broad universe of time and units since the Preceptor Table and the data are, themselves, often far apart from each other.

### Exercise 17

In your own words, define the assumption of "stability" when employed in the context of data science.

```{r example-1-17}
question_text(NULL,
	message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The longer the time period covered by the Preceptor Table (and the data), the more suspect the assumption of stability becomes. 

### Exercise 18

Provide one reason why the assumption of stability might not be true in this case.

```{r example-1-18}
question_text(NULL,
	message = "The topics that the SAT covers, as well as the formatting and structure of the test, has changed a lot since 2015. This means that it may have been easier or harder to score higher back in the day, making the results unstable over time.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

A change in time or the distribution of the data does not, in and of itself, demonstrate a violation of stability. Stability is about the parameters: $\beta_0$, $\beta_1$ and so on. Stability means these parameters are the same in the data as they are in the population as they are in the Preceptor Table.

### Exercise 19

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: `Data -> Population -> Preceptor Table`. In your own words, define the assumption of "representativeness" when employed in the context of data science.

```{r example-1-19}
question_text(NULL,
	message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Ideally, we would like both the Preceptor Table *and* our data to be random samples from the population. Sadly, this is almost never the case.

### Exercise 20

We do not use the data, directly, to estimate missing values in the Preceptor Table. Instead, we use the data to learn about the overall population. Provide one reason, involving the relationship between the data and the population, for why the assumption of representativeness might not be true in this case. 

```{r example-1-20}
question_text(NULL,
	message = "We are interested in the SAT scores of all high school juniors. The data only includes information from students who graduated and were going to college, which leaves out students who didn't fit the requirements to graduate or were going on to do other things. It is not representative of the larger population of juniors.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The reason that representativeness is important is because, when it is violated, the estimates for the model parameters might be biased. 

### Exercise 21

We use information about the population to make inferences about the Preceptor Table. Provide one reason, involving the relationship between the Population and the Preceptor Table, for why the assumption of representativeness might not be true in this case.

```{r example-1-21}

question_text(NULL,
	message = "The current juniors, in the Preceptor Table, may have different demographics and better curriculum than those in the Population Table.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Stability looks across time periods. Representativeness looks within time periods, for the most part. 

### Exercise 22


In your own words, define the assumption of "unconfoundedness" when employed in the context of data science.

```{r example-1-22}
question_text(NULL,
	message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

This assumption is only relevant for causal models. We describe a model as "confounded" if this is not true. The easiest way to ensure unconfoundedness is to assign treatment randomly.

### Exercise 23


In your own words, describe the components of the virtue of Courage for analyzing data.

```{r example-1-23}
question_text(NULL,
	message = "Courage starts with math, explores models, and then creates the data generating mechanism.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

###

Since this is an introduction to the different virtues and things to think about, we are not yet ready to code. Don't worry about Courage for right now.

### Exercise 24

In your own words, describe the use of Temperance in data science.

```{r example-1-24}
question_text(NULL,
	message = "Temperance uses the data generating mechanism to answer the question with which we began. Humility reminds us that this answer is always a lie. We can also use the DGM to calculate many similar quantities of interest, displaying the results graphically.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Courage gave us the data generating mechanism. Temperance guides us in the use of the DGM — or the “model” — we have created to answer the question(s) with which we began. We create posteriors for the quantities of interest. 

Once again, we aren't making a data generating mechanism, so we can't answer our question! However, if we were to do so, we would have considered all the important concepts to concluding with a more accurate answer.

## Example 2
###

Imagine you are a high school principal at a different high school who is considering providing an SAT tutoring class to all the juniors in her school. What will happen to SAT scores if he does so?

The data is from a high school from the class of 2015. It includes all the members of the graduating class who went on to college. The outcome of interest is each student's SAT score (math + verbal, scores can range from 400 to 1600). The covariates include their sex, high school grades, PSAT scores, the average SAT scores of their parents, their family income, how many hours they studied for the SAT, whether or not they took an SAT tutoring class, and where they were going to college.

### Exercise 1

What are the units for this problem?

```{r example-2-1}
question_text(NULL,
	message = "The individual students.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### Exercise 2

What is the outcome variable for this problem?

```{r example-2-2}
question_text(NULL,
	message = "SAT scores.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### Exercise 3

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r example-2-3}
question_text(NULL,
	message = "How many hours each student studied for the SAT is a good indication of how well a student will do on the SAT, as if they study more, they are likely to do better.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The term "covariates" is used in at least three ways in data science. First, it is all the variables which might be useful, regardless of whether or not we have the data. Second, it is all the variables for which we have data. Third, it is the set of variables in the data which we end up using in the model.

### Exercise 4

What are the treatments, if any, for this problem?

```{r example-2-4}
question_text(NULL,
	message = "Whether or not the student took an SAT tutoring class.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Remember that a treatment is just another covariate which, for the purposes of this specific problem, we are assuming can be manipulated, thereby, creating two or more different potential outcomes for each unit.

### Exercise 5

What moment in time does the Preceptor Table refer to?


```{r example-2-5}
question_text(NULL,
	message = "The Preceptor Table refers to the current moment, since we are trying to predict the future SAT scores of current juniors.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### Exercise 6

How does the motto "No causation without manipulation" apply in this problem?

```{r example-2-6}
question_text(NULL,
	message = "We can only find the causal effect if we manipulate the treatment (either give or don't give the student SAT tutoring). It is easy to do this, as we can offer it to some students and not others.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### Exercise 7

Describe in words the Preceptor Table for this problem.


```{r example-2-7}
question_text(NULL,
	message = "The rows (units) are individual students. The outcome is the SAT score. There needs to be a column for the treatment, which is whether or not the student recieved SAT tutong. Other key covariates include high school grades, PSAT scores, how many hours they studied for the SAT, and whether or not they took an SAT tutoring class.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### Exercise 8

Provide one reason why the assumption of validity might not hold for the outcome variable (SAT score) or for one of the covariates. Use the words "column" or "columns" in your answer.

```{r example-2-8}
question_text(NULL,
	message = "SAT tutoring styles has advanced a lot since 2015, so we are probably not covering the same content as before. In this case, the SAT tutoring column from the Preceptor Table and the data may not correspond completely.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

In order to consider the Preceptor Table and the data to be drawn from the same population, the columns from one must have a *valid correspondence* with the columns in the other. Validity, if true (or at least reasonable), allows us to construct the Population Table, which is the first step in Justice.

Because we control the Preceptor Table and, to a lesser extent, the original question, we can adjust those variables to be “closer” to the data that we actually have. This is another example of the iterative nature of data science. If the data is not close enough to the question, then we check with our boss/colleague/customer to see if we can modify the question in order to make the match between the data and the Preceptor Table close enough for validity to hold.

### Exercise 9

Provide one reason why the assumption of stability might not be true in this case.


```{r example-2-9}
question_text(NULL,
	message = "What if the student class of 2015 was smarter, on average, than the current junior class that we are trying to calculate scores for? In this case, the stability assumption would not hold.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

A change in time or the distribution of the data does not, in and of itself, demonstrate a violation of stability. Stability is about the parameters: $\beta_0$, $\beta_1$ and so on. Stability means these parameters are the same in the data as they are in the population as they are in the Preceptor Table.

### Exercise 10

We do not use the data, directly, to estimate missing values in the Preceptor Table. Instead, we use the data to learn about the overall population. Provide one reason, involving the relationship between the data and the population, for why the assumption of representativeness might not be true in this case. 


```{r example-2-10}
question_text(NULL,
	message = "We are interested in the SAT scores of all high school juniors. The data only includes information from students who graduated and were going to college, which leaves out students who didn't fit the requirements to graduate or were going on to do other things. It is not representative of the larger population of juniors.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The reason that representativeness is important is because, when it is violated, the estimates for the model parameters might be biased. 

### Exercise 11

We use information about the population to make inferences about the Preceptor Table. Provide one reason, involving the relationship between the Population and the Preceptor Table, for why the assumption of representativeness might not be true in this case.

```{r example-2-11}

question_text(NULL,
	message = "The current juniors, in the Preceptor Table, may have different demographics and better curriculum than those in the Population Table.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

Stability looks across time periods. Representativeness looks within time periods, for the most part. 

### Exercise 12

In your own words, define the assumption of "unconfoundedness" when employed in the context of data science.

```{r example-2-12}
question_text(NULL,
	message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

This assumption is only relevant for causal models. We describe a model as "confounded" if this is not true. The easiest way to ensure unconfoundedness is to assign treatment randomly.

### Exercise 13

Provide one reason why the assumption of unconfoundedness might not be true (or relevant) in this case.

```{r example-2-13}
question_text(NULL,
	message = "The assignment might not be completely random. SAT tutoring is most helpful to those in desperate need of help, so students who may not be as intelligent may be more likely to take it than students who are smarter. This might lead to some false conclusions, as the average of their SAT scores after tutoring may be lower, because they didn't have much to build off in the first place.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

The great advantage of randomized assignment of treatment is that it guarantees unconfoundedness, *if the randomization is done correctly*. There is no way for treatment assignment to be correlated with anything, including potential outcomes, if treatment assignment is random, and *if the experimental set up worked as designed.* Sadly, in the real world, there are sometimes problems.


## Summary
### 

This tutorial covered [Chapter 2: Probability](https://ppbds.github.io/primer/probability.html) of [*Preceptor’s Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).  

This tutorial introduced the four Cardinal Virtues of Wisdom, Justice, Courage, and Temperance. It covered key concepts such as validity, stability, representativeness, and unconfoundedness.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
