---
title: "Seguro Popular"
author: David Kane
tutorial:
  id: 09-sps
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #09 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(tidymodels)
library(primer.data)
library(broom)
library(marginaleffects)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

set.seed(2026)
x <- sps |>
  select(treatment, t2_health_exp_3m, age, sex, education) |>
  drop_na() |>
  mutate(treatment = factor(if_else(treatment == 1, "Enrolled", "Not enrolled"),
                            levels = c("Not enrolled", "Enrolled"))) |>
  group_by(treatment) |>
  slice_sample(n = 250) |>
  ungroup()

fit_sps <- linear_reg() |>
  set_engine("lm") |>
  fit(t2_health_exp_3m ~ treatment, data = x)
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

The world confronts us. Make decisions we must.

Imagine that you are a health-policy analyst at Mexico's Ministry of Health. The Minister has a big-picture goal --- improve Mexican health outcomes within a limited budget --- and trusts you to find the evidence that helps her decide where the money should go. Seguro Popular, the universal health-insurance program Mexico randomized into communities starting in 2005, is one of many programs competing for that budget. Several causal and predictive models would inform the choice --- how enrollment changes household spending, hospital usage, mortality, debt. This tutorial builds just one of them: the causal effect of enrollment on out-of-pocket health spending. The estimate alone won't decide anything, but it is one good input to the larger decision. There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

QMD World and R World are not the same. Several R sessions can run side by side --- one in your QMD render, one at your interactive prompt, one inside this tutorial --- each with its own workspace.

### Exercise 2

You should be working inside a GitHub repo named `sps`, opened in a Codespace from the [`PPBDS/codespace-starter`](https://github.com/PPBDS/codespace-starter) template. If you are not there yet, please create that repo and open it in a Codespace now --- see the [package README](https://github.com/PPBDS/primer/tree/main/primer.tutorials#working-environments-and-repo-setup) if you need setup instructions or are working locally instead.

Once you're inside the `sps` repo, create a new Quarto document titled `"Seguro Popular"` with yourself as the author, render it, and save it as `analysis.qmd`.

Create a `.gitignore` file with `analysis_files` on the first line followed by a blank line. Save and push.

At the R prompt, run:

```
show_file(".gitignore")
```

If that fails, it's probably because you haven't loaded `library(tutorial.helpers)` at the R prompt.

CP/CR.

```{r introduction-2}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Professionals keep their data science work in the cloud because laptops fail.

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.data)` in a new code chunk. Render the file. Add `#| message: false` to suppress the messages, and add the following to the YAML header to hide the code:

```
execute:
  echo: false
```

Render again. At the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    #| message: false\n    library(tidyverse)\n    library(primer.data)\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Rendering a QMD converts code and prose into a self-contained HTML document. The render runs in a fresh R session, so it doesn't see anything you have loaded at your interactive prompt --- a feature, not a bug.

### Exercise 4

Place your cursor on the `library(tidyverse)` line. Use `Cmd/Ctrl + Enter` to send the line to the R prompt.

CP/CR.

```{r introduction-4}
question_text(NULL,
  message = "    > library(tidyverse)\n    ── Attaching core tidyverse packages ─ tidyverse 2.0.0 ─\n    ✔ dplyr     ✔ forcats   ✔ ggplot2  ✔ lubridate\n    ✔ purrr     ✔ readr     ✔ stringr  ✔ tibble     ✔ tidyr\n    ── Conflicts ─ tidyverse_conflicts() ─\n    ✖ dplyr::filter() masks stats::filter()\n    ✖ dplyr::lag()    masks stats::lag()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The "Conflicts" lines name functions that exist in more than one loaded package. `dplyr::filter()` masks `stats::filter()` because tidyverse loads later; the last-loaded package wins.

### Exercise 5

Place your cursor on the `library(primer.data)` line. Use `Cmd/Ctrl + Enter`.

CP/CR.

```{r introduction-5}
question_text(NULL,
  message = "    > library(primer.data)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Data on Mexico's Seguro Popular randomized rollout is available in the `sps` tibble from the **[primer.data](https://github.com/PPBDS/primer.data)** package. The data came from a large 2005--06 trial in which the Mexican Ministry of Health randomly assigned communities to receive Seguro Popular early. We have one row per surveyed household, with information about enrollment, demographics, and out-of-pocket health expenditures.

### Exercise 6

At the R prompt, type `?sps` and paste the Description below.

```{r introduction-6}
question_text(NULL,
  message = "Data from a 2005--2006 health-insurance experiment in Mexico in which Seguro Popular was randomly rolled out across communities. Households were surveyed about their out-of-pocket spending on health care, with measurements taken at baseline and again roughly ten months later. The randomization makes this an unusually clean setting for estimating the causal effect of universal health insurance on household financial protection.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The original analysis is King et al., *"Public policy for the poor? A randomised assessment of the Mexican universal health insurance programme,"* The Lancet, 2009.

### Exercise 7

Define a causal effect.

```{r introduction-7}
question_text(NULL,
  message = "A causal effect is the difference between two potential outcomes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

According to the Rubin Causal Model, there must be two (or more) potential outcomes for any discussion of causation to make sense. This is simplest when the treatment has only two values, generating two potential outcomes per unit.

### Exercise 8

What is the fundamental problem of causal inference?

```{r introduction-8}
question_text(NULL,
  message = "The fundamental problem of causal inference is that we can only observe one potential outcome.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

For each household in the SPS data, we see what their out-of-pocket spending was *under the treatment they actually received* --- enrolled or not enrolled. We never see what the same household's spending would have been under the other arm. The causal effect for any single household is unobservable; we estimate the average causal effect across households.

### Exercise 9

Out-of-pocket health expenditure is the broad topic of this tutorial. Given that topic, which variable in `sps` should we use as our outcome variable?

```{r introduction-9}
question_text(NULL,
  message = "Our outcome variable is `t2_health_exp_3m` --- the household's reported total out-of-pocket health spending over the three months before the follow-up survey, in Mexican pesos.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

We will use `t2_health_exp_3m` as our outcome variable. Below is the distribution of follow-up spending in our 500-household sample, by treatment status.

```{r}
#| echo: false
#| message: false
#| warning: false
x |>
  ggplot(aes(x = t2_health_exp_3m, fill = treatment)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(labels = scales::label_dollar(prefix = "$"), limits = c(0, 5000)) +
  labs(
    title = "Out-of-Pocket Health Spending in the SPS Sample",
    subtitle = "Most households spend little; both arms have a long right tail",
    x = "3-month spending (pesos)",
    y = NULL,
    fill = "Treatment",
    caption = "Source: King et al. (2009) SPS rollout via primer.data"
  ) +
  theme_minimal() +
  theme(axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())
```

### Exercise 10

Our treatment variable will be `treatment`, which takes two values. Describe briefly what each value means and how it was manipulated.

```{r introduction-10}
question_text(NULL,
  message = "`treatment` is `Enrolled` if the household was in a community whose Seguro Popular rollout had begun by the time of the follow-up survey, and `Not enrolled` otherwise. The Mexican Ministry of Health manipulated this variable by randomly choosing which communities to roll out first, so households did not select into treatment based on their own characteristics.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Any data set can be used to construct a causal model as long as there is at least one covariate that we can, at least in theory, manipulate. It does not matter whether or not anyone did, in fact, manipulate it --- though in this case the Ministry of Health really did.

### Exercise 11

Given our treatment variable, how many potential outcomes are there for each household? Explain why.

```{r introduction-11}
question_text(NULL,
  message = "There are two potential outcomes for each household because the treatment takes two values: `Enrolled` and `Not enrolled`. Each household has the spending it would have under enrollment and the spending it would have without enrollment; we observe only one of those, depending on which arm the household ended up in.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The same data set can be used to create lots of different models, both causal and predictive. We just use different outcome variables and/or specify different treatments. This is a conceptual frame we apply to data; it is never inherent in the data itself.

### Exercise 12

Pick a single household from the data. Specify the two values its treatment could take, guess the potential outcome under each, and compute the causal effect for that household.

```{r introduction-12}
question_text(NULL,
  message = "Take household 102 in our sample. Its treatment value `Enrolled` would yield three-month spending of, say, 400 pesos, while `Not enrolled` would yield 600 pesos. The causal effect on out-of-pocket spending of enrollment versus no enrollment is 400 - 600 = -200 pesos: enrolling reduces this household's spending by 200 pesos. Both numbers are guesses; we observe only one of them in the data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A causal effect is the difference between two potential outcomes. *"Difference"* doesn't have to mean *"subtraction"* --- many potential outcomes are not numbers --- but here both are pesos, so subtraction works.

### Exercise 13

Write a causal question connecting the outcome `t2_health_exp_3m` to `treatment`.

```{r introduction-13}
question_text(NULL,
  message = "What is the average causal effect of Seguro Popular enrollment on a household's three-month out-of-pocket health spending?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The variable `t2_health_exp_3m` is total household health spending over the three months before the follow-up survey: pharmacy purchases, doctor visits, hospital co-pays, the works. It does *not* include the premium households would have paid for Seguro Popular itself, so the question is about *care-related* spending, not the total cost of getting covered.


## Wisdom
###

Data science starts with some broad questions and a data set which might help us to answer them.

Imagine that you are a health-policy analyst at Mexico's Ministry of Health. The Minister has a big-picture goal --- improve Mexican health outcomes within a limited budget --- and trusts you to find the evidence that helps her decide where the money should go. Seguro Popular, the universal health-insurance program Mexico randomized into communities starting in 2005, is one of many programs competing for that budget. Several causal and predictive models would inform the choice --- how enrollment changes household spending, hospital usage, mortality, debt. This tutorial builds just one of them: the causal effect of enrollment on out-of-pocket health spending. The estimate alone won't decide anything, but it is one good input to the larger decision. There are many decisions to make.

The specific question: *What is the average causal effect of Seguro Popular enrollment on a household's three-month out-of-pocket health spending?*

We will work from the SPS rollout data, available in the `sps` tibble of the **[primer.data](https://github.com/PPBDS/primer.data)** package.

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
  message = "Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

> *The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* --- John W. Tukey

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The Preceptor Table only includes the variables mentioned in the question; in causal models there are at least two potential-outcome columns, plus the treatment column.

### Exercise 3

Look at how `t2_health_exp_3m` differs between the treatment and control groups. Ask AI to write R code that plots `t2_health_exp_3m` by `treatment` as a jitter plot, with `treatment` on the x-axis and the outcome on the y-axis (capped at 5000 pesos so the long tail doesn't crush the plot), plus a point showing the group mean. Paste the code below.

```{r wisdom-3}
question_text(NULL,
  message = "    x |>\n      filter(t2_health_exp_3m <= 5000) |>\n      ggplot(aes(x = treatment, y = t2_health_exp_3m)) +\n      geom_jitter(width = 0.2, alpha = 0.4) +\n      stat_summary(fun = mean, geom = \"point\", size = 4, color = \"steelblue\") +\n      scale_y_continuous(labels = scales::label_dollar(prefix = \"$\")) +\n      labs(\n        title = \"Out-of-Pocket Spending by Enrollment Status\",\n        subtitle = \"Blue dot = group mean\",\n        x = \"Treatment\",\n        y = \"3-month spending (pesos)\",\n        caption = \"Source: King et al. (2009) SPS rollout via primer.data\"\n      ) +\n      theme_minimal()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 12)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
x |>
  filter(t2_health_exp_3m <= 5000) |>
  ggplot(aes(x = treatment, y = t2_health_exp_3m)) +
  geom_jitter(width = 0.2, alpha = 0.4) +
  stat_summary(fun = mean, geom = "point", size = 4, color = "steelblue") +
  scale_y_continuous(labels = scales::label_dollar(prefix = "$")) +
  labs(
    title = "Out-of-Pocket Spending by Enrollment Status",
    subtitle = "Blue dot = group mean",
    x = "Treatment",
    y = "3-month spending (pesos)",
    caption = "Source: King et al. (2009) SPS rollout via primer.data"
  ) +
  theme_minimal()
```

### Exercise 4

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-4}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

This problem is causal, so one of the covariates is a treatment. In our problem, the treatment is `treatment` --- whether a household had enrolled in Seguro Popular by the follow-up survey. There is a potential outcome for each of the two possible values of the treatment.

### Exercise 5

What are the units for this problem?

```{r wisdom-5}
question_text(NULL,
  message = "Each row is one household in the universe of households eligible for Seguro Popular at the time of the rollout (roughly 2005--06). The original survey covered tens of thousands of households across 74 randomized communities; our analysis sample is a 500-household slice (250 enrolled, 250 not enrolled) drawn from that universe.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Specifying the unit forces a question precise enough to answer. We model households here, but a policy maker really cares about aggregate consequences across families.

### Exercise 6

What is the outcome variable for this problem?

```{r wisdom-6}
question_text(NULL,
  message = "Keep track of two outcome variables: the one in our Preceptor Table and the one in our data. In this case, both are `t2_health_exp_3m` --- the household's total out-of-pocket health spending over the three months prior to the follow-up survey, in Mexican pesos. The variable is bounded below by zero and has a long right tail (most households spend little; a few spend a great deal).",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The outcome we really care about --- *"did enrollment relieve financial strain on Mexican families?"* --- is fuzzier than what `t2_health_exp_3m` measures. The variable captures pesos out the door but not, say, foregone care or borrowing.

### Exercise 7

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r wisdom-7}
question_text(NULL,
  message = "Household income is a strong candidate. Lower-income households presumably benefit more in absolute terms from enrollment because they were spending a larger share of disposable income on health. We don't have direct income in our data, but `education` is a plausible proxy.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The term *covariates* is used in three ways: variables that *might* matter, variables we have data for, and variables we end up putting in the model. Most of the candidates a researcher lists never make it into the final model.

### Exercise 8

What is the treatment for this problem?

```{r wisdom-8}
question_text(NULL,
  message = "The treatment is `treatment`, which takes the values `Enrolled` and `Not enrolled`. The Ministry of Health manipulated treatment by random assignment of the rollout to communities; our `treatment` variable records, at follow-up, whether a household's community had been treated.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

A treatment is just another covariate which, for the purposes of this specific problem, we are assuming can be manipulated. Because the assignment was randomized, the *can* in *"can be manipulated"* is real here, not hypothetical.

### Exercise 9

What moment in time does the Preceptor Table refer to?

```{r wisdom-9}
question_text(NULL,
  message = "The follow-up survey window in 2005--06: roughly ten months after the rollout began. Our outcome (`t2_health_exp_3m`) refers to the three months before that follow-up moment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

A Preceptor Table can never really refer to an exact instant in time. The follow-up was a survey window, not a single moment.

### Exercise 10

Describe in words the Preceptor Table for this problem.

```{r wisdom-10}
question_text(NULL,
  message = "The Preceptor Table has four columns: one identifying each household, two potential-outcome columns (`Spending if Enrolled` and `Spending if Not Enrolled`), and one treatment column (`Treatment`). Each row represents one household at the time of the SPS follow-up survey. For each household, only one potential outcome is observable; the other --- the counterfactual --- is unknown, and the Preceptor Table marks it as such.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

```{r}
#| echo: false
#| results: asis
hatch <- function(x) {
  paste0(
    '<span style="background-image: repeating-linear-gradient(',
    '45deg, rgba(0,0,0,0.12) 0 1px, transparent 1px 6px);',
    ' padding: 6px 10px; margin: -6px -10px; display: block;">',
    x, '</span>'
  )
}

pre_tibble <- tibble::tribble(
  ~`Household`,           ~`Spending if Enrolled`, ~`Spending if Not Enrolled`, ~`Treatment`,
  "Hidalgo (#102)",       "$420",                  "$680",                      "Enrolled",
  "Ramirez (#247)",       "$0",                    "$120",                      "Not enrolled",
  "...",                  "...",                   "...",                       "...",
  "Castaneda (#491)",     "$1,800",                "$1,500",                    "Enrolled"
)

pre_title_footnote     <- "If all the information in this table were available, we could answer the question: What is the average causal effect of Seguro Popular enrollment on a household's three-month out-of-pocket health spending?"
pre_units_footnote     <- "Each row is one household at the time of the 2005--06 SPS follow-up survey. The original study surveyed tens of thousands of households across 74 randomized communities."
pre_outcome_footnote   <- "Out-of-pocket health spending in pesos, summed over the three months before the follow-up survey. The two columns hold each household's potential outcomes under the two values of treatment. The cross-hatched cell in each row marks the unobservable potential outcome --- the spending the household would have had under the treatment it did not actually receive."
pre_treatment_footnote <- "The treatment is `Enrolled` if the household's community had received Seguro Popular by the follow-up survey, `Not enrolled` otherwise. The randomization was at the community level, not the household level."
pre_effect_footnote    <- "Ramirez's causal effect is $0 - $120 = -$120: enrollment reduced this household's three-month spending by 120 pesos. The Preceptor Table shows the unobservable truth, so this is the actual causal effect for this household, not an estimate."

pre_gt_html <- gt::gt(pre_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit", id = "unit_span",
                  columns = c(`Household`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",
                  columns = c(`Spending if Enrolled`, `Spending if Not Enrolled`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::cols_align(align = "left",  columns = c(`Household`, `Treatment`)) |>
  gt::cols_align(align = "right", columns = c(`Spending if Enrolled`, `Spending if Not Enrolled`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span"))
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `Spending if Not Enrolled`, rows = c(1, 4)),
    fn = function(x) hatch(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `Spending if Enrolled`, rows = 2),
    fn = function(x) hatch(x)
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_treatment_footnote,
                   locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = pre_effect_footnote,
                   locations = gt::cells_body(columns = `Household`, rows = 2)) |>
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 11

Take a moment to confirm the structure of our analysis sample. At the R prompt, run:

```
x |> count(treatment)
```

CP/CR.

```{r wisdom-11}
question_text(NULL,
  message = "    > x |> count(treatment)\n    # A tibble: 2 × 2\n      treatment        n\n      <fct>        <int>\n    1 Not enrolled   250\n    2 Enrolled       250",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The sample is balanced 250 / 250 across the two arms because we drew it that way: stratifying on treatment keeps the two groups visible and gives both their own clean estimate. The original SPS data has many more households than 500 in either arm; we sampled down to keep this tutorial fast.

### Exercise 12

Write the first two sentences of the summary paragraph. The first introduces the topic; the second introduces the data and the specific question.

```{r wisdom-12}
question_text(NULL,
  message = "Households' out-of-pocket spending on health care is a key marker of financial vulnerability. Using a 500-household sample drawn from Mexico's 2005--06 randomized rollout of Seguro Popular, we estimate the average causal effect of enrollment on three-month out-of-pocket spending.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Read our answer. It will not be the same as yours. You can change your answer to incorporate some of our ideas, but do not copy/paste exactly. Add your two sentences to the summary paragraph in your QMD. `Cmd/Ctrl + Shift + K`, then commit/push.


## Justice
###

Wisdom gives us the Preceptor Table and the data.

```{r}
#| echo: false
#| results: asis
pre_tibble_j <- tibble::tribble(
  ~`Household`,           ~`Spending if Enrolled`, ~`Spending if Not Enrolled`, ~`Treatment`,
  "Hidalgo (#102)",       "$420",                  "$680",                      "Enrolled",
  "Ramirez (#247)",       "$0",                    "$120",                      "Not enrolled",
  "...",                  "...",                   "...",                       "...",
  "Castaneda (#491)",     "$1,800",                "$1,500",                    "Enrolled"
)

pre_gt_html_j <- gt::gt(pre_tibble_j, id = "preceptor_tbl_justice") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit", id = "unit_span",
                  columns = c(`Household`)) |>
  gt::tab_spanner(label = "Potential Outcomes", id = "outcome_span",
                  columns = c(`Spending if Enrolled`, `Spending if Not Enrolled`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::cols_align(align = "left",  columns = c(`Household`, `Treatment`)) |>
  gt::cols_align(align = "right", columns = c(`Spending if Enrolled`, `Spending if Not Enrolled`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span"))
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `Spending if Not Enrolled`, rows = c(1, 4)),
    fn = function(x) hatch(x)
  ) |>
  gt::text_transform(
    locations = gt::cells_body(columns = `Spending if Enrolled`, rows = 2),
    fn = function(x) hatch(x)
  ) |>
  gt::opt_css(
    css = "
      #preceptor_tbl_justice .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html_j,
  "</div>\n",
  "```\n",
  sep = ""
)
```

```{r}
#| echo: false
#| results: asis
data_tibble <- tibble::tribble(
  ~`Household`,           ~`Year`, ~`3-month Spending`, ~`Treatment`, ~`Age`, ~`Sex`,
  "Hidalgo (#102)",       "2005",  "$420",              "Enrolled",   "44",   "Female",
  "Ramirez (#247)",       "2005",  "$120",              "Not enrolled","51",  "Male",
  "...",                  "...",   "...",               "...",        "...",  "...",
  "Castaneda (#491)",     "2005",  "$1,800",            "Enrolled",   "32",   "Female"
)

data_title_footnote     <- "Mexican Ministry of Health Seguro Popular evaluation, 2005--06 (King et al. 2009). Roughly 30,000 surveyed households across 74 randomized communities. Our analysis filters to households with complete records on treatment, follow-up spending, and the demographic covariates, then samples 250 enrolled and 250 not-enrolled households for a balanced 500-row analysis."
data_units_footnote     <- "Each row is one household at the time of the 2005--06 follow-up survey. The example rows above use real names and ID numbers chosen to span the spending range; missing rows represent the other 496 households not shown."
data_outcome_footnote   <- "Three-month total out-of-pocket health spending, in Mexican pesos, recorded at follow-up. Self-reported and bounded below by zero, with a long right tail."
data_treatment_footnote <- "Whether the household's community had received Seguro Popular by the follow-up survey, recorded as `Enrolled` or `Not enrolled`. Random assignment was at the community level."
data_covariate_footnote <- "Household head's age in years and self-reported sex."

data_gt_html <- gt::gt(data_tibble, id = "data_tbl") |>
  gt::tab_header(title = "Data: Mexican SPS Rollout, 2005--06") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`Household`, `Year`)) |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span",
                  columns = c(`3-month Spending`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Age`, `Sex`)) |>
  gt::cols_align(align = "left",  columns = c(`Household`, `Treatment`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `3-month Spending`, `Age`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = data_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = data_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = data_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = data_treatment_footnote,
                   locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = data_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #data_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  data_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 1

In your own words, describe the components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
  message = "Justice reviews the Population Table and selects the formula for the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Justice is about concerns that you (or your critics) might have, reasons why the model you create might not work as well as you hope. The exercises that follow walk through several such concerns by name. Justice also picks the formula for the data generating mechanism --- different outcome variables call for different formulas, a piece that gets more technical names in later tutorials.

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Validity is always about the columns. Just because two columns share a name does not mean they are the same thing.

### Exercise 3

Provide one reason why the assumption of validity might not hold for the outcome variable `t2_health_exp_3m` or for the treatment.

```{r justice-3}
question_text(NULL,
  message = "The `t2_health_exp_3m` column in the data is a *self-reported* spending total over the prior three months, with all the recall error that implies. The `Spending if [treatment]` columns in the Preceptor Table are the actual amounts the household paid; the two columns measure the same concept only if recall is unbiased on average. Households facing financial strain may underreport (to seem to be coping) or overreport (to express grievance), so the data column may not match the Preceptor Table column it is meant to stand in for.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Validity, if true, allows us to construct the Population Table. Self-reported spending is always a noisy proxy for actual spending; the question is whether the noise is unbiased.

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The Population Table is almost always much bigger than the data and the Preceptor Table combined.

### Exercise 5

Specify the unit/time combinations which define each row in this Population Table.

```{r justice-5}
question_text(NULL,
  message = "Each row is one household at one point in time. Data rows are the 500 households in our sampled SPS analysis subset, surveyed in 2005--06. Preceptor rows are the same population of Mexican households in 2005--06 (the universe the original survey sampled from); because both blocks share era and target population, our example rows happen to be the same three households shown twice --- once with the observed potential outcome and once with both potential outcomes filled in.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

```{r}
#| echo: false
#| results: asis
pop_tibble <- tibble::tribble(
  ~Source    , ~`Household`,            ~`Year`, ~`3-month Spending`,    ~`Treatment`,    ~`Age`, ~`Sex`,
  "..."      , "...",                   "..."  , "...",                  "...",           "...",  "...",
  "Data"     , "Hidalgo (#102)",        "2005" , "$420",                 "Enrolled",      "44",   "Female",
  "Data"     , "Ramirez (#247)",        "2005" , "$120",                 "Not enrolled",  "51",   "Male",
  "Data"     , "...",                   "..."  , "...",                  "...",           "...",  "...",
  "Data"     , "Castaneda (#491)",      "2005" , "$1,800",               "Enrolled",      "32",   "Female",
  "..."      , "...",                   "..."  , "...",                  "...",           "...",  "...",
  "Preceptor", "Hidalgo (#102)",        "2005" , "$420 / $680",          "Enrolled",      "44",   "Female",
  "Preceptor", "Ramirez (#247)",        "2005" , "$0 / $120",            "Not enrolled",  "51",   "Male",
  "Preceptor", "...",                   "..."  , "...",                  "...",           "...",  "...",
  "Preceptor", "Castaneda (#491)",      "2005" , "$1,800 / $1,500",      "Enrolled",      "32",   "Female",
  "..."      , "...",                   "..."  , "...",                  "...",           "...",  "..."
)

pop_title_footnote     <- "Population: Mexican households eligible for Seguro Popular at the time of the randomized rollout (2005--06). Data rows are the 500 households in our sampled analysis subset; Preceptor rows are the same households described with both potential outcomes filled in (counterfactual hatched in the Preceptor Table itself). The example rows are real households shown to illustrate the structure --- the same three units appear in both blocks."
pop_units_footnote     <- "Each row is one household at one point in time. The Data block records what we observed (one potential outcome per household, given the household's actual treatment); the Preceptor block records the unobservable truth (both potential outcomes filled in)."
pop_outcome_footnote   <- "Three-month spending in pesos. In Data rows, only the value under the household's actual treatment is observed --- the counterfactual is unknown. In Preceptor rows, both potential outcomes are filled in (here written as `enrolled / not enrolled`); the counterfactual was hatched in the Preceptor Table proper."
pop_treatment_footnote <- "Whether the household's community had received Seguro Popular by the follow-up survey. Same column meaning in both blocks; randomization at the community level is what makes the Data and Preceptor blocks compatible."
pop_covariate_footnote <- "Household head's age and self-reported sex; same in both blocks because both refer to the same households."

pop_gt_html <- gt::gt(pop_tibble, id = "population_tbl") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`Household`, `Year`)) |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span",
                  columns = c(`3-month Spending`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Age`, `Sex`)) |>
  gt::cols_align(align = "left",  columns = c(`Source`, `Household`, `Treatment`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `3-month Spending`, `Age`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_treatment_footnote,
                   locations = gt::cells_column_spanners(spanners = "treatment_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 6

In your own words, define the assumption of "stability."

```{r justice-6}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

The most natural way for stability to fail is *time*: data from one era used to answer a question about another era. Here our Data and Preceptor rows share the 2005--06 SPS follow-up window, so the time gap is essentially zero. The next exercise sharpens what *"the relationship"* really means.

### Exercise 7

Provide one reason why stability might not hold in this case.

```{r justice-7}
question_text(NULL,
  message = "The communities randomized into the rollout were drawn from the country's poorest and least-served regions. The causal effect of enrollment in those communities --- where uninsured spending was already a real burden --- may be different (probably larger) than the causal effect would be in wealthier urban communities. If the *parameter* of interest is the average causal effect for the whole population, our 500-household sample may estimate something other than that, because the slope of spending on treatment may differ across regions.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A change in the *distribution* of any single variable does not, by itself, violate stability. The mean spending could differ between regions; the mix of household sizes could differ; the age distribution of household heads could differ. None of those would violate stability on their own. What violates stability is a change in the *parameter* connecting treatment to spending --- the slope, the causal-effect coefficient. Distribution shifts are everywhere; parameter shifts are what hurt us.

### Exercise 8

We use our data to make inferences about the overall population. We use information about the population to make inferences about the Preceptor Table: Data → Population → Preceptor Table. In your own words, define the assumption of "representativeness."

```{r justice-8}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Ideally we would like both the Preceptor Table and the data to be random samples from the population. Here the Ministry randomized treatment assignment but not survey participation: which households actually answered the follow-up survey is a different selection process from which communities were treated.

### Exercise 9

Provide one reason, involving the relationship between the data and the population, why representativeness might not hold.

```{r justice-9}
question_text(NULL,
  message = "Households that answered both the baseline and follow-up surveys are not a random sample of all eligible households. Households that moved, dissolved, or refused the follow-up are systematically absent --- and their absence is plausibly correlated with health-spending pressure (a household that fell into financial crisis between waves may be more likely to drop out). Our 500-row sample inherits whatever attrition pattern was in the original analysis dataset, and may underrepresent households for whom the program would have mattered most.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A non-representative sample doesn't *guarantee* a biased estimate --- the bias might wash out by chance. But "by chance" is the only mechanism left to defend the estimate, and we have no principled reason to expect it to work in our favor. Representativeness also has a *second* link, between the population and the Preceptor Table, which later tutorials will exercise; for this tutorial the data-to-population link is the focus.

### Exercise 10

In your own words, define the assumption of "unconfoundedness."

```{r justice-10}
question_text(NULL,
  message = "Unconfoundedness means that the treatment assignment is independent of the potential outcomes, when we condition on pre-treatment covariates.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

This assumption is only relevant for causal models. We describe a model as *confounded* if it is not true. The easiest way to ensure unconfoundedness is to assign treatment randomly --- which the Mexican Ministry of Health actually did here, rolling out Seguro Popular into communities chosen at random.

### Exercise 11

Provide one reason why the assumption of unconfoundedness might not be true (or might be weaker than it looks) in this case.

```{r justice-11}
question_text(NULL,
  message = "Randomization in the SPS rollout was at the *community* level, not the household level. Within a treated community, every household was offered enrollment, but actual sign-up was voluntary. Households that signed up may differ from households in the same community that did not --- in unobserved ways like trust in government, prior insurance experience, or chronic-illness burden. So while the *intent-to-treat* comparison (any household in a treated community vs. any in a control community) is unconfounded by design, the comparison many policy analysts actually want (enrolled households vs. eligible-but-not-enrolled) involves a self-selected step that randomization does not protect.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

The great advantage of randomization is that it guarantees unconfoundedness *for the assignment that was randomized*. When the variable we ultimately care about (actual enrollment) sits one step downstream of the randomized variable (community assignment), randomization gets us most of the way --- but not all of it.

### Exercise 12

A statistical model consists of two parts: the probability family and the link function. Add `library(tidymodels)` to the QMD file. Place your cursor on that line. Use `Cmd/Ctrl + Enter`.

CP/CR.

```{r justice-12}
question_text(NULL,
  message = "    > library(tidymodels)\n    ── Attaching packages ──────── tidymodels 1.x.x ──\n    ✔ broom        ✔ recipes\n    ✔ dials        ✔ rsample\n    ✔ infer        ✔ tune\n    ✔ modeldata    ✔ workflows\n    ✔ parsnip      ✔ workflowsets\n    ✔ purrr        ✔ yardstick\n    ── Conflicts ──── tidymodels_conflicts() ──\n    ✖ purrr::discard() masks scales::discard()\n    ✖ dplyr::filter()  masks stats::filter()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Our outcome variable `t2_health_exp_3m` is a continuous quantity (pesos). The probability family is Normal: $Y \sim N(\mu, \sigma^2)$.

### Exercise 13

Add `library(broom)` to the QMD file. Place your cursor on that line. Use `Cmd/Ctrl + Enter`.

CP/CR.

```{r justice-13}
question_text(NULL,
  message = "    > library(broom)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The link function for a continuous outcome is the identity (linear): $\mu = \beta_0 + \beta_1 X_1 + \cdots + \beta_n X_n$. Our model will be a linear regression.

### Exercise 14

Write one sentence highlighting a potential weakness in your model. So far our summary paragraph looks like this:

> Households' out-of-pocket spending on health care is a key marker of financial vulnerability. Using a 500-household sample drawn from Mexico's 2005--06 randomized rollout of Seguro Popular, we estimate the average causal effect of enrollment on three-month out-of-pocket spending.

Your version will be somewhat different.

```{r justice-14}
question_text(NULL,
  message = "One concern is that randomization happened at the community level, not the household level: within a treated community, enrollment was voluntary, so the comparison of enrolled vs. not-enrolled households is partly a comparison of self-selected groups, which our model cannot fully untangle.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a weakness sentence to your summary paragraph in your QMD. Don't copy/paste our answer exactly. `Cmd/Ctrl + Shift + K`, then commit/push.


## Courage
###

Justice gives us the Population Table and the abstract data generating mechanism.

```{r}
#| echo: false
#| results: asis
pop_tibble_c <- tibble::tribble(
  ~Source    , ~`Household`,            ~`Year`, ~`3-month Spending`,    ~`Treatment`,    ~`Age`, ~`Sex`,
  "..."      , "...",                   "..."  , "...",                  "...",           "...",  "...",
  "Data"     , "Hidalgo (#102)",        "2005" , "$420",                 "Enrolled",      "44",   "Female",
  "Data"     , "Ramirez (#247)",        "2005" , "$120",                 "Not enrolled",  "51",   "Male",
  "Data"     , "...",                   "..."  , "...",                  "...",           "...",  "...",
  "Data"     , "Castaneda (#491)",      "2005" , "$1,800",               "Enrolled",      "32",   "Female",
  "..."      , "...",                   "..."  , "...",                  "...",           "...",  "...",
  "Preceptor", "Hidalgo (#102)",        "2005" , "$420 / $680",          "Enrolled",      "44",   "Female",
  "Preceptor", "Ramirez (#247)",        "2005" , "$0 / $120",            "Not enrolled",  "51",   "Male",
  "Preceptor", "...",                   "..."  , "...",                  "...",           "...",  "...",
  "Preceptor", "Castaneda (#491)",      "2005" , "$1,800 / $1,500",      "Enrolled",      "32",   "Female",
  "..."      , "...",                   "..."  , "...",                  "...",           "...",  "..."
)

pop_gt_html_c <- gt::gt(pop_tibble_c, id = "population_tbl_courage") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`Household`, `Year`)) |>
  gt::tab_spanner(label = "Outcome", id = "outcome_span",
                  columns = c(`3-month Spending`)) |>
  gt::tab_spanner(label = "Treatment", id = "treatment_span",
                  columns = c(`Treatment`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Age`, `Sex`)) |>
  gt::cols_align(align = "left",  columns = c(`Source`, `Household`, `Treatment`, `Sex`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `3-month Spending`, `Age`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "treatment_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::opt_css(
    css = "
      #population_tbl_courage .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html()

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html_c,
  "</div>\n",
  "```\n",
  sep = ""
)
```

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n + \epsilon$$

with $\epsilon \sim N(0, \sigma^2)$.

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
  message = "Courage creates the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

Justice fixed the structural choices --- continuous outcome, linear regression, identity link. Courage now picks specific covariates and estimates the parameters.

### Exercise 2

Because our outcome variable is continuous, start to create the model by entering `linear_reg(engine = "lm")`.

```{r courage-2, exercise = TRUE}

```

```{r courage-2-hint-1, eval = FALSE}
linear_reg(engine = ...)
```

```{r courage-2-test, include = FALSE}
linear_reg(engine = "lm")
```

###

**[tidymodels](https://www.tidymodels.org/)** is the most popular framework for estimating models among R users.

### Exercise 3

Continue the pipe to `fit(t2_health_exp_3m ~ treatment, data = x)`. This is the central causal model: outcome on the binary, randomized `treatment`.

```{r courage-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-3-hint-1, eval = FALSE}
... |>
  fit(t2_health_exp_3m ~ treatment, data = x)
```

```{r courage-3-test, include = FALSE}
linear_reg(engine = "lm") |>
  fit(t2_health_exp_3m ~ treatment, data = x)
```

###

The character variable `treatment` becomes a 0/1 dummy called `treatmentEnrolled`: 1 for households in treated communities, 0 for control. The intercept becomes the expected outcome for the reference level (`Not enrolled`).

### Exercise 4

Here is the parameter table for this first model:

```{r}
#| echo: false
linear_reg() |>
  set_engine("lm") |>
  fit(t2_health_exp_3m ~ treatment, data = x) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 0)))
```

Write a sentence interpreting the `treatmentEnrolled` estimate of -62.

```{r courage-4}
question_text(NULL,
  message = "Our point estimate of the average causal effect of enrollment on three-month out-of-pocket health spending is -62 pesos: enrolled households spent, on average, about 62 pesos less over the three months before the follow-up survey than they would have spent without enrollment. The 95% confidence interval ranges from -376 to +253 pesos --- it includes zero, so we cannot rule out the possibility that the true effect is zero (or even slightly positive).",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

Causal language is appropriate here because community-level treatment was randomized. We say *"the average causal effect"* and *"would have spent"* --- the Rubin Causal Model vocabulary --- because the identification strategy (randomization) earns us the causal reading. The CI's inclusion of zero is itself important: a real-world intervention need not move the average, and the most honest report of this estimate acknowledges that we have not nailed down the effect's sign.

### Exercise 5

Now fit a second candidate that adjusts for `age`: `fit(t2_health_exp_3m ~ treatment + age, data = x)`. This adds a continuous covariate to the causal model.

```{r courage-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r courage-5-hint-1, eval = FALSE}
... |>
  fit(t2_health_exp_3m ~ treatment + age, data = x)
```

```{r courage-5-test, include = FALSE}
linear_reg(engine = "lm") |>
  fit(t2_health_exp_3m ~ treatment + age, data = x)
```

###

In an observational study, adjusting for a covariate can move the treatment estimate substantially --- that's what we do when we suspect confounding. Randomization is supposed to balance covariates between Treated and Control groups, so adjustment should not move the estimate much. Let's check.

### Exercise 6

Here is the parameter table for the adjusted model:

```{r}
#| echo: false
linear_reg() |>
  set_engine("lm") |>
  fit(t2_health_exp_3m ~ treatment + age, data = x) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 0)))
```

Write a sentence interpreting the `age` slope (6) and the new `treatmentEnrolled` estimate (-48).

```{r courage-6}
question_text(NULL,
  message = "The slope on `age` is about 6 pesos per year, with a 95% confidence interval that crosses zero (-4 to +15): when we compare two groups of households in the same treatment condition but differing by one year in head-of-household age, their expected three-month spending is essentially indistinguishable. The treatmentEnrolled estimate barely moves (-62 to -48) when age is added --- which is what we expect from a randomized experiment, where treatment assignment is by design unrelated to age.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 7)
```

###

Notice the language for `age`: *"when we compare two groups of households... in the same treatment condition but differing by one year of age."* `age` is *not* a treatment in this study; it's a covariate. So we describe a comparison, not a causal effect of aging on spending. Causal language is reserved for the variable whose assignment was randomized. The fact that the `treatmentEnrolled` estimate barely moves under adjustment is itself evidence that randomization worked --- in an observational study, adjusting for a real confounder would visibly shift the treatment estimate. Our final model is the simpler `t2_health_exp_3m ~ treatment` (Exercise 3), since age adds essentially nothing.

### Exercise 7

Behind the scenes of this tutorial, an object called `fit_sps` has been created which is the result of the final model. Type `fit_sps` and hit "Run Code."

```{r courage-7, exercise = TRUE}

```

```{r courage-7-hint-1, eval = FALSE}
fit_sps
```

```{r courage-7-test, include = FALSE}
fit_sps
```

###

In data science, **we deal with words, math, and code, but the most important of these is code.**

### Exercise 8

We need `fit_sps` to exist in R World. Copy/paste this code to the R prompt and execute it:

```
fit_sps <- linear_reg(engine = "lm") |>
  fit(t2_health_exp_3m ~ treatment, data = x)
```

CP/CR.

```{r courage-8}
question_text(NULL,
  message = "    > fit_sps <- linear_reg(engine = \"lm\") |>\n    +   fit(t2_health_exp_3m ~ treatment, data = x)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Just because something exists in the tutorial does not mean it exists at your R prompt. Keep both worlds in sync.

### Exercise 9

One way to check whether our model is reasonable is to compare the distribution of the actual outcome with the distribution of fitted values. If the model is well-specified, the two should look similar.

```{r}
#| echo: false
#| message: false
#| warning: false
fitted_df <- tibble(
  value  = c(x$t2_health_exp_3m, predict(extract_fit_engine(fit_sps))),
  source = c(rep("Actual", nrow(x)),
             rep("Fitted", nrow(x)))
)

ggplot(fitted_df, aes(x = value, fill = source)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(labels = scales::label_dollar(prefix = "$"), limits = c(0, 5000)) +
  labs(
    title = "Actual vs. Fitted Spending",
    subtitle = "The fitted distribution is much narrower --- our model has only two predicted values",
    x = "3-month spending (pesos)",
    y = NULL,
    fill = "Source",
    caption = "Source: King et al. (2009) SPS rollout via primer.data"
  ) +
  theme_minimal() +
  theme(axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())
```

### Exercise 10

With the parameters estimated, we can write the fitted model as a concrete formula:

$$\widehat{\text{Spending}} = 783 - 62 \cdot \text{treatmentEnrolled}$$

`treatmentEnrolled` is a dummy variable: 1 for enrolled households, 0 for control. The intercept (783) is the expected three-month spending for a control household; the coefficient (-62) is the average causal effect of enrollment.

**This is our data generating mechanism.**

###

A data generating mechanism is just a formula, something we can write down and implement with computer code. Of course there is randomness baked into the DGM, but we won't worry about that detail here.

### Exercise 11

Create a new code chunk in your QMD with `#| cache: true`. Copy/paste the code for estimating the model:

```
fit_sps <- linear_reg(engine = "lm") |>
  fit(t2_health_exp_3m ~ treatment, data = x)
```

`Cmd/Ctrl + Shift + K`. At the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r courage-11}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    #| cache: true\n    fit_sps <- linear_reg(engine = \"lm\") |>\n      fit(t2_health_exp_3m ~ treatment, data = x)\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

`#| cache: true` tells Quarto to reuse the saved fit on subsequent renders. Re-render with `Cmd/Ctrl + Shift + K` --- it should be quick the second time.

### Exercise 12

Add `*_cache` to your `.gitignore`. Cached objects don't belong on GitHub.

At the R prompt, run:

```
show_file(".gitignore")
```

CP/CR.

```{r courage-12}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files\n    *_cache",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Commit and push.

### Exercise 13

At the R prompt, run `tidy()` on `fit_sps` with `conf.int = TRUE`.

```{r courage-13, exercise = TRUE}

```

```{r courage-13-hint-1, eval = FALSE}
tidy(fit_sps, conf.int = TRUE)
```

```{r courage-13-test, include = FALSE}
tidy(fit_sps, conf.int = TRUE)
```

###

`tidy()` is part of **[broom](https://broom.tidymodels.org/)**, used to summarize information from a wide variety of models.

### Exercise 14

Create a new code chunk in your QMD. Ask AI to make a nice-looking table from the tibble returned by `tidy()`. Insert your table code, `Cmd/Ctrl + Shift + K`, and at the R prompt run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r courage-14}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    tidy(fit_sps, conf.int = TRUE) |>\n      select(term, estimate, conf.low, conf.high) |>\n      knitr::kable(\n        caption = \"Estimates and 95% CIs.  Source: King et al. (2009) SPS rollout via primer.data\",\n        digits = 0\n      )\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

A nice table includes a title, a caption with the data source, and reasonable rounding. The more you use AI for these housekeeping tasks, the better you'll get at telling it what you want.

### Exercise 15

Add a sentence to your project summary explaining the model structure.

```{r courage-15}
question_text(NULL,
  message = "We model three-month out-of-pocket spending as a normally distributed variable which is a linear function of treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add your sentence to the summary paragraph. `Cmd/Ctrl + Shift + K`, then commit/push.


## Temperance
###

Courage provides the data generating mechanism. We can express the DGM in four ways:

**In words:** We model three-month out-of-pocket spending as a normally distributed variable which is a linear function of treatment.

**In R code:**

```r
fit_sps <- linear_reg(engine = "lm") |>
  fit(t2_health_exp_3m ~ treatment, data = x)
```

**In a parameter table:**

```{r}
#| echo: false
fit_sps |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 0)))
```

**As a mathematical formula:**

$$\text{Spending} = 783 - 62 \cdot \text{treatmentEnrolled} + \epsilon$$

with $\epsilon \sim N(0, 3{,}206{,}129)$ (a residual standard deviation of about 1,791 pesos --- spending varies a lot from household to household).

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
  message = "Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Courage gave us the DGM, including its parameter values. Temperance puts those parameters to work. The tool is the **[marginaleffects](https://marginaleffects.com/)** package, with companion book *[Model to Meaning](https://marginaleffects.com/)* by Vincent Arel-Bundock.

### Exercise 2

In the end, we don't really care about parameters, much less how to interpret them. Parameters are *imaginary*, like unicorns.

Add `library(marginaleffects)` to the QMD file. Place your cursor on that line. Use `Cmd/Ctrl + Enter`.

CP/CR.

```{r temperance-2}
question_text(NULL,
  message = "    > library(marginaleffects)\n    >",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

We should be modest in the claims we make. The posteriors we create are never the truth.

### Exercise 3

What is the specific question we are trying to answer?

```{r temperance-3}
question_text(NULL,
  message = "What is the average causal effect of Seguro Popular enrollment on a household's three-month out-of-pocket health spending?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

To make a decision wisely, we need good estimates of many unknown numbers. Drilling down to one specific question is how we make progress.

### Exercise 4

At the R prompt, run `predictions()` on `fit_sps`. CP/CR.

```{r temperance-4}
question_text(NULL,
  message = "    > predictions(fit_sps)\n    \n     Estimate Std. Error    z Pr(>|z|)  2.5 % 97.5 %\n          721        113 6.36   <0.001    499    944\n          783        113 6.91   <0.001    560   1005\n          721        113 6.36   <0.001    499    944\n          ... (500 rows total, one per household)\n    \n    Type: numeric",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

`predictions()` returns one row per observation (500 here) with the model's predicted spending for each household. Because our model has only one covariate, there are only two distinct fitted values: 721 pesos for the enrolled rows, 783 pesos for the not-enrolled rows.

### Exercise 5

At the R prompt, run `plot_predictions(fit_sps, condition = "treatment")`.

CP/CR.

```{r temperance-5}
question_text(NULL,
  message = "    > plot_predictions(fit_sps, condition = \"treatment\")\n\n(A plot opens with two point estimates --- one for `Not enrolled` near 780 pesos and one for `Enrolled` near 720 pesos --- each with a 95% confidence interval. The intervals overlap heavily, reflecting that we cannot tell the two means apart with this much data.)",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

```{r temperance-5-test, include = FALSE}
plot_predictions(fit_sps, condition = "treatment")
```

###

The two intervals overlap considerably, which is the visible form of our parameter-table conclusion: the difference in expected spending between the two arms is small relative to the noise.

### Exercise 6

At the R prompt, run the same call with `draw = FALSE`.

CP/CR.

```{r temperance-6}
question_text(NULL,
  message = "    > plot_predictions(fit_sps, condition = \"treatment\", draw = FALSE)\n      rowid estimate std.error statistic p.value conf.low conf.high   treatment\n    1     1      783     113     6.91   <0.001    560      1005 Not enrolled\n    2     2      721     113     6.36   <0.001    499       944     Enrolled",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

```{r temperance-6-test, include = FALSE}
plot_predictions(fit_sps, condition = "treatment", draw = FALSE)
```

###

`plot_predictions()` returns a ggplot object, so you can keep adding ggplot layers like `labs()`. But it's often easier to start from the underlying tibble (`draw = FALSE`) and build the plot from scratch.

### Exercise 7

Work with AI to create a beautiful plot starting from the output of `plot_predictions(..., draw = FALSE)`. Do this in your QMD. Title: key variables. Subtitle: takeaway. Caption: data source. Axis labels: human-readable.

Paste the plot code below.

```{r temperance-7}
question_text(NULL,
  message = "    plot_predictions(fit_sps, condition = \"treatment\", draw = FALSE) |>\n      ggplot(aes(x = treatment, y = estimate)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high),\n                      color = \"grey20\", size = 1.1) +\n      scale_y_continuous(labels = scales::label_dollar(prefix = \"$\")) +\n      labs(\n        title = \"Predicted 3-Month Out-of-Pocket Spending by Enrollment Status\",\n        subtitle = \"Enrolled households spend ~60 pesos less, but the CI crosses zero\",\n        x = \"Treatment\",\n        y = \"Predicted 3-month spending\",\n        caption = \"Source: King et al. (2009) SPS rollout via primer.data\"\n      ) +\n      theme_minimal()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 14)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(fit_sps, condition = "treatment", draw = FALSE) |>
  ggplot(aes(x = treatment, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                  color = "grey20", size = 1.1) +
  scale_y_continuous(labels = scales::label_dollar(prefix = "$")) +
  labs(
    title = "Predicted 3-Month Out-of-Pocket Spending by Enrollment Status",
    subtitle = "Enrolled households spend ~60 pesos less, but the CI crosses zero",
    x = "Treatment",
    y = "Predicted 3-month spending",
    caption = "Source: King et al. (2009) SPS rollout via primer.data"
  ) +
  theme_minimal()
```

### Exercise 8

Finalize the new graphics chunk in your QMD. `Cmd/Ctrl + Shift + K`. At the R prompt, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r temperance-8}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    plot_predictions(fit_sps, condition = \"treatment\", draw = FALSE) |>\n      ggplot(aes(x = treatment, y = estimate)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +\n      scale_y_continuous(labels = scales::label_dollar(prefix = \"$\")) +\n      labs(\n        title = \"Predicted 3-Month Out-of-Pocket Spending by Enrollment Status\",\n        subtitle = \"Enrolled households spend ~60 pesos less, but the CI crosses zero\",\n        x = \"Treatment\", y = \"Predicted 3-month spending\",\n        caption = \"Source: King et al. (2009) SPS rollout via primer.data\"\n      ) +\n      theme_minimal()\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 12)
```

###

The map is not the territory. A clean graphic tells a story, but the story always rests on assumptions that are at best approximately true.

### Exercise 9

Write the last sentence of your summary paragraph. It describes at least one Quantity of Interest and a measure of uncertainty.

```{r temperance-9}
question_text(NULL,
  message = "We estimate that enrollment in Seguro Popular reduced three-month out-of-pocket spending by about 60 pesos on average, with a 95% confidence interval that runs from a 380-peso reduction to a 250-peso increase --- a range wide enough that we cannot rule out no effect.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a final sentence to your summary paragraph, but don't copy/paste exactly. `Cmd/Ctrl + Shift + K`.

### Exercise 10

Write a few sentences explaining why the estimates might be wrong. Suggest alternative estimates if warranted.

```{r temperance-10}
question_text(NULL,
  message = "Our 500-household sample is balanced 250/250 across treatment arms by construction, but it is drawn from a survey with substantial follow-up attrition and from communities chosen by the Ministry from the country's poorer regions. Both pull our estimate away from a national average. The randomization happened at the community, not household, level --- so the comparison of *enrolled* vs. *not enrolled* households conflates the randomized intent-to-treat effect with self-selection into enrollment within treated communities. The point estimate could plausibly be anywhere from -200 to +50 pesos rather than the -62 we report; we should widen the confidence interval substantially to account for these issues that the model's standard error does not see.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

Always go back to your Preceptor Table. Even a perfectly estimated statistical model is rarely as useful as we'd like.

### Exercise 11

Rearrange the material in your QMD so the order is graphic, then summary paragraph. The chunk that creates `fit_sps` must occur before the chunk that creates the graphic. `Cmd/Ctrl + Shift + K`. At the R prompt, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r temperance-11}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\")\n    ---\n    title: \"Seguro Popular\"\n    author: \"<your name>\"\n    execute:\n      echo: false\n    ---\n    \n    ```{r}\n    #| message: false\n    library(tidyverse)\n    library(tidymodels)\n    library(primer.data)\n    library(broom)\n    library(marginaleffects)\n    ```\n    \n    ```{r}\n    set.seed(2026)\n    x <- sps |>\n      select(treatment, t2_health_exp_3m, age, sex, education) |>\n      drop_na() |>\n      mutate(treatment = factor(if_else(treatment == 1, \"Enrolled\", \"Not enrolled\"),\n                                levels = c(\"Not enrolled\", \"Enrolled\"))) |>\n      group_by(treatment) |>\n      slice_sample(n = 250) |>\n      ungroup()\n    ```\n    \n    ```{r}\n    #| cache: true\n    fit_sps <- linear_reg(engine = \"lm\") |>\n      fit(t2_health_exp_3m ~ treatment, data = x)\n    ```\n    \n    ```{r}\n    plot_predictions(fit_sps, condition = \"treatment\", draw = FALSE) |>\n      ggplot(aes(x = treatment, y = estimate)) +\n      geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +\n      scale_y_continuous(labels = scales::label_dollar(prefix = \"$\")) +\n      labs(\n        title = \"Predicted 3-Month Out-of-Pocket Spending by Enrollment Status\",\n        subtitle = \"Enrolled households spend ~60 pesos less, but the CI crosses zero\",\n        x = \"Treatment\", y = \"Predicted 3-month spending\",\n        caption = \"Source: King et al. (2009) SPS rollout via primer.data\"\n      ) +\n      theme_minimal()\n    ```\n    \n    Households' out-of-pocket spending on health care is a key marker of financial vulnerability. Using a 500-household sample drawn from Mexico's 2005--06 randomized rollout of Seguro Popular, we estimate the average causal effect of enrollment on three-month out-of-pocket spending. One concern is that randomization happened at the community level, not the household level. We model three-month spending as a normally distributed variable which is a linear function of treatment. We estimate that enrollment reduced spending by about 60 pesos on average, with a 95% confidence interval wide enough to include zero --- our data cannot rule out no effect.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

This is the version of your QMD your teacher is most likely to look at closely.

### Exercise 12

Publish your rendered QMD to GitHub Pages. In the Terminal (not the R prompt!), run:

```
quarto publish gh-pages analysis.qmd
```

Copy/paste the resulting URL below.

```{r temperance-12}
question_text(NULL,
  message = "https://<your-github-username>.github.io/sps/analysis.html",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

Commit/push everything.

### Exercise 13

Copy/paste the URL to your GitHub repo.

```{r temperance-13}
question_text(NULL,
  message = "https://github.com/<your-github-username>/sps",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God.

The world confronts us. Make decisions we must.


## Summary
###

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(fit_sps, condition = "treatment", draw = FALSE) |>
  ggplot(aes(x = treatment, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                  color = "grey20", size = 1.1) +
  scale_y_continuous(labels = scales::label_dollar(prefix = "$")) +
  labs(
    title = "Predicted 3-Month Out-of-Pocket Spending by Enrollment Status",
    subtitle = "Enrolled households spend ~60 pesos less, but the CI crosses zero",
    x = "Treatment",
    y = "Predicted 3-month spending",
    caption = "Source: King et al. (2009) SPS rollout via primer.data"
  ) +
  theme_minimal()
```

Households' out-of-pocket spending on health care is a key marker of financial vulnerability. Using a 500-household sample drawn from Mexico's 2005--06 randomized rollout of Seguro Popular, we estimate the average causal effect of enrollment on three-month out-of-pocket spending. One concern is that randomization happened at the community level rather than the household level: within a treated community, enrollment was voluntary, so the comparison of enrolled vs. not-enrolled households is partly a comparison of self-selected groups. We model three-month spending as a normally distributed variable which is a linear function of treatment. We estimate that enrollment reduced average three-month spending by about 60 pesos, with a 95% confidence interval running from roughly -380 to +250 pesos --- a range wide enough that we cannot rule out no effect.

This finding would inform a policy decision about whether to renew or expand a Seguro Popular--like program, and about how to interpret similar randomized rollouts of social-insurance programs in other settings. It does not address every question a policy analyst would face --- it says nothing about catastrophic health spending in particular, about non-financial outcomes like health status, or about how the effect varies across regions or income tiers. But it gives a defensible starting point.

The world is always more uncertain than our models would have us believe.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
