---
title: Colleges
author: David Kane
tutorial:
  id: 07-colleges
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: "Tutorial #07 for Preceptor's Primer"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(gt)

library(tidyverse)
library(tidymodels)
library(primer.data)
library(broom)
library(marginaleffects)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = "90%")
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

x <- colleges |>
  select(tuition, grad_rate, selectivity) |>
  filter(tuition > 2) |>
  drop_na() |>
  mutate(selectivity = factor(as.character(selectivity),
                              levels = c("Elite", "Highly Selective",
                                         "Moderately Selective",
                                         "Lowly Selective", "Non-selective")))

fit_colleges <- linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition + selectivity, data = x)
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial supports the [Colleges](https://ppbds.github.io/primer/07-colleges.html) chapter of [*Preceptor's Primer for Bayesian Data Science: Using the Cardinal Virtues for Inference*](https://ppbds.github.io/primer/) by [David Kane](https://davidkane.info/).

The world confronts us. Make decisions we must.

Imagine that you are a data scientist at a non-profit that helps high-school students choose a college. The Executive Director has a big-picture goal --- make sure students who use the service end up at colleges where they're more likely to graduate, earn back their tuition, and not leave with crushing debt --- and trusts you to figure out what evidence-based recommendations the service should give. Many causal and predictive models would help: predicted graduation rate, predicted post-graduation earnings, predicted debt at graduation, the relationship between major choice and labor-market outcomes. This tutorial builds just one of them: the predicted graduation rate as a function of tuition and selectivity. The estimate alone won't determine which colleges the service recommends, but the graduation-rate piece is one good input to the larger picture. There are many decisions to make.

### Exercise 1

What are the four [Cardinal Virtues](https://en.wikipedia.org/wiki/Cardinal_virtues), in order, which we use to guide our data science work?

```{r introduction-1}
question_text(NULL,
  message = "Wisdom, Justice, Courage, and Temperance.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

The dataset we'll work from is `colleges` in **[primer.data](https://github.com/PPBDS/primer.data)** --- a snapshot of IPEDS (the Integrated Postsecondary Education Data System), the U.S. Department of Education's annual survey of every accredited college and university. The `tuition` column is *sticker price*, not what a typical student actually pays after financial aid; at many private four-year institutions, the average net price is a third or less of the sticker number we are about to model.

### Exercise 2

You should be working inside a GitHub repo named `colleges`, opened in a Codespace from the [`PPBDS/codespace-starter`](https://github.com/PPBDS/codespace-starter) template. If you are not there yet, please create that repo and open it in a Codespace now --- see the [package README](https://github.com/PPBDS/primer/tree/main/primer.tutorials#working-environments-and-repo-setup) if you need setup instructions or are working locally instead.

Once you're inside the `colleges` repo, create a new Quarto document titled `"Colleges"` with yourself as the author, render it, and save it as `analysis.qmd`.

Create a `.gitignore` file with `analysis_files` on the first line followed by a blank line. Save and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The `colleges` tibble holds nearly 900 U.S. four-year institutions from the 2013 IPEDS snapshot; the `tuition` column is sticker price (in $10,000s) and `grad_rate` is the six-year graduation rate, the standard federal accountability measure.

### Exercise 3

In your QMD, put `library(tidyverse)` and `library(primer.data)` in a new code chunk. Render the file.

Add `#| message: false` to the setup chunk to suppress the package-load messages. Add the following to the YAML header:

```
execute:
  echo: false
```

Render again. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    #| message: false\n    library(tidyverse)\n    library(primer.data)\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Rendering a QMD converts the code and prose into an HTML document that shows only what you want readers to see.

### Exercise 4

In the R Terminal, type `?colleges` and paste the Description below.

```{r introduction-4}
question_text(NULL,
  message = "A dataset from the Department of Education's 2013 IPEDS database containing information on U.S. four-year colleges and universities. Variables include tuition (in $10,000), graduation rate, SAT score, enrollment, and a five-level selectivity rating.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Note that `tuition` is recorded in units of $10,000 per year --- a tuition value of `3` means $30,000/year, not $3. Easy to miss; easy to misinterpret later.

### Exercise 5

Graduation rate is the broad topic of this tutorial. Given that topic, which variable in `colleges` should we use as our outcome variable?

```{r introduction-5}
question_text(NULL,
  message = "Our outcome variable is `grad_rate`. It records each college's six-year graduation rate as a proportion between 0 and 1.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

We will use `grad_rate` as our outcome variable.

```{r}
#| echo: false
#| message: false
#| warning: false
colleges |>
  filter(tuition > 2) |>
  drop_na(grad_rate) |>
  ggplot(aes(x = grad_rate)) +
  geom_density(fill = "grey70", color = "grey30") +
  scale_x_continuous(labels = scales::label_percent()) +
  labs(
    title = "Graduation Rates Across US Colleges",
    subtitle = "Most colleges graduate 50% to 80% of their students within six years",
    x = "Graduation Rate",
    y = NULL,
    caption = "Source: IPEDS 2013 via primer.data"
  ) +
  theme_minimal() +
  theme(axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())
```

### Exercise 6

Which variable in `colleges` do you think might have an important connection to `grad_rate`?

```{r introduction-6}
question_text(NULL,
  message = "One candidate is `tuition`. Colleges that charge more tend to have more resources per student, higher admission standards, and students from wealthier families --- all of which correlate with graduation rates. Another candidate is `selectivity`, which directly captures how choosy the college is about whom it admits.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

With a predictive model, each individual unit has only one observed outcome. Predictive models have no "treatments" --- only covariates.

### Exercise 7

Specify two different groups of colleges which have different values for `tuition` and which might have different average values for `grad_rate`.

```{r introduction-7}
question_text(NULL,
  message = "Consider two groups: colleges with `tuition` below $15,000 per year, and colleges with `tuition` above $40,000 per year. These two groups might have different average graduation rates.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

In predictive models, do not use "cause," "influence," "impact," or anything else which suggests causation. The best phrasing is in terms of "differences" between groups of units with different values for a covariate of interest.

### Exercise 8

Write a predictive question connecting the outcome `grad_rate` to `tuition`, the covariate of interest.

```{r introduction-8}
question_text(NULL,
  message = "What is the difference in expected graduation rates between colleges with tuition of $20,000 and colleges with tuition of $30,000?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

School tuition in this data is the *stated* tuition --- the sticker price. At many colleges, most students receive significant financial aid, so the average tuition actually paid is far lower than the recorded number. Our question therefore compares colleges by sticker price, not by net price. A version of the question that compared net prices would need a different dataset.


## Wisdom
###

Data science starts with some broad questions and a data set which might help us to answer them.

The specific question: *What is the difference in expected graduation rates between colleges with tuition of $20,000 and colleges with tuition of $30,000?*

We will work from the Department of Education's 2013 IPEDS database, available in the `colleges` tibble of the **primer.data** package.

### Exercise 1

In your own words, describe the key components of Wisdom when working on a data science problem.

```{r wisdom-1}
question_text(NULL,
  message = "Wisdom begins with a question and then moves on to the creation of a Preceptor Table and an examination of our data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

> *The combination of some data and an aching desire for an answer does not ensure that a reasonable answer can be extracted from a given body of data.* --- John W. Tukey

### Exercise 2

Define a Preceptor Table.

```{r wisdom-2}
question_text(NULL,
  message = "A Preceptor Table is the smallest possible table of data with rows and columns such that, if there is no missing data, we can easily calculate the quantity of interest.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The Preceptor Table does not include all the covariates which you will eventually include in your model. It only includes, along with the outcome(s), covariates which are mentioned in your question.

### Exercise 3

We saw the distribution of `grad_rate` on its own in the Introduction. Now look at how `grad_rate` varies with `tuition`. Write R code that plots `grad_rate` on the y-axis against `tuition` on the x-axis using points with `alpha = 0.4`, a smooth trend line, and descriptive labels. Paste the code below.

```{r wisdom-3}
question_text(NULL,
  message = "    colleges |>\n      filter(tuition > 2) |>\n      drop_na(grad_rate, tuition) |>\n      ggplot(aes(x = tuition, y = grad_rate)) +\n      geom_point(alpha = 0.4, color = \"grey30\") +\n      geom_smooth(method = \"lm\", color = \"steelblue\", se = FALSE) +\n      scale_x_continuous(labels = scales::label_dollar(scale = 10000)) +\n      scale_y_continuous(labels = scales::label_percent()) +\n      labs(\n        title = \"Graduation Rate by Tuition\",\n        subtitle = \"Higher-tuition colleges tend to have higher graduation rates\",\n        x = \"Tuition\", y = \"Graduation Rate\",\n        caption = \"Source: IPEDS 2013 via primer.data\"\n      ) +\n      theme_minimal()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
colleges |>
  filter(tuition > 2) |>
  drop_na(grad_rate, tuition) |>
  ggplot(aes(x = tuition, y = grad_rate)) +
  geom_point(alpha = 0.4, color = "grey30") +
  geom_smooth(method = "lm", color = "steelblue", se = FALSE) +
  scale_x_continuous(labels = scales::label_dollar(scale = 10000)) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(
    title = "Graduation Rate by Tuition",
    subtitle = "Higher-tuition colleges tend to have higher graduation rates",
    x = "Tuition",
    y = "Graduation Rate",
    caption = "Source: IPEDS 2013 via primer.data"
  ) +
  theme_minimal()
```

### Exercise 4

Describe the key components of Preceptor Tables in general, without worrying about this specific problem. Use words like "units," "outcomes," and "covariates."

```{r wisdom-4}
question_text(NULL,
  message = "The rows of the Preceptor Table are the units. The outcome is at least one of the columns. If the problem is causal, there will be at least two (potential) outcome columns. The other columns are covariates. If the problem is causal, at least one of the covariates will be considered a treatment.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

This problem is predictive, so there is one outcome column: `grad_rate`. The key covariate is `tuition`.

### Exercise 5

What are the units for this problem?

```{r wisdom-5}
question_text(NULL,
  message = "Each row is one US four-year college or university *today* --- in 2026, when the non-profit is helping its students choose. The Preceptor Table would have a row for every such institution, roughly 900 in total. Note the time mismatch: our data is from 2013, so the Preceptor Table and the data refer to the same set of colleges but at very different points in time.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

We model units, but we only really care about aggregates.

### Exercise 6

What is the outcome variable for this problem?

```{r wisdom-6}
question_text(NULL,
  message = "`grad_rate` --- the proportion of entering students who graduate within six years, a number between 0 and 1.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The outcome we really care about --- "did this college's graduates succeed in life?" --- is not measured in our data. `grad_rate` is a narrow proxy: it tells us completion, not outcomes after graduation.

### Exercise 7

What is a covariate which you think might be useful for this problem, regardless of whether or not it might be included in the data?

```{r wisdom-7}
question_text(NULL,
  message = "`selectivity` is a strong candidate --- more selective schools admit students who are more likely to graduate regardless of the school's support. SAT score distribution, percent of first-generation students, and endowment per student are others.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

The term "covariates" is used in at least three ways. First, it is all the variables which might be useful, regardless of whether or not we have the data. Second, it is all the variables for which we have data. Third, it is the set of variables in the data which we end up using in the model.

### Exercise 8

What are the treatments, if any, for this problem?

```{r wisdom-8}
question_text(NULL,
  message = "There are no treatments. This is a predictive model --- nothing in the `colleges` dataset can be manipulated (we cannot force a college to charge more tuition or to become more selective). Predictive models have covariates only.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Remember that a treatment is just another covariate which, for the purposes of a specific problem, we are assuming can be manipulated.

### Exercise 9

What moment in time does the Preceptor Table refer to?

```{r wisdom-9}
question_text(NULL,
  message = "Today, in 2026 --- when the non-profit's students are choosing colleges. The graduation rate is the six-year rate, so each row's `grad_rate` would refer to the cohort that entered in 2020. Our data is from 2013, but the Preceptor Table is about now.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

A Preceptor Table can never really refer to an exact instant in time.

### Exercise 10

Describe in words the Preceptor Table for this problem.

```{r wisdom-10}
question_text(NULL,
  message = "The Preceptor Table has three columns: one identifying each college, one for `grad_rate`, and one for `tuition`. Each row represents one US four-year institution in 2026 --- the year of the question, not the year of the data.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

```{r}
#| echo: false
#| results: asis
pre_tibble <- tibble::tribble(
  ~`College`,            ~`Graduation Rate`, ~`Tuition ($10,000)`,
  "Amherst College",     "0.95",             "7.4",
  "University of Iowa",  "0.72",             "3.3",
  "...",                 "...",              "...",
  "Wagner College",      "0.55",             "5.5"
)

pre_title_footnote     <- "A Preceptor Table is the smallest table such that, if every cell were filled with its true value, answering the question 'What is the difference in expected graduation rates between colleges with tuition of $20,000 and colleges with tuition of $30,000?' would be easy."
pre_units_footnote     <- "Each row is one US four-year college or university in 2026. The three example rows are real institutions chosen to span the selectivity range; missing rows represent the other roughly 900 institutions not shown."
pre_outcome_footnote   <- "Six-year graduation rate, a proportion between 0 and 1. Reflects the entering class from six years before the snapshot year."
pre_covariate_footnote <- "Annual tuition and fees in units of $10,000. A value of 3 means $30,000 per year."

pre_gt_html <- gt::gt(pre_tibble, id = "preceptor_tbl") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`College`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Graduation Rate`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Tuition ($10,000)`)) |>
  gt::cols_align(align = "left" , columns = c(`College`)) |>
  gt::cols_align(align = "right", columns = c(`Graduation Rate`, `Tuition ($10,000)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = "unit_span")
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = c("outcome_span", "covariates_span"))
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html(inline_css = FALSE)

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 11

We need a cleaned tibble. In `analysis.qmd`, add a new code chunk that creates a tibble called `x` by taking `colleges`, selecting `tuition`, `grad_rate`, and `selectivity`, filtering to rows with tuition above $20,000 (so `tuition > 2`), dropping rows with missing values, and recoding `selectivity` as a factor with the levels in the order `Elite`, `Highly Selective`, `Moderately Selective`, `Lowly Selective`, `Non-selective`. Then on the next line, print `head(x)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r wisdom-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r wisdom-11-test}
#| echo: true
x <- colleges |>
  select(tuition, grad_rate, selectivity) |>
  filter(tuition > 2) |>
  drop_na() |>
  mutate(selectivity = factor(as.character(selectivity),
                              levels = c("Elite", "Highly Selective",
                                         "Moderately Selective",
                                         "Lowly Selective", "Non-selective")))
head(x)
```

###

`filter(tuition > 2)` restricts to schools charging at least $20,000/year, which removes a handful of community-college-priced entries in the IPEDS snapshot that don't fit the four-year residential model we care about. `drop_na()` silently removes rows missing any of the three variables --- a simplification Justice will revisit. The factor recoding pins the selectivity levels in the order we want them to appear in later tables and plots --- otherwise they would sort alphabetically.

### Exercise 12

The pipeline that builds `x` runs every time you render. It's quick here, but `x` will be used by every later modeling exercise, so it earns its own cached create chunk. The next three exercises set that up: clean the chunk, turn caching on, add the cache directory to `.gitignore`.

First, clean up. In your chunk from Exercise 11, delete the `head(x)` line at the bottom, leaving just the `x <- colleges ...` pipeline. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r wisdom-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

<pre><code>x <- colleges |>
  select(tuition, grad_rate, selectivity) |>
  filter(tuition > 2) |>
  drop_na() |>
  mutate(selectivity = factor(as.character(selectivity),
                              levels = c("Elite", "Highly Selective",
                                         "Moderately Selective",
                                         "Lowly Selective", "Non-selective")))
</code></pre>

###

A chunk ready to cache should contain only the artifact it builds.

### Exercise 13

Add `#| cache: true` as the first line inside the chunk. Render the QMD. Rendering creates an `analysis_cache/` directory next to your `analysis.qmd`, which holds the saved tibble so Quarto can reload it on subsequent renders rather than re-running the pipeline every time. To confirm the directory is there, run `ls` in the **bash Terminal**. CP/CR.

```{r wisdom-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<pre><code>$ ls
analysis.qmd  analysis_cache  analysis_files  analysis.html
</code></pre>

###

The savings here are small; the same pattern matters more when the cached pipeline takes seconds (or minutes) rather than milliseconds.

### Exercise 14

The Source Control change count just jumped, because `analysis_cache/` and everything inside it now look like untracked work to git. Cached files should never go to GitHub --- they're machine-specific and regenerated on every render.

Add `analysis_cache` to your `.gitignore` on its own line. In the R Terminal, run `show_file(".gitignore")`. CP/CR.

```{r wisdom-14}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<pre><code>analysis_files
analysis_cache
</code></pre>

###

The Source Control change count should drop back to whatever it was before the render.

### Exercise 15

Write the first two sentences of the summary paragraph. The first sentence is a general statement about the overall topic, mentioning the general class of outcome and at least one covariate. The second sentence introduces the data source and the specific question.

```{r wisdom-15}
question_text(NULL,
  message = "Where students go to college and what they pay for it varies enormously, and so does the likelihood that they will finish. We help today's students choose by examining how graduation rate varies with tuition and selectivity, using 2013 IPEDS data on roughly 900 US four-year colleges and universities --- the most recent reliable national snapshot we have.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Read our answer. It will not be the same as yours. Do not copy/paste exactly. Add your two sentences to the summary paragraph. Render the QMD, then commit/push.


## Justice
###

Wisdom gives us the Preceptor Table and the data.

```{r}
#| echo: false
#| results: asis
pre_tibble_j <- tibble::tribble(
  ~`College`,            ~`Graduation Rate`, ~`Tuition ($10,000)`,
  "Amherst College",     "0.95",             "7.4",
  "University of Iowa",  "0.72",             "3.3",
  "...",                 "...",              "...",
  "Wagner College",      "0.55",             "5.5"
)

pre_title_footnote_j     <- "A Preceptor Table is the smallest table such that, if every cell were filled with its true value, answering the question 'What is the difference in expected graduation rates between colleges with tuition of $20,000 and colleges with tuition of $30,000?' would be easy."
pre_units_footnote_j     <- "Each row is one US four-year college or university."
pre_outcome_footnote_j   <- "Six-year graduation rate, 0 to 1."
pre_covariate_footnote_j <- "Annual tuition in units of $10,000."

pre_gt_html_j <- gt::gt(pre_tibble_j, id = "preceptor_tbl_justice") |>
  gt::tab_header(title = "Preceptor Table") |>
  gt::tab_spanner(label = "Unit",      id = "unit_span",
                  columns = c(`College`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Graduation Rate`)) |>
  gt::tab_spanner(label = "Covariate", id = "covariates_span",
                  columns = c(`Tuition ($10,000)`)) |>
  gt::cols_align(align = "left" , columns = c(`College`)) |>
  gt::cols_align(align = "right", columns = c(`Graduation Rate`, `Tuition ($10,000)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = "unit_span")
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = c("outcome_span", "covariates_span"))
  ) |>
  gt::tab_footnote(footnote = pre_title_footnote_j,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pre_units_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pre_outcome_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pre_covariate_footnote_j,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #preceptor_tbl_justice .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html(inline_css = FALSE)

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pre_gt_html_j,
  "</div>\n",
  "```\n",
  sep = ""
)
```

```{r}
#| echo: false
#| results: asis
data_tibble <- tibble::tribble(
  ~`College`,            ~`Year`, ~`Graduation Rate`, ~`Tuition ($10,000)`, ~`Selectivity`,
  "Amherst College",     "2013",  "0.95",             "4.6",                "Elite",
  "University of Iowa",  "2013",  "0.71",             "2.7",                "Moderately Selective",
  "...",                 "...",   "...",              "...",                "...",
  "Wagner College",      "2013",  "0.51",             "3.6",                "Lowly Selective"
)

data_title_footnote     <- "Integrated Postsecondary Education Data System (IPEDS), US Department of Education, 2013 snapshot. Covers ~900 US four-year colleges and universities reporting graduation rates, tuition, selectivity, and other institutional metrics. After filtering to tuition > $20,000 and dropping rows with missing values, we have 542 colleges. The example rows above are the same three institutions shown in the Preceptor Table, recorded here at their 2013 values."
data_units_footnote     <- "Each row is one US four-year college in the 2013 reporting year. The graduation rate is the six-year rate for the cohort that entered six years before."
data_outcome_footnote   <- "Graduation rate is the proportion of first-time, full-time bachelor's-degree-seeking students who graduated within six years. IPEDS collects this from institutional reports; definitions have been consistent across reporting years."
data_covariate_footnote <- "Tuition is reported annual tuition and fees, converted to units of $10,000. `Selectivity` is a five-level ordered factor assigned by IPEDS based on admissions data: Elite, Highly Selective, Moderately Selective, Lowly Selective, Non-selective."

data_gt_html <- gt::gt(data_tibble, id = "data_tbl") |>
  gt::tab_header(title = "Data: IPEDS US Four-Year Colleges, 2013") |>
  gt::tab_spanner(label = "Unit/Time",  id = "unit_span",
                  columns = c(`College`, `Year`)) |>
  gt::tab_spanner(label = "Outcome",    id = "outcome_span",
                  columns = c(`Graduation Rate`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Tuition ($10,000)`, `Selectivity`)) |>
  gt::cols_align(align = "left" , columns = c(`College`, `Selectivity`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Graduation Rate`, `Tuition ($10,000)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = data_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = data_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = data_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = data_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #data_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html(inline_css = FALSE)

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  data_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 1

In your own words, describe the components of Justice when working on a data science problem.

```{r justice-1}
question_text(NULL,
  message = "Justice reviews the Population Table and selects the formula for the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Justice is about concerns that you (or your critics) might have, reasons why the model you create might not work as well as you hope. The exercises that follow walk through several such concerns by name. Justice also picks the formula for the data generating mechanism --- different outcome variables call for different formulas, a piece that gets more technical names in later tutorials.

### Exercise 2

In your own words, define "validity" as we use the term.

```{r justice-2}
question_text(NULL,
  message = "Validity is the consistency, or lack thereof, in the columns of the data set and the corresponding columns in the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Validity is always about the columns in the Preceptor Table and the data. Just because two columns share a name does not mean that they are the same thing.

### Exercise 3

Provide one reason why the assumption of validity might not hold for the outcome variable `grad_rate` or for the covariate `tuition`. Use the words "column" or "columns" in your answer.

```{r justice-3}
question_text(NULL,
  message = "The `grad_rate` column in IPEDS measures six-year completion of *a* bachelor's degree, but not necessarily at the college where the student started (IPEDS only counts students who graduate from the original institution). A student who transfers and finishes elsewhere counts as a non-graduation, so the column understates the true outcome. That is different from the Preceptor Table's 'did this student finish college in six years?' --- the two columns are not quite the same concept.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Validity allows us to construct the Population Table --- the bridge from data to Preceptor Table.

### Exercise 4

In your own words, define a Population Table.

```{r justice-4}
question_text(NULL,
  message = "The Population Table includes a row for each unit/time combination in the underlying population from which both the Preceptor Table and the data are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The Population Table is almost always much bigger than the combination of the Preceptor Table and the data.

### Exercise 5

Specify the unit/time combinations which define each row in this Population Table.

```{r justice-5}
question_text(NULL,
  message = "Each row is one US four-year college in a given year. Data rows are the 542 colleges in our filtered 2013 IPEDS snapshot. Preceptor rows are US four-year colleges in 2026 --- the year the non-profit's students are choosing. Many of the same institutions appear in both blocks, but the rows describe them at different times: the 2013 row for Amherst and the 2026 row for Amherst are different rows in the Population Table even though they share an institution.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

```{r}
#| echo: false
#| results: asis
pop_tibble <- tibble::tribble(
  ~Source    , ~`College`           , ~`Year`, ~`Graduation Rate`, ~`Tuition ($10,000)`, ~`Selectivity`,
  "..."      , "..."                , "..."  , "..."             , "..."               , "..."         ,
  "Data"     , "Amherst College"    , "2013" , "0.95"            , "4.6"               , "Elite"       ,
  "Data"     , "University of Iowa" , "2013" , "0.71"            , "2.7"               , "Moderately Selective",
  "Data"     , "..."                , "..."  , "..."             , "..."               , "..."         ,
  "Data"     , "Wagner College"     , "2013" , "0.51"            , "3.6"               , "Lowly Selective",
  "..."      , "..."                , "..."  , "..."             , "..."               , "..."         ,
  "Preceptor", "Amherst College"    , "2026" , "0.95"            , "7.4"               , "Elite"       ,
  "Preceptor", "University of Iowa" , "2026" , "0.72"            , "3.3"               , "Moderately Selective",
  "Preceptor", "..."                , "..."  , "..."             , "..."               , "..."         ,
  "Preceptor", "Wagner College"     , "2026" , "0.55"            , "5.5"               , "Lowly Selective",
  "..."      , "..."                , "..."  , "..."             , "..."               , "..."
)

pop_title_footnote     <- "This table combines the 2013 IPEDS snapshot with the Preceptor Table's view of the same colleges in 2026, when the non-profit's students are choosing. The data and Preceptor rows refer to overlapping institutions but at different points in time."
pop_units_footnote     <- "Each row is one college at one point in time. The same three institutions --- Amherst College, the University of Iowa, and Wagner College --- appear in both the Data block (2013) and the Preceptor block (2026). Their tuition values rise and their graduation rates barely move over the 13-year gap; the assumption that the *parameters* of the model are stable across the gap is what stability is about."
pop_outcome_footnote   <- "Six-year graduation rate, 0 to 1."
pop_covariate_footnote <- "Tuition in units of $10,000. Selectivity is IPEDS's five-level ordered factor."

pop_gt_html <- gt::gt(pop_tibble, id = "population_tbl") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`College`, `Year`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Graduation Rate`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Tuition ($10,000)`, `Selectivity`)) |>
  gt::cols_align(align = "left" , columns = c(`Source`, `College`, `Selectivity`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Graduation Rate`, `Tuition ($10,000)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html(inline_css = FALSE)

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html,
  "</div>\n",
  "```\n",
  sep = ""
)
```

### Exercise 6

In your own words, define "stability."

```{r justice-6}
question_text(NULL,
  message = "Stability means that the relationship between the columns in the Population Table is the same for three categories of rows: the data, the Preceptor Table, and the larger population from which both are drawn.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Time is what makes stability hard. Our IPEDS snapshot is from 2013, but the question is about 2026 --- a gap of 13 years over which federal aid policy, the cost-of-living, online education, and selective-college admissions practices have all changed. Stability asks whether the relationship among the columns of the Population Table is the same across data and Preceptor Table. We sharpen what "the relationship" means in the next exercise.

### Exercise 7

Provide one reason why stability might not hold.

```{r justice-7}
question_text(NULL,
  message = "The slope of `grad_rate` on `tuition` may not be the same in 2026 as it was in 2013. Two specific mechanisms could move that slope: (a) the rise of large institutional aid budgets at high-tuition private colleges has loosened the link between sticker price and per-student spending, which would *flatten* the slope; (b) the growth of subsidized federal loans and income-share agreements has changed who can afford which schools, which could either steepen or flatten the slope depending on how it interacts with selectivity. The point is the *coefficient* connecting `tuition` to `grad_rate` --- not the level of tuition itself, which has obviously risen but is a distribution change, not a stability violation.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Stability is about parameters, not distributions. Tuition has surely risen between 2013 and 2026 --- that is a *distribution* change, not a stability violation. Even a change in the spread of tuition values, or in the mix of selectivity tiers represented across colleges, is not by itself a stability violation. What *would* break stability is a change in the slope of `grad_rate` on `tuition` --- the parameter $\beta_1$. The level and shape of the tuition distribution can drift however they like; what has to stay constant is the coefficient connecting tuition to graduation rate.

### Exercise 8

In your own words, define representativeness.

```{r justice-8}
question_text(NULL,
  message = "Representativeness, or the lack thereof, concerns two relationships among the rows in the Population Table. The first is between the data and the other rows. The second is between the other rows and the Preceptor Table.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Ideally we would like both the Preceptor Table and the data to be random samples from the population. Sadly, this is almost never the case.

### Exercise 9

Provide one reason, involving the relationship between the data and the population, why representativeness might not hold.

```{r justice-9}
question_text(NULL,
  message = "IPEDS reports from colleges that participate in federal financial-aid programs, which is nearly all accredited four-year institutions but not quite all. Colleges that do not participate --- a handful of religious and specialty schools --- are systematically absent. More importantly, our filter to `tuition > 2` drops low-tuition and community-college-priced schools, so the 542 rows we fit on are a deliberately non-representative slice.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

A non-representative sample doesn't *guarantee* a biased estimate --- it's possible for the bias to wash out by chance. But "by chance" is the only mechanism left to us; we have no principled reason to defend the estimate to a critic. Treat representativeness violations as a reason the parameter *might* be biased, and adjust uncertainty accordingly. Representativeness also has a *second* link --- between the population and the Preceptor Table --- which later tutorials will exercise; here we focus on the data-to-population link.

### Exercise 10

A statistical model consists of two parts: the probability family and the link function. The probability family is the probability distribution that generates the randomness in our data. Our outcome variable is `grad_rate` --- continuous. What probability family should we use?

```{r justice-10}
question_text(NULL,
  message = "Normal: $Y \\sim N(\\mu, \\sigma^2)$. The Normal distribution is the standard choice for continuous outcomes.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

While you're here, also add `library(tidymodels)` to the setup chunk in `analysis.qmd` so the rest of Courage can call `linear_reg()` and `fit()`.

### Exercise 11

The link function is the mathematical formula that links our data to the unknown parameters. What is the link function for a continuous outcome?

```{r justice-11}
question_text(NULL,
  message = "Linear: $\\mu = \\beta_0 + \\beta_1 X_1 + \\ldots + \\beta_n X_n$. Our model will be a linear regression.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 4)
```

###

While you're here, also add `library(broom)` to the setup chunk in `analysis.qmd` so we can call `tidy()` on the fitted model in Courage.

### Exercise 12

Write one sentence highlighting a potential weakness in your model. So far our summary paragraph looks like this:

> Where students go to college and what they pay for it varies enormously, and so does the likelihood that they will finish. We help today's students choose by examining how graduation rate varies with tuition and selectivity, using 2013 IPEDS data on roughly 900 US four-year colleges and universities --- the most recent reliable national snapshot we have.

```{r justice-12}
question_text(NULL,
  message = "One concern is the 13-year gap between our data (2013) and the question (2026). Tuition has risen much faster than inflation over that span, and the relationship between price and graduation rate may have shifted. The 2013 coefficients we estimate may not reflect what they would be if we could fit the same model on today's colleges.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a weakness sentence to your summary paragraph. Render the QMD, then commit/push.


## Courage
###

Justice gives us the Population Table and the abstract data generating mechanism.

```{r}
#| echo: false
#| results: asis
pop_tibble_c <- tibble::tribble(
  ~Source    , ~`College`           , ~`Year`, ~`Graduation Rate`, ~`Tuition ($10,000)`, ~`Selectivity`,
  "..."      , "..."                , "..."  , "..."             , "..."               , "..."         ,
  "Data"     , "Amherst College"    , "2013" , "0.95"            , "4.6"               , "Elite"       ,
  "Data"     , "University of Iowa" , "2013" , "0.71"            , "2.7"               , "Moderately Selective",
  "Data"     , "..."                , "..."  , "..."             , "..."               , "..."         ,
  "Data"     , "Wagner College"     , "2013" , "0.51"            , "3.6"               , "Lowly Selective",
  "..."      , "..."                , "..."  , "..."             , "..."               , "..."         ,
  "Preceptor", "Amherst College"    , "2026" , "0.95"            , "7.4"               , "Elite"       ,
  "Preceptor", "University of Iowa" , "2026" , "0.72"            , "3.3"               , "Moderately Selective",
  "Preceptor", "..."                , "..."  , "..."             , "..."               , "..."         ,
  "Preceptor", "Wagner College"     , "2026" , "0.55"            , "5.5"               , "Lowly Selective",
  "..."      , "..."                , "..."  , "..."             , "..."               , "..."
)

pop_title_footnote_c     <- "The 2013 IPEDS snapshot (Data rows) combined with the same colleges in 2026 (Preceptor rows). Same institutions, different points in time."
pop_units_footnote_c     <- "Each row is one college at one point in time. The three colleges shown appear in both the Data and Preceptor blocks."
pop_outcome_footnote_c   <- "Graduation rate, 0 to 1."
pop_covariate_footnote_c <- "Tuition in units of $10,000. Selectivity is IPEDS's five-level factor."

pop_gt_html_c <- gt::gt(pop_tibble_c, id = "population_tbl_courage") |>
  gt::tab_header(title = "Population Table") |>
  gt::tab_spanner(label = "Unit/Time", id = "unit_span",
                  columns = c(`College`, `Year`)) |>
  gt::tab_spanner(label = "Outcome",   id = "outcome_span",
                  columns = c(`Graduation Rate`)) |>
  gt::tab_spanner(label = "Covariates", id = "covariates_span",
                  columns = c(`Tuition ($10,000)`, `Selectivity`)) |>
  gt::cols_align(align = "left" , columns = c(`Source`, `College`, `Selectivity`)) |>
  gt::cols_align(align = "right", columns = c(`Year`, `Graduation Rate`, `Tuition ($10,000)`)) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::tab_options(
    heading.title.font.size   = "1.5em",
    heading.title.font.weight = "bolder",
    column_labels.font.weight = "normal"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_spanners()
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "left"),
    locations = gt::cells_column_spanners(spanners = c("unit_span", "covariates_span"))
  ) |>
  gt::tab_style(
    style     = gt::cell_text(align = "right"),
    locations = gt::cells_column_spanners(spanners = "outcome_span")
  ) |>
  gt::tab_footnote(footnote = pop_title_footnote_c,
                   locations = gt::cells_title()) |>
  gt::tab_footnote(footnote = pop_units_footnote_c,
                   locations = gt::cells_column_spanners(spanners = "unit_span")) |>
  gt::tab_footnote(footnote = pop_outcome_footnote_c,
                   locations = gt::cells_column_spanners(spanners = "outcome_span")) |>
  gt::tab_footnote(footnote = pop_covariate_footnote_c,
                   locations = gt::cells_column_spanners(spanners = "covariates_span")) |>
  gt::opt_css(
    css = "
      #population_tbl_courage .gt_footnote {
        max-width: 1px;
        word-break: break-word;
      }
    "
  ) |>
  gt::as_raw_html(inline_css = FALSE)

cat(
  "```{=html}\n",
  '<div style="display: inline-block; width: auto; max-width: 100%;">',
  pop_gt_html_c,
  "</div>\n",
  "```\n",
  sep = ""
)
```

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n + \epsilon$$

with $\epsilon \sim N(0, \sigma^2)$.

### Exercise 1

In your own words, describe the components of the virtue of Courage for analyzing data.

```{r courage-1}
question_text(NULL,
  message = "Courage creates the data generating mechanism.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The abstract data generating mechanism shown in the preamble uses generic variables --- $Y$, $X_1$, $X_2$, and so on --- because our purpose is to describe the general mathematical structure. At the end of Justice we have decided the functional family (linear with normal errors), but not the number or identity of covariates we will ultimately include. Courage fills in the missing pieces.

### Exercise 2

Because our outcome variable is continuous, add a code chunk to `analysis.qmd` that begins building the model with `linear_reg(engine = "lm")`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-2-test}
#| echo: true
linear_reg(engine = "lm")
```

###

Within the R modeling community, **[tidymodels](https://www.tidymodels.org/)** is the framework most authors reach for first. It bundles a consistent interface for linear regression, logistic regression, multinomial models, and many more, all sharing the same `fit()` syntax.

### Exercise 3

Extend the chunk in `analysis.qmd` to `linear_reg(engine = "lm") |> fit(grad_rate ~ selectivity, data = x)`. This fits a first candidate model using only the five-level `selectivity` factor. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-3-test}
#| echo: true
linear_reg(engine = "lm") |>
  fit(grad_rate ~ selectivity, data = x)
```

###

A categorical covariate with five values produces four dummy 0/1 variables --- one short of the number of levels, because the "missing" category is absorbed into the intercept. For `selectivity`, the reference level is `Elite`; each of the other four rows in the parameter table is the offset from Elite.

### Exercise 4

Here is the parameter table for this first model:

```{r}
#| echo: false
linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ selectivity, data = x) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

Looking at the `selectivityLowly Selective` row, write a sentence interpreting the -0.376 estimate.

```{r courage-4}
question_text(NULL,
  message = "When we compare colleges in the Lowly Selective tier with colleges in the Elite tier (the reference category), the Lowly Selective colleges have a graduation rate about 37.6 percentage points lower.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

For a multi-level categorical, every dummy coefficient is read *relative to the reference level*, which is `Elite` here. Notice the language: *"when we compare,"* not *"raising selectivity raises graduation rate"* --- selectivity isn't something a college can be "raised"; we are comparing two groups of colleges that differ in their selectivity tier.

### Exercise 5

Change the formula to `fit(grad_rate ~ tuition, data = x)`. This refits with the continuous covariate `tuition` instead of the selectivity factor. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-5-test}
#| echo: true
linear_reg(engine = "lm") |>
  fit(grad_rate ~ tuition, data = x)
```

###

A continuous covariate produces a single slope coefficient. The reading is per-unit comparison: how does the predicted outcome differ between two groups whose covariate values differ by one?

### Exercise 6

Here is the parameter table for the second candidate model:

```{r}
#| echo: false
linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition, data = x) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

Write a sentence interpreting the slope on `tuition` (0.164).

```{r courage-6}
question_text(NULL,
  message = "When we compare two groups of colleges differing by $10,000 in tuition, the higher-tuition group has a graduation rate about 16.4 percentage points higher. This is the *unadjusted* relationship --- we have not yet held selectivity constant.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

The slope is **a comparison, not a change**: we are comparing two groups of colleges that differ in `tuition`, not describing what happens when one college's tuition rises. *Predictive* models warrant comparison language only --- nothing here says raising the price *causes* better graduation outcomes; it says higher-priced colleges, on average, graduate more of their students.

### Exercise 7

Fit the final model in `analysis.qmd` with the formula `fit(grad_rate ~ tuition + selectivity, data = x)`. This combines the two candidates. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-7-test}
#| echo: true
linear_reg(engine = "lm") |>
  fit(grad_rate ~ tuition + selectivity, data = x)
```

###

When a model includes two or more covariates, every interpretation gains an *adjustment clause*: each coefficient is read holding the other covariates constant.

### Exercise 8

Here is the parameter table for the final model:

```{r}
#| echo: false
fit_colleges |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

Write a sentence interpreting the slope on `tuition` (0.090) in this final model. Notice that the value has shrunk from the 0.164 in the previous model.

```{r courage-8}
question_text(NULL,
  message = "When we compare two groups of colleges in the same selectivity tier but differing by $10,000 in tuition, the higher-tuition group has a graduation rate about 9.0 percentage points higher, adjusting for selectivity. The unadjusted slope was 16.4 percentage points; adjusting for selectivity has shrunk it to 9.0. Much of what looked like a tuition effect in the previous model was really a selectivity effect.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

The shrinkage from 16.4 to 9.0 percentage points is the effect of adjustment. Selectivity and tuition are correlated --- elite colleges tend to charge more --- so the unadjusted slope on tuition was capturing some of selectivity's signal. Once we put both in the model, each coefficient describes a comparison *holding the other constant*. We care if a confidence interval excludes zero; the tuition CI [0.074, 0.106] does, so the model says higher-tuition colleges have a higher *expected* graduation rate than lower-tuition colleges in the same selectivity tier. That does not mean every higher-tuition college has a higher graduation rate than every lower-tuition one --- the two distributions overlap, and there are plenty of low-tuition colleges that graduate more of their students than some high-tuition colleges.

### Exercise 9

We've settled on the final fit. In `analysis.qmd`, add a new code chunk that assigns the result of `linear_reg() |> set_engine("lm") |> fit(grad_rate ~ tuition + selectivity, data = x)` to `fit_colleges`, then on the next line prints `fit_colleges`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-9-test}
#| echo: true
fit_colleges <- linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition + selectivity, data = x)
fit_colleges
```

###

In data science, **we deal with words, math, and code, but the most important of these is code.**

### Exercise 10

Re-fitting the model on every render is wasteful. `fit_colleges` earns its own cached create chunk. As with the first cache, we clean up first, then turn caching on. (No `.gitignore` step this time --- `analysis_cache` is already ignored.)

Clean up. In your chunk from Exercise 9, delete the `fit_colleges` line at the bottom, leaving just the `fit_colleges <- linear_reg() ...` assignment. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

<pre><code>fit_colleges <- linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition + selectivity, data = x)
</code></pre>

###

Ready to cache.

### Exercise 11

Add `#| cache: true` as the first line inside the chunk. Render the QMD. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

<pre><code>#| cache: true
fit_colleges <- linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition + selectivity, data = x)
</code></pre>

###

`#| cache: true` tells Quarto to store the fitted object and reload it from disk on subsequent renders rather than re-fitting every time.

### Exercise 12

One way to check whether our model is reasonable is to compare the distribution of our actual outcome (`grad_rate`) with the distribution of fitted values from the model. If the model is well-specified, the two distributions should look similar.

```{r}
#| echo: false
#| message: false
#| warning: false
fitted_df <- tibble(
  value  = c(x$grad_rate, predict(extract_fit_engine(fit_colleges))),
  source = c(rep("Actual", nrow(x)),
             rep("Fitted", nrow(x)))
)

ggplot(fitted_df, aes(x = value, fill = source)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(labels = scales::label_percent()) +
  labs(
    title = "Actual vs. Fitted Graduation Rates",
    subtitle = "The fitted distribution is narrower --- our model smooths out the tails",
    x = "Graduation Rate",
    y = NULL,
    fill = "Source",
    caption = "Source: IPEDS 2013 via primer.data"
  ) +
  theme_minimal() +
  theme(axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())
```

###

In a good model, the distribution of fitted values looks like the distribution of the actual outcome. Big divergences mean our model is missing something --- here, the fitted values cluster more tightly than the actual graduation rates do, which means our model can't reproduce the very-high and very-low extremes. That's a real limitation; whether it matters depends on which questions we ask the fit later.

### Exercise 13

With the parameters estimated, we can write the fitted model as a concrete formula:

$$
\begin{aligned}
\widehat{\text{Graduation Rate}} =\ & 0.487 + 0.090 \cdot \text{Tuition} \\
&{} - 0.078 \cdot \text{Highly Selective} \\
&{} - 0.128 \cdot \text{Moderately Selective} \\
&{} - 0.213 \cdot \text{Lowly Selective} \\
&{} - 0.205 \cdot \text{Non-selective}
\end{aligned}
$$

Each selectivity term is a dummy 0/1 variable (e.g. `Highly Selective` is 1 if the college is Highly Selective, 0 otherwise). The baseline case --- an Elite college with every selectivity dummy = 0 --- has predicted graduation rate $0.487 + 0.090 \cdot \text{tuition}$. Values rounded to three decimal places, fine enough that each estimate and its confidence-interval bounds are visibly distinct.

**This is our data generating mechanism.**

###

A data generating mechanism is just a formula, something which we can write down and implement with computer code.

### Exercise 14

Add `*_cache` to `.gitignore`. In the R Terminal, run `show_file(".gitignore")`. CP/CR.

```{r courage-14}
question_text(NULL,
  message = "    > show_file(\".gitignore\")\n    analysis_files\n    *_cache",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Cached objects are often large and don't belong on GitHub.

### Exercise 15

In `analysis.qmd`, add a chunk that runs `tidy(fit_colleges, conf.int = TRUE)`. This returns 95% intervals for all the parameters in our model. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r courage-15}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r courage-15-test}
#| echo: true
tidy(fit_colleges, conf.int = TRUE)
```

###

`tidy()` returns a tibble with one row per parameter --- intercept plus one per covariate (with 4 dummy rows for the 5-level `selectivity`).

### Exercise 16

Here is a nicer-looking version of that parameter table, rendered with **[knitr](https://yihui.org/knitr/)**:

```{r}
#| echo: false
library(knitr)

tidy(fit_colleges, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  kable(
    caption = "Estimates and 95% Confidence Intervals for Parameters.  Source: IPEDS 2013 via primer.data"
  )
```

###

We care about using our model to estimate quantities of interest, not about the parameters themselves.

### Exercise 17

Add a sentence to your project summary explaining the structure of the model.

```{r courage-17}
question_text(NULL,
  message = "We model graduation rate as a normally distributed variable which is a linear function of tuition and selectivity.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add your sentence to the summary paragraph. Render the QMD, then commit/push.


## Temperance
###

Courage provides the data generating mechanism. We can express the DGM in four ways:

**In words:** We model graduation rate as a normally distributed variable which is a linear function of tuition and selectivity.

**In R code:**

```r
fit_colleges <- linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition + selectivity, data = x)
```

**In a parameter table:**

```{r}
#| echo: false
fit_colleges |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

**As a mathematical formula:**

$$
\begin{aligned}
\text{Graduation Rate} =\ & 0.487 + 0.090 \cdot \text{Tuition} \\
&{} - 0.078 \cdot \text{Highly Selective} \\
&{} - 0.128 \cdot \text{Moderately Selective} \\
&{} - 0.213 \cdot \text{Lowly Selective} \\
&{} - 0.205 \cdot \text{Non-selective} + \epsilon
\end{aligned}
$$

with $\epsilon \sim N(0, 0.008)$.

### Exercise 1

In your own words, describe the use of Temperance in data science.

```{r temperance-1}
question_text(NULL,
  message = "Temperance interprets the data generating mechanism and then uses it to answer, with the help of graphics, the question(s) with which we began. Humility reminds us that this answer is always false.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Courage gave us the data generating mechanism, including all of its parameter values. Temperance guides us in the use of the DGM to answer the question(s) with which we began. Our tool for doing so is the **[marginaleffects](https://marginaleffects.com/)** package, with companion book *[Model to Meaning](https://marginaleffects.com/)* by Vincent Arel-Bundock.

### Exercise 2

In the end, we don't really care about parameters, much less how to interpret them. Parameters are *imaginary*, like unicorns.

Add `library(marginaleffects)` to the setup chunk in `analysis.qmd` and render.

CP/CR.

```{r temperance-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

<pre><code>
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(marginaleffects)
</code></pre>

###

We should be modest in the claims we make. The posteriors we create are never the "truth."

### Exercise 3

What is the specific question we are trying to answer?

```{r temperance-3}
question_text(NULL,
  message = "What is the difference in expected graduation rates between colleges with tuition of $20,000 and colleges with tuition of $30,000?",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 3)
```

###

Data science projects begin with a decision which we face. To make that decision wisely, we drill down to one specific question.

### Exercise 4

In `analysis.qmd`, add a chunk that runs `predictions(fit_colleges)`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r temperance-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r temperance-4-test}
#| echo: true
predictions(fit_colleges)
```

###

`predictions()` returns one row per observation --- 542 here --- with the model's predicted graduation rate for each college.

### Exercise 5

In `analysis.qmd`, add a chunk that runs `plot_predictions(fit_colleges, condition = c("tuition", "selectivity"))`. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r temperance-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r temperance-5-test}
#| echo: true
plot_predictions(fit_colleges, condition = c("tuition", "selectivity"))
```

###

Conditioning on two variables gives us a two-dimensional predicted surface --- here rendered as five parallel lines, one per selectivity tier.

### Exercise 6

Modify the previous chunk to add `draw = FALSE` to the `plot_predictions()` call. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r temperance-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r temperance-6-test}
#| echo: true
plot_predictions(fit_colleges, condition = c("tuition", "selectivity"), draw = FALSE)
```

###

Because `plot_predictions()` returns a ggplot object, you can continue with ggplot commands. But it can be useful to see the underlying values.

### Exercise 7

Create a beautiful plot starting from the output above. Do this in your QMD.

Paste the plot code below.

```{r temperance-7}
question_text(NULL,
  message = "    plot_predictions(fit_colleges, condition = c(\"tuition\", \"selectivity\"), draw = FALSE) |>\n      ggplot(aes(x = tuition, y = estimate, color = selectivity)) +\n      geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = selectivity),\n                  alpha = 0.15, color = NA) +\n      geom_line(size = 1) +\n      scale_x_continuous(labels = scales::label_dollar(scale = 10000)) +\n      scale_y_continuous(labels = scales::label_percent()) +\n      labs(\n        title = \"Predicted Graduation Rate by Tuition and Selectivity\",\n        subtitle = \"Within each selectivity tier, every $10k of tuition adds ~9 percentage points\",\n        x = \"Tuition\", y = \"Predicted Graduation Rate\",\n        color = \"Selectivity\", fill = \"Selectivity\",\n        caption = \"Source: IPEDS 2013 via primer.data\"\n      ) +\n      theme_minimal()",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 12)
```

###

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(fit_colleges, condition = c("tuition", "selectivity"), draw = FALSE) |>
  ggplot(aes(x = tuition, y = estimate, color = selectivity)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = selectivity),
              alpha = 0.15, color = NA) +
  geom_line(size = 1) +
  scale_x_continuous(labels = scales::label_dollar(scale = 10000)) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(
    title = "Predicted Graduation Rate by Tuition and Selectivity",
    subtitle = "Within each selectivity tier, every $10k of tuition adds ~9 percentage points",
    x = "Tuition",
    y = "Predicted Graduation Rate",
    color = "Selectivity",
    fill = "Selectivity",
    caption = "Source: IPEDS 2013 via primer.data"
  ) +
  theme_minimal()
```

### Exercise 8

Finalize the graphics chunk in your QMD. Render the QMD. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r temperance-8}
question_text(NULL,
  message = "    > show_file(\"analysis.qmd\", chunk = \"Last\")\n    ```{r}\n    plot_predictions(fit_colleges, condition = c(\"tuition\", \"selectivity\"), draw = FALSE) |>\n      ggplot(aes(x = tuition, y = estimate, color = selectivity)) +\n      geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = selectivity),\n                  alpha = 0.15, color = NA) +\n      geom_line(size = 1) +\n      scale_x_continuous(labels = scales::label_dollar(scale = 10000)) +\n      scale_y_continuous(labels = scales::label_percent()) +\n      labs(\n        title = \"Predicted Graduation Rate by Tuition and Selectivity\",\n        subtitle = \"Within each selectivity tier, every $10k of tuition adds ~9 percentage points\",\n        x = \"Tuition\", y = \"Predicted Graduation Rate\",\n        color = \"Selectivity\", fill = \"Selectivity\",\n        caption = \"Source: IPEDS 2013 via primer.data\"\n      ) +\n      theme_minimal()\n    ```",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 8)
```

###

The map is not the territory. A beautiful graphic tells a story, but that story is always an imperfect representation of reality.

### Exercise 9

Write the last sentence of your summary paragraph. It describes at least one Quantity of Interest and a measure of uncertainty.

```{r temperance-9}
question_text(NULL,
  message = "We estimate that within a given selectivity tier, each $10,000 of tuition is associated with about 9 percentage points more graduation rate, with a 95% confidence interval of 7 to 11 percentage points; Elite colleges graduate about 20 percentage points more students than Lowly Selective ones at comparable tuition.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 5)
```

###

Add a final sentence to your summary paragraph. Render the QMD.

### Exercise 10

Write a few sentences explaining why the estimates might be wrong.

```{r temperance-10}
question_text(NULL,
  message = "The biggest concern is the 13-year gap between our 2013 data and the 2026 question --- the slope we estimate may not be the slope today. Beyond that, our filter to `tuition > 2` drops low-tuition colleges, so even within 2013 the estimated slope doesn't apply to schools charging under $20,000. Our `grad_rate` outcome excludes students who transfer and graduate elsewhere, which undercounts completion. And selectivity is only a five-level summary of a much richer admissions picture. The confidence intervals the model reports reflect sampling variation around the 542 rows; they don't account for any of those framing, measurement, or time issues.",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Always go back to your Preceptor Table. Even a perfectly estimated statistical model is rarely as useful as we might like.

### Exercise 11

Rearrange the material in your QMD so the order is graphic, then summary paragraph. The chunk that creates `fit_colleges` must occur before the chunk that creates the graphic. Render the QMD. In the R Terminal, run `show_file("analysis.qmd")`. CP/CR.

```{r temperance-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

<pre><code>
---
title: "Colleges"
author: "&lt;your name&gt;"
execute:
  echo: false
---

#| message: false
library(tidyverse)
library(tidymodels)
library(primer.data)
library(broom)
library(marginaleffects)

x <- colleges |>
  select(tuition, grad_rate, selectivity) |>
  filter(tuition > 2) |>
  drop_na()

#| cache: true
fit_colleges <- linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition + selectivity, data = x)

plot_predictions(fit_colleges, condition = c("tuition", "selectivity"), draw = FALSE) |>
  ggplot(aes(x = tuition, y = estimate, color = selectivity)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = selectivity),
              alpha = 0.15, color = NA) +
  geom_line(size = 1) +
  labs(
    title = "Predicted Graduation Rate by Tuition and Selectivity",
    subtitle = "Within each selectivity tier, every $10k of tuition adds ~9 percentage points",
    x = "Tuition ($10,000)", y = "Predicted Graduation Rate",
    color = "Selectivity", fill = "Selectivity",
    caption = "Source: IPEDS 2013 via primer.data"
  ) +
  theme_minimal()

Where students go to college and what they pay for it varies enormously, and so does the likelihood that they will finish. We help today's students choose by examining how graduation rate varies with tuition and selectivity, using 2013 IPEDS data on roughly 900 US four-year colleges and universities --- the most recent reliable national snapshot we have. The 13-year gap between our data and the question is a real concern: tuition has risen much faster than inflation since 2013, and the price/graduation relationship may have shifted. We model graduation rate as a normally distributed variable which is a linear function of tuition and selectivity. We estimate that within a given selectivity tier, each $10,000 of tuition is associated with about 9 percentage points higher graduation rate, with a 95% confidence interval of 7 to 11 percentage points; Elite colleges have graduation rates about 20 percentage points higher than Lowly Selective ones at comparable tuition.
</code></pre>

###

This is the version of your QMD file at which your teacher is most likely to look closely.

### Exercise 12

Publish your rendered QMD to GitHub Pages. In the bash Terminal (not the R Terminal!), run:

```
quarto publish gh-pages analysis.qmd
```

Copy/paste the resulting URL below.

```{r temperance-12}
question_text(NULL,
  message = "https://<your-github-username>.github.io/colleges/analysis.html",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

Commit/push everything.

### Exercise 13

Copy/paste the URL to your GitHub repo.

```{r temperance-13}
question_text(NULL,
  message = "https://github.com/<your-github-username>/colleges",
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 2)
```

###

We can never know all the entries in the Preceptor Table. That knowledge is reserved for God.

The world confronts us. Make decisions we must.


## Summary
###

```{r}
#| echo: false
#| message: false
#| warning: false
plot_predictions(fit_colleges, condition = c("tuition", "selectivity"), draw = FALSE) |>
  ggplot(aes(x = tuition, y = estimate, color = selectivity)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = selectivity),
              alpha = 0.15, color = NA) +
  geom_line(size = 1) +
  scale_x_continuous(labels = scales::label_dollar(scale = 10000)) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(
    title = "Predicted Graduation Rate by Tuition and Selectivity",
    subtitle = "Within each selectivity tier, every $10k of tuition adds ~9 percentage points",
    x = "Tuition",
    y = "Predicted Graduation Rate",
    color = "Selectivity",
    fill = "Selectivity",
    caption = "Source: IPEDS 2013 via primer.data"
  ) +
  theme_minimal()
```

Where students go to college and what they pay for it varies enormously, and so does the likelihood that they will finish. We help today's students choose by examining how graduation rate varies with tuition and selectivity, using 2013 IPEDS data on roughly 900 US four-year colleges and universities --- the most recent reliable national snapshot we have. The 13-year gap between our data and the question is a real concern: tuition has risen much faster than inflation since 2013, and the price/graduation relationship may have shifted. We model graduation rate as a normally distributed variable which is a linear function of tuition and selectivity. We estimate that within a given selectivity tier, each $10,000 of tuition is associated with about 9 percentage points higher graduation rate, with a 95% confidence interval of 7 to 11 percentage points; Elite colleges have graduation rates about 20 percentage points higher than Lowly Selective ones at comparable tuition.

This forecast would help a student gauge whether a specific school's graduation rate is typical for its tier or unusual. It does not address every question a student might face --- it says nothing about outcomes after graduation, nor about the experience of students within a given college. But it gives a defensible starting point.

The world is always more uncertain than our models would have us believe.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
